### 2014-04-09 binary/LaTeX/RareBinary.tex
Initial Commit
Starting point: Copy from SVN

\documentclass[11pt]{article}
\usepackage{fullpage}
\usepackage{amssymb, amsthm, amsmath, amsfonts}
\usepackage{doublespace}
\usepackage{bm}
\usepackage{graphicx}
\usepackage[authoryear]{natbib}
\usepackage{bm}
\usepackage{verbatim}
\usepackage{lineno}
\usepackage{times}

\linespread{1.0}

\newcommand{\btheta}{ \mbox{\boldmath $\theta$}}
\newcommand{\bmu}{ \mbox{\boldmath $\mu$}}
\newcommand{\balpha}{ \mbox{\boldmath $\alpha$}}
\newcommand{\bbeta}{ \mbox{\boldmath $\beta$}}
\newcommand{\bdelta}{ \mbox{\boldmath $\delta$}}
\newcommand{\blambda}{ \mbox{\boldmath $\lambda$}}
\newcommand{\bgamma}{ \mbox{\boldmath $\gamma$}}
\newcommand{\brho}{ \mbox{\boldmath $\rho$}}
\newcommand{\bpsi}{ \mbox{\boldmath $\psi$}}
\newcommand{\bepsilon}{ \mbox{\boldmath $\epsilon$}}
\newcommand{\bomega}{ \mbox{\boldmath $\omega$}}
\newcommand{\bDelta}{ \mbox{\boldmath $\Delta$}}
\newcommand{\bSigma}{ \mbox{\boldmath $\Sigma$}}
\newcommand{\bA}{ \mbox{\bf A}}
\newcommand{\bP}{ \mbox{\bf P}}
\newcommand{\bx}{ \mbox{\bf x}}
\newcommand{\bX}{ \mbox{\bf X}}
\newcommand{\bB}{ \mbox{\bf B}}
\newcommand{\bZ}{ \mbox{\bf Z}}
\newcommand{\by}{ \mbox{\bf y}}
\newcommand{\bY}{ \mbox{\bf Y}}
\newcommand{\bz}{ \mbox{\bf z}}
\newcommand{\bh}{ \mbox{\bf h}}
\newcommand{\br}{ \mbox{\bf r}}
\newcommand{\bt}{ \mbox{\bf t}}
\newcommand{\bs}{ \mbox{\bf s}}
\newcommand{\bb}{ \mbox{\bf b}}
\newcommand{\bL}{ \mbox{\bf L}}
\newcommand{\bu}{ \mbox{\bf u}}
\newcommand{\bv}{ \mbox{\bf v}}
\newcommand{\bV}{ \mbox{\bf V}}
\newcommand{\bG}{ \mbox{\bf G}}
\newcommand{\bH}{ \mbox{\bf H}}
\newcommand{\bw}{ \mbox{\bf w}}
\newcommand{\bo}{ \mbox{\bf o}}
\newcommand{\bfe}{ \mbox{\bf e}}
\newcommand{\iid}{\stackrel{iid}{\sim}}
\newcommand{\indep}{\stackrel{indep}{\sim}}
\newcommand{\calR}{{\cal R}}
\newcommand{\calG}{{\cal G}}
\newcommand{\calD}{{\cal D}}
\newcommand{\calS}{{\cal S}}
\newcommand{\calB}{{\cal B}}
\newcommand{\calA}{{\cal A}}
\newcommand{\calT}{{\cal T}}
\newcommand{\calO}{{\cal O}}
\newcommand{\argmax}{{\mathop{\rm arg\, max}}}
\newcommand{\argmin}{{\mathop{\rm arg\, min}}}
\newcommand{\Frechet}{\mbox{Fr$\acute{\mbox{e}}$chet}}
\newcommand{\Matern}{ \mbox{Mat$\acute{\mbox{e}}$rn}}

\newcommand{\beq}{ \begin{equation}}
\newcommand{\eeq}{ \end{equation}}
\newcommand{\beqn}{ \begin{eqnarray}}
\newcommand{\eeqn}{ \end{eqnarray}}

\begin{document}\linenumbers

\begin{center}
{\Large {\bf A spatial model for rare binary events}}\\
\today
\end{center}

\section{Introduction}\label{s:intro}

\section{New Model?}\label{s:model}

Let $Y_i\in\{0,1\}$ be the binary response at spatial location $\bs_i\in\calD$, and $\bX_i$ be the associated $p$-vector of covariates with first element equal to one for the intercept.  We relate the covariates with the response using the link function $g$ so that P$(Y_i=1) = p_i= g(\bX_i\bbeta)$, where $\bbeta$ is the $p$-vector of regression coefficients.  For example, \cite{Wang-2010} propose the GEV link function
$p_i=1-\exp\left[\left(1-\xi\bX_i\bbeta\right)^{-1/\xi}\right]$ for rare binary data. We will also consider logit and probit links.

-- Not quite sure why the article uses this. I think we should use
\begin{align}
	p_i = 1 - \exp \left[ - \left(1 + \xi \bX_i \bbeta \right)^{-1/\xi} \right] 
\end{align}

We propose a copula \citep{nelsen-1999} to account for spatial dependence while preserving the marginal event probabilities. Let $Y_i = I(Z_i>z_i)$, where $Z_i$ is a continuous latent variable and $z_i$ is the appropriate threshold so that $P(Y_i=1)=p_i$.  The latent $Z_i$ is modeled using spatial extreme value analysis methods to capture dependence between rare events.  We assume $Z$ follows the max-stable spatial process of \cite{reich-2012}.  Under this model, the marginal distribution of each $Z_i$ is GEV(1,1,1) with P$(Z_i>c) = 1-\exp(-1/c)$.  Therefore, we must set $z_i=-1/\log(1-p_i)$ so that $P(Y_i=1)=p_i$.

Spatial dependence is determined by the joint distribution of $\bZ = (Z_1,\ldots,Z_n)$, 
\beq\label{jointCDF}
 G(\bz) =  \mbox{P}[Z_1<z_1,\ldots,Z_n<z_n] = \exp\left\{-\sum_{l=1}^L\left[\sum_{i=1}^n\left(\frac{w_{l}(\bs_i)}{z_i}\right)^{1/\alpha}\right]^{\alpha}\right\},
\eeq
where $\bz = (z_1,\ldots,z_n)$. This is a special case of the multivariate GEV distribution with asymmetric Laplace dependence function \citep{Tawn-1990}.  The parameter $\alpha\in(0,1)$ determines the strength of dependence, with $\alpha$ near zero giving strong dependence and $\alpha=1$ giving joint independence. The weights $w_{li}>0$ determine the spatial dependence structure, and are discussed in detail in Section \ref{s:spatial}.  Many weight functions are possible, but the weights must be constrained so that $\sum_{l=1}^L w_{l}(\bs_i)=1$ for all $i=1,\ldots,n$ to preserve the marginal GEV distribution.

\section{Spatial dependence}\label{s:spatial}

The weights $w_{l}(\bs_i)$ in (\ref{jointCDF}) should vary smoothly across space to induce spatial dependence.  For example, \cite{reich-2012} take the weights to be scaled Gaussian kernels with knots $\bv_l$, that is
\beq\label{w}
   w_{l}(\bs_i) = \frac{\exp\left[-0.5\left(||\bs_i-\bv_l||/\rho\right)^2\right]}
                 {\sum_{j=1}^L\exp\left[-0.5\left(||\bs_i-\bv_j||/\rho\right)^2\right]}.
\eeq
To kernel bandwidth $\rho>0$ determines the spatial range of the dependence, with large $\rho$ giving long-range dependence and vice versa.

Then in a bivariate setting, the probability of observing a joint exceedances as a function of $\alpha$ is

\begin{align} \label{biv}
  \text{P}(Y_i = 1, & Y_j = 1) = 1 - \exp\left\{ - \frac{ 1 }{ z_i } \right\} - \exp \left\{ - \frac{ 1 }{z_j} \right\} + \exp \left\{ - \sum_{ l = 1 }^{ L } \left[ \left( \frac{ w_{ l }(\bs_i) }{ z_i } \right)^{1/\alpha} + \left( \frac{ w_{l }(\bs_i)}{z_j} \right)^{1/\alpha} \right]^{\alpha} \right\} \nonumber \\
  		&= p_i + p_j - \left( 1 - \exp \left \{ - \sum_{ l = 1 }^{ L } \left( \left[ -\log(1-p_i) w_{l}(\bs_i) \right]^{ 1/\alpha } + \left[ -\log(1 - p_j) w_{ l}(\bs_j) \right]^{ 1/\alpha } \right)^{\alpha} \right \} \right).
\end{align}

To describe the tail dependence, we use the $\chi$ statistic of \citet{Coles-1999}.
Assume that $Y_i$ and $Y_j$ have the same marginal distributions, then $p_i = p_j = p$ for all $i, j$.
As shown in Appendix A.2,

\begin{align}
\chi = 2 -  \vartheta(\bs_i, \bs_j).
\end{align}
where $\vartheta (\bs_i, \bs_j) = \sum_{ l = 1 }^{ L } \left[ w_{l}(\bs_i)^{ 1/\alpha } +  w_{ l}(\bs_j)^{ 1/\alpha } \right]^\alpha$ is the pairwise extremal coefficient given by \citet{reich-2012}.
In the case of complete dependence, $\chi = 1$, and in the case of complete independence, $\chi = 0$. 
This is relatively easy to show for $\alpha = 1$, but I don't know of a way to prove $\lim_{\alpha \rightarrow 0} \chi = 1$. Any thoughts?

\section{Computation}\label{s:comp}

As shown in Appendix A.1, the joint probability mass function of $\bY=(Y_1,\ldots,Y_n)$ has a convenient form when the number of events is small.  Let $K=\sum_{i=1}^nY_i$ be the number of events, and assume without loss of generality the data are ordered so that the $Y_1=\ldots=Y_K=1$.  Then
\beq\label{pmf}
   P(Y_1=y_1,\ldots,Y_n=y_n) =  \left\{
                               \begin{array}{ll}
                                 G(\bz) & K=0 \\
                                 G(\bz_{(1)})-G(\bz) & K=1 \\
                                 G(\bz_{(12)})-G(\bz_{(1)})-G(\bz_{(2)})+G(\bz) & K=2
                               \end{array}
                             \right.
\eeq
where $G(\bz_{(1)}) = P(Z_2<z_2,\ldots,Z_n<z_n)$, $G(\bz_{(2)}) = P(Z_1<z_1,Z_3<z_3,\ldots,Z_n<z_n)$, and $G(\bz_{(12)}) = P(Z_3<z_3,\ldots,Z_n<z_n)$.  Similar expressions can be derived for all $K$, but become cumbersome for large $K$.  Therefore, for small $K$ we can evaluate the likelihood directly.  Most days in our dataset have $K<4$, so we use this expression for those days.  However for days with many events, we must use the latent variable scheme described below (unless you can think of a better way!).
I think it should be more computationally efficient to use (\ref{pmf}) for any $K$. At most, we have to calculate the $\left( \frac{ w_{l}(\bs_i) }{ z_i } \right)^{1/\alpha}$, for all $i, l$. In the random effects model, the expression for the joint density conditional on $\theta$ is 
\beq
	P(Y_1=y_1,\ldots,Y_n=y_n) = \prod_{ i = 1 }^{ n } \left[ \exp \left\{ \sum_{ l = 1 }^{L} A_l \left( \frac{ w_{l}(\bs_i) }{ z_i } \right)^{ 1/\alpha} \right\} \right]^{ 1 - Y_i } \left[ 1 - \exp \left\{ \sum_{ l = 1 }^{L} A_l \left( \frac{ w_{l}(\bs_i) }{ z_i } \right)^{ 1/\alpha} \right\} \right]^{ Y_i }.
\eeq
So we still need to compute $\left( \frac{ w_{l}(\bs_i) }{ z_i } \right)^{1/\alpha}$, but we also need to do the sampling for all the $A_l$ terms as well.

\section{Simulation study}\label{s:sim}

\section{Data analysis}\label{s:analysis}

\section{Conclusions}\label{s:con}

\section*{Acknowledgments}

\section*{Appendix A.1: Derivation of the likelihood}
We use the hierarchical max-stable spatial model given by \citet{reich-2012}. If at each margin, $Z_i \sim $ GEV$(1,1,1)$, then $Z_i | \theta_i \indep $ GEV$(\theta, \alpha \theta, \alpha)$. As defined in section \ref{s:comp}, we reorder the data such that $Y_1=\ldots=Y_K=1$, and $Y_{K+1} = \ldots = Y_n = 0$. Then the joint likelihood conditional on the random effect $\theta$ is

\begin{align} \label{joint_cond}
	P(Y_1=y_1,\ldots,Y_n=y_n) &= \prod_{ i \le K } \left\{ 1 - \exp \left[ - \left( \frac{ \theta_i }{ z_i } \right)^{ 1/\alpha} \right] \right \} \prod_{ i > K } \exp \left[ -\left( \frac{ \theta_i }{ z_i } \right)^{1/\alpha} \right] \nonumber \\[0.5em]
		&= \exp \left[ -\sum_{ i = K+1}^{ n }\left( \frac{ \theta_i }{ z_i } \right)^{1/\alpha} \right] - \exp \left[ -\sum_{ i = K+1}^{ n }\left( \frac{ \theta_i }{ z_i } \right)^{1/\alpha} \right] \sum_{ i = 1}^{K} \exp\left[ -\left( \frac{ \theta_i }{ z_i } \right)^{ 1/\alpha} \right] \nonumber\\
		&  + \exp \left[ -\sum_{ i = K+1}^{ n }\left( \frac{ \theta_i }{ z_i } \right)^{1/\alpha} \right] \sum_{ 1 < i < j \le K } \left\{ \exp \left[ - \left( \frac{ \theta_i }{ z_i } \right)^{ 1/\alpha} - \left( \frac{ \theta_j }{ z_j } \right)^{ 1/\alpha } \right] \right \} \nonumber \\[0.5em]
		& + \cdots + (-1)^K \exp\left[ - \sum_{ i = 1 }^{ n }\left( \frac{ \theta_i }{ z_i } \right)^{ 1/\alpha} \right]
\end{align}

Finally marginalizing over the random effect, we obtain

\begin{align} \label{joint}
    P(Y_1=y_1,\ldots,Y_n=y_n) &=\int G(\bz | \bA) p( \bA | \alpha) d\bA. \nonumber\\[0.5em]
			&= \int \exp \left[ -\sum_{ i = K+1}^{ n }\left( \frac{ \theta_i }{ z_i } \right)^{1/\alpha} \right] - \exp \left[ -\sum_{ i = K+1}^{ n }\left( \frac{ \theta_i }{ z_i } \right)^{1/\alpha} \right] \sum_{ i = 1}^{K} \exp\left[ -\left( \frac{ \theta_i }{ z_i } \right)^{ 1/\alpha} \right] \nonumber\\
		&  + \exp \left[ -\sum_{ i = K+1}^{ n }\left( \frac{ \theta_i }{ z_i } \right)^{1/\alpha} \right] \sum_{ 1 < i < j \le K } \left\{ \exp \left[ - \left( \frac{ \theta_i }{ z_i } \right)^{ 1/\alpha} - \left( \frac{ \theta_j }{ z_j } \right)^{ 1/\alpha } \right] \right \} \nonumber \\[0.5em]
		& + \cdots + (-1)^K \exp\left[ - \sum_{ i = 1 }^{ n }\left( \frac{ \theta_i }{ z_i } \right)^{ 1/\alpha} \right] p( \bA | \alpha) d\bA.
\end{align}

Consider the first term in the summation,

\begin{align}
	\int \exp \left\{ -\sum_{ i = K+1}^{ n }\left( \frac{ \theta_i }{ z_i } \right)^{1/\alpha} \right\} p( \bA | \alpha) d\bA &= \int \exp \left\{ - \sum_{ i = K + 1 }^n \left( \frac{ \left[ \sum_{ l = 1 }^L  A_l w_{l}(\bs_i)^{1/\alpha} \right)^\alpha }{ z_i} \right]^{ 1/\alpha } \right \} p( \bA | \alpha) d\bA \nonumber \\[0.5em]
	 &= \int \exp \left\{ -\sum_{ i = K + 1}^n \sum_{ l = 1}^L A_l \left( \frac{ w_l (\bs_i) }{ z_i } \right)^{1/\alpha} \right \} p( \bA | \alpha) d\bA \nonumber \\[0.5em]
	 &=\exp\left\{-\sum_{ l = 1}^L \left[ \sum_{ i = K + 1 }^n \left( \frac{ w_l(\bs_i)}{ z_i} \right)^{1/\alpha} \right]^\alpha \right\}.
\end{align}

The remaining terms in equation (\ref{joint}) are straightforward to obtain, and after integrating out the random effect, the joint density is the density given in (\ref{pmf}).

\section*{Appendix A.2: Derivation of the $\chi$ statistic}
\begin{align} \label{chi}
  \chi &= \lim_{p \rightarrow 0 }\text{P}(Y_i = 1|Y_j = 1)\nonumber \\
   &= \lim_{p \rightarrow \infty }\frac{ p + p - \left( 1 - \exp \left \{ - \sum_{ l = 1 }^{ L } \left[ \left( -\log(1-p) w_{l}(\bs_i) \right)^{ 1/\alpha } + \left( -\log(1 - p) w_{ l}(\bs_j) \right)^{ 1/\alpha } \right]^{\alpha} \right \} \right) }{ p } \nonumber \\[0.5em]
  &= \lim_{p \rightarrow 0 } \frac{ 2p - \left( 1 - \exp \left \{ \log(1-p) \sum_{ l = 1 }^{ L } \left[  w_{l}(\bs_i) ^{ 1/\alpha } +  w_{ l}(\bs_j) ^{ 1/\alpha } \right]^{\alpha} \right \} \right) }{ p } \nonumber \\[0.5em]
  &= \lim_{p \rightarrow 0 } \frac{ 2p - \left( 1 - (1-p)^{ \sum_{ l = 1 }^{ L } \left[ \left( w_{l}(\bs_i) \right)^{ 1/\alpha } + \left( w_{ l}(\bs_j) \right)^{ 1/\alpha } \right]^\alpha } \right) }{ p } \nonumber \\[0.5em]
  &=\lim_{p \rightarrow 0 } 2 -\sum_{ l = 1 }^{ L } \left[ w_{l}(\bs_i) ^{ 1/\alpha } +  w_{ l} (\bs_j)^{ 1/\alpha } \right]^\alpha ( 1 - p)^{ -1 + \sum_{ l = 1 }^{ L } \left[  w_{l}(\bs_i) ^{ 1/\alpha } +  w_{ l}(\bs_j)^{ 1/\alpha } \right]^\alpha } \nonumber \\[0.5em]
  &= 2 -  \sum_{ l = 1 }^{ L } \left[ w_{l}(\bs_i)^{ 1/\alpha } +  w_{ l}(\bs_j) ^{ 1/\alpha } \right]^\alpha.
\end{align}

\begin{singlespace}
\bibliographystyle{rss}
\bibliography{spatialbinary}
\end{singlespace}

\end{document}

### 2014-04-09 pot/LaTeX/Schlather.tex
Initial Commit
Starting point: Copy from SVN

\documentclass[11pt]{article}
\usepackage{amssymb, amsthm, amsmath}
\usepackage{bm}
\usepackage{graphicx}
\usepackage[authoryear]{natbib}
\usepackage{bm}
\usepackage{verbatim}
\usepackage{lineno}
\usepackage{times}
\usepackage{soul}
\usepackage{color}

\usepackage[left=1in,top=1in,right=1in]{geometry}
\pdfpageheight 11in
\pdfpagewidth 8.5in
\linespread{1.0}
\input{mycommands.sty}

\begin{document}\linenumbers

\begin{center}
{\Large {\bf A new spatial model for points above a threshold}}\\
\today
\end{center}

\section{Introduction}\label{s:intro}

\section{Statistical model}\label{s:model}

Our approach is to take the construction of \cite{schlather2002} for a max-stable model for block maxima as a literal model for extreme daily data.  We will show that his has nice theoretical and computational properties on the daily scale, and gives the popular \cite{schlather2002} as the limit of the block maximum data.

Let $Y_t(\bs)\in \calR$ be the observed value at location $\bs$ on day $t$.  To avoid bias in estimating tail parameters, we model the thresholded data
\beq\label{Yt}
  \Yt_t(\bs) = \left\{
          \begin{array}{ll}
            Y_t(\bs) & Y_t(\bs)>T \\
            T & Y_t(\bs)\le T
          \end{array}
        \right.
\eeq
where $T$ is a pre-specified threshold.

We first specify a model for the complete data, $Y_t(\bs)$, and then study the induced model for thresholded data, $\Yt_t(\bs)$.  
The full data model is given in Section \ref{s:model1} assuming a multivariate normal distribution with a different variance each day.
Computationally, the values below the threshold are updated using standard Bayesian missing data methods as described in Section \ref{s:comp}.

\subsection{Complete data}\label{s:model}
Consider the model
\begin{align} \label{fullmodel}
  Y_t(\bs) &= X_t(\bs) \beta + e_t(\bs)\\
  e_t(\bs) &= \sigma \delta | u_t(\bs) | + v_t(\bs)
\end{align}
where $u_t(\bs)$ has marginal $N(0, 1)$ distribution, $\delta \in (-1, 1)$ controls skew, and $v_t(\bs)$ is a spatial process with mean zero and variance $\sigma^2(1 - \delta^2)$.

% \subsection{Complete data}\label{s:model1}
% Consider a zero-mean Gaussian process $\epsilon_t(\bs)$ with $\Matern$ correlation structure.  
% Consider the model
% \beq\label{fullmodel}
%   Y_t(\bs) = \mu(\bs) + r_t W_t(\bs).
% \eeq
% where $r_t$ are daily random effects and $W_t$ are spatial processes with mean E$[W_t(\bs)]=0$, variance V$[W_t(\bs)]=1$, and $\Matern$ spatial correlation Cor$[W_t(\bs_1),W_t(\bs_2)]=M(||\bs_1-\bs_2||;\rho,\nu)$ with spatial range $\rho>0$, smoothness $\nu>0$.
% Let $\alpha$ be the proportion of the total variance, $\sigma^2_W$, that is accounted for by the partial sill of the spatial process.
% Then we denote the spatial $\Matern$ process model for $W_t$ as $W_t\sim \mbox{Mat}(0,1,\rho,\nu, \alpha)$.
% So for a single day, 
% \beq \label{eq:singleday}
% Y_t(\bs)|r_t \sim \mbox{Mat}[\mu(\bs), r_t, \rho, \nu, \alpha].
% \eeq
% We consider two different random effects distributions, $r_t \iid \mbox{GPD}(0, \sigma_r, \xi_r)$ and $r_t \iid \mbox{Inv. Gamma}(\xi_r, \sigma_r)$.

%\subsubsection{Special case}\label{s.powerlaw}
%Consider a special case where $r_t \iid GPD(\sigma_r/\xi_r, \sigma_r, \xi_r)$. Then $r_t$ follows a power law distribution. 
%The algebra works out and we can use an IG kernel to marginalize out the $r_t$ terms. 
%Similar results come from letting $r_t \iid GEV(\sigma_r/\xi_r, \sigma_r, \xi_r)$.
%Otherwise, I don't really have a good place to start for marginalizing over the random effect term. 
%After marginalizing out $r_t$ in this special case, we are left with (I'll add in the verification to the appendix)
%\beq
%f_Z(z) = \left[ \frac{ 1 }{ 2\sigma^2_W }(\bZ^T \Sigma^{-1} \bZ)\right]^{-\frac{ n }{ 2}} \Gamma\left(\frac{ n }{ 2 } \right) \left[ \frac{\xi_r}{\sigma_r} \right]^{-1/\xi_r - 1}  (2\pi)^{ -n/2} \sigma_W^{-n} \left| \Sigma \right|^{-1}\exp\left\{ -\frac{ 1 }{ 2 \sigma^2_W } \left(\bZ^T \Sigma^{-1} \bZ\right) \right\}
%\eeq
%which looks strangely like a GEV distribution.
%In fact, marginally at each site, I'm pretty sure this will follow a univariate GEV distribution.
%Also, to figure out the bivariate joint density from this distribution shouldn't all that challenging either.
%It seems like there's a relationship between the GEV and the Generalized Gamma Distribution, but as of right now, I can only get them to match up when we have the relationship that $\mu_r = \frac{ \xi_r }{ \sigma_r }$.
%The pdf for the Generalized Gamma Distribution is
%\beq \label{eq:ggamma}
% \frac{ p / a^d }{ \Gamma(d/p) } x^{ d-1 }\exp\{ -(x/a)^p \}
%\eeq
%and I also found a four-parameter version that gives a shift parameter with density
%\beq
% \frac{ \theta^\alpha c }{\Gamma(\alpha) } (x - \gamma)^{ c \alpha - 1} \exp\{ -\theta(x - \gamma)^c \}
%\eeq
%which is equivalent to (\ref{eq:ggamma}) where $\alpha = d/p$, $\theta=(1/a)^p$ and $c=p$
%Do you think I should continue pursuing this, or start focusing on modifying the MCMC now?

\subsubsection{GPD random effect}
Let $r_t \iid \mbox{GPD}(0, \sigma_r, \xi_r)$

\subsubsection{Power law random effect}
Let $r_t \iid \mbox{GPD}(\sigma_r/\xi_r, \sigma_r, \xi_r)$. Then by marginalizing out the $r_t$ terms, we find that 
\beq
	f_Y(y) = \frac{ 1 }{ \sigma_r }\left( \frac{ \xi_r }{ \sigma_r } \right)^{ -1/\xi_r - 1 } (2\pi)^{ -\frac{ n }{ 2 } }\beta^{ -\left( \frac{ 1 }{ \xi_r } + \frac{ n }{ 2 } \right) } \gamma\left( \frac{ 1 }{ \xi_r } + \frac{ n }{ 2 }, \frac{ \xi_r }{ \sigma_r } \beta \right)
\eeq
where $\beta = \frac{ 1 }{ 2 } (\bY - \bmu)^T \Sigma^{-1} (\bY - \bmu)$ and $\gamma(s, x)$ is the lower incomplete gamma function. 
We also find that marginally, 
\beqn
	f(y) = \frac{ 1 }{ \sigma_r }\left( \frac{ \xi_r }{ \sigma_r } \right)^{ -1/\xi_r - 1 } (2\pi)^{ -\frac{ 1 }{ 2 } }\beta^{ -\left( \frac{ 1 }{ \xi_r } + \frac{ 1 }{ 2 } \right) } \gamma\left( \frac{ 1 }{ \xi_r } + \frac{ 1 }{ 2 }, \frac{ \xi_r }{ \sigma_r } \beta \right)
\eeqn
where $\beta = \frac{ (y - \mu)^2 }{ 2 }$.

\subsubsection{Gamma random effect}
	Let $r_t \iid \mbox{IG}(\xi_r, \sigma_r)$. Then after marginalizing out the random effect, we find that $Y$ follows a multivariate T distribution with $2 \xi_r$ degrees of freedom, $\mu = \mu(\bs)$, and $\widehat{\Sigma} = \frac{ \sigma_r }{ \xi_r }{ \Sigma }$ where $\Sigma$ is the correlation matrix from (\ref{eq:singleday}). 
Then, for a single location 
\beqn\label{marg}
	P[Y_t(\bs) > y] &=& 1 - F(y)
\eeqn
where $F(y)$ is the distribution function for a t-distribution with $2 \xi_r$ degrees of freedom and $\mu = \mu$.
The joint is 
\beqn\label{joint}
	P[Y_t(\bs_1) > y, Y_t(\bs_2) > y] &=& 1 - F(y) - F(y) + F_{12}(y) 
\eeqn
where $F_{12}(y)$ is the multivariate t-distribution described above.
In this model,
\beqn\label{chi}
	P[Y_t(\bs_1) > y | Y_t(\bs_2) > y] &=& \frac{ 1 - F(y) - F(y) + F_{12}(y) }{ 1 - F(y) }\nonumber\\
	&=& 1 - \frac{ F(y) - F_{12}(y) }{ 1 - F(y) } = \chi(h)
\eeqn
where $h = ||\bs_1 - \bs_2||$. It's not entirely obvious here, but the $h$ comes in from the $\widehat{\Sigma}$ in the t-distribution.
%From here on out is from the old version of the paper.
%Then if $\mu_W$ and $\sigma_W$ satisfy $E[\exp(W_t(\bs)] = \exp(\mu_W + \sigma_W^2/2) = 1$,
%\beqn\label{marg}
%  P[Z_t(\bs)>z] &=& P\left\{r_t>z\exp[-W_t(\bs)]\right\}\\
%  &=& \mbox{E}\{\exp[W_t(s)]\}/z = 1/z,\nonumber
%\eeqn
%so $Z_t(\bs)$ has GPD(1,1,1) marginal.
%
%The joint is
%\beqn\label{joint}
%  P[Z_t(\bs_1)>z,Z_t(\bs_2)>z] &=& P\left\{r_t>z\exp[-W_t(\bs_1)],r_t>z\exp[-W_t(\bs_2)]\right\}\\
%  &=& P\left\{r_t>z\exp[-\min\{W_t(\bs_1),W_t(\bs_2)\}]\right\}\nonumber\\
%  &=& \mbox{E}\{\exp[\min\{W_t(\bs_1),W_t(\bs_2)\}]\}/z = \chi(||\bs_1-\bs_2||)/z,\nonumber
%\eeqn
%where $\chi(||\bs_1-\bs_2||) = \mbox{E}\{\exp[\min\{W_t(\bs_1),W_t(\bs_2)\}]\}$.  In this model,
%\beq\label{chi}
%  P[Z_t(\bs_1)>z|Z_t(\bs_2)>z] = \frac{P[Z_t(\bs_1)>z,Z_t(\bs_2)>z]}{P[Z_t(\bs_2)>z]} = \chi(||\bs_1-\bs_2||).
%\eeq
%We'll have to work it out, but I'm guessing this is the Schlather model's extremal coefficient.

\subsubsection{Addressing long-range dependence}
In (\ref{chi}), $\chi(h)$ does not decrease to zero as distance $h=||\bs_1-\bs_2||$ goes to infinity. 
This is because all spatial locations share a common random effect $r_t$, and thus locations separated by very large distances are still dependent.  
%Can we address this as in the Brown-Resnick version of the max-stable process?  Let's start by ignoring this, but keep in the back of our minds.

To remove long-range dependence, we allow the random effect $r_t$ to vary by subregion.  
We partition the spatial domain into $J$ subregions and allow the random effects to vary by subregion, $r_t(\bs) = r_{tj}$ if $\bs$ is in subregion $j$. 
The partitioning is treated as a random process.
Let $\bv_{t1},...,\bv_{tJ}$ be spatial knots and assign
\beqn\label{partition}
  r_t(\bs) = r_{tj} \mbox{\ \ if \ \ } j=\argmin_l\{||\bs-\bv_{tl}||\}.
\eeqn 
The knots $\bv_{t1},...,\bv_{tJ}$ are given a homogeneous Poisson process and $r_{tj}\iid \mbox{IG}(\xi_r, \sigma_r)$.
Denote $\pi(h)$ as the probability that two points separated by distance $h$ are in the same subregion.

Given that two locations are in the same subdomain, their joint survival function is (\ref{marg}).  If not,
\beqn\label{marg2}
  P[Y_t(\bs_1) > y, Y_t(\bs_2) > y ] &=& [1 - F(y)][ 1 - F(y)].
\eeqn
Combining both cases gives
\beqn\label{chi2}
  P[Y_t(\bs_1)>y|Y_t(\bs_2)>y] = [1 - \pi(h)][1 - F(y)] + \pi(h)\left[1 - \frac{ F(y) - F_{12}(y) }{ 1 - F(y) } \right].
\eeqn
Then
\beqn\label{chi2}
  \lim_{y\rightarrow \infty}P[Y_t(\bs_1)>y|Y_t(\bs_2)>y] = \pi(h) \chi(h),
\eeqn
and 
\beqn\label{chi3}
  \lim_{y\rightarrow \infty , h\rightarrow \infty}P[Y_t(\bs_1)>y|Y_t(\bs_2)>y] = 0 
\eeqn
if $\lim_{h\rightarrow \infty}\pi(h) = 0 $. 
A proof of this is given in Appendix A.1
Under this model, (\ref{chi2}) shows there is asymptotic dependence, and (\ref{chi3}) shows that long-range spatial dependence is eliminated.

%\subsection{Complete data - general GPD marginals}\label{s:model2}
%Returning to the complete data, $Y_t(\bs)$, if $Z_t(\bs)\sim$ GPD(1,1,1), then let $Y_t(\bs) = \mu(\bs) + \sigma g_{\xi}[Z_t(\bs)]$, where $g_{\xi}(z) = (z^{\xi}-1)/\xi$.  After this transformation, the marginal distribution of $Y_t$ is
%\beqn\label{Ymarg}
%	Y_t(\bs) &\sim& \mbox{GPD}[\mu^*(\bs),\sigma^*(\bs),\xi]
%\eeqn
%where $\mu(\bs)^* = \mu(\bs) + \frac{ \sigma }{ \xi } \{ \exp[\xi W_t(\bs)] - 1 \}$ and $\sigma(\bs)^* = \sigma \exp[ \xi W_t(\bs) ]$. This has the same tail dependence as in (\ref{chi}).  So, the final complete data model is
%\beqn\label{Y}
%  Y_t(\bs) &=& \mu(\bs) + \sigma g_{\xi}\left\{r_t\exp[W_t(\bs)]\right\}\\
%  r_t      &\iid&  \mbox{GPD}(1,1,1)\nonumber\\
%  W_t      &\iid&  \mbox{Mat}(\mu_W,\sigma_W,\rho,\nu, \alpha).\nonumber
%\eeqn
%Conditioned on the daily effect $r_t$, this model is essentially a Gaussian copula.  However, asymptotic dependence is induced by marginalizing over $r_t$.

%\textbf{Hierarchical:}
%Would it be worth exploring something like this?
%If $Z_t(\bs) \sim \mbox{GPD}(1,1,1)$ marginally, then $Y_t(\bs) = \mu(\bs) + \frac{\sigma}{\xi}[Z^(\bs)^\xi - 1]$ is marginally distributed as $\mbox{GPD}[\mu(\bs),\sigma,\xi]$. By defining $Z_t(\bs) = r_t \exp[ W_t(\bs)]$ the $Y$ terms exhibit the same tail dependence as in (\ref{chi}).  After marginalizing over $r_t$, this gives the hierarchical model
%\beqn\label{Y}
%  Y_t(\bs) | W_t(\bs) &\iid& \mbox{GPD}(\mu(\bs)^*, \sigma(\bs)^*, \xi)\\
%  W_t      &\iid&  \mbox{Mat}(\mu_W,\sigma_W,\rho,\nu, \alpha).\nonumber
%\eeqn
%where $\mu(\bs)^* = \mu(\bs) + \frac{ \sigma }{ \xi } \{ \exp[\xi W_t(\bs)] - 1 \}$ and $\sigma(\bs)^* = \sigma \exp[ \xi W_t(\bs) ]$.
%I think this would be an equivalent model to one where we have a different $r_t$ for each site.
%My hunch would be that this would suffer from some kind of asymptotic independence, but would something like this be helpful when dealing with the long-range dependence?

\subsection{Points above a threshold model}\label{s:pot}

So, considering (\ref{fullmodel}), the marginal distribution for the thresholded observations $\Yt_t(\bs)$ is  
\beqn\label{Ytmarg}
  P(\Yt_t(\bs) = T) &=& \Phi\left(\frac{ T - \mu(\bs) }{r_t \sigma_y} \right)\\
  P(\Yt_t(\bs)|\Yt_t(\bs) > T)  &=& \frac{ 1 - \Phi\left(\frac{ Y_t(\bs) - \mu(\bs) }{r_t \sigma_y} \right) }{ 1 - \Phi\left(\frac{ T - \mu(\bs) }{r_t \sigma_y} \right) }\nonumber
\eeqn
where $\Phi(\bZ(\bs))$ is a multivariate normal distribution that follows $\mbox{Mat}(0,\sigma_W,\rho,\nu, \alpha)$. The tail dependence is the same as in (\ref{chi}).  By allowing the location parameter $\mu(\bs)$ to vary spatially, the marginal probability of threshold exceedance vary by location.  We let $\mu(\bs)$ be a Gaussian process, or maybe just a function of spatial covariates like CMAQ?  We could also let $\sigma$ vary spatially to get an even more flexible model.

\section{Computation}\label{s:comp}

There are three main parts to the MCMC: (1) imputing the thresholded $Y_t(\bs)$; (2) updating the model parameters $\Theta=\{r_1,\ldots,r_m,\mu(\bs),\sigma_Y,\rho, \nu, \alpha, \mu_r, \sigma_r, \xi_r\}$; and (3) making spatial predictions.  All three require the joint distribution for complete data given $\Theta$.  Denote $\bY_t = [Y_t(\bs_1),...,Y_t(\bs_n)]^T$, and $Y_t(\bs)=G_{\Theta}[W_t(\bs)]$ defined by solving (\ref{fullmodel}).  The distribution of $Y_t|\Theta$ is the usual multivariate normal distribution
\beqn\label{copula}
   f(Y_t|\Theta) = \phi_n[Y_t|\mu(\bs),\sigma_Y,\rho,\nu_Y, r_1, \ldots, r_m]
\eeqn
where $\phi_n$ is the multivariate normal pdf with $\Matern$ correlation matrix.

\subsection{Imputation}\label{s:impute}

We can use Gibbs sampling to update $Y_t(\bs)$ for observations with $\Yt_t(\bs)=T$.
Given $\Theta$, $Y_t(\bs)$ has truncated normal full conditional with these parameter values.
So we sample $Y_t(\bs)\sim\mbox{TN}_{(-\infty,T)}(a,b^2)$.
Imputation of the $Y_t(\bs)$ terms is done based upon the conditional multivariate normal. 
If we consider the full distribution of $Y_t(\bs)$, then
\beqn
	Y_t(\bs) \sim \mbox{Mat}(\mu(\bs), \sigma_Y, \rho, \nu, \alpha).
\eeqn
where $\sigma_Y = r_t * \sigma_W$.
So, to impute at each location below the threshold, the standard multivariate normal theory holds. 
So, conditional upon all other locations, at location $i$
\beqn
	Y_t(\bs_{(i)}) | Y_t(\bs_{(-i)}) \sim \mbox{TN}(a, b^2)
\eeqn
where
\beqn
	a 	&=& \mu_Y - \Sigma_{12} \Sigma^{-1}_{22}(y_{(-i)}(\bs) - \mu_Y)\\
	b^2	&=& \Sigma_{11} - \Sigma_{12}\Sigma^{-1}_{22}\Sigma_{21}.
\eeqn

\subsection{Parameter updates}
To update $\Theta$ given the current value of the compute data $\bY_1,...,\bY_{n_t}$, we use a standard Metropolis update using the likelihood (\ref{copula}).

\subsection{Spatial prediction}
Given $\bY_t$ the usual Kriging equations give the predictive distribution for $Y_t(\bs^*)$ at prediction location $\bs^*$.

\section{Simulation study}\label{s:sim}
Quantile scores. $T=0.8$ is data that has been thresholded at the 80th sample quantile.
\begin{table}[htdp]
\caption{Quantile Scores for $\alpha=.5$ settings}
\begin{center}
\begin{tabular}{r|rrr|rrr|rrr|}
	\hline
	&&	Setting 1&	&	&Setting 2	&	&	&Setting 3	&	\\
	\hline
	&$\xi=-0.25$	&	&$T=0.8$& $\xi=0.25$ & & $T=0.8$& $\xi=-0.25$&&$T=0.9$\\
	\hline
	&GPD& Gamma & Fixed & GPD & Gamma & Fixed & GPD &Gamma & Fixed\\
	\hline
	0.900 & 0.182 & 0.181 & 0.182	&0.271 & 0.270 & 0.280	& 0.121 & 0.121 & 0.127\\
	\hline
	0.950 & 0.131 & 0.131 & 0.133	&0.202 & 0.202 & 0.202	& 0.095 & 0.095 & 0.095\\
	\hline
	0.970	&	0.098	&	0.098	&	0.105	&	0.154	&	0.154	&	0.157	&	0.075	&	0.074	&	0.077\\
	\hline
0.990	&	0.050	&	0.049	&	0.066	&	0.077	&	0.077	&	0.092	&	0.039	&	0.038	&	0.050\\
	\hline
0.995	&	0.032	&	0.031	&	0.051	&	0.048	&	0.048	&	0.067	&	0.025	&	0.024	&	0.040\\
	\hline
0.999	&	0.012	&	0.011	&	0.032	&	0.015	&	0.015	&	0.037	&	0.008	&	0.008	&	0.026\\
	\hline
\end{tabular}
\end{center}
\label{lbl:quantsim5}
\end{table}%

\begin{table}[htdp]
\caption{Quantile Scores for $\alpha=.9$ settings}
\begin{center}
\begin{tabular}{r|rrr|rrr|rrr|}
	\hline
	&&	Setting 4&	&	&Setting 5	&	&	&Setting 6	&	\\
	\hline
	&$\xi=-0.25$	&	&$T=0.8$& $\xi=0.25$ & & $T=0.8$& $\xi=-0.25$&&$T=0.9$\\
	\hline
	&GPD& Gamma & Fixed & GPD & Gamma & Fixed & GPD &Gamma & Fixed\\
	\hline
0.900	&	0.179	&	0.179	&	0.180	&	0.229	&	0.229	&	0.231	&	0.117	&	0.117	&	0.122\\
	\hline
0.950	&	0.128	&	0.128	&	0.129	&	0.163	&	0.163	&	0.164	&	0.092	&	0.092	&	0.091\\
	\hline
0.970	&	0.095	&	0.095	&	0.098	&	0.119	&	0.119	&	0.123	&	0.071	&	0.071	&	0.072\\
	\hline
0.990	&	0.046	&	0.045	&	0.054	&	0.055	&	0.054	&	0.064	&	0.036	&	0.035	&	0.043\\
	\hline
0.995	&	0.027	&	0.026	&	0.038	&	0.032	&	0.032	&	0.043	&	0.022	&	0.022	&	0.032\\
	\hline
0.999	&	0.008	&	0.008	&	0.019	&	0.008	&	0.008	&	0.021	&	0.007	&	0.006	&	0.019\\
	\hline
\end{tabular}
\end{center}
\label{lbl:quantsim9}
\end{table}%

\section{Data analysis}\label{s:analysis}
Posteriors of $\beta$ terms. $\beta_0$ is intercept, $\beta_1$ is x coordinate, $\beta_2$ is y coordinate, $\beta_3$ is CMAC coordinate.
\begin{table}[htdp]
\caption{Threshold: 0.80 - GPD - Beta Posteriors}
\begin{center}
\begin{tabular}{|c|rrr|rrr|rrr|}
	 &\multicolumn{3}{c|}{GPD}& \multicolumn{3}{c|}{Gamma}& \multicolumn{3}{c|}{Fixed}\\
	\hline
				& 0.025		&	0.500	&	0.975& 0.025		&	0.500	&	0.975& 0.025		&	0.500	&	0.975\\
	\hline
	\hline
	$\beta_0$	& 17.885	&	27.317	&	34.651 & 16.446	&	27.187	&	35.351& 9.801		&	17.565	&	24.322\\
	\hline
	$\beta_1$	& -4.000	&	-1.352	&	1.209& -3.993	&	-1.380	&	1.123& -3.894	&	-1.240	&	1.152\\
	\hline
	$\beta_2$	& -1.092	&	0.694	&	2.516& -1.017	&	0.720	&	2.516& -1.269	&	0.975	&	3.064\\
	\hline
	$\beta_3$	& 0.456		&	0.569	&	0.707& 0.444		&	0.570	&	0.727& 0.631		&	0.726	&	0.834\\
	\hline	
\end{tabular}
\end{center}
\label{lbl:beta80}
\end{table}%

\begin{table}[htdp]
\caption{Threshold: 0.90 - GPD - Beta Posteriors}
\begin{center}
\begin{tabular}{|c|rrr|rrr|rrr|}
	 &\multicolumn{3}{c|}{GPD}& \multicolumn{3}{c|}{Gamma}& \multicolumn{3}{c|}{Fixed}\\
	\hline
			& 0.025		&	0.500	&	0.975& 0.025		&	0.500	&	0.975& 0.025		&	0.500	&	0.975\\
	\hline
	\hline
	$\beta_0$	& 11.377	&	24.969	&	35.848& 12.893	&	25.372	&	37.379& -3.209	&	11.777	&	22.724\\
	\hline
	$\beta_1$	& -4.299	&	0.060	&	5.691& -4.513	&	-0.171	&	4.965& -4.983	&	-0.801	&	4.534\\
	\hline
	$\beta_2$	& -6.198	&	-2.082	&	0.963& -5.644	&	-1.789	&	1.288& -6.164	&	-2.269	&	1.175\\
	\hline
	$\beta_3$	& 0.514		&	0.663	&	0.829& 0.492		&	0.649	&	0.825& 0.696		&	0.845	&	1.028\\
	\hline	
\end{tabular}
\end{center}
\label{lbl:beta90}
\end{table}%

\begin{table}[htdp]
\caption{Threshold: 0.95 - GPD - Beta Posteriors}
\begin{center}
\begin{tabular}{|c|rrr|rrr|rrr|}
	 &\multicolumn{3}{c|}{GPD}& \multicolumn{3}{c|}{Gamma}& \multicolumn{3}{c|}{Fixed}\\
	\hline
			& 0.025		&	0.500	&	0.975& 0.025		&	0.500	&	0.975& 0.025		&	0.500	&	0.975\\
	\hline
	\hline
	$\beta_0$	& -21.936	&	16.245	&	29.874& -14.792	&	11.460	&	32.541& -47.356	&	-8.725	&	12.051\\
	\hline
	$\beta_1$	& 2.948		&	12.280	&	26.863& 3.497		&	13.160	&	27.162& 2.788		&	13.244	&	29.326\\
	\hline
	$\beta_2$	& -14.984	&	-8.785	&	-0.523& -17.402	&	-8.016	&	0.920& -19.640	&	-9.499	&	-0.563\\
	\hline
	$\beta_3$	& 0.599		&	0.773	&	1.164& 0.567		&	0.807	&	1.073& 0.807		&	1.057	&	1.454\\
	\hline	
\end{tabular}
\end{center}
\label{lbl:beta95}
\end{table}%

Table \ref{lbl:quant-5fold} has quantile scores. $T=0.8$ is data that has been thresholded at the 80th sample quantile.
\begin{table}[htdp]
\caption{Quantile Scores for 5-fold cross-validation}
\begin{center}
\begin{tabular}{r|rrr|rrr|rrr|}
	\hline
	&&	$T=0.80$&	&	&$T=0.95$	&	&	&$T=0.99$	&	\\
	\hline
			&GPD& Gamma & Fixed & GPD & Gamma & Fixed & GPD &Gamma & Fixed\\
	\hline
	0.900	&4.870	&4.870	&4.842	&4.925	&4.914	&4.900	&5.501	&5.229	&5.629 \\
	\hline
	0.950	&2.861	&2.861	&2.837	&2.871	&2.865	&2.861	&3.244	&3.077	&3.290 \\
	\hline
	0.970	&1.891	&1.890	&1.881	&1.898	&1.896	&1.891	&2.155	&2.047	&2.175 \\
	\hline
	0.990	&0.742	&0.742	&0.743	&0.749	&0.748	&0.741	&0.853	&0.816	&0.855 \\
	\hline
	0.995	&0.401	&0.400	&0.403	&0.405	&0.404	&0.399	&0.461	&0.442	&0.464 \\
	\hline
	0.999	&0.092	&0.093	&0.093	&0.096	&0.096	&0.094	&0.102	&0.096	&0.106 \\
	\hline
\end{tabular}
\end{center}
\label{lbl:quant-5fold}
\end{table}%

\section{Conclusions}\label{s:con}

\section*{Acknowledgments}

%\section*{Appendix A.1: Derivation of $\tilde{\sigma}(\bs)$}
%Over the threshold, $\Yt_t(\bs) = Y_t(\bs)$. So, to derive the parameters for the density of $\Yt(\bs)$ we need to consider the truncated likelihood of $Y(\bs)$. So
%\begin{align*}
%	F(\Yt(\bs) | \Yt(\bs) > T) &= \frac{ \left[ 1 + \frac{ \xi }{ \sigma^*(\bs)} (Y(\bs)- \mu^*(\bs)) \right]^{-1/\xi} }{ \left[ 1 + \frac{ \xi}{\sigma^*(\bs)} (T - \mu^*(\bs) ) \right]^{-1/\xi}} = \left[ \frac{ 1 + \frac{ \xi }{ \sigma^*(\bs) }(Y(\bs) - \mu^*(\bs) )}{1 + \frac{ \xi }{ \sigma^*(\bs) } (T - \mu^*(\bs) ) }\right]^{-1/\xi}\\
%		&= \left[ \frac{ 1 + \frac{\xi}{\sigma^*(\bs)} (Y(\bs) - T + T - \mu^*(\bs) ) }{ 1 + \frac{ \xi}{\sigma^*(\bs)}(T - \mu^*(\bs) ) } \right]^{-1/\xi}\\
%		&= \left[ 1 + \frac{ \frac{ \xi }{ \sigma^*(\bs)} (Y(\bs) - T ) }{ 1 + \frac{ \xi }{ \sigma^*(\bs) }(T - \mu^*(\bs) )} \right]^{ -1/\xi}\\
%		&= \left[ 1 + \frac{ \xi }{ \tilde{\sigma}(\bs) } (Y(\bs) - T) \right]^{ -1/\xi}
%\end{align*}
%where 
%\begin{align*}
%	\tilde{\sigma}(\bs) = \frac{ \sigma^*(\bs) }{ 1 + \frac{ \xi }{ \sigma^*(\bs) } (T - \mu^*(\bs) ) } = \sigma^*(\bs) \theta(\bs)^{ \xi }.
%\end{align*}
\newpage
\section*{Appendix A.1: Proof that $\lim_{h \rightarrow \infty} \pi(h) = 0$}
Consider two points $\bs_1, \bs_2 \in \calD \subset \calR^2$ with $h = || \bs_1 - \bs_2 ||$, a set of random knots $\bv_1, \ldots, \bv_J$.
Define $g(\bs_i) = \argmin_j ||\bs_i - \bv_j ||$ to be the partition in which $\bs_i$ resides. 
Let $\pi(h) = \Pr[ g(\bs_1) = g(\bs_2) ]$. Then $\pi(h)$ is the probability that $\bv_j$ is the only knot in $(\ballunion \cap \calD)$ where $B_r(\bs)$ is a ball of radius $r > 0$ centered at point $\bs$.
Assume that $g(\bs_1) = g(\bs_2) = j$, and let $a = ||\bs_1 - \bv_j ||$ and $b = ||\bs_2 - \bv_j ||$. 
Then
\beqn
	\pi(h) = \Pr\left[\bv_2, \ldots, \bv_J \in (\ballunion)^C\right ] = \left( 1 - \frac{ | \ballunion | }{ | \calD |} \right)^{J - 1}
\eeqn
where $| \calS |$ is the area of $\calS \cap \calD$.
We know that $\max( |\ballunion |)$ occurs when $B_a(\bs_1) \cap B_b(\bs_2) = 0$ and $\min( | \ballunion | )$ occurs when either $B_a(\bs_1) \subset B_b(\bs_2)$ or $B_b(\bs_2) \subset B_a(\bs_1)$.
Therefore
\beqn
	\left( 1 - \frac{ 2 \pi ( a^2 + b^2) }{ | \calD |} \right)^{J - 1} \le \pi(h) \le \left( 1 - \frac{ 2 \pi \max(a^2,b^2) }{ | \calD |} \right)^{J - 1}.
\eeqn
Then by the triangle inequality, $h \le a + b$, so if $h \rightarrow \infty$, then $\max(a, b) \rightarrow \infty$.
Then because $h \rightarrow \infty$ then $\max(a^2, b^2) \rightarrow \infty$ and 
\beqn
	\lim_{h \rightarrow \infty} \pi(h) \le \lim_{\max(a, b) \rightarrow \infty} \left( 1 - \frac{ 2 \pi \max(a^2, b^2) }{ | \calD |}\right)^{ J - 1 } = \left( 1 - \frac{ | \calD |}{ | \calD |} \right)^{ J - 1 } = 0.
\eeqn
Therefore $\lim_{h \rightarrow \infty} \pi(h) = 0$.

\section*{Appendix A.2: MCMC Details}

\subsection*{Priors}
For a given day
\beqn
	r_{j} &\iid& \mbox{IG}(\xi_r, \sigma_r)\nonumber\\ 
	\sigma_r &\sim& \mbox{Gamma}(0.1, 0.1)\nonumber\\
	\xi_r &\sim& \mbox{Discrete Uniform}(0.5, 30)\nonumber\\
	\bv_{j} &\iid& \mbox{Uniform}(\calD)\nonumber\\
	\mu(\bs) &\sim& \mbox{MVN}(0, \mbox{diag}(10))\nonumber\\
	\log(\rho) &\sim& \mbox{N}(0, 10)\nonumber\\
	\log(\nu) &\sim& \mbox{N}(-1, 1)\nonumber\\
	\alpha &\sim& \mbox{Unif}(0, 1)\nonumber
\eeqn
where $v_j$ are the locations of the spatial knots over $\calD$, $\alpha$ is a parameter controlling the proportion of $r^2_{j}$ that is attributed to the nugget and partial sill. 
If $\alpha = 0$, then $r^2_{j}$ can be entirely attributed to the nugget effect, and if $\alpha = 1$, then $r^2_{tj}$ can be entirely attributed to the partial sill.
We use Gibbs sampling for $r_j, \sigma_r$, and $\mu(\bs)$. 
All other parameters are sampled using a random-walk Metropolis Hastings algorithm.

\bibliographystyle{rss}
\bibliography{Schlather}

\end{document}

### 2014-04-09 pot/LaTeX/pot-annotatedbib.tex
Initial Commit
Starting point: Copy from SVN

\documentclass[11pt]{article}
\usepackage{amssymb, amsthm, amsmath}
\usepackage{bm}
\usepackage{graphicx}
\usepackage[authoryear]{natbib}
\usepackage{bm}
\usepackage{verbatim}
\usepackage{lineno}
\usepackage{times}
\usepackage{soul}
\usepackage{color}

\usepackage[left=1in,top=1in,right=1in]{geometry}
\input{mycommands.sty}

\pdfpageheight 11in
\pdfpagewidth 8.5in
\linespread{1.0}

\begin{document}\linenumbers
\section*{\citep{Azzalini1996}}
Consider $\bY \in \calR^k$ that is distributed as MVN and is independent from $Y_0 \sim N(0, 1)$. Then for
\begin{align}
\left( \begin{array}{c}
    Y_0\\
    \bY
\end{array} \right)
\sim
N_{k + 1} \left\{ 0, \left( 
    \begin{array}{c c}
    1 & 0 \\
    0 & \bPsi
    \end{array}
\right) \right\}
\end{align}
Then, $Z_j = \delta_j |Y_0| + \sqrt{ 1 - \delta_j^2} Y_j, j = 1, \ldots, k$ is skewed normal and its density is
\begin{align}
    f_k(z) = 2 \phi_k (z; \bOmega) \Phi (\balpha^t \bz) \quad \bz \in \calR^k
\end{align}

\section*{\citep{Azzalini1999}}

\section*{\citep{Branco2001}}
Modeling distributions that can account for skewness and heavy tails.

\subsection*{Multivariate elliptical}
Notation: $\bX \sim El_k(\bmu, \bSigma; \phi)$ means that $\bX \in \calR^k$ follows an elliptical distribution with location vector $\bmu \in \calR^k$, a dispersion matrix $\bSigma \in \calR^{k \times k}$ and characteristic function $\phi$.
If the density exists, then it is given by
\beqn
    f(\bx \mid \bmu, \bSigma) = | \bSigma |^{-1/2} g^{(k)} [ ( \bx - \bmu)^T \bSigma^{-1} (\bx - \bmu)]
\eeqn

\subsection*{Multivariate skew elliptical}
Notation: $\bY \sim SE_k(\bmu, \bOmega, \bdelta; \phi)$ means that $\bY \in \calR^k$ follows a skew-elliptical distribution with location vector $\bmu \in \calR^k$, a dispersion matrix $\bSigma \in \calR^{k \times k}$, characteristic function $\phi$ and skewness parameter $\bdelta$.
If the density exists, then it is given by
\beqn
    f_{\bY}(\by) = 2 f_{g^(k)} (\by) F_{g_{q(\by)}} (\blambda^T (\by - \bmu))
\eeqn

\subsection*{Multivariate skew normal distribution}
For a multivariate skew normal distribution, the density function is
\beqn
    f_{\bY}(\by) = 2 \phi_k ( \by; \bmu, \bOmega) \Phi(\blambda^T (\by - \bmu))
\eeqn
where
\beqn
    \blambda^T = \frac{ \bdelta^T \bOmega^{-1}}{(1 - \bdelta^T \bOmega^{-1} \bdelta)^{1/2}}
\eeqn

\subsection*{Multivariate skew $t$ distribution}
For a multivariate skew $t$ distribution, the density function is 
\beqn
    f_{\bY}(\by) = 2f_{\nu, \tau}(\by; \bmu; \bOmega) F_{\nu^*, \tau^*}(\blambda^T(\by - \bmu))
\eeqn

\subsection*{Other densities mentioned}
\begin{itemize}
    \item Skew logistic
    \item Skew stable distribution
    \item Skew exponential power distribution
    \item Skew Pearson Type II distribution
\end{itemize}

\section*{\citep{Sahu2003}}

\section*{\citep{Gupta2004}}
p. 189
The general multivariate skew normal distribution has density function
\beqn
    2 \phi_k (z; \Omega) \Phi (\alpha^{\text{T}} z) \qquad (z \in \calR^k)
\eeqn
where $\phi_k(z; \Omega)$ is a $k$-dimensional process with mean zero, and correlation matrix $\Omega$, $\Phi( \cdot)$ is the $N(0, 1)$ distribution, and $\alpha \in \calR^k$ is a shape term.

\section*{\citep{Allard2007}}

\section*{\citep{Zhang2010}}

\section*{\citep{Minozzo2012}}
p. 164
Let $U_t \sim N(0, 1)$ and $V_t(\bs) \sim MVN$ with mean 0, and variance 1, then 
\beqn
    \bY_t(\bs) = \sigma \delta |U_t| + \sigma \sqrt{1 - \delta^2} V_t(\bs)
\eeqn
follows a

\bibliographystyle{rss}
\bibliography{Schlather}
\end{document}

### 2014-04-09 thresh/LaTeX/threshold-annotatedbib.tex
Initial Commit
Starting point: Copy from SVN

\documentclass[11pt]{article}
\usepackage{amssymb, amsthm, amsmath}
\usepackage{bm}
\usepackage{graphicx}
\usepackage[authoryear]{natbib}
\usepackage{bm}
\usepackage{verbatim}
\usepackage{lineno}
\usepackage{times}
\usepackage{soul}
\usepackage{color}

\usepackage[left=1in,top=1in,right=1in]{geometry}
\input{mycommands.sty}

\pdfpageheight 11in
\pdfpagewidth 8.5in
\linespread{1.0}

\begin{document}\linenumbers
\citep{Gneiting2011}

p. 415 (Table 4) gives suggested threshold weighting for Brier and Quantile scores

\section{Continuous Ranked Probability Scores (CRPS)}
\begin{itemize}
    \item For each of the CRPS versions, they still integrate over all $q$ and $\tau$
    \item For these to be strictly proper, we need a finite first moment ($\xi < 1$)
\end{itemize}

\subsection{Brier scores}
The Brier probability scores is given by 
\begin{align}
    \text{PS}[ F(q), I(y \le q) ] = [ F(q) - I(y \le q) ]^2
\end{align}
where $q$ is a quantile of interest.

The CRPS with Brier score is
\begin{align}
    S(f, y) = \int_{-\infty}^{\infty} \text{PS}[F(q), I(y \le q) ] u(q) \text{d}q
\end{align}
where $u(q)$ is a weighting function (see \ref{sec:weighting}).

Discretized (integral w.r.t. a discrete Stieltjes measure)
\begin{align}
    S(f, y) = \frac{ q_u - q_\ell }{ I - 1 } \sum_{i = 1}^I u(q_i) \text{PS}[ F(q), I(y \le q) ] 
\end{align}
where $(q_\ell, q_u)$ is the range of interest

\subsection{Quantile scores}
The quantile score is given by
\begin{align}
    \text{QS}_\tau [ q(\tau), y ] = 2 \{ I[y \le q(\tau) ] \} [q(\tau) - y]
\end{align}
where $q(\tau)$ is the $\tau$th quantile of the density.

The CRPS with quantile score is
\begin{align}
    S(f, y) = \int_{0}^{1} \text{QS}_\tau (q(\tau), y) v(\tau) \text{d}\tau
\end{align}
where $v(\tau)$ is a weighting function (see \ref{sec:weighting}).

Discretized (integral w.r.t. a discrete Stieltjes measure)
\begin{align}
    S(f, y) = \frac{ 1 }{ J - 1} \sum_{ j = 1 }^{ J-1 } v(\tau_j) \text{QS}_\tau [ F^{-1}(\tau), y ]
\end{align}
where $\tau_j = \frac{ j }{ J }$

\section{Suggested threshold weighting functions} \label{sec:weighting}
This comes from p. 415 - Table 4

\begin{align}
    & \text{Brier Scores} && \text{Quantile Scores}\\
    \text{Tails:} \qquad & u(q) = 1 - \phi_{a, b}(q) / \phi_{a,b}(a) && v(\tau) = (2 \tau - 1)^2\\
    \text{Right Tail:} \qquad & u(q) = \Phi_{a, b}(q) && v(\tau) = \tau^2\\
    \text{Left Tail:} \qquad & u(q) = 1 - \Phi_{a, b} && v(\tau) = (1 - \tau)^2
\end{align}
where $\phi_{a, b}$ and $\Phi_{a, b}$ are the pdf and cdf of a $N(a, b)$ distribution, and $a$ and $b$ are parameters to fine-tune the weighting function.

\bibliographystyle{rss}
\bibliography{threshold}
\end{document}

### 2014-04-09 thresh/LaTeX/threshold.tex
Initial Commit
Starting point: Copy from SVN

\documentclass[11pt]{article}
\usepackage{amssymb, amsthm, amsmath}
\usepackage{bm}
\usepackage{graphicx}
\usepackage[authoryear]{natbib}
\usepackage{bm}
\usepackage{verbatim}
\usepackage{lineno}
\usepackage{times}
\usepackage{soul}
\usepackage{color}

\usepackage[left=1in,top=1in,right=1in]{geometry}
\input{mycommands.sty}
\pdfpageheight 11in
\pdfpagewidth 8.5in
\linespread{1.0}

\begin{document}\linenumbers

\begin{center}
{\Large {\bf Estimation of the threshold using a proper scoring rule}}\\
\today
\end{center}

\section{Introduction}\label{s:intro}
When using the generalized Pareto distribution, the choice of threshold is very important.
Typically, the selection of threshold is done through the use of diagnostic plots.
The goal of this paper is to explore the use of proper scoring rules in order to select an ideal threshold for GPD analysis.

\section{Decision framework}\label{s:framework}
In extremes analysis, the role of the threshold is very important.
If the selected threshold is too low, it can result in high bias when estimating extreme return levels.
If the selected threshold is too high, it can result in high variance for the estimated return levels.
Furthermore if a predictive distribution is reasonable for the data, the $\tau$th quantile of the predictive distribution, $q(\tau)$, should be reasonably close to $\tau$th quantile in the observed data, $q_n(\tau)$.
To select an optimal threshold, we propose two scoring functions.
The first is a modification to continuous ranked probability score (CRPS) \citep{Gneiting2007}. 
The second is an integrated quantile score.
In the traditional framework scoring functions should be maximized, but we use the negation which treats these as loss functions.

\subsection{Continuous ranked probability score}
The CRPS \citep{Gneiting2007} is given by
\beqn
    CRPS = -\int_{-\infty}^{\infty} \left[ F(q) - I(x \le q(\tau)) \right]^2 \dq.
\eeqn
We propose
\beqn
    CRPS_m = \int_{-T^*}^{\infty} \left[ F(q) - I(x \le q(\tau)) \right]^2 \dq.
\eeqn

\subsection{Integrated quantile score}
The Brier score does not account for the closeness of a predicted quantile when assigning weight.
The quantile score \citep{Gneiting2007} is a modification of the Brier score that replaces one of the quadratic terms by the distance between the predicted quantile and the observed value.
The quantile score is given by
\beqn
    S(q, x) = [I(x \le q(\tau)) - \tau] (x - q(\tau)).
\eeqn
We propose
\beqn
    iQS = \int_{\tau^*}^1 (q(\tau) - x) [I(x \le q(\tau)) - \tau] \dtau
\eeqn
When considering the quantile function of the GPD, iQS is only finite when $\xi \le 1$.

\section{Computation}\label{s:comp}

\section{Simulation study}\label{s:sim}

\section{Conclusions}\label{s:con}

\section*{Acknowledgments}

\bibliographystyle{rss}
\bibliography{threshold}

\end{document}

### 2014-04-09 binary/code/mcmc.R

@@ -1,0 +1,621 @@
+#########|#########|#########|#########|#########|#########|#########|#########
+######################################################
+########      THE MAIN MCMC FUNCTION       ###########
+######################################################
+
+
+Bayes_GPD <-function(
+	y,								#data
+	S,								#locations
+	knots=NULL,						#knots
+    X=NULL,							#covariates
+    Sp=NULL, 						#Prediction locations
+    Xp=NULL,						#Prediction covariates
+    thresh=100,						#Threshold value
+    init.beta=c(0,.1, -2.94),		#Initial GPD parameters - logscale, shape, pat
+    init.alpha=.5,					#Initial alpha
+    init.range=1,					#Initial range
+    init.bw=1,						#Initial bandwidth
+    init.logs=NULL,
+    vary=c(F,F,F),					#Varying GPD parameters by location (logsig, xi, pat)
+    pri.mn.range=-2,				#Range
+    pri.sd.range=1,
+    pri.mn.bw=0,					#Bandwidth
+    pri.sd.bw=1,
+    pri.var.a=0.1,					#a and b (gamma parameters for var)
+    pri.var.b=0.1,
+    pri.alpha.a=1,					#a and b (beta parameters for alpha)
+    pri.alpha.b=1,
+    #pri.mn.Phi.alpha=0,
+    #pri.sd.Phi.alpha=0.25,
+    pri.sd.beta=10,					#beta sd
+    pri.mn.gev=c(0,0,-2.5),			    #mean for gev
+    pri.sd.gev=c(1,0.25,0.5),		    #sd for gev params
+    keep.samples=T,					#Retain samples
+    nthreads = 2,
+    iters=200,burn=50,update=50, nthin=1){
+
+    start.time <- proc.time()
+
+    if(is.null(knots)){
+    	knots<-S
+    }
+    if(is.null(X)){
+    	X<-cbind(1,S)
+    }
+
+	y[y <= thresh] <- thresh		#Everything below the threshold = threshold
+
+    #### Some bookkeeping values
+    n <- nrow(y)					#Number of sites
+    nt <- ncol(y)					#Number of days
+    nF <- nrow(knots)				#Number of knots
+    p <- ncol(X)					#Number of covariates
+
+    d<-rdist(S,S)					#Distance between locations
+    diag(d)<-0
[…]

### 2014-04-09 binary/code/plots.R

@@ -1,0 +1,127 @@
+###############################################################################
+#### Basic setup                                                           ####
+###############################################################################
+#setwd('~/svn-repos/dissertation/extreme-value/code')
+#setwd('S:/Documents/Dissertation/extreme-value/code')
+
+library(fields)
+library(emulator)
+library(evd)
+library(POT)
+library(geoR)
+
+load(file="./simdata.RData")
+
+#### Model parameters                
+alpha <- c(.3, .5, .7)  # Spatial correlation 
+                            # (0 = spatial dep, 1 = spatial ind)     
+logrange <- 0         				    # Kernel bandwidth 
+
+for(set in 1:nsets){
+	for(day in 1:nt){
+		file1 <- paste("dataplots/sim1day",day,".pdf", sep="")
+		file2 <- paste("dataplots/sim2day",day,".pdf", sep="")
+		file3 <- paste("dataplots/sim3day",day,".pdf", sep="")
+		dev.new(width=10, height=5)
+		par(mfrow=c(1, 2))
+		z <- matrix(y.1[,day,set], length(s2), length(s2))
+		image.plot(s1, s2, z, 
+			main=bquote(paste(
+				alpha == .(alpha[1]), ", ", xi==.(xi), " I(Z > Thresh)")
+			)
+		)
+		z <- matrix(latent.1[,day,set], length(s2), length(s2))
+		image.plot(s1, s2, z, 
+			main=bquote(paste(
+				alpha == .(alpha[1]), ", ", xi==.(xi), " Latent Process")
+			)
+		)
+		dev.print(file=file1, device=pdf)
+		
+		z <- matrix(y.2[,day,set], length(s2), length(s2))
+		image.plot(s1, s2, z, 
+			main=bquote(paste(
+				alpha == .(alpha[2]), ", ", xi==.(xi), " I(Z > Thresh)")
+			)
+		)
+		z <- matrix(latent.2[,day,set], length(s2), length(s2))
+		image.plot(s1, s2, z, 
+			main=bquote(paste(
+				alpha == .(alpha[2]), ", ", xi==.(xi), " Latent Process")
+			)
+		)
[…]

### 2014-04-09 pot/code/analysis/simstudy/pot3.R

@@ -1,0 +1,144 @@
+#########################################################################
+# A simulation study to determine if a thresholded method improves
+# return-level estimation and threshold exceedance prediction over
+# standard kriging methods
+#
+# data settings:
+#	1 - Gaussian
+#	2 - t
+#	3 - t w/partition (3 knots)
+#	4 - 1/2 Gaussian, 1/2 t
+#	5 - 1/2 Gaussian, 1/2 t w/partition (3 knots)
+#	6 - 1/2 Gaussian (range = 0.40), 1/2 t (range = 0.10)
+#	7 - 1/2 Gaussian (range = 0.10), 1/2 t (range = 0.40)
+#	8 - 90% Gaussian, 10% t
+#	9 - 95% Gaussian, 5% t
+#
+# analysis methods:
+#	1 - Gaussian kriging
+#	2 - t
+#	3 - t w/partition (3 knots)
+#	4 - t (thresh = 0.95)
+#	5 - t w/partition (3 knots, thresh = 0.95)
+#	
+#########################################################################
+
+library(fields)
+library(geoR)
+library(mvtnorm)
+library(evd)
+
+#### Load simdata
+rm(list = ls())
+load(file='./simdata.RData')
+
+setting <- 3
+iters <- 20000; burn <- 10000
+
+pred.1 <- pred.2 <- pred.3 <- pred.4 <- pred.5 <- array(NA, dim=c(ntest, nt, iters-burn, nsets))
+params <- array(NA, dim=c(iters, 6, nsets))
+beta.1 <- beta.2 <- beta.3 <- beta.4 <- beta.5 <- array(NA, dim=c(iters-burn, 3, nsets))
+y.validate <- array(NA, dim=c(ntest, nt, nsets))
+
+outputfile <- paste("pot-", setting, ".RData", sep="")
+set.seed(setting)
+
+start <- proc.time()
+
+for(d in 1:nsets){
+	
+	set.seed(setting * 100 + d)
+	
+	y.d <- y[, , d, setting]
+	obs <- rep(c(T, F), 100)[1:ns]
+	y.o <- y.d[obs, ]
+	x.o <- x[obs, , ]
+	s.o <- s[obs, ]
+	
+	y.validate[, , d] <- y.d[!obs, ]
+	x.p <- x[!obs, , ]
[…]

### 2014-04-09 pot/code/analysis/simstudy/pot4.R

@@ -1,0 +1,144 @@
+#########################################################################
+# A simulation study to determine if a thresholded method improves
+# return-level estimation and threshold exceedance prediction over
+# standard kriging methods
+#
+# data settings:
+#	1 - Gaussian
+#	2 - t
+#	3 - t w/partition (3 knots)
+#	4 - 1/2 Gaussian, 1/2 t
+#	5 - 1/2 Gaussian, 1/2 t w/partition (3 knots)
+#	6 - 1/2 Gaussian (range = 0.40), 1/2 t (range = 0.10)
+#	7 - 1/2 Gaussian (range = 0.10), 1/2 t (range = 0.40)
+#	8 - 90% Gaussian, 10% t
+#	9 - 95% Gaussian, 5% t
+#
+# analysis methods:
+#	1 - Gaussian kriging
+#	2 - t
+#	3 - t w/partition (3 knots)
+#	4 - t (thresh = 0.95)
+#	5 - t w/partition (3 knots, thresh = 0.95)
+#	
+#########################################################################
+
+library(fields)
+library(geoR)
+library(mvtnorm)
+library(evd)
+
+#### Load simdata
+rm(list = ls())
+load(file='./simdata.RData')
+
+setting <- 4
+iters <- 20000; burn <- 10000
+
+pred.1 <- pred.2 <- pred.3 <- pred.4 <- pred.5 <- array(NA, dim=c(ntest, nt, iters-burn, nsets))
+params <- array(NA, dim=c(iters, 6, nsets))
+beta.1 <- beta.2 <- beta.3 <- beta.4 <- beta.5 <- array(NA, dim=c(iters-burn, 3, nsets))
+y.validate <- array(NA, dim=c(ntest, nt, nsets))
+
+outputfile <- paste("pot-", setting, ".RData", sep="")
+set.seed(setting)
+
+start <- proc.time()
+
+for(d in 1:nsets){
+	
+	set.seed(setting * 100 + d)
+	
+	y.d <- y[, , d, setting]
+	obs <- rep(c(T, F), 100)[1:ns]
+	y.o <- y.d[obs, ]
+	x.o <- x[obs, , ]
+	s.o <- s[obs, ]
+	
+	y.validate[, , d] <- y.d[!obs, ]
+	x.p <- x[!obs, , ]
[…]

### 2014-04-09 pot/code/analysis/simstudy/pot6.R

@@ -1,0 +1,144 @@
+#########################################################################
+# A simulation study to determine if a thresholded method improves
+# return-level estimation and threshold exceedance prediction over
+# standard kriging methods
+#
+# data settings:
+#	1 - Gaussian
+#	2 - t
+#	3 - t w/partition (3 knots)
+#	4 - 1/2 Gaussian, 1/2 t
+#	5 - 1/2 Gaussian, 1/2 t w/partition (3 knots)
+#	6 - 1/2 Gaussian (range = 0.40), 1/2 t (range = 0.10)
+#	7 - 1/2 Gaussian (range = 0.10), 1/2 t (range = 0.40)
+#	8 - 90% Gaussian, 10% t
+#	9 - 95% Gaussian, 5% t
+#
+# analysis methods:
+#	1 - Gaussian kriging
+#	2 - t
+#	3 - t w/partition (3 knots)
+#	4 - t (thresh = 0.95)
+#	5 - t w/partition (3 knots, thresh = 0.95)
+#	
+#########################################################################
+
+library(fields)
+library(geoR)
+library(mvtnorm)
+library(evd)
+
+#### Load simdata
+rm(list = ls())
+load(file='./simdata.RData')
+
+setting <- 6
+iters <- 20000; burn <- 10000
+
+pred.1 <- pred.2 <- pred.3 <- pred.4 <- pred.5 <- array(NA, dim=c(ntest, nt, iters-burn, nsets))
+params <- array(NA, dim=c(iters, 6, nsets))
+beta.1 <- beta.2 <- beta.3 <- beta.4 <- beta.5 <- array(NA, dim=c(iters-burn, 3, nsets))
+y.validate <- array(NA, dim=c(ntest, nt, nsets))
+
+outputfile <- paste("pot-", setting, ".RData", sep="")
+set.seed(setting)
+
+start <- proc.time()
+
+for(d in 1:nsets){
+	
+	set.seed(setting * 100 + d)
+	
+	y.d <- y[, , d, setting]
+	obs <- rep(c(T, F), 100)[1:ns]
+	y.o <- y.d[obs, ]
+	x.o <- x[obs, , ]
+	s.o <- s[obs, ]
+	
+	y.validate[, , d] <- y.d[!obs, ]
+	x.p <- x[!obs, , ]
[…]

### 2014-04-09 pot/code/analysis/simstudy/pot7.R

@@ -1,0 +1,144 @@
+#########################################################################
+# A simulation study to determine if a thresholded method improves
+# return-level estimation and threshold exceedance prediction over
+# standard kriging methods
+#
+# data settings:
+#	1 - Gaussian
+#	2 - t
+#	3 - t w/partition (3 knots)
+#	4 - 1/2 Gaussian, 1/2 t
+#	5 - 1/2 Gaussian, 1/2 t w/partition (3 knots)
+#	6 - 1/2 Gaussian (range = 0.40), 1/2 t (range = 0.10)
+#	7 - 1/2 Gaussian (range = 0.10), 1/2 t (range = 0.40)
+#	8 - 90% Gaussian, 10% t
+#	9 - 95% Gaussian, 5% t
+#
+# analysis methods:
+#	1 - Gaussian kriging
+#	2 - t
+#	3 - t w/partition (3 knots)
+#	4 - t (thresh = 0.95)
+#	5 - t w/partition (3 knots, thresh = 0.95)
+#	
+#########################################################################
+
+library(fields)
+library(geoR)
+library(mvtnorm)
+library(evd)
+
+#### Load simdata
+rm(list = ls())
+load(file='./simdata.RData')
+
+setting <- 7
+iters <- 20000; burn <- 10000
+
+pred.1 <- pred.2 <- pred.3 <- pred.4 <- pred.5 <- array(NA, dim=c(ntest, nt, iters-burn, nsets))
+params <- array(NA, dim=c(iters, 6, nsets))
+beta.1 <- beta.2 <- beta.3 <- beta.4 <- beta.5 <- array(NA, dim=c(iters-burn, 3, nsets))
+y.validate <- array(NA, dim=c(ntest, nt, nsets))
+
+outputfile <- paste("pot-", setting, ".RData", sep="")
+set.seed(setting)
+
+start <- proc.time()
+
+for(d in 1:nsets){
+	
+	set.seed(setting * 100 + d)
+	
+	y.d <- y[, , d, setting]
+	obs <- rep(c(T, F), 100)[1:ns]
+	y.o <- y.d[obs, ]
+	x.o <- x[obs, , ]
+	s.o <- s[obs, ]
+	
+	y.validate[, , d] <- y.d[!obs, ]
+	x.p <- x[!obs, , ]
[…]

### 2014-04-09 pot/code/analysis/simstudy/pot8.R

@@ -1,0 +1,144 @@
+#########################################################################
+# A simulation study to determine if a thresholded method improves
+# return-level estimation and threshold exceedance prediction over
+# standard kriging methods
+#
+# data settings:
+#	1 - Gaussian
+#	2 - t
+#	3 - t w/partition (3 knots)
+#	4 - 1/2 Gaussian, 1/2 t
+#	5 - 1/2 Gaussian, 1/2 t w/partition (3 knots)
+#	6 - 1/2 Gaussian (range = 0.40), 1/2 t (range = 0.10)
+#	7 - 1/2 Gaussian (range = 0.10), 1/2 t (range = 0.40)
+#	8 - 90% Gaussian, 10% t
+#	9 - 95% Gaussian, 5% t
+#
+# analysis methods:
+#	1 - Gaussian kriging
+#	2 - t
+#	3 - t w/partition (3 knots)
+#	4 - t (thresh = 0.95)
+#	5 - t w/partition (3 knots, thresh = 0.95)
+#	
+#########################################################################
+
+library(fields)
+library(geoR)
+library(mvtnorm)
+library(evd)
+
+#### Load simdata
+rm(list = ls())
+load(file='./simdata.RData')
+
+setting <- 8
+iters <- 20000; burn <- 10000
+
+pred.1 <- pred.2 <- pred.3 <- pred.4 <- pred.5 <- array(NA, dim=c(ntest, nt, iters-burn, nsets))
+params <- array(NA, dim=c(iters, 6, nsets))
+beta.1 <- beta.2 <- beta.3 <- beta.4 <- beta.5 <- array(NA, dim=c(iters-burn, 3, nsets))
+y.validate <- array(NA, dim=c(ntest, nt, nsets))
+
+outputfile <- paste("pot-", setting, ".RData", sep="")
+set.seed(setting)
+
+start <- proc.time()
+
+for(d in 1:nsets){
+	
+	set.seed(setting * 100 + d)
+	
+	y.d <- y[, , d, setting]
+	obs <- rep(c(T, F), 100)[1:ns]
+	y.o <- y.d[obs, ]
+	x.o <- x[obs, , ]
+	s.o <- s[obs, ]
+	
+	y.validate[, , d] <- y.d[!obs, ]
+	x.p <- x[!obs, , ]
[…]

### 2014-04-09 pot/code/analysis/simstudy/pottroub.R

@@ -1,0 +1,166 @@
+#########################################################################
+# A simulation study to determine if a thresholded method improves
+# return-level estimation and threshold exceedance prediction over
+# standard kriging methods
+#
+# data settings:
+#	1 - Gaussian
+#	2 - t
+#	3 - t w/partition (3 knots)
+#	4 - 1/2 Gaussian, 1/2 t
+#	5 - 1/2 Gaussian, 1/2 t w/partition (3 knots)
+#	6 - 1/2 Gaussian (range = 0.20), 1/2 t (range = 0.05)
+#	7 - 1/2 Gaussian (range = 0.05), 1/2 t (range = 0.20)
+#	8 - 90% Gaussian, 10% t
+#	9 - 95% Gaussian, 5% t
+#
+# analysis methods:
+#	1 - Gaussian kriging
+#	2 - t
+#	3 - t w/partition (3 knots)
+#	4 - t (thresh = 0.95)
+#	5 - t w/partition (3 knots, thresh = 0.95)
+#	
+#########################################################################
+
+library(fields)
+library(geoR)
+library(mvtnorm)
+library(evd)
+
+#### Load simdata
+rm(list = ls())
+load(file='./simdata.RData')
+source(file='../../R/mcmc.R')
+source(file='../../R/auxfunctions.R')
+
+setting <- 3
+iters <- 20000; burn <- 10000
+
+pred.1 <- pred.2 <- pred.3 <- pred.4 <- pred.5 <- array(NA, dim=c(ntest, nt, iters-burn, nsets))
+params <- array(NA, dim=c(iters, 6, nsets))
+beta.1 <- beta.2 <- beta.3 <- beta.4 <- beta.5 <- array(NA, dim=c(iters-burn, 3, nsets))
+y.validate <- array(NA, dim=c(ntest, nt, nsets))
+
+outputfile <- paste("pot-", setting, ".RData", sep="")
+set.seed(setting)
+
+start <- proc.time()
+
+for(d in 1:nsets){
+	
+	set.seed(setting * 100 + d)
+	r.inv.init <- r.inv.t[[setting]][, , d]
+	
+	y.d <- y[, , d, setting]
+	obs <- rep(c(T, F), 100)[1:ns]
[…]

### 2014-04-09 thresh/code/simstudy/plots.R

@@ -1,0 +1,216 @@
+#### Quantile scores
+
+library(evd)
+rm(list=ls())
+
+set.seed(2087)
+xi <- 0.25
+sig <- 1
+n <- 16
+x <- round(rgpd(n, 0, scale=sig, shape=xi), 3)
+x <- sort(x)
+
+tau <- seq(0, 1, 0.001)
+
+quant <- sig / xi * ((1-tau)^(-xi) - 1)
+
+par(mfrow=c(4, 4))
+par(oma=c(2, 2, 2, 0))
+par(mar=c(3, 3, 2, 1))
+for (i in 1:n) {
+	quant.score <- (quant - x[i]) * ((x[i] < quant) - tau)
+	
+	plot(tau, quant.score, type="l", main=paste("x = ", x[i]), xlab="", ylab="")
+
+}
+mtext("tau", side="1", outer=T)
+mtext("quantile scores", side="2", outer=T)
+mtext("tau in (0, 1), xi = 1", side="3", outer=T)
+dev.print(file="./plots/qscores/qscore-xipos-all.pdf", device=pdf)
+
+tau <- seq(0.9, 1, 0.001)
+xi <- 0.25
+sig <- 1
+quant <- sig / xi * ((1-tau)^(-xi) - 1)
+
+par(mfrow=c(4, 4))
+par(oma=c(2, 2, 2, 0))
+par(mar=c(3, 3, 1, 1))
+for (i in 1:n) {
+	quant.score <- (quant - x[i]) * ((x[i] < quant) - tau)
+	
+	plot(tau, quant.score, type="l", main=paste("x = ", x[i]), xlab="", ylab="")
+
+}
+mtext("tau", side="1", outer=T)
+mtext("quantile scores", side="2", outer=T)
+mtext("tau in (0.90, 1), xi = 1", side="3", outer=T)
+dev.print(file="./plots/qscores/qscore-xipos-highquant.pdf", device=pdf)
+
+set.seed(2087)
+xi <- 1
+sig <- 1
+n <- 16
+x <- round(rgpd(n, 0, scale=sig, shape=xi), 3)
+x <- sort(x)
+
+tau <- seq(0, 1, 0.001)
+
+quant <- sig / xi * ((1-tau)^(-xi) - 1)
+
+par(mfrow=c(4, 4))
+par(oma=c(2, 2, 2, 0))
+par(mar=c(3, 3, 2, 1))
+for (i in 1:n) {
[…]

### 2014-04-09 thresh/code/simstudy/setting2-low.R

@@ -1,0 +1,48 @@
+library(evd)
+rm(list=ls())
+load(file='simdata.RData')
+
+setting <- 2
+
+n <- dim(y)[1]
+nsets <- dim(y)[2]
+iters <- 6000; burn <- 2000
+thresh.probs <- seq(0.50, 0.95, 0.05)
+
+nthreshs <- length(thresh.probs)
+outputfile <- paste("thresh-low", setting, ".RData", sep="")
+
+
+# want to store all of the predicted datasets - list that is nsets long
+preds <- beta.prob <- beta.scale <- beta.shape <- list()
+
+for (set in 1:nsets) {
+  cat("Dataset", set, "started \n")
+  tic <- proc.time()
+  y.set <- y[, set, setting]
+  
+  threshs <- quantile(y.set, probs=thresh.probs)
+  
+  preds[[set]] <- array(NA, dim=c(iters, n, nthreshs))
+  beta.prob[[set]] <- beta.scale[[set]] <- beta.shape[[set]] <- array(NA, dim=c(iters, ncol(X), nthreshs))
+  
+  for (t in 1:nthreshs){
+    fit <- Bayes_GPD(y=y.set, X.prob=X, X.sig=X, X.xi=X, 
+                     Xp.prob=X, Xp.sig=X, Xp.xi=X,
+                     thresh=threshs[t], iters=iters, burn=burn, update=1000,
+                     iterplot=F)
+  
+    beta.prob[[set]][, , t] <- fit$beta.prob
+    beta.scale[[set]][, , t] <- fit$beta.sig
+    beta.shape[[set]][, , t] <- fit$beta.xi
+    preds[[set]][, , t] <- fit$yp
+  	cat("  thresh", t, "finished \n")
+  	
+  	save(beta.prob, beta.scale, beta.shape, preds,
+  	     file=outputfile)
+  }
+  
+  elap.time.set <- (proc.time() - tic)[3]
+  avg.time.set  <- elap.time.set / set
+  cat("Dataset", set, "finished ", round(avg.time.set, 2), "per dataset \n")
+}

### 2014-04-10 pot/LaTeX/Schlather.tex
Rough draft for manuscript
Rough draft for manuscript that now includes big picture idea for MCMC
with skew normal.

@@
 
-Our approach is to take the construction of \cite{schlather2002} for a max-stable model for block maxima as a literal model for extreme daily data.
-We will show that his has nice theoretical and computational properties on the daily scale, and gives the popular \cite{schlather2002} as the limit of the block maximum data.
-
 Let $Y_t(\bs)\in \calR$ be the observed value at location $\bs$ on day $t$.  To avoid bias in estimating tail parameters, we model the thresholded data
@@
 We first specify a model for the complete data, $Y_t(\bs)$, and then study the induced model for thresholded data, $\Yt_t(\bs)$.  
-The full data model is given in Section \ref{s:model1} assuming a multivariate normal distribution with a different variance each day.
+The full data model is given in Section \ref{s:model} assuming a multivariate normal distribution with a different variance each day.
 Computationally, the values below the threshold are updated using standard Bayesian missing data methods as described in Section \ref{s:comp}.
@@
 \subsection{Complete data}\label{s:model}
-Consider the model
+Consider the spatial process
 \begin{align} \label{fullmodel}
@@
 \end{align}
-where $u_t(\bs)$ has marginal $N(0, 1)$ distribution, $\delta \in (-1, 1)$ controls skew, and $v_t(\bs)$ is a spatial process with mean zero and variance $\sigma^2(1 - \delta^2)$.
-
-% \subsection{Complete data}\label{s:model1}
-% Consider a zero-mean Gaussian process $\epsilon_t(\bs)$ with $\Matern$ correlation structure.  
-% Consider the model
-% \beq\label{fullmodel}
-%   Y_t(\bs) = \mu(\bs) + r_t W_t(\bs).
-% \eeq
-% where $r_t$ are daily random effects and $W_t$ are spatial processes with mean E$[W_t(\bs)]=0$, variance V$[W_t(\bs)]=1$, and $\Matern$ spatial correlation Cor$[W_t(\bs_1),W_t(\bs_2)]=M(||\bs_1-\bs_2||;\rho,\nu)$ with spatial range $\rho>0$, smoothness $\nu>0$.
-% Let $\alpha$ be the proportion of the total variance, $\sigma^2_W$, that is accounted for by the partial sill of the spatial process.
-% Then we denote the spatial $\Matern$ process model for $W_t$ as $W_t\sim \mbox{Mat}(0,1,\rho,\nu, \alpha)$.
-% So for a single day, 
-% \beq \label{eq:singleday}
-% Y_t(\bs)|r_t \sim \mbox{Mat}[\mu(\bs), r_t, \rho, \nu, \alpha].
-% \eeq
-% We consider two different random effects distributions, $r_t \iid \mbox{GPD}(0, \sigma_r, \xi_r)$ and $r_t \iid \mbox{Inv. Gamma}(\xi_r, \sigma_r)$.
-
-%\subsubsection{Special case}\label{s.powerlaw}
-%Consider a special case where $r_t \iid GPD(\sigma_r/\xi_r, \sigma_r, \xi_r)$. Then $r_t$ follows a power law distribution. 
-%The algebra works out and we can use an IG kernel to marginalize out the $r_t$ terms. 
-%Similar results come from letting $r_t \iid GEV(\sigma_r/\xi_r, \sigma_r, \xi_r)$.
-%Otherwise, I don't really have a good place to start for marginalizing over the random effect term. 
-%After marginalizing out $r_t$ in this special case, we are left with (I'll add in the verification to the appendix)
-%\beq
-%f_Z(z) = \left[ \frac{ 1 }{ 2\sigma^2_W }(\bZ^T \Sigma^{-1} \bZ)\right]^{-\frac{ n }{ 2}} \Gamma\left(\frac{ n }{ 2 } \right) \left[ \frac{\xi_r}{\sigma_r} \right]^{-1/\xi_r - 1}  (2\pi)^{ -n/2} \sigma_W^{-n} \left| \Sigma \right|^{-1}\exp\left\{ -\frac{ 1 }{ 2 \sigma^2_W } \left(\bZ^T \Sigma^{-1} \bZ\right) \right\}
-%\eeq
-%which looks strangely like a GEV distribution.
-%In fact, marginally at each site, I'm pretty sure this will follow a univariate GEV distribution.
-%Also, to figure out the bivariate joint density from this distribution shouldn't all that challenging either.
-%It seems like there's a relationship between the GEV and the Generalized Gamma Distribution, but as of right now, I can only get them to match up when we have the relationship that $\mu_r = \frac{ \xi_r }{ \sigma_r }$.
-%The pdf for the Generalized Gamma Distribution is
-%\beq \label{eq:ggamma}
-% \frac{ p / a^d }{ \Gamma(d/p) } x^{ d-1 }\exp\{ -(x/a)^p \}
-%\eeq
-%and I also found a four-parameter version that gives a shift parameter with density
-%\beq
-% \frac{ \theta^\alpha c }{\Gamma(\alpha) } (x - \gamma)^{ c \alpha - 1} \exp\{ -\theta(x - \gamma)^c \}
-%\eeq
-%which is equivalent to (\ref{eq:ggamma}) where $\alpha = d/p$, $\theta=(1/a)^p$ and $c=p$
-%Do you think I should continue pursuing this, or start focusing on modifying the MCMC now?
-
-\subsubsection{GPD random effect}
-Let $r_t \iid \mbox{GPD}(0, \sigma_r, \xi_r)$
-
-\subsubsection{Power law random effect}
-Let $r_t \iid \mbox{GPD}(\sigma_r/\xi_r, \sigma_r, \xi_r)$. Then by marginalizing out the $r_t$ terms, we find that 
-\beq
-	f_Y(y) = \frac{ 1 }{ \sigma_r }\left( \frac{ \xi_r }{ \sigma_r } \right)^{ -1/\xi_r - 1 } (2\pi)^{ -\frac{ n }{ 2 } }\beta^{ -\left( \frac{ 1 }{ \xi_r } + \frac{ n }{ 2 } \right) } \gamma\left( \frac{ 1 }{ \xi_r } + \frac{ n }{ 2 }, \frac{ \xi_r }{ \sigma_r } \beta \right)
-\eeq
-where $\beta = \frac{ 1 }{ 2 } (\bY - \bmu)^T \Sigma^{-1} (\bY - \bmu)$ and $\gamma(s, x)$ is the lower incomplete gamma function. 
-We also find that marginally, 
-\beqn
-	f(y) = \frac{ 1 }{ \sigma_r }\left( \frac{ \xi_r }{ \sigma_r } \right)^{ -1/\xi_r - 1 } (2\pi)^{ -\frac{ 1 }{ 2 } }\beta^{ -\left( \frac{ 1 }{ \xi_r } + \frac{ 1 }{ 2 } \right) } \gamma\left( \frac{ 1 }{ \xi_r } + \frac{ 1 }{ 2 }, \frac{ \xi_r }{ \sigma_r } \beta \right)
-\eeqn
-where $\beta = \frac{ (y - \mu)^2 }{ 2 }$.
-
-\subsubsection{Gamma random effect}
-	Let $r_t \iid \mbox{IG}(\xi_r, \sigma_r)$.
-Then after marginalizing out the random effect, we find that $Y$ follows a multivariate T distribution with $2 \xi_r$ degrees of freedom, $\mu = \mu(\bs)$, and $\widehat{\Sigma} = \frac{ \sigma_r }{ \xi_r }{ \Sigma }$ where $\Sigma$ is the correlation matrix from (\ref{eq:singleday}).
-Then, for a single location 
-\beqn\label{marg}
-	P[Y_t(\bs) > y] &=& 1 - F(y)
-\eeqn
-where $F(y)$ is the distribution function for a t-distribution with $2 \xi_r$ degrees of freedom and $\mu = \mu$.
-The joint is 
-\beqn\label{joint}
-	P[Y_t(\bs_1) > y, Y_t(\bs_2) > y] &=& 1 - F(y) - F(y) + F_{12}(y) 
-\eeqn
-where $F_{12}(y)$ is the multivariate t-distribution described above.
-In this model,
-\beqn\label{chi}
-	P[Y_t(\bs_1) > y | Y_t(\bs_2) > y] &=& \frac{ 1 - F(y) - F(y) + F_{12}(y) }{ 1 - F(y) }\nonumber\\
-	&=& 1 - \frac{ F(y) - F_{12}(y) }{ 1 - F(y) } = \chi(h)
-\eeqn
-where $h = ||\bs_1 - \bs_2||$. It's not entirely obvious here, but the $h$ comes in from the $\widehat{\Sigma}$ in the t-distribution.
-%From here on out is from the old version of the paper.
-%Then if $\mu_W$ and $\sigma_W$ satisfy $E[\exp(W_t(\bs)] = \exp(\mu_W + \sigma_W^2/2) = 1$,
-%\beqn\label{marg}
-%  P[Z_t(\bs)>z] &=& P\left\{r_t>z\exp[-W_t(\bs)]\right\}\\
-%  &=& \mbox{E}\{\exp[W_t(s)]\}/z = 1/z,\nonumber
-%\eeqn
-%so $Z_t(\bs)$ has GPD(1,1,1) marginal.
-%
-%The joint is
-%\beqn\label{joint}
-%  P[Z_t(\bs_1)>z,Z_t(\bs_2)>z] &=& P\left\{r_t>z\exp[-W_t(\bs_1)],r_t>z\exp[-W_t(\bs_2)]\right\}\\
-%  &=& P\left\{r_t>z\exp[-\min\{W_t(\bs_1),W_t(\bs_2)\}]\right\}\nonumber\\
-%  &=& \mbox{E}\{\exp[\min\{W_t(\bs_1),W_t(\bs_2)\}]\}/z = \chi(||\bs_1-\bs_2||)/z,\nonumber
-%\eeqn
-%where $\chi(||\bs_1-\bs_2||) = \mbox{E}\{\exp[\min\{W_t(\bs_1),W_t(\bs_2)\}]\}$.  In this model,
-%\beq\label{chi}
-%  P[Z_t(\bs_1)>z|Z_t(\bs_2)>z] = \frac{P[Z_t(\bs_1)>z,Z_t(\bs_2)>z]}{P[Z_t(\bs_2)>z]} = \chi(||\bs_1-\bs_2||).
-%\eeq
-%We'll have to work it out, but I'm guessing this is the Schlather model's extremal coefficient.
-
-\subsubsection{Addressing long-range dependence}
-In (\ref{chi}), $\chi(h)$ does not decrease to zero as distance $h=||\bs_1-\bs_2||$ goes to infinity. 
-This is because all spatial locations share a common random effect $r_t$, and thus locations separated by very large distances are still dependent.  
-%Can we address this as in the Brown-Resnick version of the max-stable process?  Let's start by ignoring this, but keep in the back of our minds.
-
-To remove long-range dependence, we allow the random effect $r_t$ to vary by subregion.  
-We partition the spatial domain into $J$ subregions and allow the random effects to vary by subregion, $r_t(\bs) = r_{tj}$ if $\bs$ is in subregion $j$. 
-The partitioning is treated as a random process.
-Let $\bv_{t1},...,\bv_{tJ}$ be spatial knots and assign
-\beqn\label{partition}
-  r_t(\bs) = r_{tj} \mbox{\ \ if \ \ } j=\argmin_l\{||\bs-\bv_{tl}||\}.
-\eeqn 
-The knots $\bv_{t1},...,\bv_{tJ}$ are given a homogeneous Poisson process and $r_{tj}\iid \mbox{IG}(\xi_r, \sigma_r)$.
-Denote $\pi(h)$ as the probability that two points separated by distance $h$ are in the same subregion.
-
-Given that two locations are in the same subdomain, their joint survival function is (\ref{marg}).  If not,
-\beqn\label{marg2}
-  P[Y_t(\bs_1) > y, Y_t(\bs_2) > y ] &=& [1 - F(y)][ 1 - F(y)].
-\eeqn
-Combining both cases gives
-\beqn\label{chi2}
-  P[Y_t(\bs_1)>y|Y_t(\bs_2)>y] = [1 - \pi(h)][1 - F(y)] + \pi(h)\left[1 - \frac{ F(y) - F_{12}(y) }{ 1 - F(y) } \right].
-\eeqn
+where $u_t(\bs) = u_{tl}$ if $s \in P_{tl}$ where $P_{t1}, \ldots, P_{tL}$ form a partition, and $u_{tl} \iid N(0, 1)$, $\delta \in (-1, 1)$ controls skew, and $v_t(\bs)$ is a spatial process with mean zero and variance $\sigma^2(1 - \delta^2)$.
+Then $Y_t(\bs)$ is skew normal within each partition \citep{Minozzo2012}.
+We model this with a Bayesian hierarchical model as follows.
+Let $w_{t1}, \ldots, w_{tL}$ be partition centers so that $P_{tl}$ includes all spatial locations $\bs$ that are within the partition.
 Then
-\beqn\label{chi2}
-  \lim_{y\rightarrow \infty}P[Y_t(\bs_1)>y|Y_t(\bs_2)>y] = \pi(h) \chi(h),
-\eeqn
-and 
-\beqn\label{chi3}
-  \lim_{y\rightarrow \infty , h\rightarrow \infty}P[Y_t(\bs_1)>y|Y_t(\bs_2)>y] = 0 
-\eeqn
-if $\lim_{h\rightarrow \infty}\pi(h) = 0 $. 
-A proof of this is given in Appendix A.1
-Under this model, (\ref{chi2}) shows there is asymptotic dependence, and (\ref{chi3}) shows that long-range spatial dependence is eliminated.
-
-%\subsection{Complete data - general GPD marginals}\label{s:model2}
-%Returning to the complete data, $Y_t(\bs)$, if $Z_t(\bs)\sim$ GPD(1,1,1), then let $Y_t(\bs) = \mu(\bs) + \sigma g_{\xi}[Z_t(\bs)]$, where $g_{\xi}(z) = (z^{\xi}-1)/\xi$.
-After this transformation, the marginal distribution of $Y_t$ is
-%\beqn\label{Ymarg}
-%	Y_t(\bs) &\sim& \mbox{GPD}[\mu^*(\bs),\sigma^*(\bs),\xi]
-%\eeqn
-%where $\mu(\bs)^* = \mu(\bs) + \frac{ \sigma }{ \xi } \{ \exp[\xi W_t(\bs)] - 1 \}$ and $\sigma(\bs)^* = \sigma \exp[ \xi W_t(\bs) ]$.
-This has the same tail dependence as in (\ref{chi}).
-So, the final complete data model is
-%\beqn\label{Y}
-%  Y_t(\bs) &=& \mu(\bs) + \sigma g_{\xi}\left\{r_t\exp[W_t(\bs)]\right\}\\
-%  r_t      &\iid&  \mbox{GPD}(1,1,1)\nonumber\\
-%  W_t      &\iid&  \mbox{Mat}(\mu_W,\sigma_W,\rho,\nu, \alpha).\nonumber
-%\eeqn
-%Conditioned on the daily effect $r_t$, this model is essentially a Gaussian copula.  However, asymptotic dependence is induced by marginalizing over $r_t$.
-
-%\textbf{Hierarchical:}
-%Would it be worth exploring something like this?
-%If $Z_t(\bs) \sim \mbox{GPD}(1,1,1)$ marginally, then $Y_t(\bs) = \mu(\bs) + \frac{\sigma}{\xi}[Z^(\bs)^\xi - 1]$ is marginally distributed as $\mbox{GPD}[\mu(\bs),\sigma,\xi]$.
-By defining $Z_t(\bs) = r_t \exp[ W_t(\bs)]$ the $Y$ terms exhibit the same tail dependence as in (\ref{chi}).
-After marginalizing over $r_t$, this gives the hierarchical model
-%\beqn\label{Y}
-%  Y_t(\bs) | W_t(\bs) &\iid& \mbox{GPD}(\mu(\bs)^*, \sigma(\bs)^*, \xi)\\
-%  W_t      &\iid&  \mbox{Mat}(\mu_W,\sigma_W,\rho,\nu, \alpha).\nonumber
-%\eeqn
-%where $\mu(\bs)^* = \mu(\bs) + \frac{ \sigma }{ \xi } \{ \exp[\xi W_t(\bs)] - 1 \}$ and $\sigma(\bs)^* = \sigma \exp[ \xi W_t(\bs) ]$.
-%I think this would be an equivalent model to one where we have a different $r_t$ for each site.
-%My hunch would be that this would suffer from some kind of asymptotic independence, but would something like this be helpful when dealing with the long-range dependence?
-
-\subsection{Points above a threshold model}\label{s:pot}
-
-So, considering (\ref{fullmodel}), the marginal distribution for the thresholded observations $\Yt_t(\bs)$ is  
-\beqn\label{Ytmarg}
-  P(\Yt_t(\bs) = T) &=& \Phi\left(\frac{ T - \mu(\bs) }{r_t \sigma_y} \right)\\
-  P(\Yt_t(\bs)|\Yt_t(\bs) > T)  &=& \frac{ 1 - \Phi\left(\frac{ Y_t(\bs) - \mu(\bs) }{r_t \sigma_y} \right) }{ 1 - \Phi\left(\frac{ T - \mu(\bs) }{r_t \sigma_y} \right) }\nonumber
-\eeqn
-where $\Phi(\bZ(\bs))$ is a multivariate normal distribution that follows $\mbox{Mat}(0,\sigma_W,\rho,\nu, \alpha)$.
-The tail dependence is the same as in (\ref{chi}).
-By allowing the location parameter $\mu(\bs)$ to vary spatially, the marginal probability of threshold exceedance vary by location.
-We let $\mu(\bs)$ be a Gaussian process, or maybe just a function of spatial covariates like CMAQ?
-We could also let $\sigma$ vary spatially to get an even more flexible model.
+\begin{align} \label{eq:hier}
+    Y_t(\bs) \mid \Theta &= \mu_t(\bs) + v_t(\bs) \\
+    \mu_t(\bs) &= X_t(\bs) \beta + \sigma \delta | u_{tl} |
+\end{align}
+where $l = \argmin_j ||\bs - w_j ||$ and $\Theta = \{ u_{t1}, \ldots, u_{tL}, w_{t1}, \ldots, w_{tL}, \beta, \rho, \nu, \sigma \}$ are the random effects, knot locations, and parameters for the mean, and spatial covariance.
 
 \section{Computation}\label{s:comp}
-
-There are three main parts to the MCMC: (1) imputing the thresholded $Y_t(\bs)$; (2) updating the model parameters $\Theta=\{r_1,\ldots,r_m,\mu(\bs),\sigma_Y,\rho, \nu, \alpha, \mu_r, \sigma_r, \xi_r\}$; and (3) making spatial predictions.
-All three require the joint distribution for complete data given $\Theta$.
-Denote $\bY_t = [Y_t(\bs_1),...,Y_t(\bs_n)]^T$, and $Y_t(\bs)=G_{\Theta}[W_t(\bs)]$ defined by solving (\ref{fullmodel}).
-The distribution of $Y_t|\Theta$ is the usual multivariate normal distribution
-\beqn\label{copula}
-   f(Y_t|\Theta) = \phi_n[Y_t|\mu(\bs),\sigma_Y,\rho,\nu_Y, r_1, \ldots, r_m]
-\eeqn
-where $\phi_n$ is the multivariate normal pdf with $\Matern$ correlation matrix.
+The MCMC for this model is fairly straightforward.
+First, we impute values below the threshold.
+Then, we update $\Theta$ using random walk MH or Gibbs sampling when appropriate.
+Finally, we make spatial predictions.
+Each requires the joint distribution for the complete data given $\Theta$.
+As defined in \ref{eq:hier}, the distribution of $Y_t(\bs) \mid \Theta$ is the usual multivariate normal distribution with a \Matern spatial covariance structure.
 
 \subsection{Imputation}\label{s:impute}
+We can use Gibbs sampling to update $\Yt_t(\bs)$ for observations that are below $T$, the thresholded value.
+Given $\Theta$, $Y_t(\bs)$ has truncated normal full conditional with these parameter values.
+So we sample $Y_t(\bs) \sim \text{TN}_{(-\infty, T)}$
 
-We can use Gibbs sampling to update $Y_t(\bs)$ for observations with $\Yt_t(\bs)=T$.
-Given $\Theta$, $Y_t(\bs)$ has truncated normal full conditional with these parameter values.
-So we sample $Y_t(\bs)\sim\mbox{TN}_{(-\infty,T)}(a,b^2)$.
-Imputation of the $Y_t(\bs)$ terms is done based upon the conditional multivariate normal. 
-If we consider the full distribution of $Y_t(\bs)$, then
-\beqn
-	Y_t(\bs) \sim \mbox{Mat}(\mu(\bs), \sigma_Y, \rho, \nu, \alpha).
-\eeqn
-where $\sigma_Y = r_t * \sigma_W$.
-So, to impute at each location below the threshold, the standard multivariate normal theory holds. 
-So, conditional upon all other locations, at location $i$
-\beqn
-	Y_t(\bs_{(i)}) | Y_t(\bs_{(-i)}) \sim \mbox{TN}(a, b^2)
-\eeqn
-where
-\beqn
-	a 	&=& \mu_Y - \Sigma_{12} \Sigma^{-1}_{22}(y_{(-i)}(\bs) - \mu_Y)\\
-	b^2	&=& \Sigma_{11} - \Sigma_{12}\Sigma^{-1}_{22}\Sigma_{21}.
-\eeqn
+\subsection{Parameter updates}\label{s:params}
+To update $\Theta$ given the current value of the complete data $\bY_1, \ldots, \bY_T$, we use a standard Metropolis update.
 
-\subsection{Parameter updates}
-To update $\Theta$ given the current value of the compute data $\bY_1,...,\bY_{n_t}$, we use a standard Metropolis update using the likelihood (\ref{copula}).
-
-\subsection{Spatial prediction}
-Given $\bY_t$ the usual Kriging equations give the predictive distribution for $Y_t(\bs^*)$ at prediction location $\bs^*$.
-
-\section{Simulation study}\label{s:sim}
-Quantile scores. $T=0.8$ is data that has been thresholded at the 80th sample quantile.
-\begin{table}[htdp]
-\caption{Quantile Scores for $\alpha=.5$ settings}
-\begin{center}
-\begin{tabular}{r|rrr|rrr|rrr|}
-	\hline
-	&&	Setting 1&	&	&Setting 2	&	&	&Setting 3	&	\\
-	\hline
-	&$\xi=-0.25$	&	&$T=0.8$& $\xi=0.25$ & & $T=0.8$& $\xi=-0.25$&&$T=0.9$\\
-	\hline
-	&GPD& Gamma & Fixed & GPD & Gamma & Fixed & GPD &Gamma & Fixed\\
-	\hline
-	0.900 & 0.182 & 0.181 & 0.182	&0.271 & 0.270 & 0.280	& 0.121 & 0.121 & 0.127\\
-	\hline
-	0.950 & 0.131 & 0.131 & 0.133	&0.202 & 0.202 & 0.202	& 0.095 & 0.095 & 0.095\\
-	\hline
-	0.970	&	0.098	&	0.098	&	0.105	&	0.154	&	0.154	&	0.157	&	0.075	&	0.074	&	0.077\\
-	\hline
-0.990	&	0.050	&	0.049	&	0.066	&	0.077	&	0.077	&	0.092	&	0.039	&	0.038	&	0.050\\
-	\hline
-0.995	&	0.032	&	0.031	&	0.051	&	0.048	&	0.048	&	0.067	&	0.025	&	0.024	&	0.040\\
-	\hline
-0.999	&	0.012	&	0.011	&	0.032	&	0.015	&	0.015	&	0.037	&	0.008	&	0.008	&	0.026\\
-	\hline
-\end{tabular}
-\end{center}
-\label{lbl:quantsim5}
-\end{table}%
-
-\begin{table}[htdp]
-\caption{Quantile Scores for $\alpha=.9$ settings}
-\begin{center}
-\begin{tabular}{r|rrr|rrr|rrr|}
-	\hline
-	&&	Setting 4&	&	&Setting 5	&	&	&Setting 6	&	\\
-	\hline
-	&$\xi=-0.25$	&	&$T=0.8$& $\xi=0.25$ & & $T=0.8$& $\xi=-0.25$&&$T=0.9$\\
-	\hline
-	&GPD& Gamma & Fixed & GPD & Gamma & Fixed & GPD &Gamma & Fixed\\
-	\hline
-0.900	&	0.179	&	0.179	&	0.180	&	0.229	&	0.229	&	0.231	&	0.117	&	0.117	&	0.122\\
-	\hline
-0.950	&	0.128	&	0.128	&	0.129	&	0.163	&	0.163	&	0.164	&	0.092	&	0.092	&	0.091\\
-	\hline
-0.970	&	0.095	&	0.095	&	0.098	&	0.119	&	0.119	&	0.123	&	0.071	&	0.071	&	0.072\\
-	\hline
-0.990	&	0.046	&	0.045	&	0.054	&	0.055	&	0.054	&	0.064	&	0.036	&	0.035	&	0.043\\
-	\hline
-0.995	&	0.027	&	0.026	&	0.038	&	0.032	&	0.032	&	0.043	&	0.022	&	0.022	&	0.032\\
-	\hline
-0.999	&	0.008	&	0.008	&	0.019	&	0.008	&	0.008	&	0.021	&	0.007	&	0.006	&	0.019\\
-	\hline
-\end{tabular}
-\end{center}
-\label{lbl:quantsim9}
-\end{table}%
+\subsection{Spatial prediction}\label{s:pred}
+Given $\bY_t$ the usual Kriging equations give the predictive distribution for $Y_t(\bs^*)$ at prediction location $(\bs^*)$
 
 \section{Data analysis}\label{s:analysis}
-Posteriors of $\beta$ terms. $\beta_0$ is intercept, $\beta_1$ is x coordinate, $\beta_2$ is y coordinate, $\beta_3$ is CMAC coordinate.
-\begin{table}[htdp]
-\caption{Threshold: 0.80 - GPD - Beta Posteriors}
-\begin{center}
-\begin{tabular}{|c|rrr|rrr|rrr|}
-	 &\multicolumn{3}{c|}{GPD}& \multicolumn{3}{c|}{Gamma}& \multicolumn{3}{c|}{Fixed}\\
-	\hline
-				& 0.025		&	0.500	&	0.975& 0.025		&	0.500	&	0.975& 0.025		&	0.500	&	0.975\\
-	\hline
-	\hline
-	$\beta_0$	& 17.885	&	27.317	&	34.651 & 16.446	&	27.187	&	35.351& 9.801		&	17.565	&	24.322\\
-	\hline
-	$\beta_1$	& -4.000	&	-1.352	&	1.209& -3.993	&	-1.380	&	1.123& -3.894	&	-1.240	&	1.152\\
-	\hline
-	$\beta_2$	& -1.092	&	0.694	&	2.516& -1.017	&	0.720	&	2.516& -1.269	&	0.975	&	3.064\\
-	\hline
-	$\beta_3$	& 0.456		&	0.569	&	0.707& 0.444		&	0.570	&	0.727& 0.631		&	0.726	&	0.834\\
-	\hline	
-\end{tabular}
-\end{center}
-\label{lbl:beta80}
-\end{table}%
-
-\begin{table}[htdp]
-\caption{Threshold: 0.90 - GPD - Beta Posteriors}
-\begin{center}
-\begin{tabular}{|c|rrr|rrr|rrr|}
-	 &\multicolumn{3}{c|}{GPD}& \multicolumn{3}{c|}{Gamma}& \multicolumn{3}{c|}{Fixed}\\
-	\hline
-			& 0.025		&	0.500	&	0.975& 0.025		&	0.500	&	0.975& 0.025		&	0.500	&	0.975\\
-	\hline
-	\hline
-	$\beta_0$	& 11.377	&	24.969	&	35.848& 12.893	&	25.372	&	37.379& -3.209	&	11.777	&	22.724\\
-	\hline
-	$\beta_1$	& -4.299	&	0.060	&	5.691& -4.513	&	-0.171	&	4.965& -4.983	&	-0.801	&	4.534\\
-	\hline
-	$\beta_2$	& -6.198	&	-2.082	&	0.963& -5.644	&	-1.789	&	1.288& -6.164	&	-2.269	&	1.175\\
-	\hline
-	$\beta_3$	& 0.514		&	0.663	&	0.829& 0.492		&	0.649	&	0.825& 0.696		&	0.845	&	1.028\\
-	\hline	
-\end{tabular}
-\end{center}
-\label{lbl:beta90}
-\end{table}%
-
-\begin{table}[htdp]
-\caption{Threshold: 0.95 - GPD - Beta Posteriors}
-\begin{center}
-\begin{tabular}{|c|rrr|rrr|rrr|}
-	 &\multicolumn{3}{c|}{GPD}& \multicolumn{3}{c|}{Gamma}& \multicolumn{3}{c|}{Fixed}\\
-	\hline
-			& 0.025		&	0.500	&	0.975& 0.025		&	0.500	&	0.975& 0.025		&	0.500	&	0.975\\
-	\hline
-	\hline
-	$\beta_0$	& -21.936	&	16.245	&	29.874& -14.792	&	11.460	&	32.541& -47.356	&	-8.725	&	12.051\\
-	\hline
-	$\beta_1$	& 2.948		&	12.280	&	26.863& 3.497		&	13.160	&	27.162& 2.788		&	13.244	&	29.326\\
-	\hline
-	$\beta_2$	& -14.984	&	-8.785	&	-0.523& -17.402	&	-8.016	&	0.920& -19.640	&	-9.499	&	-0.563\\
-	\hline
-	$\beta_3$	& 0.599		&	0.773	&	1.164& 0.567		&	0.807	&	1.073& 0.807		&	1.057	&	1.454\\
-	\hline	
-\end{tabular}
-\end{center}
-\label{lbl:beta95}
-\end{table}%
-
-Table \ref{lbl:quant-5fold} has quantile scores. $T=0.8$ is data that has been thresholded at the 80th sample quantile.
-\begin{table}[htdp]
-\caption{Quantile Scores for 5-fold cross-validation}
-\begin{center}
-\begin{tabular}{r|rrr|rrr|rrr|}
-	\hline
-	&&	$T=0.80$&	&	&$T=0.95$	&	&	&$T=0.99$	&	\\
-	\hline
-			&GPD& Gamma & Fixed & GPD & Gamma & Fixed & GPD &Gamma & Fixed\\
-	\hline
-	0.900	&4.870	&4.870	&4.842	&4.925	&4.914	&4.900	&5.501	&5.229	&5.629 \\
-	\hline
-	0.950	&2.861	&2.861	&2.837	&2.871	&2.865	&2.861	&3.244	&3.077	&3.290 \\
-	\hline
-	0.970	&1.891	&1.890	&1.881	&1.898	&1.896	&1.891	&2.155	&2.047	&2.175 \\
-	\hline
-	0.990	&0.742	&0.742	&0.743	&0.749	&0.748	&0.741	&0.853	&0.816	&0.855 \\
-	\hline
-	0.995	&0.401	&0.400	&0.403	&0.405	&0.404	&0.399	&0.461	&0.442	&0.464 \\
-	\hline
-	0.999	&0.092	&0.093	&0.093	&0.096	&0.096	&0.094	&0.102	&0.096	&0.106 \\
-	\hline
-\end{tabular}
-\end{center}
-\label{lbl:quant-5fold}
-\end{table}%
 
@@
 
-%\section*{Appendix A.1: Derivation of $\tilde{\sigma}(\bs)$}
-%Over the threshold, $\Yt_t(\bs) = Y_t(\bs)$. So, to derive the parameters for the density of $\Yt(\bs)$ we need to consider the truncated likelihood of $Y(\bs)$. So
-%\begin{align*}
-%	F(\Yt(\bs) | \Yt(\bs) > T) &= \frac{ \left[ 1 + \frac{ \xi }{ \sigma^*(\bs)} (Y(\bs)- \mu^*(\bs)) \right]^{-1/\xi} }{ \left[ 1 + \frac{ \xi}{\sigma^*(\bs)} (T - \mu^*(\bs) ) \right]^{-1/\xi}} = \left[ \frac{ 1 + \frac{ \xi }{ \sigma^*(\bs) }(Y(\bs) - \mu^*(\bs) )}{1 + \frac{ \xi }{ \sigma^*(\bs) } (T - \mu^*(\bs) ) }\right]^{-1/\xi}\\
-%		&= \left[ \frac{ 1 + \frac{\xi}{\sigma^*(\bs)} (Y(\bs) - T + T - \mu^*(\bs) ) }{ 1 + \frac{ \xi}{\sigma^*(\bs)}(T - \mu^*(\bs) ) } \right]^{-1/\xi}\\
-%		&= \left[ 1 + \frac{ \frac{ \xi }{ \sigma^*(\bs)} (Y(\bs) - T ) }{ 1 + \frac{ \xi }{ \sigma^*(\bs) }(T - \mu^*(\bs) )} \right]^{ -1/\xi}\\
-%		&= \left[ 1 + \frac{ \xi }{ \tilde{\sigma}(\bs) } (Y(\bs) - T) \right]^{ -1/\xi}
-%\end{align*}
-%where 
-%\begin{align*}
-%	\tilde{\sigma}(\bs) = \frac{ \sigma^*(\bs) }{ 1 + \frac{ \xi }{ \sigma^*(\bs) } (T - \mu^*(\bs) ) } = \sigma^*(\bs) \theta(\bs)^{ \xi }.
-%\end{align*}
-\newpage
-\section*{Appendix A.1: Proof that $\lim_{h \rightarrow \infty} \pi(h) = 0$}
-Consider two points $\bs_1, \bs_2 \in \calD \subset \calR^2$ with $h = || \bs_1 - \bs_2 ||$, a set of random knots $\bv_1, \ldots, \bv_J$.
-Define $g(\bs_i) = \argmin_j ||\bs_i - \bv_j ||$ to be the partition in which $\bs_i$ resides. 
-Let $\pi(h) = \Pr[ g(\bs_1) = g(\bs_2) ]$.
-Then $\pi(h)$ is the probability that $\bv_j$ is the only knot in $(\ballunion \cap \calD)$ where $B_r(\bs)$ is a ball of radius $r > 0$ centered at point $\bs$.
-Assume that $g(\bs_1) = g(\bs_2) = j$, and let $a = ||\bs_1 - \bv_j ||$ and $b = ||\bs_2 - \bv_j ||$. 
-Then
-\beqn
-	\pi(h) = \Pr\left[\bv_2, \ldots, \bv_J \in (\ballunion)^C\right ] = \left( 1 - \frac{ | \ballunion | }{ | \calD |} \right)^{J - 1}
-\eeqn
-where $| \calS |$ is the area of $\calS \cap \calD$.
-We know that $\max( |\ballunion |)$ occurs when $B_a(\bs_1) \cap B_b(\bs_2) = 0$ and $\min( | \ballunion | )$ occurs when either $B_a(\bs_1) \subset B_b(\bs_2)$ or $B_b(\bs_2) \subset B_a(\bs_1)$.
-Therefore
-\beqn
-	\left( 1 - \frac{ 2 \pi ( a^2 + b^2) }{ | \calD |} \right)^{J - 1} \le \pi(h) \le \left( 1 - \frac{ 2 \pi \max(a^2,b^2) }{ | \calD |} \right)^{J - 1}.
-\eeqn
-Then by the triangle inequality, $h \le a + b$, so if $h \rightarrow \infty$, then $\max(a, b) \rightarrow \infty$.
-Then because $h \rightarrow \infty$ then $\max(a^2, b^2) \rightarrow \infty$ and 
-\beqn
-	\lim_{h \rightarrow \infty} \pi(h) \le \lim_{\max(a, b) \rightarrow \infty} \left( 1 - \frac{ 2 \pi \max(a^2, b^2) }{ | \calD |}\right)^{ J - 1 } = \left( 1 - \frac{ | \calD |}{ | \calD |} \right)^{ J - 1 } = 0.
-\eeqn
-Therefore $\lim_{h \rightarrow \infty} \pi(h) = 0$.
-
-\section*{Appendix A.2: MCMC Details}
+\section*{Appendix A.1: MCMC Details}
 

### 2014-04-13 pot/LaTeX/Schlather.tex
updated manuscript
Added in include command for the conditional distribution for u | y

@@
 Consider the spatial process
-\begin{align} \label{fullmodel}
+\begin{align} \label{eq:fullmodel}
   Y_t(\bs) &= X_t(\bs) \beta + e_t(\bs)\\
@@
 \subsection{Parameter updates}\label{s:params}
-To update $\Theta$ given the current value of the complete data $\bY_1, \ldots, \bY_T$, we use a standard Metropolis update.
+To update $\Theta$ given the current value of the complete data $\bY_1, \ldots, \bY_T$, we use a standard Gibbs updates for all parameters except for the knot locations which are done using a Metropolis update.
+See Appendix A.1 for details regarding Gibbs sampling and $|u_t(\bs)|$.
 
@@
 
-\section*{Appendix A.1: MCMC Details}
+\section*{Appendix A.1: Half-Normal results}
+\input{halfnormal.tex}
+
+\section*{Appendix A.2: MCMC Details}
 
 \subsection*{Priors}
-For a given day
-\beqn
-	r_{j} &\iid& \mbox{IG}(\xi_r, \sigma_r)\nonumber\\ 
-	\sigma_r &\sim& \mbox{Gamma}(0.1, 0.1)\nonumber\\
-	\xi_r &\sim& \mbox{Discrete Uniform}(0.5, 30)\nonumber\\
-	\bv_{j} &\iid& \mbox{Uniform}(\calD)\nonumber\\
-	\mu(\bs) &\sim& \mbox{MVN}(0, \mbox{diag}(10))\nonumber\\
-	\log(\rho) &\sim& \mbox{N}(0, 10)\nonumber\\
-	\log(\nu) &\sim& \mbox{N}(-1, 1)\nonumber\\
-	\alpha &\sim& \mbox{Unif}(0, 1)\nonumber
-\eeqn
-where $v_j$ are the locations of the spatial knots over $\calD$, $\alpha$ is a parameter controlling the proportion of $r^2_{j}$ that is attributed to the nugget and partial sill. 
-If $\alpha = 0$, then $r^2_{j}$ can be entirely attributed to the nugget effect, and if $\alpha = 1$, then $r^2_{tj}$ can be entirely attributed to the partial sill.
-We use Gibbs sampling for $r_j, \sigma_r$, and $\mu(\bs)$. 
-All other parameters are sampled using a random-walk Metropolis Hastings algorithm.
 

### 2014-04-17 pot/LaTeX/skewnormal.tex
Working on the algebra for the posterior distributions

--- 2014-04-13 pot/LaTeX/Schlather.tex
@@
 We first specify a model for the complete data, $Y_t(\bs)$, and then study the induced model for thresholded data, $\Yt_t(\bs)$.  
-The full data model is given in Section \ref{s:model} assuming a multivariate normal distribution with a different variance each day.
+The full data model is given in Section \ref{s:model} assuming a skew normal distribution with a different variance each day.
 Computationally, the values below the threshold are updated using standard Bayesian missing data methods as described in Section \ref{s:comp}.
+The skew normal representation is from \citep{Minozzo2012} and is the sum of a normal and half-normal random variable.
+\subsection{Half-normal}\label{s:hn}
+Let $u = \xi + \sqrt{\eta} |z|$ where $Z \sim N(0, 1)$.
+Then $U$ follows a half-normal distribution, $U \sim \mbox{HN}(\xi, \eta )$ \citep{Wiper2008}, and the density is given by 
+\begin{align}
+  f_U(u) = \frac{ \sqrt{\pi} }{ \sqrt{2 \eta} } \exp \left( - \frac{ (u-\xi)^2 }{ 2 \eta } \right), \quad u > \xi.
+\end{align}
 
@@
   Y_t(\bs) &= X_t(\bs) \beta + e_t(\bs)\\
-  e_t(\bs) &= \sigma \delta | u_t(\bs) | + v_t(\bs)
+  e_t(\bs) &= u_t(\bs) + v_t(\bs)
 \end{align}
-where $u_t(\bs) = u_{tl}$ if $s \in P_{tl}$ where $P_{t1}, \ldots, P_{tL}$ form a partition, and $u_{tl} \iid N(0, 1)$, $\delta \in (-1, 1)$ controls skew, and $v_t(\bs)$ is a spatial process with mean zero and variance $\sigma^2(1 - \delta^2)$.
+where $u_t(\bs) = \sigma \delta |z_t|$, $z_t(\bs) = z_{tl}$ if $s \in P_{tl}$ where $P_{t1}, \ldots, P_{tL}$ form a partition, and $z_{tl} \iid N(0, 1)$, $\delta \in (-1, 1)$ controls skew, and $v_t(\bs)$ is a spatial process with mean zero and variance $\sigma^2(1 - \delta^2)$.
 Then $Y_t(\bs)$ is skew normal within each partition \citep{Minozzo2012}.
@@
 \begin{align} \label{eq:hier}
-    Y_t(\bs) \mid \Theta &= \mu_t(\bs) + v_t(\bs) \\
-    \mu_t(\bs) &= X_t(\bs) \beta + \sigma \delta | u_{tl} |
+    Y_t(\bs) \mid \Theta &= X_t(\bs) \beta + u_t(\bs) + v_t(\bs) \\
+    u_t(\bs) \mid \Theta &\sim \mbox{Half-Normal}(0, \sigma \delta)\\
+    v_t(\bs) \mid \Theta & \sim \Matern(0, \Sigma)\\
+    \sigma &\sim \mbox{IG}(\alpha, \beta)\\
+    \delta &\sim \mbox{U}(-1, 1)
 \end{align}
-where $l = \argmin_j ||\bs - w_j ||$ and $\Theta = \{ u_{t1}, \ldots, u_{tL}, w_{t1}, \ldots, w_{tL}, \beta, \rho, \nu, \sigma \}$ are the random effects, knot locations, and parameters for the mean, and spatial covariance.
+where $l = \argmin_j ||\bs - w_j ||$, $\Theta = \{u_{t1}, \ldots, u_{tL}, w_{t1}, \ldots, w_{tL}, \beta, \sigma, \delta, \rho, \nu \}$ and $\Sigma$ is a \Matern covariance matrix with variance $\sigma^2 (1 - \delta^2)$
 
@@
 
-\section*{Appendix A.1: Half-Normal results}
-\input{halfnormal.tex}
+\section*{Appendix A.1: Posterior distributions}
+\input{upost.tex}
+\input{betapost.tex}
+\input{sigpost.tex}
 
@@
 \bibliographystyle{rss}
-\bibliography{Schlather}
+\bibliography{skewnormal}
 

### 2014-04-21 pot/LaTeX/skewnormal.tex
basic edits per Brian to manuscript and fixed line numbering issue

@@
 \subsection{Half-normal}\label{s:hn}
-Let $u = \xi + \sqrt{\eta} |z|$ where $Z \sim N(0, 1)$.
+Let $u = \xi + \eta |z|$ where $Z \sim N(0, 1)$.
 Then $U$ follows a half-normal distribution, $U \sim \mbox{HN}(\xi, \eta )$ \citep{Wiper2008}, and the density is given by 
 \begin{align}
-  f_U(u) = \frac{ \sqrt{\pi} }{ \sqrt{2 \eta} } \exp \left( - \frac{ (u-\xi)^2 }{ 2 \eta } \right), \quad u > \xi.
+  f_U(u) = \frac{ \sqrt{\pi} }{ \sqrt{2 \eta^2} } \exp \left( - \frac{ (u-\xi)^2 }{ 2 \eta^2 } \right), \quad u > \xi.
 \end{align}
@@
   Y_t(\bs) &= X_t(\bs) \beta + e_t(\bs)\\
-  e_t(\bs) &= u_t(\bs) + v_t(\bs)
+  e_t(\bs) &= \delta z_t(\bs) + v_t(\bs)
 \end{align}
-where $u_t(\bs) = \sigma \delta |z_t|$, $z_t(\bs) = z_{tl}$ if $s \in P_{tl}$ where $P_{t1}, \ldots, P_{tL}$ form a partition, and $z_{tl} \iid N(0, 1)$, $\delta \in (-1, 1)$ controls skew, and $v_t(\bs)$ is a spatial process with mean zero and variance $\sigma^2(1 - \delta^2)$.
+where $z_t(\bs) = z_{tl}$ if $s \in P_{tl}$ where $P_{t1}, \ldots, P_{tL}$ form a partition, and $z_{tl} \iid \mbox{HN}(0, \sigma_t^2)$, $\delta \in (-1, 1)$ controls skew, and $v_t(\bs)$ is a spatial Gaussian process with mean zero and variance $\sigma_t^2(1 - \delta^2)$.
 Then $Y_t(\bs)$ is skew normal within each partition \citep{Minozzo2012}.
 We model this with a Bayesian hierarchical model as follows.
-Let $w_{t1}, \ldots, w_{tL}$ be partition centers so that $P_{tl}$ includes all spatial locations $\bs$ that are within the partition.
+Let $w_{t1}, \ldots, w_{tL}$ be partition centers so that \mbox{$P_{tl} = \{\bs_t : l = \argmin_k || \bs_t - w_{tk} ||\}$}.
 Then
-\begin{align} \label{eq:hier}
-    Y_t(\bs) \mid \Theta &= X_t(\bs) \beta + u_t(\bs) + v_t(\bs) \\
-    u_t(\bs) \mid \Theta &\sim \mbox{Half-Normal}(0, \sigma \delta)\\
-    v_t(\bs) \mid \Theta & \sim \Matern(0, \Sigma)\\
+\begin{align}
+    Y_t(\bs) \mid \Theta &= X_t(\bs) \beta + \delta z_t(\bs) + v_t(\bs) \label{eq:hier}\\
+    z_{tl}(\bs) \mid \Theta &\sim \mbox{Half-Normal}(0, \sigma^2_t)\\
+    v_t(\bs) \mid \Theta &\sim \Matern(0, \Sigma)\\
     \sigma &\sim \mbox{IG}(\alpha, \beta)\\
-    \delta &\sim \mbox{U}(-1, 1)
+    \delta &\sim \mbox{U}(-1, 1)\\
+    w_{tk} &\sim \mbox{U}(\calD)
 \end{align}
-where $l = \argmin_j ||\bs - w_j ||$, $\Theta = \{u_{t1}, \ldots, u_{tL}, w_{t1}, \ldots, w_{tL}, \beta, \sigma, \delta, \rho, \nu \}$ and $\Sigma$ is a \Matern covariance matrix with variance $\sigma^2 (1 - \delta^2)$
+where $\Theta = \{z_{t1}, \ldots, z_{tL}, w_{t1}, \ldots, w_{tL}, \beta, \sigma_t, \delta, \rho, \nu \}$; $l = \argmin_k ||\bs - w_k ||$; $\Sigma_t$ is a \Matern covariance matrix with variance $\sigma_t^2 (1 - \delta^2)$, spatial range $\rho$ and smoothness $\nu$; and $\calD$ is the spatial domain of interest.
 

### 2014-04-22 pot/LaTeX/skewnormal.tex
Updates to manuscript
Suggested changes from Brian for the manuscript.

@@
 The skew normal representation is from \citep{Minozzo2012} and is the sum of a normal and half-normal random variable.
-\subsection{Half-normal}\label{s:hn}
-Let $u = \xi + \eta |z|$ where $Z \sim N(0, 1)$.
-Then $U$ follows a half-normal distribution, $U \sim \mbox{HN}(\xi, \eta )$ \citep{Wiper2008}, and the density is given by 
+\subsection{Half-normal distribution}\label{s:hn}
+Let $u = |z|$ where $Z \sim N(\mu, \sigma^2)$.
+Specifically, we consider the case where $\mu = 0$. Then $U$ follows a half-normal distribution which we denote as $U \sim HN(0, 1)$, and the density is given by 
 \begin{align}
-  f_U(u) = \frac{ \sqrt{\pi} }{ \sqrt{2 \eta^2} } \exp \left( - \frac{ (u-\xi)^2 }{ 2 \eta^2 } \right), \quad u > \xi.
+  f_U(u) = \frac{ \sqrt{2} }{ \sqrt{\pi \sigma^2} } \exp \left( - \frac{ u^2 }{ 2 \sigma^2 } \right) I(u > 0)
 \end{align}
+When $\mu = 0$, the half-normal distribution is also equivalent to a $N_{(0, \infty)}(0, \sigma^2)$ where $N_{(a, b)}(\mu, \sigma^2)$ represents a normal distribution with mean $\mu$ and standard deviation $\sigma$ that has been truncated below at $a$ and above at $b$.
 
@@
 \end{align}
-where $z_t(\bs) = z_{tl}$ if $s \in P_{tl}$ where $P_{t1}, \ldots, P_{tL}$ form a partition, and $z_{tl} \iid \mbox{HN}(0, \sigma_t^2)$, $\delta \in (-1, 1)$ controls skew, and $v_t(\bs)$ is a spatial Gaussian process with mean zero and variance $\sigma_t^2(1 - \delta^2)$.
+where $z_t(\bs) = z_{tl}$ if $s \in P_{tl}$ where $P_{t1}, \ldots, P_{tL}$ form a partition, and $z_{tl} \iid N_{(0, \infty)}(0, \sigma_t^2)$, $\delta \in (-1, 1)$ controls skew, and $v_t(\bs)$ is a spatial Gaussian process with mean zero and variance $\sigma_t^2(1 - \delta^2)$.
+
 Then $Y_t(\bs)$ is skew normal within each partition \citep{Minozzo2012}.
 We model this with a Bayesian hierarchical model as follows.
-Let $w_{t1}, \ldots, w_{tL}$ be partition centers so that \mbox{$P_{tl} = \{\bs_t : l = \argmin_k || \bs_t - w_{tk} ||\}$}.
+Let $w_{t1}, \ldots, w_{tL}$ be partition centers so that
+\begin{align*}
+  P_{tl} = \{\bs_t : l = \argmin_k || \bs_t - w_{tk} ||\}.
+\end{align*}
 Then
 \begin{align}
-    Y_t(\bs) \mid \Theta &= X_t(\bs) \beta + \delta z_t(\bs) + v_t(\bs) \label{eq:hier}\\
-    z_{tl}(\bs) \mid \Theta &\sim \mbox{Half-Normal}(0, \sigma^2_t)\\
+    Y_t(\bs) \mid \Theta, z_{t1}, \ldots, z_{tL} &= X_t(\bs) \beta + \delta z_t(\bs) + v_t(\bs) \label{eq:hier}\\
+    z_{tl}(\bs) \mid \Theta &\sim N_{(0, \infty)}(0, \sigma^2_t)\\
     v_t(\bs) \mid \Theta &\sim \Matern(0, \Sigma)\\
-    \sigma &\sim \mbox{IG}(\alpha, \beta)\\
-    \delta &\sim \mbox{U}(-1, 1)\\
-    w_{tk} &\sim \mbox{U}(\calD)
+    \sigma &\sim IG(\alpha, \beta)\\
+    \delta &\sim Unif(-1, 1)\\
+    w_{tk} &\sim Unif(\calD)
 \end{align}
-where $\Theta = \{z_{t1}, \ldots, z_{tL}, w_{t1}, \ldots, w_{tL}, \beta, \sigma_t, \delta, \rho, \nu \}$; $l = \argmin_k ||\bs - w_k ||$; $\Sigma_t$ is a \Matern covariance matrix with variance $\sigma_t^2 (1 - \delta^2)$, spatial range $\rho$ and smoothness $\nu$; and $\calD$ is the spatial domain of interest.
+where $\Theta = \{w_{t1}, \ldots, w_{tL}, \beta, \sigma_t, \delta, \rho, \nu \}$; $l = \argmin_k ||\bs - w_k ||$; $\Sigma_t$ is a \Matern covariance matrix with variance $\sigma_t^2 (1 - \delta^2)$, spatial range $\rho$ and smoothness $\nu$; and $\calD$ is the spatial domain of interest.
 
@@
 Given $\Theta$, $Y_t(\bs)$ has truncated normal full conditional with these parameter values.
-So we sample $Y_t(\bs) \sim \text{TN}_{(-\infty, T)}$
+So we sample $Y_t(\bs) \sim N_{(-\infty, T)}$
 
@@
 \section*{Appendix A.1: Posterior distributions}
-\input{upost.tex}
+\input{zpost.tex}
 \input{betapost.tex}

### 2014-04-29 pot/code/R/mcmc.R
Found some mistakes in the MCMC
Made an adjustment so all calculations account for the fact that the z
term already incorporate sigma when it's generated. Also general
adjustments to incorporate misc bugs in the code.

@@ -5,13 +5,11 @@
 #########################################################################
 
 mcmc <- function(y, s, x, s.pred=NULL, x.pred=NULL, 
-                 thresh=0, thresh.quant=T, 
-                 r.model = "gamma", #also allow "fixed"
-                 nknots=1,          
+                 thresh=0, thresh.quant=T, nknots=1,          
                  iters=5000, burn=1000, update=100, thin=1, scale=T,
                  iterplot=F, plotname=NULL,
                  # initial values
-                 knots.init, z.init, beta.init=NULL, sigma.init=1,
+                 beta.init=NULL, sigma.init=1,
                  rho.init=0.5, nu.init=0.5, alpha.init=0.5,
                  delta.init=0,
                  # priors
@@ -19,7 +17,7 @@ mcmc <- function(y, s, x, s.pred=NULL, x.pred=NULL,
                  logrho.m=-2, logrho.s=1,
                  lognu.m=-1.2, lognu.s=1,
                  # debugging settings
-                 debug=F, 
+                 debug=F, knots.init, z.init, 
                  fixknots=F, fixz=F, fixbeta=F, fixsigma=F,
                  fixrho=F, fixnu=F, fixalpha=F,
                  fixdelta=F){
@@ -131,7 +129,7 @@ mcmc <- function(y, s, x, s.pred=NULL, x.pred=NULL,
   lognu <- log(nu)
   alpha <- alpha.init
   
-  if (length(sigma) == 1) {
+  if (length(sigma.init) == 1) {
     sigma <- rep(sigma.init, nt)
   } else {
     sigma <- sigma.init
@@ -150,7 +148,7 @@ mcmc <- function(y, s, x, s.pred=NULL, x.pred=NULL,
   
   # MH tuning params
   acc.w <- att.w <- mh.w <- rep(0.1, nt)  # knot locations
-  acc.delta <- att.delta <- mh.delta <- 0.1  
+  acc.delta <- att.delta <- mh.delta <- 10  
   acc.rho   <- att.rho   <- mh.rho   <- 0.1
   acc.nu    <- att.nu    <- mh.nu    <- 0.1
   acc.alpha <- att.alpha <- mh.alpha <- 0.1
[…]

### 2014-04-29 pot/code/R/plots/delta-0.9.R

@@ -42,7 +42,7 @@ y[[1]] <- data$y
 z.knots.t[[1]] <- data$z.knots
 z.sites.t[[1]] <- data$z.sites
 knots.t[[1]] <- data$knots
-#hist(y, main="delta = 0")
+# hist(y[[1]], main="delta = 0")
 
 data <- rpotspatial(nt=nt, s=s, x=x, beta=beta.t, sigma=sigma.t, delta=delta.t[2],
                    rho=rho.t, nu=nu.t, alpha=alpha.t)          
@@ -50,7 +50,7 @@ y[[2]] <- data$y
 z.knots.t[[2]] <- data$z.knots
 z.sites.t[[2]] <- data$z.sites
 knots.t[[2]] <- data$knots
-#hist(y, main="delta = 0.5")
+# hist(y[[2]], main="delta = 0.5")
 
 data <- rpotspatial(nt=nt, s=s, x=x, beta=beta.t, sigma=sigma.t, delta=delta.t[3],
                    rho=rho.t, nu=nu.t, alpha=alpha.t)
@@ -58,15 +58,15 @@ y[[3]] <- data$y
 z.knots.t[[3]] <- data$z.knots
 z.sites.t[[3]] <- data$z.sites
 knots.t[[3]] <- data$knots
-#hist(y, main="delta = 0.9")
+# hist(y[[3]], main="delta = 0.9")
 
-data <- rpotspatial(nt=nt, s=s, x=x, beta=beta.t, sigma=sigma.t, delta=delta.t[4],
+data <- rpotspatial(nt=nt, s=s, x=x, beta=beta.t, sigma=sigma.t, delta=0.95,
                    rho=rho.t, nu=nu.t, alpha=alpha.t)
 y[[4]] <- data$y
 z.knots.t[[4]] <- data$z.knots
 z.sites.t[[4]] <- data$z.sites
 knots.t[[4]] <- data$knots
-#hist(y, main="delta = 0.95")
+# hist(y[[4]], main="delta = 0.95")
 
 source("auxfunctions.R")
 source("mcmc.R")
@@ -77,6 +77,30 @@ fit1 <- mcmc(y=y[[1]], s=s, x=x, thresh=0, nknots=1,
              beta.init=beta.t, sigma.init=sigma.t, rho.init=rho.t,
              nu.init=nu.t, alpha.init=alpha.t, delta.init=delta.t[1],
[…]

### 2014-05-13 pot/code/R/mcmc.R
typos and plots
corrected a few typos and added in some iteration plots

@@ -285,7 +285,7 @@ mcmc <- function(y, s, x, s.pred=NULL, x.pred=NULL,
     # update beta
     if (debug) { print("beta") }
     if (!fixbeta) {  # debug
-      vvv <- diag(p) * beta.s^2
+      vvv <- diag(p) / beta.s^2
       mmm <- 0
       
       for (t in 1:nt) {
@@ -458,6 +458,9 @@ mcmc <- function(y, s, x, s.pred=NULL, x.pred=NULL,
       
     }
   }  # fi np > 0
+  
+  cur.ll <- LLike(y=y, x.beta=x.beta, sigma=sigma, delta=delta, 
+                  prec=prec, log.det=log.det, z.sites=z.sites, log=T)
   
   ##############################################
   # Keep track of iterations

### 2014-05-13 pot/code/R/pottest.R

@@ -446,7 +446,7 @@ for (t in 1:nt) {
   x[, t, 3] <- s[, 2]
 }
 
-beta.t <- c(0, 0, 0)
+beta.t <- c(10, -2, 3)
 rho.t <- 0.1
 nu.t <- 0.5
 delta.t <- 0
@@ -469,3 +469,54 @@ fit <- mcmc(y=y, s=s, x=x, thresh=0, nknots=1,
             debug=F, knots.init=knots.t, z.init=z.knots.t,
             fixknots=T, fixz=T, fixbeta=F, fixsigma=T, 
             fixrho=T, fixnu=T, fixalpha=T, fixdelta=T)
+#estimates beta pretty well
+
+rm(list=ls())
+
+library(fields)
+library(geoR)
+library(mvtnorm)
+library(evd)
+
+source("auxfunctions.R")
+source("mcmc.R")
+
+# set.seed(1234)
+
+# iid n(0, 1)
+# data settings
+s <- cbind(runif(50), runif(50))
+ns <- nrow(s)
+nt <- 100
+nsets <- 1
+nknots <- 1
+
+x <- array(1, c(ns, nt, 3))
+for (t in 1:nt) {
+  x[, t, 2] <- s[, 1]
+  x[, t, 3] <- s[, 2]
+}
+
+beta.t <- c(10, -2, 3)
+rho.t <- 0.1
+nu.t <- 0.5
+delta.t <- 0.5
+sigma.t <- rep(1, nt)
+alpha.t <- 0.8
+
+data <- rpotspatial(nt=nt, s=s, x=x, beta=beta.t, sigma=sigma.t, delta=delta.t,
+                    rho=rho.t, nu=nu.t, alpha=alpha.t, nknots=1)
+
+y <- data$y
+z.knots.t <- data$z.knots
+z.sites.t <- data$z.sites
+knots.t <- data$knots
+# hist(y, breaks=30)
+
+fit <- mcmc(y=y, s=s, x=x, thresh=0, nknots=1,
+            iters=15000, burn=10000, update=1000, iterplot=T,
+            beta.init=c(10, 5, 7), sigma.init=rep(1, nt), rho.init=rho.t,
+            nu.init=nu.t, alpha.init=alpha.t, delta.init=delta.t,
+            debug=F, knots.init=knots.t, z.init=z.knots.t,
+            fixknots=T, fixz=T, fixbeta=F, fixsigma=F, 
+            fixrho=F, fixnu=F, fixalpha=F, fixdelta=F)

### 2014-05-22 pot/code/analysis/ozone/OzoneResults.R
Troubleshooting
Troubleshooting issues related to estimating rho and nu when threshold
is high and delta is near 1.

@@ -3,20 +3,82 @@ rm(list=ls())
 load("OzoneMCMCnothresh.RData")
 load("OzoneMCMCthresh.RData")
 
+start <- 25001
+end   <- 30000
+
+# posterior distribution delta
 delta <- matrix(0, nrow=5000, ncol=4)
-delta[, 1] <- fit.normal.nothresh$delta[15001:20000]
-delta[, 2] <- fit.normal.thresh90$delta[25001:30000]
-delta[, 3] <- fit.skew.nothresh$delta[15001:20000]
-delta[, 4] <- fit.skew.thresh90$delta[25001:30000]
+delta[, 1] <- fit.normal.nothresh$delta[start:end]
+delta[, 2] <- fit.normal.thresh90$delta[start:end]
+delta[, 3] <- fit.skew.nothresh$delta[start:end]
+delta[, 4] <- fit.skew.thresh90$delta[start:end]
 colnames(delta) <- c("Normal (T=0)", "Normal (T=0.9)", "Skew (T=0)", "Skew (T=0.9)")
-boxplot(delta)
+boxplot(delta, main="Delta posterior")
 
+# iteration plot delta
+par(mfrow=c(2, 1))
+plot(fit.skew.nothresh$delta, type="l", main="Delta iteration plot (T=0)", ylab="delta")
+abline(v=25000, lty=2)
+plot(fit.skew.thresh90$delta, type="l", main="Delta iteration plot (T=0.9)", ylab="delta")
+abline(v=25000, lty=2)
+
+# posterior distribution alpha
 alpha <- matrix(0, nrow=5000, ncol=4)
-alpha[, 1] <- fit.normal.nothresh$alpha[15001:20000]
-alpha[, 2] <- fit.normal.thresh90$alpha[25001:30000]
-alpha[, 3] <- fit.skew.nothresh$alpha[15001:20000]
-alpha[, 4] <- fit.skew.thresh90$alpha[25001:30000]
+alpha[, 1] <- fit.normal.nothresh$alpha[start:end]
+alpha[, 2] <- fit.normal.thresh90$alpha[start:end]
+alpha[, 3] <- fit.skew.nothresh$alpha[start:end]
+alpha[, 4] <- fit.skew.thresh90$alpha[start:end]
 colnames(alpha) <- c("Normal (T=0)", "Normal (T=0.9)", "Skew (T=0)", "Skew (T=0.9)")
-boxplot(alpha)
[…]

### 2014-06-13 pot/code/R/auxfunctions.R
Multiple sigma bugfix and update
Found a bug in the update for the z terms, and adjusted the manuscript
to reflect recent updates for the multiple knot setup.

@@ -76,11 +76,11 @@ rpotspatial <- function(nt, s, x, beta, sigma.alpha.t, sigma.beta.t,
   sigma.knots <- matrix(sigma.knots, nrow=nknots, ncol=nt)
   sigma.sites <- SigmaSites(sigma.knots, partition, nknots)
   
+  z.knots <- sqrt(sigma.knots) * abs(rnorm(n=(nknots * nt), mean=0, sd=1))
+  z.knots <- matrix(z.knots, nrow=nknots, ncol=nt)
+  z.sites <- ZBySites(z.knots, partition, nknots)
+  
   for (t in 1:nt) {
-  	sigma.knots.t   <- sigma.knots[, t]
-  	z.knots.t <- sqrt(sigma.knots.t) * abs(rnorm(n=nknots, mean=0, sd=1))
-    z.sites.t <- ZBySites(z.knots.t, partition[, t])
-    
     x.beta.t <- x[, t, ] %*% beta
      
     if (alpha != 0) {
@@ -91,11 +91,10 @@ rpotspatial <- function(nt, s, x, beta, sigma.alpha.t, sigma.beta.t,
     }
     
     sigma.sites.t <- sigma.sites[, t]
+    z.sites.t <- z.sites[, t]
     v.nug <- rnorm(ns, 0, sqrt(1 - alpha))
     v.t <- sqrt(sigma.sites.t) * sqrt(1 - delta^2) * ( v.nug + v.spat) 
     y[, t]       <- x.beta.t + delta * z.sites.t + v.t
-    z.knots[, t] <- z.knots.t
-    z.sites[, t] <- z.sites.t
     v[, t]       <- v.t
     
   }
@@ -232,21 +231,32 @@ CorFx <- function(d, alpha, rho, nu, cov=F){
 
 ################################################################
 # Arguments:
-#   z.knots(nknots): vector of random effects with one entry per knot
-#   partition(ns): partition membership vector
+#   z.knots(nknots, nt): matrix of random effects with one entry per knot/day
+#   partition(ns, nt): partition membership matrix
 #
 # Returns:
-#   z.sites(ns): vector of random effects with one entry per site
+#   z.sites(ns, nt): matrix of random effects with one entry per site/day
 ################################################################
-ZBySites <- function(z.knots, partition){
-  nknots  <- length(z.knots)
-  ns      <- length(partition)
-  z.sites <- rep(NA, ns)
-  
-  for (k in 1:nknots) {
-  	these <- which(partition == k)
[…]

### 2014-06-13 pot/code/R/mcmc.R

@@ -102,11 +102,8 @@ mcmc <- function(y, s, x, s.pred=NULL, x.pred=NULL,
   
   # initialize random effects
   z.knots <- matrix(1, nrow=nknots, ncol=nt)
-  z.sites <- matrix(NA, nrow=ns, ncol=nt)
   if (fixz) {z.knots <- z.init}
-  for (t in 1:nt) {
-  	z.sites[, t] <- ZBySites(z.knots[, t], partition[, t])
-  }
+  z.sites <- ZBySites(z.knots, partition, nknots)
     
   # initialize parameters
   if (is.null(beta.init)) {
@@ -226,13 +223,14 @@ mcmc <- function(y, s, x, s.pred=NULL, x.pred=NULL,
             r2       <- y[-these, t] - x.beta[-these, t] - delta * z.sites[-these, t]
             
             # vvv      <- sigma[, t] * (1 - delta^2)
-            prec.11  <- prec[these, these]
-            prec.21  <- prec[-these, these]
+            prec.t     <- diag(1 / (sqrt(sigma.sites[, t]))) %*% prec %*% diag(1 / (sqrt(sigma.sites[, t])))
+            prec.11.t  <- prec.t[these, these]
+            prec.21.t  <- prec.t[-these, these]
             
-            mu.z     <- delta * sum(t(r1) %*% prec.11 + t(r2) %*% prec.21)
-            lambda.z <- delta^2 * sum(prec.11) + 1 - delta^2
+            mu.z     <- delta * sum(t(r1) %*% prec.11.t + t(r2) %*% prec.21.t)
+            lambda.z <- delta^2 * sum(prec.11.t) + (1 - delta^2) / sigma.knots[k, t]
             mn.z     <- mu.z / lambda.z
-            sd.z     <- sqrt(sigma.knots[k, t] * (1 - delta^2) / lambda.z)
+            sd.z     <- sqrt((1 - delta^2) / lambda.z)
         
             z.new   <- rTNorm(mn=mn.z, sd=sd.z, lower=0, upper=Inf) 
             # if any z come back as Inf it's because P(Y > T) = 0
@@ -273,26 +271,32 @@ mcmc <- function(y, s, x, s.pred=NULL, x.pred=NULL,
         for (t in 1:nt) {
           can.knots[[1]] <- matrix(NA, nrow=nknots, ncol=2)
           att.w[t] <- att.w[t] + 1
-          ss.t <- t((y[, t] - mu[, t]) / sigma.sites[, t]) %*% prec %*% 
[…]

### 2014-06-16 pot/code/R/auxfunctions.R
Change to precision instead of variance
In an attempt to find the bug, I am changing things from precision
terms to variance terms.

@@ -30,8 +30,8 @@ rTNorm <- function(mn, sd, lower=-Inf, upper=Inf, fudge=0){
 #   s(ns, 2): spatial locations
 #   x(ns, nt, np): Matrix of spatial covariates
 #   beta.y(np): parameters for mu.y
-#   sigma.alpha.t(1): hyperparameter for sigma terms
-#   sigma.beta.t(1): hyperparameter for sigma terms
+#   tau.alpha.t(1): hyperparameter for tau terms
+#   tau.beta.t(1): hyperparameter for tau terms
 #   delta(1): skewness parameter
 #   rho(1): spatial range parameter
 #   nu.y(1): smoothness parameter for Matern covariance
@@ -46,7 +46,7 @@ rTNorm <- function(mn, sd, lower=-Inf, upper=Inf, fudge=0){
 #		knots(nknots, 2): knot locations
 #       fixr(nknots, nt): whether it was a gaussian or gamma random effect
 #########################################################################
-rpotspatial <- function(nt, s, x, beta, sigma.alpha.t, sigma.beta.t, 
+rpotspatial <- function(nt, s, x, beta, tau.alpha.t, tau.beta.t, 
                         delta, rho, nu, alpha, nknots=1){
     
   # initial setup
@@ -72,11 +72,13 @@ rpotspatial <- function(nt, s, x, beta, sigma.alpha.t, sigma.beta.t,
   }
   
   partition    <- Membership(s=s, knots=knots)
-  sigma.knots <- 1 / rgamma(nt * nknots, sigma.alpha.t, sigma.beta.t)
-  sigma.knots <- matrix(sigma.knots, nrow=nknots, ncol=nt)
-  sigma.sites <- SigmaSites(sigma.knots, partition, nknots)
-  
+  tau.knots <- rgamma(nt * nknots, tau.alpha.t, tau.beta.t)
+  sigma.knots <- 1 / tau.knots
   z.knots <- sqrt(sigma.knots) * abs(rnorm(n=(nknots * nt), mean=0, sd=1))
+  
+  tau.knots <- matrix(sigma.knots, nrow=nknots, ncol=nt)
+  tau.sites <- TauSites(tau.knots, partition, nknots)
+  sigma.sites <- 1 / tau.sites
   z.knots <- matrix(z.knots, nrow=nknots, ncol=nt)
   z.sites <- ZBySites(z.knots, partition, nknots)
   
@@ -130,28 +132,28 @@ Membership <- function(s, knots){
 
 ################################################################
[…]

### 2014-06-16 pot/code/R/pottest.R

@@ -880,12 +880,12 @@ for (t in 1:nt) {
 beta.t <- c(10, -2, 3)
 rho.t <- 0.1
 nu.t <- 0.5
-delta.t <- 0.8
+delta.t <- 0
 
-sigma.alpha.t <- 5
-sigma.beta.t <- 2
+sigma.alpha.t <- 3
+sigma.beta.t <- 1
 
-alpha.t <- 0.8
+alpha.t <- 0.9
 
 data <- rpotspatial(nt=nt, s=s, x=x, beta=beta.t, 
                     sigma.alpha=sigma.alpha.t, sigma.beta=sigma.beta.t,
@@ -898,9 +898,9 @@ knots.t <- data$knots
 sigma.knots.t <- data$sigma.knots
 sigma.sites.t <- data$sigma.sites
 # hist(y, breaks=30)
-# sigma1,1 = 0.395, sigma3,21 = 1.071
-# z1,8 = 0.090, z3,21 = 1.318
 
+# sigma1,1 = 0.342, sigma3,20 = 1.960
+# z1,8 = 1.208, z3,21 = 0.008
 
 source("auxfunctions.R")
 source("mcmc.R")
@@ -912,12 +912,12 @@ fit <- mcmc(y=y, s=s, x=x, thresh=0, nknots=nknots,
             sigma.alpha.init=sigma.alpha.t, 
             sigma.beta.init=sigma.beta.t, rho.init=rho.t,
             nu.init=nu.t, alpha.init=alpha.t, delta.init=delta.t,
-            sigma.beta.a=0.1, sigma.beta.b=0.1,
+            sigma.beta.a=0.01, sigma.beta.b=0.01,
             debug=F, 
             knots.init=knots.t, z.init=z.knots.t,
-            fixknots=T, fixz=F, fixbeta=T, 
-            fixsigma=T, fixsigma.alpha=T, fixsigma.beta=T,
-            fixrho=T, fixnu=T, fixalpha=T, fixdelta=T,
+            fixknots=T, fixz=T, fixbeta=T, 
+            fixsigma=F, fixsigma.alpha=F, fixsigma.beta=F,
+            fixrho=F, fixnu=F, fixalpha=F, fixdelta=T,
             sigma.by.knots=T)
 end.time <- proc.time()
 

### 2014-06-30 pot/LaTeX/skewnormal.tex
Found some mistakes in the MCMC
Made an adjustment so all calculations account for the fact that the z
term already incorporate sigma when it's generated. Also general
adjustments to incorporate misc bugs in the code.
Updates to simplify the mcmc

@@
 \subsection{Complete data}\label{s:model}
-Consider the spatial process
+Consider a skew Gaussian spatial process
 \begin{align} \label{eq:fullmodel}
-  Y_t(\bs) &= X_t(\bs) \beta + e_t(\bs)\\
-  e_t(\bs) &= \delta z_t(\bs) + v_t(\bs)
+  Y_t(\bs) &= X_t(\bs) \beta + \sigma_1 z_t(\bs) + v_t(\bs)
 \end{align}
-where $z_t(\bs) = z_{tl}$ if $s \in P_{tl}$ where $P_{t1}, \ldots, P_{tL}$ form a partition, and $z_{tl} \iid N_{(0, \infty)}(0, \sigma_t^2)$, $\delta \in (-1, 1)$ controls skew, and $v_t(\bs)$ is a spatial Gaussian process with mean zero and variance $\sigma_t^2(1 - \delta^2)$.
+where $z_t(\bs) = z_{tl}$ if $s \in P_{tl}$ where $P_{t1}, \ldots, P_{tL}$ form a partition, and $z_{tl} \iid N_{(0, \infty)}(0, 1)$, $\sigma_1 \in \calR$, $\sigma_2, \sigma_0 \in \calR^+$, and $v_t(\bs)$ is a spatial Gaussian process with mean zero and variance $\sigma_{tl}^2$.
+It can be shown \citep{Zhang2010} that $Y_t(\bs)$ follows a skew normal distribution with skewness parameter $\alpha = \frac{ \sigma_1 }{ \sigma_{tl} }$.
+We can then reexpress the model in (\ref{eq:fullmodel}) as
+\begin{align}
+  Y_t(\bs) &= X_t(\bs) \beta + \alpha z_t(\bs) + v_t(\bs)
+\end{align}
+where $z_t(\bs) = z_{tl}$, $z_{tl} \iid N_{(0, \infty)}(0, \sigma_{tl}^2)$, and $v_t(\bs)$ is defined as before.
 
-Then $Y_t(\bs)$ is skew normal within each partition \citep{Minozzo2012}.
 We model this with a Bayesian hierarchical model as follows.
@@
 \begin{align}
-    Y_t(\bs) \mid \Theta, z_{t1}, \ldots, z_{tL} &= X_t(\bs) \beta + \delta z_t(\bs) + v_t(\bs) \label{eq:hier}\\
-    z_{tl}(\bs) \mid \Theta &\sim N_{(0, \infty)}(0, \sigma^2_t)\\
+    Y_t(\bs) \mid \Theta, z_{t1}, \ldots, z_{tL} &= X_t(\bs) \beta + \alpha z_t(\bs) + v_t(\bs) \label{eq:hier}\\
+    z_{tl}(\bs) \mid \Theta &\sim N_{(0, \infty)}(0, \sigma^2_{tl})\\
     v_t(\bs) \mid \Theta &\sim \Matern(0, \Sigma)\\
-    \sigma &\sim IG(\alpha, \beta)\\
-    \delta &\sim Unif(-1, 1)\\
+    \sigma^2_{tl} &\iid IG(\alpha, \beta)\\
+    \alpha &\sim N(0, 10)\\
     w_{tk} &\sim Unif(\calD)
 \end{align}
-where $\Theta = \{w_{t1}, \ldots, w_{tL}, \beta, \sigma_t, \delta, \rho, \nu \}$; $l = \argmin_k ||\bs - w_k ||$; $\Sigma_t$ is a \Matern covariance matrix with variance $\sigma_t^2 (1 - \delta^2)$, spatial range $\rho$ and smoothness $\nu$; and $\calD$ is the spatial domain of interest.
+where $\Theta = \{w_{t1}, \ldots, w_{tL}, \beta, \sigma_t, \delta, \rho, \nu \}$; $l = \argmin_k ||\bs - w_k ||$; $\Sigma_t$ is a \Matern covariance matrix with variance $\sigma_{tl}^2 (1 - \delta^2)$, spatial range $\rho$ and smoothness $\nu$; and $\calD$ is the spatial domain of interest.
 
@@
 To update $\Theta$ given the current value of the complete data $\bY_1, \ldots, \bY_T$, we use a standard Gibbs updates for all parameters except for the knot locations which are done using a Metropolis update.
-See Appendix A.1 for details regarding Gibbs sampling and $|u_t(\bs)|$.
+See Appendix A.1 for details regarding Gibbs sampling.
 
@@
 \input{sigpost.tex}
+\input{alphapost.tex}
 

### 2014-06-30 pot/code/analysis/ozone/OzoneSetup.R

@@ -92,19 +92,8 @@ tau.init <- 1
 
 save.image(file="cv-setup.RData")
 
-# setting 1:  T = 0.00,  1 partition
-# setting 2:  T = 0.00,  1 partition, rhofixed
-# setting 3:  T = 0.80,  1 partition
-# setting 4:  T = 0.80,  1 partition, rhofixed
-# setting 5:  T = 0.95,  1 partition
-# setting 6:  T = 0.95,  1 partition, rhofixed
-# setting 7:  T = 0.99,  1 partition
-# setting 8:  T = 0.99,  1 partition, rhofixed
-# setting 9:  T = 0.00,  5 partitions
-# setting 10: T = 0.80,  5 partitions
-# setting 11: T = 0.95,  5 partitions
-# setting 12: T = 0.99,  5 partitions
-# setting 13: T = 0.00, 10 partitions
-# setting 14: T = 0.80, 10 partitions
-# setting 15: T = 0.95, 10 partitions
-# setting 16: T = 0.99, 10 partitions
+# setting 1: gaussian
+# setting 2: t0
+# setting 3: t5
+# setting 4: t0 T = 0.9
+# setting 5: t5 T = 0.9

### 2014-06-30 pot/code/analysis/ozone/cv5-10.R

@@ -1,49 +1,0 @@
-library(fields)
-library(geoR)
-library(mvtnorm)
-
-rm(list=ls())
-load('cv-setup.RData')
-source('../../R/mcmc.R')
-source('../../R/auxfunctions.R')
-
-setting <- 10
-threshold <- 0.80
-nknots <- 5
-outputfile <- paste("cv-", setting, ".RData", sep="")
-start <- proc.time()
-
-fit <- vector(mode="list", length=5)
-
-for(val in 1:5){
-	
-	set.seed(setting*100 + val)
-	
-	cat("CV", val, "started \n")
-	val.idx <- cv.lst[[val]]
-	y.o <- y[-val.idx,]
-	X.o <- X[-val.idx,,]
-	S.o <- s[-val.idx,]
-	
-	rho.ml <- RhoML(S.o, y.o, cov=X.o[, , 4])
-	
-	y.p <- y[val.idx,]
-	X.p <- X[val.idx,,]
-	S.p <- s[val.idx,]
-	
-	tic.set <- proc.time()
-	fit[[val]] <- mcmc(y=y.o, s=S.o, x=X.o, x.pred=X.p, s.pred=S.p, 
-	                   thresh=threshold, nknots=nknots, 
-                       iters=30000, burn=25000, update=1000, iterplot=F,
-                       beta.init=beta.init, tau.init=tau.init, 
-                       rho.init=rho.ml, fixrho=T,
-                       nu.init=0.5, alpha.init=0.5, delta.init=0, fixdelta=T,
-                       fixz=T, z.init=matrix(0, nrow=nknots, ncol=nt), scale=F)
-	toc.set <- proc.time()
-	time.set <- (toc.set - tic.set)[3]
-	
-	elap.time.val <- (proc.time() - start)[3]
-	avg.time.val <- elap.time.val / val
-	cat("CV", val, "finished", round(avg.time.val, 2), "per dataset \n")
-	save(fit, file=outputfile)
-}

### 2014-06-30 pot/code/analysis/ozone/cv5-11.R

@@ -1,49 +1,0 @@
-library(fields)
-library(geoR)
-library(mvtnorm)
-
-rm(list=ls())
-load('cv-setup.RData')
-source('../../R/mcmc.R')
-source('../../R/auxfunctions.R')
-
-setting <- 11
-threshold <- 0.95
-nknots <- 5
-outputfile <- paste("cv-", setting, ".RData", sep="")
-start <- proc.time()
-
-fit <- vector(mode="list", length=5)
-
-for(val in 1:5){
-	
-	set.seed(setting*100 + val)
-	
-	cat("CV", val, "started \n")
-	val.idx <- cv.lst[[val]]
-	y.o <- y[-val.idx,]
-	X.o <- X[-val.idx,,]
-	S.o <- s[-val.idx,]
-	
-	rho.ml <- RhoML(S.o, y.o, cov=X.o[, , 4])
-	
-	y.p <- y[val.idx,]
-	X.p <- X[val.idx,,]
-	S.p <- s[val.idx,]
-	
-	tic.set <- proc.time()
-	fit[[val]] <- mcmc(y=y.o, s=S.o, x=X.o, x.pred=X.p, s.pred=S.p, 
-	                   thresh=threshold, nknots=nknots, 
-                       iters=30000, burn=25000, update=1000, iterplot=F,
-                       beta.init=beta.init, tau.init=tau.init, 
-                       rho.init=rho.ml, fixrho=T,
-                       nu.init=0.5, alpha.init=0.5, delta.init=0, fixdelta=T,
-                       fixz=T, z.init=matrix(0, nrow=nknots, ncol=nt), scale=F)
-	toc.set <- proc.time()
-	time.set <- (toc.set - tic.set)[3]
-	
-	elap.time.val <- (proc.time() - start)[3]
-	avg.time.val <- elap.time.val / val
-	cat("CV", val, "finished", round(avg.time.val, 2), "per dataset \n")
-	save(fit, file=outputfile)
-}

### 2014-06-30 pot/code/analysis/ozone/cv5-13.R

@@ -1,47 +1,0 @@
-library(fields)
-library(geoR)
-library(mvtnorm)
-
-rm(list=ls())
-load('cv-setup.RData')
-source('../../R/mcmc.R')
-source('../../R/auxfunctions.R')
-
-setting <- 13
-threshold <- 0
-nknots <- 10
-outputfile <- paste("cv-", setting, ".RData", sep="")
-start <- proc.time()
-
-fit <- vector(mode="list", length=5)
-
-for(val in 1:5){
-	
-	set.seed(setting*100 + val)
-	
-	cat("CV", val, "started \n")
-	val.idx <- cv.lst[[val]]
-	y.o <- y[-val.idx,]
-	X.o <- X[-val.idx,,]
-	S.o <- s[-val.idx,]
-	
-	y.p <- y[val.idx,]
-	X.p <- X[val.idx,,]
-	S.p <- s[val.idx,]
-	
-	tic.set <- proc.time()
-	fit[[val]] <- mcmc(y=y.o, s=S.o, x=X.o, x.pred=X.p, s.pred=S.p, 
-	                   thresh=threshold, nknots=nknots, 
-                       iters=30000, burn=25000, update=1000, iterplot=F,
-                       beta.init=beta.init, tau.init=tau.init, rho.init=0.5,
-                       nu.init=0.5, alpha.init=0.5, delta.init=0, fixdelta=T, 
-                       fixz=T, z.init=matrix(0, nrow=nknots, ncol=nt), scale=F)
-
-	toc.set <- proc.time()
-	time.set <- (toc.set - tic.set)[3]
-	
-	elap.time.val <- (proc.time() - start)[3]
-	avg.time.val <- elap.time.val / val
-	cat("CV", val, "finished", round(avg.time.val, 2), "per dataset \n")
-	save(fit, file=outputfile)
-}

### 2014-07-01 pot/code/R/mcmc.R
Bugfix and cleanup
Fixed bugs in the MCMC for the skewness parameters. Seems to be working
now. Also removed some old files for the ozone analysis.

@@ -10,7 +10,7 @@ mcmc <- function(y, s, x, s.pred=NULL, x.pred=NULL,
                  thresh=0, thresh.quant=T, nknots=1, 
                  iters=5000, burn=1000, update=100, thin=1,
                  iterplot=F, plotname=NULL, method="t",
-                # initial values
+                 # initial values
                  beta.init=NULL, 
                  tau.init=1,
                  tau.alpha.init=0.1, 
@@ -18,13 +18,17 @@ mcmc <- function(y, s, x, s.pred=NULL, x.pred=NULL,
                  rho.init=0.5, 
                  nu.init=0.5, 
                  alpha.init=0.5,
-                # priors
+                 # priors
                  beta.m=0, beta.s=10, 
                  tau.alpha.m=0, tau.alpha.s=1, 
                  tau.beta.a=0.1, tau.beta.b=0.1,
                  logrho.m=-2, logrho.s=1,
                  lognu.m=-1.2, lognu.s=1,
-                 alpha.m=0, alpha.s=1
+                 alpha.m=0, alpha.s=1,
+                 # skew inits
+                 z.init=1, z.alpha.init=0,
+                 # skew priors
+                 z.alpha.m=0, z.alpha.s=2, skew=T
         ){
     
   library(geoR)
@@ -103,6 +107,24 @@ mcmc <- function(y, s, x, s.pred=NULL, x.pred=NULL,
     }
   }
   
+  zg <- matrix(0, ns, nt)
+  if (length(z.init) == 1) {
+    cat("initializing all z terms to", z.init, "\n")
+  }
+  z <- matrix(z.init, nknots, nt)
+  for (t in 1:nt) {
+    zg[, t] <- z[g[, t], t]
+  }
+  
+  if (skew) {
+    z.alpha <- z.alpha.init
+  } else {
+  	if (z.alpha.init != 0) {
+  	  warning("z.alpha.init being ignored since skew=F")
+  	}
+    z.alpha <- 0
+  }
+  
   # easier to keep calculations in the precision scale for MCMC
   sigma2    <- 1 / tau
   sigma2g   <- 1 / taug
@@ -129,7 +151,7 @@ mcmc <- function(y, s, x, s.pred=NULL, x.pred=NULL,
[…]

### 2014-07-01 pot/code/analysis/ozone/OzoneSetup.R

@@ -136,11 +136,78 @@ for (thresh in 1:length(threshs)){
   exceed[, thresh] <- exceed.thresh
 }
 
+par(mfrow=c(1, 2))
 xplot <- (0:12) + 0.5
-plot(xplot, exceed[, 1], type="b", ylim=c(min(exceed), max(exceed)), ylab="exceed", xaxt="n", xlab="bin distance", pch=1, lty=3)
+plot(xplot, exceed[, 1], type="b", ylim=c(0, 0.75), ylab="exceed", xaxt="n", xlab="bin distance", pch=1, lty=3, main="chi-plot ozone")
 axis(1, at=0:12, labels=bins[-14])
 for (line in 2:9) { 
 	lines(xplot, exceed[, line], lty=3)
 	points(xplot, exceed[, line], pch=line)
 }
-legend("topright", lty=1:9, pch=1:9, legend=probs, title="sample quantiles")
+legend("topright", lty=1:9, pch=1:9, legend=probs, title="sample quants")
+
+# Making the chi plot
+# Residuals after lm
+y.lm <- y[, 1]
+x.lm <- X[, 1, ]
+for (t in 2:nt) {
+  y.lm <- c(y.lm, y[, t])
+  x.lm <- rbind(x.lm, X[, t, ]) 
+}
+ozone.lm <- lm(y.lm ~ x.lm)
+ozone.res <- residuals(ozone.lm)
+ozone.int <- ozone.lm$coefficients[1]
+ozone.beta1 <- ozone.lm$coefficients[3]
+ozone.beta2 <- ozone.lm$coefficients[4]
+ozone.cmaq <- ozone.lm$coefficients[5]
+ozone.beta <- c(ozone.int, ozone.beta1, ozone.beta2, ozone.cmaq)
+res <- matrix(NA, ns, nt)
+for (t in 1:nt) {
+  res[, t] <- y[, t] - X[, t, ] %*% ozone.beta
+}
+
+d <- as.vector(dist(s))
+j <- 1:54
+i <- 2:55
+ij <- expand.grid(i, j)
+ij <- ij[(ij[1] > ij[2]), ]
+sites <- cbind(d, ij)
+dist <- rdist(s)
+diag(dist) <- 0
+
+bins <- seq(0, 600, 50)
+bins <- c(bins, 900)
+
+probs <- c(0.7, 0.8, 0.9, 0.95, 0.96, 0.97, 0.98, 0.99, 0.995)
+threshs <- quantile(res, probs=probs, na.rm=T)
[…]

### 2014-07-29 pot/LaTeX/jsm2014.tex
formatting changes
font tweaks

\documentclass[t]{beamer}
\usepackage{beamerthemesplit}
\usepackage{pstricks}
\usepackage{graphicx}
\usepackage{mdwlist}

\usepackage{amssymb,latexsym,amsmath,amsthm,bbm}
\usepackage{tikz}
\usepackage[english]{babel}
\usepackage[latin1]{inputenc}
\usepackage{multirow}
\usepackage{verbatim}
\usepackage{alltt}
\usepackage{mycommands}

\usepackage{cmbright}
\renewcommand*\familydefault{\sfdefault}
\usepackage[T1]{fontenc}

\definecolor{wp-red}{RGB}{204,0,0}
\definecolor{wp-gray}{RGB}{51,51,51}
\definecolor{reynolds-red}{RGB}{153,0,0}
\definecolor{pyroman-flame}{RGB}{209,81,34}
\definecolor{hunt-yellow}{RGB}{253,215,38}
\definecolor{genomic-green}{RGB}{125,140,31}
\definecolor{innovation-blue}{RGB}{66,126,147}
\definecolor{bio-indigo}{RGB}{65,86,161}

\setbeamercolor{structure}{fg=wp-red}
\setbeamercolor{title}{bg=white, fg=wp-red}  % changes color on title page
\setbeamerfont{title}{series=\bfseries, size=\LARGE}
\setbeamerfont{author}{series=\bfseries}

\setbeamercolor{frametitle}{bg=wp-red, fg=white}  % changes color at top of frame
\setbeamerfont{frametitle}{series=\mdseries}
\setbeamercolor{title in head/foot}{fg=white, bg=wp-red}  % changes color for title in footer
\setbeamerfont{title in head/foot}{series=\mdseries}
\setbeamercolor{author in head/foot}{fg=white,bg=wp-gray}  % changes color for author in footer
\setbeamerfont{author in head/foot}{series=\mdseries}

\title[Spatiotemporal Modeling of Extreme Events] % (optional, use only with long paper titles)
{
  Spatiotemporal Modeling of Extreme Events
}
\author[B. Reich and S. Morris]{Brian Reich and Samuel Morris}
\institute[NCSU]{North Carolina State University}
\date{August 7, 2014}

\begin{document}

\begin{frame}
\begin{center}
  \titlepage
\end{center}
\end{frame}

\begin{frame}{Motivation}
  \begin{itemize}
    \item Average behavior is important to understand, but it does not paint the whole picture.
    \begin{itemize}
      \item e.g. When constructing river levees, engineers need to be able to estimate a 100-year or 1000-year flood levels.
    \end{itemize}
    \item In geostatistical analysis, kriging uses spatial correlation to help inform prediction at unknown locations.
    \item Want to explore ways to incorporate this spatial correlation when estimating the tails of the distribution.
  \end{itemize}
\end{frame}

\begin{frame}{Introduction to extremes}
  \begin{itemize} \itemsep 1.5em
  \item Two common methods for extremes:
  \begin{enumerate}[1]
    \item Block maxima
    \begin{itemize}
      \item Only use maxima of independent blocks (e.g. yearly maxima)
      \item Lose information by throwing away non-maxima
    \end{itemize}
    \item Peaks over threshold
    \begin{itemize}
      \item Use all observations over a threshold
      \item
    \end{itemize}
  \end{enumerate}

  \end{itemize}
\end{frame}

\end{document}

### 2014-07-30 pot/LaTeX/jsm2014.tex
minor tweaks to styling
typo correction
Accidentally did not run cross-validation for skew-t-5 (T=0.9)

@@
-\documentclass[t]{beamer}
+\documentclass{beamer}
 \usepackage{beamerthemesplit}
-\usepackage{pstricks}
+% \usepackage{pstricks}
 \usepackage{graphicx}
 \usepackage{mdwlist}
+\usepackage{lineno, hyperref}
 
 \usepackage{amssymb,latexsym,amsmath,amsthm,bbm}
+\usepackage{hyperref}
 \usepackage{tikz}
@@
 \setbeamercolor{title}{bg=white, fg=wp-red}  % changes color on title page
-\setbeamerfont{title}{series=\bfseries, size=\LARGE}
-\setbeamerfont{author}{series=\bfseries}
+\setbeamerfont{title}{series=\bfseries, size=\huge}
+\setbeamerfont{author}{series=\bfseries, size=\Large}
+\setbeamerfont{institute}{series=\mdseries, size=\large}
 
 \setbeamercolor{frametitle}{bg=wp-red, fg=white}  % changes color at top of frame
-\setbeamerfont{frametitle}{series=\mdseries}
+\setbeamerfont{frametitle}{series=\bfseries}
 \setbeamercolor{title in head/foot}{fg=white, bg=wp-red}  % changes color for title in footer
-\setbeamerfont{title in head/foot}{series=\mdseries}
+\setbeamerfont{title in head/foot}{series=\bfseries}
 \setbeamercolor{author in head/foot}{fg=white,bg=wp-gray}  % changes color for author in footer
-\setbeamerfont{author in head/foot}{series=\mdseries}
+\setbeamerfont{author in head/foot}{series=\bfseries}
 
@@
 \institute[NCSU]{North Carolina State University}
-\date{August 7, 2014}
+\date{}
 
@@
 
-\begin{frame}
+\begin{frame}\frametitle{\ }
 \begin{center}
-  \titlepage
+	\maketitle
 \end{center}
@@
 \begin{frame}{Motivation}
-  \begin{itemize}
+  \begin{itemize} \setlength{\itemsep}{1em}
     \item Average behavior is important to understand, but it does not paint the whole picture.
@@
 \begin{frame}{Introduction to extremes}
-  \begin{itemize} \itemsep 1.5em
-  \item Two common methods for extremes:
-  \begin{enumerate}[1]
-    \item Block maxima
+  \begin{itemize} \setlength{\itemsep}{0.5em}
+    \item Max-stable processes (Cooley et al., 2012):
     \begin{itemize}
-      \item Only use maxima of independent blocks (e.g. yearly maxima)
-      \item Lose information by throwing away non-maxima
+      \item Consider a spatial process $x_t(\bs)$, $t = 1, \ldots, T$.
+      \item Let $M_T(\bs) = \left\{ \bigvee_{t=1}^T x_t(\bs_1), \ldots, \bigvee_{t=1}^T x_t(\bs_n) \right\}$
+      \item If there exists normalizing sequences $a_T(\bs)$ and $b_T(\bs)$ such
+      that for all sites, $\bs_i, i = 1, \ldots, d$,
+      \begin{align*}
+        a_T^{-1}(\bs) \left\{ M_T(\bs) - b_T(\bs) \right\} \converged Y(\bs)
+      \end{align*}
+      which has a non-degenerate distribution, then $Y(\bs)$ is a max-stable process.
     \end{itemize}
-    \item Peaks over threshold
+  \end{itemize}
+\end{frame}
+
+\begin{frame}{General approaches}
   \begin{itemize}
-      \item Use all observations over a threshold
-      \item
+    \item Two standard approaches:
+    \begin{enumerate}[1.] \setlength{\itemsep}{0.5em}
+      \item Block maxima:
+      \begin{itemize}
+        \item Uses yearly maxima
+        \item Discards many observations
+      \end{itemize}
+      \item Peaks-over-threshold
+      \begin{itemize}
+        \item Incorporates more data than block maxima
+        \item Autocorrelation may be an issue between observations (e.g. flood levels don't dissipate overnight)
       \end{itemize}
     \end{enumerate}
+  \end{itemize}
+\end{frame}
 
+\begin{frame}{Univariate distribution functions}
+  \begin{itemize}
+    \item Generalized extreme value distribution (GEV):
+    \begin{align*}
+      \Pr(Y_j < y) = G_j(y) = \exp \left\{ -\left[ \left(1 + \xi_j \frac{y - \mu_j}{\sigma_j}\right)_+^{-1/\xi_j} \right] \right\}
+    \end{align*}
+    \item Generalized Pareto distribution (GPD):
+    \begin{align*}
+      \Pr(Y_j > y | Y_j > T) = F_j(y) = \left( 1 + \xi_j \frac{y - T}{\sigma_j} \right)_+^{-1/\xi_j}
+    \end{align*}
+  \end{itemize}
+\end{frame}
+
+\begin{frame}{Multivariate representations}
+  \begin{itemize}
+    \item Multivariate distributions:
+    \begin{itemize}
+      \item Assume common standardized max-stable marginal, like unit-Fr\'{e}chet
+      \begin{align*}
+        \Pr(Z < z) = exp(-z^{-1})
+      \end{align*}
+      \item The multivariate representation for the GEV is
+      \begin{align*}
+        \Pr(\bZ \le \bz)  &= G^*(\bz) = \exp(-V(\bz))\\
+                V(\bs)    &= d \int_{\Delta_d} \bigvee_{i = 1}^d \frac{w_i}{z_i} H(\ddd w)
+      \end{align*}
+      where
+      \begin{itemize}
+        \item $\Delta_d = \{ \bw \in \calR^d_+ \mid w_1 + \cdots + w_d = 1\}$
+        \item $H$ is a probability measure on $\Delta_d$
+        \item $\int_{\Delta_d}w_i H(\ddd w) = 1 / d$ for $i = 1, \ldots, d$.
+      \end{itemize}
+    \end{itemize}
+  \end{itemize}
+\end{frame}
+
+\begin{frame}{}
+  \begin{itemize} \setlength{\itemsep}{0.5em}
+    \item Although $V(\bs)$ can be flexible, the number of parameters is unwieldy.
+    \begin{itemize}
+      \item For an asymmetric logistic dependence, we have $2^d - d - 1$ free parameters
+    \end{itemize}
+    \item New model builds on existing geostatistical methods.
+  \end{itemize}
+\end{frame}
+
+\begin{frame}{Skew-normal distribution}
+  \begin{itemize} \setlength{\itemsep}{0.5em}
+    \item Assume observed data $Y_t(\bs)$ come from a skew-normal (Zhang and El-Shaarawi, 2012)
+    \begin{align*}
+      Y_t(\bs) = X_t(\bs)\beta + \alpha z_t + v_t(\bs)
+    \end{align*}
+    where
+    \begin{itemize} \setlength{\itemsep}{0.25em}
+      \item $\alpha \in \calR$ controls the skew
+      \item $z_t \iid N_{(0, \infty)}(0, \sigma^2_t)$ is a random effect
+      \item $v_t(\bs) \sim \text{MVN}(0, \Sigma)$ where $\Sigma_t$ is a spatial covariance matrix with variance $\sigma^2_t$
+    \end{itemize}
+  \end{itemize}
+\end{frame}
+
+\begin{frame}{Skew-normal distribution}
+  \begin{itemize} \setlength{\itemsep}{0.5em}
+    \item By conditioning on $\alpha$ and $z_t$
+    \begin{align*}
+      Y_t(\bs) \mid \alpha, z_t \sim \text{MVN}(\mu_t(\bs), \Sigma_t)
+    \end{align*}
+    where $\mu_t(\bs) = X_t(\bs) \beta + \alpha z_t$
+    \item When $\sigma_t^2 \iid \text{IG}(a, b)$, then this becomes a multivariate $t$-distribution
+    after marginalizing over the $\sigma_t^2$ terms.
+    \item Can use standard geostatistical methods to fit this model.
+    \item Predictions can be made through kriging.
+  \end{itemize}
+\end{frame}
+
+\begin{frame}{Random daily partition}
+  \begin{itemize} \setlength{\itemsep}{0.5em}
+    \item Consider a set of daily knots $\{w_{t1}, \ldots, w_{tK}\}$ that define a daily partition
+    $\{P_{t1}, \ldots, P_{tK}\}$ such that
+    \begin{align*}
+      P_{tk} = \{s : k = \argmin_\ell|| \bs - w_{t\ell}|| \}
+    \end{align*}
+    \item For $\bs \in P_{tk}$, the model becomes
+    \begin{itemize}
+      \item $\mu_t(\bs) = X_t(\bs) \beta + \alpha z_{tk}$
+      \item $\sigma^2_t = \sigma^2_{tk}$
+    \end{itemize}
+    \item After marginalizing over the $\sigma_{tk}^2$ terms, $Y_t(\bs)$ is a multivariate $t$-distribution
+    within the partition.
+  \end{itemize}
+\end{frame}
+
+\begin{frame}{Thresholding data}
+  \begin{itemize} \setlength{\itemsep}{0.5em}
+    \item Tails of the distribution should speak for themselves.
+    \item We threshold the observed data at a suitably high threshold $T$.
+    \item Observations below the threshold are imputed during MCMC.
+  \end{itemize}
+\end{frame}
+
+\begin{frame}{MCMC details}
+  \begin{itemize} \setlength{\itemsep}{0.5em}
+    \item Three main steps:
+    \begin{enumerate}[1.]
+      \item Impute missing observations and data below $T$
+      \item Update parameters with random walk Metropolis Hastings or Gibbs sampling
+      \item Make spatial predictions
+    \end{enumerate}
+  \end{itemize}
+\end{frame}
+
+\begin{frame}{Data analysis}
+  \begin{itemize} \setlength{\itemsep}{0.5em}
+    \item Ozone analysis at 85 sites in NC, SC, and GA for 92 days
+    \includegraphics[width=1\linewidth]{./plots/ozone_station.pdf}
+  \end{itemize}
+\end{frame}
+
+\begin{frame}{Model comparisons}
+  \begin{itemize} \setlength{\itemsep}{0.5em}
+    \item 9 different analysis methods
+    \begin{enumerate}[1.]
+      \item Gaussian
+      \item $t$, $K=1$, $T=0$
+      \item $t$, $K=5$, $T=0$
+      \item $t$, $K=1$, $T=0.9$
+      \item $t$, $K=5$, $T=0.9$
+      \item skew-$t$, $K=1$, $T=0$
+      \item skew-$t$, $K=1$, $T=0.9$
+      \item skew-$t$, $K=5$, $T=0$
+      \item skew-$t$, $K=5$, $T=0.9$
+    \end{enumerate}
+    \item Compare quantile scores using 5-fold cross validation
+  \end{itemize}
+\end{frame}
+
+\begin{frame}{References}
+  \begin{itemize} \setlength{\itemsep}{0.5em}
+    \item Cooley, D., Cisewski, J., Erhardt, R.
+J., Mannshardt, E., Omolo, B.
+O.
+and Sun, Y.
+(2012) A survey of spatial extremes: Measuring spatial dependence and modeling spatial effects.
+{\it REVSTAT}, 10, 135--165.
+    \item Zhang, H. and El-Shaarawi, A. (2010) On spatial skew-Gaussian processes and applications. {\it Environmetrics}, 33--47.
   \end{itemize}

### 2014-07-30 pot/code/analysis/ozone/SE/cv5-10SE.R

@@ -9,7 +9,7 @@ source('../../../R/mcmc.R')
 source('../../../R/auxfunctions.R')
 
 setting <- 10
-method <- "t"
+method <- "gaussian"
 nknots <- 1
 threshold <- 0
 outputfile <- paste("cv5-", setting, "SE.RData", sep="")

### 2014-07-30 pot/code/analysis/ozone/SE/cv5-11SE.R

@@ -10,7 +10,7 @@ source('../../../R/auxfunctions.R')
 
 setting <- 11
 method <- "t"
-nknots <- 5
+nknots <- 1
 threshold <- 0
 outputfile <- paste("cv5-", setting, "SE.RData", sep="")
 X <- X[, , c(1, 2, 3)]

### 2014-07-30 pot/code/analysis/ozone/SE/cv5-12SE.R

@@ -10,8 +10,8 @@ source('../../../R/auxfunctions.R')
 
 setting <- 12
 method <- "t"
-nknots <- 1
-threshold <- 0.90
+nknots <- 5
+threshold <- 0
 outputfile <- paste("cv5-", setting, "SE.RData", sep="")
 X <- X[, , c(1, 2, 3)]
 start <- proc.time()

### 2014-07-30 pot/code/analysis/ozone/SE/cv5-13SE.R

@@ -10,7 +10,7 @@ source('../../../R/auxfunctions.R')
 
 setting <- 13
 method <- "t"
-nknots <- 5
+nknots <- 1
 threshold <- 0.90
 outputfile <- paste("cv5-", setting, "SE.RData", sep="")
 X <- X[, , c(1, 2, 3)]

### 2014-07-30 pot/code/analysis/ozone/SE/cv5-18SE.R

@@ -33,12 +33,28 @@ for(val in 1:5){
 	S.p <- s.scale[val.idx,]
 	
 	tic.set <- proc.time()
-	fit[[val]] <- mcmc(y=y.o, s=S.o, x=X.o, x.pred=X.p, s.pred=S.p,
-	                   method=method, skew=F,
-	                   thresh=threshold, nknots=nknots, 
-                       iters=30000, burn=25000, update=1000, iterplot=F,
-                       beta.init=beta.init, tau.init=tau.init, rho.init=0.5,
-                       nu.init=0.5, alpha.init=0.5)
+	fit[[val]] <- tryCatch(
+	                mcmc(y=y.o, s=S.o, x=X.o, x.pred=X.p, s.pred=S.p,
+	                method=method, skew=T,
+	                thresh=threshold, nknots=nknots, 
+                    iters=30000, burn=25000, update=1000, iterplot=F,
+                    beta.init=beta.init, tau.init=tau.init, rho.init=0.5,
+                    nu.init=0.5, alpha.init=0.5, cov.model="exponential", rho.prior="cont"),
+                    error = function(e) {
+                      tryCatch(
+                        mcmc(y=y.o, s=S.o, x=X.o, x.pred=X.p, s.pred=S.p,
+	                    method=method, skew=T,
+	                    thresh=threshold, nknots=nknots, 
+                        iters=30000, burn=25000, update=1000, iterplot=F,
+                        beta.init=beta.init, tau.init=tau.init, rho.init=0.5,
+                        nu.init=0.5, alpha.init=0.5, cov.model="exponential", rho.prior="cont"),
+                        error = function(e) {
+                          cat("dataset", d, "not working \n")
+                              "no results"
+                          }
+                      )
+                    }
+                  )
 	toc.set <- proc.time()
 	time.set <- (toc.set - tic.set)[3]
 	

### 2014-07-31 pot/LaTeX/skewnormal.tex

@@
 \section{Introduction}\label{s:intro}
+In most climatological applications, researchers are interested in learning about the average behavior of different climate variables (e.g. ozone, temperature, ranfall).
+However, averages do not help regulators prepare for the unusual events that only happen once every 100 years.
+For example, it is important to have an idea of how much rain will come in a 100-year floor in order to construct strong enough river levees to protect lands from flooding.
+Unlike multivariate normal distributions, it is challenging to model multivariate extreme value distributions (e.g.
+generalized extreme value and generalized Pareto distribution) because few closed-form expressions exist for the density in more than two-dimensions \citep{Coles1991}.
+
+In a bivariate setting, the extremal coefficient describes the pairwise dependence between two spatial locations \citep{Smith1990}.
+Consider a spatial process $Y(\bs) \in \calR^n$ observed at locations $s \in \calD \subset \calR^2$.
+Then the extremal coefficient, $\theta(\bs_i, \bs_j) \in [1, 2]$, is defined as
+\begin{align}
+  \Pr(Y(\bs_i) < c, Y(\bs_j) < c) = \Pr(Y(\bs_i) < c)^{\theta(\bs_i, \bs_j) }. \label{eq:extremalcoef}
+\end{align}
+One way to characterize the dependence over the entire set of spatial locations is to calculate all of the pairwise extremal coefficients.
+Although this method provides information regarding the spatial structure of the observations, it does not fully characterize the joint spatial dependence.
+One way to account for this limitation is the use of pairwise composite likelihoods \citep{Padoan2010,Blanchet2011,Huser2013}.
+
+Recently, skew eliptical distributions have become a popular alternative to model correlated extreme values \citep{Genton2004,Zhang2010,Padoan2011}.
+Specifically, the skew-normal and skew-t distribution offer a flexible way to handle non-symmetric data within a framework of multivariate normal and multivariate t-distributions.
+In this paper, we present a modified version of the multivariate skew-t distribution that incorporates thresholding and random spatial partitions.
+The advantage of using a thresholded model as opposed to a non-thresholded model is that is allows for the tails of the distribution to inform the predictions in the tails \citep{DuMouchel1983}.
+The random spatial partitions increase the flexibility of the model by giving us a way to account for observations in spatial regions that may have higher variability due to certain regional climate variables (e.g.
+forest fires).
 
@@
 We first specify a model for the complete data, $Y_t(\bs)$, and then study the induced model for thresholded data, $\Yt_t(\bs)$.
-The full data model is given in Section \ref{s:model} assuming a skew normal distribution with a different variance each day.
+The full data model is given in Section \ref{s:data} assuming a skew normal distribution with a different variance each day.
 Computationally, the values below the threshold are updated using standard Bayesian missing data methods as described in Section \ref{s:comp}.
-The skew normal representation is from \citep{Minozzo2012} and is the sum of a normal and half-normal random variable.
+The skew normal representation is from \citep{Zhang2010}.
 \subsection{Half-normal distribution}\label{s:hn}
@@
 
-\subsection{Complete data}\label{s:model}
+\subsection{Complete data}\label{s:data}
 Consider a skew Gaussian spatial process
@@
 \end{align}
-where $z_t(\bs) = z_{tl}$ if $s \in P_{tl}$ where $P_{t1}, \ldots, P_{tL}$ form a partition, and $z_{tl} \iid N_{(0, \infty)}(0, 1)$, $\sigma_1 \in \calR$, $\sigma_2, \sigma_0 \in \calR^+$, and $v_t(\bs)$ is a spatial Gaussian process with mean zero and variance $\sigma_{tl}^2$.
+where $z_t(\bs) = z_{tl}$ if $s \in P_{tl}$ where $P_{t1}, \ldots, P_{tL}$ form a partition, and $z_{tl} \iid N_{(0, \infty)}(0, 1)$, $\sigma_1 \in \calR$, and $v_t(\bs)$ is a spatial Gaussian process with mean zero and variance $\sigma_{tl}^2$.
 It can be shown \citep{Zhang2010} that $Y_t(\bs)$ follows a skew normal distribution with skewness parameter $\alpha = \frac{ \sigma_1 }{ \sigma_{tl} }$.
@@
     z_{tl}(\bs) \mid \Theta &\sim N_{(0, \infty)}(0, \sigma^2_{tl})\\
-    v_t(\bs) \mid \Theta &\sim \Matern(0, \Sigma)\\
+    v_t(\bs) \mid \Theta &\sim \Matern(0, \Sigma_t(\lambda))\\
     \sigma^2_{tl} &\iid IG(\alpha, \beta)\\
@@
 \end{align}
-where $\Theta = \{w_{t1}, \ldots, w_{tL}, \beta, \sigma_t, \delta, \rho, \nu \}$; $l = \argmin_k ||\bs - w_k ||$; $\Sigma_t$ is a \Matern covariance matrix with variance $\sigma_{tl}^2 (1 - \delta^2)$, spatial range $\rho$ and smoothness $\nu$; and $\calD$ is the spatial domain of interest.
+where $\Theta = \{w_{t1}, \ldots, w_{tL}, \beta, \sigma_t, \alpha, \lambda, \rho, \nu \}$; $l = \argmin_k ||\bs - w_k ||$; $\Sigma_t(\lambda)$ is a \Matern covariance matrix with variance $\sigma_{tl}^2$, spatial range $\rho$, smoothness $\nu$, and $\lambda$ is the percentage of variation that can be attributed to the partial sill; and $\calD$ is the spatial domain of interest.
 
@@
 Then, we update $\Theta$ using random walk MH or Gibbs sampling when appropriate.
-Finally, we make spatial predictions.
-Each requires the joint distribution for the complete data given $\Theta$.
-As defined in \ref{eq:hier}, the distribution of $Y_t(\bs) \mid \Theta$ is the usual multivariate normal distribution with a \Matern spatial covariance structure.
+Finally, we make spatial precictions using conditional multivariate normal results and the fact that the distribution of $Y_t(\bs) \mid \Theta, z_{tl}$ is the usual multivariate normal distribution with a \Matern spatial covariance structure.
 
 \subsection{Imputation}\label{s:impute}
-We can use Gibbs sampling to update $\Yt_t(\bs)$ for observations that are below $T$, the thresholded value.
-Given $\Theta$, $Y_t(\bs)$ has truncated normal full conditional with these parameter values.
+We can use Gibbs sampling to update $Y_t(\bs)$ for observations that are below $T$, the thresholded value. Given $\Theta$, $Y_t(\bs)$ has truncated normal full conditional with these parameter values.
 So we sample $Y_t(\bs) \sim N_{(-\infty, T)}$

### 2014-07-31 pot/LaTeX/WNS.tex

\documentclass[slidestop,dvips]{beamer}
\usepackage{beamerthemesplit}
%\usepackage{graphics}
\usepackage{pstricks}
\usepackage{graphicx}

\setbeamercolor{title}{fg=red}
\setbeamercolor{frametitle}{fg=white}
\setbeamercolor{structure}{fg=red}

\title{Policy optimization for dynamic spatiotemporal systems \hspace{1em} \insertframenumber}
\author{Brian Reich, NCSU}

\newcommand{\btheta}{ \mbox{\boldmath $\theta$}}
\newcommand{\blambda}{ \mbox{\boldmath $\lambda$}}
\newcommand{\bmu}{ \mbox{\boldmath $\mu$}}
\newcommand{\bsigma}{ \mbox{\boldmath $\sigma$}}
\newcommand{\balpha}{ \mbox{\boldmath $\alpha$}}
\newcommand{\bbeta}{ \mbox{\boldmath $\beta$}}
\newcommand{\bdelta}{ \mbox{\boldmath $\delta$}}
\newcommand{\bgamma}{ \mbox{\boldmath $\gamma$}}
\newcommand{\brho}{ \mbox{\boldmath $\rho$}}
\newcommand{\bnu}{ \mbox{\boldmath $\nu$}}
\newcommand{\bpsi}{ \mbox{\boldmath $\psi$}}
\newcommand{\btau}{ \mbox{\boldmath $\tau$}}
\newcommand{\bepsilon}{ \mbox{\boldmath $\epsilon$}}
\newcommand{\beps}{ \mbox{\boldmath $\varepsilon$}}
\newcommand{\bX}{ \mbox{\boldmath X}}
\newcommand{\bS}{ \mbox{\boldmath S}}
\newcommand{\bx}{ {\bf x} }
\newcommand{\by}{ \mbox{\boldmath y}}
\newcommand{\bY}{ \mbox{\boldmath Y}}
\newcommand{\bz}{ \mbox{\boldmath z}}
\newcommand{\bZ}{ \mbox{\boldmath Z}}
\newcommand{\bW}{ \mbox{\boldmath W}}
\newcommand{\bb}{ \mbox{\boldmath b}}
\newcommand{\bg}{ \mbox{\boldmath g}}
\newcommand{\bL}{ \mbox{\boldmath L}}
\newcommand{\bs}{ \mbox{\boldmath s}}
\newcommand{\bt}{ \mbox{\boldmath t}}
\newcommand{\bu}{ \mbox{\boldmath u}}
\newcommand{\bv}{ \mbox{\boldmath v}}
\newcommand{\bw}{ \mbox{\boldmath w}}
\newcommand{\bfe}{ \mbox{\boldmath e}}
\newcommand{\calA}{{\cal A}}
\newcommand{\calI}{{\cal I}}

\newcommand{\argmin}{{\mathop{\rm arg\, min}}}
\newcommand{\argmax}{{\mathop{\rm arg\, max}}}
\newcommand{\iid}{\stackrel{iid}{\sim}}

\begin{document}

\frame{\frametitle{}
  \begin{center}
   \huge{Policy optimization for dynamic spatiotemporal systems}\vspace{24pt}

   \Large{Brian Reich}

   \vspace{12pt}

   \includegraphics[width=4.5in]{NCSU.ps}
  \end{center}

}

\frame{\frametitle{Joint work with}

\hspace{0.5in}   \includegraphics[width=1.25in, height=1.25in]{Drake.ps} \hspace{.5in}
   \includegraphics[width=1.25in, height=1.25in, trim=0in 0.5in 0in 0in, clip=true]{Laber.ps}\\
\hspace{0.5in}John Drake, UGA \hspace{0.55in}
Eric Laber, NCSU\vspace{12pt}

\hspace{0.5in}   \includegraphics[width=1.25in, height=1.25in]{Nick.ps} \hspace{.5in}
   \includegraphics[width=1.25in, height=1.25in]{Krishna.ps}\\
\hspace{0.5in}Nick Meyer, NCSU \hspace{0.5in}
Krishna Pacifici, NCSU

}

\frame{\frametitle{Motivating example: White nose syndrome}
\begin{itemize}
 \item White-nose syndrome (WNS) is an emerging disease affecting hibernating bats.\vspace{12pt}
 \item The first case was reported in 2007 in NY, and it has since spread rapidly throughout the Eastern US.\vspace{12pt}
  \item It is estimated that the bat population has declined by 80\% in affected areas. \vspace{12pt}
 \item This directly affects agriculture, as bats help control insects.\vspace{12pt}
 \item This pest control is extremely valuable. \vspace{12pt}
 \end{itemize}
}

\frame{\frametitle{Spread of white nose syndrome}
\vspace{-1in}
\begin{center}
  \includegraphics[height=4.5in,width=4.5in]{WNSmap.ps}
\end{center}
}

\frame{\frametitle{Some potential actions being considered and applied}

Both active and preventative treatments are being considered.\vspace{12pt}

 \begin{itemize}
   \item Close infected caves to human access.\vspace{12pt}
   \item Close uninfected caves to human access (currently being implemented by the Rocky Mountain region of USFWS).\vspace{12pt}
   \item Biocontrol measures, such as introducing beneficial bacteria into caves.\vspace{12pt}
   \item Create structures for bats that replace caves and that can be monitored and cleaned. \vspace{12pt}
 \end{itemize}

}

\frame{\frametitle{Statement of the general problem}
\begin{itemize}
 \item Our objective is to develop a procedure for identifying the {\bf optimal policy} for allocating treatments in space and time.\vspace{12pt}
 \item We define a specific reward criteria and try to find the policy which maximizes the expected reward. \vspace{12pt}
 \item The policy must balance short-term and long-term rewards.\vspace{12pt}
 \item We need the method to be interpretable and applicable in very high dimensions.\vspace{12pt}
 \end{itemize}
}

\frame{\frametitle{Other applications of this general methodology}
\begin{itemize}
 \item {\bf Conservation}: select lands to protect or designate as easements.\vspace{12pt}
 \item {\bf Wildlife management}: allocating catch sizes in each county.\vspace{12pt}
 \item {\bf Forest fires}: deciding where and when to do controlled burns.\vspace{12pt}
 \item {\bf Crime prevention}: assigning police officers to prevent crime.\vspace{12pt}
 \item {\bf Sampling}: determining where to take the next observation.
 \end{itemize}
}

\frame{\frametitle{Notation}
\begin{itemize}
 \item {\bf State}: $S_t = \{$all information collected by time $t\}$, including status of each county, covariates, spatial information, etc. \vspace{12pt}
 \item {\bf Action}: $A_t = \{$the counties to treat at time $t\}$. \vspace{12pt}
 \item {\bf Policy}: $\pi(S|\bgamma)$ ($\bgamma$ are parameters) maps states to actions $$A_t = \pi(S_t|\bgamma).$$
 \item {\bf Reward}: $R$, e.g., the number of uninfected counties in 2050. \vspace{12pt}
 \item {\bf Value}: $V_{\pi} = \mbox{E}_{\pi}(R)$ is the expected reward if we adhere to $\pi$.
 \end{itemize}
}

\frame{\frametitle{Reinforcement learning (RL)}
 There are many algorithms from RL (Sutton and Barto, 1998) that dynamically update the policy as data (state/action/reward) accrue, including:\vspace{12pt}
  \begin{itemize}
 \item Stochastic dynamic programming \vspace{12pt}
 \item Q-learning \vspace{12pt}
 \item SARSA \vspace{12pt}
 \item Actor-critic\vspace{12pt}
 \item Policy search\vspace{12pt}
 \item many, many more.
 \end{itemize}
}

\frame{\frametitle{Ms. Pacman}
\vspace{-0.3in}
\begin{center}
  \includegraphics[height=9in,width=9in]{Pacman.ps}
\end{center}
}

\frame{\frametitle{Ms. Pacman's policy}
\begin{itemize}
 \item Her actions are to move left, right, up, down or nowhere.\vspace{12pt}
 \item The states are the locations of Ms. Pacman ($s_{current}$), the balls, the bad guys, etc.\vspace{12pt}
 \item The reward is whether she wins or loses.\vspace{12pt}
 \item The policy is the set of rules that determine her movements. \vspace{12pt}
 \item If you had some time to kill, you could play the game thousands of times to estimate the value (probability of winning the game) of any policy you could dream up.\vspace{12pt}
 \item Based on these data, you could pick the best policy to use in the future.
 \end{itemize}
}

\frame{\frametitle{What might Ms. Pacman's policy look like?}
\begin{itemize}
 \item We need to decide where to go for any configuration of balls and bad guys.\vspace{12pt}
 \item We could simplify the problem by extracting a few features from the states.\vspace{12pt}
 \begin{enumerate}
  \item $F_1(S_t,A) =$ distance from $s_{current}+A$ to the nearest ball
  \item $F_2(S_t,A) =$ distance from $s_{current}+A$ to the nearest bad guy.\vspace{12pt}
 \end{enumerate}
 \item Policy:
 $$
    A_t= \pi(S_t;\bgamma) = \underset{A}{\argmax} \left\{F_1(S_t,A)\gamma_1 + F_2(S_t,A)\gamma_2\right\}.
 $$
 \item We could then find the $\bgamma$ that maximizes the value to obtain the optimal policy in this interpretable class of policies.

 \end{itemize}
}

\frame{\frametitle{How is white nose syndrome different than Ms. Pacman?}
\begin{itemize}
 \item We only get to play the game once, and not for long.\vspace{6pt}
 \item We have some time to make recommendations.\vspace{6pt}
 \item Our state and action spaces are spatial and huge (2$^n$ states)!\vspace{6pt}
 \item We want to exploit the scientific knowledge of our collaborators.\vspace{6pt}
 \item We want a policy with a simple form to share with stake-holders.\vspace{6pt}
 \item We need a policy that evolves over time and accounts for uncertainty.\vspace{6pt}
 \end{itemize}

Our contribution is to combine methods from RL, ecology, and statistics to accomplish these objectives within a unified framework.

}

\frame{\frametitle{Proposed policy search method}
   We select a simple parametric form for the policy in terms of features constructed from the current state and potential actions:
$$A_t = \pi(S_t;\bgamma) = \underset{A\in\calA_t}{\argmax} \sum_{l=1}^pF_l(S_t,A)\gamma_l.$$
\begin{itemize}
 \item $A_t$ is the recommended action, for example, the set of $m$ counties to be treated.\vspace{12pt}

 \item $\calA_t$ is the set of possible actions at time $t$, for example, all combinations of $m$ counties of uninfected counties.
 \end{itemize}
}

\frame{\frametitle{Proposed policy search method}
   We select a simple parametric form for the policy in terms of features constructed from the current state and potential actions:
$$A_t = \pi(S_t;\bgamma) = \underset{A\in\calA_t}{\argmax} \sum_{l=1}^pF_l(S_t,A)\gamma_l.$$
\begin{itemize}
 \item $F_l$ are the features and can be any function of the states (all information available at time $t$) and actions.\vspace{12pt}
 \item Feature selection has the usual regression tradeoffs:
 \begin{itemize}
    \item Selecting many features gives a flexible class of policies
     \item Selecting only a few features favors transparency.\vspace{12pt}
 \end{itemize}
 \item $\bgamma$ weights the features and must be optimized.
 \end{itemize}
}

\frame{\frametitle{Proposed policy search method}
   We select a simple parametric form for the policy in terms of features constructed from the current state and potential actions:
$$A_t = \pi(S_t;\bgamma) = \underset{A\in\calA_t}{\argmax} \sum_{l=1}^pF_l(S_t,A)\gamma_l.$$
\begin{itemize}
 \item The ``myopic policy'' would take $p=1$ and $F_1(S_t,A)$ to be the posterior predictive mean number of sites infected at time $t+1$ if we treated sites $A$.\vspace{12pt}
 \item Computing this feature requires fitting a spatiotemporal model.\vspace{12pt}
 \item This is optimal if we are only concerned with the next time step, but we can do better by considering future time steps.

 \end{itemize}
}

\frame{\frametitle{Feature selection}
 Which features will help us beat the myopic policy?  We use information about medium-term predictions and network structure.\vspace{12pt}
 \begin{itemize}
 \item Average transmission probability from treated counties to other counties.\vspace{12pt}
 \item Subgraph connectivity of the treated sites.\vspace{12pt}
 \item Betweenness connectivity of the treated sites.\vspace{12pt}
 \item Interactions.\vspace{12pt}
 \end{itemize}
}

\frame{\frametitle{Outline of the computational algorithm}
\begin{enumerate}
 \item Fit a spatiotemporal model using data available at time $t$.\vspace{12pt}
  \begin{itemize}
    \item Gravity model (Maher et al., 2012) with time-evolving parameters.\vspace{12pt}
  \end{itemize}
 \item Obtain feature weights $\bgamma$ for use at time $t$.\vspace{12pt}
  \begin{itemize}
    \item The value of the policy indexed by a candidate $\bgamma$ is estimated using Monte Carlo sampling from the posterior predictive distribution.\vspace{12pt}
    \item The optimal weight is found using stochastic gradient descent.\vspace{12pt}
  \end{itemize}
 \item Apply recommended treatments.\vspace{12pt}
 \item Repeat steps 1-3 each year until the end of the study.
 \end{enumerate}
}

\frame{\frametitle{1. Spatiotemporal gravity model}
\begin{itemize}
 \item Let $Y_{jt} = 1$ if county $j$ is infected in year $t$ and $Y_{jt}=0$ otherwise.\vspace{12pt}

\item Given the history up to time $t$ ($H_{t}$), the probability of an infection in county $j$ at time $t$ is
$$  \mbox{Prob}(Y_{jt}=1|H_{t}) = \left\{
                       \begin{array}{ll}
                         1 & Y_{jt-1}=1 \\
                         1-\prod_{l\in \calI_{t-1}}(1-p_{jlt}) & Y_{jt-1}=0
                       \end{array}
                     \right.$$
  \item $\calI_t$ is the set of indices of counties that are infected in year $t$.\vspace{12pt}
 \item $p_{jlt}$ is the probability of a spread from county $l$ to county $j$ in year $t$.
 \end{itemize}
}

\frame{\frametitle{1. Spatiotemporal gravity model}

The infection probability is modeled as:
$$\mbox{logit}(p_{jlt}) = \bX_{jt}^T\bbeta_t + \alpha_t A_{jt}- \rho_t \frac{d_{jl}}{(m_jm_l)^{\nu}} .$$
\begin{itemize}
 \item $\bX_{jt}$ are covariates and $\bbeta_t$ are dynamic coefficients.\vspace{12pt}
 \item $A_{jt}$ indicates treatment, and $\alpha_t$ is the treatment effect.\vspace{12pt}
 \item The final term controls spatial dependence:\vspace{12pt}
 \begin{itemize}
  \item $d_{jl}$ is the distance between counties,\vspace{6pt}
   \item $m_j$ is the number of caves in county $j$,\vspace{6pt}
   \item $\rho_t$ and $\nu$ are unknown parameters.
  \end{itemize}
 \end{itemize}
}

\frame{\frametitle{2. Obtain feature weights $\bgamma$ for use at time $t$}

\begin{itemize}
 \item The value of policy $\pi(S|\bgamma)$, denoted $V_{\bgamma}$, is the expected reward if we adhere to $\pi(S|\bgamma)$.\vspace{12pt}
 \item To approximate the value, we sample many realizations of the process conditional on the current state with actions taken following $\pi(S|\bgamma)$.\vspace{12pt}
 \item These posterior predictive draws account for uncertainty in the model parameters.\vspace{12pt}
 \item These simulations could also account for any other sources of uncertainty that can be encoded in the statistical model, such as changes in the climate or land use.\vspace{12pt}
 \end{itemize}
}

\frame{\frametitle{2. Obtain feature weights $\bgamma$ for use at time $t$}
Optimization is done through stochastic gradient descent:\vspace{12pt}
\begin{enumerate}
\setcounter{enumi}{-1}
 \item Set some initial value $\bgamma_0$ and compute its value $V_0$.\vspace{12pt}
 \item Generate $\bgamma^* \sim \mbox{N}(\bgamma_t,c I_p)$ and compute its value $V^*$.\vspace{12pt}
 \item Set $\bgamma_{t+1} = \bgamma_t+a_t(V_t-V^*)(\bgamma^*-\bgamma_t)$.\vspace{12pt}
 \item Repeat 1-2 until convergence.\vspace{12pt}
 \end{enumerate}
 $a_t$ and $c$ are tuning parameters.
}

\frame{\frametitle{3. Apply recommended treatments}

\begin{itemize}
 \item Given a model and policy (which give the features and $\gamma_l$), the recommended action is
 $$A_t = \pi(S_t;\bgamma) = \underset{A\in\calA_t}{\argmax} \sum_{l=1}^pF_l(S_t,A)\gamma_l.$$
 \item This is a challenging discrete optimization problem.\vspace{12pt}
 \item It must be done efficiently because it is also used in Monte Carlo simulations.\vspace{12pt}
 \item We begin by adding sites sequentially one at a time.\vspace{12pt}
 \item Once a full set is proposed, we cycle through the proposed sites and optimize them one at a time with the others fixed until convergence.
 \end{itemize}
}

\frame{\frametitle{Results: Model comparisons}

We compared six variants of the gravity model
$$\mbox{logit}(p_{jlt}) = \bX_{jt}^T\bbeta_t + \rho_t \frac{d_{jl}}{(m_jm_l)^{\nu}} .$$

\begin{enumerate}
 \item Static ($\bbeta_t\equiv \bbeta$), non-spatial ($\rho_t\equiv 0$)\vspace{12pt}
 \item Static, diffusion ($\nu\equiv 0$)\vspace{12pt}
 \item Static, gravity\vspace{12pt}
 \item Dynamic, non-spatial\vspace{12pt}
 \item Dynamic, diffusion\vspace{12pt}
 \item Dynamic, gravity\vspace{12pt}
 \end{enumerate}
}

\frame{\frametitle{Results: Model comparisons}

The dynamic gravity model has the smallest (best) DIC.\vspace{24pt}

\begin{center}\begin{tabular}{llccc}
Spatial model & Coefficient model & $DIC$ & ${\bar D}$ & $p_D$ \\\hline
Non-spatial & Static  & 1307 & 1302 &  4.9 \\
Diffusion   & Static  & 1003 &  997 &  5.8 \\
Gravity     & Static  &  939 &  933 &  6.4 \\
Non-spatial & Dynamic & 1223 & 1206 & 16.7 \\
Diffusion   & Dynamic &  911 &  893 & 17.7 \\
Gravity     & Dynamic &  878 &  860 & 18.0 \\
\end{tabular}\end{center}}

\frame{\frametitle{Results: Model adequacy}

\begin{itemize}
  \item In addition to comparing a few models, we want to show the dynamic gravity model is adequate.\vspace{12pt}
  \item This is crucial, since our entire strategy relies on having a good predictive model.\vspace{12pt}
  \item We use the Bayesian p-value:\vspace{12pt}
  \begin{itemize}
  \item Let $D_0$ be some summary statistic of the dataset, e.g., the number of infected counties in 2010.\vspace{12pt}
  \item We sample $N = 10,000$ draws from the process, and compute $D_i$ for simulated dataset $i$.\vspace{12pt}
  \item The p-value is then $p = \frac{1}{N}\sum_{i=1}^NI(D_0>D_i)$.
\end{itemize}
\end{itemize}

}

\frame{\frametitle{Results: Bayes p-values for the dynamic gravity model}

\begin{center}\begin{tabular}{lccc}
Statistic & Observed value & Predictive distribution & P-value \\\hline
n inf & 148 & 149 ( 60 , 325 ) & 0.49 \\
n inf 2007 & 1 & 1 ( 0 , 6 ) & 0.35 \\
n inf 2008 & 14 & 10 ( 3 , 27 ) & 0.70 \\
n inf 2009 & 26 & 26 ( 9 , 67 ) & 0.50 \\
n inf 2010 & 35 & 36 ( 13 , 97 ) & 0.46 \\
n inf 2011 & 38 & 37 ( 13 , 82 ) & 0.52 \\
n inf 2012 & 33 & 36 ( 13 , 68 ) & 0.40 \\
mean year & 2010 & 2010 ( 2010 , 2011 ) & 0.37 \\
mean long & 1399 & 1347 ( 879 , 1605 ) & 0.64 \\
mean lat & 393 & 388 ( 225 , 595 ) & 0.52 \\
%mean dist from start & 608 & 653 ( 373 , 1095 ) & 0.38 \\
min long & -290 & -97 ( -1816 , 817 ) & 0.47 \\
min lat & -258 & -335 ( -826 , -187 ) & 0.80 \\
max long & 2156 & 2051 ( 1950 , 2201 ) & 0.82 \\
max lat & 1118 & 1118 ( 1016 , 1325 ) & 0.49 \\
%max dist from start & 2218 & 2072 ( 1325 , 3666 ) & 0.51
\end{tabular}\end{center}}

\frame{\frametitle{Results: Bayes p-values for the static gravity model}

\begin{center}\begin{tabular}{lccc}
Statistic & Observed value & Predictive distribution & P-value \\\hline
n inf & 148 & 97 ( 25 , 222 ) & 0.81 \\
n inf 2007 & 1 & 2 ( 0 , 6 ) & 0.11 \\
n inf 2008 & 14 & 4 ( 0 , 11 ) & 0.99 \\
n inf 2009 & 26 & 8 ( 1 , 22 ) & 0.99 \\
n inf 2010 & 35 & 15 ( 3 , 40 ) & 0.95 \\
n inf 2011 & 38 & 26 ( 6 , 65 ) & 0.76 \\
n inf 2012 & 33 & 40 ( 10 , 91 ) & 0.35 \\
mean year & 2010 & 2011 ( 2010 , 2011 ) & 0.01 \\
mean long & 1399 & 1415 ( 738 , 1730 ) & 0.47 \\
mean lat & 393 & 447 ( 264 , 742 ) & 0.32 \\
%mean dist from start & 608 & 568 ( 237 , 1208 ) & 0.57 \\
min long & -290 & 192 ( -1766 , 1346 ) & 0.43 \\
min lat & -258 & -265 ( -764 , 162 ) & 0.51 \\
max long & 2156 & 2014 ( 1900 , 2156 ) & 0.88 \\
max lat & 1118 & 1106 ( 965 , 1325 ) & 0.61 \\
%max dist from start & 2218 & 1768 ( 728 , 3548 ) & 0.57 \\
\end{tabular}\end{center}}

\frame{\frametitle{Results: DLM coefficients $\bbeta_t$}

\vspace{-0.3in}
\begin{center}
  \includegraphics[height=3.3in,width=3.3in]{dlm.ps}
\end{center}
}

\frame{\frametitle{Simulation study}

\begin{itemize}
  \item To illustrate the proposed method, we estimate the value of several policies for data similar to the WNS data.\vspace{12pt}
  \item We generate data from the gravity model with parameters set to the posterior mean.\vspace{12pt}
  \item We consider both active and preventative treatments, with effect size that roughly halves the infection probabilities.\vspace{12pt}
  \item Data are generated for 2007-2022 with treatments starting in 2012.\vspace{12pt}
  \item The reward is the number of uninfected counties in 2022.\vspace{12pt}
  \item We are allowed 30 treatments of each type each year.
\end{itemize}

}

\frame{\frametitle{Methods for comparison}

We compare 3 policies:\vspace{12pt}
\begin{itemize}
  \item No treatments.\vspace{12pt}
  \item Nearest neighbor: Treat the counties nearest to an infected county, and infected counties nearest an uninfected county.\vspace{12pt}
  \item The proposed policy search method.\vspace{12pt}
\end{itemize}

We generate 100 datasets and record the number of uninfected counties for each method.
}

\frame{\frametitle{\ }
   \begin{center}
   {\huge One realization with no treatments}
   \end{center}
}
\frame{\frametitle{\ }
   \begin{center}
   {\huge One realization using the nearest neighbor policy}
   \end{center}
}
\frame{\frametitle{\ }
   \begin{center}
   {\huge One realization using policy search}
   \end{center}
}

\frame{\frametitle{Proportion of infected counties}

\vspace{-0.3in}
\begin{center}
  \includegraphics[height=3.3in,width=3.3in]{valuePlot.ps}
\end{center}

}

\frame{\frametitle{Monte Carlo 90\% intervals for $\bgamma$}

\begin{center}\begin{tabular}{l|c}
Feature & 90\% Interval\\\hline
AIP over S & (-1.89 , -1.23 )\\
AIP over neighbors of T  & (-3.73 , -3.36 )\\
AIP over S weighted by SC & (-5.18 , -4.64 )\\
AIP over S weighted by BC & (-1.81 , -1.30 )\\
AIP over NI weighed by SC & (\ 2.58 , \ 2.85 )\\
AIP over NI weighed by BC & (-2.01 , -1.53 )\\
AIP over S weighted by WM & (-0.93 , -0.63 )\\
AIP over T weighted by WM & (-1.63 , -1.05 )\\
\end{tabular}\end{center}

\begin{itemize}
  \item AIP = average infection probability.
  \item Set of counties: S = susceptible; T = treated; \\
        NI = susceptible neighbors of infected counties.
  \item Connectivity measures:  SC = subgraph; BC = betweenness; WM = walk measure.
\end{itemize}

}

\frame{\frametitle{Summary}

 \begin{itemize}
 \item We have proposed a general framework for policy optimization for spatial decision problems. \vspace{6pt}
 \item The key features are that we attain an interpretable policy and can handle non-stationarity and high dimensions.\vspace{6pt}
 \item Our simulations illustrate the potential of this approach. \vspace{6pt}
 \item Limitations:\vspace{0pt}
 \begin{itemize}
    \item Entirely model-based.\vspace{6pt}
    \item Doesn't find the globally optimal policy.\vspace{6pt}
    \item Not applicable in real time.\vspace{6pt}
    \item Practical constraints?\vspace{6pt}
 \end{itemize}
 \item Future work:\vspace{0pt}
 \begin{itemize}
    \item Generalizing.\vspace{6pt}
    \item Applications!\vspace{6pt}
 \end{itemize}

 \end{itemize}

}

\frame{\frametitle{Conservation in Puerto Rico}

\vspace{-0.1in}
\begin{center}
  \includegraphics[height=3in,width=3in,  trim=0.5in 0.5in 0.5in 0.5in, clip=true]{PR.ps}
\end{center}
}

\frame{\frametitle{\ }
  \vspace{1in}
  \begin{center}
   {\Huge Thank you for your attention!}
  \end{center}
}

\end{document}

### 2014-07-31 pot/LaTeX/jsm2014.tex

@@
 }
-\author[B. Reich and S. Morris]{Brian Reich and Samuel Morris}
+\author[S. Morris and B. Reich]{Samuel Morris and Brian Reich}
 \institute[NCSU]{North Carolina State University}
@@
     \item In geostatistical analysis, kriging uses spatial correlation to help inform prediction at unknown locations.
-    \item Want to explore ways to incorporate this spatial correlation when estimating the tails of the distribution.
+    \item Want to explore computationally easy methods that are available in higher dimensions
   \end{itemize}
@@
 
-\begin{frame}{Introduction to extremes}
+% \begin{frame}{Introduction to extremes}
+%   \begin{itemize} \setlength{\itemsep}{0.5em}
+%     \item Max-stable processes (Cooley et al., 2012):
+%     \begin{itemize}
+%       \item Consider a spatial process $x_t(\bs)$, $t = 1, \ldots, T$.
+%       \item Let $M_T(\bs) = \left\{ \bigvee_{t=1}^T x_t(\bs_1), \ldots, \bigvee_{t=1}^T x_t(\bs_n) \right\}$
+%       \item If there exists normalizing sequences $a_T(\bs)$ and $b_T(\bs)$ such
+%       that for all sites, $\bs_i, i = 1, \ldots, d$,
+%       \begin{align*}
+%         a_T^{-1}(\bs) \left\{ M_T(\bs) - b_T(\bs) \right\} \converged Y(\bs)
+%       \end{align*}
+%       which has a non-degenerate distribution, then $Y(\bs)$ is a max-stable process.
+%     \end{itemize}
+%   \end{itemize}
+% \end{frame}
+
+\begin{frame}{Standard non-spatial analysis}
   \begin{itemize} \setlength{\itemsep}{0.5em}
-    \item Max-stable processes (Cooley et al., 2012):
-    \begin{itemize}
-      \item Consider a spatial process $x_t(\bs)$, $t = 1, \ldots, T$.
-      \item Let $M_T(\bs) = \left\{ \bigvee_{t=1}^T x_t(\bs_1), \ldots, \bigvee_{t=1}^T x_t(\bs_n) \right\}$
-      \item If there exists normalizing sequences $a_T(\bs)$ and $b_T(\bs)$ such
-      that for all sites, $\bs_i, i = 1, \ldots, d$,
-      \begin{align*}
-        a_T^{-1}(\bs) \left\{ M_T(\bs) - b_T(\bs) \right\} \converged Y(\bs)
-      \end{align*}
-      which has a non-degenerate distribution, then $Y(\bs)$ is a max-stable process.
-    \end{itemize}
-  \end{itemize}
-\end{frame}
-
-\begin{frame}{General approaches}
-  \begin{itemize}
-    \item Two standard approaches:
-    \begin{enumerate}[1.] \setlength{\itemsep}{0.5em}
     \item Block maxima:
@@
       \item Discards many observations
+      \item Models are fit using the generalized extreme value distribution
     \end{itemize}
-      \item Peaks-over-threshold
-      \begin{itemize}
-        \item Incorporates more data than block maxima
-        \item Autocorrelation may be an issue between observations (e.g. flood levels don't dissipate overnight)
-      \end{itemize}
-    \end{enumerate}
-  \end{itemize}
-\end{frame}
-
-\begin{frame}{Univariate distribution functions}
-  \begin{itemize}
     \item Generalized extreme value distribution (GEV):
@@
     \end{align*}
+  \end{itemize}
+\end{frame}
+
+\begin{frame}{Standard non-spatial analysis}
+  \begin{itemize}
+    \item Peaks-over-threshold:
+    \begin{itemize}
+      \item Incorporates more data than block maxima
+      \item Select a threshold, $T$, and fit data above the threshold using the generalized Pareto distribution
+      \item Autocorrelation may be an issue between observations (e.g. flood levels don't dissipate overnight)
+    \end{itemize}
     \item Generalized Pareto distribution (GPD):
@@
 
-\begin{frame}{Multivariate representations}
+% \begin{frame}{Multivariate representations}
+%   \begin{itemize}
+%     \item Multivariate distributions:
+%     \begin{itemize}
+%       \item Assume common standardized max-stable marginal, like unit-Fr\'{e}chet
+%       \begin{align*}
+%         \Pr(Z < z) = exp(-z^{-1})
+%       \end{align*}
+%       \item The multivariate representation for the GEV is
+%       \begin{align*}
+%         \Pr(\bZ \le \bz)  &= G^*(\bz) = \exp(-V(\bz))\\
+%                 V(\bs)    &= d \int_{\Delta_d} \bigvee_{i = 1}^d \frac{w_i}{z_i} H(\ddd w)
+%       \end{align*}
+%       where
+%       \begin{itemize}
+%         \item $\Delta_d = \{ \bw \in \calR^d_+ \mid w_1 + \cdots + w_d = 1\}$
+%         \item $H$ is a probability measure on $\Delta_d$
+%         \item $\int_{\Delta_d}w_i H(\ddd w) = 1 / d$ for $i = 1, \ldots, d$.
+%       \end{itemize}
+%     \end{itemize}
+%   \end{itemize}
+% \end{frame}
+
+\begin{frame}{Multivariate analysis}
+  \begin{itemize} \setlength{\itemsep}{0.5em}
+    \item Multivariate max-stable and GPD models have nice features, but they are
     \begin{itemize}
-    \item Multivariate distributions:
+      \item computationally hard to work with
+      \item joint distribution only available in low dimension
+    \end{itemize}
+    \item Pairwise likelihood approach (Huser and Davison, 2014)
+  \end{itemize}
+\end{frame}
+
+\begin{frame}{Model objectives}
+  \begin{itemize} \setlength{\itemsep}{0.5em}
+    \item Our objective is to build a model that
     \begin{itemize}
-      \item Assume common standardized max-stable marginal, like unit-Fr\'{e}chet
-      \begin{align*}
-        \Pr(Z < z) = exp(-z^{-1})
-      \end{align*}
-      \item The multivariate representation for the GEV is
-      \begin{align*}
-        \Pr(\bZ \le \bz)  &= G^*(\bz) = \exp(-V(\bz))\\
-                V(\bs)    &= d \int_{\Delta_d} \bigvee_{i = 1}^d \frac{w_i}{z_i} H(\ddd w)
-      \end{align*}
-      where
-      \begin{itemize}
-        \item $\Delta_d = \{ \bw \in \calR^d_+ \mid w_1 + \cdots + w_d = 1\}$
-        \item $H$ is a probability measure on $\Delta_d$
-        \item $\int_{\Delta_d}w_i H(\ddd w) = 1 / d$ for $i = 1, \ldots, d$.
-      \end{itemize}
+      \item has a flexible tail
+      \item has asymptotic spatial dependence
+      \item computation on the order of Gaussian models for large space-time datasets
     \end{itemize}
@@
 
-\begin{frame}{}
+\begin{frame}{Thresholding data}
   \begin{itemize} \setlength{\itemsep}{0.5em}
-    \item Although $V(\bs)$ can be flexible, the number of parameters is unwieldy.
-    \begin{itemize}
-      \item For an asymmetric logistic dependence, we have $2^d - d - 1$ free parameters
-    \end{itemize}
-    \item New model builds on existing geostatistical methods.
+    \item We threshold the observed data at a high threshold $T$.
+    \item Thresholded data:
+    \begin{align*}
+      Y_t^*(\bs) = \left\{ \begin{array}{ll}
+          Y_t(\bs) \quad & Y_t(\bs) > T\\
+          T \quad & Y_t(\bs) \le T
+      \end{array}\right.
+    \end{align*}
+    \item Allows tails of the distribution to speak for themselves.
   \end{itemize}
@@
 
-\begin{frame}{Skew-normal distribution}
+\begin{frame}{Spatial skew-$t$ distribution}
   \begin{itemize} \setlength{\itemsep}{0.5em}
-    \item Assume observed data $Y_t(\bs)$ come from a skew-normal (Zhang and El-Shaarawi, 2012)
+    \item Assume observed data $Y_t(\bs)$ come from a skew-$t$ (Zhang and El-Shaarawi, 2012)
     \begin{align*}
@@
       \item $z_t \iid N_{(0, \infty)}(0, \sigma^2_t)$ is a random effect
-      \item $v_t(\bs) \sim \text{MVN}(0, \Sigma)$ where $\Sigma_t$ is a spatial covariance matrix with variance $\sigma^2_t$
+      \item $v_t(\bs)$ is a Gaussian process with variance $\sigma^2_t$ and \Matern correlation
+      \item $\sigma_t^2 \iid \text{IG}(a, b)$
     \end{itemize}
@@
   \begin{itemize} \setlength{\itemsep}{0.5em}
-    \item By conditioning on $\alpha$ and $z_t$
-    \begin{align*}
-      Y_t(\bs) \mid \alpha, z_t \sim \text{MVN}(\mu_t(\bs), \Sigma_t)
-    \end{align*}
-    where $\mu_t(\bs) = X_t(\bs) \beta + \alpha z_t$
-    \item When $\sigma_t^2 \iid \text{IG}(a, b)$, then this becomes a multivariate $t$-distribution
-    after marginalizing over the $\sigma_t^2$ terms.
+    \item Conditioned on $z_t$ and $\sigma^2_t$, $Y_t(\bs)$ is Gaussian
+    \begin{itemize}
+      \item After marginalizing over $\sigma^2_t$, this becomes a multivariate $t$-distribution with $2a$ degrees of freedom.
+    \end{itemize}
     \item Can use standard geostatistical methods to fit this model.
     \item Predictions can be made through kriging.
+    \item Marginalizing over $z_t$ and $\sigma^2_t$ (via MCMC),
+    \begin{align*}
+      Y_t(\bs) \sim \text{skew-t}(\mu, \alpha, \text{df}=2a)
+    \end{align*}
+  \end{itemize}
+\end{frame}
+
+\begin{frame}{Long-range dependence}
+  \begin{itemize} \setlength{\itemsep}{0.5em}
+    \item The $\chi$ coefficient is a measure of extremal spatial correlation
+    \begin{align*}
+      \chi(\bh) = \Pr(Y_t(\bs) > c \mid Y_t(\bs + \bh) > c)
+    \end{align*}
+    \item This value shows asymptotic dependence that does not approach 0 as $\bh \rightarrow \infty$.
+    \item Deal with this through a daily random partition.
   \end{itemize}
@@
   \begin{itemize} \setlength{\itemsep}{0.5em}
+    \item Daily random partition allows $z_t$ and $\sigma^2_t$ to vary by site.
+    \begin{align*}
+      Y_t(\bs) = X_t(\bs) \beta + \alpha z_t(\bs) + v_t(\bs)
+    \end{align*}
     \item Consider a set of daily knots $\{w_{t1}, \ldots, w_{tK}\}$ that define a daily partition
@@
     \end{align*}
-    \item For $\bs \in P_{tk}$, the model becomes
-    \begin{itemize}
-      \item $\mu_t(\bs) = X_t(\bs) \beta + \alpha z_{tk}$
-      \item $\sigma^2_t = \sigma^2_{tk}$
-    \end{itemize}
-    \item After marginalizing over the $\sigma_{tk}^2$ terms, $Y_t(\bs)$ is a multivariate $t$-distribution
-    within the partition.
-  \end{itemize}
-\end{frame}
-
-\begin{frame}{Thresholding data}
-  \begin{itemize} \setlength{\itemsep}{0.5em}
-    \item Tails of the distribution should speak for themselves.
-    \item We threshold the observed data at a suitably high threshold $T$.
-    \item Observations below the threshold are imputed during MCMC.
+    \item For $\bs \in P_{tk}$
+    \begin{align*}
+      z_t(\bs) &= z_{tk}\\
+      \sigma^2_t(\bs) &= \sigma^2_{tk}
+    \end{align*}
+    \item Within each partition $Y_t(\bs)$ has the same distribution as before.
   \end{itemize}
@@
     \end{enumerate}
+    \item Priors are selected to be conjugate when possible.
   \end{itemize}
@@
   \begin{itemize} \setlength{\itemsep}{0.5em}
-    \item 9 different analysis methods
+    \item 9 different analysis methods incorporating
+    \begin{itemize}
+      \item Gaussian vs $t$
+      \item 1 partition vs 5 partitions
+      \item No thresholding vs thresholded
+      \begin{itemize}
+        \item Thresholded data at 0.90 sample quantile
+      \end{itemize}
+    \end{itemize}
+    \item All methods use a \Matern or exponential covariance ($\nu = 0.5$)
+    \item Compare quantile and Brier scores using 5-fold cross validation
+    \item Mean function modeled using a first-order spatial trend
+  \end{itemize}
+\end{frame}
+
+\begin{frame}{Cross-validation results}
+  \begin{table}[htbp]
+    \small
+    \centering
+    \begin{tabular}{lrrrrrr}
+           & \multicolumn{6}{c}{Quantile}\\
+           \cline{2-7}
+    & 0.900 & 0.950 & 0.980 & 0.990 & 0.995 & 0.999\\
+  \hline
+Gaussian & 16.38 & 15.76 & 15.02 & 14.52 & 14.08 & 13.22\\
+$t$-1 ($T=0$) & 16.15 & 15.51 & 14.62 & 14.00 & 13.43 & 12.32\\
+$t$-5 ($T=0$) & 13.61 & 12.66 & 11.61 & 10.96 & 10.40 & 9.34\\
+$t$-1 ($T=0.9$) & 5.52  & 3.58  & 2.28  & 1.77  & 1.47  & 1.10\\
+$t$-5 ($T=0.9$) & 5.98  & 4.27  & 2.99  & 2.41  & 2.03  & 1.49\\
+skew $t$-1 ($T=0$) & 9.24  & 7.27  & 5.26  & 4.13  & 3.27  & 1.96\\
+skew $t$-1 ($T=0.9$) & {\bf 4.91}  & {\bf 3.16} & {\bf 1.94}  & {\bf 1.45}  & {\bf 1.16}  & {\bf 0.82}\\
+skew $t$-5 ($T=0$) & 15.81 & 14.46 & 12.74 & 11.57 & 10.57 & 8.60\\
+skew $t$-5 ($T=0.9$) & & & & & &\\
+\hline
+    \end{tabular}
+  \end{table}
+  \begin{itemize} \setlength{\itemsep}{0.5em}
+    \item Quantile scores -- Still waiting on last set
+  \end{itemize}
+\end{frame}
+
+\begin{frame}{Simulation study settings}
+  \begin{itemize} \setlength{\itemsep}{0.5em}
+    \item Data generated from 6 different settings.
     \begin{enumerate}[1.]
       \item Gaussian
-      \item $t$, $K=1$, $T=0$
-      \item $t$, $K=5$, $T=0$
-      \item $t$, $K=1$, $T=0.9$
-      \item $t$, $K=5$, $T=0.9$
-      \item skew-$t$, $K=1$, $T=0$
-      \item skew-$t$, $K=1$, $T=0.9$
-      \item skew-$t$, $K=5$, $T=0$
-      \item skew-$t$, $K=5$, $T=0.9$
+      \item $t$-1 with 4 degrees of freedom
+      \item $t$-5 with 4 degrees of freedom
+      \item skew $t$-1 with 4 degrees of freedom ($\alpha = 3$)
+      \item skew $t$-5 with 4 degrees of freedom ($\alpha = 3$)
+      \item Half-Gaussian, Half $t$-1 (spatial range = 0.4)
     \end{enumerate}
-    \item Compare quantile scores using 5-fold cross validation
+    \item Spatial settings
+    \begin{itemize}
+      \item $\bs \in [0, 1] \times [0, 1]$
+      \item Exponential covariance with range: 0.1
+      \item Ratio of spatial to nugget error: 0.9
+    \end{itemize}
+  \end{itemize}
+\end{frame}
+
+\begin{frame}{Simulation study methods}
+  \begin{itemize} \setlength{\itemsep}{0.5em}
+    \item 5 different analysis methods
+    \begin{enumerate}[1.]
+      \item Gaussian
+      \item Skew $t$-1 ($T=0$)
+      \item Skew $t$-1 ($T=0.9$)
+      \item Skew $t$-5 ($T=0$)
+      \item Skew $t$-5 ($T=0.9$)
+    \end{enumerate}
+    \item All methods use a \Matern covariance structure except for method 5 which uses an exponential covariance ($\nu = 0.5$)
+  \end{itemize}
+\end{frame}
+
+\begin{frame}{Simulation results}
+  Still rerunning a few of the MCMC runs because quantile scores for 4. and 5. are
+  unusually high for method 5
+\end{frame}
+
+\begin{frame}{Future work}
+  \begin{itemize} \setlength{\itemsep}{0.5em}
+    \item Comparison with extreme value analysis methods
+    \item Including time in the model
+    \begin{itemize}
+      \item AR(1): $Y_t(\bs) = X_t(\bs) \beta + \phi Y_{t-1}(\bs) + \alpha z_t(\bs) + v_t(\bs)$
+    \end{itemize}
+  \end{itemize}
+\end{frame}
+
+\begin{frame}{Questions}
+  \begin{itemize} \setlength{\itemsep}{0.5em}
+    \item Any questions?
+    \item Thank you for your attention.
   \end{itemize}
@@
 (2012) A survey of spatial extremes: Measuring spatial dependence and modeling spatial effects.
-{\it REVSTAT}, 10, 135--165.
-    \item Zhang, H. and El-Shaarawi, A. (2010) On spatial skew-Gaussian processes and applications. {\it Environmetrics}, 33--47.
+{\it REVSTAT}, {\bf 10}, 135--165.
+    \item Huser, R. and Davison, A. C. (2014) Space-time modelling of extreme events. {\it Journal of the Royal Statistical Society: Series B (Statistical Methodology)}, {\bf 76}, 439--461
+    \item Zhang, H. and El-Shaarawi, A. (2010) On spatial skew-Gaussian processes and applications. {\it Environmetrics}, {\bf 21} 33--47.
   \end{itemize}

### 2014-08-02 pot/LaTeX/jsm2014.tex
presentation updates

@@
     \begin{itemize} \setlength{\itemsep}{0.25em}
-      \item $\alpha \in \calR$ controls the skew
+      \item $\alpha \in \calR$ controls the skewness
       \item $z_t \iid N_{(0, \infty)}(0, \sigma^2_t)$ is a random effect
@@
 
-\begin{frame}{Skew-normal distribution}
+\begin{frame}{Spatial skew-$t$ distribution}
   \begin{itemize} \setlength{\itemsep}{0.5em}
-    \item Conditioned on $z_t$ and $\sigma^2_t$, $Y_t(\bs)$ is Gaussian
-    \begin{itemize}
-      \item After marginalizing over $\sigma^2_t$, this becomes a multivariate $t$-distribution with $2a$ degrees of freedom.
-    \end{itemize}
+    \item \alert{Conditioned} on $z_t$ and $\sigma^2_t$, $Y_t(\bs)$ is Gaussian
     \item Can use standard geostatistical methods to fit this model.
     \item Predictions can be made through kriging.
-    \item Marginalizing over $z_t$ and $\sigma^2_t$ (via MCMC),
+    \item \alert{Marginalizing} over $z_t$ and $\sigma^2_t$ (via MCMC),
     \begin{align*}
@@
     \begin{align*}
-      Y_t(\bs) = X_t(\bs) \beta + \alpha z_t(\bs) + v_t(\bs)
+      Y_t(\bs) = X_t(\bs) \beta + \alpha z_t(\bs) + \sigma(\bs) v_t(\bs)
     \end{align*}
     \item Consider a set of daily knots $\{w_{t1}, \ldots, w_{tK}\}$ that define a daily partition
-    $\{P_{t1}, \ldots, P_{tK}\}$ such that
+    $P_{t1}, \ldots, P_{tK}$ such that
     \begin{align*}
@@
     \end{align*}
-    \item Within each partition $Y_t(\bs)$ has the same distribution as before.
+    \item Within each partition $Y_t(\bs)$ has the same MVT distribution as before.
   \end{itemize}
@@
       \item Impute missing observations and data below $T$
-      \item Update parameters with random walk Metropolis Hastings or Gibbs sampling
+      \item Update parameters with standard random walk Metropolis Hastings or Gibbs sampling
       \item Make spatial predictions
@@
     \begin{itemize}
-      \item Gaussian vs $t$
-      \item 1 partition vs 5 partitions
+      \item Gaussian vs $t$ vs skew-$t$ marginal distribution
+      \item $K=1$ partition vs $K=5$ partitions
       \item No thresholding vs thresholded
       \begin{itemize}
-        \item Thresholded data at 0.90 sample quantile
+        \item Thresholded data at $T=0.90$ sample quantile
       \end{itemize}
@@
     \centering
-    \begin{tabular}{lrrrrrr}
-           & \multicolumn{6}{c}{Quantile}\\
-           \cline{2-7}
-    & 0.900 & 0.950 & 0.980 & 0.990 & 0.995 & 0.999\\
+    \begin{tabular}{l|c|r|rrrrr}
+          \multicolumn{3}{c}{\ } & \multicolumn{5}{c}{Quantile}\\
            \hline
-Gaussian & 16.38 & 15.76 & 15.02 & 14.52 & 14.08 & 13.22\\
-$t$-1 ($T=0$) & 16.15 & 15.51 & 14.62 & 14.00 & 13.43 & 12.32\\
-$t$-5 ($T=0$) & 13.61 & 12.66 & 11.61 & 10.96 & 10.40 & 9.34\\
-$t$-1 ($T=0.9$) & 5.52  & 3.58  & 2.28  & 1.77  & 1.47  & 1.10\\
-$t$-5 ($T=0.9$) & 5.98  & 4.27  & 2.99  & 2.41  & 2.03  & 1.49\\
-skew $t$-1 ($T=0$) & 9.24  & 7.27  & 5.26  & 4.13  & 3.27  & 1.96\\
-skew $t$-1 ($T=0.9$) & {\bf 4.91}  & {\bf 3.16} & {\bf 1.94}  & {\bf 1.45}  & {\bf 1.16}  & {\bf 0.82}\\
-skew $t$-5 ($T=0$) & 15.81 & 14.46 & 12.74 & 11.57 & 10.57 & 8.60\\
-skew $t$-5 ($T=0.9$) & & & & & &\\
+  Marginal & $K$ & $T$  & 0.900 & 0.950 & 0.990 & 0.995 & 0.999\\
+  \hline
+Gaussian & 1 & 0 & 16.38 & 15.76 & 14.52 & 14.08 & 13.22\\
+$t$ & 1 & 0 & 16.15 & 15.51 & 14.00 & 13.43 & 12.32\\
+$t$ & 5 & 0 & 13.61 & 12.66 & 10.96 & 10.40 & 9.34\\
+skew $t$ & 1 & 0 & 9.24  & 7.27 & 4.13  & 3.27  & 1.96\\
+skew $t$ & 5 & 0 & 15.81 & 14.46 & 11.57 & 10.57 & 8.60\\
+$t$ & 1 & 0.9 & 5.52  & 3.58  & 1.77  & 1.47  & 1.10\\
+$t$ & 5 & 0.9 & 5.98  & 4.27  & 2.41  & 2.03  & 1.49\\
+ skew $t$ & 1 & 0.9 & {\bf 4.91}  & {\bf 3.16} & {\bf 1.45}  & {\bf 1.16}  & {\bf 0.82}\\
+skew $t$ & 3 & 0.9 & 5.58 & 3.78 & 1.93& 1.58& 1.11\\
 \hline
@@
   \end{table}
-  \begin{itemize} \setlength{\itemsep}{0.5em}
-    \item Quantile scores -- Still waiting on last set
-  \end{itemize}
 \end{frame}

### 2014-08-04 pot/LaTeX/jsm2014.tex
typo correction

@@
     \begin{align*}
-      Y_t(\bs) \sim \text{skew-t}(\mu, \alpha, \text{df}=2a)
+      Y_t(\bs) \sim \text{skew-t}(\mu, \Sigma^*, \alpha, \text{df}=2a)
     \end{align*}
+    where
+    \begin{itemize}
+    	\item $\mu$ is the location
+	\item $a$, $b$ are the IG parameters for $\sigma^2_t$
+	\item $\Sigma^* = \frac{ b }{ a } \Sigma$ is a scale matrix, and $\Sigma$ is a \Matern covariance matrix
+	\item $\alpha \in \calR$ controls the skewness
+    \end{itemize}
   \end{itemize}
@@
     \end{align*}
-    \item This value shows asymptotic dependence that does not approach 0 as $\bh \rightarrow \infty$.
+    \item This value shows asymptotic dependence that does not approach 0 as $\bh \rightarrow \infty$ (Padoan, 2011)
     \item Deal with this through a daily random partition.
   \end{itemize}
+\end{frame}
+
+\begin{frame}{Simulated $\chi$ plots}
+  \centering
+  \includegraphics[width=1\linewidth]{./plots/chi-plots.pdf}\\[-0.25in]
+  \includegraphics[width=0.2\linewidth]{./plots/chi-legend.pdf}
 \end{frame}
@@
     \item All methods use a \Matern or exponential covariance ($\nu = 0.5$)
-    \item Compare quantile and Brier scores using 5-fold cross validation
+    \item Compare quantile and Brier scores using 5-fold cross validation (Gneiting and Raftery, 2007)
     \item Mean function modeled using a first-order spatial trend
@@
 
-\begin{frame}{Cross-validation results}
+\begin{frame}{Quantile score}
+  \begin{itemize} \setlength{\itemsep}{0.5em}
+    \item The quantile score for the $\tau$th quantile is
+    \begin{align*}
+      2 \{ I[y < \widehat{q}(\tau)] - \tau\} (\widehat{q} - y)
+    \end{align*}
+    where:
+    \begin{itemize}
+      \item $y$ is a test set value
+      \item $\widehat{q}(\tau)$ is the estimated $\tau$th quantile
+    \end{itemize}
+  \end{itemize}
+\end{frame}
+
+\begin{frame}{Brier score}
+  \begin{itemize} \setlength{\itemsep}{0.5em}
+	\item Brier score for predicting exceedance of threshold $c$
+	\begin{align*}
+	  [e(c) - P(c)]^2
+	\end{align*}
+	where 
+	\begin{itemize}
+		\item $y$ is a test set value
+		\item $e(c) = I[y > c]$
+		\item $P(c)$ is the predicted probability of exceeding $c$
+	\end{itemize}
+  \end{itemize}
+\end{frame}
+
+\begin{frame}{Five-fold cross-validation results}
   \begin{table}[htbp]
@@
   \end{table}
+  \begin{itemize}
+  	\item Brier score results are similar.
+  \end{itemize}
 \end{frame}
@@
 
-\begin{frame}{Simulation results}
-  Still rerunning a few of the MCMC runs because quantile scores for 4. and 5. are
-  unusually high for method 5
+\begin{frame}{Simulation study results}
+  \begin{itemize} \setlength{\itemsep}{0.5em}
+    \item Results are similar to the results from the data analysis
+    \item Biggest gains come from thresholding.
+    \item Using skew models give additional gain, but small relative to gain for thresholding.
+  \end{itemize}
 \end{frame}
@@
   \begin{itemize} \setlength{\itemsep}{0.5em}
-    \item Cooley, D., Cisewski, J., Erhardt, R.
-J., Mannshardt, E., Omolo, B.
-O.
-and Sun, Y.
-(2012) A survey of spatial extremes: Measuring spatial dependence and modeling spatial effects.
-{\it REVSTAT}, {\bf 10}, 135--165.
-    \item Huser, R. and Davison, A. C. (2014) Space-time modelling of extreme events. {\it Journal of the Royal Statistical Society: Series B (Statistical Methodology)}, {\bf 76}, 439--461
-    \item Zhang, H. and El-Shaarawi, A. (2010) On spatial skew-Gaussian processes and applications. {\it Environmetrics}, {\bf 21} 33--47.
+    \item Huser, R. and Davison, A. C. (2014) Space-time modelling of extreme events. {\it Journal of the Royal Statistical Society: Series B (Statistical Methodology)}, {\bf 76}, 439--461.
+    \item Padoan, S. A. (2011) Multivariate extreme models based on underlying skew-$t$ and skew-normal distributions. {\it Journal of Multivariate Analysis}, {\bf 102}, 977--991.
+    \item Zhang, H. and El-Shaarawi, A. (2010) On spatial skew-Gaussian processes and applications. {\it Environmetrics}, {\bf 21}, 33--47.
   \end{itemize}

### 2014-08-04 pot/LaTeX/plots/chi-plots.R

@@ -3,8 +3,8 @@ library(fields)
 library(geoR)
 
 # necessary functions
-source('../../R/mcmc.R')
-source('../../R/auxfunctions.R')
+source('../../code/R/mcmc.R')
+source('../../code/R/auxfunctions.R')
 
 # data settings
 beta.t <- c(10, 2, -3)
@@ -75,11 +75,11 @@ diag(dist) <- 0
 bins <- seq(0, 1, length=10)
 bins <- c(bins, 1.5)
 
-y.set <- y[, , 1, 2]
+y.set <- y[, , 1, 1]
 
 probs <- c(0.7, 0.8, 0.9, 0.95, 0.96, 0.97, 0.98, 0.99, 0.995)
 threshs <- quantile(y.set, probs=probs, na.rm=T)
-exceed <- matrix(NA, nrow=(length(bins) - 1), ncol=length(threshs))
+exceed.1 <- matrix(NA, nrow=(length(bins) - 1), ncol=length(threshs))
 for (thresh in 1:length(threshs)){
   exceed.thresh <- att <- acc <- rep(0, (length(bins) - 1))
   for (t in 1:nt) {
@@ -98,7 +98,33 @@ for (thresh in 1:length(threshs)){
       }
     }
   }
-  exceed[, thresh] <- exceed.thresh
+  exceed.1[, thresh] <- exceed.thresh
+}
+
+y.set <- y[, , 1, 2]
+
+probs <- c(0.7, 0.8, 0.9, 0.95, 0.96, 0.97, 0.98, 0.99, 0.995)
+threshs <- quantile(y.set, probs=probs, na.rm=T)
+exceed.2 <- matrix(NA, nrow=(length(bins) - 1), ncol=length(threshs))
+for (thresh in 1:length(threshs)){
+  exceed.thresh <- att <- acc <- rep(0, (length(bins) - 1))
+  for (t in 1:nt) {
+    these <- which(y.set[, t] > threshs[thresh])
+    n.na <- sum(is.na(y.set[, t]))
+    for (b in 1:(length(bins) - 1)) {
+      for (site in these){
+        inbin <- (dist[site, ] > bins[b]) & (dist[site, ] < bins[b+1])
+        att[b] <- att[b] + sum(inbin) - sum(is.na(y.set[inbin, t]))
+        acc[b] <- acc[b] + sum(y.set[inbin, t] > threshs[thresh], na.rm=T)
+      }
[…]

### 2014-08-04 pot/LaTeX/jsm2014.tex
maps and troubleshooting
map updates

@@
 
+\begin{frame}{Example daily partition}
+	Two sample partitions
+    \centering
+    \includegraphics[width=0.54\linewidth]{./plots/example-partition-1.pdf}
+    \includegraphics[width=0.54\linewidth]{./plots/example-partition-2.pdf}
+\end{frame}
+
 \begin{frame}{MCMC details}
@@
   \begin{itemize} \setlength{\itemsep}{0.5em}
-    \item Ozone analysis at 85 sites in NC, SC, and GA for 92 days
+    \item Max 8-hour ozone measurements at 85 sites in NC, SC, and GA for 92 days
     \includegraphics[width=1\linewidth]{./plots/ozone_station.pdf}
   \end{itemize}
+\end{frame}
+
+\begin{frame}{$\chi$-plot for selected ozone sample quantiles}
+	\centering
+    \includegraphics[width=1\linewidth]{./plots/chi-plot-ozone-res.pdf}
 \end{frame}

### 2014-08-04 pot/LaTeX/jsm2014.tex
pull new info from ozone results

@@
 \begin{frame}{Data analysis}
-  \begin{itemize} \setlength{\itemsep}{0.5em}
-    \item Max 8-hour ozone measurements at 85 sites in NC, SC, and GA for 92 days
-    \includegraphics[width=1\linewidth]{./plots/ozone_station.pdf}
-  \end{itemize}
+	\centering
+	Ozone monitoring station locations
+    \includegraphics[width=0.8\linewidth]{./plots/ozone_station.pdf}
+
 \end{frame}
 
-\begin{frame}{$\chi$-plot for selected ozone sample quantiles}
+\begin{frame}{Data analysis}
   \centering
+  Max 8-hour ozone measurements at 85 sites in NC, SC, and GA for days 5 and 34.
+    \includegraphics[width=1\linewidth]{./plots/ozone-day.pdf}
+
+\end{frame}
+
+\begin{frame}{Exploratory data analysis}
+	\centering
+	$\chi$-plot for residuals selected ozone sample quantiles
     \includegraphics[width=1\linewidth]{./plots/chi-plot-ozone-res.pdf}
@@
 
-\begin{frame}{Simulation study settings}
+\begin{frame}{Simulation study}
   \begin{itemize} \setlength{\itemsep}{0.5em}
-    \item Data generated from 6 different settings.
-    \begin{enumerate}[1.]
-      \item Gaussian
-      \item $t$-1 with 4 degrees of freedom
-      \item $t$-5 with 4 degrees of freedom
-      \item skew $t$-1 with 4 degrees of freedom ($\alpha = 3$)
-      \item skew $t$-5 with 4 degrees of freedom ($\alpha = 3$)
-      \item Half-Gaussian, Half $t$-1 (spatial range = 0.4)
-    \end{enumerate}
-    \item Spatial settings
+    \item 6 different data settings:
     \begin{itemize}
-      \item $\bs \in [0, 1] \times [0, 1]$
-      \item Exponential covariance with range: 0.1
-      \item Ratio of spatial to nugget error: 0.9
+    	\item Gaussian vs $t$ vs skew-$t$ marginal distribution
+        \item $K=1$ partition vs $K=5$ partitions
     \end{itemize}
-  \end{itemize}
-\end{frame}
-
-\begin{frame}{Simulation study methods}
-  \begin{itemize} \setlength{\itemsep}{0.5em}
-    \item 5 different analysis methods
-    \begin{enumerate}[1.]
-      \item Gaussian
-      \item Skew $t$-1 ($T=0$)
-      \item Skew $t$-1 ($T=0.9$)
-      \item Skew $t$-5 ($T=0$)
-      \item Skew $t$-5 ($T=0.9$)
-    \end{enumerate}
-    \item All methods use a \Matern covariance structure except for method 5 which uses an exponential covariance ($\nu = 0.5$)
-  \end{itemize}
-\end{frame}
-
-\begin{frame}{Simulation study results}
-  \begin{itemize} \setlength{\itemsep}{0.5em}
     \item Results are similar to the results from the data analysis

### 2014-08-05 pot/LaTeX/jsm2014.tex
mapping, slide updates
Edit to mapping file. Pull out different information. Slide updates

@@
     \end{align*}
+    \item For a spatial analysis, max-stable processes give an appropriate limiting distribution.
   \end{itemize}
@@
 
+\begin{frame}{$\chi$ coefficient}
+  \begin{itemize} \setlength{\itemsep}{0.5em}
+   \item The $\chi$ coefficient is a measure of extremal correlation
+   \item Specifically, we focus on $\chi(c, \bh)$ for the upper tail given by
+    \begin{align*}
+      \chi(c, \bh) = \Pr(Y(\bs) > c \mid Y(\bs + \bh) > c)
+    \end{align*}
+    \item If $\lim_{c \rightarrow \infty} \chi(c) = 0$, then high extreme values are not correlated.
+  \end{itemize}
+\end{frame}
+
+\begin{frame}{Gaussian spatial model}
+  \begin{itemize} \setlength{\itemsep}{0.5em}
+    \item In geostatistical models, observations, $Y(\bs)$ are often modeled using a Gaussian spatial model with a likelihood of
+      \begin{align*}
+    	\hspace{-2em}
+    	f(y(\bs)) = \frac{ 1 }{ (2 \pi)^{n/2} } | \Sigma |^{-1/2} \exp \left[ -\frac{ 1 }{ 2 } (Y(\bs) - \mu(\bs) )^T \Sigma^{-1} (Y(\bs) - \mu(\bs)) \right]
+    \end{align*}
+    where
+    \begin{itemize}
+    	\item $\mu(\bs)$ is a mean vector
+	\item $\Sigma$ is a covariance matrix
+    \end{itemize}
+    \item Model properties:
+    \begin{itemize}
+      \item Nice computing properties
+    \item For a Gaussian spatial model $\lim_{c \rightarrow \infty} \chi(c) = 0$ regardless of the strength of the correlation in the bulk of the distribution.
+    \item Tail is not flexible
+    \end{itemize}
+    \end{itemize}
+\end{frame}
+
 \begin{frame}{Spatial skew-$t$ distribution}
@@
   \begin{itemize} \setlength{\itemsep}{0.5em}
-    \item \alert{Conditioned} on $z_t$ and $\sigma^2_t$, $Y_t(\bs)$ is Gaussian
+   \item \alert{Conditioned} on $z_t$ and $\sigma^2_t$, $Y_t(\bs)$ is a Gaussian spatial model
     \item Can use standard geostatistical methods to fit this model.
@@
     \end{itemize}
+
   \end{itemize}
@@
 
-\begin{frame}{Long-range dependence}
+\begin{frame}{Spatial skew-$t$ distribution}
   \begin{itemize} \setlength{\itemsep}{0.5em}
-    \item The $\chi$ coefficient is a measure of extremal spatial correlation
-    \begin{align*}
-      \chi(\bh) = \Pr(Y_t(\bs) > c \mid Y_t(\bs + \bh) > c)
-    \end{align*}
-    \item This value shows asymptotic dependence that does not approach 0 as $\bh \rightarrow \infty$ (Padoan, 2011)
+      \item Model properties
+    \begin{itemize}
+    	\item Has flexible tail controlled by $\alpha$ and degrees of freedom
+    	\item For a skew-$t$ distribution $\lim_{c \rightarrow \infty} \chi(c) > 0$ Padoan, 2011)
+    	\item Computation that is on the order of Gaussian computation
+    \end{itemize}
+    \item For this distribution, $\chi(c, \bh)$ shows asymptotic dependence that does not approach 0 as $\bh \rightarrow \infty$ 
     \item Deal with this through a daily random partition.
@@
 
-\begin{frame}{Simulated $\chi$ plots}
-  \centering
-  \includegraphics[width=1\linewidth]{./plots/chi-plots.pdf}\\[-0.25in]
-  \includegraphics[width=0.2\linewidth]{./plots/chi-legend.pdf}
-\end{frame}
-
-\begin{frame}{Random daily partition}
+\begin{frame}{Daily random partition}
   \begin{itemize} \setlength{\itemsep}{0.5em}
@@
   \end{itemize}
+\end{frame}
+
+\begin{frame}{Simulated $\chi$ plots}
+  \centering
+  \includegraphics[width=1\linewidth]{./plots/chi-plots.pdf}\\[-0.25in]
+  \includegraphics[width=0.2\linewidth]{./plots/chi-legend.pdf}
 \end{frame}
@@
     \item Comparison with extreme value analysis methods
+    \item Reporting of compliance with EPA ozone standards
+    \item Spatial mapping of high quantiles
     \item Including time in the model
@@
   \begin{itemize} \setlength{\itemsep}{0.5em}
+    \item Demarta, S. and McNeil, A. J. (2007) The $t$ copula and related copulas. {\it International Statistical Review}, {\bf 73}, 111--129.
     \item Huser, R. and Davison, A. C. (2014) Space-time modelling of extreme events. {\it Journal of the Royal Statistical Society: Series B (Statistical Methodology)}, {\bf 76}, 439--461.

### 2014-08-05 pot/code/analysis/simstudy/setup.R
updates
generate sample datasets and plot them

@@ -86,6 +86,51 @@ for (setting in 1:nsettings) {
   knots.t[[setting]] <- knots.t.setting
 }
 
+# plot the data
+s1 <- s2 <- seq(0, 1, length=50)
+s <- expand.grid(s1, s2)
+X <- array(1, dim=c(nrow(s), 2, 3))
+X.temp <- cbind(rep(1, nrow(s)), s[, 1], s[, 2])
+X[, 1, ] <- X.temp
+X[, 2, ] <- X.temp
+y.gau <- rpotspat(nt=2, x=X, s=s, beta=c(10, 0, 0), alpha=alpha.t, nu=nu.t,
+                  gau.rho=0.1, t.rho=0.1, mixprob=0, z.alpha=0, tau.alpha=tau.alpha.t,
+                  tau.beta=tau.beta.t, nknots=1)
+par(mfrow=c(1, 2))
+quilt.plot(s[, 1], s[, 2], z=y.gau$y[, 1], nx=50, ny=50, main="Gaussian")
+hist(y.gau$y[, 1], main="Histogram", xlab="")
+
+y.t1 <- rpotspat(nt=2, x=X, s=s, beta=c(10, 0, 0), alpha=alpha.t, nu=nu.t,
+                  gau.rho=0.1, t.rho=0.1, mixprob=1, z.alpha=0, tau.alpha=tau.alpha.t,
+                  tau.beta=tau.beta.t, nknots=1)
+par(mfrow=c(1, 2))
+quilt.plot(s[, 1], s[, 2], z=y.t1$y[, 1], nx=50, ny=50, main="t, K=1")
+hist(y.t1$y[, 1], main="Histogram", xlab="")
+
+y.t3 <- rpotspat(nt=2, x=X, s=s, beta=c(10, 0, 0), alpha=alpha.t, nu=nu.t,
+                  gau.rho=0.1, t.rho=0.1, mixprob=1, z.alpha=0, tau.alpha=tau.alpha.t,
+                  tau.beta=tau.beta.t, nknots=3)
+par(mfrow=c(1, 2))
+quilt.plot(s[, 1], s[, 2], z=y.t3$y[, 1], nx=50, ny=50, main="t, K=3")
+hist(y.t3$y[, 1], main="Histogram", xlab="")
+
+y.st1 <- rpotspat(nt=2, x=X, s=s, beta=c(10, 0, 0), alpha=alpha.t, nu=nu.t,
[…]

### 2014-08-05 pot/LaTeX/jsm2014.tex
mapping2
updates
generate sample datasets and plot them
bugfix predictions
Found a fairly major bug in the mcmc

@@
 \setbeamerfont{title}{series=\bfseries, size=\huge}
-\setbeamerfont{author}{series=\bfseries, size=\Large}
+\setbeamerfont{author}{series=\bfseries, size=\large}
 \setbeamerfont{institute}{series=\mdseries, size=\large}
@@
 }
-\author[S. Morris and B. Reich]{Samuel Morris and Brian Reich}
-\institute[NCSU]{North Carolina State University}
-\date{}
+\author[S. Morris, B. Reich, and D. Cooley]{Samuel Morris (NC State), Brian Reich (NC State), and Dan Cooley (Colorado State)}
+\institute[NCSU, CSU]{}
+\date{JSM 2014, Boston}
 
@@
     \end{itemize}
-    \item In geostatistical analysis, kriging uses spatial correlation to help inform prediction at unknown locations.
-    \item Want to explore computationally easy methods that are available in higher dimensions
+    \item Spatial extreme borrows strength across space to estimate return levels and make predictions at unknown locations..
+    \item In geostatistical analysis, kriging uses spatial correlation for prediction.
+    \item Want to explore similar methods for extremes.
   \end{itemize}
@@
 
-\begin{frame}{Standard non-spatial analysis}
+\begin{frame}{Standard analysis - Block maxima}
   \begin{itemize} \setlength{\itemsep}{0.5em}
-    \item Block maxima:
-    \begin{itemize}
       \item Uses yearly maxima
@@
       \item Models are fit using the generalized extreme value distribution
-    \end{itemize}
-    \item Generalized extreme value distribution (GEV):
-    \begin{align*}
-      \Pr(Y_j < y) = G_j(y) = \exp \left\{ -\left[ \left(1 + \xi_j \frac{y - \mu_j}{\sigma_j}\right)_+^{-1/\xi_j} \right] \right\}
-    \end{align*}
-    \item For a spatial analysis, max-stable processes give an appropriate limiting distribution.
+    \item For a spatial analysis, max-stable processes give an appropriate limiting distribution
   \end{itemize}
@@
 
-\begin{frame}{Standard non-spatial analysis}
-  \begin{itemize}
-    \item Peaks-over-threshold:
+\begin{frame}{Standard analysis - Peaks over threshold}
   \begin{itemize}
@@
       \item Select a threshold, $T$, and fit data above the threshold using the generalized Pareto distribution
-      \item Autocorrelation may be an issue between observations (e.g. flood levels don't dissipate overnight)
-    \end{itemize}
-    \item Generalized Pareto distribution (GPD):
-    \begin{align*}
-      \Pr(Y_j > y | Y_j > T) = F_j(y) = \left( 1 + \xi_j \frac{y - T}{\sigma_j} \right)_+^{-1/\xi_j}
-    \end{align*}
+      \item Temporal dependence may be an issue between observations (e.g. flood levels don't dissipate overnight)
+    \item Generalized Pareto distribution (GPD) is used for the exceedances.
   \end{itemize}
@@
     \begin{itemize}
-      \item computationally hard to work with
+      \item computationally challenging to work with
       \item joint distribution only available in low dimension
@@
     \begin{itemize}
-      \item has a flexible tail
+      \item has marginal distribution with a flexible tail
       \item has asymptotic spatial dependence
-      \item computation on the order of Gaussian models for large space-time datasets
+      \item has computation on the order of Gaussian models for large space-time datasets
     \end{itemize}
@@
   \begin{itemize} \setlength{\itemsep}{0.5em}
-   \item The $\chi$ coefficient is a measure of extremal correlation
-   \item Specifically, we focus on $\chi(c, \bh)$ for the upper tail given by
+   \item The $\chi$ coefficient is a measure of extremal dependence
+   \item Specifically, we focus on $\chi(\bh)$ for the upper tail given by
     \begin{align*}
-      \chi(c, \bh) = \Pr(Y(\bs) > c \mid Y(\bs + \bh) > c)
+      \chi(\bh) = \lim_{c \rightarrow \infty}\Pr(Y(\bs) > c \mid Y(\bs + \bh) > c)
     \end{align*}
-    \item If $\lim_{c \rightarrow \infty} \chi(c) = 0$, then high extreme values are not correlated.
+    \item If $ \chi(\bh) = 0$, then observations are asymptotically independent at distance $\bh$.
+    \item We expect $\lim_{\bh \rightarrow \infty}\chi(h)=0$.
   \end{itemize}
@@
   \begin{itemize} \setlength{\itemsep}{0.5em}
-    \item In geostatistical models, observations, $Y(\bs)$ are often modeled using a Gaussian spatial model with a likelihood of
-      \begin{align*}
-    	\hspace{-2em}
-    	f(y(\bs)) = \frac{ 1 }{ (2 \pi)^{n/2} } | \Sigma |^{-1/2} \exp \left[ -\frac{ 1 }{ 2 } (Y(\bs) - \mu(\bs) )^T \Sigma^{-1} (Y(\bs) - \mu(\bs)) \right]
-    \end{align*}
-    where
-    \begin{itemize}
-    	\item $\mu(\bs)$ is a mean vector
-	\item $\Sigma$ is a covariance matrix
-    \end{itemize}
+    \item In geostatistics $Y(\bs)$ are often modeled using a Gaussian process with mean function $\mu(\bs)$ and covariance function
+$\rho(\bh)$.
     \item Model properties:
     \begin{itemize}
-      \item Nice computing properties
+      \item Nice computing properties (closed-form likelihood)
     \item For a Gaussian spatial model $\lim_{c \rightarrow \infty} \chi(c) = 0$ regardless of the strength of the correlation in the bulk of the distribution.
-    \item Tail is not flexible
+    \item Tail is not flexible (Gaussian is light tailed)
     \end{itemize}
@@
     \item Can use standard geostatistical methods to fit this model.
-    \item Predictions can be made through kriging.
+    \item Predictions can be made through Kriging.
     \item \alert{Marginalizing} over $z_t$ and $\sigma^2_t$ (via MCMC),
@@
     \end{itemize}
-
   \end{itemize}
@@
     \begin{itemize}
-    	\item Has flexible tail controlled by $\alpha$ and degrees of freedom
+    	\item Has flexible tail controlled by skewness $\alpha$ and degrees of freedom $2a$
     	\item For a skew-$t$ distribution $\lim_{c \rightarrow \infty} \chi(c) > 0$ Padoan, 2011)
@@
     \end{itemize}
-    \item For this distribution, $\chi(c, \bh)$ shows asymptotic dependence that does not approach 0 as $\bh \rightarrow \infty$ 
-    \item Deal with this through a daily random partition.
+    \item For this distribution, $\chi(\bh)$ shows asymptotic dependence that does not approach 0 as $\bh \rightarrow \infty$ 
+   \item This occurs because all observations (near and far) share the same $z_t$ and $\sigma_t^2$.
+    \item We deal with this through a daily random partition (similar to Huser and Davison).
   \end{itemize}
@@
     \end{align*}
-    \item Consider a set of daily knots $\{w_{t1}, \ldots, w_{tK}\}$ that define a daily partition
+    \item Consider a set of daily knots $w_{tk} \sim$ Uniform that define a random daily partition
     $P_{t1}, \ldots, P_{tK}$ such that
@@
     \end{align*}
-    \item Within each partition $Y_t(\bs)$ has the same MVT distribution as before.
+    \item Within each partition $Y_t(\bs)$ has the same MV skew-t distribution as before.
   \end{itemize}
+\end{frame}
+
+\begin{frame}{Example daily partition}
+	Two sample partitions (number is at partition center)
+    \centering
+    \includegraphics[width=0.54\linewidth]{./plots/example-partition-1.pdf}
+    \includegraphics[width=0.54\linewidth]{./plots/example-partition-2.pdf}
 \end{frame}
@@
 
-\begin{frame}{Example daily partition}
-	Two sample partitions
+\begin{frame}{Sample simulated datasets}
   \centering
-    \includegraphics[width=0.54\linewidth]{./plots/example-partition-1.pdf}
-    \includegraphics[width=0.54\linewidth]{./plots/example-partition-2.pdf}
+  \includegraphics[width=1\linewidth]{./plots/gauss-vs-skew-t3.pdf}
+\end{frame}
+
+\begin{frame}{Spatiotemporal Model}
+  \begin{itemize} \setlength{\itemsep}{0.5em}
+    \item We can account for time in one of two ways
+    \begin{itemize}
+    	\item The mean: e.g. AR(1)
+	\item Three dimensional covariance model (e.g. Huser and Davison, 2014) 
+    \end{itemize}
+  \end{itemize}
 \end{frame}
@@
     \begin{enumerate}[1.]
-      \item Impute missing observations and data below $T$
+      \item Impute missing observations and censored data below $T$
       \item Update parameters with standard random walk Metropolis Hastings or Gibbs sampling
@@
       \item $K=1$ partition vs $K=5$ partitions
-      \item No thresholding vs thresholded
-      \begin{itemize}
-        \item Thresholded data at $T=0.90$ sample quantile
-      \end{itemize}
+      \item No thresholding vs thresholding at $T=0.90$ sample quantile
     \end{itemize}
@@
 
-\begin{frame}{Quantile score}
+\begin{frame}{Quantile score for cross-validation}
   \begin{itemize} \setlength{\itemsep}{0.5em}
@@
   \begin{itemize} \setlength{\itemsep}{0.5em}
-	\item Brier score for predicting exceedance of threshold $c$
+	\item The Brier score for predicting exceedance of threshold $c$ is
 	\begin{align*}
@@
     \item Reporting of compliance with EPA ozone standards
-    \item Spatial mapping of high quantiles
-    \item Including time in the model
-    \begin{itemize}
-      \item AR(1): $Y_t(\bs) = X_t(\bs) \beta + \phi Y_{t-1}(\bs) + \alpha z_t(\bs) + v_t(\bs)$
-    \end{itemize}
+    \item Including time in the model via standard spatiotemporal Gaussian models.  
   \end{itemize}
@@
   \begin{itemize} \setlength{\itemsep}{0.5em}
-    \item Any questions?
+    \item Questions?
     \item Thank you for your attention.
+    \item Acknowledgment: This work was funded by EPA STAR award R835228.
   \end{itemize}

### 2014-08-06 pot/LaTeX/jsm2014.tex
typo bounds
forgot to scale covariates

@@
 \begin{frame}{Example daily partition}
-	Two sample partitions (number is at partition center)
     \centering
+    \begin{figure}
     \includegraphics[width=0.54\linewidth]{./plots/example-partition-1.pdf}
     \includegraphics[width=0.54\linewidth]{./plots/example-partition-2.pdf}
+    \caption{Two sample partitions (number is at partition center)}
+    \end{figure}
 \end{frame}
@@
   \centering
+  \begin{figure}
   \includegraphics[width=1\linewidth]{./plots/chi-plots.pdf}\\[-0.25in]
   \includegraphics[width=0.2\linewidth]{./plots/chi-legend.pdf}
+  \caption{$\chi$ plots from simulated datasets}
+  \end{figure}
 \end{frame}
@@
   \centering
+  \begin{figure}
   \includegraphics[width=1\linewidth]{./plots/gauss-vs-skew-t3.pdf}
+  \caption{Gaussian and skew-$t$ with 3 partitions}
+  \end{figure}
 \end{frame}
@@
 	\centering
-	Ozone monitoring station locations
+	\begin{figure}
     \includegraphics[width=0.8\linewidth]{./plots/ozone_station.pdf}
+    \caption{Ozone monitoring station locations}
+    \end{figure}
 
@@
   \centering
-  Max 8-hour ozone measurements at 85 sites in NC, SC, and GA for days 5 and 34.
+  \begin{figure}
     \includegraphics[width=1\linewidth]{./plots/ozone-day.pdf}
+    \caption{Max 8-hour ozone measurements at 85 sites in NC, SC, and GA for days 5 and 34.}
+   \end{figure}
 
@@
 	\centering
-	$\chi$-plot for residuals selected ozone sample quantiles
+  \begin{figure}
     \includegraphics[width=1\linewidth]{./plots/chi-plot-ozone-res.pdf}
+    \caption{$\chi$-plot for residuals of ozone sample quantiles}
+  \end{figure}
 \end{frame}
@@
       \item Gaussian vs $t$ vs skew-$t$ marginal distribution
-      \item $K=1$ partition vs $K=5$ partitions
+      \item $K=1$ partition vs $K=3$ partitions
       \item No thresholding vs thresholding at $T=0.90$ sample quantile
@@
   \hline
-Gaussian & 1 & 0 & 16.38 & 15.76 & 14.52 & 14.08 & 13.22\\
-$t$ & 1 & 0 & 16.15 & 15.51 & 14.00 & 13.43 & 12.32\\
-$t$ & 5 & 0 & 13.61 & 12.66 & 10.96 & 10.40 & 9.34\\
-skew $t$ & 1 & 0 & 9.24  & 7.27 & 4.13  & 3.27  & 1.96\\
-skew $t$ & 5 & 0 & 15.81 & 14.46 & 11.57 & 10.57 & 8.60\\
-$t$ & 1 & 0.9 & 5.52  & 3.58  & 1.77  & 1.47  & 1.10\\
-$t$ & 5 & 0.9 & 5.98  & 4.27  & 2.41  & 2.03  & 1.49\\
- skew $t$ & 1 & 0.9 & {\bf 4.91}  & {\bf 3.16} & {\bf 1.45}  & {\bf 1.16}  & {\bf 0.82}\\
-skew $t$ & 3 & 0.9 & 5.58 & 3.78 & 1.93& 1.58& 1.11\\
+Gaussian & 1 & 0 & 3.807 & 2.981 & 1.797 & 1.493 & 1.044\\
+$t$ & 1 & 0 & 2.622 & 1.584 & 0.436 & 0.244 & 0.063 \\
+$t$ & 3 & 0 & 2.643 & 1.601 & 0.455 & 0.265 & 0.085 \\
+skew-$t$ & 1 & 0 & 2.766 & 1.683 & 0.467 & 0.257 & 0.063 \\
+$t$ & 1 & 0.9 & 4.514 & 2.422 & 0.577 & 0.318 & 0.086 \\
+skew-$t$ & 1 & 0.9 & 4.396 & 2.388 & 0.573 & 0.311 & 0.079 \\
 \hline
     \end{tabular}
+    \caption{Selected five-fold cross-validation results}
   \end{table}
   \begin{itemize}
+  	\item Results for 
   	\item Brier score results are similar.

### 2014-08-06 pot/code/analysis/ozone/SE/OzoneFull-se5.R

@@ -25,11 +25,11 @@ s.scale.preds[,2] <- (s.preds[,2] - range(s[,2])[1])/(range(s[,2])[2] - range(s[
 X <- array(1, dim=c(ns, nt, 6))
 X.preds <- array(1, dim=c(nrow(s.preds), nt, 6))
 for (t in 1:nt) {
-  X[, t, 2] <- s[, 1]
-  X[, t, 3] <- s[, 2]
-  X[, t, 4] <- s[, 1]^2
-  X[, t, 5] <- s[, 2]^2
-  X[, t, 6] <- s[, 1] * s[, 2]
+  X[, t, 2] <- s.scale[, 1]
+  X[, t, 3] <- s.scale[, 2]
+  X[, t, 4] <- s.scale[, 1]^2
+  X[, t, 5] <- s.scale[, 2]^2
+  X[, t, 6] <- s.scale[, 1] * s.scale[, 2]
   
   X.preds[, t, 2] <- s.scale.preds[, 1]
   X.preds[, t, 3] <- s.scale.preds[, 2]

### 2014-08-07 pot/LaTeX/jsm2014.tex
small changes and tweak to mapping file
jsm updates

@@
       \item e.g. When constructing river levees, engineers need to be able to estimate a 100-year or 1000-year flood levels.
+      \item e.g. Probability of exceeding a certain threshold level
     \end{itemize}
-    \item Spatial extreme borrows strength across space to estimate return levels and make predictions at unknown locations..
-    \item In geostatistical analysis, kriging uses spatial correlation for prediction.
+    \item Spatial methods borrow information across space to estimate spatial correlation and make predictions by Kriging at unknown locations.
     \item Want to explore similar methods for extremes.
@@
 \begin{frame}{Standard analysis - Peaks over threshold}
-  \begin{itemize}
+  \begin{itemize}  \setlength{\itemsep}{0.5em}
       \item Incorporates more data than block maxima
-      \item Select a threshold, $T$, and fit data above the threshold using the generalized Pareto distribution
+      \item Select a threshold, $T$, and use the Generalized Pareto distribution (GPD) to model the exceedances.
       \item Temporal dependence may be an issue between observations (e.g. flood levels don't dissipate overnight)
-    \item Generalized Pareto distribution (GPD) is used for the exceedances.
   \end{itemize}
@@
     \item If $ \chi(\bh) = 0$, then observations are asymptotically independent at distance $\bh$.
-    \item We expect $\lim_{\bh \rightarrow \infty}\chi(h)=0$.
+    \item We expect $\lim_{\bh \rightarrow \infty}\chi(\bh)=0$.
   \end{itemize}
@@
       \item Nice computing properties (closed-form likelihood)
-    \item For a Gaussian spatial model $\lim_{c \rightarrow \infty} \chi(c) = 0$ regardless of the strength of the correlation in the bulk of the distribution.
+    \item For a Gaussian spatial model $\lim_{c \rightarrow \infty} \chi(\bh) = 0$ regardless of the strength of the correlation in the bulk of the distribution.
     \item Tail is not flexible (Gaussian is light tailed)
@@
     	\item Has flexible tail controlled by skewness $\alpha$ and degrees of freedom $2a$
-    	\item For a skew-$t$ distribution $\lim_{c \rightarrow \infty} \chi(c) > 0$ Padoan, 2011)
+    	\item For a skew-$t$ distribution $\lim_{c \rightarrow \infty} \chi(\bh) > 0$ (Padoan, 2011)
     	\item Computation that is on the order of Gaussian computation
@@
     \end{align*}
-    \item Consider a set of daily knots $w_{tk} \sim$ Uniform that define a random daily partition
+    \item Consider a set of daily knots $\bw_{tk} \sim$ Uniform that define a random daily partition
     $P_{t1}, \ldots, P_{tK}$ such that
     \begin{align*}
-      P_{tk} = \{s : k = \argmin_\ell|| \bs - w_{t\ell}|| \}
+      P_{tk} = \{s : k = \argmin_\ell|| \bs - \bw_{t\ell}|| \}
     \end{align*}
@@
 
-\begin{frame}{Simulated $\chi$ plots}
+\begin{frame}{Simulated $\widehat{\chi}(\bh)$ plots}
   \centering
@@
   \begin{figure}
+  \includegraphics[width=1\linewidth]{./plots/gauss-vs-t3.pdf}
+  \caption{Gaussian and $t$ with 3 partitions}
+  \end{figure}
+\end{frame}
+
+\begin{frame}{Sample simulated datasets}
+  \centering
+  \begin{figure}
   \includegraphics[width=1\linewidth]{./plots/gauss-vs-skew-t3.pdf}
@@
   \end{figure}
-\end{frame}
-
-\begin{frame}{Spatiotemporal Model}
-  \begin{itemize} \setlength{\itemsep}{0.5em}
-    \item We can account for time in one of two ways
-    \begin{itemize}
-    	\item The mean: e.g. AR(1)
-	\item Three dimensional covariance model (e.g. Huser and Davison, 2014) 
-    \end{itemize}
-  \end{itemize}
 \end{frame}
@@
     \begin{enumerate}[1.]
-      \item Impute missing observations and censored data below $T$
+      \item Impute censored data below $T$
       \item Update parameters with standard random walk Metropolis Hastings or Gibbs sampling
@@
 \begin{frame}{Data analysis}
-	\centering
+\begin{columns}[c]
+\column{.45 \linewidth}
+	\begin{itemize} \setlength{\itemsep}{0.5em}
+	\item Data analysis uses
+	\begin{itemize}
+		\item max 8-hour ozone measurements
+		\item 85 sites
+		\item 92 days
+	\end{itemize}
+	\end{itemize}
+
+	\column{.55\linewidth}
 	\begin{figure}
-    \includegraphics[width=0.8\linewidth]{./plots/ozone_station.pdf}
+    \includegraphics[width=1\linewidth]{./plots/ozone_station.pdf}
     \caption{Ozone monitoring station locations}
     \end{figure}
-
+\end{columns}
 \end{frame}
@@
     \includegraphics[width=1\linewidth]{./plots/chi-plot-ozone-res.pdf}
-    \caption{$\chi$-plot for residuals of ozone sample quantiles}
+    \caption{$\widehat{\chi}$-plot for sample quantiles of ozone observations}
   \end{figure}
@@
     \item Compare quantile and Brier scores using 5-fold cross validation (Gneiting and Raftery, 2007)
-    \item Mean function modeled using a first-order spatial trend
+    \item Mean function modeled as
+    \begin{align*}
+    	\beta_0 + \beta_1 \cdot \text{lat} + \beta_2 \cdot \text{long} +  \beta_3 \cdot \text{lat}^2 + \beta_4 \cdot \text{long}^2 + \beta_5 \cdot \text{lat} \cdot \text{long} 
+    \end{align*}
    \end{itemize}
@@
     \begin{tabular}{l|c|r|rrrrr}
-          \multicolumn{3}{c}{\ } & \multicolumn{5}{c}{Quantile}\\
+          \multicolumn{3}{c}{\ } & \multicolumn{5}{c}{$\tau$}\\
            \hline
-  Marginal & $K$ & $T$  & 0.900 & 0.950 & 0.990 & 0.995 & 0.999\\
+  Marginal & $K$ & $T$  & 0.950 & 0.980 & 0.990 & 0.995 & 0.999\\
   \hline
-Gaussian & 1 & 0 & 3.807 & 2.981 & 1.797 & 1.493 & 1.044\\
-$t$ & 1 & 0 & 2.622 & 1.584 & 0.436 & 0.244 & 0.063 \\
-$t$ & 3 & 0 & 2.643 & 1.601 & 0.455 & 0.265 & 0.085 \\
-skew-$t$ & 1 & 0 & 2.766 & 1.683 & 0.467 & 0.257 & 0.063 \\
-$t$ & 1 & 0.9 & 4.514 & 2.422 & 0.577 & 0.318 & 0.086 \\
-skew-$t$ & 1 & 0.9 & 4.396 & 2.388 & 0.573 & 0.311 & 0.079 \\
+Gaussian 	& 1 & 0 	& 39.820 & 17.539 & 9.167 & 4.720 & 1.057\\
+$t$		& 1 & 0 	& {\bf 31.008} & {\bf 13.898} & 7.229 & {\bf 3.405} & 0.879\\
+$t$		& 3 & 0	& 31.213 & 13.920 & {\bf 7.218} & 3.498 & 0.918\\
+$t$		& 1 & 0.9	& 32.221 & 14.519 & 7.549 & 3.604 & 0.896\\
+$t$		& 3 & 0.9 	& 38.842 & 16.781 & 8.434 & 4.180 & 1.020\\
+skew-$t$	& 1 & 0	& 31.845 & 14.542 & 7.533 & 3.645 & {\bf 0.844}\\
+skew-$t$	& 1 & 0.9	& 32.132 & 14.296 & 7.484 & 3.497 & 0.890\\
+skew-$t$	& 3 & 0	& 33.653 & 15.453 & 8.119 & 4.338 & 1.188\\
+skew-$t$	& 3 & 0.9 	& 32.157 & 14.727 & 7.794 & 3.825 & 0.917\\
 \hline
     \end{tabular}
-    \caption{Selected five-fold cross-validation results}
-  \end{table}
+    \caption{Brier score for predicting exceedance of $c = \hat{q}(\tau)$ from five-fold cross-validation ($\times 1000$)}
+  \end{table} \vspace{-1.5em}
   \begin{itemize}
-  	\item Results for 
-  	\item Brier score results are similar.
+  	\item Quantile score results are similar.
   \end{itemize}
+\end{frame}
+
+\begin{frame}{Predicted 95th quantile}
+\centering
+\begin{figure}
+    \includegraphics[width=.5\linewidth]{./plots/quantile-95-gau.pdf}
+    \includegraphics[width=.5\linewidth]{./plots/quantile-95-t19.pdf}
+    \caption{Predicted 95th quantile using Gaussian and $t$ thresholded at $T=0.9$}
+\end{figure}
+\end{frame}
+
+\begin{frame}{Predicted 99th quantile}
+\centering
+\begin{figure}
+    \includegraphics[width=.5\linewidth]{./plots/quantile-99-gau.pdf}
+    \includegraphics[width=.5\linewidth]{./plots/quantile-99-t19.pdf}
+    \caption{Predicted 95th quantile using Gaussian and $t$ thresholded at $T=0.9$}
+\end{figure}
+\end{frame}
+
+\begin{frame}{Probability of exceedance}
+\centering
+\begin{figure}
+    \includegraphics[width=.5\linewidth]{./plots/p-exceed-std-gau.pdf}
+    \includegraphics[width=.5\linewidth]{./plots/p-exceed-std-t19.pdf}
+    \caption{Probability of exceeding the 75 ppb ozone standard using Gaussian and $t$ thresholded at $T=0.9$}
+\end{figure}
 \end{frame}
@@
     \end{itemize}
-    \item Results are similar to the results from the data analysis
-    \item Biggest gains come from thresholding.
-    \item Using skew models give additional gain, but small relative to gain for thresholding.
+    \item Preliminary results suggest that there may be a modest improvement in quantile scores from thresholding when the data come from a $t$ distribution with $K=5$ partitions.
   \end{itemize}
@@
 
-\begin{frame}{Future work}
+\begin{frame}{Future Work}
   \begin{itemize} \setlength{\itemsep}{0.5em}
+    \item Different ways to incorporate the temporal dependence
+    \begin{itemize}
+    	\item Three dimensional covariance model for $v_t(\bs)$ (e.g. Huser and Davison, 2014)
+    	\item Use a temporal structure for $z_t(\bs)$:
+	\begin{itemize}
+		\item AR(1)
+		\item Moving average 
+		\item Association between $\bw_{t, k}$ and $\bw_{t+1, k}$
+	\end{itemize}
+    \end{itemize}
      \item Comparison with extreme value analysis methods
-    \item Reporting of compliance with EPA ozone standards
-    \item Including time in the model via standard spatiotemporal Gaussian models.  
   \end{itemize}

### 2014-08-07 pot/code/analysis/simstudy/results.R

@@ -24,7 +24,7 @@ rm(list=ls())
 load("simdata.RData")
 ns <- dim(y)[1]
 nt <- dim(y)[2]
-nsets <- 5
+nsets <- 2
 ngroups <- 10
 nsettings <- dim(y)[4]
 nmethods <- 5

### 2014-08-07 pot/LaTeX/jsm2014.tex
Draft updates for JSM

@@
     \includegraphics[width=.5\linewidth]{./plots/quantile-95-gau.pdf}
+    \includegraphics[width=.5\linewidth]{./plots/quantile-95-t10.pdf}
+    \caption{Predicted 95th quantile using Gaussian and $t$}
+\end{figure}
+\end{frame}
+
+\begin{frame}{Predicted 95th quantile}
+\centering
+\begin{figure}
+    \includegraphics[width=.5\linewidth]{./plots/quantile-95-t10.pdf}
     \includegraphics[width=.5\linewidth]{./plots/quantile-95-t19.pdf}
-    \caption{Predicted 95th quantile using Gaussian and $t$ thresholded at $T=0.9$}
+    \caption{Predicted 95th quantile using $t$ and $t$ thresholded at $T=0.9$}
 \end{figure}
@@
     \includegraphics[width=.5\linewidth]{./plots/quantile-99-gau.pdf}
+    \includegraphics[width=.5\linewidth]{./plots/quantile-99-t10.pdf}
+    \caption{Predicted 99th quantile using Gaussian and $t$}
+\end{figure}
+\end{frame}
+
+\begin{frame}{Predicted 99th quantile}
+\centering
+\begin{figure}
+    \includegraphics[width=.5\linewidth]{./plots/quantile-99-t10.pdf}
     \includegraphics[width=.5\linewidth]{./plots/quantile-99-t19.pdf}
-    \caption{Predicted 95th quantile using Gaussian and $t$ thresholded at $T=0.9$}
+    \caption{Predicted 99th quantile using $t$ and $t$ thresholded at $T=0.9$}
 \end{figure}
@@
     \includegraphics[width=.5\linewidth]{./plots/p-exceed-std-gau.pdf}
+    \includegraphics[width=.5\linewidth]{./plots/p-exceed-std-t10.pdf}
+    \caption{Probability of exceeding the 75 ppb ozone standard using Gaussian and $t$}
+\end{figure}
+\end{frame}
+
+\begin{frame}{Probability of exceedance}
+\centering
+\begin{figure}
+    \includegraphics[width=.5\linewidth]{./plots/p-exceed-std-t10.pdf}
     \includegraphics[width=.5\linewidth]{./plots/p-exceed-std-t19.pdf}
-    \caption{Probability of exceeding the 75 ppb ozone standard using Gaussian and $t$ thresholded at $T=0.9$}
+    \caption{Probability of exceeding the 75 ppb ozone standard using $t$ and $t$ thresholded at $T=0.9$}
 \end{figure}
@@
     \end{itemize}
-    \item Preliminary results suggest that there may be a modest improvement in quantile scores from thresholding when the data come from a $t$ distribution with $K=5$ partitions.
+    \item Preliminary results are inconclusive.
   \end{itemize}
@@
     \item Thank you for your attention.
-    \item Acknowledgment: This work was funded by EPA STAR award R835228.
+    \item Acknowledgment: This work was funded by EPA STAR award R835228
   \end{itemize}

### 2014-08-15 pot/LaTeX/skewnormal.tex
manuscript updates

@@
 \section{Introduction}\label{s:intro}
-In most climatological applications, researchers are interested in learning about the average behavior of different climate variables (e.g. ozone, temperature, ranfall).
+In most climatological applications, researchers are interested in learning about the average behavior of different climate variables (e.g. ozone, temperature, rainfall).
 However, averages do not help regulators prepare for the unusual events that only happen once every 100 years.
 For example, it is important to have an idea of how much rain will come in a 100-year floor in order to construct strong enough river levees to protect lands from flooding.
-Unlike multivariate normal distributions, it is challenging to model multivariate extreme value distributions (e.g.
-generalized extreme value and generalized Pareto distribution) because few closed-form expressions exist for the density in more than two-dimensions \citep{Coles1991}.
 
-In a bivariate setting, the extremal coefficient describes the pairwise dependence between two spatial locations \citep{Smith1990}.
+The extremal coefficient describes the pairwise dependence between spatial locations \citep{Smith1990}.
 Consider a spatial process $Y(\bs) \in \calR^n$ observed at locations $s \in \calD \subset \calR^2$.
-Then the extremal coefficient, $\theta(\bs_i, \bs_j) \in [1, 2]$, is defined as
+Then the bivariate extremal coefficient, $\theta(\bs_i, \bs_j) \in [1, 2]$, is defined as
 \begin{align}
@@
 Although this method provides information regarding the spatial structure of the observations, it does not fully characterize the joint spatial dependence.
-One way to account for this limitation is the use of pairwise composite likelihoods \citep{Padoan2010,Blanchet2011,Huser2013}.
 
-Recently, skew eliptical distributions have become a popular alternative to model correlated extreme values \citep{Genton2004,Zhang2010,Padoan2011}.
-Specifically, the skew-normal and skew-t distribution offer a flexible way to handle non-symmetric data within a framework of multivariate normal and multivariate t-distributions.
-In this paper, we present a modified version of the multivariate skew-t distribution that incorporates thresholding and random spatial partitions.
+Alternatively, one can use the $\chi$ coefficient as a measure of extremal dependence.
+The $\chi$ coefficient for the upper tail is given by
+\begin{align*}
+	\chi = \lim_{c \rightarrow \infty} \Pr(Y(\bs_1) > c | Y(\bs_2) > c)
+\end{align*}
+In a stationary spatial process, we can write the $\chi$ coefficient as
+\begin{align*}
+	\chi(\bh) = \lim_{c \rightarrow \infty} \Pr(Y(0) > c | Y(\bh) > c)
+\end{align*}
+where $\bh = ||\bs_1 - \bs_2||$.
+If $\chi(\bh) = 0$, then observations are asymptotically independent at distance $\bh$.
+For Gaussian processes, $\chi(\bh) = 0$ regardless of the the distance, so they are not suitable for modeling spatially dependent extremes.
+
+Unlike multivariate normal distributions, it is challenging to model multivariate extreme value distributions (e.g.
+generalized extreme value and generalized Pareto distribution) because few closed-form expressions exist for the density in more than two-dimensions \citep{Coles1991}.
+Given this limitation, pairwise composite likelihoods have been used when modeling dependent extremes \citep{Padoan2010,Blanchet2011,Huser2013}.
+One way around the multi-dimensional limitation of multivariate extreme value distributions is to use skew elliptical distributions to model dependent extreme values \citep{Genton2004,Zhang2010,Padoan2011}.
+Due to their flexibility, the skew-normal and skew-$t$ distribution offer a flexible way to handle non-symmetric data within a framework of multivariate normal and multivariate t-distributions.
+As with the spatial Gaussian process, the skew-normal distribution is also asymptotically independent; however, the skew-$t$ does demonstrate asymptotic dependence \citep{Padoan2011}.
+
+In this paper, we present a model that has marginal distributions with flexible tails, demonstrates asymptotic dependence for $\bh$ small, and has computation on the order of Gaussian models for large space-time datasets.
+Specifically, we use a multivariate skew-$t$ distribution that incorporates thresholding and random spatial partitions.
 The advantage of using a thresholded model as opposed to a non-thresholded model is that is allows for the tails of the distribution to inform the predictions in the tails \citep{DuMouchel1983}.
@@
 
-Let $Y_t(\bs)\in \calR$ be the observed value at location $\bs$ on day $t$.  To avoid bias in estimating tail parameters, we model the thresholded data
+Let $Y_t(\bs)\in \calR$ be the observed value at location $\bs$ on day $t$.
+To avoid bias in estimating tail parameters, we model thresholded data
 \beq\label{Yt}
@@
 where $T$ is a pre-specified threshold.
-
-We first specify a model for the complete data, $Y_t(\bs)$, and then study the induced model for thresholded data, $\Yt_t(\bs)$.
-The full data model is given in Section \ref{s:data} assuming a skew normal distribution with a different variance each day.
-Computationally, the values below the threshold are updated using standard Bayesian missing data methods as described in Section \ref{s:comp}.
-The skew normal representation is from \citep{Zhang2010}.
-\subsection{Half-normal distribution}\label{s:hn}
-Let $u = |z|$ where $Z \sim N(\mu, \sigma^2)$.
-Specifically, we consider the case where $\mu = 0$. Then $U$ follows a half-normal distribution which we denote as $U \sim HN(0, 1)$, and the density is given by
-\begin{align}
-  f_U(u) = \frac{ \sqrt{2} }{ \sqrt{\pi \sigma^2} } \exp \left( - \frac{ u^2 }{ 2 \sigma^2 } \right) I(u > 0)
-\end{align}
-When $\mu = 0$, the half-normal distribution is also equivalent to a $N_{(0, \infty)}(0, \sigma^2)$ where $N_{(a, b)}(\mu, \sigma^2)$ represents a normal distribution with mean $\mu$ and standard deviation $\sigma$ that has been truncated below at $a$ and above at $b$.
+Then, assuming the full data follow a skew-$t$ distribution, we update values below the threshold using standard Bayesian missing data methods as described in Section \ref{s:comp}.
 
 \subsection{Complete data}\label{s:data}
-Consider a skew Gaussian spatial process
+We assume the data can be modeled as skew-$t$ \citep{Zhang2010}
 \begin{align} \label{eq:fullmodel}
-  Y_t(\bs) &= X_t(\bs) \beta + \sigma_1 z_t(\bs) + v_t(\bs)
+  Y_t(\bs) &= X_t(\bs) \beta + \alpha z_t + \sigma_t v_t(\bs)
 \end{align}
-where $z_t(\bs) = z_{tl}$ if $s \in P_{tl}$ where $P_{t1}, \ldots, P_{tL}$ form a partition, and $z_{tl} \iid N_{(0, \infty)}(0, 1)$, $\sigma_1 \in \calR$, and $v_t(\bs)$ is a spatial Gaussian process with mean zero and variance $\sigma_{tl}^2$.
-It can be shown \citep{Zhang2010} that $Y_t(\bs)$ follows a skew normal distribution with skewness parameter $\alpha = \frac{ \sigma_1 }{ \sigma_{tl} }$.
-We can then reexpress the model in (\ref{eq:fullmodel}) as
+where $\alpha \in \calR$ controls the skewness, $z_{t} \ind N_{(0, \infty)}(0, \sigma^2_t)$ are a random effect from a half-normal distribution (see appendix A.3), $v_t(\bs)$ is a Gaussian process with mean zero, variance one, and \Matern correlation, and $\sigma^2_t \iid \text{IG}(a, b)$.
+When marginalizing over the $z_t$ and $\sigma^2_t$ terms,
+\begin{align*}
+  Y_t(\bs) \sim \text{skew-}t (\mu, \Sigma^*, \alpha, \text{df} = 2a)
+\end{align*}
+where $\mu$ is the location, $\Sigma^* = \frac{ b }{ a } \Sigma$, $\Sigma$ is a \Matern covariance matrix, $a, b$ are the parameters from the inverse gamma distribution for $\sigma^2_t$, and $\alpha \in \calR$ controls the skewness.
+The skew-$t$ process is desirable because it has a flexible tail that is controlled by both the skewness parameter $\alpha$ and the degrees of freedom.
+This process also has the desirable characteristic that $\chi(\bh) > 0$ \citep{Padoan2011}.
+
+\subsection{Random daily partition}\label{s:part}
+One problem with the skew-$t$ distribution is that $\chi(\bh)$ does not suggest asymptotic independence as the distance increases.
+This occurs because all observations, both near and far, share the same $z_t$ and $\sigma^2_t$ terms.
+We handle this problem with a daily random partition similar to \citet{Huser2014} that allows $z_t$ and $\sigma^2_t$ to vary by site.
+The model then becomes
 \begin{align}
-  Y_t(\bs) &= X_t(\bs) \beta + \alpha z_t(\bs) + v_t(\bs)
+  Y_t(\bs) &= X_t(\bs) \beta + \alpha z_t(\bs) + \sigma_t(\bs) v_t(\bs)
 \end{align}
-where $z_t(\bs) = z_{tl}$, $z_{tl} \iid N_{(0, \infty)}(0, \sigma_{tl}^2)$, and $v_t(\bs)$ is defined as before.
+where $\alpha$ controls the skewness, $z_t(\bs) \ind \text{N}_{(0, \infty)}(0, \sigma^2_t(\bs))$, $v_t(\bs)$ is defined as before, and $\sigma^2_t(\bs) \iid \text{IG}(a, b)$.
+Consider a set of daily knots $\bw_{tk} \sim \text{Uniform}$ that define a random daily partition $P_{t1}, \ldots, P_{tK}$ such that
+\begin{align*}
+  P_{tk} = \{ \bs : k = \argmin_\ell || \bs - \bw_{t\ell} \}.
+\end{align*}
+So, for $\bs \in P_{tk}$, let
+\begin{align}
+  z_t(\bs) &= z_{tk} \label{eq:sitez}\\
+  \sigma^2_t(\bs) &= \sigma^2_{tk} \label{eq:sitesig}.
+\end{align}
+Then within each partition, $Y_t(\bs)$ follows the distribution given in (\ref{eq:fullmodel}).
 
+\subsection{Hierarchical model}\label{s:hier}
+Conditioned on $z_t(\bs)$ and $\sigma^2_t(\bs)$, $Y_t(\bs)$ is a Gaussian spatial model.
+Thus, standard geostatistical methods can be used to fit the model, and predictions can be made by Kriging at unobserved locations.
 We model this with a Bayesian hierarchical model as follows.
-Let $w_{t1}, \ldots, w_{tL}$ be partition centers so that
+Let $w_{t1}, \ldots, w_{tK}$ be partition centers so that
 \begin{align*}
-  P_{tl} = \{\bs_t : l = \argmin_k || \bs_t - w_{tk} ||\}.
+ P_{tk} = \{\bs : k = \argmin_\ell || \bs_t - w_{t\ell} ||\}.
 \end{align*}
@@
 \begin{align}
-    Y_t(\bs) \mid \Theta, z_{t1}, \ldots, z_{tL} &= X_t(\bs) \beta + \alpha z_t(\bs) + v_t(\bs) \label{eq:hier}\\
-    z_{tl}(\bs) \mid \Theta &\sim N_{(0, \infty)}(0, \sigma^2_{tl})\\
-    v_t(\bs) \mid \Theta &\sim \Matern(0, \Sigma_t(\lambda))\\
-    \sigma^2_{tl} &\iid IG(\alpha, \beta)\\
+   Y_t(\bs) \mid \Theta, z_{t}(\bs) &= X_t(\bs) \beta + \alpha z_t(\bs) + \sigma
+   _t(\bs) v_t(\bs) \label{eq:hier}\\
+   z_{tk} \mid \sigma^2_{tk} &\sim N_{(0, \infty)}(0, \sigma^2_{tk})\\
+   \sigma^2_{tk} &\iid IG(\alpha, \beta)\\
+   v_t(\bs) \mid \Theta &\sim \Matern(0, \Sigma)\\
    \alpha &\sim N(0, 10)\\
-    w_{tk} &\sim Unif(\calD)
+   w_{tk} &\sim Unif(\calR^2)
 \end{align}
-where $\Theta = \{w_{t1}, \ldots, w_{tL}, \beta, \sigma_t, \alpha, \lambda, \rho, \nu \}$; $l = \argmin_k ||\bs - w_k ||$; $\Sigma_t(\lambda)$ is a \Matern covariance matrix with variance $\sigma_{tl}^2$, spatial range $\rho$, smoothness $\nu$, and $\lambda$ is the percentage of variation that can be attributed to the partial sill; and $\calD$ is the spatial domain of interest.
+where $\Theta = \{w_{t1}, \ldots, w_{tK}, \beta, \sigma_t, \alpha, \lambda, \rho, \nu \}$; $z_t(\bs)$ and $\sigma^2_t(\bs)$ are defined as in (\ref{eq:sitez}) and (\ref{eq:sitesig}); \mbox{$k = \argmin_\ell ||\bs - \bw_\ell ||$}; and $\Sigma$ is a \Matern covariance matrix with variance one, spatial range $\rho$, smoothness $\nu$.
 
@@
 We can use Gibbs sampling to update $Y_t(\bs)$ for observations that are below $T$, the thresholded value. Given $\Theta$, $Y_t(\bs)$ has truncated normal full conditional with these parameter values.
-So we sample $Y_t(\bs) \sim N_{(-\infty, T)}$
+So we sample $Y_t(\bs) \sim N_{(-\infty, T)}(\mu(\bs), \Sigma)$
 
@@
 
-\section*{Appendix A.2: MCMC Details}
+\section*{Appendix A.2: MCMC details}
 
-\subsection*{Priors}
+\section*{Appendix A.3: Half-normal distribution}
+Let $u = |z|$ where $Z \sim N(\mu, \sigma^2)$.
+Specifically, we consider the case where $\mu = 0$. Then $U$ follows a half-normal distribution which we denote as $U \sim HN(0, 1)$, and the density is given by
+\begin{align}
+  f_U(u) = \frac{ \sqrt{2} }{ \sqrt{\pi \sigma^2} } \exp \left( - \frac{ u^2 }{ 2 \sigma^2 } \right) I(u > 0)
+\end{align}
+When $\mu = 0$, the half-normal distribution is also equivalent to a $N_{(0, \infty)}(0, \sigma^2)$ where $N_{(a, b)}(\mu, \sigma^2)$ represents a normal distribution with mean $\mu$ and standard deviation $\sigma$ that has been truncated below at $a$ and above at $b$.
 

### 2014-08-19 pot/LaTeX/skewnormal.tex
Include intercept in X

@@
 
-The extremal coefficient describes the pairwise dependence between spatial locations \citep{Smith1990}.
-Consider a spatial process $Y(\bs) \in \calR^n$ observed at locations $s \in \calD \subset \calR^2$.
-Then the bivariate extremal coefficient, $\theta(\bs_i, \bs_j) \in [1, 2]$, is defined as
-\begin{align}
-  \Pr(Y(\bs_i) < c, Y(\bs_j) < c) = \Pr(Y(\bs_i) < c)^{\theta(\bs_i, \bs_j) }. \label{eq:extremalcoef}
-\end{align}
-One way to characterize the dependence over the entire set of spatial locations is to calculate all of the pairwise extremal coefficients.
-Although this method provides information regarding the spatial structure of the observations, it does not fully characterize the joint spatial dependence.
+Unlike multivariate normal distributions, it is challenging to model multivariate extreme value distributions (e.g.
+generalized extreme value and generalized Pareto distribution) because few closed-form expressions exist for the density in more than two-dimensions \citep{Coles1991}.
+Given this limitation, pairwise composite likelihoods have been used when modeling dependent extremes \citep{Padoan2010,Blanchet2011,Huser2013}.
+One way around the multi-dimensional limitation of multivariate extreme value distributions is to use skew elliptical distributions to model dependent extreme values \citep{Genton2004,Zhang2010,Padoan2011}.
+Due to their flexibility, the skew-normal and skew-$t$ distribution offer a flexible way to handle non-symmetric data within a framework of multivariate normal and multivariate t-distributions.
+As with the spatial Gaussian process, the skew-normal distribution is also asymptotically independent; however, the skew-$t$ does demonstrate asymptotic dependence \citep{Padoan2011}.
 
-Alternatively, one can use the $\chi$ coefficient as a measure of extremal dependence.
+In this paper, we present a model that has marginal distributions with flexible tails, demonstrates asymptotic dependence for small $\bh$, and has computation on the order of Gaussian models for large space-time datasets.
+Specifically, our contribution is to incorporate thresholding and random spatial partitions using a multivariate skew-$t$ distribution.
+The advantage of using a thresholded model as opposed to a non-thresholded model is that is allows for the tails of the distribution to inform the predictions in the tails \citep{DuMouchel1983}.
+The random spatial partitions increase the flexibility of the model by giving us a way to account for observations in spatial regions that may have higher variability due to certain regional climate variables (e.g.
+forest fires).
+
+\section{Extreme distributions}
+Maybe some information about max-stable processes and standard analyses?
+
+\subsection{Extremal dependence}
+A popular measure of extremal dependence is the $\chi$.
 The $\chi$ coefficient for the upper tail is given by
@@
 For Gaussian processes, $\chi(\bh) = 0$ regardless of the the distance, so they are not suitable for modeling spatially dependent extremes.
+However, for the process described in Section \ref{s:data}, $\chi(\bh) > 0$ \citep{Padoan2011}.
 
-Unlike multivariate normal distributions, it is challenging to model multivariate extreme value distributions (e.g.
-generalized extreme value and generalized Pareto distribution) because few closed-form expressions exist for the density in more than two-dimensions \citep{Coles1991}.
-Given this limitation, pairwise composite likelihoods have been used when modeling dependent extremes \citep{Padoan2010,Blanchet2011,Huser2013}.
-One way around the multi-dimensional limitation of multivariate extreme value distributions is to use skew elliptical distributions to model dependent extreme values \citep{Genton2004,Zhang2010,Padoan2011}.
-Due to their flexibility, the skew-normal and skew-$t$ distribution offer a flexible way to handle non-symmetric data within a framework of multivariate normal and multivariate t-distributions.
-As with the spatial Gaussian process, the skew-normal distribution is also asymptotically independent; however, the skew-$t$ does demonstrate asymptotic dependence \citep{Padoan2011}.
-
-In this paper, we present a model that has marginal distributions with flexible tails, demonstrates asymptotic dependence for $\bh$ small, and has computation on the order of Gaussian models for large space-time datasets.
-Specifically, we use a multivariate skew-$t$ distribution that incorporates thresholding and random spatial partitions.
-The advantage of using a thresholded model as opposed to a non-thresholded model is that is allows for the tails of the distribution to inform the predictions in the tails \citep{DuMouchel1983}.
-The random spatial partitions increase the flexibility of the model by giving us a way to account for observations in spatial regions that may have higher variability due to certain regional climate variables (e.g.
-forest fires).
+Another common measure of extremal spatial dependence is the extremal coefficient which describes the pairwise dependence between spatial locations \citep{Smith1990}.
+Consider a spatial process $Y(\bs) \in \calR^n$ observed at locations $s \in \calD \subset \calR^2$.
+Then the bivariate extremal coefficient, $\theta(\bs_i, \bs_j) \in [1, 2]$, is defined as
+\begin{align}
+  \Pr(Y(\bs_i) < c, Y(\bs_j) < c) = \Pr(Y(\bs_i) < c)^{\theta(\bs_i, \bs_j) }. \label{eq:extremalcoef}
+\end{align}
+One way to characterize the dependence over the entire set of spatial locations is to calculate all of the pairwise extremal coefficients.
+Although this method provides information regarding the spatial structure of the observations, it does not fully characterize the joint spatial dependence.
 
 \section{Statistical model}\label{s:model}
-
 Let $Y_t(\bs)\in \calR$ be the observed value at location $\bs$ on day $t$.
-To avoid bias in estimating tail parameters, we model thresholded data
+To avoid bias in estimating tail parameters, we model censored data
 \beq\label{Yt}
@@
 \end{align*}
-where $\mu$ is the location, $\Sigma^* = \frac{ b }{ a } \Sigma$, $\Sigma$ is a \Matern covariance matrix, $a, b$ are the parameters from the inverse gamma distribution for $\sigma^2_t$, and $\alpha \in \calR$ controls the skewness.
-The skew-$t$ process is desirable because it has a flexible tail that is controlled by both the skewness parameter $\alpha$ and the degrees of freedom.
-This process also has the desirable characteristic that $\chi(\bh) > 0$ \citep{Padoan2011}.
+where $\mu$ is the location, $\Sigma^* = \frac{ b }{ a } \Sigma$, $\Sigma$ is a \Matern covariance matrix, and $\alpha \in \calR$ controls the skewness.
+The skew-$t$ process is desirable because it has a flexible tail that is controlled by both the skewness parameter $\alpha$ and the degrees of freedom $2a$.
 
@@
 where $\alpha$ controls the skewness, $z_t(\bs) \ind \text{N}_{(0, \infty)}(0, \sigma^2_t(\bs))$, $v_t(\bs)$ is defined as before, and $\sigma^2_t(\bs) \iid \text{IG}(a, b)$.
-Consider a set of daily knots $\bw_{tk} \sim \text{Uniform}$ that define a random daily partition $P_{t1}, \ldots, P_{tK}$ such that
+Consider a set of daily spatial knots $\bw_{tk} \sim \text{Uniform}(\calD)$ that define a random daily partition $P_{t1}, \ldots, P_{tK}$ of the spatial domain of interest $\calD \subset \calR^2$ such that
 \begin{align*}
@@
 We model this with a Bayesian hierarchical model as follows.
-Let $w_{t1}, \ldots, w_{tK}$ be partition centers so that
+Let $w_{t1}, \ldots, w_{tK}$ be a set of daily spatial knots in a spatial domain of interest, $\calD$, so that
 \begin{align*}
@@
    \alpha &\sim N(0, 10)\\
-   w_{tk} &\sim Unif(\calR^2)
+   w_{tk} &\sim Unif(\calD)
 \end{align}
@@
 
+\section{Simulation study}\label{s:simstudy}
+In this section, we conduct a simulation study to investigate how the number of partitions and the level of thresholding impact the accuracy of predictions made by the model.
+For all simulation designs, we generate data from the model presented in Section \ref{s:part} using $n=130$ sites and $T=50$ independent days.
+The sites are generated Uniform$([0, 1] \times [0, 1])$.
+We generate data from 6 different simulation designs:
+\begin{enumerate} \setlength{\itemsep}{-0.5em}
+  \item Gaussian marginal, $K=1$ knots
+  \item $t$ marginal, $K=1$ knots
+  \item $t$ marginal, $K=5$ knots
+  \item skew-$t$ marginal, $K=1$ knots
+  \item skew-$t$ marginal, $K=5$ knots
+  \item Mixture of Gaussian marginal and $t$ marginal, $K=1$ knots.
+\end{enumerate}
+In the first five designs, the $v_t(\bs)$ terms are generated using a \Matern covariance with parameters smoothness ($\nu = 0.5$), spatial range ($\rho = 0.1$), and variance $\sigma^2_t$ where $\sigma^2_t \iid \text{IG}(2, 8)$.
+In each covariance matrix, the proportion of the variance accounted for by the spatial variation is $\gamma = 0.9$ while the proportion of the variance accounted for by the nugget effect is $0.1$.
+In the first design, the same $\sigma^2$ is used for all days.
+For skew marginals, $\alpha = 3$ was used, and the $z_t$ are generated as described in (\ref{eq:sitez}).
+In the sixth design, each day has either a Gaussian or a $t$ marginal with probability of 0.5.
+All days in the sixth design with Gaussian marginal distributions use the same $\sigma^2$ with $v_t(\bs)$ generated as described for the other five designs.
+For days in the sixth design with $t$ marginal distributions, the spatial range is four times as large ($\rho = 0.4$) as the days with Gaussian marginals.
+In all six designs, the mean ($\mu(\bs)$) is assumed to be constant across space.
+
+$M = 50$ data sets are generated for each design.
+For each data set we fit the data using
+\begin{enumerate} \setlength{\itemsep}{-0.5em}
+  \item Gaussian marginal, $K=1$ knots
+  \item skew-$t$ marginal, $K=1$ knots, $T=0.0$
+  \item skew-$t$ marginal, $K=1$ knots, $T=0.9$
+  \item skew-$t$ marginal, $K=5$ knots, $T=0.0$
+  \item skew-$t$ marginal, $K=5$ knots, $T=0.9$
+\end{enumerate}
+where $T=0.9$ indicates that data are thresholded at the 90th sample quantile of the data.
+The design matrix $\bX$ includes on the intercept with a prior of $\beta \sim \text{N}(0, 10)$.
+The spatial covariance parameters have priors $\log(\nu) \sim \text{N}(-1.2, 1)$, $\gamma \sim \text{U}(0, 1)$, $\log(\rho) \sim \text{N}(-2, 1)$.
+The skewness parameter has prior $\alpha \sim \text{N}(0, 2)$.
+The residual variance terms have priors $\sigma^2_t(\bs) \sim \text{IG}(0.1, 0.1)$.
+
+\subsection{Cross validation}\label{s:modelselect}
+Models were compared using two-fold cross validation where data were split with 50\% being used as a training set and 50\% as a testing set.
+The model was fit using the training set, and predictions were generated at the testing site locations.
+Because one of the primary goals of this model is to predict extreme events, we quantify use Brier scores and quantile scores to select the model that best fits the data \citep{Gneiting2007}.
+The Brier score for predicting exceedance of a threshold $c$ is given by $[e(c) - P(c)]^2$ where $e(c) = I[y>c]$ is an indicator function indicating that a test set value, $y$, has exceeded the threshold, and $P(c)$ is the predicted probability of exceeding $c$.
+The quantile score for the $\tau$th quantile is $2\{ I[y < \widehat{q}(\tau)] - \tau\} (\widehat{q} - y)$ where $y$ is an AQS test set value and $\widehat{q}(\tau)$ is the estimated $\tau$th quantile.
+For both the Brier score and the quantile score, a lower score indicates a better fit.
+These scores were averaged over all sites and days to obtain a single quantile score for each dataset.
+
 \section{Data analysis}\label{s:analysis}
+To illustrate this method, we consider the daily maximum 8-hour ozone measurements for July 2005 at 735 Air Quality System (AQS) monitoring sites in the eastern United States as the response.
+For each site, we also have covariate information containing the estimated ozone from the Community Multi-scale Air Quality (CMAQ) modeling system.
+We fit the model using Gaussian and skew-$t$ marginal distributions, $K=1, 5, 10, 15$ partitions, with $Y(\bs)$ censored at $T = 0, 50, 75, 90$ ppb as described in Section \ref{s:data}.
+We also include a max-stable analysis using the method by ????
+All methods assume the location can be expressed as
+\begin{align}
+  \mu_t(\bs) = \beta_0 + \beta_1 \cdot \text{CMAQ}_t(\bs).
+\end{align}
+To explore the extremal dependence both over space and time, we plot $\chi(\bs)$ and $\chi(t)$.
+For each model, Brier scores and quantile scores were were averaged over all sites and days to obtain a single quantile score for each dataset.
+At a particular threshold or quantile level, the model that fits the best is the one with the lowest score.
+
+\subsection{Results}\label{s:results}
 

### 2014-08-23 pot/LaTeX/jsm2014/jsm2014.tex
New plots, tweaks to simstudy

--- 2014-08-07 pot/LaTeX/jsm2014.tex
@@
   \begin{itemize} \setlength{\itemsep}{1em}
-    \item Average behavior is important to understand, but it does not paint the whole picture.
+    \item Average behavior is important to understand, but it does not paint the whole picture
     \begin{itemize}
-      \item e.g. When constructing river levees, engineers need to be able to estimate a 100-year or 1000-year flood levels.
+      \item e.g. When constructing river levees, engineers need to be able to estimate a 100-year or 1000-year flood levels
       \item e.g. Probability of exceeding a certain threshold level
     \end{itemize}
-    \item Spatial methods borrow information across space to estimate spatial correlation and make predictions by Kriging at unknown locations.
-    \item Want to explore similar methods for extremes.
+    \item Spatial methods borrow information across space to estimate spatial correlation and make predictions by Kriging at unknown locations
+    \item Want to explore similar methods for extremes
   \end{itemize}
@@
       \item Incorporates more data than block maxima
-      \item Select a threshold, $T$, and use the Generalized Pareto distribution (GPD) to model the exceedances.
+      \item Select a threshold, $T$, and use the Generalized Pareto distribution (GPD) to model the exceedances
       \item Temporal dependence may be an issue between observations (e.g. flood levels don't dissipate overnight)
@@
       \item Nice computing properties (closed-form likelihood)
-    \item For a Gaussian spatial model $\lim_{c \rightarrow \infty} \chi(\bh) = 0$ regardless of the strength of the correlation in the bulk of the distribution.
+    \item For a Gaussian spatial model $\lim_{c \rightarrow \infty} \chi(\bh) = 0$ regardless of the strength of the correlation in the bulk of the distribution
     \item Tail is not flexible (Gaussian is light tailed)
@@
    \item \alert{Conditioned} on $z_t$ and $\sigma^2_t$, $Y_t(\bs)$ is a Gaussian spatial model
-    \item Can use standard geostatistical methods to fit this model.
-    \item Predictions can be made through Kriging.
+    \item Can use standard geostatistical methods to fit this model
+    \item Predictions can be made through Kriging
     \item \alert{Marginalizing} over $z_t$ and $\sigma^2_t$ (via MCMC),
@@
     \item For this distribution, $\chi(\bh)$ shows asymptotic dependence that does not approach 0 as $\bh \rightarrow \infty$
-   \item This occurs because all observations (near and far) share the same $z_t$ and $\sigma_t^2$.
-    \item We deal with this through a daily random partition (similar to Huser and Davison).
+   \item This occurs because all observations (near and far) share the same $z_t$ and $\sigma_t^2$
+    \item We deal with this through a daily random partition (similar to Huser and Davison)
   \end{itemize}
@@
   \begin{itemize} \setlength{\itemsep}{0.5em}
-    \item Daily random partition allows $z_t$ and $\sigma^2_t$ to vary by site.
+    \item Daily random partition allows $z_t$ and $\sigma^2_t$ to vary by site
     \begin{align*}
@@
     \end{align*}
-    \item Within each partition $Y_t(\bs)$ has the same MV skew-t distribution as before.
+    \item Within each partition $Y_t(\bs)$ has the same MV skew-t distribution as before
   \end{itemize}
@@
     \end{enumerate}
-    \item Priors are selected to be conjugate when possible.
+    \item Priors are selected to be conjugate when possible
   \end{itemize}
@@
     \includegraphics[width=1\linewidth]{./plots/ozone-day.pdf}
-    \caption{Max 8-hour ozone measurements at 85 sites in NC, SC, and GA for days 5 and 34.}
+    \caption{Max 8-hour ozone measurements at 85 sites in NC, SC, and GA for days 5 and 34}
    \end{figure}
@@
   \begin{itemize}
-  	\item Quantile score results are similar.
+  	\item Quantile score results are similar
   \end{itemize}
@@
     \end{itemize}
-    \item Preliminary results are inconclusive.
+    \item Preliminary results are inconclusive
   \end{itemize}

### 2014-08-23 pot/code/analysis/simstudy/pot2-b.R

@@ -35,7 +35,7 @@ analysis <- "b"
 iters <- 20000; burn <- 10000; update <- 1000; thin <- 1
 nsets <- 5
 
-for (g in 1:2) {
+for (g in 1:10) {
   fit.1 <- vector(mode="list", length=nsets)
   y.validate <- array(NA, dim=c(ntest, nt, nsets))
   outputfile <- paste(setting, "-", analysis, "-", g, ".RData", sep="")

### 2014-09-08 pot/LaTeX/proof_dist.tex
minor changes to manuscript.

Consider two points $\bs_1, \bs_2 \in \calD \subset \calR^2$. Let $\calD = B_{2h}(\bm)$ which is a ball of radius $h > 0$ centered at $\bm$, the midpoint between $\bs_1$ and $\bs_2$, and $h = ||\bs_1 - \bs_2||$. Let $w_1, \ldots, w_K \sim \text{Unif}(\calD)$ be a set of random knots uniformly distributed over $\calD$. Define $g(\bs_i) = \argmin_j || \bs_i - w_j ||$ to be the partition in which $\bs_i$ resides. Let $\pi(h) = \Pr[ g(\bs_1) = g(\bs_2)]$ be the probability that both sites are in the same partition. Then $\pi(h)$ is the probability that $w_j$ is the only knot in $(B_a(\bs_1) \cup B_b(\bs_2) \cap \calD)$ where $a = ||\bs_1 - w_j||$ and $b = ||\bs_2 - w_j||$.

### 2014-09-10 pot/LaTeX/skewnormal.tex
minor changes to manuscript.
manuscript updates
added in proof about spatial locations

@@
 Given this limitation, pairwise composite likelihoods have been used when modeling dependent extremes \citep{Padoan2010,Blanchet2011,Huser2013}.
+
 One way around the multi-dimensional limitation of multivariate extreme value distributions is to use skew elliptical distributions to model dependent extreme values \citep{Genton2004,Zhang2010,Padoan2011}.
@@
 As with the spatial Gaussian process, the skew-normal distribution is also asymptotically independent; however, the skew-$t$ does demonstrate asymptotic dependence \citep{Padoan2011}.
+Although asymptotic dependence is desirable between sites that are near one another, one drawback to the skew-$t$ is that sites remain asymptotically dependent even at far distances.
 
@@
 The advantage of using a thresholded model as opposed to a non-thresholded model is that is allows for the tails of the distribution to inform the predictions in the tails \citep{DuMouchel1983}.
-The random spatial partitions increase the flexibility of the model by giving us a way to account for observations in spatial regions that may have higher variability due to certain regional climate variables (e.g.
-forest fires).
+The random spatial partition alleviates the long-range spatial dependence seen by the skew-$t$.
 
-\section{Extreme distributions}
-Maybe some information about max-stable processes and standard analyses?
+\section{Statistical model}\label{s:model}
+Let $Y_t(\bs)\in \calR$ be the observed value at location $\bs$ and timepoint $t$.
+To avoid bias in estimating tail parameters, we model censored data
+\beq\label{Yt}
+  \Yt_t(\bs) = \left\{
+          \begin{array}{ll}
+            Y_t(\bs) & Y_t(\bs)>T \\
+            T & Y_t(\bs)\le T
+          \end{array}
+        \right.
+\eeq
+where $T$ is a pre-specified threshold.
+Then, assuming the full data follow a skew-$t$ distribution, we update values censored below the threshold using standard Bayesian missing data methods as described in Section \ref{s:comp}.
+
+\subsection{Skew-$t$ process}\label{s:data}
+We assume the data can be modeled as skew-$t$.
+\citet{Zhang2010} show that the skew-$t$ can be written as the hierarchical model
+\begin{align} \label{eq:fullmodel}
+  Y_t(\bs) &= X_t(\bs) \beta + \alpha z_t + \sigma_t v_t(\bs)
+\end{align}
+where $\alpha \in \calR$ controls the skewness, $z_{t} \ind N_{(0, \infty)}(0, \sigma^2_t)$ are a random effect from a half-normal distribution (see appendix A.3), $v_t(\bs)$ is a Gaussian process with mean zero, variance one, and \Matern correlation, and $\sigma^2_t \iid \text{IG}(a, b)$.
+When marginalizing over the $z_t$ and $\sigma^2_t$ terms,
+\begin{align*}
+  Y_t(\bs) \sim \text{skew-}t (\mu, \Sigma^*, \alpha, \text{df} = 2a)
+\end{align*}
+where $\mu$ is the location, $\Sigma^* = \frac{ b }{ a } \Sigma$, $\Sigma$ is a \Matern covariance matrix, and $\alpha \in \calR$ controls the skewness.
+The skew-$t$ process is desirable because of its flexible tail, controlled by both the skewness parameter $\alpha$ and the degrees of freedom $2a$, and its joint distribution is a multivariate skew-$t$ distribution.
 
 \subsection{Extremal dependence}
-A popular measure of extremal dependence is the $\chi$.
+One common measure of extremal spatial dependence is the extremal coefficient which describes the pairwise dependence between spatial locations \citep{Smith1990}.
+Consider a spatial process $Y(\bs) \in \calR^n$ observed at locations $s \in \calD \subset \calR^2$.
+Then the bivariate extremal coefficient, $\theta(\bs_i, \bs_j) \in [1, 2]$, is defined as
+\begin{align}
+  \Pr(Y(\bs_i) < c, Y(\bs_j) < c) = \Pr(Y(\bs_i) < c)^{\theta(\bs_i, \bs_j) }. \label{eq:extremalcoef}
+\end{align}
+One way to characterize the dependence over the entire set of spatial locations is to calculate all of the pairwise extremal coefficients.
+Although this method provides information regarding the spatial structure of the observations, it does not fully characterize the joint spatial dependence.
+
+Another popular measure of extremal dependence is the $\chi$ statistic.
 The $\chi$ coefficient for the upper tail is given by
@@
 
-Another common measure of extremal spatial dependence is the extremal coefficient which describes the pairwise dependence between spatial locations \citep{Smith1990}.
-Consider a spatial process $Y(\bs) \in \calR^n$ observed at locations $s \in \calD \subset \calR^2$.
-Then the bivariate extremal coefficient, $\theta(\bs_i, \bs_j) \in [1, 2]$, is defined as
-\begin{align}
-  \Pr(Y(\bs_i) < c, Y(\bs_j) < c) = \Pr(Y(\bs_i) < c)^{\theta(\bs_i, \bs_j) }. \label{eq:extremalcoef}
-\end{align}
-One way to characterize the dependence over the entire set of spatial locations is to calculate all of the pairwise extremal coefficients.
-Although this method provides information regarding the spatial structure of the observations, it does not fully characterize the joint spatial dependence.
-
-\section{Statistical model}\label{s:model}
-Let $Y_t(\bs)\in \calR$ be the observed value at location $\bs$ on day $t$.
-To avoid bias in estimating tail parameters, we model censored data
-\beq\label{Yt}
-  \Yt_t(\bs) = \left\{
-          \begin{array}{ll}
-            Y_t(\bs) & Y_t(\bs)>T \\
-            T & Y_t(\bs)\le T
-          \end{array}
-        \right.
-\eeq
-where $T$ is a pre-specified threshold.
-Then, assuming the full data follow a skew-$t$ distribution, we update values below the threshold using standard Bayesian missing data methods as described in Section \ref{s:comp}.
-
-\subsection{Complete data}\label{s:data}
-We assume the data can be modeled as skew-$t$ \citep{Zhang2010}
-\begin{align} \label{eq:fullmodel}
-  Y_t(\bs) &= X_t(\bs) \beta + \alpha z_t + \sigma_t v_t(\bs)
-\end{align}
-where $\alpha \in \calR$ controls the skewness, $z_{t} \ind N_{(0, \infty)}(0, \sigma^2_t)$ are a random effect from a half-normal distribution (see appendix A.3), $v_t(\bs)$ is a Gaussian process with mean zero, variance one, and \Matern correlation, and $\sigma^2_t \iid \text{IG}(a, b)$.
-When marginalizing over the $z_t$ and $\sigma^2_t$ terms,
-\begin{align*}
-  Y_t(\bs) \sim \text{skew-}t (\mu, \Sigma^*, \alpha, \text{df} = 2a)
-\end{align*}
-where $\mu$ is the location, $\Sigma^* = \frac{ b }{ a } \Sigma$, $\Sigma$ is a \Matern covariance matrix, and $\alpha \in \calR$ controls the skewness.
-The skew-$t$ process is desirable because it has a flexible tail that is controlled by both the skewness parameter $\alpha$ and the degrees of freedom $2a$.
-
 \subsection{Random daily partition}\label{s:part}
-One problem with the skew-$t$ distribution is that $\chi(\bh)$ does not suggest asymptotic independence as the distance increases.
+One problem with the skew-$t$ distribution is that all sites are asymptotically dependent regardless of their spatial separation.
 This occurs because all observations, both near and far, share the same $z_t$ and $\sigma^2_t$ terms.
@@
 \begin{align}
-  Y_t(\bs) &= X_t(\bs) \beta + \alpha z_t(\bs) + \sigma_t(\bs) v_t(\bs)
+  Y_t(\bs) &= X_t(\bs) \beta + \alpha z_t(\bs) + \sigma_t(\bs) v_t(\bs).
 \end{align}
-where $\alpha$ controls the skewness, $z_t(\bs) \ind \text{N}_{(0, \infty)}(0, \sigma^2_t(\bs))$, $v_t(\bs)$ is defined as before, and $\sigma^2_t(\bs) \iid \text{IG}(a, b)$.
+In this extension of (\ref{eq:fullmodel}), $z_t(\bs)$ and $\sigma_t(\bs)$ are allowed to vary by site.
+Their spatial variation is determined by the random partition model defined below.
 Consider a set of daily spatial knots $\bw_{tk} \sim \text{Uniform}(\calD)$ that define a random daily partition $P_{t1}, \ldots, P_{tK}$ of the spatial domain of interest $\calD \subset \calR^2$ such that
@@
 Then within each partition, $Y_t(\bs)$ follows the distribution given in (\ref{eq:fullmodel}).
+When incorporating the random daily partition, the $chi$ statistic becomes
+\begin{align}
+  \chi(\bh) = \lim_{c \rightarrow \infty} \pi(\bh) \chi(\bh)
+\end{align}
+where $\pi(h)$ is the probability that two sites separated by distance $h$ are in the same partition.
+Asymptotic dependence is eliminated as $h$ increases, because
+\begin{align}
+  \lim_{h \rightarrow \infty} \chi(\bh) = \lim_{h \rightarrow \infty} \pi(\bh) \chi(\bh) = 0.
+\end{align}
+A proof of this is given in Appendix A.2.
 
 \subsection{Hierarchical model}\label{s:hier}
-Conditioned on $z_t(\bs)$ and $\sigma^2_t(\bs)$, $Y_t(\bs)$ is a Gaussian spatial model.
+Conditioned on $z_{tk}(\bs) \iid $ $N(0, \sigma^2_{tk})$, $\sigma^2_{tk}(\bs) \sim IG(a, b)$, and $P_tk$, the marginal distributions are skew-$t$ and the joint distribution within a partition is multivariate skew-$t$.
+However, we do not fix the partitions, they are treated as unknown and updated in the MCMC.
 Thus, standard geostatistical methods can be used to fit the model, and predictions can be made by Kriging at unobserved locations.
@@
    _t(\bs) v_t(\bs) \label{eq:hier}\\
+   z_t(\bs) = z_{tk} \text{ if } \bs \in P_{tk}\\
+   \sigma^2_{t}(\bs) = \sigma^2_{tk} \text{ if } \bs \in P_{tk}
    z_{tk} \mid \sigma^2_{tk} &\sim N_{(0, \infty)}(0, \sigma^2_{tk})\\
@@
 \end{align}
-where $\Theta = \{w_{t1}, \ldots, w_{tK}, \beta, \sigma_t, \alpha, \lambda, \rho, \nu \}$; $z_t(\bs)$ and $\sigma^2_t(\bs)$ are defined as in (\ref{eq:sitez}) and (\ref{eq:sitesig}); \mbox{$k = \argmin_\ell ||\bs - \bw_\ell ||$}; and $\Sigma$ is a \Matern covariance matrix with variance one, spatial range $\rho$, smoothness $\nu$.
+where $\Theta = \{w_{t1}, \ldots, w_{tK}, \beta, \sigma_t, \alpha, \lambda, \rho, \nu \}$; \mbox{$k = \argmin_\ell ||\bs - \bw_\ell ||$}; and $\Sigma$ is a \Matern covariance matrix with variance one, spatial range $\rho$, smoothness $\nu$.
 
@@
 In this section, we conduct a simulation study to investigate how the number of partitions and the level of thresholding impact the accuracy of predictions made by the model.
-For all simulation designs, we generate data from the model presented in Section \ref{s:part} using $n=130$ sites and $T=50$ independent days.
-The sites are generated Uniform$([0, 1] \times [0, 1])$.
+
+\subsection{Design}\label{s:simdesign}
+For all simulation designs, we generate data from the model presented in Section \ref{s:part} using $n_s=130$ sites and $n_t=50$ independent days.
+The sites are generated Uniform$([0, 10] \times [0, 10])$.
 We generate data from 6 different simulation designs:
 \begin{enumerate} \setlength{\itemsep}{-0.5em}
-  \item Gaussian marginal, $K=1$ knots
-  \item $t$ marginal, $K=1$ knots
+  \item Gaussian marginal, $K=1$ knot
+  \item $t$ marginal, $K=1$ knot
   \item $t$ marginal, $K=5$ knots
@@
   \item skew-$t$ marginal, $K=5$ knots
-  \item Mixture of Gaussian marginal and $t$ marginal, $K=1$ knots.
+  \item Max-stable.
 \end{enumerate}
-In the first five designs, the $v_t(\bs)$ terms are generated using a \Matern covariance with parameters smoothness ($\nu = 0.5$), spatial range ($\rho = 0.1$), and variance $\sigma^2_t$ where $\sigma^2_t \iid \text{IG}(2, 8)$.
-In each covariance matrix, the proportion of the variance accounted for by the spatial variation is $\gamma = 0.9$ while the proportion of the variance accounted for by the nugget effect is $0.1$.
-In the first design, the same $\sigma^2$ is used for all days.
-For skew marginals, $\alpha = 3$ was used, and the $z_t$ are generated as described in (\ref{eq:sitez}).
-In the sixth design, each day has either a Gaussian or a $t$ marginal with probability of 0.5.
-All days in the sixth design with Gaussian marginal distributions use the same $\sigma^2$ with $v_t(\bs)$ generated as described for the other five designs.
-For days in the sixth design with $t$ marginal distributions, the spatial range is four times as large ($\rho = 0.4$) as the days with Gaussian marginals.
+In the first five designs, the $v_t(\bs)$ terms are generated using a \Matern covariance with smoothness parameter, $\nu = 0.5$, and spatial range, $\rho = 0.1$.
+For the covariance matrices in designs 1 -- 5, the proportion of the variance accounted for by the spatial variation is $\gamma = 0.9$ while the proportion of the variance accounted for by the nugget effect is $0.1$.
+In the first design, $\sigma^2 = 2$ is used for all days.
+For designs 2 -- 4, $sigma^2_{tk} \iid \text{IG}(3, 8)$
+For designs 1 -- 3, we set $\alpha = 0$.
+For designs four and five, $\alpha = 3$ was used, and the $z_t$ are generated as described in (\ref{eq:sitez}).
+In the sixth design, we generate from a spatial max-stable distribution \citep{Reich2012} with parameters $\mu = 1, \sigma=1, \xi=0.2$ and 144 spatial knots on a regular lattice in the square $[1, 9] \times [1, 9]$.
 In all six designs, the mean ($\mu(\bs)$) is assumed to be constant across space.
@@
   \item Gaussian marginal, $K=1$ knots
-  \item skew-$t$ marginal, $K=1$ knots, $T=0.0$
-  \item skew-$t$ marginal, $K=1$ knots, $T=0.9$
-  \item skew-$t$ marginal, $K=5$ knots, $T=0.0$
-  \item skew-$t$ marginal, $K=5$ knots, $T=0.9$
+  \item skew-$t$ marginal, $K=1$ knots, $T=-\infty$
+  \item skew-$t$ marginal, $K=1$ knots, $T=q(0.90)$
+  \item skew-$t$ marginal, $K=5$ knots, $T=-\infty$
+  \item skew-$t$ marginal, $K=5$ knots, $T=q(0.9)$
 \end{enumerate}
-where $T=0.9$ indicates that data are thresholded at the 90th sample quantile of the data.
+where $q(0.9)$ is the 90th sample quantile of the data.
 The design matrix $\bX$ includes on the intercept with a prior of $\beta \sim \text{N}(0, 10)$.
@@
 \subsection{Cross validation}\label{s:modelselect}
-Models were compared using two-fold cross validation where data were split with 50\% being used as a training set and 50\% as a testing set.
+Models were compared using cross validation with 100 sites used as training sites and 30 sites witheld for testing.
 The model was fit using the training set, and predictions were generated at the testing site locations.
 Because one of the primary goals of this model is to predict extreme events, we quantify use Brier scores and quantile scores to select the model that best fits the data \citep{Gneiting2007}.
-The Brier score for predicting exceedance of a threshold $c$ is given by $[e(c) - P(c)]^2$ where $e(c) = I[y>c]$ is an indicator function indicating that a test set value, $y$, has exceeded the threshold, and $P(c)$ is the predicted probability of exceeding $c$.
-The quantile score for the $\tau$th quantile is $2\{ I[y < \widehat{q}(\tau)] - \tau\} (\widehat{q} - y)$ where $y$ is an AQS test set value and $\widehat{q}(\tau)$ is the estimated $\tau$th quantile.
+The Brier score for predicting exceedance of a threshold $c$ is given by $[e(c) - P(c)]^2$ where $e(c) = I[y>c]$ is an indicator function indicating that a test set value, $y$, has exceeded the threshold, $c$, and $P(c)$ is the predicted probability of exceeding $c$.
+The quantile score for the $\tau$th quantile is $2\{ I[y < \widehat{q}(\tau)] - \tau\} (\widehat{q}(\tau) - y)$ where $y$ is a AQS test set value and $\widehat{q}(\tau)$ is the estimated $\tau$th quantile.
 For both the Brier score and the quantile score, a lower score indicates a better fit.
 These scores were averaged over all sites and days to obtain a single quantile score for each dataset.
+
+\subsection{Results}\label{s:simresults}
 
@@
 
-\section*{Appendix A.2: MCMC details}
+\section*{Appendix A.2: Proof that $\lim_{h \rightarrow \infty} \pi(h) = 0$}
+\input{proof_dist.tex}
 

### 2014-09-11 pot/LaTeX/skewnormal.tex
results compilation

@@
 
-In this paper, we present a model that has marginal distributions with flexible tails, demonstrates asymptotic dependence for small $\bh$, and has computation on the order of Gaussian models for large space-time datasets.
+In this paper, we present a model that has marginal distributions with flexible tails, demonstrates asymptotic dependence for observations at sites that are near to one another, and has computation on the order of Gaussian models for large space-time datasets.
 Specifically, our contribution is to incorporate thresholding and random spatial partitions using a multivariate skew-$t$ distribution.
@@
 The random spatial partition alleviates the long-range spatial dependence seen by the skew-$t$.
+
+The paper is organized as follows.
+In section \ref{s:model}, we describe the skew-$t$ process with partitioning and thresholding.
+The computing is described in section \ref{s:comp}.
+In section \ref{s:simstudy}, we present a simulation study that examines the predictive capabilities of this model compared with a na\:{i}ve Gaussian method.
+We then compare our method to Gaussian and max-stable methods with a data analysis of ozone measurements from the eastern US in section \ref{s:analysis}.
+The final section provides brief discussion and direction for future research.
 
@@
 where $T$ is a pre-specified threshold.
-Then, assuming the full data follow a skew-$t$ distribution, we update values censored below the threshold using standard Bayesian missing data methods as described in Section \ref{s:comp}.
+Then, assuming the full data are observations from a skew-$t$ process, we update values censored below the threshold using standard Bayesian missing data methods as described in Section \ref{s:comp}.
 
 \subsection{Skew-$t$ process}\label{s:data}
-We assume the data can be modeled as skew-$t$.
-\citet{Zhang2010} show that the skew-$t$ can be written as the hierarchical model
+Many types of data demonstrate some level of skewness, and therefore should be modeled with distributions that allow for non-symmetry.
+The skew-elliptical family of distributions provides models that are mathematically tractable while introducing additional paremeters to account for non-symmetric data \citep{Genton2004}.
+One member of this family, is the skew-$t$ distribution \citep{Azzalini2003}.
+We assume the data can be modeled using a skew-$t$ distribution.
+\citet{Zhang2010} show that the skew-$t$ can be expressed as a hierarchical model
 \begin{align} \label{eq:fullmodel}
@@
 \end{align}
-where $\alpha \in \calR$ controls the skewness, $z_{t} \ind N_{(0, \infty)}(0, \sigma^2_t)$ are a random effect from a half-normal distribution (see appendix A.3), $v_t(\bs)$ is a Gaussian process with mean zero, variance one, and \Matern correlation, and $\sigma^2_t \iid \text{IG}(a, b)$.
+where $\alpha \in \calR$ controls the skewness, $z_{t} \ind N_{(0, \infty)}(0, \sigma^2_t)$ are a random effect from a half-normal distribution (see appendix A.3), $v_t(\bs)$ is a Gaussian process with mean zero, variance one, \Matern correlation, and $\sigma^2_t \iid \text{IG}(a, b)$.
 When marginalizing over the $z_t$ and $\sigma^2_t$ terms,
@@
 where $\mu$ is the location, $\Sigma^* = \frac{ b }{ a } \Sigma$, $\Sigma$ is a \Matern covariance matrix, and $\alpha \in \calR$ controls the skewness.
-The skew-$t$ process is desirable because of its flexible tail, controlled by both the skewness parameter $\alpha$ and the degrees of freedom $2a$, and its joint distribution is a multivariate skew-$t$ distribution.
+The skew-$t$ process is desirable because of its flexible tail that is controlled by the skewness parameter, $\alpha$, and the degrees of freedom, $2a$.
+Furthermore, if $\bY$ follows a multivariate skew-$t$ distribution, the marginal distributions also follow a skew-$t$ distribution \citep{Azzalini2003}.
 
@@
 
-Another popular measure of extremal dependence is the $\chi$ statistic.
-The $\chi$ coefficient for the upper tail is given by
+Another measure of extremal dependence is the $\chi$ statistic.
+The $\chi$ statistic for the upper tail is given by
 \begin{align*}
-  \chi = \lim_{c \rightarrow \infty} \Pr(Y(\bs_1) > c | Y(\bs_2) > c)
+  \chi = \lim_{c \rightarrow \infty} \Pr(Y(\bs_1) > c | Y(\bs_2) > c),
 \end{align*}
@@
 \subsection{Random daily partition}\label{s:part}
-One problem with the skew-$t$ distribution is that all sites are asymptotically dependent regardless of their spatial separation.
+One problem with the spatial skew-$t$ process is that all sites are asymptotically dependent regardless of their spatial separation.
 This occurs because all observations, both near and far, share the same $z_t$ and $\sigma^2_t$ terms.
@@
 \begin{align*}
-  P_{tk} = \{ \bs : k = \argmin_\ell || \bs - \bw_{t\ell} \}.
+  P_{tk} = \{ \bs : k = \argmin_\ell || \bs - \bw_{t\ell} || \}.
 \end{align*}
@@
 \end{align}
-A proof of this is given in Appendix A.2.
+A proof of this can be found in Appendix A.2.
 
@@
 However, we do not fix the partitions, they are treated as unknown and updated in the MCMC.
-Thus, standard geostatistical methods can be used to fit the model, and predictions can be made by Kriging at unobserved locations.
 We model this with a Bayesian hierarchical model as follows.
@@
    _t(\bs) v_t(\bs) \label{eq:hier}\\
-   z_t(\bs) = z_{tk} \text{ if } \bs \in P_{tk}\\
-   \sigma^2_{t}(\bs) = \sigma^2_{tk} \text{ if } \bs \in P_{tk}
+   z_t(\bs) &= z_{tk} \text{ if } \bs \in P_{tk}\\
+   \sigma^2_{t}(\bs) &= \sigma^2_{tk} \text{ if } \bs \in P_{tk}\\
    z_{tk} \mid \sigma^2_{tk} &\sim N_{(0, \infty)}(0, \sigma^2_{tk})\\
@@
 
-\subsection{Imputation}\label{s:impute}
-We can use Gibbs sampling to update $Y_t(\bs)$ for observations that are below $T$, the thresholded value. Given $\Theta$, $Y_t(\bs)$ has truncated normal full conditional with these parameter values.
-So we sample $Y_t(\bs) \sim N_{(-\infty, T)}(\mu(\bs), \Sigma)$
-
-\subsection{Parameter updates}\label{s:params}
-To update $\Theta$ given the current value of the complete data $\bY_1, \ldots, \bY_T$, we use a standard Gibbs updates for all parameters except for the knot locations which are done using a Metropolis update.
-See Appendix A.1 for details regarding Gibbs sampling.
-
-\subsection{Spatial prediction}\label{s:pred}
-Given $\bY_t$ the usual Kriging equations give the predictive distribution for $Y_t(\bs^*)$ at prediction location $(\bs^*)$
+We can use Gibbs sampling to update $Y_t(\bs)$ for censored observations that are below the threshold $T$.
+After conditioning on $\alpha$, $z_t(\bs)$ and non-censored observations, $Y_t(\bs)$ has truncated normal full conditionals.
+So we sample $Y_t(\bs) \sim N_{(-\infty, T)}(\mu(\bs), \Sigma)$.
+After imputing the censored observations, we update the model parameters.
+To update the model parameters, we use standard Gibbs updates for parameters when possible.
+In the case Gibbs sampling is not possible, parameters are updated using a random-walk Metropolis Hastings algorithm.
+See Appendix A.1 for details regarding Gibbs sampling and Metropolis update rules.
+The final step of the computation is to use Bayesian Kriging to generate a predictive distribution for $Y_t(\bs^*)$ at prediction location $(\bs^*)$.
+This step is similar to the imputation for censored observations except that the full conditionals are no longer truncated at $T$.
 
@@
   \item skew-$t$ marginal, $K=1$ knots, $T=-\infty$
-  \item skew-$t$ marginal, $K=1$ knots, $T=q(0.90)$
+  \item skew-$t$ marginal, $K=1$ knots, $T=q(0.80)$
   \item skew-$t$ marginal, $K=5$ knots, $T=-\infty$
-  \item skew-$t$ marginal, $K=5$ knots, $T=q(0.9)$
+  \item skew-$t$ marginal, $K=5$ knots, $T=q(0.80)$
 \end{enumerate}
-where $q(0.9)$ is the 90th sample quantile of the data.
-The design matrix $\bX$ includes on the intercept with a prior of $\beta \sim \text{N}(0, 10)$.
+where $q(0.80)$ is the 80th sample quantile of the data.
+The design matrix $\bX$ includes an the intercept with a prior of $\beta \sim \text{N}(0, 10)$.
 The spatial covariance parameters have priors $\log(\nu) \sim \text{N}(-1.2, 1)$, $\gamma \sim \text{U}(0, 1)$, $\log(\rho) \sim \text{N}(-2, 1)$.
@@
 For each site, we also have covariate information containing the estimated ozone from the Community Multi-scale Air Quality (CMAQ) modeling system.
+Initially, we fit a linear regression assuming that the mean function can be expressed as
+\begin{align}
+  \mu_t(\bs) = \beta_0 + \beta_1 \cdot \text{CMAQ}_t(\bs). \label{eq:datamean}
+\end{align}
+The data from 5 July are shown in Figure \ref{fig:ozone} along with a Q-Q plot of the residuals compared to a skew-$t$ distribution with 8 d.f. and $\alpha = 1$.
+\begin{center}
+\begin{figure}
+  \includegraphics[width=0.5\linewidth]{plots/ozone-5jul.pdf}
+  \includegraphics[width=0.5\linewidth]{plots/qq-res.pdf}
+  \caption{Ozone values on 5 July 2005 (left) Q-Q plot of the residuals (right)}
+  \label{fig:ozone}
+\end{figure}
+\end{center}
+The $\chi(h)$ and $\chi(t)$ plots indicate that the residuals still demonstrate spatial and temporal dependence (see Figure \ref{fig:chi-st})
+\begin{center}
+\begin{figure}
+  \includegraphics[width=0.5\linewidth]{plots/chi-h-ozone.pdf}
+  \includegraphics[width=0.5\linewidth]{plots/chi-t-ozone.pdf}
+  \caption{$\chi(h)$ plot for the residuals. The vertical line indicates distance after which observations no longer demonstrate dependence (left) $\chi(t)$ plot for the residuals (right)}
+  \label{fig:chi-st}
+\end{figure}
+\end{center}
+
+\subsection{Model comparisons}
 We fit the model using Gaussian and skew-$t$ marginal distributions, $K=1, 5, 10, 15$ partitions, with $Y(\bs)$ censored at $T = 0, 50, 75, 90$ ppb as described in Section \ref{s:data}.
 We also include a max-stable analysis using the method by ????
-All methods assume the location can be expressed as
-\begin{align}
-  \mu_t(\bs) = \beta_0 + \beta_1 \cdot \text{CMAQ}_t(\bs).
-\end{align}
-To explore the extremal dependence both over space and time, we plot $\chi(\bs)$ and $\chi(t)$.
+All methods assume the mean function given in (\ref{eq:datamean}).
 For each model, Brier scores and quantile scores were were averaged over all sites and days to obtain a single quantile score for each dataset.
 At a particular threshold or quantile level, the model that fits the best is the one with the lowest score.
+Looking at the heatplot in Figure \ref{fig:heatplot}, we can see that the methods using no thresholding tend to outperform those methods that do use thresholding.
+\hl{Brian, I'm not sure that this is the best way to show this.
+Maybe a better heatplot would be quantile scores for $T=0$ for a few different quantiles.
+Then you can get a sense how the partitions impact the scores.
+The main problem is that $K=10, T=q(0.90)$ is much larger (5 - 7 times) than the rest of the quantile scores (across all values), so it really has an impact on the color scale}
 
 \subsection{Results}\label{s:results}
+\begin{figure}
+  \includegraphics[width=0.5\linewidth]{plots/heatplot.pdf}
+  \includegraphics[width=0.5\linewidth]{plots/qscore-best.pdf}
+  \caption{Heatplot of quantile scores for the 99th quantile (left) Quantile scores of Gaussian, skew-$t$ (K=10, T=q(0)), and Max-stable (right)}
+  \label{fig:heatplot}
+\end{figure}
 

### 2014-09-15 pot/LaTeX/skewnormal.tex
more data analysis settings

@@
 Looking at the heatplot in Figure \ref{fig:heatplot}, we can see that the methods using no thresholding tend to outperform those methods that do use thresholding.
-\hl{Brian, I'm not sure that this is the best way to show this.
-Maybe a better heatplot would be quantile scores for $T=0$ for a few different quantiles.
-Then you can get a sense how the partitions impact the scores.
-The main problem is that $K=10, T=q(0.90)$ is much larger (5 - 7 times) than the rest of the quantile scores (across all values), so it really has an impact on the color scale}
 

### 2014-09-15 pot/code/R/mcmc.R

@@ -135,7 +135,7 @@ mcmc <- function(y, s, x, s.pred=NULL, x.pred=NULL,
   }
   
   zg <- matrix(0, ns, nt)
-  if (length(z.init) == 1) {
+  if (length(z.init) == 1 && skew) {
     cat("\t initializing all z terms to", z.init, "\n")
   }
   z <- matrix(z.init, nknots, nt)

### 2014-09-15 pot/code/analysis/ozone/US/cv5-21US.R

@@ -1,0 +1,53 @@
+options(warn=2)
+library(fields)
+library(geoR)
+library(mvtnorm)
+
+rm(list=ls())
+load('cv-setup-us.RData')
+source('../../../R/mcmc.R')
+source('../../../R/auxfunctions.R')
+
+setting <- 21
+method <- "t"
+nknots <- 5
+keep.knots <- F
+threshold <- 75
+thresh.quant <- F
+skew <- F
+outputfile <- paste("cv5-", setting, "US.RData", sep="")
+
+start <- proc.time()
+
+fit <- vector(mode="list", length=2)
+
+for(val in 1:2){
+	
+	set.seed(setting*100 + val)
+	
+	cat("CV", val, "started \n")
+	val.idx <- cv.lst[[val]]
+	y.o <- Y[-val.idx,]
+	X.o <- X[-val.idx, , ]
+	S.o <- S[-val.idx,]
+	
+	y.p <- Y[val.idx,]
+	X.p <- X[val.idx, , ]
+	S.p <- S[val.idx,]
+	
+	tic.set <- proc.time()
+	fit[[val]] <- mcmc(y=y.o, s=S.o, x=X.o, x.pred=X.p, s.pred=S.p,
+	                   method=method, skew=skew, keep.knots=keep.knots,
+	                   thresh.all=threshold, thresh.quant=thresh.quant, nknots=nknots, 
+                       iters=30000, burn=25000, update=500, iterplot=F,
+                       beta.init=beta.init, tau.init=tau.init, rho.init=1,
+                       nu.init=0.5, alpha.init=0.5)
+	toc.set <- proc.time()
+	time.set <- (toc.set - tic.set)[3]
+	
+	elap.time.val <- (proc.time() - start)[3]
+	avg.time.val <- elap.time.val / val
+	cat("CV", val, "finished", round(avg.time.val, 2), "per dataset \n")
+	save(fit, file=outputfile)
+}
+

### 2014-09-15 pot/code/analysis/ozone/US/cv5-26US.R

@@ -1,0 +1,53 @@
+options(warn=2)
+library(fields)
+library(geoR)
+library(mvtnorm)
+
+rm(list=ls())
+load('cv-setup-us.RData')
+source('../../../R/mcmc.R')
+source('../../../R/auxfunctions.R')
+
+setting <- 26
+method <- "t"
+nknots <- 15
+keep.knots <- F
+threshold <- 90
+thresh.quant <- F
+skew <- F
+outputfile <- paste("cv5-", setting, "US.RData", sep="")
+
+start <- proc.time()
+
+fit <- vector(mode="list", length=2)
+
+for(val in 1:2){
+	
+	set.seed(setting*100 + val)
+	
+	cat("CV", val, "started \n")
+	val.idx <- cv.lst[[val]]
+	y.o <- Y[-val.idx,]
+	X.o <- X[-val.idx, , ]
+	S.o <- S[-val.idx,]
+	
+	y.p <- Y[val.idx,]
+	X.p <- X[val.idx, , ]
+	S.p <- S[val.idx,]
+	
+	tic.set <- proc.time()
+	fit[[val]] <- mcmc(y=y.o, s=S.o, x=X.o, x.pred=X.p, s.pred=S.p,
+	                   method=method, skew=skew, keep.knots=keep.knots,
+	                   thresh.all=threshold, thresh.quant=thresh.quant, nknots=nknots, 
+                       iters=30000, burn=25000, update=500, iterplot=F,
+                       beta.init=beta.init, tau.init=tau.init, rho.init=1,
+                       nu.init=0.5, alpha.init=0.5)
+	toc.set <- proc.time()
+	time.set <- (toc.set - tic.set)[3]
+	
+	elap.time.val <- (proc.time() - start)[3]
+	avg.time.val <- elap.time.val / val
+	cat("CV", val, "finished", round(avg.time.val, 2), "per dataset \n")
+	save(fit, file=outputfile)
+}
+

### 2014-09-15 pot/LaTeX/skewnormal.tex
updates to manuscript

@@
 \end{align}
-where $\alpha \in \calR$ controls the skewness, $z_{t} \ind N_{(0, \infty)}(0, \sigma^2_t)$ are a random effect from a half-normal distribution (see appendix A.3), $v_t(\bs)$ is a Gaussian process with mean zero, variance one, \Matern correlation, and $\sigma^2_t \iid \text{IG}(a, b)$.
+where $\alpha \in \calR$ controls the skewness, $z_{t} \ind N_{(0, \infty)}(0, \sigma^2_t)$ are a random effect from a half-normal distribution (see appendix A.4), $v_t(\bs)$ is a Gaussian process with mean zero, variance one, \Matern correlation, and $\sigma^2_t \iid \text{IG}(a, b)$.
 When marginalizing over the $z_t$ and $\sigma^2_t$ terms,
@@
 \end{align}
-A proof of this can be found in Appendix A.2.
+A proof of this can be found in Appendix A.3.
 
@@
 Then, we update $\Theta$ using random walk MH or Gibbs sampling when appropriate.
-Finally, we make spatial precictions using conditional multivariate normal results and the fact that the distribution of $Y_t(\bs) \mid \Theta, z_{tl}$ is the usual multivariate normal distribution with a \Matern spatial covariance structure.
+Finally, we make spatial predictions using conditional multivariate normal results and the fact that the distribution of $Y_t(\bs) \mid \Theta, z_{tl}$ is the usual multivariate normal distribution with a \Matern spatial covariance structure.
 
@@
 In the case Gibbs sampling is not possible, parameters are updated using a random-walk Metropolis Hastings algorithm.
-See Appendix A.1 for details regarding Gibbs sampling and Metropolis update rules.
+See Appendices A.1 and A.2 for details regarding the MCMC.
 The final step of the computation is to use Bayesian Kriging to generate a predictive distribution for $Y_t(\bs^*)$ at prediction location $(\bs^*)$.
@@
   \item skew-$t$ marginal, $K=1$ knots, $T=-\infty$
-  \item skew-$t$ marginal, $K=1$ knots, $T=q(0.80)$
+  \item $t$ marginal, $K=1$ knots, $T=q(0.80)$
   \item skew-$t$ marginal, $K=5$ knots, $T=-\infty$
-  \item skew-$t$ marginal, $K=5$ knots, $T=q(0.80)$
+  \item $t$ marginal, $K=5$ knots, $T=q(0.80)$
 \end{enumerate}
@@
 
-\section*{Appendix A.1: Posterior distributions}
+\section*{Appendix A.1: MCMC details}
+The MCMC sampling for the model \ref{s:hier} is done using {\tt R} (http://www.r-project.org).
+Whenever possible, we select conjugate priors (see Appendix A.2); however, for some of the parameters, no conjugate prior distributions exist.
+When no conjugate prior distribution exists, we use a random walk Metropolis Hastings update step.
+In each Metropolis Hastings update, we tune the algorithm to give acceptance rates near 0.40.d
+
+\subsection*{Spatial knot locations}
+For each day, we update the spatial knot locations, $\bw_1, \ldots, \bw_K$, using a Metropolis Hastings block update.
+First, we generate candidate knots locations using a random walk Gaussian candidate distribution
+\begin{align*}
+
+\end{align*}
+
+\subsection*{Spatial covariance parameters}
+We update the three spatial covariance parameters, $\log(\rho)$, $\log(\nu)$, $\gamma$, using a Metropolis Hastings block update step.
+First, we generate a candidate using a random walk Gaussian candidate distribution 
+\begin{align*}
+	\log(\rho)^{(c)} \sim \text{N}(\log(\rho)^{(r - 1)}, s^2)
+\end{align*}
+where $\log(\rho)^{(r-1)}$ is the value at MCMC iteration $r - 1$, and $s$ is a tuning parameter.
+Candidates are generated for $\log(\nu)$ and $\gamma$ in a similar fashion.
+The acceptance ratio is 
+\begin{align*}
+	R = \left\{ \frac{ \prod_{t = 1}^{T} l[Y_t(\bs | \rho^{(c)}, \nu^{(c)}, \gamma^{(c)}, \ldots] }{\prod_{t = 1}^{T} l[Y_t(\bs | \rho^{(r-1)}, \nu^{(r-1)}, \gamma^{(r-1)}, \ldots] } \right\} \times \left\{ \frac{ p[\rho^{(c)}] }{ p[\rho^{(r - 1)] } } \right\} \times \left\{ \frac{ p[\nu^{(c)}] }{ p[\nu^{(r - 1)}] } \right\} \times \left\{ \frac{ p[ \gamma^{(c)} ] }{ p[\gamma^{(r - 1)} ] } \right\}
+\end{align*}
+where $l$ is the likelihood given in \ref{eq:hier}.
+All three candidates are accepted with probability $\min\{R, 1\}$.
+
+\section*{Appendix A.2: Posterior distributions}
 \input{zpost.tex}
@@
 
-\section*{Appendix A.2: Proof that $\lim_{h \rightarrow \infty} \pi(h) = 0$}
+\section*{Appendix A.3: Proof that $\lim_{h \rightarrow \infty} \pi(h) = 0$}
 \input{proof_dist.tex}
 
-\section*{Appendix A.3: Half-normal distribution}
+\section*{Appendix A.4: Half-normal distribution}
 Let $u = |z|$ where $Z \sim N(\mu, \sigma^2)$.

### 2014-09-17 pot/LaTeX/skewnormal.tex
manuscript updates

@@
 Then within each partition, $Y_t(\bs)$ follows the distribution given in (\ref{eq:fullmodel}).
-When incorporating the random daily partition, the $chi$ statistic becomes
+When incorporating the random daily partition, the $\chi$ statistic becomes
 \begin{align}
@@
 In the first design, $\sigma^2 = 2$ is used for all days.
-For designs 2 -- 4, $sigma^2_{tk} \iid \text{IG}(3, 8)$
+For designs 2 -- 4, $\sigma^2_{tk} \iid \text{IG}(3, 8)$
 For designs 1 -- 3, we set $\alpha = 0$.
@@
 \end{align}
-The data from 5 July are shown in Figure \ref{fig:ozone} along with a Q-Q plot of the residuals compared to a skew-$t$ distribution with 8 d.f. and $\alpha = 1$.
+The data from 10 July are shown in Figure \ref{fig:ozone} along with a Q-Q plot of the residuals compared to a skew-$t$ distribution with 8 d.f. and $\alpha = 1$.
 \begin{center}
 \begin{figure}
-  \includegraphics[width=0.5\linewidth]{plots/ozone-5jul.pdf}
+  \includegraphics[width=0.5\linewidth]{plots/ozone-10jul.pdf}
   \includegraphics[width=0.5\linewidth]{plots/qq-res.pdf}
-  \caption{Ozone values on 5 July 2005 (left) Q-Q plot of the residuals (right)}
+  \caption{Ozone values on 10 July 2005 (left) Q-Q plot of the residuals (right)}
   \label{fig:ozone}
@@
 When no conjugate prior distribution exists, we use a random walk Metropolis Hastings update step.
-In each Metropolis Hastings update, we tune the algorithm to give acceptance rates near 0.40.d
+In each Metropolis Hastings update, we tune the algorithm to give acceptance rates near 0.40.
 
@@
 For each day, we update the spatial knot locations, $\bw_1, \ldots, \bw_K$, using a Metropolis Hastings block update.
-First, we generate candidate knots locations using a random walk Gaussian candidate distribution
+Let $\Phi_2$ represent a bivariate normal density function, and let $\bw^*_k = h[\Phi_2^{-1}(\bs_k)]$ where $h(\cdot)$ is a transformation to shift and scale the transformed knots so that $\bs^*_k \in \calD$.
+Because the spatial domain is bounded, we generate candidate knots using $\bw^*_1, \ldots, \bw^*_K$ and a random walk bivariate Gaussian candidate distribution
 \begin{align*}
-
+	{\bw^*_k}^{(c)} \sim \text{N}({\bw^*_k}^{(r - 1)}, s^2 I_2)
 \end{align*}
+where $\bw*_k^{(r - 1)}$ is the location for the transformed knot at MCMC iteration $r - 1$, $s$ is a tuning parameter, and $I_2$ is an identity matrix.
+After candidates have been generated for all $K$ knots, the acceptance ratio is
+\begin{align*}
+  R = \left\{ \frac{ l[ Y_t(\bs | \bw_1^{(c)}, \ldots, \bw_K^{(c)}, \ldots)] }{l[ Y_t(\bs | \bw_1^{(r - 1)}, \ldots, \bw_K^{(r - 1)}, \ldots)]} \right\} \times \left\{ \frac{ \prod_{k = 1}^{K}\phi(\bw_k^{(c)})}{ \prod_{k = 1}^{K}\phi(\bw_k^{(r - 1)})} \right\}
+\end{align*}
+where $l$ is the likelihood given in (\ref{eq:hier}).
+The candidate knots are accepted with probability $\min\{R, 1\}$.
 
@@
 \begin{align*}
-	R = \left\{ \frac{ \prod_{t = 1}^{T} l[Y_t(\bs | \rho^{(c)}, \nu^{(c)}, \gamma^{(c)}, \ldots] }{\prod_{t = 1}^{T} l[Y_t(\bs | \rho^{(r-1)}, \nu^{(r-1)}, \gamma^{(r-1)}, \ldots] } \right\} \times \left\{ \frac{ p[\rho^{(c)}] }{ p[\rho^{(r - 1)] } } \right\} \times \left\{ \frac{ p[\nu^{(c)}] }{ p[\nu^{(r - 1)}] } \right\} \times \left\{ \frac{ p[ \gamma^{(c)} ] }{ p[\gamma^{(r - 1)} ] } \right\}
+	R = \left\{ \frac{ \prod_{t = 1}^{T} l[Y_t(\bs | \rho^{(c)}, \nu^{(c)}, \gamma^{(c)}, \ldots] }{\prod_{t = 1}^{T} l[Y_t(\bs | \rho^{(r-1)}, \nu^{(r-1)}, \gamma^{(r-1)}, \ldots] } \right\} \times \left\{ \frac{ p[\rho^{(c)}] }{ p[\rho^{(r - 1)] } } \right\} \times \left\{ \frac{ p[\nu^{(c)}] }{ p[\nu^{(r - 1)}] } \right\} \times \left\{ \frac{ p[ \gamma^{(c)} ] }{ p[\gamma^{(r - 1)} ] } \right\}.
 \end{align*}
-where $l$ is the likelihood given in \ref{eq:hier}.
 All three candidates are accepted with probability $\min\{R, 1\}$.

### 2014-09-19 pot/LaTeX/skewnormal.tex
Manuscript updates

@@
 \subsection{Hierarchical model}\label{s:hier}
-Conditioned on $z_{tk}(\bs) \iid $ $N(0, \sigma^2_{tk})$, $\sigma^2_{tk}(\bs) \sim IG(a, b)$, and $P_tk$, the marginal distributions are skew-$t$ and the joint distribution within a partition is multivariate skew-$t$.
+Conditioned on $z_{tk}(\bs) \iid $ $N(0, \sigma^2_{tk})$, $\sigma^2_{tk}(\bs) \sim IG(a, b)$, and $P_{tk}$, the marginal distributions are skew-$t$ and the joint distribution within a partition is multivariate skew-$t$.
 However, we do not fix the partitions, they are treated as unknown and updated in the MCMC.
@@
 \begin{align}
-   Y_t(\bs) \mid \Theta, z_{t}(\bs) &= X_t(\bs) \beta + \alpha z_t(\bs) + \sigma
-   _t(\bs) v_t(\bs) \label{eq:hier}\\
+   Y_t(\bs) \mid z_{t}(\bs), \sigma^2(\bs), P_{tk}, \alpha, \beta, \Theta &= X_t(\bs) \beta + \alpha z_t(\bs) + \sigma_t(\bs) v_t(\bs) \label{eq:hier}\\
    z_t(\bs) &= z_{tk} \text{ if } \bs \in P_{tk}\\
    \sigma^2_{t}(\bs) &= \sigma^2_{tk} \text{ if } \bs \in P_{tk}\\
+   \alpha &\sim N(0, 10)\\
+   v_t(\bs) \mid \Theta &\sim \Matern(0, \Sigma)\\
    z_{tk} \mid \sigma^2_{tk} &\sim N_{(0, \infty)}(0, \sigma^2_{tk})\\
    \sigma^2_{tk} &\iid IG(\alpha, \beta)\\   
-   v_t(\bs) \mid \Theta &\sim \Matern(0, \Sigma)\\
-   \alpha &\sim N(0, 10)\\
    w_{tk} &\sim Unif(\calD)
 \end{align}
-where $\Theta = \{w_{t1}, \ldots, w_{tK}, \beta, \sigma_t, \alpha, \lambda, \rho, \nu \}$; \mbox{$k = \argmin_\ell ||\bs - \bw_\ell ||$}; and $\Sigma$ is a \Matern covariance matrix with variance one, spatial range $\rho$, smoothness $\nu$.
+where $\Theta = \{\rho, \nu, \gamma\}$; \mbox{$k = \argmin_\ell ||\bs - \bw_\ell ||$}; and $\Sigma$ is a \Matern covariance matrix with variance one, spatial range $\rho$, smoothness $\nu$, and the proportion of variance accounted for by the spatial variation is $\gamma$.
 

### 2014-09-25 pot/LaTeX/skewnormal.tex
manuscript updates and tweaks to results files

@@
 
+\subsection{Accounting for temporal dependence}
+We account for temporal dependence through a AR(1) time series on the spatial knots and $z_t(\bs)$ terms.
+These are modeled as
+\begin{align}
+  \bw_{t, k} | \bw_{t-1, k} &\sim N_2(\phi_w \bw_{t-1, k}, (1 - \phi_w^2) ) \\
+  z_{t, k} | z_{t-1, k} &\sim N(\phi_z z_{t-1, k}, (1 - \phi_z^2))
+\end{align}
+where $|\phi_w|, |\phi_z| < 1$.
+
 \subsection{Hierarchical model}\label{s:hier}
@@
 We model this with a Bayesian hierarchical model as follows.
-Let $w_{t1}, \ldots, w_{tK}$ be a set of daily spatial knots in a spatial domain of interest, $\calD$, so that
+Let $\bw_{t1}, \ldots, \bw_{tK}$ be a set of daily spatial knots in a spatial domain of interest, $\calD$, so that
 \begin{align*}
- P_{tk} = \{\bs : k = \argmin_\ell || \bs_t - w_{t\ell} ||\}.
+ P_{tk} = \{\bs : k = \argmin_\ell || \bs_t - \bw_{t\ell} ||\}.
 \end{align*}
+Because the spatial domain is bounded, we do a transformation on the knot locations to $\calR^2$.
+Let $\Phi_2$ represent a bivariate standard normal density function, and let $\bw_k = \min(\bs) + \text{range}(\bs) * \Phi_2^{-1}(\bw^*_k)$.
+Then $\bw^*_k \in \calR^2$.
 Then
 \begin{align}
-   Y_t(\bs) \mid z_{t}(\bs), \sigma^2(\bs), P_{tk}, \alpha, \beta, \Theta &= X_t(\bs) \beta + \alpha z_t(\bs) + \sigma_t(\bs) v_t(\bs) \label{eq:hier}\\
+   Y_t(\bs) \mid z_{t}(\bs), \sigma^2(\bs), P_{tk}, \alpha, \beta, \Theta &= X_t(\bs) \beta + \alpha |z_t(\bs)| + \sigma_t(\bs) v_t(\bs) \label{eq:hier}\\
    z_t(\bs) &= z_{tk} \text{ if } \bs \in P_{tk}\\
@@
    v_t(\bs) \mid \Theta &\sim \Matern(0, \Sigma)\\
-   z_{tk} \mid \sigma^2_{tk} &\sim N_{(0, \infty)}(0, \sigma^2_{tk})\\
+   z_{tk} \mid z_{t-1, k}, \sigma^2_{tk} &\sim N(\phi_z z_{t-1, k}, \sigma^2_{tk} (1 - \phi_z^2))\\
    \sigma^2_{tk} &\iid IG(\alpha, \beta)\\
-   w_{tk} &\sim Unif(\calD)
+   \bw^*_{tk} \mid \bw^*_{t-1, k} &\sim N_2(\phi_w \bw^*_{t-1, k}, (1 - \phi_z^2))
 \end{align}
@@
 \subsection{Results}\label{s:simresults}
+When comparing quantile scores, in almost all of the data settings, the biggest gains are seen when using thresholded methods.
+In all data settings except for the max-stable setting, method 3 performs the best.
+However, in most of the data settings, method 5 also performs reasonably well.
+In fact, for setting 5, paired $t$-tests suggest that the difference in performance is not significant for any quantile score.
+Furthermore, paired $t$-tests results indicate that the difference in quantile scores between methods 3 and 5 is only statistically significant after q(0.97) for setting 4 and after q(0.93) for setting 1.
+
+\begin{figure}
+  \includegraphics[width=\linewidth]{plots/quantileplots.pdf}
+  \caption{Quantile score plots for simulation study results}
+  \label{fig:simquantscores}
+\end{figure}
 
@@
 For each day, we update the spatial knot locations, $\bw_1, \ldots, \bw_K$, using a Metropolis Hastings block update.
-Let $\Phi_2$ represent a bivariate normal density function, and let $\bw^*_k = h[\Phi_2^{-1}(\bs_k)]$ where $h(\cdot)$ is a transformation to shift and scale the transformed knots so that $\bs^*_k \in \calD$.
-Because the spatial domain is bounded, we generate candidate knots using $\bw^*_1, \ldots, \bw^*_K$ and a random walk bivariate Gaussian candidate distribution
+Because the spatial domain is bounded, we generate candidate knots using the transformed knots $\bw^*_1, \ldots, \bw^*_K$ (see section \ref{s:hier}) and a random walk bivariate Gaussian candidate distribution
 \begin{align*}
@@
 
+\subsection*{Spatial random effects}
+If there is temporal dependence amongst the observations, then we update $z_tk$, using a Metropolis Hastings update.
+First, we generate a candidate using a random walk Gaussian candidate distribution
+\begin{align*}
+  z_{tk}^{(c)} \sim \text{N}(z_{tk}^{(r - 1)}, s^2)
+\end{align*}
+where $z_{tk}^{(r-1)}$ is the value at MCMC iteration $r - 1$, and $s$ is a tuning parameter.
+The acceptance ratio is
+\begin{align*}
+  R = \left\{ \frac{ l[Y_t(\bs) | z_{tk}^{(c)}, \ldots] }{ l[Y_t(\bs) | z_{tk}^{(r - 1)}]} \right\} \times \left\{ \frac{ p[ z_{tk}^{(c)} ] }{ p[ z_{tk}^{(r - 1)}]}\right\}.
+\end{align*}
+The candidate is accepted with probability $\min\{R, 1\}$
+
 \subsection*{Spatial covariance parameters}
@@
 \begin{align*}
-	R = \left\{ \frac{ \prod_{t = 1}^{T} l[Y_t(\bs | \rho^{(c)}, \nu^{(c)}, \gamma^{(c)}, \ldots] }{\prod_{t = 1}^{T} l[Y_t(\bs | \rho^{(r-1)}, \nu^{(r-1)}, \gamma^{(r-1)}, \ldots] } \right\} \times \left\{ \frac{ p[\rho^{(c)}] }{ p[\rho^{(r - 1)] } } \right\} \times \left\{ \frac{ p[\nu^{(c)}] }{ p[\nu^{(r - 1)}] } \right\} \times \left\{ \frac{ p[ \gamma^{(c)} ] }{ p[\gamma^{(r - 1)} ] } \right\}.
+	R = \left\{ \frac{ \prod_{t = 1}^{T} l[Y_t(\bs) | \rho^{(c)}, \nu^{(c)}, \gamma^{(c)}, \ldots] }{\prod_{t = 1}^{T} l[Y_t(\bs) | \rho^{(r-1)}, \nu^{(r-1)}, \gamma^{(r-1)}, \ldots] } \right\} \times \left\{ \frac{ p[\rho^{(c)}] }{ p[\rho^{(r - 1)] } } \right\} \times \left\{ \frac{ p[\nu^{(c)}] }{ p[\nu^{(r - 1)}] } \right\} \times \left\{ \frac{ p[ \gamma^{(c)} ] }{ p[\gamma^{(r - 1)} ] } \right\}.
 \end{align*}

### 2014-09-25 pot/code/analysis/ozone/US/cv-results-us.R

@@ -14,7 +14,7 @@ source("../../../R/auxfunctions.R")
 # # abline(0,1)
 # # print(mean(y.full.p>lo & y.full.p<hi))
 
-probs <- c(0.9, 0.91, 0.92, 0.93, 0.94, 0.95, 0.96, 0.97, 0.98, 0.99, 0.995, 0.999)
+probs <- c(0.9, 0.91, 0.92, 0.93, 0.94, 0.95, 0.96, 0.97, 0.98, 0.99, 0.995)
 thresholds <- quantile(Y, probs=probs, na.rm=T)
 nsets <- 2 # Number of cv sets
 nbetas <- 2 # number of betas
@@ -25,7 +25,7 @@ brier.score <- array(NA, dim=c(length(thresholds), nsets, 26))
 beta.0 <- array(NA, dim=c(5000, nsets, 26))
 beta.1 <- array(NA, dim=c(5000, nsets, 26)) 
 
-done <- c(1:25)
+done <- c(1:26)
 
 for (i in 1:26) {
   file <- paste("cv5-", i, "US.RData", sep="")
@@ -86,10 +86,10 @@ for (i in 1:26) {
   }
 }
 
-quant.score.mean[c(1:10, 13, 14, 17:19, 21, 23, 25), ]
+quant.score.mean[c(1:10, 13, 14, 17:19, 21, 23, 25, 26), ]
 for (i in 1:12) {
- print(which(quant.score.mean[c(1:25), i] == min(quant.score.mean[c(1:25), i])))
- print(which(brier.score.mean[c(1:25), i] == min(brier.score.mean[c(1:25), i])))
+ print(which(quant.score.mean[c(1:26), i] == min(quant.score.mean[c(1:26), i])))
+ print(which(brier.score.mean[c(1:26), i] == min(brier.score.mean[c(1:26), i])))
 }
 
 library(fields)

### 2014-10-10 pot/LaTeX/skewnormal.tex
minor changes to manuscript

@@
 \begin{align} \label{eq:fullmodel}
-  Y_t(\bs) &= X_t(\bs) \beta + \alpha z_t + \sigma_t v_t(\bs)
+  Y_t(\bs) &= X_t(\bs) \beta + \alpha \sigma_t z_t + \sigma_t v_t(\bs)
 \end{align}
-where $\alpha \in \calR$ controls the skewness, $z_{t} \ind N_{(0, \infty)}(0, \sigma^2_t)$ are a random effect from a half-normal distribution (see appendix A.4), $v_t(\bs)$ is a Gaussian process with mean zero, variance one, \Matern correlation, and $\sigma^2_t \iid \text{IG}(a, b)$.
+where $\alpha \in \calR$ controls the skewness, $z_{t} \ind N_{(0, \infty)}(0, 1)$ are a random effect from a half-normal distribution (see appendix A.4), $v_t(\bs)$ is a Gaussian process with mean zero, variance one, \Matern correlation, and $\sigma^2_t \iid \text{IG}(a, b)$.
 When marginalizing over the $z_t$ and $\sigma^2_t$ terms,
@@
 
+\section*{Appendix A.5: Multivariate skew-$t$ distribution}
+\input{mvskewt.tex}
+
 \bibliographystyle{rss}

### 2014-10-16 pot/LaTeX/mvskewt.tex

Let $Z_{0t} \sim \text{SN}_d(0, \bar{\Omega}, \alpha)$ be a $d$-dimensional skew-normal random variable.
Note, the relationship between $\bar{\Omega}$ and the \Matern covariance matrix given for (\ref{eq:fullmodel}) is given as follows
\begin{align*}
  \alpha &= \frac{ \delta }{ \sqrt{1 - \delta^2} } \\
  D_\delta &= (diag(1 - \delta))^{1/2}\\
  \bar{\Omega} &= D_\delta (\Sigma + \lambda \lambda^T) D_\delta \\
  \lambda &= D_\delta^{-1} \delta\\
\end{align*}
Let $\sigma_{tk}^2 \sim \chi^2_\nu / \nu$ where $\nu$ is the degrees of freedom.
Let $Z_{tk} = \sigma_{tk}^{-1} Z_{0, tk}$.
After marginalizing out the $\sigma_{tk}^2$ terms, $Z$ follows a multivariate skew-$t$ distribution, and the density is given by
\begin{align*}
  f_z(z) = 2 t_d(z; \bar{\Omega}, \nu) T \left( \alpha^T z \sqrt{ \frac{\nu + d}{\nu + Q(z)} }; \nu + d\right)
\end{align*}
where $Q(z) = z^T \bar{\Omega}^{-1}z$ and $T(\cdot; \rho)$ denotes the univariate Student's $t$ distribution function on $\rho$ degrees of freedom \citep{Azzalini2013}.

### 2014-10-20 pot/LaTeX/skewnormal.tex
time series updates

@@
 \begin{align} \label{eq:fullmodel}
-  Y_t(\bs) &= X_t(\bs) \beta + \alpha \sigma_t z_t + \sigma_t v_t(\bs)
+  Y_t(\bs) &= X_t(\bs) \beta + \alpha |z_t| + \sigma_t v_t(\bs)
 \end{align}
-where $\alpha \in \calR$ controls the skewness, $z_{t} \ind N_{(0, \infty)}(0, 1)$ are a random effect from a half-normal distribution (see appendix A.4), $v_t(\bs)$ is a Gaussian process with mean zero, variance one, \Matern correlation, and $\sigma^2_t \iid \text{IG}(a, b)$.
+where $\alpha \in \calR$ controls the skewness, $z_{t} \ind N(0, \sigma^2_t)$ are a random effect from a half-normal distribution (see appendix A.4), $v_t(\bs)$ is a Gaussian process with mean zero, variance one, \Matern correlation, and $\sigma^2_t \iid \text{IG}(a, b)$.
 When marginalizing over the $z_t$ and $\sigma^2_t$ terms,
@@
 This occurs because all observations, both near and far, share the same $z_t$ and $\sigma^2_t$ terms.
-We handle this problem with a daily random partition similar to \citet{Huser2014} that allows $z_t$ and $\sigma^2_t$ to vary by site.
+We handle this problem with a daily random partition similar to \citet{Kim2005} that allows $z_t$ and $\sigma^2_t$ to vary by site.
 The model then becomes
 \begin{align}
-  Y_t(\bs) &= X_t(\bs) \beta + \alpha z_t(\bs) + \sigma_t(\bs) v_t(\bs).
+  Y_t(\bs) &= X_t(\bs) \beta + \alpha |z_t(\bs)| + \sigma_t(\bs) v_t(\bs).
 \end{align}
@@
 
-\subsection{Accounting for temporal dependence}
-We account for temporal dependence through a AR(1) time series on the spatial knots and $z_t(\bs)$ terms.
-These are modeled as
+\subsection{Accounting for temporal dependence} \label{s:temporal}
+In a standard block-maxima approach, the temporal dependence is assumed to be negligible because blocks (e.g. yearly maxima) are taken to be large enough to assume independence.
+However, when using daily meausrements, the assumption of temporal independence is no longer appropriate.
+We account for temporal dependence through an AR(1) time series on the spatial knots, $z_t(\bs)$ terms, and $\sigma^2_t(\bs)$ terms.
+We first, transform the spatial knots from $\calD$ to $\calR^2$ as follows.
+Let
+\begin{align*}
+  \bw^*_k = \Phi^{-1}_2\left[ \frac{ (\bw_k - \min(\bs))}{ \text{range}(\bs) } \right].
+\end{align*}
+where $\Phi_2$ represents a bivariate standard normal density function.
+Then $\bw^*_k \in \calR^2$.
+We use a copula on the $\sigma^2_t(\bs)$ to ensure that the marginal distributions of $\sigma^2_t(\bs)$ are inverse gamma.
+Let
+\begin{align*}
+  \sigma^{2*}_t(\bs) =\Phi^{-1}\left\{ \text{IG}[\sigma^2_t(\bs)] \right\}
+\end{align*}
+where $\Phi$ is a univariate standard normal density function, and IG is the distribution function for an IG$(a, b)$ random variable.
+Then the time series is modeled as
 \begin{align}
-  \bw_{t, k} | \bw_{t-1, k} &\sim N_2(\phi_w \bw_{t-1, k}, (1 - \phi_w^2) ) \\
-  z_{t, k} | z_{t-1, k} &\sim N(\phi_z z_{t-1, k}, (1 - \phi_z^2))
+  \bw^*_{t, k} | \bw^*_{t-1, k} &\sim N_2\left[\phi_w \bw^*_{t-1, k}, (1 - \phi_w^2) \right] \\
+  z_{t, k} | z_{t-1, k} &\sim N \left[\phi_z z_{t-1, k}, \sigma^2_{t,k} (1 - \phi_z^2)\right] \\
+  \sigma^{2*}_{t, k} | \sigma^{2*}_{t-1, k} &\sim N \left[\phi_\sigma \sigma^{2*}_{t-1, k}, (1 - \phi_\sigma^2) \right]
 \end{align}
-where $|\phi_w|, |\phi_z| < 1$.
+where $|\phi_w| < 1$, $0 < \phi_z < 1$, and $|\phi_\sigma| < 1$.
 
 \subsection{Hierarchical model}\label{s:hier}
-Conditioned on $z_{tk}(\bs) \iid $ $N(0, \sigma^2_{tk})$, $\sigma^2_{tk}(\bs) \sim IG(a, b)$, and $P_{tk}$, the marginal distributions are skew-$t$ and the joint distribution within a partition is multivariate skew-$t$.
+Conditioned on $z_{t,k}(\bs) \iid $ $N(0, \sigma^2_{t,k})$, $\sigma^2_{t,k}(\bs) \sim IG(a, b)$, and $P_{t,k}$, the marginal distributions are skew-$t$ and the joint distribution within a partition is multivariate skew-$t$.
 However, we do not fix the partitions, they are treated as unknown and updated in the MCMC.
 We model this with a Bayesian hierarchical model as follows.
-Let $\bw_{t1}, \ldots, \bw_{tK}$ be a set of daily spatial knots in a spatial domain of interest, $\calD$, so that
+Let $\bw_{t,1}, \ldots, \bw_{t,K}$ be a set of daily spatial knots in a spatial domain of interest, $\calD$, so that
 \begin{align*}
- P_{tk} = \{\bs : k = \argmin_\ell || \bs_t - \bw_{t\ell} ||\}.
+ P_{t,k} = \{\bs : k = \argmin_\ell || \bs_t - \bw_{t\ell} ||\}.
 \end{align*}
-Because the spatial domain is bounded, we do a transformation on the knot locations to $\calR^2$.
-Let $\Phi_2$ represent a bivariate standard normal density function, and let $\bw_k = \min(\bs) + \text{range}(\bs) * \Phi_2^{-1}(\bw^*_k)$.
-Then $\bw^*_k \in \calR^2$.
+
 Then
 \begin{align}
-   Y_t(\bs) \mid z_{t}(\bs), \sigma^2(\bs), P_{tk}, \alpha, \beta, \Theta &= X_t(\bs) \beta + \alpha |z_t(\bs)| + \sigma_t(\bs) v_t(\bs) \label{eq:hier}\\
-   z_t(\bs) &= z_{tk} \text{ if } \bs \in P_{tk}\\
-   \sigma^2_{t}(\bs) &= \sigma^2_{tk} \text{ if } \bs \in P_{tk}\\
+   Y_t(\bs) \mid z_{t}(\bs), \sigma^2(\bs), P_{t,k}, \alpha, \beta, \Theta &= X_t(\bs) \beta + \alpha |z_t(\bs)| + \sigma_t(\bs) v_t(\bs) \label{eq:hier}\\
+   z_t(\bs) &= z_{t,k} \text{ if } \bs \in P_{t,k}\\
+   \sigma^2_{t}(\bs) &= \sigma^2_{t,k} \text{ if } \bs \in P_{t,k}\\
    \alpha &\sim N(0, 10)\\
    v_t(\bs) \mid \Theta &\sim \Matern(0, \Sigma)\\
-   z_{tk} \mid z_{t-1, k}, \sigma^2_{tk} &\sim N(\phi_z z_{t-1, k}, \sigma^2_{tk} (1 - \phi_z^2))\\
-   \sigma^2_{tk} &\iid IG(\alpha, \beta)\\
-   \bw^*_{tk} \mid \bw^*_{t-1, k} &\sim N_2(\phi_w \bw^*_{t-1, k}, (1 - \phi_z^2))
+   z_{t,k} \mid z_{t-1, k}, \sigma^2_{tk} &\sim N(\phi_z z_{t-1, k}, \sigma^2_{t,k} (1 - \phi_z^2))\\
+   \sigma^{2*}_{t,k} \mid \sigma^{2*}_{t-1, k} &\sim N(\phi_\sigma \sigma^{2*}_{t-1, k}, (1 - \phi_\sigma^2))\\
+   \bw^*_{t,k} \mid \bw^*_{t-1, k} &\sim N_2(\phi_w \bw^*_{t-1, k}, (1 - \phi_w^2))
 \end{align}
@@
 The Brier score for predicting exceedance of a threshold $c$ is given by $[e(c) - P(c)]^2$ where $e(c) = I[y>c]$ is an indicator function indicating that a test set value, $y$, has exceeded the threshold, $c$, and $P(c)$ is the predicted probability of exceeding $c$.
-The quantile score for the $\tau$th quantile is $2\{ I[y < \widehat{q}(\tau)] - \tau\} (\widehat{q}(\tau) - y)$ where $y$ is a AQS test set value and $\widehat{q}(\tau)$ is the estimated $\tau$th quantile.
+The quantile score for the $\tau$th quantile is $2\{ I[y < \widehat{q}(\tau)] - \tau\} (\widehat{q}(\tau) - y)$ where $y$ is a test set value and $\widehat{q}(\tau)$ is the estimated $\tau$th quantile.
 For both the Brier score and the quantile score, a lower score indicates a better fit.
@@
 \subsection*{Spatial random effects}
-If there is temporal dependence amongst the observations, then we update $z_tk$, using a Metropolis Hastings update.
+If there is temporal dependence amongst the observations, then we update $z_{tk}$, using a Metropolis Hastings update.
 First, we generate a candidate using a random walk Gaussian candidate distribution

### 2014-10-20 pot/code/analysis/ozone/US/cv5-27US.R

@@ -59,8 +59,8 @@ source('../../../R/auxfunctions.R')
 fit[[val]] <- mcmc(y=y.o, s=S.o, x=X.o, x.pred=X.p, s.pred=S.p,
                    temporalw=T, temporalz=T, temporaltau=T,
                    method=method, skew=T, keep.knots=keep.knots,
-	               thresh.all=0, thresh.quant=thresh.quant, nknots=nknots,
-                   iters=10000, burn=5000, update=10, iterplot=T,
+	               thresh.all=0, thresh.quant=thresh.quant, nknots=5,
+                   iters=25000, burn=20000, update=100, iterplot=T,
                    beta.init=beta.init, tau.init=0.05, rho.init=1,
                    nu.init=0.5, alpha.init=0.5)
 # Rprof(NULL)

### 2014-10-21 pot/LaTeX/skewnormal.tex
time series tweaks

@@
 where $\Phi$ is a univariate standard normal density function, and IG is the distribution function for an IG$(a, b)$ random variable.
-Then the time series is modeled as
+Then, when $t > 1$ the time series is modeled as
 \begin{align}
@@
 where $|\phi_w| < 1$, $0 < \phi_z < 1$, and $|\phi_\sigma| < 1$.
+All three of the time series are initialized using $N(0, 1)$.
 
@@
 For each day, we update the spatial knot locations, $\bw_1, \ldots, \bw_K$, using a Metropolis Hastings block update.
-Because the spatial domain is bounded, we generate candidate knots using the transformed knots $\bw^*_1, \ldots, \bw^*_K$ (see section \ref{s:hier}) and a random walk bivariate Gaussian candidate distribution
+Because the spatial domain is bounded, we generate candidate knots using the transformed knots $\bw^*_1, \ldots, \bw^*_K$ (see section \ref{s:temporal}) and a random walk bivariate Gaussian candidate distribution
 \begin{align*}
@@
 \end{align*}
-where $\bw*_k^{(r - 1)}$ is the location for the transformed knot at MCMC iteration $r - 1$, $s$ is a tuning parameter, and $I_2$ is an identity matrix.
+where ${\bw^*_k}^{(r - 1)}$ is the location for the transformed knot at MCMC iteration $r - 1$, $s$ is a tuning parameter, and $I_2$ is an identity matrix.
 After candidates have been generated for all $K$ knots, the acceptance ratio is
 \begin{align*}
-  R = \left\{ \frac{ l[ Y_t(\bs | \bw_1^{(c)}, \ldots, \bw_K^{(c)}, \ldots)] }{l[ Y_t(\bs | \bw_1^{(r - 1)}, \ldots, \bw_K^{(r - 1)}, \ldots)]} \right\} \times \left\{ \frac{ \prod_{k = 1}^{K}\phi(\bw_k^{(c)})}{ \prod_{k = 1}^{K}\phi(\bw_k^{(r - 1)})} \right\}
+  R = \left\{ \frac{ l[ Y_t(\bs | \bw_1^{(c)}, \ldots, \bw_K^{(c)}, \ldots)] }{l[ Y_t(\bs | \bw_1^{(r - 1)}, \ldots, \bw_K^{(r - 1)}, \ldots)]} \right\} \times \left\{ \frac{ \prod_{k = 1}^{K}\phi(\bw_k^{(c)})}{ \prod_{k = 1}^{K}\phi(\bw_k^{(r - 1)})} \right\} \times \left\{ \frac{ \prod_{k = 1}^{K} p({\bw^*_k}^{(c)})}{ \prod_{k = 1}^{K} p({\bw^*_k}^{(r - 1)})}\right\}
 \end{align*}
-where $l$ is the likelihood given in (\ref{eq:hier}).
+where $l$ is the likelihood given in (\ref{eq:hier}), and $p(\cdot)$ is the prior either taken from the time series given in (\ref{s:temporal}) or assumed to be uniform over $\calD$.
 The candidate knots are accepted with probability $\min\{R, 1\}$.
@@
 \subsection*{Spatial random effects}
-If there is temporal dependence amongst the observations, then we update $z_{tk}$, using a Metropolis Hastings update.
+If there is temporal dependence amongst the observations, then we update $z_{tk}$ using a Metropolis Hastings update.
 First, we generate a candidate using a random walk Gaussian candidate distribution
@@
 \end{align*}
-The candidate is accepted with probability $\min\{R, 1\}$
+The candidate is accepted with probability $\min\{R, 1\}$.
+
+\subsection*{Variance terms}
+When there is more than one site in a partition, then we update $\sigma^2_{tk}$ using a Metropolis Hastings update.
+First, we generate from one of two candidate distributions.
+In the case of temporal dependence, we generate a candidate for ${\sigma^*_{tk}}^2$ using a random walk Gaussian candidate distribution
+\begin{align*}
+  {\sigma^{*2}_{tk}}^{(c)} \sim \text{N}({\sigma^{*2}_{tk}}^{(r - 1)}, s^2)
+\end{align*}
+where ${{\sigma^*_{tk}}^2}^{(r - 1)}$ is the value at MCMC iteration $r - 1$, and $s$ is a tuning parameter.
+In the case where there is no temporal dependence, we generate a candidate for $\sigma^2_{tk}$ using an IG$(a^*/s, b^*/s)$ candidate distribution where $a^* = n_{tk} / 2 + a$, $b^* = [Y_{tk}^T \Sigma^{-1}_{tk} Y_{tk}] / 2 + b$, $n_{tk}$ is the number of sites in partition $k$ on day $t$, and $Y_{tk}$ and $\Sigma^{-1}_{tk}$ are the observations and precision matrix for partition $k$ on day $t$.
+The acceptance ratio is
+\begin{align*}
+  R = \left\{
+    \frac{ l[Y_t(\bs) | {\sigma^2_{tk}}^{(c)}, \ldots] }
+         { l[Y_t(\bs) | {\sigma^2_{tk}}^{(r - 1)}]}
+    \right\} \times \left\{
+    \frac{ l[z_tk | {\sigma^2_{tk}}^{(c)}, \ldots] }
+         { l[z_tk | {\sigma^2_{tk}}^{(r - 1)}, \ldots] }
+    \right\} \times \left\{
+    \frac{ p[ {\sigma^2_{tk}}^{(c)} ] }
+         { p[ {\sigma^2_{tk}}^{(r - 1)}] }
+    \right\} \times \left\{
+    \frac{ c[ {\sigma^2_{tk}}^{(r - 1)}] }
+         { c[ {\sigma^2_{tk}}^{(c)}]}
+    \right\}
+\end{align*}
+where $p[\cdot]$ is the prior either taken from the time series given in (\ref{s:temporal}) or assumed to be IG$(a, b)$, and $c[\cdot]$ is the candidate distribution.
+The candidate is accepted with probability $\min\{R, 1\}$.
 

### 2014-10-21 pot/code/R/auxfunctions.R

@@ -301,12 +301,12 @@ QuantScore <- function(preds, probs, validate) {
   nprobs <- length(probs)  # number of quantiles to find quantile score
   
   # we need to know the predicted quantiles for each site and day in the validation set
-  pred.quants <- apply(preds, c(2, 3), quantile, probs=probs, na.rm=T)  # gives nprobs x np x nt
+  pred.quants <- apply(preds, 2, quantile, probs=probs, na.rm=T)  # gives nprobs x np x nt
   
   scores.sites <- array(NA, dim=c(nprobs, np, nt))
   
   for (q in 1:nprobs) {
-    diff <- pred.quants[q, , ] - validate
+    diff <- pred.quants[q, ] - validate
     i <- diff >= 0  # diff >= 0 means qhat is larger
     scores.sites[q, , ] <- 2 * (i - probs[q]) * diff
   }

### 2014-10-21 pot/code/R/mcmc.R

@@ -222,9 +222,6 @@ mcmc <- function(y, s, x, s.pred=NULL, x.pred=NULL,
     phi.tau <- 0
     acc.phi.tau <- att.phi.tau <- 1
     mh.phi.tau <- 1
-    tau.s <- 1
-    tau.s.a <- 2
-    tau.s.b <- 8
   }
   if (temporalz) {
   	z.star <- z  # need a place to keep track of normal values for time series
@@ -243,7 +240,7 @@ mcmc <- function(y, s, x, s.pred=NULL, x.pred=NULL,
   # candidate distributions for tau depend on the percentage of sites in the 
   # partition. 
   acc.tau.ns <- att.tau.ns <- rep(1, ns)
-  mh.tau.ns  <- rep(0.3, 10)
+  mh.tau.ns  <- rep(0.1, 10)
   mh.tau.parts <- c(0, 0.05, 0.10, 0.20, 0.40, 0.60, 0.80)
   acc.tau    <- att.tau   <- matrix(1, nrow=nknots, ncol=nt) 
   mh.tau <- matrix(0.05, nknots, nt)
@@ -473,7 +470,6 @@ mcmc <- function(y, s, x, s.pred=NULL, x.pred=NULL,
                                 taug=taug, phi=phi.tau, att.phi=att.phi.tau, 
                                 acc.phi=acc.phi.tau, mh.phi=mh.phi.tau,  
                                 tau.alpha=tau.alpha, tau.beta=tau.beta,
-                                s=tau.s, s.a=tau.s.a, s.b=tau.s.b,
                                 res=res, prec.cor=prec.cor, g=g, z=z)
 
         tau <- ts.tau$tau
@@ -482,18 +478,12 @@ mcmc <- function(y, s, x, s.pred=NULL, x.pred=NULL,
         phi.tau <- ts.tau$phi
         att.phi.tau <- ts.tau$att.phi
         acc.phi.tau <- ts.tau$acc.phi
-        # print(tau[1, 1])
-        # print(tau[1, 10])
-        # print(tau[1, 21])
         if ((att.phi.tau > 50) & (iter < (burn / 2))) {
           if (acc.phi.tau / att.phi.tau < 0.25) { mh.phi.tau <- mh.phi.tau * 0.8 }
           if (acc.phi.tau / att.phi.tau > 0.50) { mh.phi.tau <- mh.phi.tau * 1.2 }
-          # print(acc.phi.tau / att.phi.tau)
[…]

### 2014-10-25 pot/LaTeX/skewnormal.tex
Analysis updates
Tweak to initial tau values. Incorporation of time series in the data
analysis.

@@
 The computing is described in section \ref{s:comp}.
-In section \ref{s:simstudy}, we present a simulation study that examines the predictive capabilities of this model compared with a na\:{i}ve Gaussian method.
+In section \ref{s:simstudy}, we present a simulation study that examines the predictive capabilities of this model compared with a na{\"i}ve Gaussian method.
 We then compare our method to Gaussian and max-stable methods with a data analysis of ozone measurements from the eastern US in section \ref{s:analysis}.
@@
 One member of this family, is the skew-$t$ distribution \citep{Azzalini2003}.
-We assume the data can be modeled using a skew-$t$ distribution.
 \citet{Zhang2010} show that the skew-$t$ can be expressed as a hierarchical model
@@
 where $\pi(h)$ is the probability that two sites separated by distance $h$ are in the same partition.
+Figure \ref{fig:simchi} shows $\chi(h)$ for different sample quantiles from simulated datasets where $\bs \in [0, 1] \times [0, 1]$.
+These plots show how the partitioning allows for short range extremal dependence while reducing extremal dependence as the distance between sites increases.
+\begin{figure}
+  \includegraphics[width=\linewidth]{plots/sim-chi.pdf}
+  \caption{$\chi(h)$ for simulated datasets.}
+  \label{fig:simchi}
+\end{figure}
+
 Asymptotic dependence is eliminated as $h$ increases, because
@@
 The model was fit using the training set, and predictions were generated at the testing site locations.
-Because one of the primary goals of this model is to predict extreme events, we quantify use Brier scores and quantile scores to select the model that best fits the data \citep{Gneiting2007}.
+Because one of the primary goals of this model is to predict extreme events, we quantify use Brier scores to select the model that best fits the data \citep{Gneiting2007}.
 The Brier score for predicting exceedance of a threshold $c$ is given by $[e(c) - P(c)]^2$ where $e(c) = I[y>c]$ is an indicator function indicating that a test set value, $y$, has exceeded the threshold, $c$, and $P(c)$ is the predicted probability of exceeding $c$.
-The quantile score for the $\tau$th quantile is $2\{ I[y < \widehat{q}(\tau)] - \tau\} (\widehat{q}(\tau) - y)$ where $y$ is a test set value and $\widehat{q}(\tau)$ is the estimated $\tau$th quantile.
-For both the Brier score and the quantile score, a lower score indicates a better fit.
-These scores were averaged over all sites and days to obtain a single quantile score for each dataset.
+For the Brier score, a lower score indicates a better fit.
 
 \subsection{Results}\label{s:simresults}
-When comparing quantile scores, in almost all of the data settings, the biggest gains are seen when using thresholded methods.
-In all data settings except for the max-stable setting, method 3 performs the best.
-However, in most of the data settings, method 5 also performs reasonably well.
-In fact, for setting 5, paired $t$-tests suggest that the difference in performance is not significant for any quantile score.
-Furthermore, paired $t$-tests results indicate that the difference in quantile scores between methods 3 and 5 is only statistically significant after q(0.97) for setting 4 and after q(0.93) for setting 1.
+We compared the Brier scores for exceeding 11 different thresholds for each dataset.
+For each threshold, Brier score were calculated for each day and site in the testing set.
+The Brier scores were then averaged over all the days and sites to obtain a single Brier score for threshold $c$.
+The thresholds used for the Brier scores are extreme quantiles from the simulated data.
+Figure \ref{fig:simbrierscores} gives the Brier score relative to the Brier score for the Gaussian method calculated as
+\begin{align*}
+  \text{BS}_{\text{rel}} = \frac{\text{BS}_{\text{method}}}{\text{BS}_{\text{Gaussian}}}.
+\end{align*}
+For data that are relatively symmetric, the standard Gaussian method outperforms the other methods.
+However, when the data come from a skewed distribution (settings 3 -- 6), the partitioned and thresholded methods outperform the Gaussian method.
+Additionally, for both of the skew-$t$ settings ($K = 1$, $K = 5$), the skew-$t$ method with the appropriate number of partitions performs the best.
+Finally, the simulation results suggest that for data from a skew-$t$ distribution, thresholded methods provide some improvement over non-thresholded models for the extreme thresholds.
+With the exception of the Brier score for $c = q(0.995)$ in setting 6, the improvement over the Gaussian method is statistically significant using a Wilcoxon signed rank test ($\alpha = 0.05$).
 
 \begin{figure}
-  \includegraphics[width=\linewidth]{plots/quantileplots.pdf}
-  \caption{Quantile score plots for simulation study results}
-  \label{fig:simquantscores}
+  \includegraphics[width=\linewidth]{plots/bsplots-mean.pdf}
+  \caption{Brier scores relative to the Gaussian method for simulation study results. A ratio lower than 1 indicates that the method outperforms relative to the Gaussian method.}
+  \label{fig:simbrierscores}
 \end{figure}

### 2014-10-25 pot/LaTeX/jsm2014/plots/plots.R

@@ -3,8 +3,8 @@ library(fields)
 library(geoR)
 
 # necessary functions
-source('../../code/R/mcmc.R')
-source('../../code/R/auxfunctions.R')
+source('../../../code/R/mcmc.R')
+source('../../../code/R/auxfunctions.R')
 
 # data settings
 beta.t <- c(10, 2, -3)
@@ -241,7 +241,7 @@ bg <- c("firebrick1", "firebrick1", "dodgerblue1", "firebrick1", "dodgerblue1")
 col <- c("firebrick4", "firebrick4", "dodgerblue4", "firebrick4", "dodgerblue4")
 pch <- c(24, 22, 22, 22, 22)
 lty <- c(1, 1, 1, 3, 3)
-quartz()
+
 par(mfrow=c(2, 2))
 # chi-plot sample quantile 0.90
 xplot <- bins[-11]
@@ -281,7 +281,7 @@ lines(xplot, exceed.3[, 9], type="o", lty=lty[3], pch=pch[3], col=col[3], bg=bg[
 lines(xplot, exceed.4[, 9], type="o", lty=lty[4], pch=pch[4], col=col[4], bg=bg[4])
 lines(xplot, exceed.5[, 9], type="o", lty=lty[5], pch=pch[5], col=col[5], bg=bg[5])
 abline(h=0.005, lty=2)
-# legend("topright", legend=methods, lty=lty, col=col, pch=pch, pt.bg=bg)
+legend("topright", legend=methods, lty=lty, col=col, pch=pch, pt.bg=bg, cex=1.5)
 
 plot(xplot, exceed.1[, 9], type="n", axes=F, xlab="", ylab="")
 legend("center", legend=methods, lty=lty, col=col, pch=pch, pt.bg=bg, cex=4, bty="n")

### 2014-10-25 pot/code/analysis/ozone/US/cv5-16US.R

@@ -13,6 +13,7 @@ method <- "t"
 nknots <- 15
 keep.knots <- F
 threshold <- 75
+tau.init <- 0.05
 thresh.quant <- F
 skew <- T
 outputfile <- paste("cv5-", setting, "US.RData", sep="")

### 2014-10-25 pot/code/analysis/ozone/US/cv5-18US.R

@@ -13,6 +13,7 @@ method <- "t"
 nknots <- 1
 keep.knots <- F
 threshold <- 75
+tau.init <- 0.05
 thresh.quant <- F
 skew <- F
 outputfile <- paste("cv5-", setting, "US.RData", sep="")

### 2014-10-25 pot/code/analysis/ozone/US/cv5-24US.R

@@ -13,6 +13,7 @@ method <- "t"
 nknots <- 15
 keep.knots <- F
 threshold <- 75
+tau.init <- 0.05
 thresh.quant <- F
 skew <- F
 outputfile <- paste("cv5-", setting, "US.RData", sep="")

### 2014-10-25 pot/code/analysis/ozone/US/cv5-29US.R

@@ -7,15 +7,16 @@ rm(list=ls())
 load('cv-setup-us.RData')
 source('../../../R/mcmc.R')
 source('../../../R/auxfunctions.R')
+source('../../../R/time_series.R')
 
-setting <- 20
+setting <- 29
 method <- "t"
-nknots <- 5
+nknots <- 1
 keep.knots <- F
 threshold <- 75
 tau.init <- 0.05
 thresh.quant <- F
-skew <- F
+skew <- T
 outputfile <- paste("cv5-", setting, "US.RData", sep="")
 
 start <- proc.time()
@@ -23,32 +24,32 @@ start <- proc.time()
 fit <- vector(mode="list", length=2)
 
 for(val in 1:2){
-	
+
 	set.seed(setting*100 + val)
-	
+
 	cat("CV", val, "started \n")
 	val.idx <- cv.lst[[val]]
 	y.o <- Y[-val.idx,]
 	X.o <- X[-val.idx, , ]
 	S.o <- S[-val.idx,]
-	
+
 	y.p <- Y[val.idx,]
 	X.p <- X[val.idx, , ]
 	S.p <- S[val.idx,]
-	
+
 	tic.set <- proc.time()
 	fit[[val]] <- mcmc(y=y.o, s=S.o, x=X.o, x.pred=X.p, s.pred=S.p,
-	                   method=method, skew=skew, keep.knots=keep.knots,
-	                   thresh.all=threshold, thresh.quant=thresh.quant, nknots=nknots, 
-                       iters=30000, burn=25000, update=500, iterplot=F,
-                       beta.init=beta.init, tau.init=tau.init, rho.init=1,
-                       nu.init=0.5, alpha.init=0.5)
+                   temporalw=T, temporalz=T, temporaltau=T,
+                   method=method, skew=skew, keep.knots=keep.knots,
+	               thresh.all=threshold, thresh.quant=thresh.quant, nknots=nknots,
+                   iters=30000, burn=25000, update=500, iterplot=F,
+                   beta.init=beta.init, tau.init=tau.init, rho.init=1,
+                   nu.init=0.5, alpha.init=0.5)
 	toc.set <- proc.time()
 	time.set <- (toc.set - tic.set)[3]
-	
+
 	elap.time.val <- (proc.time() - start)[3]
 	avg.time.val <- elap.time.val / val
 	cat("CV", val, "finished", round(avg.time.val, 2), "per dataset \n")
 	save(fit, file=outputfile)
 }
-

### 2014-10-25 pot/code/analysis/ozone/US/cv5-2US.R

@@ -13,6 +13,7 @@ method <- "t"
 nknots <- 1
 keep.knots <- F
 threshold <- 0
+tau.init <- 0.05
 thresh.quant <- F
 skew <- T
 outputfile <- paste("cv5-", setting, "US.RData", sep="")

### 2014-10-25 pot/code/analysis/ozone/US/cv5-30US.R

@@ -7,15 +7,16 @@ rm(list=ls())
 load('cv-setup-us.RData')
 source('../../../R/mcmc.R')
 source('../../../R/auxfunctions.R')
+source('../../../R/time_series.R')
 
-setting <- 20
+setting <- 30
 method <- "t"
-nknots <- 5
+nknots <- 1
 keep.knots <- F
-threshold <- 75
+threshold <- 90
 tau.init <- 0.05
 thresh.quant <- F
-skew <- F
+skew <- T
 outputfile <- paste("cv5-", setting, "US.RData", sep="")
 
 start <- proc.time()
@@ -23,32 +24,32 @@ start <- proc.time()
 fit <- vector(mode="list", length=2)
 
 for(val in 1:2){
-	
+
 	set.seed(setting*100 + val)
-	
+
 	cat("CV", val, "started \n")
 	val.idx <- cv.lst[[val]]
 	y.o <- Y[-val.idx,]
 	X.o <- X[-val.idx, , ]
 	S.o <- S[-val.idx,]
-	
+
 	y.p <- Y[val.idx,]
 	X.p <- X[val.idx, , ]
 	S.p <- S[val.idx,]
-	
+
 	tic.set <- proc.time()
 	fit[[val]] <- mcmc(y=y.o, s=S.o, x=X.o, x.pred=X.p, s.pred=S.p,
-	                   method=method, skew=skew, keep.knots=keep.knots,
-	                   thresh.all=threshold, thresh.quant=thresh.quant, nknots=nknots, 
-                       iters=30000, burn=25000, update=500, iterplot=F,
-                       beta.init=beta.init, tau.init=tau.init, rho.init=1,
-                       nu.init=0.5, alpha.init=0.5)
+                   temporalw=T, temporalz=T, temporaltau=T,
+                   method=method, skew=skew, keep.knots=keep.knots,
+	               thresh.all=threshold, thresh.quant=thresh.quant, nknots=nknots,
+                   iters=30000, burn=25000, update=500, iterplot=F,
+                   beta.init=beta.init, tau.init=tau.init, rho.init=1,
+                   nu.init=0.5, alpha.init=0.5)
 	toc.set <- proc.time()
 	time.set <- (toc.set - tic.set)[3]
-	
+
 	elap.time.val <- (proc.time() - start)[3]
 	avg.time.val <- elap.time.val / val
 	cat("CV", val, "finished", round(avg.time.val, 2), "per dataset \n")
 	save(fit, file=outputfile)
 }
-

### 2014-10-25 pot/code/analysis/ozone/US/cv5-31US.R

@@ -7,15 +7,16 @@ rm(list=ls())
 load('cv-setup-us.RData')
 source('../../../R/mcmc.R')
 source('../../../R/auxfunctions.R')
+source('../../../R/time_series.R')
 
-setting <- 20
+setting <- 31
 method <- "t"
 nknots <- 5
 keep.knots <- F
-threshold <- 75
+threshold <- 0
 tau.init <- 0.05
 thresh.quant <- F
-skew <- F
+skew <- T
 outputfile <- paste("cv5-", setting, "US.RData", sep="")
 
 start <- proc.time()
@@ -23,32 +24,32 @@ start <- proc.time()
 fit <- vector(mode="list", length=2)
 
 for(val in 1:2){
-	
+
 	set.seed(setting*100 + val)
-	
+
 	cat("CV", val, "started \n")
 	val.idx <- cv.lst[[val]]
 	y.o <- Y[-val.idx,]
 	X.o <- X[-val.idx, , ]
 	S.o <- S[-val.idx,]
-	
+
 	y.p <- Y[val.idx,]
 	X.p <- X[val.idx, , ]
 	S.p <- S[val.idx,]
-	
+
 	tic.set <- proc.time()
 	fit[[val]] <- mcmc(y=y.o, s=S.o, x=X.o, x.pred=X.p, s.pred=S.p,
-	                   method=method, skew=skew, keep.knots=keep.knots,
-	                   thresh.all=threshold, thresh.quant=thresh.quant, nknots=nknots, 
-                       iters=30000, burn=25000, update=500, iterplot=F,
-                       beta.init=beta.init, tau.init=tau.init, rho.init=1,
-                       nu.init=0.5, alpha.init=0.5)
+                   temporalw=T, temporalz=T, temporaltau=T,
+                   method=method, skew=skew, keep.knots=keep.knots,
+	               thresh.all=threshold, thresh.quant=thresh.quant, nknots=nknots,
+                   iters=30000, burn=25000, update=500, iterplot=F,
+                   beta.init=beta.init, tau.init=tau.init, rho.init=1,
+                   nu.init=0.5, alpha.init=0.5)
 	toc.set <- proc.time()
 	time.set <- (toc.set - tic.set)[3]
-	
+
 	elap.time.val <- (proc.time() - start)[3]
 	avg.time.val <- elap.time.val / val
 	cat("CV", val, "finished", round(avg.time.val, 2), "per dataset \n")
 	save(fit, file=outputfile)
 }
-

### 2014-10-25 pot/code/analysis/ozone/US/cv5-37US.R

@@ -7,15 +7,16 @@ rm(list=ls())
 load('cv-setup-us.RData')
 source('../../../R/mcmc.R')
 source('../../../R/auxfunctions.R')
+source('../../../R/time_series.R')
 
-setting <- 20
+setting <- 37
 method <- "t"
-nknots <- 5
+nknots <- 10
 keep.knots <- F
 threshold <- 75
 tau.init <- 0.05
 thresh.quant <- F
-skew <- F
+skew <- T
 outputfile <- paste("cv5-", setting, "US.RData", sep="")
 
 start <- proc.time()
@@ -23,32 +24,32 @@ start <- proc.time()
 fit <- vector(mode="list", length=2)
 
 for(val in 1:2){
-	
+
 	set.seed(setting*100 + val)
-	
+
 	cat("CV", val, "started \n")
 	val.idx <- cv.lst[[val]]
 	y.o <- Y[-val.idx,]
 	X.o <- X[-val.idx, , ]
 	S.o <- S[-val.idx,]
-	
+
 	y.p <- Y[val.idx,]
 	X.p <- X[val.idx, , ]
 	S.p <- S[val.idx,]
-	
+
 	tic.set <- proc.time()
 	fit[[val]] <- mcmc(y=y.o, s=S.o, x=X.o, x.pred=X.p, s.pred=S.p,
-	                   method=method, skew=skew, keep.knots=keep.knots,
-	                   thresh.all=threshold, thresh.quant=thresh.quant, nknots=nknots, 
-                       iters=30000, burn=25000, update=500, iterplot=F,
-                       beta.init=beta.init, tau.init=tau.init, rho.init=1,
-                       nu.init=0.5, alpha.init=0.5)
+                   temporalw=T, temporalz=T, temporaltau=T,
+                   method=method, skew=skew, keep.knots=keep.knots,
+	               thresh.all=threshold, thresh.quant=thresh.quant, nknots=nknots,
+                   iters=30000, burn=25000, update=500, iterplot=F,
+                   beta.init=beta.init, tau.init=tau.init, rho.init=1,
+                   nu.init=0.5, alpha.init=0.5)
 	toc.set <- proc.time()
 	time.set <- (toc.set - tic.set)[3]
-	
+
 	elap.time.val <- (proc.time() - start)[3]
 	avg.time.val <- elap.time.val / val
 	cat("CV", val, "finished", round(avg.time.val, 2), "per dataset \n")
 	save(fit, file=outputfile)
 }
-

### 2014-10-25 pot/code/analysis/ozone/US/cv5-38US.R

@@ -7,15 +7,16 @@ rm(list=ls())
 load('cv-setup-us.RData')
 source('../../../R/mcmc.R')
 source('../../../R/auxfunctions.R')
+source('../../../R/time_series.R')
 
-setting <- 20
+setting <- 38
 method <- "t"
-nknots <- 5
+nknots <- 10
 keep.knots <- F
-threshold <- 75
+threshold <- 90
 tau.init <- 0.05
 thresh.quant <- F
-skew <- F
+skew <- T
 outputfile <- paste("cv5-", setting, "US.RData", sep="")
 
 start <- proc.time()
@@ -23,32 +24,32 @@ start <- proc.time()
 fit <- vector(mode="list", length=2)
 
 for(val in 1:2){
-	
+
 	set.seed(setting*100 + val)
-	
+
 	cat("CV", val, "started \n")
 	val.idx <- cv.lst[[val]]
 	y.o <- Y[-val.idx,]
 	X.o <- X[-val.idx, , ]
 	S.o <- S[-val.idx,]
-	
+
 	y.p <- Y[val.idx,]
 	X.p <- X[val.idx, , ]
 	S.p <- S[val.idx,]
-	
+
 	tic.set <- proc.time()
 	fit[[val]] <- mcmc(y=y.o, s=S.o, x=X.o, x.pred=X.p, s.pred=S.p,
-	                   method=method, skew=skew, keep.knots=keep.knots,
-	                   thresh.all=threshold, thresh.quant=thresh.quant, nknots=nknots, 
-                       iters=30000, burn=25000, update=500, iterplot=F,
-                       beta.init=beta.init, tau.init=tau.init, rho.init=1,
-                       nu.init=0.5, alpha.init=0.5)
+                   temporalw=T, temporalz=T, temporaltau=T,
+                   method=method, skew=skew, keep.knots=keep.knots,
+	               thresh.all=threshold, thresh.quant=thresh.quant, nknots=nknots,
+                   iters=30000, burn=25000, update=500, iterplot=F,
+                   beta.init=beta.init, tau.init=tau.init, rho.init=1,
+                   nu.init=0.5, alpha.init=0.5)
 	toc.set <- proc.time()
 	time.set <- (toc.set - tic.set)[3]
-	
+
 	elap.time.val <- (proc.time() - start)[3]
 	avg.time.val <- elap.time.val / val
 	cat("CV", val, "finished", round(avg.time.val, 2), "per dataset \n")
 	save(fit, file=outputfile)
 }
-

### 2014-10-25 pot/code/analysis/ozone/US/cv5-47US.R

@@ -7,10 +7,11 @@ rm(list=ls())
 load('cv-setup-us.RData')
 source('../../../R/mcmc.R')
 source('../../../R/auxfunctions.R')
+source('../../../R/time_series.R')
 
-setting <- 20
+setting <- 47
 method <- "t"
-nknots <- 5
+nknots <- 10
 keep.knots <- F
 threshold <- 75
 tau.init <- 0.05
@@ -23,32 +24,32 @@ start <- proc.time()
 fit <- vector(mode="list", length=2)
 
 for(val in 1:2){
-	
+
 	set.seed(setting*100 + val)
-	
+
 	cat("CV", val, "started \n")
 	val.idx <- cv.lst[[val]]
 	y.o <- Y[-val.idx,]
 	X.o <- X[-val.idx, , ]
 	S.o <- S[-val.idx,]
-	
+
 	y.p <- Y[val.idx,]
 	X.p <- X[val.idx, , ]
 	S.p <- S[val.idx,]
-	
+
 	tic.set <- proc.time()
 	fit[[val]] <- mcmc(y=y.o, s=S.o, x=X.o, x.pred=X.p, s.pred=S.p,
-	                   method=method, skew=skew, keep.knots=keep.knots,
-	                   thresh.all=threshold, thresh.quant=thresh.quant, nknots=nknots, 
-                       iters=30000, burn=25000, update=500, iterplot=F,
-                       beta.init=beta.init, tau.init=tau.init, rho.init=1,
-                       nu.init=0.5, alpha.init=0.5)
+                   temporalw=T, temporalz=F, temporaltau=T,
+                   method=method, skew=skew, keep.knots=keep.knots,
+	               thresh.all=threshold, thresh.quant=thresh.quant, nknots=nknots,
+                   iters=30000, burn=25000, update=500, iterplot=F,
+                   beta.init=beta.init, tau.init=tau.init, rho.init=1,
+                   nu.init=0.5, alpha.init=0.5)
 	toc.set <- proc.time()
 	time.set <- (toc.set - tic.set)[3]
-	
+
 	elap.time.val <- (proc.time() - start)[3]
 	avg.time.val <- elap.time.val / val
 	cat("CV", val, "finished", round(avg.time.val, 2), "per dataset \n")
 	save(fit, file=outputfile)
 }
-

### 2014-10-25 pot/code/analysis/ozone/US/cv5-49US.R

@@ -7,10 +7,11 @@ rm(list=ls())
 load('cv-setup-us.RData')
 source('../../../R/mcmc.R')
 source('../../../R/auxfunctions.R')
+source('../../../R/time_series.R')
 
-setting <- 20
+setting <- 49
 method <- "t"
-nknots <- 5
+nknots <- 15
 keep.knots <- F
 threshold <- 75
 tau.init <- 0.05
@@ -23,32 +24,32 @@ start <- proc.time()
 fit <- vector(mode="list", length=2)
 
 for(val in 1:2){
-	
+
 	set.seed(setting*100 + val)
-	
+
 	cat("CV", val, "started \n")
 	val.idx <- cv.lst[[val]]
 	y.o <- Y[-val.idx,]
 	X.o <- X[-val.idx, , ]
 	S.o <- S[-val.idx,]
-	
+
 	y.p <- Y[val.idx,]
 	X.p <- X[val.idx, , ]
 	S.p <- S[val.idx,]
-	
+
 	tic.set <- proc.time()
 	fit[[val]] <- mcmc(y=y.o, s=S.o, x=X.o, x.pred=X.p, s.pred=S.p,
-	                   method=method, skew=skew, keep.knots=keep.knots,
-	                   thresh.all=threshold, thresh.quant=thresh.quant, nknots=nknots, 
-                       iters=30000, burn=25000, update=500, iterplot=F,
-                       beta.init=beta.init, tau.init=tau.init, rho.init=1,
-                       nu.init=0.5, alpha.init=0.5)
+                   temporalw=T, temporalz=F, temporaltau=T,
+                   method=method, skew=skew, keep.knots=keep.knots,
+	               thresh.all=threshold, thresh.quant=thresh.quant, nknots=nknots,
+                   iters=30000, burn=25000, update=500, iterplot=F,
+                   beta.init=beta.init, tau.init=tau.init, rho.init=1,
+                   nu.init=0.5, alpha.init=0.5)
 	toc.set <- proc.time()
 	time.set <- (toc.set - tic.set)[3]
-	
+
 	elap.time.val <- (proc.time() - start)[3]
 	avg.time.val <- elap.time.val / val
 	cat("CV", val, "finished", round(avg.time.val, 2), "per dataset \n")
 	save(fit, file=outputfile)
 }
-

### 2014-10-25 pot/code/analysis/ozone/US/cv5-50US.R

@@ -7,12 +7,13 @@ rm(list=ls())
 load('cv-setup-us.RData')
 source('../../../R/mcmc.R')
 source('../../../R/auxfunctions.R')
+source('../../../R/time_series.R')
 
-setting <- 20
+setting <- 50
 method <- "t"
-nknots <- 5
+nknots <- 15
 keep.knots <- F
-threshold <- 75
+threshold <- 90
 tau.init <- 0.05
 thresh.quant <- F
 skew <- F
@@ -23,32 +24,32 @@ start <- proc.time()
 fit <- vector(mode="list", length=2)
 
 for(val in 1:2){
-	
+
 	set.seed(setting*100 + val)
-	
+
 	cat("CV", val, "started \n")
 	val.idx <- cv.lst[[val]]
 	y.o <- Y[-val.idx,]
 	X.o <- X[-val.idx, , ]
 	S.o <- S[-val.idx,]
-	
+
 	y.p <- Y[val.idx,]
 	X.p <- X[val.idx, , ]
 	S.p <- S[val.idx,]
-	
+
 	tic.set <- proc.time()
 	fit[[val]] <- mcmc(y=y.o, s=S.o, x=X.o, x.pred=X.p, s.pred=S.p,
-	                   method=method, skew=skew, keep.knots=keep.knots,
-	                   thresh.all=threshold, thresh.quant=thresh.quant, nknots=nknots, 
-                       iters=30000, burn=25000, update=500, iterplot=F,
-                       beta.init=beta.init, tau.init=tau.init, rho.init=1,
-                       nu.init=0.5, alpha.init=0.5)
+                   temporalw=T, temporalz=F, temporaltau=T,
+                   method=method, skew=skew, keep.knots=keep.knots,
+	               thresh.all=threshold, thresh.quant=thresh.quant, nknots=nknots,
+                   iters=30000, burn=25000, update=500, iterplot=F,
+                   beta.init=beta.init, tau.init=tau.init, rho.init=1,
+                   nu.init=0.5, alpha.init=0.5)
 	toc.set <- proc.time()
 	time.set <- (toc.set - tic.set)[3]
-	
+
 	elap.time.val <- (proc.time() - start)[3]
 	avg.time.val <- elap.time.val / val
 	cat("CV", val, "finished", round(avg.time.val, 2), "per dataset \n")
 	save(fit, file=outputfile)
 }
-

### 2014-10-25 pot/code/analysis/ozone/US/cv5-7US.R

@@ -13,6 +13,7 @@ method <- "t"
 nknots <- 5
 keep.knots <- F
 threshold <- 50
+tau.init <- 0.05
 thresh.quant <- F
 skew <- T
 outputfile <- paste("cv5-", setting, "US.RData", sep="")

### 2014-10-25 pot/code/analysis/ozone/US/cv5-9US.R

@@ -13,6 +13,7 @@ method <- "t"
 nknots <- 5
 keep.knots <- F
 threshold <- 90
+tau.init <- 0.05
 thresh.quant <- F
 skew <- T
 outputfile <- paste("cv5-", setting, "US.RData", sep="")

### 2014-10-25 pot/code/analysis/simstudy/results.R

@@ -140,33 +140,86 @@ for (i in 1:length(probs)) { for (j in 1:(nmethods-1)) { for (k in 1:nsettings) 
 
 round(paired.results, 4)
 
+# wilcoxon signed-rank test for Brier scores
+wilcox.results <- array(NA, dim=c(length(probs), (nmethods-1), nsettings))
+best.methods <- vector(mode="list")
+for (i in 1:length(probs)) { for (k in 1:nsettings) {
+  ref.j <- best.brier[i, k]
+  cmp.j <- (1:5)[-ref.j]
+  for (j in 1:(nmethods-1)) {
+    wilcox.results[i, j, k] <- wilcox.test(brier.score[i, , ref.j, k], brier.score[i, , cmp.j[j], k], paired=T)$p.value
+  }
+}}
 
-setting.title <- c("Gaussian", "t-1", "t-5", "skew t-1 (alpha = 3)", "skew t-5 (alpha = 3)", "max-stable")
-methods <- c("Gaussian", "skew-t, K = 1, T = q(0.0)", "t, K = 1, T = q(0.8)", "skew-t, K = 5, T = q(0.0)", "t, K = 5, T = q(0.8)")
-bg <- c("firebrick1", "firebrick1", "dodgerblue1", "firebrick1", "dodgerblue1")
-col <- c("firebrick4", "firebrick4", "dodgerblue4", "firebrick4", "dodgerblue4")
-pch <- c(24, 22, 22, 22, 22)
-lty <- c(1, 1, 1, 3, 3)
+# get single brier scores and quantile scores for each setting x method x quantile
+quant.score.med <- apply(quant.score, c(1, 3, 4), median, na.rm=T)
+brier.score.med <- apply(brier.score, c(1, 3, 4), median, na.rm=T)
+
+bs.med.ref.gau <- array(NA, dim=c(11, 4, 6))
+bs.mean.ref.gau <- array(NA, dim=c(11, 4, 6))
+for (j in 1:4) {
+  bs.med.ref.gau[, j, ] <- brier.score.med[, (j + 1), ] / brier.score.med[, 1, ]
+  bs.mean.ref.gau[, j, ] <- brier.score.mean[, (j + 1), ] / brier.score.mean[, 1, ]
+}
+
+
+setting.title <- c("Gaussian", "t (K = 1)", "t (K = 5)", "skew t (K = 1, alpha = 3)", "skew t (K = 5, alpha = 3)", "max-stable")
[…]

### 2014-11-02 pot/LaTeX/skewnormal.tex
update results file with new plots

@@
 The paper is organized as follows.
-In section \ref{s:model}, we describe the skew-$t$ process with partitioning and thresholding.
+In section \ref{s:spatial}, we describe a spatial skew-$t$ process with partitioning and thresholding.
+In section \ref{s:temporal}, we provide a way to model temporal correlation.
 The computing is described in section \ref{s:comp}.
@@
 
-\section{Statistical model}\label{s:model}
-Let $Y_t(\bs)\in \calR$ be the observed value at location $\bs$ and timepoint $t$.
-To avoid bias in estimating tail parameters, we model censored data
+\section{Spatial model}\label{s:spatial}
+To avoid bias in estimating tail parameters, we model censored data.
+Let
 \beq\label{Yt}
-  \Yt_t(\bs) = \left\{
-          \begin{array}{ll}
-            Y_t(\bs) & Y_t(\bs)>T \\
-            T & Y_t(\bs)\le T
-          \end{array}
-        \right.
+  \tilde{Y}(\bs) = \delta(\bs) Y(\bs) + (1 - \delta(\bs)) T
 \eeq
-where $T$ is a pre-specified threshold.
-Then, assuming the full data are observations from a skew-$t$ process, we update values censored below the threshold using standard Bayesian missing data methods as described in Section \ref{s:comp}.
+be the censored observation at site $\bs$ where $Y(\bs)$ is the uncensored observation, $\delta(\bs) = I[Y(\bs) > T]$, and $T$ is a pre-specified threshold value.
+Then, assuming the uncensored data are observations from a skew-$t$ process, we update values censored below the threshold using standard Bayesian missing data methods as described in Section \ref{s:comp}.
 
@@
 The skew-elliptical family of distributions provides models that are mathematically tractable while introducing additional paremeters to account for non-symmetric data \citep{Genton2004}.
-One member of this family, is the skew-$t$ distribution \citep{Azzalini2003}.
+One member of this family, is the skew-$t$ distribution \citep[see Appendix A.5]{Azzalini2003}.
 \citet{Zhang2010} show that the skew-$t$ can be expressed as a hierarchical model
 \begin{align} \label{eq:fullmodel}
-  Y_t(\bs) &= X_t(\bs) \beta + \alpha |z_t| + \sigma_t v_t(\bs)
+  \bY &= \bX \bbeta + \alpha |z| + \sigma \bv
 \end{align}
-where $\alpha \in \calR$ controls the skewness, $z_{t} \ind N(0, \sigma^2_t)$ are a random effect from a half-normal distribution (see appendix A.4), $v_t(\bs)$ is a Gaussian process with mean zero, variance one, \Matern correlation, and $\sigma^2_t \iid \text{IG}(a, b)$.
-When marginalizing over the $z_t$ and $\sigma^2_t$ terms,
+where $\alpha \in \calR$ controls the skewness; $z \sim N(0, \sigma^2)$ is a random effect from a half-normal distribution (see appendix A.4); $\bv$ is a Gaussian process with mean zero, variance one, \Matern spatial correlation with spatial range $\rho$, smoothness $\nu$, and $\gamma$ is the proportion of variance accounted for by the spatial variation; and $\sigma^2 \sim \text{IG}(a, b)$.
+When marginalizing over the $z$ and $\sigma^2$ terms, the joint distribution for all sites becomes
 \begin{align*}
-  Y_t(\bs) \sim \text{skew-}t (\mu, \Sigma^*, \alpha, \text{df} = 2a)
+  \bY \sim \text{skew-}t (\mu, \bSigma^*, \alpha, \text{df} = 2a)
 \end{align*}
-where $\mu$ is the location, $\Sigma^* = \frac{ b }{ a } \Sigma$, $\Sigma$ is a \Matern covariance matrix, and $\alpha \in \calR$ controls the skewness.
+where $\mu$ is the location, $\bSigma^* = \frac{ b }{ a } \bSigma$, $\bSigma$ is a \Matern covariance matrix, and $\alpha \in \calR$ controls the skewness.
 The skew-$t$ process is desirable because of its flexible tail that is controlled by the skewness parameter, $\alpha$, and the degrees of freedom, $2a$.
@@
 \subsection{Extremal dependence}
-One common measure of extremal spatial dependence is the extremal coefficient which describes the pairwise dependence between spatial locations \citep{Smith1990}.
-Consider a spatial process $Y(\bs) \in \calR^n$ observed at locations $s \in \calD \subset \calR^2$.
-Then the bivariate extremal coefficient, $\theta(\bs_i, \bs_j) \in [1, 2]$, is defined as
-\begin{align}
-  \Pr(Y(\bs_i) < c, Y(\bs_j) < c) = \Pr(Y(\bs_i) < c)^{\theta(\bs_i, \bs_j) }. \label{eq:extremalcoef}
-\end{align}
-One way to characterize the dependence over the entire set of spatial locations is to calculate all of the pairwise extremal coefficients.
-Although this method provides information regarding the spatial structure of the observations, it does not fully characterize the joint spatial dependence.
-
-Another measure of extremal dependence is the $\chi$ statistic.
+One measure of extremal dependence is the $\chi$ statistic \citep{Padoan2011}.
 The $\chi$ statistic for the upper tail is given by
 \begin{align*}
-  \chi = \lim_{c \rightarrow \infty} \Pr(Y(\bs_1) > c | Y(\bs_2) > c),
+  \chi = \lim_{c \rightarrow \infty} \Pr(Y(\bs_1) > c | Y(\bs_2) > c).
 \end{align*}
@@
 \begin{align*}
-  \chi(\bh) = \lim_{c \rightarrow \infty} \Pr(Y(0) > c | Y(\bh) > c)
+  \chi(h) = \lim_{c \rightarrow \infty} \Pr(Y(0) > c | Y(h) > c)
 \end{align*}
-where $\bh = ||\bs_1 - \bs_2||$.
-If $\chi(\bh) = 0$, then observations are asymptotically independent at distance $\bh$.
-For Gaussian processes, $\chi(\bh) = 0$ regardless of the the distance, so they are not suitable for modeling spatially dependent extremes.
-However, for the process described in Section \ref{s:data}, $\chi(\bh) > 0$ \citep{Padoan2011}.
+where $h = ||\bs_1 - \bs_2||$.
+If $\chi(h) = 0$, then observations are asymptotically independent at distance $h$.
+For Gaussian processes, $\chi(h) = 0$ regardless of the the distance, so they are not suitable for modeling spatially dependent extremes.
+However, for the skew-$t$ process described in \ref{s:data}, $\chi(h) > 0$ \citep[see Appendix A.5 for the expression]{Padoan2011}.
 
@@
 One problem with the spatial skew-$t$ process is that all sites are asymptotically dependent regardless of their spatial separation.
-This occurs because all observations, both near and far, share the same $z_t$ and $\sigma^2_t$ terms.
-We handle this problem with a daily random partition similar to \citet{Kim2005} that allows $z_t$ and $\sigma^2_t$ to vary by site.
-The model then becomes
+This occurs because all observations, both near and far, share the same $z$ and $\sigma^2$ terms.
+Even more problematic is that $\lim_{h \rightarrow \infty} \chi(h) \neq 0$.
+We handle this problem with a random partition similar to \citet{Kim2005} that allows $z$ and $\sigma^2$ to vary by site.
+
+Consider a set of spatial knots $\bw_{k} \sim \text{Uniform}(\calD)$ that define a random daily partition $P_{1}, \ldots, P_{K}$ of the spatial domain of interest $\calD \subset \calR^2$ such that
+\begin{align*}
+  P_{k} = \{ \bs : k = \argmin_\ell || \bs - \bw_{\ell} || \}.
+\end{align*}
+Then, the observation at each site becomes
 \begin{align}
-  Y_t(\bs) &= X_t(\bs) \beta + \alpha |z_t(\bs)| + \sigma_t(\bs) v_t(\bs).
+  Y(\bs) &= X(\bs) \bbeta + \alpha |z(\bs)| + \sigma v(\bs),
 \end{align}
-In this extension of (\ref{eq:fullmodel}), $z_t(\bs)$ and $\sigma_t(\bs)$ are allowed to vary by site.
-Their spatial variation is determined by the random partition model defined below.
-Consider a set of daily spatial knots $\bw_{tk} \sim \text{Uniform}(\calD)$ that define a random daily partition $P_{t1}, \ldots, P_{tK}$ of the spatial domain of interest $\calD \subset \calR^2$ such that
-\begin{align*}
-  P_{tk} = \{ \bs : k = \argmin_\ell || \bs - \bw_{t\ell} || \}.
-\end{align*}
-So, for $\bs \in P_{tk}$, let
+and, for $\bs \in P_{k}$,
 \begin{align}
-  z_t(\bs) &= z_{tk} \label{eq:sitez}\\
-  \sigma^2_t(\bs) &= \sigma^2_{tk} \label{eq:sitesig}.
+  z(\bs) &= z_{tk} \label{eq:sitez}\\
+  \sigma^2(\bs) &= \sigma^2_{k} \label{eq:sitesig}.
 \end{align}
-Then within each partition, $Y_t(\bs)$ follows the distribution given in (\ref{eq:fullmodel}).
+So, within each partition, $\bY$ follows the distribution given in (\ref{eq:fullmodel}).
+
 When incorporating the random daily partition, the $\chi$ statistic becomes
 \begin{align}
-  \chi(\bh) = \lim_{c \rightarrow \infty} \pi(\bh) \chi(\bh)
+  \chi(h) = \pi(h) \chi(h) + (1 - \pi(h)) \chi(h)
 \end{align}
 where $\pi(h)$ is the probability that two sites separated by distance $h$ are in the same partition.
+Asymptotic dependence is eliminated as $h$ increases, because
+\begin{align}
+  \lim_{h \rightarrow \infty} \chi(h) = \lim_{h \rightarrow \infty} \pi(h) \chi(h) = 0.
+\end{align}
+A proof of this can be found in Appendix A.3.
+
 Figure \ref{fig:simchi} shows $\chi(h)$ for different sample quantiles from simulated datasets where $\bs \in [0, 1] \times [0, 1]$.
@@
 \end{figure}
-
-Asymptotic dependence is eliminated as $h$ increases, because
-\begin{align}
-  \lim_{h \rightarrow \infty} \chi(\bh) = \lim_{h \rightarrow \infty} \pi(\bh) \chi(\bh) = 0.
-\end{align}
-A proof of this can be found in Appendix A.3.
 

### 2014-11-02 pot/code/analysis/simstudy/results6.R

@@ -26,10 +26,10 @@ load("simdata.RData")
 ns <- dim(y)[1]
 nt <- dim(y)[2]
 nsets <- 5
-ngroups <- 10
+ngroups <- 1
 nsettings <- dim(y)[4]
 nmethods <- 5
-obs <- c(rep(T, 100), rep(F, 30))
+obs <- c(rep(T, 100), rep(F, 44))
 
 setting <- 6
 filename <- paste("scores", setting, ".RData", sep="")

### 2014-11-10 pot/LaTeX/skewnormal.tex
tweaks for time series

@@
 In section \ref{s:spatial}, we describe a spatial skew-$t$ process with partitioning and thresholding.
-In section \ref{s:temporal}, we provide a way to model temporal correlation.
+In section \ref{s:temporal}, we expand the model to incorporate temporal dependence.
 The computing is described in section \ref{s:comp}.
@@
 For Gaussian processes, $\chi(h) = 0$ regardless of the the distance, so they are not suitable for modeling spatially dependent extremes.
-However, for the skew-$t$ process described in \ref{s:data}, $\chi(h) > 0$ \citep[see Appendix A.5 for the expression]{Padoan2011}.
+However, for the skew-$t$ process described in \ref{s:data}, $\chi(h) > 0$ \citep[they give an expression that is the sum of two survival functions from the skew-$t$ distribution.
+The expression isn't terribly informative, but I can include it if you think it would be helpful.]{Padoan2011}.
 
@@
   z(\bs) &= z_{tk} \label{eq:sitez}\\
-  \sigma^2(\bs) &= \sigma^2_{k} \label{eq:sitesig}.
+  \sigma(\bs) &= \sigma_{k} \label{eq:sitesig}.
 \end{align}
@@
 
-When incorporating the random daily partition, the $\chi$ statistic becomes
-\begin{align}
-  \chi(h) = \pi(h) \chi(h) + (1 - \pi(h)) \chi(h)
-\end{align}
+When incorporating the random daily partition, for two sites in the same partition, the $\chi$ statistic is $\chi(h) = \chi_{\text{skew-}t}(h)$.
+However, for two sites in different partitions, the $\chi$ statistic should be the same as the Gaussian $\chi(h)$, because the only correlation between the two sites, is the correlation given by $\bSigma$.
+So, to show that $\lim_{h \rightarrow \infty} \chi(h) = 0$, we need only know that $\lim_{h \rightarrow \infty} \pi(h) = 0$
 where $\pi(h)$ is the probability that two sites separated by distance $h$ are in the same partition.
-Asymptotic dependence is eliminated as $h$ increases, because
-\begin{align}
-  \lim_{h \rightarrow \infty} \chi(h) = \lim_{h \rightarrow \infty} \pi(h) \chi(h) = 0.
-\end{align}
-A proof of this can be found in Appendix A.3.
+A proof of this is given in Appendix A.3.
 
@@
 
-\subsection{Accounting for temporal dependence} \label{s:temporal}
+\section{Accounting for temporal dependence} \label{s:temporal}
 In a standard block-maxima approach, the temporal dependence is assumed to be negligible because blocks (e.g. yearly maxima) are taken to be large enough to assume independence.
 However, when using daily meausrements, the assumption of temporal independence is no longer appropriate.
-We account for temporal dependence through an AR(1) time series on the spatial knots, $z_t(\bs)$ terms, and $\sigma^2_t(\bs)$ terms.
+There are several places where temporal dependence could be included, including the residual $v(\bs)$.
+However, we choose to allow for temporal dependence in the $\bw$, $z(\bs)$, and $\sigma(\bs)$ terms because these terms derive the tail behavior.
+Specifically, we incorporate an AR(1) time series.
 We first, transform the spatial knots from $\calD$ to $\calR^2$ as follows.
@@
 \end{align*}
+\hl{What's the best way to write the $\bw$ transformation as 1-D? $w_x$ and $w_y$ or $w_1$ and $w_2$?}
 where $\Phi_2$ represents a bivariate standard normal density function.
@@
 Then, when $t > 1$ the time series is modeled as
-\begin{align}
-  \bw^*_{t, k} | \bw^*_{t-1, k} &\sim N_2\left[\phi_w \bw^*_{t-1, k}, (1 - \phi_w^2) \right] \\
-  z_{t, k} | z_{t-1, k} &\sim N \left[\phi_z z_{t-1, k}, \sigma^2_{t,k} (1 - \phi_z^2)\right] \\
-  \sigma^{2*}_{t, k} | \sigma^{2*}_{t-1, k} &\sim N \left[\phi_\sigma \sigma^{2*}_{t-1, k}, (1 - \phi_\sigma^2) \right]
-\end{align}
+\begin{align*}
+  \bw^*_{1, k} & \sim N_2(0, 1)\\
+  z_{1k} & \sim N(0, \sigma^2_{1k})\\
+  \sigma^{2*}_{1k} &\sim N(0, 1)\\
+  \bw^*_{tk} | \bw^*_{t-1, k} &\sim N_2\left[\phi_w \bw^*_{t-1, k}, (1 - \phi_w^2) \right] \\
+  z_{tk} | z_{t-1, k} &\sim N \left[\phi_z z_{t-1, k}, \sigma^2_{tk} (1 - \phi_z^2)\right] \\
+  \sigma^{2*}_{tk} | \sigma^{2*}_{t-1, k} &\sim N \left[\phi_\sigma \sigma^{2*}_{t-1, k}, (1 - \phi_\sigma^2) \right]
+\end{align*}
 where $|\phi_w| < 1$, $0 < \phi_z < 1$, and $|\phi_\sigma| < 1$.
-All three of the time series are initialized using $N(0, 1)$.
+These are stationary time series models with marginal distributions
+\begin{align*}
+  \bw^*_{k} & \sim N_2(0, 1)\\
+  z_{k} & \sim N(0, \sigma^2_{k})\\
+  \sigma^{2*}_{k} &\sim N(0, 1).
+\end{align*}
+For each day, the model is identical to the spatial-only model in Seciton \ref{s:part}.
 
 \subsection{Hierarchical model}\label{s:hier}
-Conditioned on $z_{t,k}(\bs) \iid $ $N(0, \sigma^2_{t,k})$, $\sigma^2_{t,k}(\bs) \sim IG(a, b)$, and $P_{t,k}$, the marginal distributions are skew-$t$ and the joint distribution within a partition is multivariate skew-$t$.
+Conditioned on $z_{tk}(\bs)$, $\sigma^2_{t,k}(\bs)$, and $P_{t,k}$, the marginal distributions are Gaussian and the joint distribution multivariate Gaussian.
 However, we do not fix the partitions, they are treated as unknown and updated in the MCMC.
 We model this with a Bayesian hierarchical model as follows.
-Let $\bw_{t,1}, \ldots, \bw_{t,K}$ be a set of daily spatial knots in a spatial domain of interest, $\calD$, so that
-\begin{align*}
- P_{t,k} = \{\bs : k = \argmin_\ell || \bs_t - \bw_{t\ell} ||\}.
-\end{align*}
+Let $\bw_{t,1}, \ldots, \bw_{t,K}$ be a set of daily spatial knots in a spatial domain of interest, $\calD$, and $P_{tk}$ as defined in \ref{s:part}.
 
@@
 \begin{align}
-   Y_t(\bs) \mid z_{t}(\bs), \sigma^2(\bs), P_{t,k}, \alpha, \beta, \Theta &= X_t(\bs) \beta + \alpha |z_t(\bs)| + \sigma_t(\bs) v_t(\bs) \label{eq:hier}\\
-   z_t(\bs) &= z_{t,k} \text{ if } \bs \in P_{t,k}\\
-   \sigma^2_{t}(\bs) &= \sigma^2_{t,k} \text{ if } \bs \in P_{t,k}\\
+   Y_t(\bs) \mid z_{t}(\bs), \sigma^2(\bs), P_{tk}, \alpha, \beta, \Theta &= X_t(\bs) \beta + \alpha |z_t(\bs)| + \sigma_t(\bs) v_t(\bs) \label{eq:hier}\\
+   z_t(\bs) &= z_{tk} \text{ if } \bs \in P_{tk}\\
+   \sigma^2_{t}(\bs) &= \sigma^2_{tk} \text{ if } \bs \in P_{tk}\\
    \alpha &\sim N(0, 10)\\
    v_t(\bs) \mid \Theta &\sim \Matern(0, \Sigma)\\
-   z_{t,k} \mid z_{t-1, k}, \sigma^2_{tk} &\sim N(\phi_z z_{t-1, k}, \sigma^2_{t,k} (1 - \phi_z^2))\\
-   \sigma^{2*}_{t,k} \mid \sigma^{2*}_{t-1, k} &\sim N(\phi_\sigma \sigma^{2*}_{t-1, k}, (1 - \phi_\sigma^2))\\
-   \bw^*_{t,k} \mid \bw^*_{t-1, k} &\sim N_2(\phi_w \bw^*_{t-1, k}, (1 - \phi_w^2))
+   z_{tk} \mid z_{t-1, k}, \sigma^2_{tk} &\sim N(\phi_z z_{t-1, k}, \sigma^2_{tk} (1 - \phi_z^2))\\
+   \sigma^{2*}_{tk} \mid \sigma^{2*}_{t-1, k} &\sim N(\phi_\sigma \sigma^{2*}_{t-1, k}, (1 - \phi_\sigma^2))\\
+   \bw^*_{tk} \mid \bw^*_{t-1, k} &\sim N_2(\phi_w \bw^*_{t-1, k}, (1 - \phi_w^2))
 \end{align}
@@
 \section{Computation}\label{s:comp}
-The MCMC for this model is fairly straightforward.
 First, we impute values below the threshold.
-Then, we update $\Theta$ using random walk MH or Gibbs sampling when appropriate.
-Finally, we make spatial predictions using conditional multivariate normal results and the fact that the distribution of $Y_t(\bs) \mid \Theta, z_{tl}$ is the usual multivariate normal distribution with a \Matern spatial covariance structure.
+Then, we update $\Theta$ using Metropolis Hastings or Gibbs sampling when appropriate.
+Finally, we make spatial predictions using conditional multivariate normal results and the fact that the distribution of $Y_t(\bs) \mid \Theta, z(\bs)$ is the usual multivariate normal distribution with a \Matern spatial covariance structure.
 
@@
 After conditioning on $\alpha$, $z_t(\bs)$ and non-censored observations, $Y_t(\bs)$ has truncated normal full conditionals.
-So we sample $Y_t(\bs) \sim N_{(-\infty, T)}(\mu(\bs), \Sigma)$.
+So we sample $Y_t(\bs) \sim N_{(-\infty, T)}(\mu(\bs), \bSigma)$.
 After imputing the censored observations, we update the model parameters.
@@
 See Appendices A.1 and A.2 for details regarding the MCMC.
-The final step of the computation is to use Bayesian Kriging to generate a predictive distribution for $Y_t(\bs^*)$ at prediction location $(\bs^*)$.
+The final step of the computation is to use Bayesian Kriging to generate a predictive distribution for $Y_t(\bs^*)$ at prediction location $\bs^*$.
 This step is similar to the imputation for censored observations except that the full conditionals are no longer truncated at $T$.
@@
 \subsection{Design}\label{s:simdesign}
-For all simulation designs, we generate data from the model presented in Section \ref{s:part} using $n_s=130$ sites and $n_t=50$ independent days.
+For all simulation designs, we generate data from the model in Section \ref{s:part} using $n_s=144$ sites and $n_t=50$ independent days.
 The sites are generated Uniform$([0, 10] \times [0, 10])$.
@@
   \item Gaussian marginal, $K=1$ knot
-  \item $t$ marginal, $K=1$ knot
-  \item $t$ marginal, $K=5$ knots
-  \item skew-$t$ marginal, $K=1$ knots
-  \item skew-$t$ marginal, $K=5$ knots
+  \item $T$ marginal, $K=1$ knot
+  \item $T$ marginal, $K=5$ knots
+  \item Skew-$t$ marginal, $K=1$ knots
+  \item Skew-$t$ marginal, $K=5$ knots
   \item Max-stable.
 \end{enumerate}
-In the first five designs, the $v_t(\bs)$ terms are generated using a \Matern covariance with smoothness parameter, $\nu = 0.5$, and spatial range, $\rho = 0.1$.
+In the first five designs, the $v_t(\bs)$ terms are generated using a \Matern covariance with smoothness parameter $\nu = 0.5$ and spatial range $\rho = 0.1$.
 For the covariance matrices in designs 1 -- 5, the proportion of the variance accounted for by the spatial variation is $\gamma = 0.9$ while the proportion of the variance accounted for by the nugget effect is $0.1$.
@@
 In the sixth design, we generate from a spatial max-stable distribution \citep{Reich2012} with parameters $\mu = 1, \sigma=1, \xi=0.2$ and 144 spatial knots on a regular lattice in the square $[1, 9] \times [1, 9]$.
-In all six designs, the mean ($\mu(\bs)$) is assumed to be constant across space.
+\hl{I know I need more here about the max-stable method}
+In all six designs, the mean $\mu(\bs) = 10$ is assumed to be constant across space.
 
@@
   \item Gaussian marginal, $K=1$ knots
-  \item skew-$t$ marginal, $K=1$ knots, $T=-\infty$
-  \item $t$ marginal, $K=1$ knots, $T=q(0.80)$
-  \item skew-$t$ marginal, $K=5$ knots, $T=-\infty$
-  \item $t$ marginal, $K=5$ knots, $T=q(0.80)$
+  \item Skew-$t$ marginal, $K=1$ knots, $T=-\infty$
+  \item $T$ marginal, $K=1$ knots, $T=q(0.80)$
+  \item Skew-$t$ marginal, $K=5$ knots, $T=-\infty$
+  \item $T$ marginal, $K=5$ knots, $T=q(0.80)$
 \end{enumerate}
@@
 The residual variance terms have priors $\sigma^2_t(\bs) \sim \text{IG}(0.1, 0.1)$.
+The knots have priors $\bw \sim \text{Unif}(\calD)$.
+We do not fit the data using the max-stable methods from \citet{Reich2012} because of the time it takes.
 
@@
 The model was fit using the training set, and predictions were generated at the testing site locations.
-Because one of the primary goals of this model is to predict extreme events, we quantify use Brier scores to select the model that best fits the data \citep{Gneiting2007}.
+Because one of the primary goals of this model is to predict extreme events, we use Brier scores to select the model that best fits the data \citep{Gneiting2007}.
 The Brier score for predicting exceedance of a threshold $c$ is given by $[e(c) - P(c)]^2$ where $e(c) = I[y>c]$ is an indicator function indicating that a test set value, $y$, has exceeded the threshold, $c$, and $P(c)$ is the predicted probability of exceeding $c$.
+We average the Brier scores over all test sites and days.
 For the Brier score, a lower score indicates a better fit.
@@
 We compared the Brier scores for exceeding 11 different thresholds for each dataset.
-For each threshold, Brier score were calculated for each day and site in the testing set.
-The Brier scores were then averaged over all the days and sites to obtain a single Brier score for threshold $c$.
 The thresholds used for the Brier scores are extreme quantiles from the simulated data.
@@
 \section{Data analysis}\label{s:analysis}
-To illustrate this method, we consider the daily maximum 8-hour ozone measurements for July 2005 at 735 Air Quality System (AQS) monitoring sites in the eastern United States as the response.
+To illustrate this method, we consider the daily maximum 8-hour ozone measurements for July 2005 at 735 Air Quality System (AQS) monitoring sites in the eastern United States as the response (see Figure \ref{fig:ozone}).
 For each site, we also have covariate information containing the estimated ozone from the Community Multi-scale Air Quality (CMAQ) modeling system.
-Initially, we fit a linear regression assuming that the mean function can be expressed as
+Initially, we fit a linear regression assuming a mean function of
 \begin{align}
@@
 \end{align}
-The data from 10 July are shown in Figure \ref{fig:ozone} along with a Q-Q plot of the residuals compared to a skew-$t$ distribution with 8 d.f. and $\alpha = 1$.
+The data from July 10 are shown in Figure \ref{fig:ozone} along with a Q-Q plot of the residuals compared to a skew-$t$ distribution with 8 d.f. and $\alpha = 1$.
 \begin{center}
@@
 \end{center}
-The $\chi(h)$ and $\chi(t)$ plots indicate that the residuals still demonstrate spatial and temporal dependence (see Figure \ref{fig:chi-st})
+The $\widehat{\chi}(h)$ and $\widehat{\chi}(t)$ plots estimate the spatial and temporal dependence using sample data.
+For $\widehat{\chi}(h)$, we take the sample proportion of all pairs of sites within distance $h$ of one another that are above a threshold $T$.
+\hl{I know I need another sentence or two here about $\widehat{\chi}(t)$}
+indicate that the residuals still demonstrate spatial and temporal dependence (see Figure \ref{fig:chi-st})
 \begin{center}
@@
   \includegraphics[width=0.5\linewidth]{plots/chi-t-ozone.pdf}
-  \caption{$\chi(h)$ plot for the residuals. The vertical line indicates distance after which observations no longer demonstrate dependence (left) $\chi(t)$ plot for the residuals (right)}
+  \caption{$\widehat{\chi}(h)$ plot for the residuals.
+The vertical line indicates the distance after which observations no longer demonstrate dependence (left) $\widehat{\chi}(t)$ plot for the residuals (right)}
   \label{fig:chi-st}
@@
 where ${{\sigma^*_{tk}}^2}^{(r - 1)}$ is the value at MCMC iteration $r - 1$, and $s$ is a tuning parameter.
-In the case where there is no temporal dependence, we generate a candidate for $\sigma^2_{tk}$ using an IG$(a^*/s, b^*/s)$ candidate distribution where $a^* = n_{tk} / 2 + a$, $b^* = [Y_{tk}^T \Sigma^{-1}_{tk} Y_{tk}] / 2 + b$, $n_{tk}$ is the number of sites in partition $k$ on day $t$, and $Y_{tk}$ and $\Sigma^{-1}_{tk}$ are the observations and precision matrix for partition $k$ on day $t$.
+In the case where there is no temporal dependence, we generate a candidate for $\sigma^2_{tk}$ using an IG$(a^*/s, b^*/s)$ candidate distribution in an independence Metropolis Hastings update where $a^* = (n_{tk} + 1) / 2 + a$, $b^* = [Y_{tk}^T \Sigma^{-1}_{tk} Y_{tk} + z_{tk}^2] / 2 + b$, $n_{tk}$ is the number of sites in partition $k$ on day $t$, and $Y_{tk}$ and $\Sigma^{-1}_{tk}$ are the observations and precision matrix for partition $k$ on day $t$.
 The acceptance ratio is
@@
 \section*{Appendix A.3: Proof that $\lim_{h \rightarrow \infty} \pi(h) = 0$}
-\input{proof_dist.tex}
+Let $N(A)$ be the number of knots in $A$, the area between sites $\bs_1$ and $\bs_2$.
+Consider a spatial Poisson process with intensity $\mu(A)$.
+So,
+\begin{align*}
+  P[ N(A) = k] = \frac{ \mu(A)^k \exp\{ -\mu(A)\}}{k!}.
+\end{align*}
+Then for any finite $k$, $\lim_{h \rightarrow \infty} P[N(A) = k] = 0$ because $\lim_{h \rightarrow \infty} \mu(A) = \infty$.
+With each additional knot in $A$, the chance that $\bs_1$ and $\bs_2$ will be be in the same partition will decrease, because partition membership is defined by the closest knot to a site.
+Therefore, $\lim_{h \rightarrow \infty} \pi(h) = 0$.
 

### 2014-11-14 pot/LaTeX/skewnormal.tex
plot updates and sim study tweak

@@
 The paper is organized as follows.
-In section \ref{s:spatial}, we describe a spatial skew-$t$ process with partitioning and thresholding.
-In section \ref{s:temporal}, we expand the model to incorporate temporal dependence.
+Section \ref{s:skewt} is a brief review of the spatial skew-$t$ process.
+In section \ref{s:temporal}, we build upon the traditional skew-$t$ by incorporating censoring to focus on tails, partitioning to remove long-range asymptotic dependence, and extending the model to space-time data.
 The computing is described in section \ref{s:comp}.
@@
 
-\section{Spatial model}\label{s:spatial}
-To avoid bias in estimating tail parameters, we model censored data.
-Let
-\beq\label{Yt}
-  \tilde{Y}(\bs) = \delta(\bs) Y(\bs) + (1 - \delta(\bs)) T
-\eeq
-be the censored observation at site $\bs$ where $Y(\bs)$ is the uncensored observation, $\delta(\bs) = I[Y(\bs) > T]$, and $T$ is a pre-specified threshold value.
-Then, assuming the uncensored data are observations from a skew-$t$ process, we update values censored below the threshold using standard Bayesian missing data methods as described in Section \ref{s:comp}.
+\section{Spatial skew processes}\label{s:skewt}
+Many types of data demonstrate some level of skewness and therefore should be modeled with distributions that allow for non-symmetry.
+The skew-elliptical family of distributions provides models that are mathematically tractable while introducing a slant parameter $\alpha$ to account for non-symmetric data \citep{Genton2004}.
 
-\subsection{Skew-$t$ process}\label{s:data}
-Many types of data demonstrate some level of skewness, and therefore should be modeled with distributions that allow for non-symmetry.
-The skew-elliptical family of distributions provides models that are mathematically tractable while introducing additional paremeters to account for non-symmetric data \citep{Genton2004}.
-One member of this family, is the skew-$t$ distribution \citep[see Appendix A.5]{Azzalini2003}.
-\citet{Zhang2010} show that the skew-$t$ can be expressed as a hierarchical model
-\begin{align} \label{eq:fullmodel}
-  \bY &= \bX \bbeta + \alpha |z| + \sigma \bv
+\subsection{Skew-normal process}
+Let $U(\bs)$ be a normalized skew-normal process with slant $\alpha$. We write $U(\bs) \sim \text{SN}_d(\Omega, \alpha)$, and the density is given as
+\begin{align}
+  2 \phi_d (U(\bs); \Omega) \Phi(\alpha \bOne^T U(\bs))
 \end{align}
-where $\alpha \in \calR$ controls the skewness; $z \sim N(0, \sigma^2)$ is a random effect from a half-normal distribution (see appendix A.4); $\bv$ is a Gaussian process with mean zero, variance one, \Matern spatial correlation with spatial range $\rho$, smoothness $\nu$, and $\gamma$ is the proportion of variance accounted for by the spatial variation; and $\sigma^2 \sim \text{IG}(a, b)$.
-When marginalizing over the $z$ and $\sigma^2$ terms, the joint distribution for all sites becomes
+where $\phi_d$ is the density function for a $N_d(0, \Sigma)$ random variable, and $\Phi$ is a standard normal cumulative distribution function.
+There is an additive representation \citep{Zhang2010} to construct a standardized skew-normal process given by
+\begin{align}
+  U(\bs) = \alpha |z| + v(\bs)
+\end{align}
+where $\alpha \in \calR$ is a parameter controlling skew, $z \sim N(0, 1)$, and $v(\bs)$ is a Gaussian process with mean zero, variance one, \Matern spatial correlation with spatial range $\rho$, smoothness $\nu$, and $\gamma$ is the proportion of variance accounted for by the spatial variation.
+So,
+\begin{align}
+  \text{cor}(U(\bs), U(\bs)) = \gamma I(s = t) + (1 - \gamma)  \frac{ 1 }{ \Gamma(\nu) 2^{ \nu - 1}} \left( \sqrt{2\nu} \frac{ ||\bs - \bt|| }{ \rho } \right)^{\nu} K_{\nu} \left( \sqrt{2\nu} \frac{ ||\bs - \bt|| }{ \rho } \right).
+\end{align}
+Let $Y(\bs) = \bX(\bs)^T + \sigma U(\bs)$.
+Then, after marginalizing over the $z$ term, we find that
+\begin{align}
+  Y(\bs) \sim \text{SN}_d(\bX^T(\bs)\bbeta, \Omega, \alpha)
+\end{align}
+where $\Omega = \sigma^2 (\Sigma + \alpha^2 \bOne \bOne^T$).
+
+\subsection{Skew-$t$ process}
+If $\sigma^2 \sim \text{IG}(a, b)$, after over the $\sigma^2$ terms, we have a skew-$t$ process with distribution function
+
 \begin{align*}
@@
 Furthermore, if $\bY$ follows a multivariate skew-$t$ distribution, the marginal distributions also follow a skew-$t$ distribution \citep{Azzalini2003}.
+
+One member of this family, is the skew-$t$ distribution \citep[see Appendix A.5]{Azzalini2003}.
+\citet{Zhang2010} show that the skew-$t$ process can be expressed as a hierarchical model
+\begin{align} \label{eq:fullmodel}
+  Y(\bs) &= \bX(\bs)^T \bbeta + \alpha |z| + \sigma v(\bs)
+\end{align}
+where $\alpha \in \calR$ controls the skewness; $z \sim N(0, \sigma^2)$ is a normal random effect; $v(\bs)$ is a Gaussian process with mean zero, variance one, \Matern spatial correlation with spatial range $\rho$, smoothness $\nu$, and $\gamma$ is the proportion of variance accounted for by the spatial variation; and $\sigma^2 \sim \text{IG}(a, b)$.
+So,
+\begin{align}
+  \text{cor}(Y(\bs), Y(\bs)) = \gamma I(s = t) + (1 - \gamma)  \frac{ 1 }{ \Gamma(\nu) 2^{ \nu - 1}} \left( \sqrt{2\nu} \frac{ ||\bs - \bt|| }{ \rho } \right)^{\nu} K_{\nu} \left( \sqrt{2\nu} \frac{ ||\bs - \bt|| }{ \rho } \right)
+\end{align}
 
@@
 For Gaussian processes, $\chi(h) = 0$ regardless of the the distance, so they are not suitable for modeling spatially dependent extremes.
-However, for the skew-$t$ process described in \ref{s:data}, $\chi(h) > 0$ \citep[they give an expression that is the sum of two survival functions from the skew-$t$ distribution.
+However, for the skew-$t$ process described in \ref{s:skewt}, $\chi(h) > 0$ \citep[they give an expression that is the sum of two survival functions from the skew-$t$ distribution.
 The expression isn't terribly informative, but I can include it if you think it would be helpful.]{Padoan2011}.
 
-\subsection{Random daily partition}\label{s:part}
+\section{Spatiotemporal skew-$t$ model for extremes}\label{s:spatial}
+In this section, we describe the
+
+\subsection{Censoring to focus on the tails}
+To avoid bias in estimating tail parameters, we model censored data.
+Let
+\beq\label{Yt}
+  \tilde{Y}(\bs) = \left\{ \begin{array}{ll}
+      Y(\bs) \quad & \delta(\bs) = 1 \\
+      T & \delta(\bs) = 0
+  \end{array} \right.
+\eeq
+be the censored observation at site $\bs$ where $Y(\bs)$ is the uncensored observation, $\delta(\bs) = I[Y(\bs) > T]$, and $T$ is a pre-specified threshold value.
+Then, assuming the uncensored data $Y(\bs)$ are observations from a skew-$t$ process, we update values censored below the threshold using standard Bayesian missing data methods as described in Section \ref{s:comp}.
+
+\subsection{Partitioning to remove long-range asymptotic dependence}\label{s:part}
 One problem with the spatial skew-$t$ process is that all sites are asymptotically dependent regardless of their spatial separation.
@@
 
-\section{Accounting for temporal dependence} \label{s:temporal}
+\subsection{Extension to space-time data} \label{s:temporal}
 In a standard block-maxima approach, the temporal dependence is assumed to be negligible because blocks (e.g. yearly maxima) are taken to be large enough to assume independence.

### 2014-11-21 pot/LaTeX/skewnormal.tex
tweaks to results

@@
 
-\section{Spatial skew processes}\label{s:skewt}
+\section{Spatial skew processes}\label{s:spatialskew}
 Many types of data demonstrate some level of skewness and therefore should be modeled with distributions that allow for non-symmetry.
 The skew-elliptical family of distributions provides models that are mathematically tractable while introducing a slant parameter $\alpha$ to account for non-symmetric data \citep{Genton2004}.
+A brief review of the additive process by which a skew-$t$ process is created is given here.
 
-\subsection{Skew-normal process}
-Let $U(\bs)$ be a normalized skew-normal process with slant $\alpha$. We write $U(\bs) \sim \text{SN}_d(\Omega, \alpha)$, and the density is given as
+\subsection{Skew-normal process} \label{s:skewnorm}
+A skew-normal process can be constructed using an additive representation \citep{Zhang2010} given by
 \begin{align}
-  2 \phi_d (U(\bs); \Omega) \Phi(\alpha \bOne^T U(\bs))
+  U(\bs) = \lambda |z| + v(\bs)
 \end{align}
-where $\phi_d$ is the density function for a $N_d(0, \Sigma)$ random variable, and $\Phi$ is a standard normal cumulative distribution function.
-There is an additive representation \citep{Zhang2010} to construct a standardized skew-normal process given by
-\begin{align}
-  U(\bs) = \alpha |z| + v(\bs)
-\end{align}
-where $\alpha \in \calR$ is a parameter controlling skew, $z \sim N(0, 1)$, and $v(\bs)$ is a Gaussian process with mean zero, variance one, \Matern spatial correlation with spatial range $\rho$, smoothness $\nu$, and $\gamma$ is the proportion of variance accounted for by the spatial variation.
+where $\lambda \in \calR$ is a parameter controlling skew, $z \sim N(0, 1)$, and $v(\bs)$ is a Gaussian process with mean zero, variance one, \Matern spatial correlation with spatial range $\rho$, smoothness $\nu$, and $\gamma$ is the proportion of variance accounted for by the spatial variation.
 So,
 \begin{align}
-  \text{cor}(U(\bs), U(\bs)) = \gamma I(s = t) + (1 - \gamma)  \frac{ 1 }{ \Gamma(\nu) 2^{ \nu - 1}} \left( \sqrt{2\nu} \frac{ ||\bs - \bt|| }{ \rho } \right)^{\nu} K_{\nu} \left( \sqrt{2\nu} \frac{ ||\bs - \bt|| }{ \rho } \right).
+  \text{cor}(v(\bs), v(\bt)) = \gamma I(s = t) + (1 - \gamma)  \frac{ 1 }{ \Gamma(\nu) 2^{ \nu - 1}} \left( \sqrt{2\nu} \frac{ ||\bs - \bt|| }{ \rho } \right)^{\nu} K_{\nu} \left( \sqrt{2\nu} \frac{ ||\bs - \bt|| }{ \rho } \right).
 \end{align}
-Let $Y(\bs) = \bX(\bs)^T + \sigma U(\bs)$.
 Then, after marginalizing over the $z$ term, we find that
 \begin{align}
-  Y(\bs) \sim \text{SN}_d(\bX^T(\bs)\bbeta, \Omega, \alpha)
+  U(\bs) \sim \text{SN}_d(0, \bar{\bOmega}, \balpha)
 \end{align}
-where $\Omega = \sigma^2 (\Sigma + \alpha^2 \bOne \bOne^T$).
+where $\bar{\bOmega} = (\Sigma + \lambda^2 \bOne \bOne^T$), $\balpha = \lambda (1 + \lambda^2 \bOne^T \bSigma^{-1} \bOne)^{-1/2} \bOne^T \bSigma^{-1}$ is a vector of slant parameters, and SN is a skew-normal process with correlation matrix $\bar{\bOmega}$ \citep{Azzalini2013}.
 
-\subsection{Skew-$t$ process}
-If $\sigma^2 \sim \text{IG}(a, b)$, after over the $\sigma^2$ terms, we have a skew-$t$ process with distribution function
+\subsection{Skew-$t$ process} \label{s:skewt}
+Let
+\begin{align}
+  Y(\bs) = \bX(\bs)^T \bbeta + \sigma \lambda |z| + \sigma \bv(\bs)
+\end{align}
+where $\sigma^2 \sim \text{IG}(a, b)$ is an inverse gamma random variable.
+Then after marginalizing over both the $z$ and $\sigma$ terms, we find that
+\begin{align}
+  Y(\bs) \sim \text{ST}(\bX(\bs)^T \bbeta, \bOmega, \balpha, 2a) \label{eq:fullmodel}
+\end{align}
+where $\bOmega = \bomega \bar{\bOmega} \bomega$, $\bomega = \text{diag}\left(\frac{1}{ \sqrt{ab}}, \ldots, \frac{1}{\sqrt{ab}}\right)$, $\bar{\bOmega}$ and $\balpha$ are given in Section \ref{s:skewnorm}, and ST$(\mu(\bs), \bOmega, \balpha, 2a)$ represents a $d$-dimensional skew-$t$ process with location $\mu(\bs)$, correlation matrix $\bOmega$, slant parameters $\balpha$ and degrees of freedom $2a$.
 
-\begin{align*}
-  \bY \sim \text{skew-}t (\mu, \bSigma^*, \alpha, \text{df} = 2a)
-\end{align*}
-where $\mu$ is the location, $\bSigma^* = \frac{ b }{ a } \bSigma$, $\bSigma$ is a \Matern covariance matrix, and $\alpha \in \calR$ controls the skewness.
-The skew-$t$ process is desirable because of its flexible tail that is controlled by the skewness parameter, $\alpha$, and the degrees of freedom, $2a$.
-Furthermore, if $\bY$ follows a multivariate skew-$t$ distribution, the marginal distributions also follow a skew-$t$ distribution \citep{Azzalini2003}.
-
-One member of this family, is the skew-$t$ distribution \citep[see Appendix A.5]{Azzalini2003}.
-\citet{Zhang2010} show that the skew-$t$ process can be expressed as a hierarchical model
-\begin{align} \label{eq:fullmodel}
-  Y(\bs) &= \bX(\bs)^T \bbeta + \alpha |z| + \sigma v(\bs)
+This kind of process is desirable because of its flexible tail that is controlled by the skewness parameter $\lambda$ and degrees of freedom $2a$.
+Furthermore, the marginal distributions also follow a skew-$t$ distribution \citep{Azzalini2013}.
+If we partition a $d$-dimensional vector of observations $Y(\bs)$ into two components $Y(\bs_1)$ and $Y(\bs_2)$ of size $p$ and $d - p$ respectively, then after marginalizing over $Y(\bs_2)$, we get that
+\begin{align}
+  Y(\bs_1) \sim \text{ST}_p(\bX(\bs_1)^T \bbeta, \bOmega_{11}, \balpha_{1(2)}, 2a)
 \end{align}
-where $\alpha \in \calR$ controls the skewness; $z \sim N(0, \sigma^2)$ is a normal random effect; $v(\bs)$ is a Gaussian process with mean zero, variance one, \Matern spatial correlation with spatial range $\rho$, smoothness $\nu$, and $\gamma$ is the proportion of variance accounted for by the spatial variation; and $\sigma^2 \sim \text{IG}(a, b)$.
-So,
+where
 \begin{align}
-  \text{cor}(Y(\bs), Y(\bs)) = \gamma I(s = t) + (1 - \gamma)  \frac{ 1 }{ \Gamma(\nu) 2^{ \nu - 1}} \left( \sqrt{2\nu} \frac{ ||\bs - \bt|| }{ \rho } \right)^{\nu} K_{\nu} \left( \sqrt{2\nu} \frac{ ||\bs - \bt|| }{ \rho } \right)
+  \bX(\bs) &= \left( \begin{array}{c}
+    \bX(\bs_1) \\
+    \bX(\bs_2)
+  \end{array} \right), \quad
+  \bOmega = \left( \begin{array}{cc}
+    \bOmega_{11} & \bOmega_{12} \\
+    \bOmega_{21} & \bOmega_{22}
+  \end{array} \right), \quad
+  \balpha = \left(\begin{array}{c}
+    \balpha_1\\
+    \balpha_2
+  \end{array} \right),\\
+  \balpha_{1(2)} &= (1 = \balpha_2^T \bar{\bOmega}_{22 \cdot 1} \balpha)^{-1/2}(\balpha_1 + \bar{\bOmega}_{11}^{-1} \bar{\bOmega}_{12} \balpha_{2}),\\
+  \bar{\bOmega}_{11}^{-1} &= (\bar{\bOmega})^{-1}, \qquad \bar{\bOmega}_{22 \cdot 1} = \bar{\bOmega}_{22} - \bar{\bOmega}_{21} \bar{\bOmega}_{11}^{-1} \bar{\bOmega}_{12}.
 \end{align}
@@
 The $\chi$ statistic for the upper tail is given by
-\begin{align*}
-  \chi = \lim_{c \rightarrow \infty} \Pr(Y(\bs_1) > c | Y(\bs_2) > c).
-\end{align*}
+\begin{align}
+  \chi(\bs_1, \bs_2) = \lim_{c \rightarrow \infty} \Pr(Y(\bs_1) > c | Y(\bs_2) > c).
+\end{align}
 In a stationary spatial process, we can write the $\chi$ coefficient as
-\begin{align*}
+\begin{align}
   \chi(h) = \lim_{c \rightarrow \infty} \Pr(Y(0) > c | Y(h) > c)
-\end{align*}
+\end{align}
 where $h = ||\bs_1 - \bs_2||$.
@@
 For Gaussian processes, $\chi(h) = 0$ regardless of the the distance, so they are not suitable for modeling spatially dependent extremes.
-However, for the skew-$t$ process described in \ref{s:skewt}, $\chi(h) > 0$ \citep[they give an expression that is the sum of two survival functions from the skew-$t$ distribution.
-The expression isn't terribly informative, but I can include it if you think it would be helpful.]{Padoan2011}.
+One problem with the spatial skew-$t$ process is that all sites are asymptotically dependent regardless of their spatial separation.
+This occurs because all observations, both near and far, share the same $z$ and $\sigma^2$ terms.
+Even more problematic is that $\lim_{h \rightarrow \infty} \chi(h) > 0$ \citep[see Appendix A.5]{Padoan2011}.
 
 \section{Spatiotemporal skew-$t$ model for extremes}\label{s:spatial}
-In this section, we describe the
+In this section, we introduce censoring to focus on tail behavior, and a random partition to remove long-range asymptotic dependence.
 
@@
 \subsection{Partitioning to remove long-range asymptotic dependence}\label{s:part}
-One problem with the spatial skew-$t$ process is that all sites are asymptotically dependent regardless of their spatial separation.
-This occurs because all observations, both near and far, share the same $z$ and $\sigma^2$ terms.
-Even more problematic is that $\lim_{h \rightarrow \infty} \chi(h) \neq 0$.
-We handle this problem with a random partition similar to \citet{Kim2005} that allows $z$ and $\sigma^2$ to vary by site.
-
-Consider a set of spatial knots $\bw_{k} \sim \text{Uniform}(\calD)$ that define a random daily partition $P_{1}, \ldots, P_{K}$ of the spatial domain of interest $\calD \subset \calR^2$ such that
-\begin{align*}
-  P_{k} = \{ \bs : k = \argmin_\ell || \bs - \bw_{\ell} || \}.
-\end{align*}
+We handle the problem or long-range asymptotic dependence with a random partition similar to \citet{Kim2005} for Gaussian data that allows $z$ and $\sigma^2$ to vary by site.
 Then, the observation at each site becomes
 \begin{align}
-  Y(\bs) &= X(\bs) \bbeta + \alpha |z(\bs)| + \sigma v(\bs),
+  Y(\bs) &= \bX(\bs)^T \bbeta + \lambda |z(\bs)| + \sigma v(\bs).
 \end{align}
+To model spatial variation, consider a set of spatial knots $\bw_1, \ldots, \bw_K$ from a homogeneous Poisson process with rate $\lambda$ over spatial domain $\calD \in \calR^2$
+The knots define a random daily partition $P_{1}, \ldots, P_{K}$ of $\calD$ given by
+\begin{align*}
+  P_{k} = \{ \bs : k = \argmin_\ell || \bs - \bw_{\ell} || \},
+\end{align*}
 and, for $\bs \in P_{k}$,
 \begin{align}
-  z(\bs) &= z_{tk} \label{eq:sitez}\\
+  z(\bs) &= z_{k} \label{eq:sitez}\\
   \sigma(\bs) &= \sigma_{k} \label{eq:sitesig}.
 \end{align}
-So, within each partition, $\bY$ follows the distribution given in (\ref{eq:fullmodel}).
+So, within each partition, $Y(\bs)$ follows the distribution given in (\ref{eq:fullmodel}).
+Across partitions, the $Y(\bs)$ remain correlated via the correlation function for $v(\bs)$ because it spans the partitions.
 
-When incorporating the random daily partition, for two sites in the same partition, the $\chi$ statistic is $\chi(h) = \chi_{\text{skew-}t}(h)$.
+When incorporating the random daily partition, the $\chi$ statistic is for two sites $\bs$ in partition $k_s$ and $\bt$ in partition $k_t$ is
+\begin{align}
+  \chi(h) = I(k_s = k_t) \chi_{\text{skew-}t}(h) + I(k_s \neq k_t) \chi_{\text{Gaus}}(h).
+\end{align}
 However, for two sites in different partitions, the $\chi$ statistic should be the same as the Gaussian $\chi(h)$, because the only correlation between the two sites, is the correlation given by $\bSigma$.
@@
 \begin{align}
-   Y_t(\bs) \mid z_{t}(\bs), \sigma^2(\bs), P_{tk}, \alpha, \beta, \Theta &= X_t(\bs) \beta + \alpha |z_t(\bs)| + \sigma_t(\bs) v_t(\bs) \label{eq:hier}\\
+   Y_t(\bs) \mid z_{t}(\bs), \sigma^2(\bs), P_{tk}, \alpha, \beta, \Theta &= \bX_t(\bs)^T \beta + \lambda |z_t(\bs)| + \sigma_t(\bs) v_t(\bs) \label{eq:hier}\\
    z_t(\bs) &= z_{tk} \text{ if } \bs \in P_{tk}\\
    \sigma^2_{t}(\bs) &= \sigma^2_{tk} \text{ if } \bs \in P_{tk}\\
-   \alpha &\sim N(0, 10)\\
+   \lambda &\sim N(0, 10)\\
    v_t(\bs) \mid \Theta &\sim \Matern(0, \Sigma)\\
@@
 We can use Gibbs sampling to update $Y_t(\bs)$ for censored observations that are below the threshold $T$.
-After conditioning on $\alpha$, $z_t(\bs)$ and non-censored observations, $Y_t(\bs)$ has truncated normal full conditionals.
+After conditioning on $\lambda$, $z_t(\bs)$ and non-censored observations, $Y_t(\bs)$ has truncated normal full conditionals.
 So we sample $Y_t(\bs) \sim N_{(-\infty, T)}(\mu(\bs), \bSigma)$.
@@
 For designs 2 -- 4, $\sigma^2_{tk} \iid \text{IG}(3, 8)$
-For designs 1 -- 3, we set $\alpha = 0$.
-For designs four and five, $\alpha = 3$ was used, and the $z_t$ are generated as described in (\ref{eq:sitez}).
+For designs 1 -- 3, we set $\lambda = 0$.
+For designs four and five, $\lambda = 3$ was used, and the $z_t$ are generated as described in (\ref{eq:sitez}).
 In the sixth design, we generate from a spatial max-stable distribution \citep{Reich2012} with parameters $\mu = 1, \sigma=1, \xi=0.2$ and 144 spatial knots on a regular lattice in the square $[1, 9] \times [1, 9]$.
@@
 The spatial covariance parameters have priors $\log(\nu) \sim \text{N}(-1.2, 1)$, $\gamma \sim \text{U}(0, 1)$, $\log(\rho) \sim \text{N}(-2, 1)$.
-The skewness parameter has prior $\alpha \sim \text{N}(0, 2)$.
+The skewness parameter has prior $\lambda \sim \text{N}(0, 2)$.
 The residual variance terms have priors $\sigma^2_t(\bs) \sim \text{IG}(0.1, 0.1)$.
@@
     \right\} \times \left\{
-    \frac{ l[z_tk | {\sigma^2_{tk}}^{(c)}, \ldots] }
-         { l[z_tk | {\sigma^2_{tk}}^{(r - 1)}, \ldots] }
+    \frac{ l[z_{tk} | {\sigma^2_{tk}}^{(c)}, \ldots] }
+         { l[z_{tk} | {\sigma^2_{tk}}^{(r - 1)}, \ldots] }
     \right\} \times \left\{
@@
 \input{sigpost.tex}
-\input{alphapost.tex}
+\input{lambdapost.tex}
 
@@
 
-\section*{Appendix A.5: Multivariate skew-$t$ distribution}
-\input{mvskewt.tex}
+\section*{Appendix A.5: Skew-$t$ distribution}
+\subsection*{Univariate extended skew-$t$ distribution}
+We say that $Y$ follow a univariate extended skew-$t$ distribution with location $\xi \in \calR$, scale $\omega > 0$, skew parameter $\alpha \in \calR$, extended parameter $\tau \in \calR$, and degrees of freedom $\nu$ if has distribution function
+\begin{align}
+  f_{\text{EST}}(y) = \omega^{-1} \frac{ f_T (z; \nu) }{ F_T(\tau / \sqrt{ 1 + \alpha^2}; \nu)} F_T\left( (\alpha z + \tau) \sqrt{ \frac{ \nu + 1 }{ \nu + z^2}}; 0, 1, \nu + 1 \right)
+\end{align}
+where $f_T(t; \nu)$ is a univariate Student's $t$ with $\nu$ degrees of freedom, $F_T(t; \nu) = P(T < t)$, and $z = (y - \xi) / \omega$.
+In the case that $\tau = 0$, then $Y$ follows a univariate skew-$t$ distribution.
+\subsection*{Multivariate skew-$t$ distribution}
+If $\bZ \sim \text{ST}_d(0, \bar{\bOmega}, \balpha, \nu)$ is a $d$-dimensional skew-$t$ distribution, and $\bY = \xi + \bomega \bZ$, where $\bomega = \text{diag}(\omega_1, \ldots, \omega_d)$, then the density of $Y$ at $y$ is
+\begin{align*}
+  f_y(\by) = det(\bomega)^{-1} f_z(\bz)
+\end{align*}
+where
+\begin{align*}
+  f_z(\bz) &= 2 t_d(\bz; \bar{\bOmega}, \nu) T \left( \balpha^T \bz \sqrt{ \frac{\nu + d}{\nu + Q(\bz)} }; \nu + d\right) \\
+  \bz &= \bomega^{-1}(\by - \xi)
+\end{align*}
+where $t_d(\bz; \bar{\bOmega}, \nu)$ is a $d$-dimensional Student's $t$-distribution with scale matrix $\bar{\bOmega}$ and degrees of freedom $\nu$, $Q(z) = \bz^T \bar{\Omega}^{-1}\bz$ and $T(\cdot; \nu)$ denotes the univariate Student's $t$ distribution function with $\nu$ degrees of freedom \citep{Azzalini2013}.
 
+\subsection*{Extremal dependence}
+The $\chi$ statistic \citep{Padoan2011} is given by
+\begin{align}
+  \chi = \bar{F}_{\text{EST}}\left\{ \frac{(x_1^{1 / \nu} - \omega) \sqrt{\nu + 1} }{\sqrt{1 - \omega^2}}; 0, 1, \alpha_1, \tau_1, \nu+1 \right\} + \bar{F}_{\text{EST}}\left\{ \frac{ (x_2^{1 / \nu} - \omega) \sqrt{\nu + 1} }{ \sqrt{1 - \omega^2} }; 0, 1, \alpha_2, \tau_2, \nu + 1 \right\},
+\end{align}
+where $\bar{F}_{\text{EST}}$ is the univariate survival extended skew-$t$ function with zero location, unit scale, $\alpha_j = \alpha_i \sqrt{1 - \omega^2}$, $\tau_j = \sqrt{\nu + 1}(\alpha_j + \alpha_i \omega)$, and $x_j = F_T(\bar{\alpha}_i \sqrt{\nu + 1}; 0, 1, \nu) / F_T(\bar{\alpha}_j \sqrt{\nu + 1}; 0, 1, \nu)$ with $j = 1, 2$ and $i = 2, 1$ and where $\bar{\alpha}_j = (\alpha_j + \alpha_i \omega) / \sqrt{ 1 + \alpha_i^2 (1 - \omega^2)}$
 \bibliographystyle{rss}

### 2014-11-21 pot/code/analysis/ozone/US/cv5-51US.R

@@ -1,53 +1,0 @@
-options(warn=2)
-library(fields)
-library(SpatialTools)
-library(mvtnorm)
-
-rm(list=ls())
-load('cv-setup-us.RData')
-source('../../../R/mcmc.R')
-source('../../../R/auxfunctions.R')
-
-setting <- 51
-method <- "t"
-nknots <- 2
-keep.knots <- F
-threshold <- 0
-tau.init <- 0.5
-thresh.quant <- F
-skew <- T
-outputfile <- paste("cv5-", setting, "US.RData", sep="")
-
-start <- proc.time()
-
-fit <- vector(mode="list", length=2)
-
-for(val in 1:2){
-	
-	set.seed(setting*100 + val)
-	
-	cat("CV", val, "started \n")
-	val.idx <- cv.lst[[val]]
-	y.o <- Y[-val.idx,]
-	X.o <- X[-val.idx, , ]
-	S.o <- S[-val.idx,]
-	
-	y.p <- Y[val.idx,]
-	X.p <- X[val.idx, , ]
-	S.p <- S[val.idx,]
-	
-	tic.set <- proc.time()
-	fit[[val]] <- mcmc(y=y.o, s=S.o, x=X.o, x.pred=X.p, s.pred=S.p,
-	                   method=method, skew=skew, keep.knots=keep.knots,
-	                   thresh.all=threshold, thresh.quant=thresh.quant, nknots=nknots, 
-                       iters=30000, burn=25000, update=100, iterplot=T,
-                       beta.init=beta.init, tau.init=tau.init, rho.init=1,
-                       nu.init=0.5, alpha.init=0.5)
-	toc.set <- proc.time()
-	time.set <- (toc.set - tic.set)[3]
-	
-	elap.time.val <- (proc.time() - start)[3]
-	avg.time.val <- elap.time.val / val
-	cat("CV", val, "finished", round(avg.time.val, 2), "per dataset \n")
-	save(fit, file=outputfile)
-}

### 2014-11-21 pot/code/analysis/simstudy/results.R

@@ -196,10 +196,10 @@ load("simdata.RData")
 ns <- dim(y)[1]
 nt <- dim(y)[2]
 nsets <- 5
-ngroups <- 10
+ngroups <- 3
 nsettings <- dim(y)[4]
 nmethods <- 5
-obs <- rep(c(T, F), 100)[1:ns]
+obs <- c(rep(T, 100), rep(F, 44))
 
 setting <- 1
 filename <- paste("scores", setting, ".RData", sep="")
@@ -229,8 +229,8 @@ rho.all <- array(NA, dim=c(length(intervals), (nsets * ngroups), nmethods, nsett
 nu.all <- array(NA, dim=c(length(intervals), (nsets * ngroups), nmethods, nsettings))
 alpha.all <- array(NA, dim=c(length(intervals), (nsets * ngroups), nmethods, nsettings))
 # not all methods use skew or multiple partitions
-z.alpha.all <- array(NA, dim=c(length(intervals), (nsets * ngroups), 4, nsettings))
-avgparts.all <- array(NA, dim=c(length(intervals), (nsets * ngroups), 2, nsettings))
+z.alpha.all <- array(NA, dim=c(length(intervals), (nsets * ngroups), 5, nsettings))
+avgparts.all <- array(NA, dim=c(length(intervals), (nsets * ngroups), 3, nsettings))
 
 iters <- 20000; burn <- 10000
 for (setting in 1:nsettings) {

### 2014-11-21 pot/code/analysis/simstudy/results1.R

@@ -26,7 +26,7 @@ load("simdata.RData")
 ns <- dim(y)[1]
 nt <- dim(y)[2]
 nsets <- 5
-ngroups <- 10
+ngroups <- 3
 nsettings <- dim(y)[4]
 nmethods <- 5
 obs <- c(rep(T, 100), rep(F, 44))
@@ -59,8 +59,8 @@ rho <- array(NA, dim=c(length(intervals), (nsets * ngroups), nmethods, nsettings
 nu <- array(NA, dim=c(length(intervals), (nsets * ngroups), nmethods, nsettings))
 alpha <- array(NA, dim=c(length(intervals), (nsets * ngroups), nmethods, nsettings))
 # not all methods use skew or multiple partitions
-z.alpha <- array(NA, dim=c(length(intervals), (nsets * ngroups), 4, nsettings))
-avgparts <- array(NA, dim=c(length(intervals), (nsets * ngroups), 2, nsettings))
+z.alpha <- array(NA, dim=c(length(intervals), (nsets * ngroups), 5, nsettings))
+avgparts <- array(NA, dim=c(length(intervals), (nsets * ngroups), 3, nsettings))
 
 iters <- 20000; burn <- 10000
 
@@ -167,13 +167,42 @@ for (group in 1:ngroups) {
       nu[, set.idx, 5, setting] <- quantile(fit$nu, probs=intervals)
       alpha[, set.idx, 5, setting] <- quantile(fit$alpha, probs=intervals)
       z.alpha[, set.idx, 4, setting] <- quantile(fit$z.alpha, probs=intervals)
-      avgparts[, set.idx, 1, setting] <- mean(fit$avgparts)
+      avgparts[, set.idx, 2, setting] <- mean(fit$avgparts)
     }
     
     cat("dataset", dataset, "\n")
     rm(fit, fit.1)
     gc()
+    
+    dataset <- paste(setting,"-d-",group,".RData", sep="")
+    load(dataset)
+    for(d in 1:nsets){ 
+      set.idx <- (group - 1) * 5 + d
+      thresholds <- quantile(y[, , set.idx, setting], probs=probs, na.rm=T)
+      validate <- y[!obs, , set.idx, setting]
+      
+      fit <- fit.1[[d]]  # skew t-5 (T = 0.90)
+      pred <- fit$yp
+      quant.score[, set.idx, 6, setting] <- QuantScore(pred, probs, validate) 
[…]

### 2014-12-07 pot/LaTeX/skewnormal.tex
fix typo in mh.phi.tau adjustment

--- 2014-11-10 pot/LaTeX/skewnormal.tex
@@
 
-\section{Spatial model}\label{s:spatial}
+\section{Spatial skew processes}\label{s:spatialskew}
+Many types of data demonstrate some level of skewness and therefore should be modeled with distributions that allow for asymmetry.
+The skew-elliptical family of distributions provides models that are mathematically tractable while introducing a slant parameter to account for non-symmetric data \citep{Genton2004}.
+A brief review of the additive process by which a skew-$t$ process is created is given here.
+
+\subsection{Skew-$t$ process} \label{s:skewt}
+Let $Y(\bs)$ be the observation at spatial location $\bs = (s_1, s_2)$.
+The spatial skew-$t$ process can be written
+\begin{align}
+  Y(\bs) = \bX(\bs)^T \bbeta + \lambda \sigma |z| + \sigma v(\bs)
+\end{align}
+where $\bX(\bs)$ is a set of spatial covariates at site $\bs$, $\bbeta$ is a set of regression parameters, $\lambda \in \calR$ is a parameter controlling skew, $z \sim N(0, 1)$, $\sigma^2 \sim \text{IG}(a, b)$ is an inverse gamma random variable, and $v(\bs)$ is a spatial Gaussian process with mean zero and variance one.
+It should be noted that the spatial-$t$ process is a special case of the skew-$t$ process when $\lambda = 0$.
+A common spatial correlation function is the \Matern with
+\begin{align}
+  \text{cor}(v(\bs), v(\bt)) = \gamma I(\bs = \bt) + (1 - \gamma)  \frac{ 1 }{ \Gamma(\nu) 2^{ \nu - 1}} \left( \sqrt{2\nu} \frac{ h }{ \rho } \right)^{\nu} K_{\nu} \left( \sqrt{2\nu} \frac{ h }{ \rho } \right) \label{eq:matern}
+\end{align}
+where $\rho$ is the spatial range, $\nu$ is the smoothness, $\gamma$ is the proportion of variance accounted for by the spatial variation, $K_\nu$ is a modified Bessel function of the second time, and $h = || \bs - \bt ||$.
+
+Let $\bY = [Y(\bs_1), \ldots, Y(\bs_n)]^T$ be a set of observations at a finite collection of locations $\bs_1, \ldots, \bs_n$.
+After marginalizing over both $z$ and $\sigma$,
+\begin{align}
+  \bY \sim \text{ST}_n(\bX \bbeta, \bOmega, \balpha, 2a), \label{eq:fullmodel}
+\end{align}
+that is, $\bY$ follows an $n$-dimensional skew-$t$ process with location $\bX \bbeta$, correlation matrix $\bOmega$, slant parameters $\balpha$ and degrees of freedom $2a$ where $\bX = [\bX(\bs_1)^T, \ldots, \bX(\bs_n)^T]$, $\bOmega = \bomega \bar{\bOmega} \bomega$, $\bomega = \text{diag}\left(\frac{1}{ \sqrt{ab}}, \ldots, \frac{1}{\sqrt{ab}}\right)$, $\bar{\bOmega} = (\bSigma + \lambda^2 \bOne \bOne^T$), $\bSigma$ is a positive definite correlation matrix, $\balpha = \lambda (1 + \lambda^2 \bOne^T \bSigma^{-1} \bOne)^{-1/2} \bOne^T \bSigma^{-1}$ is a vector of slant parameters.
+This process is desirable because of its flexible tail that is controlled by the skewness parameter $\lambda$ and degrees of freedom $2a$.
+Furthermore, the marginal distributions at each location also follow a univariate skew-$t$ distribution \citep{Azzalini2013}.
+
+\subsection{Extremal dependence}
+One measure of extremal dependence is the $\chi$ statistic \citep{Padoan2011}.
+The $\chi$ statistic for the upper tail is given by $\lim_{c \rightarrow \infty} \Pr(Y(\bs_1) > c | Y(\bs_2) > c)$.
+For a stationary spatial process, we can write the $\chi$ coefficient as
+\begin{align}
+  \chi(h) = \lim_{c \rightarrow \infty} \Pr(Y(\bs) > c | Y(\bt) > c).
+\end{align}
+If $\chi(h) = 0$, then observations are asymptotically independent at distance $h$.
+For Gaussian processes, $\chi(h) = 0$ regardless of the distance, so they are not suitable for modeling spatially-dependent extremes.
+Unlike the Gaussian process, the skew-$t$ process is asymptotically dependent.
+However, one problem with the spatial skew-$t$ process is that $\lim_{h \rightarrow \infty} \chi(h) > 0$.
+This occurs because all observations, both near and far, share the same $z$ and $\sigma$ terms.
+Therefore, the skew-$t$ process is not ideal for spatial analysis of large geographic regions where we expect only local spatial dependence.
+The explicit expression for $\chi(h)$ \citep{Padoan2011} and a proof of this are given in Appendix \ref{a:skewt}.
+
+\section{Spatiotemporal skew-$t$ model for extremes}\label{s:spatial}
+In this section, we propose extensions to the skew-$t$ process to model spatial extremes over a large geographic region by introducing censoring to focus on tail behavior and a random partition similar to \citet{Kim2005} to remove long-range asymptotic dependence.
+
+\subsection{Censoring to focus on the tails}
 To avoid bias in estimating tail parameters, we model censored data.
@@
 \beq\label{Yt}
-  \tilde{Y}(\bs) = \delta(\bs) Y(\bs) + (1 - \delta(\bs)) T
+  \tilde{Y}(\bs) = \left\{ \begin{array}{ll}
+      Y(\bs) \quad & \delta(\bs) = 1 \\
+      T & \delta(\bs) = 0
+  \end{array} \right.
 \eeq
 be the censored observation at site $\bs$ where $Y(\bs)$ is the uncensored observation, $\delta(\bs) = I[Y(\bs) > T]$, and $T$ is a pre-specified threshold value.
-Then, assuming the uncensored data are observations from a skew-$t$ process, we update values censored below the threshold using standard Bayesian missing data methods as described in Section \ref{s:comp}.
+Then, assuming the uncensored data $Y(\bs)$ are observations from a skew-$t$ process, we update values censored below the threshold using standard Bayesian missing data methods as described in Section \ref{s:comp}.
 
-\subsection{Skew-$t$ process}\label{s:data}
-Many types of data demonstrate some level of skewness, and therefore should be modeled with distributions that allow for non-symmetry.
-The skew-elliptical family of distributions provides models that are mathematically tractable while introducing additional paremeters to account for non-symmetric data \citep{Genton2004}.
-One member of this family, is the skew-$t$ distribution \citep[see Appendix A.5]{Azzalini2003}.
-\citet{Zhang2010} show that the skew-$t$ can be expressed as a hierarchical model
-\begin{align} \label{eq:fullmodel}
-  \bY &= \bX \bbeta + \alpha |z| + \sigma \bv
+\subsection{Partitioning to remove long-range asymptotic dependence}\label{s:part}
+We handle the problem of long-range asymptotic dependence with a random partition model.
+As discussed in Section \ref{s:spatialskew}, the source of long-range dependence is the shared $z$ and $\sigma$.
+Therefore, to alleviate this dependence, we allow $z$ and $\sigma$to vary by site.
+Then, the model becomes
+\begin{align} \label{eq:partition}
+  Y(\bs) &= \bX(\bs)^T \bbeta + \lambda \sigma(\bs) |z(\bs)| + \sigma(\bs) v(\bs).
 \end{align}
-where $\alpha \in \calR$ controls the skewness; $z \sim N(0, \sigma^2)$ is a random effect from a half-normal distribution (see appendix A.4); $\bv$ is a Gaussian process with mean zero, variance one, \Matern spatial correlation with spatial range $\rho$, smoothness $\nu$, and $\gamma$ is the proportion of variance accounted for by the spatial variation; and $\sigma^2 \sim \text{IG}(a, b)$.
-When marginalizing over the $z$ and $\sigma^2$ terms, the joint distribution for all sites becomes
-\begin{align*}
-  \bY \sim \text{skew-}t (\mu, \bSigma^*, \alpha, \text{df} = 2a)
-\end{align*}
-where $\mu$ is the location, $\bSigma^* = \frac{ b }{ a } \bSigma$, $\bSigma$ is a \Matern covariance matrix, and $\alpha \in \calR$ controls the skewness.
-The skew-$t$ process is desirable because of its flexible tail that is controlled by the skewness parameter, $\alpha$, and the degrees of freedom, $2a$.
-Furthermore, if $\bY$ follows a multivariate skew-$t$ distribution, the marginal distributions also follow a skew-$t$ distribution \citep{Azzalini2003}.
-
-\subsection{Extremal dependence}
-One measure of extremal dependence is the $\chi$ statistic \citep{Padoan2011}.
-The $\chi$ statistic for the upper tail is given by
-\begin{align*}
-  \chi = \lim_{c \rightarrow \infty} \Pr(Y(\bs_1) > c | Y(\bs_2) > c).
-\end{align*}
-In a stationary spatial process, we can write the $\chi$ coefficient as
-\begin{align*}
-  \chi(h) = \lim_{c \rightarrow \infty} \Pr(Y(0) > c | Y(h) > c)
-\end{align*}
-where $h = ||\bs_1 - \bs_2||$.
-If $\chi(h) = 0$, then observations are asymptotically independent at distance $h$.
-For Gaussian processes, $\chi(h) = 0$ regardless of the the distance, so they are not suitable for modeling spatially dependent extremes.
-However, for the skew-$t$ process described in \ref{s:data}, $\chi(h) > 0$ \citep[they give an expression that is the sum of two survival functions from the skew-$t$ distribution.
-The expression isn't terribly informative, but I can include it if you think it would be helpful.]{Padoan2011}.
-
-\subsection{Random daily partition}\label{s:part}
-One problem with the spatial skew-$t$ process is that all sites are asymptotically dependent regardless of their spatial separation.
-This occurs because all observations, both near and far, share the same $z$ and $\sigma^2$ terms.
-Even more problematic is that $\lim_{h \rightarrow \infty} \chi(h) \neq 0$.
-We handle this problem with a random partition similar to \citet{Kim2005} that allows $z$ and $\sigma^2$ to vary by site.
-
-Consider a set of spatial knots $\bw_{k} \sim \text{Uniform}(\calD)$ that define a random daily partition $P_{1}, \ldots, P_{K}$ of the spatial domain of interest $\calD \subset \calR^2$ such that
+Let $\bw = (w_1, w_2)$ be the location of a spatial knot.
+To model spatial variation, consider a set of spatial knots $\bw_1, \ldots, \bw_K$ from a homogeneous Poisson process with intensity $\mu$ over spatial domain $\calD \in \calR^2$.
+The knots define a random daily partition of $\calD$ by subregions $P_{1}, \ldots, P_{K}$ defined as
 \begin{align*}
@@
 \end{align*}
-Then, the observation at each site becomes
+All $z(\bs)$ and $\sigma(\bs)$ for sites in subregion $k$ are assigned common values
 \begin{align}
-  Y(\bs) &= X(\bs) \bbeta + \alpha |z(\bs)| + \sigma v(\bs),
+  z(\bs) &= z_{k} \label{eq:sitez}\\
+  \sigma(\bs) &= \sigma_{k} \label{eq:sitesig},
 \end{align}
-and, for $\bs \in P_{k}$,
+and the $z_k$ and $\sigma^2_k$ are distributed as $z_k \iid N(0, 1)$ and $\sigma^2 \iid$ IG$(a, b)$.
+So, within each partition, $Y(\bs)$ follows the spatial skew-$t$ process defined in Section \ref{s:spatialskew}.
+Across partitions, the $Y(\bs)$ remain correlated via the correlation function for $v(\bs)$ because it spans the partitions.
+
+When incorporating the random daily partition, conditional on knots $\bw_1, \ldots, \bw_K$, the $\chi$ statistic for two sites $\bs$ and $\bt$ in partitions $k_s$ and $k_t$ respectively is
 \begin{align}
-  z(\bs) &= z_{tk} \label{eq:sitez}\\
-  \sigma(\bs) &= \sigma_{k} \label{eq:sitesig}.
+  \chi(h) &= I(k_s = k_t) \chi_{\text{skew-}t}(h) + I(k_s \neq k_t) \chi_{\text{Gaus}}(h) \nonumber \\
+         &= I(k_s = k_t) \chi_{\text{skew-}t}(h)
 \end{align}
-So, within each partition, $\bY$ follows the distribution given in (\ref{eq:fullmodel}).
-
-When incorporating the random daily partition, for two sites in the same partition, the $\chi$ statistic is $\chi(h) = \chi_{\text{skew-}t}(h)$.
-However, for two sites in different partitions, the $\chi$ statistic should be the same as the Gaussian $\chi(h)$, because the only correlation between the two sites, is the correlation given by $\bSigma$.
+where $\chi_{\text{skew-}t}(h)$ is the $\chi$ statistic for a skew-$t$ process, $\chi_{\text{Gaus}}(h)$ is the $\chi$ statistic for a Gaussian process, and $h = ||\bs - \bt||$.
+Therefore, sites in different subregions are asymptotically independent.
+Marginally, over the knots $\bw_1, \ldots, \bw_K$, $\chi(h) = \pi(h) \chi_{\text{skew}-t}(h)$, where $\pi(h) = \Pr(k_s = k_t)$ is the probability that two sites separated by distance $h$ are in the same partition.
 So, to show that $\lim_{h \rightarrow \infty} \chi(h) = 0$, we need only know that $\lim_{h \rightarrow \infty} \pi(h) = 0$
-where $\pi(h)$ is the probability that two sites separated by distance $h$ are in the same partition.
-A proof of this is given in Appendix A.3.
+A proof of this is given in Appendix \ref{a:proofsamepartition}.
 
@@
 These plots show how the partitioning allows for short range extremal dependence while reducing extremal dependence as the distance between sites increases.
+
 \begin{figure}
   \includegraphics[width=\linewidth]{plots/sim-chi.pdf}
-  \caption{$\chi(h)$ for simulated datasets.}
+  \caption{$\chi(h)$ for simulated datasets. The dashed line indicates the $\chi(h)$ we would expect under asymptotic independence.}
   \label{fig:simchi}
@@
 
-\section{Accounting for temporal dependence} \label{s:temporal}
-In a standard block-maxima approach, the temporal dependence is assumed to be negligible because blocks (e.g. yearly maxima) are taken to be large enough to assume independence.
-However, when using daily meausrements, the assumption of temporal independence is no longer appropriate.
-There are several places where temporal dependence could be included, including the residual $v(\bs)$.
-However, we choose to allow for temporal dependence in the $\bw$, $z(\bs)$, and $\sigma(\bs)$ terms because these terms derive the tail behavior.
-Specifically, we incorporate an AR(1) time series.
-We first, transform the spatial knots from $\calD$ to $\calR^2$ as follows.
+\subsection{Extension to space-time data} \label{s:temporal}
+When using daily measurements, the assumption of temporal independence is inappropriate.
+There are several places where temporal dependence could be incorporated in the model, including the residual $v_t(\bs)$.
+However, we choose to allow for temporal dependence in the $\bw$, $z$, and $\sigma$ terms because these terms dictate the tail behavior which is our primary focus.
+In this section, we extend (\ref{eq:partition}) to the spatiotemporal setting.
 Let
 \begin{align*}
-  \bw^*_k = \Phi^{-1}_2\left[ \frac{ (\bw_k - \min(\bs))}{ \text{range}(\bs) } \right].
+  Y_t(\bs) = \bX_t(\bs)^T \bbeta + \lambda \sigma_t(\bs) |z_t(\bs)| + \sigma_t(\bs) v_t(\bs),
 \end{align*}
-\hl{What's the best way to write the $\bw$ transformation as 1-D? $w_x$ and $w_y$ or $w_1$ and $w_2$?}
-where $\Phi_2$ represents a bivariate standard normal density function.
+where $t \in \{1, \ldots, T\}$ denotes the day of each observation.
+Let \hbox{$\bw_{tk} = (w_{tk1}, w_{tk2})$} be a spatial knot on day $t$, and let $w_{t1}, \ldots, w_{tK}$ be a collection of spatial knots on day $t$.
+As in section \ref{s:part}, these knots define a daily partition $P_{t1}, \ldots, P_{tK}$, and for $\bs \in P_{tk}$,
+\begin{align}
+  z_t(\bs) = z_{tk}\\
+  \sigma_{t}(\bs) = \sigma_{tk}.
+\end{align}
+
+We use an AR(1) time series model for $w_{tk}$, $z_{tk}$, and $\sigma_{tk}$.
+The time series model must be specified after a transformation to preserve the skew-$t$ process at each time point.
+We first transform the spatial knots from $\calD$ to $\calR^2$ as follows.
+Let
+\begin{align}
+  w^*_{tki} = \Phi^{-1}\left[ \frac{ (w_{tki} - \min(s_i))}{ \text{range}(s_i) } \right], \quad i = 1, 2
+\end{align}
+where $\Phi$ is a univariate standard normal density function.
 Then $\bw^*_k \in \calR^2$.
@@
 Let
-\begin{align*}
+\begin{align}
   \sigma^{2*}_t(\bs) =\Phi^{-1}\left\{ \text{IG}[\sigma^2_t(\bs)] \right\}
-\end{align*}
-where $\Phi$ is a univariate standard normal density function, and IG is the distribution function for an IG$(a, b)$ random variable.
-Then, when $t > 1$ the time series is modeled as
-\begin{align*}
-  \bw^*_{1, k} & \sim N_2(0, 1)\\
-  z_{1k} & \sim N(0, \sigma^2_{1k})\\
-  \sigma^{2*}_{1k} &\sim N(0, 1)\\
+\end{align}
+where IG is the distribution function for an IG$(a, b)$ random variable.
+The AR(1) process for each tail parameter is $\bw^*_{1k} \sim N_w(0, 1)$, $z_{1k} \sim N(0, \sigma^2_{1k})$, $\sigma^{2*}_{1k} \sim N(0, 1)$, and for $t > 1$ the time series is modeled as
+\begin{align}
   \bw^*_{tk} | \bw^*_{t-1, k} &\sim N_2\left[\phi_w \bw^*_{t-1, k}, (1 - \phi_w^2) \right] \\
@@
   \sigma^{2*}_{tk} | \sigma^{2*}_{t-1, k} &\sim N \left[\phi_\sigma \sigma^{2*}_{t-1, k}, (1 - \phi_\sigma^2) \right]
-\end{align*}
+\end{align}
 where $|\phi_w| < 1$, $0 < \phi_z < 1$, and $|\phi_\sigma| < 1$.
-These are stationary time series models with marginal distributions
-\begin{align*}
-  \bw^*_{k} & \sim N_2(0, 1)\\
-  z_{k} & \sim N(0, \sigma^2_{k})\\
-  \sigma^{2*}_{k} &\sim N(0, 1).
-\end{align*}
-For each day, the model is identical to the spatial-only model in Seciton \ref{s:part}.
+We restrict $\phi_z$ to be positive because in the skew-$t$ distribution, we use $|z_t(\bs)|$.
+These are stationary time series models with marginal distributions $\bw^*_{k} \sim N_2(0, 1)$, $z_{k} \sim N(0, \sigma^2_{k})$, $\sigma^{2*}_{k} \sim N(0, 1)$.
+After transformation back to the original space, $\bw_{tk} \sim \text{Unif}(\calD)$, $\sigma^2 \sim \text{IG}(a, b)$.
+For each day, the model is identical to the spatial-only model in Section \ref{s:part} by construction.
 
 \subsection{Hierarchical model}\label{s:hier}
-Conditioned on $z_{tk}(\bs)$, $\sigma^2_{t,k}(\bs)$, and $P_{t,k}$, the marginal distributions are Gaussian and the joint distribution multivariate Gaussian.
+Conditioned on $z_{tk}(\bs)$, $\sigma^2_{tk}(\bs)$, and $P_{tk}$, the marginal distributions are Gaussian and the joint distribution multivariate Gaussian.
 However, we do not fix the partitions, they are treated as unknown and updated in the MCMC.
 We model this with a Bayesian hierarchical model as follows.
-Let $\bw_{t,1}, \ldots, \bw_{t,K}$ be a set of daily spatial knots in a spatial domain of interest, $\calD$, and $P_{tk}$ as defined in \ref{s:part}.
+Let $\bw_{t1}, \ldots, \bw_{tK}$ be a set of daily spatial knots in a spatial domain of interest, $\calD$, and $P_{tk}$ as defined in \ref{s:part}.
 
@@
 \end{align}
-where $\Theta = \{\rho, \nu, \gamma\}$; \mbox{$k = \argmin_\ell ||\bs - \bw_\ell ||$}; and $\Sigma$ is a \Matern covariance matrix with variance one, spatial range $\rho$, smoothness $\nu$, and the proportion of variance accounted for by the spatial variation is $\gamma$.
+where $\Theta = \{\rho, \nu, \gamma\}$; \mbox{$k = \argmin_\ell ||\bs - \bw_\ell ||$}; and $\Sigma$ is a \Matern covariance matrix as described in Section \ref{s:skewt}.
 
@@
 We can use Gibbs sampling to update $Y_t(\bs)$ for censored observations that are below the threshold $T$.
-After conditioning on $\alpha$, $z_t(\bs)$ and non-censored observations, $Y_t(\bs)$ has truncated normal full conditionals.
-So we sample $Y_t(\bs) \sim N_{(-\infty, T)}(\mu(\bs), \bSigma)$.
+After conditioning on $\lambda$, $z_t(\bs)$ and non-censored observations, $Y_t(\bs)$ has truncated normal full conditionals.
+So we sample $Y_t(\bs) \sim N_{(-\infty, T)}(\bX \bbeta, \bSigma)$.
 After imputing the censored observations, we update the model parameters.
@@
 For designs 2 -- 4, $\sigma^2_{tk} \iid \text{IG}(3, 8)$
-For designs 1 -- 3, we set $\alpha = 0$.
-For designs four and five, $\alpha = 3$ was used, and the $z_t$ are generated as described in (\ref{eq:sitez}).
+For designs 1 -- 3, we set $\lambda = 0$.
+For designs four and five, $\lambda = 3$ was used, and the $z_t$ are generated as described in (\ref{eq:sitez}).
 In the sixth design, we generate from a spatial max-stable distribution \citep{Reich2012} with parameters $\mu = 1, \sigma=1, \xi=0.2$ and 144 spatial knots on a regular lattice in the square $[1, 9] \times [1, 9]$.
 \hl{I know I need more here about the max-stable method}
-In all six designs, the mean $\mu(\bs) = 10$ is assumed to be constant across space.
+In all six designs, the mean $\bX \bbeta = 10$ is assumed to be constant across space.
 
@@
 \begin{align}
-  \mu_t(\bs) = \beta_0 + \beta_1 \cdot \text{CMAQ}_t(\bs). \label{eq:datamean}
+  \bX \bbeta = \beta_0 + \beta_1 \cdot \text{CMAQ}_t(\bs). \label{eq:datamean}
 \end{align}
@@
 
-\section*{Appendix A.1: MCMC details}
+\appendix
+\section{Appendices}
+\subsection{MCMC details} \label{a:mcmc}
 The MCMC sampling for the model \ref{s:hier} is done using {\tt R} (http://www.r-project.org).
-Whenever possible, we select conjugate priors (see Appendix A.2); however, for some of the parameters, no conjugate prior distributions exist.
+Whenever possible, we select conjugate priors (see Appendix \ref{a:posterior}); however, for some of the parameters, no conjugate prior distributions exist.
 When no conjugate prior distribution exists, we use a random walk Metropolis Hastings update step.
@@
 
-\subsection*{Spatial knot locations}
+\subsubsection*{Spatial knot locations}
 For each day, we update the spatial knot locations, $\bw_1, \ldots, \bw_K$, using a Metropolis Hastings block update.
@@
 
-\subsection*{Spatial random effects}
+\subsubsection*{Spatial random effects}
 If there is temporal dependence amongst the observations, then we update $z_{tk}$ using a Metropolis Hastings update.
@@
 
-\subsection*{Variance terms}
+\subsubsection*{Variance terms}
 When there is more than one site in a partition, then we update $\sigma^2_{tk}$ using a Metropolis Hastings update.
-First, we generate from one of two candidate distributions.
-In the case of temporal dependence, we generate a candidate for ${\sigma^*_{tk}}^2$ using a random walk Gaussian candidate distribution
-\begin{align*}
-  {\sigma^{*2}_{tk}}^{(c)} \sim \text{N}({\sigma^{*2}_{tk}}^{(r - 1)}, s^2)
-\end{align*}
-where ${{\sigma^*_{tk}}^2}^{(r - 1)}$ is the value at MCMC iteration $r - 1$, and $s$ is a tuning parameter.
-In the case where there is no temporal dependence, we generate a candidate for $\sigma^2_{tk}$ using an IG$(a^*/s, b^*/s)$ candidate distribution in an independence Metropolis Hastings update where $a^* = (n_{tk} + 1) / 2 + a$, $b^* = [Y_{tk}^T \Sigma^{-1}_{tk} Y_{tk} + z_{tk}^2] / 2 + b$, $n_{tk}$ is the number of sites in partition $k$ on day $t$, and $Y_{tk}$ and $\Sigma^{-1}_{tk}$ are the observations and precision matrix for partition $k$ on day $t$.
+First, we generate a candidate for $\sigma^2_{tk}$ using an IG$(a^*/s, b^*/s)$ candidate distribution in an independence Metropolis Hastings update where $a^* = (n_{tk} + 1) / 2 + a$, $b^* = [Y_{tk}^T \Sigma^{-1}_{tk} Y_{tk} + z_{tk}^2] / 2 + b$, $n_{tk}$ is the number of sites in partition $k$ on day $t$, and $Y_{tk}$ and $\Sigma^{-1}_{tk}$ are the observations and precision matrix for partition $k$ on day $t$.
 The acceptance ratio is
@@
 
-\subsection*{Spatial covariance parameters}
+\subsubsection*{Spatial covariance parameters}
 We update the three spatial covariance parameters, $\log(\rho)$, $\log(\nu)$, $\gamma$, using a Metropolis Hastings block update step.
@@
 
-\section*{Appendix A.2: Posterior distributions}
+\subsection{Posterior distributions} \label{a:posterior}
 \input{zpost.tex}
@@
 \input{sigpost.tex}
-\input{alphapost.tex}
+\input{lambdapost.tex}
 
-\section*{Appendix A.3: Proof that $\lim_{h \rightarrow \infty} \pi(h) = 0$}
+\subsection{Proof that $\lim_{h \rightarrow \infty} \pi(h) = 0$} \label{a:proofsamepartition}
 Let $N(A)$ be the number of knots in $A$, the area between sites $\bs_1$ and $\bs_2$.
@@
 
-\section*{Appendix A.4: Half-normal distribution}
-Let $u = |z|$ where $Z \sim N(\mu, \sigma^2)$.
-Specifically, we consider the case where $\mu = 0$. Then $U$ follows a half-normal distribution which we denote as $U \sim HN(0, 1)$, and the density is given by
+% \subsection{Half-normal distribution}
+% Let $u = |z|$ where $Z \sim N(\mu, \sigma^2)$.
+% Specifically, we consider the case where $\mu = 0$. Then $U$ follows a half-normal distribution which we denote as $U \sim HN(0, 1)$, and the density is given by
+% \begin{align}
+%   f_U(u) = \frac{ \sqrt{2} }{ \sqrt{\pi \sigma^2} } \exp \left( - \frac{ u^2 }{ 2 \sigma^2 } \right) I(u > 0)
+% \end{align}
+% When $\mu = 0$, the half-normal distribution is also equivalent to a $N_{(0, \infty)}(0, \sigma^2)$ where $N_{(a, b)}(\mu, \sigma^2)$ represents a normal distribution with mean $\mu$ and standard deviation $\sigma$ that has been truncated below at $a$ and above at $b$.
+
+\subsection{Skew-$t$ distribution} \label{a:skewt}
+\subsubsection*{Univariate extended skew-$t$ distribution}
+We say that $Y$ follow a univariate extended skew-$t$ distribution with location $\xi \in \calR$, scale $\omega > 0$, skew parameter $\alpha \in \calR$, extended parameter $\tau \in \calR$, and degrees of freedom $\nu$ if has distribution function
 \begin{align}
-  f_U(u) = \frac{ \sqrt{2} }{ \sqrt{\pi \sigma^2} } \exp \left( - \frac{ u^2 }{ 2 \sigma^2 } \right) I(u > 0)
+  f_{\text{EST}}(y) = \omega^{-1} \frac{ f_T (z; \nu) }{ F_T(\tau / \sqrt{ 1 + \alpha^2}; \nu)} F_T\left[ (\alpha z + \tau) \sqrt{ \frac{ \nu + 1 }{ \nu + z^2}}; 0, 1, \nu + 1 \right]
 \end{align}
-When $\mu = 0$, the half-normal distribution is also equivalent to a $N_{(0, \infty)}(0, \sigma^2)$ where $N_{(a, b)}(\mu, \sigma^2)$ represents a normal distribution with mean $\mu$ and standard deviation $\sigma$ that has been truncated below at $a$ and above at $b$.
+where $f_T(t; \nu)$ is a univariate Student's $t$ with $\nu$ degrees of freedom, $F_T(t; \nu) = P(T < t)$, and \hbox{$z = (y - \xi) / \omega$}.
+In the case that $\tau = 0$, then $Y$ follows a univariate skew-$t$ distribution.
 
-\section*{Appendix A.5: Multivariate skew-$t$ distribution}
-\input{mvskewt.tex}
+\subsection*{Multivariate skew-$t$ distribution}
+If $\bZ \sim \text{ST}_d(0, \bar{\bOmega}, \balpha, \eta)$ is a $d$-dimensional skew-$t$ distribution, and $\bY = \xi + \bomega \bZ$, where $\bomega = \text{diag}(\omega_1, \ldots, \omega_d)$, then the density of $Y$ at $y$ is
+\begin{align}
+  f_y(\by) = det(\bomega)^{-1} f_z(\bz)
+\end{align}
+where
+\begin{align}
+  f_z(\bz) &= 2 t_d(\bz; \bar{\bOmega}, \eta) T \left[ \balpha^T \bz \sqrt{ \frac{\eta + d}{\nu + Q(\bz)} }; \eta + d\right] \\
+  \bz &= \bomega^{-1}(\by - \xi)
+\end{align}
+where $t_d(\bz; \bar{\bOmega}, \eta)$ is a $d$-dimensional Student's $t$-distribution with scale matrix $\bar{\bOmega}$ and degrees of freedom $\eta$, $Q(z) = \bz^T \bar{\Omega}^{-1}\bz$ and $T(\cdot; \eta)$ denotes the univariate Student's $t$ distribution function with $\eta$ degrees of freedom \citep{Azzalini2013}.
+
+\subsubsection*{Extremal dependence}
+For a bivariate skew-$t$ random variable $\bY = [Y(\bs), Y(\bt)]^T$, the $\chi(h)$ statistic \citep{Padoan2011} is given by
+\begin{align}
+  \chi(h) = \bar{F}_{\text{EST}}\left\{ \frac{[x_1^{1 / \eta} - \varrho(h)] \sqrt{\eta + 1} }{\sqrt{1 - \varrho(h)^2}}; 0, 1, \alpha_1, \tau_1, \eta + 1 \right\} + \bar{F}_{\text{EST}}\left\{ \frac{ [x_2^{1 / \eta} - \varrho(h)] \sqrt{\eta + 1} }{ \sqrt{1 - \varrho(h)^2} }; 0, 1, \alpha_2, \tau_2, \eta + 1 \right\},
+\end{align}
+where $\bar{F}_{\text{EST}}$ is the univariate survival extended skew-$t$ function with zero location and unit scale, \hbox{$\varrho(h) = \text{cor}(y_1, y_2)$}, $\alpha_j = \alpha_i \sqrt{1 - \varrho^2}$, $\tau_j = \sqrt{\eta + 1}(\alpha_j + \alpha_i \varrho)$, and $x_j = F_T(\bar{\alpha}_i \sqrt{\eta + 1}; 0, 1, \eta) / F_T(\bar{\alpha}_j \sqrt{\eta + 1}; 0, 1, \eta)$ with $j = 1, 2$ and $i = 2, 1$ and where $\bar{\alpha}_j = (\alpha_j + \alpha_i \varrho) / \sqrt{ 1 + \alpha_i^2 [1 - \varrho(h)^2]}$.
+
+\subsubsection*{Proof that $\lim_{h \rightarrow \infty} \chi(h) > 0$}
+Consider the bivariate distribution of $\bY = [Y(\bs), Y(\bt)]^T$, with $\varrho(h)$ given by (\ref{eq:matern}).
+So, $\lim_{h \rightarrow \infty} \varrho(h) = 0$.
+Then
+\begin{align}
+  \lim_{h \rightarrow \infty} \chi(h) = \bar{F}_{\text{EST}}\left\{ \sqrt{\eta + 1}; 0, 1, \alpha_1, \tau_1, \eta + 1 \right\} + \bar{F}_{\text{EST}}\left\{ \sqrt{\eta + 1}; 0, 1, \alpha_2, \tau_2, \eta + 1 \right\}.
+\end{align}
+Because the extended skew-$t$ distribution is not bounded above, for all $\bar{F}_{\text{EST}}(x) = 1 - F_{\text{EST}} > 0$ for all $x < \infty$.
+Therefore, for a skew-$t$ distribution, $\lim_{h \rightarrow \infty} \chi(h) > 0$.
 

### 2014-12-09 pot/LaTeX/proof_dist.tex
Revert d132b13..1d41aaf
This rolls back to commit d132b1363ba8f49558713cb8aa9a840bfeb35ba0.

@@
-Consider two points $\bs_1, \bs_2 \in \calD \subset \calR^2$.
-Let $\calD = B_{2h}(\bm)$ which is a ball of radius $h > 0$ centered at $\bm$, the midpoint between $\bs_1$ and $\bs_2$, and $h = ||\bs_1 - \bs_2||$.
-Let $w_1, \ldots, w_K \sim \text{Unif}(\calD)$ be a set of random knots uniformly distributed over $\calD$.
-Define $g(\bs_i) = \argmin_j || \bs_i - w_j ||$ to be the partition in which $\bs_i$ resides.
-Let $\pi(h) = \Pr[ g(\bs_1) = g(\bs_2)]$ be the probability that both sites are in the same partition.
-Then $\pi(h)$ is the probability that $w_j$ is the only knot in $(B_a(\bs_1) \cup B_b(\bs_2) \cap \calD)$ where $a = ||\bs_1 - w_j||$ and $b = ||\bs_2 - w_j||$.
+Consider two spatial locations, $\bs_1, \bs_2 \in \calR^2$, and any two knots, $w_1, w_2 \in \calR^2$.
+Let $\overline{S}$ be the line segment connecting $\bs_1$ and $\bs_2$.
+Let $h = ||\bs_1 - \bs_2||$ be the distance between sites $\bs_1$ and $\bs_2$.
+Let $\overline{W}$ be the perpendicular bisector of the line connecting $w_1$ and $w_2$.
+Then $\overline{W}$ defines a partition dividing $\calR^2$ into two half-planes.
+Let $\lim_{h \rightarrow \infty}$ be a limit in the sense that both $\bs_1$ and $\bs_2$ are moving at some rate away from their midpoint.
+Provided that $\overline{W}$ and $\overline{S}$ are not parallel to one another, we are guaranteed that as $h \rightarrow \infty$, $\bs_1$ and $\bs_2$ will be in different partitions.

### 2014-12-09 pot/code/analysis/ozone/US/cv5-27US.R

@@ -1,6 +1,6 @@
 options(warn=2)
 library(fields)
-library(SpatialTools)
+library(geoR)
 library(mvtnorm)
 
 rm(list=ls())
@@ -41,8 +41,8 @@ for(val in 1:2){
 	fit[[val]] <- mcmc(y=y.o, s=S.o, x=X.o, x.pred=X.p, s.pred=S.p,
                    temporalw=T, temporalz=T, temporaltau=T,
                    method=method, skew=skew, keep.knots=keep.knots,
-	                 thresh.all=threshold, thresh.quant=thresh.quant, nknots=nknots,
-                   iters=30000, burn=25000, update=500, iterplot=T,
+	               thresh.all=threshold, thresh.quant=thresh.quant, nknots=nknots,
+                   iters=30000, burn=25000, update=500, iterplot=F,
                    beta.init=beta.init, tau.init=tau.init, rho.init=1,
                    nu.init=0.5, alpha.init=0.5)
 	toc.set <- proc.time()

### 2014-12-09 pot/code/analysis/ozone/US/ozone_setup_us.R

@@ -1,5 +1,5 @@
 library(fields)
-# library(geoR)
+library(geoR)
 library(mvtnorm)
 
 rm(list=ls())
@@ -128,13 +128,13 @@ for (thresh in 1:length(threshs)){
 
 # par(mfrow=c(1, 2))
 xplot <- (1:(length(bins)-1)) - 0.5
-plot(xplot, exceed[, 1], type="o", ylim=c(0, 0.35), ylab="exceed", xaxt="n", xlab="bin distance / 1000", pch=1, lty=3, main="chi-plot ozone")
+plot(xplot, exceed[, 1], type="o", ylim=c(0, 0.5), ylab="exceed", xaxt="n", xlab="bin distance / 1000", pch=1, lty=3, main="chi-plot ozone")
 axis(1, at=0:(length(bins)-2), labels=bins[-length(bins)])
 for (line in 2:3) { 
 	lines(xplot, exceed[, line], lty=line)
 	points(xplot, exceed[, line], pch=line)
 }
-legend("topright", lty=1:3, pch=1:3, legend=probs, title="sample quantiles")
+legend("topright", lty=1:3, pch=1:3, legend=probs, title="sample quants")
 
 # Making the chi plot
 # run preprocessing first
@@ -186,20 +186,18 @@ for (thresh in 1:length(threshs)){
   exceed.h.res[, thresh] <- exceed.thresh
 }
 
-par(mar=c(5.1, 5.1, 4.1, 2.1))
 xplot <- bins.h[-length(bins)] + 0.125
-plot(xplot, exceed.h.res[, 1], type="b", pch=1, lty=1, lwd=2,  
-     xlim=c(0, 3.5), xaxt="n", xlab="bin distance (km) / 1000", 
-     ylim=c(0, 0.35), ylab=bquote(paste(chi, "(h)")), 
-     main=bquote(paste(chi, "(h) for ozone residuals")),
-     cex.lab=1.5, cex.axis=1.5, cex.main=2
+plot(xplot, exceed.h.res[, 1], type="b", pch=1, lty=1, 
+     xlim=c(0, 3.5), xaxt="n", xlab="bin distance / 1000", 
+     ylim=c(0, 0.35), ylab="exceed", 
+     main=bquote(paste(chi, "(h) for ozone residuals"))
      )
-axis(1, at=bins, cex.axis=1.5)
+axis(1, at=bins)
 for (line in 2:3) { 
[…]

### 2014-12-09 pot/code/analysis/simstudy/pot6-b.R

@@ -15,8 +15,8 @@
 #  1 - Gaussian
 #  2 - skew t-1
 #  3 - t-1 (T = 0.80)
-#  4 - skew t-5
-#  5 - t-5 (T = 0.80)
+#  4 - skew t-3
+#  5 - t-3 (T = 0.80)
 #  6 - max-stable
 #	
 #########################################################################

### 2014-12-09 pot/code/analysis/simstudy/results4.R

@@ -26,10 +26,10 @@ load("simdata.RData")
 ns <- dim(y)[1]
 nt <- dim(y)[2]
 nsets <- 5
-ngroups <- 3
+ngroups <- 10
 nsettings <- dim(y)[4]
 nmethods <- 5
-obs <- c(rep(T, 100), rep(F, 44))
+obs <- c(rep(T, 100), rep(F, 30))
 
 setting <- 4
 filename <- paste("scores", setting, ".RData", sep="")
@@ -59,11 +59,10 @@ rho <- array(NA, dim=c(length(intervals), (nsets * ngroups), nmethods, nsettings
 nu <- array(NA, dim=c(length(intervals), (nsets * ngroups), nmethods, nsettings))
 alpha <- array(NA, dim=c(length(intervals), (nsets * ngroups), nmethods, nsettings))
 # not all methods use skew or multiple partitions
-z.alpha <- array(NA, dim=c(length(intervals), (nsets * ngroups), 5, nsettings))
-avgparts <- array(NA, dim=c(length(intervals), (nsets * ngroups), 3, nsettings))
+z.alpha <- array(NA, dim=c(length(intervals), (nsets * ngroups), 4, nsettings))
+avgparts <- array(NA, dim=c(length(intervals), (nsets * ngroups), 2, nsettings))
 
 iters <- 20000; burn <- 10000
-
 for (group in 1:ngroups) {
   # for(setting in 1:nsettings){
     dataset <- paste(setting,"-c-",group,".RData", sep="")
@@ -167,42 +166,14 @@ for (group in 1:ngroups) {
       nu[, set.idx, 5, setting] <- quantile(fit$nu, probs=intervals)
       alpha[, set.idx, 5, setting] <- quantile(fit$alpha, probs=intervals)
       z.alpha[, set.idx, 4, setting] <- quantile(fit$z.alpha, probs=intervals)
-      avgparts[, set.idx, 2, setting] <- mean(fit$avgparts)
+      avgparts[, set.idx, 1, setting] <- mean(fit$avgparts)
     }
     
     cat("dataset", dataset, "\n")
     rm(fit, fit.1)
     gc()
-    
-    dataset <- paste(setting,"-d-",group,".RData", sep="")
-    load(dataset)
-    for(d in 1:nsets){ 
-      set.idx <- (group - 1) * 5 + d
[…]

### 2014-12-09 pot/LaTeX/proof_dist.tex
Revert 1d41aaf..e1a4cbd
This rolls back to commit 1d41aafe4e88de9d164d76b14831c62d29f642a2.

--- 2014-09-08 pot/LaTeX/proof_dist.tex
@@
-Consider two points $\bs_1, \bs_2 \in \calD \subset \calR^2$.
-Let $\calD = B_{2h}(\bm)$ which is a ball of radius $h > 0$ centered at $\bm$, the midpoint between $\bs_1$ and $\bs_2$, and $h = ||\bs_1 - \bs_2||$.
-Let $w_1, \ldots, w_K \sim \text{Unif}(\calD)$ be a set of random knots uniformly distributed over $\calD$.
-Define $g(\bs_i) = \argmin_j || \bs_i - w_j ||$ to be the partition in which $\bs_i$ resides.
-Let $\pi(h) = \Pr[ g(\bs_1) = g(\bs_2)]$ be the probability that both sites are in the same partition.
-Then $\pi(h)$ is the probability that $w_j$ is the only knot in $(B_a(\bs_1) \cup B_b(\bs_2) \cap \calD)$ where $a = ||\bs_1 - w_j||$ and $b = ||\bs_2 - w_j||$.
+% Consider two spatial locations, $\bs_1, \bs_2 \in \calR^2$, and any two knots, $w_1, w_2 \in \calR^2$.
+% Let $\overline{S}$ be the line segment connecting $\bs_1$ and $\bs_2$.
+% Let $h = ||\bs_1 - \bs_2||$ be the distance between sites $\bs_1$ and $\bs_2$.
+% Let $\overline{W}$ be the perpendicular bisector of the line connecting $w_1$ and $w_2$.
+% Then $\overline{W}$ defines a partition dividing $\calR^2$ into two half-planes.
+% Let $\lim_{h \rightarrow \infty}$ be a limit in the sense that both $\bs_1$ and $\bs_2$ are moving at some rate away from their midpoint.
+% Provided that $\overline{W}$ and $\overline{S}$ are not parallel to one another, we are guaranteed that as $h \rightarrow \infty$, $\bs_1$ and $\bs_2$ will be in different partitions.
+Consider two spatial locations $\bs_1, \bs_2 \in \calR^2$, and consider spatial knots that follow a Poisson process with intensity $\lambda$.
+Let $h = ||\bs_1 - \bs_2||$, and let $\pi(h)$ be the probability that two sites are in the same partition.
+To guarantee that $\bs_1$ and $\bs_2$ are indifferent partitions, we need at least two knots to be in between $\bs_1$ and $\bs_2$.
+Let $B_{\bs}(\frac{h}{2})$ be a ball that is of radius $\frac{h}{2}$ that has $\bs_1$ and $\bs_2$ on its circumference, and consider $\overline{S}$ to be the line connecting $\bs_1$ and $\bs_2$ that divides $B_{\bs}(\frac{h}{2})$ into two sizes of equal area.

### 2014-12-09 pot/code/R/mcmc.R

@@ -63,7 +63,6 @@ mcmc <- function(y, s, x, s.pred=NULL, x.pred=NULL,
   d       <- rdist(s)  # distance between sites
   diag(d) <- 0
     
-    
   # store the loc/day for observed values below thresh.
   if (thresh.quant & thresh.all > 0) {  # threshold based on sample quantiles
     thresh.all.q <- quantile(y, thresh.all, na.rm=T)
@@ -222,17 +221,33 @@ mcmc <- function(y, s, x, s.pred=NULL, x.pred=NULL,
     phi.tau <- 0
     acc.phi.tau <- att.phi.tau <- 1
     mh.phi.tau <- 1
-    mh.tau.ns <- rep(0.1, 10)
+    acc.tau.ns <- att.tau.ns <- rep(1, (ns + 1))
+    mh.tau.ns <- seq(5, 1, length=(ns + 1))
+    acc.tau.low <- acc.tau.high <- 0
   } else {
-  	mh.tau.ns <- rep(1, 10)
+  	acc.tau.ns <- att.tau.ns <- rep(1, (ns + 1))
+  	mh.tau.ns <- seq(5, 1, length=(ns + 1))
+    acc.tau.low <- acc.tau.high <- 0
   }
+  tau.trials <- nknots * nt * 3
+  
   if (temporalz) {
   	z.star <- z  # need a place to keep track of normal values for time series
     phi.z <- 0.5
-    acc.z <- att.z <- mh.z <- matrix(1, nknots, nt)
+    acc.z <- att.z <- matrix(1, nknots, nt)
+    mh.z <- matrix(15, nknots, nt)
     acc.phi.z <- att.phi.z <- mh.phi.z <- 1
   }
-    
+  
+  # candidate distributions for tau depend on the percentage of sites in the 
+  # partition. 
+  # avg.sites.part <- 1 / nknots
+  # mh.tau.parts <- c(0, (avg.sites.part / 4), (avg.sites.part / 2), avg.sites.part, 
+                    # (2 * avg.sites.part), (3 * avg.sites.part), (4 * avg.sites.part))
+  acc.tau <- att.tau   <- matrix(1, nrow=nknots, ncol=nt) 
+  mh.tau <- matrix(0.05, nknots, nt)
+  nparts.tau <- matrix(1, nrow=nknots, ncol=nt)
+
   # MH tuning params
[…]

### 2014-12-09 pot/code/analysis/ozone/US/cv5-8US.R

@@ -1,6 +1,6 @@
 options(warn=2)
 library(fields)
-library(geoR)
+library(SpatialTools)
 library(mvtnorm)
 
 rm(list=ls())
@@ -10,7 +10,7 @@ source('../../../R/auxfunctions.R')
 
 setting <- 8
 method <- "t"
-nknots <- 5
+nknots <- 2
 keep.knots <- F
 threshold <- 75
 tau.init <- 0.05
@@ -40,7 +40,7 @@ for(val in 1:2){
 	fit[[val]] <- mcmc(y=y.o, s=S.o, x=X.o, x.pred=X.p, s.pred=S.p,
 	                   method=method, skew=skew, keep.knots=keep.knots,
 	                   thresh.all=threshold, thresh.quant=thresh.quant, nknots=nknots, 
-                       iters=30000, burn=25000, update=500, iterplot=F,
+                       iters=30000, burn=25000, update=100, iterplot=T,
                        beta.init=beta.init, tau.init=tau.init, rho.init=1,
                        nu.init=0.5, alpha.init=0.5)
 	toc.set <- proc.time()

### 2014-12-09 pot/code/analysis/simstudy/pot1-c.R

@@ -36,7 +36,7 @@ analysis <- "c"
 iters <- 20000; burn <- 10000; update <- 1000; thin <- 1
 nsets <- 5
 
-for (g in 3:10) {
+for (g in 8:10) {
   fit.1 <- fit.2 <- fit.3 <- vector(mode="list", length=nsets)
   y.validate <- array(NA, dim=c(ntest, nt, nsets))
   outputfile <- paste(setting, "-", analysis, "-", g, ".RData", sep="")

### 2014-12-09 pot/code/analysis/simstudy/pot2-d.R

@@ -31,8 +31,8 @@ load(file='./simdata.RData')
 source('../../R/mcmc.R')
 source('../../R/auxfunctions.R')
 
-setting <- 4
-analysis <- "b"
+setting <- 2
+analysis <- "d"
 iters <- 20000; burn <- 10000; update <- 1000; thin <- 1
 nsets <- 5
 
@@ -55,16 +55,29 @@ for (g in 1:10) {
     y.validate[, , d] <- y.d[!obs, ]
     x.p <- x[!obs, , ]
     s.p <- s[!obs, ]
-
-    cat("  start: skew t-5 - Set", dataset, "\n")
+    
+    cat("  start: t-5 (T=0.80) - Set", dataset, "\n")
     tic <- proc.time()
-    fit.1[[d]] <- mcmc(y=y.o, s=s.o, x=x.o, s.pred=s.p, x.pred=x.p,
-                       method="t", skew=T, thresh.all=0, thresh.quant=T,
-                       nknots=5, iterplot=F, iters=iters, burn=burn,
-                       update=update, thin=thin)
+    fit.1[[d]] <- tryCatch(
+                       mcmc(y=y.o, s=s.o, x=x.o, s.pred=s.p, x.pred=x.p,
+                       method="t", skew=T, thresh.all=0.80, thresh.quant=T,
+                       nknots=5, iterplot=F, iters=iters, burn=burn, 
+                       update=update, thin=thin,
+                       nu.init=0.5, cov.model="exponential", rho.prior="cont"),
+                       error = function(e) {
+                         tryCatch(mcmc(y=y.o, s=s.o, x=x.o, s.pred=s.p, x.pred=x.p,
+                         method="t", skew=T, thresh.all=0.80, thresh.quant=T, 
+                         nknots=5, iterplot=F, iters=iters, burn=burn, 
+                         update=update, thin=thin,
+                         nu.init=0.5, cov.model="exponential", rho.prior="disc"),
+                         error = function(e) {
+                           cat("dataset", d, "not working \n")
+                           "no results"
+                         })
+                       })
     toc <- proc.time()
-    cat("  skew t-5 took:", (toc - tic)[3], "\n")
-    cat("  end: skew t-5 \n")
+    cat("  t-5 (T=0.80) took:", (toc - tic)[3], "\n")
+    cat("  end: t-5 (T=0.80) \n")
     cat("------------------\n")
 
     save(fit.1, file=outputfile)

### 2014-12-09 pot/code/analysis/simstudy/pot3-b.R

@@ -15,8 +15,8 @@
 #  1 - Gaussian
 #  2 - skew t-1
 #  3 - t-1 (T = 0.80)
-#  4 - skew t-3
-#  5 - t-3 (T = 0.80)
+#  4 - skew t-5
+#  5 - t-5 (T = 0.80)
 #  6 - max-stable
 #	
 #########################################################################
@@ -36,7 +36,7 @@ analysis <- "b"
 iters <- 20000; burn <- 10000; update <- 1000; thin <- 1
 nsets <- 5
 
-for (g in 2:10) {
+for (g in 1:10) {
   fit.1 <- vector(mode="list", length=nsets)
   y.validate <- array(NA, dim=c(ntest, nt, nsets))
   outputfile <- paste(setting, "-", analysis, "-", g, ".RData", sep="")

### 2014-12-09 pot/code/analysis/simstudy/pot3-d.R

@@ -31,8 +31,8 @@ load(file='./simdata.RData')
 source('../../R/mcmc.R')
 source('../../R/auxfunctions.R')
 
-setting <- 4
-analysis <- "b"
+setting <- 3
+analysis <- "d"
 iters <- 20000; burn <- 10000; update <- 1000; thin <- 1
 nsets <- 5
 
@@ -55,16 +55,29 @@ for (g in 1:10) {
     y.validate[, , d] <- y.d[!obs, ]
     x.p <- x[!obs, , ]
     s.p <- s[!obs, ]
-
-    cat("  start: skew t-5 - Set", dataset, "\n")
+    
+    cat("  start: t-5 (T=0.80) - Set", dataset, "\n")
     tic <- proc.time()
-    fit.1[[d]] <- mcmc(y=y.o, s=s.o, x=x.o, s.pred=s.p, x.pred=x.p,
-                       method="t", skew=T, thresh.all=0, thresh.quant=T,
-                       nknots=5, iterplot=F, iters=iters, burn=burn,
-                       update=update, thin=thin)
+    fit.1[[d]] <- tryCatch(
+                       mcmc(y=y.o, s=s.o, x=x.o, s.pred=s.p, x.pred=x.p,
+                       method="t", skew=T, thresh.all=0.80, thresh.quant=T,
+                       nknots=5, iterplot=F, iters=iters, burn=burn, 
+                       update=update, thin=thin,
+                       nu.init=0.5, cov.model="exponential", rho.prior="cont"),
+                       error = function(e) {
+                         tryCatch(mcmc(y=y.o, s=s.o, x=x.o, s.pred=s.p, x.pred=x.p,
+                         method="t", skew=T, thresh.all=0.80, thresh.quant=T, 
+                         nknots=5, iterplot=F, iters=iters, burn=burn, 
+                         update=update, thin=thin,
+                         nu.init=0.5, cov.model="exponential", rho.prior="disc"),
+                         error = function(e) {
+                           cat("dataset", d, "not working \n")
+                           "no results"
+                         })
+                       })
     toc <- proc.time()
-    cat("  skew t-5 took:", (toc - tic)[3], "\n")
-    cat("  end: skew t-5 \n")
+    cat("  t-5 (T=0.80) took:", (toc - tic)[3], "\n")
+    cat("  end: t-5 (T=0.80) \n")
     cat("------------------\n")
 
     save(fit.1, file=outputfile)

### 2014-12-09 pot/code/analysis/simstudy/results1.R

@@ -26,10 +26,10 @@ load("simdata.RData")
 ns <- dim(y)[1]
 nt <- dim(y)[2]
 nsets <- 5
-ngroups <- 10
+ngroups <- 3
 nsettings <- dim(y)[4]
 nmethods <- 5
-obs <- c(rep(T, 100), rep(F, 30))
+obs <- c(rep(T, 100), rep(F, 44))
 
 setting <- 1
 filename <- paste("scores", setting, ".RData", sep="")
@@ -59,8 +59,8 @@ rho <- array(NA, dim=c(length(intervals), (nsets * ngroups), nmethods, nsettings
 nu <- array(NA, dim=c(length(intervals), (nsets * ngroups), nmethods, nsettings))
 alpha <- array(NA, dim=c(length(intervals), (nsets * ngroups), nmethods, nsettings))
 # not all methods use skew or multiple partitions
-z.alpha <- array(NA, dim=c(length(intervals), (nsets * ngroups), 4, nsettings))
-avgparts <- array(NA, dim=c(length(intervals), (nsets * ngroups), 2, nsettings))
+z.alpha <- array(NA, dim=c(length(intervals), (nsets * ngroups), 5, nsettings))
+avgparts <- array(NA, dim=c(length(intervals), (nsets * ngroups), 3, nsettings))
 
 iters <- 20000; burn <- 10000
 
@@ -167,13 +167,42 @@ for (group in 1:ngroups) {
       nu[, set.idx, 5, setting] <- quantile(fit$nu, probs=intervals)
       alpha[, set.idx, 5, setting] <- quantile(fit$alpha, probs=intervals)
       z.alpha[, set.idx, 4, setting] <- quantile(fit$z.alpha, probs=intervals)
-      avgparts[, set.idx, 1, setting] <- mean(fit$avgparts)
+      avgparts[, set.idx, 2, setting] <- mean(fit$avgparts)
     }
     
     cat("dataset", dataset, "\n")
     rm(fit, fit.1)
     gc()
+    
+    dataset <- paste(setting,"-d-",group,".RData", sep="")
+    load(dataset)
+    for(d in 1:nsets){ 
+      set.idx <- (group - 1) * 5 + d
+      thresholds <- quantile(y[, , set.idx, setting], probs=probs, na.rm=T)
+      validate <- y[!obs, , set.idx, setting]
+      
+      fit <- fit.1[[d]]  # skew t-5 (T = 0.90)
[…]

### 2014-12-09 pot/code/analysis/simstudy/results2.R

@@ -26,10 +26,10 @@ load("simdata.RData")
 ns <- dim(y)[1]
 nt <- dim(y)[2]
 nsets <- 5
-ngroups <- 10
+ngroups <- 3
 nsettings <- dim(y)[4]
 nmethods <- 5
-obs <- c(rep(T, 100), rep(F, 30))
+obs <- c(rep(T, 100), rep(F, 44))
 
 setting <- 2
 filename <- paste("scores", setting, ".RData", sep="")
@@ -59,10 +59,11 @@ rho <- array(NA, dim=c(length(intervals), (nsets * ngroups), nmethods, nsettings
 nu <- array(NA, dim=c(length(intervals), (nsets * ngroups), nmethods, nsettings))
 alpha <- array(NA, dim=c(length(intervals), (nsets * ngroups), nmethods, nsettings))
 # not all methods use skew or multiple partitions
-z.alpha <- array(NA, dim=c(length(intervals), (nsets * ngroups), 4, nsettings))
-avgparts <- array(NA, dim=c(length(intervals), (nsets * ngroups), 2, nsettings))
+z.alpha <- array(NA, dim=c(length(intervals), (nsets * ngroups), 5, nsettings))
+avgparts <- array(NA, dim=c(length(intervals), (nsets * ngroups), 3, nsettings))
 
 iters <- 20000; burn <- 10000
+
 for (group in 1:ngroups) {
   # for(setting in 1:nsettings){
     dataset <- paste(setting,"-c-",group,".RData", sep="")
@@ -166,13 +167,42 @@ for (group in 1:ngroups) {
       nu[, set.idx, 5, setting] <- quantile(fit$nu, probs=intervals)
       alpha[, set.idx, 5, setting] <- quantile(fit$alpha, probs=intervals)
       z.alpha[, set.idx, 4, setting] <- quantile(fit$z.alpha, probs=intervals)
-      avgparts[, set.idx, 1, setting] <- mean(fit$avgparts)
+      avgparts[, set.idx, 2, setting] <- mean(fit$avgparts)
     }
     
     cat("dataset", dataset, "\n")
     rm(fit, fit.1)
     gc()
+    
+    dataset <- paste(setting,"-d-",group,".RData", sep="")
+    load(dataset)
+    for(d in 1:nsets){ 
+      set.idx <- (group - 1) * 5 + d
[…]

### 2014-12-09 pot/code/analysis/simstudy/results4.R

@@ -26,10 +26,10 @@ load("simdata.RData")
 ns <- dim(y)[1]
 nt <- dim(y)[2]
 nsets <- 5
-ngroups <- 10
+ngroups <- 3
 nsettings <- dim(y)[4]
 nmethods <- 5
-obs <- c(rep(T, 100), rep(F, 30))
+obs <- c(rep(T, 100), rep(F, 44))
 
 setting <- 4
 filename <- paste("scores", setting, ".RData", sep="")
@@ -59,10 +59,11 @@ rho <- array(NA, dim=c(length(intervals), (nsets * ngroups), nmethods, nsettings
 nu <- array(NA, dim=c(length(intervals), (nsets * ngroups), nmethods, nsettings))
 alpha <- array(NA, dim=c(length(intervals), (nsets * ngroups), nmethods, nsettings))
 # not all methods use skew or multiple partitions
-z.alpha <- array(NA, dim=c(length(intervals), (nsets * ngroups), 4, nsettings))
-avgparts <- array(NA, dim=c(length(intervals), (nsets * ngroups), 2, nsettings))
+z.alpha <- array(NA, dim=c(length(intervals), (nsets * ngroups), 5, nsettings))
+avgparts <- array(NA, dim=c(length(intervals), (nsets * ngroups), 3, nsettings))
 
 iters <- 20000; burn <- 10000
+
 for (group in 1:ngroups) {
   # for(setting in 1:nsettings){
     dataset <- paste(setting,"-c-",group,".RData", sep="")
@@ -166,14 +167,42 @@ for (group in 1:ngroups) {
       nu[, set.idx, 5, setting] <- quantile(fit$nu, probs=intervals)
       alpha[, set.idx, 5, setting] <- quantile(fit$alpha, probs=intervals)
       z.alpha[, set.idx, 4, setting] <- quantile(fit$z.alpha, probs=intervals)
-      avgparts[, set.idx, 1, setting] <- mean(fit$avgparts)
+      avgparts[, set.idx, 2, setting] <- mean(fit$avgparts)
     }
     
     cat("dataset", dataset, "\n")
     rm(fit, fit.1)
     gc()
+    
+    dataset <- paste(setting,"-d-",group,".RData", sep="")
+    load(dataset)
+    for(d in 1:nsets){ 
+      set.idx <- (group - 1) * 5 + d
[…]

### 2014-12-11 pot/LaTeX/skewnormal.tex
adjustment to thresholded observation generation

@@
 Many types of data demonstrate some level of skewness and therefore should be modeled with distributions that allow for asymmetry.
-The skew-elliptical family of distributions provides models that are mathematically tractable while introducing a slant parameter to account for non-symmetric data \citep{Genton2004}.
+The skew-elliptical family of distributions provides models that are mathematically tractable while introducing a slant parameter to account for asymmetric data \citep{Genton2004}.
 A brief review of the additive process by which a skew-$t$ process is created is given here.
@@
 where $\bX(\bs)$ is a set of spatial covariates at site $\bs$, $\bbeta$ is a set of regression parameters, $\lambda \in \calR$ is a parameter controlling skew, $z \sim N(0, 1)$, $\sigma^2 \sim \text{IG}(a, b)$ is an inverse gamma random variable, and $v(\bs)$ is a spatial Gaussian process with mean zero and variance one.
-It should be noted that the spatial-$t$ process is a special case of the skew-$t$ process when $\lambda = 0$.
 A common spatial correlation function is the \Matern with
@@
 \end{align}
-where $\rho$ is the spatial range, $\nu$ is the smoothness, $\gamma$ is the proportion of variance accounted for by the spatial variation, $K_\nu$ is a modified Bessel function of the second time, and $h = || \bs - \bt ||$.
+where $\rho$ is the spatial range, $\nu$ is the smoothness, $\gamma$ is the proportion of variance accounted for by the spatial variation, $K_\nu$ is a modified Bessel function of the second kind, and $h = || \bs - \bt ||$.
 
@@
 \end{align}
-that is, $\bY$ follows an $n$-dimensional skew-$t$ process with location $\bX \bbeta$, correlation matrix $\bOmega$, slant parameters $\balpha$ and degrees of freedom $2a$ where $\bX = [\bX(\bs_1)^T, \ldots, \bX(\bs_n)^T]$, $\bOmega = \bomega \bar{\bOmega} \bomega$, $\bomega = \text{diag}\left(\frac{1}{ \sqrt{ab}}, \ldots, \frac{1}{\sqrt{ab}}\right)$, $\bar{\bOmega} = (\bSigma + \lambda^2 \bOne \bOne^T$), $\bSigma$ is a positive definite correlation matrix, $\balpha = \lambda (1 + \lambda^2 \bOne^T \bSigma^{-1} \bOne)^{-1/2} \bOne^T \bSigma^{-1}$ is a vector of slant parameters.
+that is, $\bY$ follows an $n$-dimensional skew-$t$ process with location $\bX \bbeta$, correlation matrix $\bOmega$, slant parameters $\balpha$ and degrees of freedom $2a$, where $\bX = [\bX(\bs_1)^T, \ldots, \bX(\bs_n)^T]$, $\bOmega = \bomega \bar{\bOmega} \bomega$, $\bomega = \text{diag}\left(\frac{1}{ \sqrt{ab}}, \ldots, \frac{1}{\sqrt{ab}}\right)$, $\bar{\bOmega} = (\bSigma + \lambda^2 \bOne \bOne^T$), $\bSigma$ is a positive definite correlation matrix, $\balpha = \lambda (1 + \lambda^2 \bOne^T \bSigma^{-1} \bOne)^{-1/2} \bOne^T \bSigma^{-1}$ is a vector of slant parameters.
 This process is desirable because of its flexible tail that is controlled by the skewness parameter $\lambda$ and degrees of freedom $2a$.
@@
 As discussed in Section \ref{s:spatialskew}, the source of long-range dependence is the shared $z$ and $\sigma$.
-Therefore, to alleviate this dependence, we allow $z$ and $\sigma$to vary by site.
+Therefore, to alleviate this dependence, we allow $z$ and $\sigma$ to vary by site.
 Then, the model becomes
@@
 The knots define a random daily partition of $\calD$ by subregions $P_{1}, \ldots, P_{K}$ defined as
-\begin{align*}
+\begin{align} \label{eq:subregions}
   P_{k} = \{ \bs : k = \argmin_\ell || \bs - \bw_{\ell} || \}.
-\end{align*}
+\end{align}
 All $z(\bs)$ and $\sigma(\bs)$ for sites in subregion $k$ are assigned common values
@@
 Marginally, over the knots $\bw_1, \ldots, \bw_K$, $\chi(h) = \pi(h) \chi_{\text{skew}-t}(h)$, where $\pi(h) = \Pr(k_s = k_t)$ is the probability that two sites separated by distance $h$ are in the same partition.
-So, to show that $\lim_{h \rightarrow \infty} \chi(h) = 0$, we need only know that $\lim_{h \rightarrow \infty} \pi(h) = 0$
+So, to show that $\lim_{h \rightarrow \infty} \chi(h) = 0$, we need only know that $\lim_{h \rightarrow \infty} \pi(h) = 0$.
 A proof of this is given in Appendix \ref{a:proofsamepartition}.
 
-Figure \ref{fig:simchi} shows $\chi(h)$ for different sample quantiles from simulated datasets where $\bs \in [0, 1] \times [0, 1]$.
-These plots show how the partitioning allows for short range extremal dependence while reducing extremal dependence as the distance between sites increases.
+Figure \ref{fig:simchi} shows $\chi(h)$ for varying data settings when $\bs \in [0, 1] \times [0, 1]$.
+The plot clearly shows that without partitioning, in a skew-$t$ process, all sites remain asymptotically dependent.
+However, upon incorporating partitioning into the model, the asymptotic dependence is only shared by sites that are close to one another.
 
 \begin{figure}
-  \includegraphics[width=\linewidth]{plots/sim-chi.pdf}
-  \caption{$\chi(h)$ for simulated datasets. The dashed line indicates the $\chi(h)$ we would expect under asymptotic independence.}
+  %\includegraphics[width=\linewidth]{plots/sim-chi.pdf}
+  \caption{$\chi(h)$ for different settings. I need to replace the plot here.}
   \label{fig:simchi}
@@
 Let
-\begin{align*}
+\begin{align} \label{eq:spatiotemp}
   Y_t(\bs) = \bX_t(\bs)^T \bbeta + \lambda \sigma_t(\bs) |z_t(\bs)| + \sigma_t(\bs) v_t(\bs),
-\end{align*}
+\end{align}
 where $t \in \{1, \ldots, T\}$ denotes the day of each observation.
@@
 \begin{align}
-  w^*_{tki} = \Phi^{-1}\left[ \frac{ (w_{tki} - \min(s_i))}{ \text{range}(s_i) } \right], \quad i = 1, 2
+  w^*_{tki} = \Phi^{-1}\left[ \frac{ (w_{tki} - \min(\bs_i))}{ \text{range}(\bs_i) } \right], \quad i = 1, 2
 \end{align}
-where $\Phi$ is a univariate standard normal density function.
-Then $\bw^*_k \in \calR^2$.
+where $\Phi$ is a univariate standard normal density function, and $\bs_i = [s_{1i}, \ldots, s_{ni}]$.
+Then $\bw^*_{tk} \in \calR^2$.
 We use a copula on the $\sigma^2_t(\bs)$ to ensure that the marginal distributions of $\sigma^2_t(\bs)$ are inverse gamma.
@@
 where $|\phi_w| < 1$, $0 < \phi_z < 1$, and $|\phi_\sigma| < 1$.
-We restrict $\phi_z$ to be positive because in the skew-$t$ distribution, we use $|z_t(\bs)|$.
+We restrict $\phi_z$ to be positive because in the model, we assume $Y(\bs)$ is a function of $|z_t(\bs)|$.
 These are stationary time series models with marginal distributions $\bw^*_{k} \sim N_2(0, 1)$, $z_{k} \sim N(0, \sigma^2_{k})$, $\sigma^{2*}_{k} \sim N(0, 1)$.
-After transformation back to the original space, $\bw_{tk} \sim \text{Unif}(\calD)$, $\sigma^2 \sim \text{IG}(a, b)$.
-For each day, the model is identical to the spatial-only model in Section \ref{s:part} by construction.
+After transformation back to the original space, $\bw_{tk} \sim \text{Unif}(\calD)$, $\sigma^2_{tk} \sim \text{IG}(a, b)$.
+For each day, the model is identical to the spatial-only model in (\ref{eq:partition}) by construction.
 
@@
 We model this with a Bayesian hierarchical model as follows.
-Let $\bw_{t1}, \ldots, \bw_{tK}$ be a set of daily spatial knots in a spatial domain of interest, $\calD$, and $P_{tk}$ as defined in \ref{s:part}.
+Let $\bw_{t1}, \ldots, \bw_{tK}$ be a set of daily spatial knots in a spatial domain of interest, $\calD$, and $P_{tk}$ as defined in (\ref{eq:subregions}).
 
@@
 \end{align}
-where $\Theta = \{\rho, \nu, \gamma\}$; \mbox{$k = \argmin_\ell ||\bs - \bw_\ell ||$}; and $\Sigma$ is a \Matern covariance matrix as described in Section \ref{s:skewt}.
+where $\Theta = \{\rho, \nu, \gamma\}$, and $\Sigma$ is a \Matern covariance matrix as described in Section \ref{s:skewt}.
 

### 2015-01-03 prelim/oral_prelim.tex
prelim files

--- 2014-08-23 pot/LaTeX/jsm2014/jsm2014.tex
@@
 
-\usepackage{cmbright}
+% \usepackage{cmbright}
 \renewcommand*\familydefault{\sfdefault}
@@
 
-\title[Spatiotemporal Modeling of Extreme Events] % (optional, use only with long paper titles)
+\title[Spatial methods for extreme value analysis] % (optional, use only with long paper titles)
 {
-  Spatiotemporal Modeling of Extreme Events
+  Spatial methods for extreme value analysis
 }
-\author[S. Morris, B. Reich, and D. Cooley]{Samuel Morris (NC State), Brian Reich (NC State), and Dan Cooley (Colorado State)}
-\institute[NCSU, CSU]{}
-\date{JSM 2014, Boston}
+\author[Samuel A. Morris]{Samuel A. Morris}
+\institute[NCSU]{}
+\date{January 9, 2015}
 
@@
 
-% \begin{frame}{Introduction to extremes}
-%   \begin{itemize} \setlength{\itemsep}{0.5em}
-%     \item Max-stable processes (Cooley et al., 2012):
-%     \begin{itemize}
-%       \item Consider a spatial process $x_t(\bs)$, $t = 1, \ldots, T$.
-%       \item Let $M_T(\bs) = \left\{ \bigvee_{t=1}^T x_t(\bs_1), \ldots, \bigvee_{t=1}^T x_t(\bs_n) \right\}$
-%       \item If there exists normalizing sequences $a_T(\bs)$ and $b_T(\bs)$ such
-%       that for all sites, $\bs_i, i = 1, \ldots, d$,
-%       \begin{align*}
-%         a_T^{-1}(\bs) \left\{ M_T(\bs) - b_T(\bs) \right\} \converged Y(\bs)
-%       \end{align*}
-%       which has a non-degenerate distribution, then $Y(\bs)$ is a max-stable process.
-%     \end{itemize}
-%   \end{itemize}
-% \end{frame}
+\begin{frame}{Introduction to extremes}
+  \begin{itemize} \setlength{\itemsep}{0.5em}
+    \item Max-stable processes (Cooley et al., 2012):
+    \begin{itemize}
+      \item Consider a spatial process $x_t(\bs)$, $t = 1, \ldots, T$.
+      \item Let $M_T(\bs) = \left\{ \bigvee_{t=1}^T x_t(\bs_1), \ldots, \bigvee_{t=1}^T x_t(\bs_n) \right\}$
+      \item If there exists normalizing sequences $a_T(\bs)$ and $b_T(\bs)$ such
+      that for all sites, $\bs_i, i = 1, \ldots, d$,
+      \begin{align*}
+        a_T^{-1}(\bs) \left\{ M_T(\bs) - b_T(\bs) \right\} \converged Y(\bs)
+      \end{align*}
+      which has a non-degenerate distribution, then $Y(\bs)$ is a max-stable process.
+    \end{itemize}
+  \end{itemize}
+\end{frame}
 
@@
 
-% \begin{frame}{Multivariate representations}
-%   \begin{itemize}
-%     \item Multivariate distributions:
-%     \begin{itemize}
-%       \item Assume common standardized max-stable marginal, like unit-Fr\'{e}chet
-%       \begin{align*}
-%         \Pr(Z < z) = exp(-z^{-1})
-%       \end{align*}
-%       \item The multivariate representation for the GEV is
-%       \begin{align*}
-%         \Pr(\bZ \le \bz)  &= G^*(\bz) = \exp(-V(\bz))\\
-%                 V(\bs)    &= d \int_{\Delta_d} \bigvee_{i = 1}^d \frac{w_i}{z_i} H(\ddd w)
-%       \end{align*}
-%       where
-%       \begin{itemize}
-%         \item $\Delta_d = \{ \bw \in \calR^d_+ \mid w_1 + \cdots + w_d = 1\}$
-%         \item $H$ is a probability measure on $\Delta_d$
-%         \item $\int_{\Delta_d}w_i H(\ddd w) = 1 / d$ for $i = 1, \ldots, d$.
-%       \end{itemize}
-%     \end{itemize}
-%   \end{itemize}
-% \end{frame}
+\begin{frame}{Multivariate representations}
+  \begin{itemize}
+    \item Multivariate distributions:
+    \begin{itemize}
+      \item Assume common standardized max-stable marginal, like unit-Fr\'{e}chet
+      \begin{align*}
+        \Pr(Z < z) = exp(-z^{-1})
+      \end{align*}
+      \item The multivariate representation for the GEV is
+      \begin{align*}
+        \Pr(\bZ \le \bz)  &= G^*(\bz) = \exp(-V(\bz))\\
+                V(\bs)    &= d \int_{\Delta_d} \bigvee_{i = 1}^d \frac{w_i}{z_i} H(\ddd w)
+      \end{align*}
+      where
+      \begin{itemize}
+        \item $\Delta_d = \{ \bw \in \calR^d_+ \mid w_1 + \cdots + w_d = 1\}$
+        \item $H$ is a probability measure on $\Delta_d$
+        \item $\int_{\Delta_d}w_i H(\ddd w) = 1 / d$ for $i = 1, \ldots, d$.
+      \end{itemize}
+    \end{itemize}
+  \end{itemize}
+\end{frame}
 
@@
     \end{itemize}
+    \item Bayesian hierarchical model
     \item Pairwise likelihood approach (Huser and Davison, 2014)

### 2015-01-03 prelim/plots/plots.R

@@ -1,0 +1,578 @@
+# libraries
+library(fields)
+library(geoR)
+
+# necessary functions
+source('../../../code/R/mcmc.R')
+source('../../../code/R/auxfunctions.R')
+
+# data settings
+beta.t <- c(10, 2, -3)
+nu.t <- 0.5
+alpha.t <- 0.9
+mixprob.t <- c(0, 1, 1, 1, 1, 0.5)  # 0: Gaussian, 1: t
+nknots.t <- c(1, 1, 5, 1, 5, 1)
+gau.rho.t <- c(0.10, 0.10, 0.10, 0.10, 0.10, 0.10)
+t.rho.t <- c(0.10, 0.10, 0.10, 0.10, 0.10, 0.40)
+z.alpha.t <- c(0, 0, 0, 3, 3, 0)
+tau.alpha.t <- 2
+tau.beta.t  <- 8
+
+# covariate data
+s         <- cbind(runif(500), runif(500))
+ns        <- nrow(s)
+nt        <- 50
+nsets     <- 1
+nsettings <- length(mixprob.t)
+ntest     <- floor(ns / 2)
+
+x <- array(1, c(ns, nt, 3))
+for (t in 1:nt) {
+    x[, t, 2] <- s[, 1]
+    x[, t, 3] <- s[, 2]
+}
+
+# Storage for datasets
+y <- array(NA, dim=c(ns, nt, nsets, nsettings))
+tau.t <- z.t <- knots.t <- vector("list", length=nsettings)
+
+for (setting in 1:nsettings) {
+  nknots <- nknots.t[setting]
+  tau.t.setting <- array(NA, dim=c(nknots, nt, nsets))
+  z.t.setting <- array(NA, dim=c(nknots, nt, nsets))
+  knots.t.setting <- array(NA, dim=c(nknots, 2, nt, nsets))
+  for (set in 1:nsets) {
+    set.seed(setting*100 + set)
+    data <- rpotspat(nt=nt, x=x, s=s, beta=beta.t, alpha=alpha.t, nu=nu.t,
+                     gau.rho=gau.rho.t[setting], t.rho=t.rho.t[setting],
+                     mixprob=mixprob.t[setting], z.alpha=z.alpha.t[setting],
+                     tau.alpha=tau.alpha.t, tau.beta=tau.beta.t,
+                     nknots=nknots.t[setting])
+
+    y[, , set, setting]        <- data$y
+    tau.t.setting[, , set]     <- data$tau
[…]

### 2015-01-05 pot/LaTeX/skewnormal.tex
manuscript and plotting updates

@@
 \begin{align}
-  \chi(h) = \lim_{c \rightarrow \infty} \Pr(Y(\bs) > c | Y(\bt) > c).
+  \chi(h) = \lim_{c \rightarrow \infty} \Pr[Y(\bs) > c | Y(\bt) > c].
 \end{align}
@@
 A proof of this is given in Appendix \ref{a:proofsamepartition}.
-
-Figure \ref{fig:simchi} shows $\chi(h)$ for varying data settings when $\bs \in [0, 1] \times [0, 1]$.
-The plot clearly shows that without partitioning, in a skew-$t$ process, all sites remain asymptotically dependent.
-However, upon incorporating partitioning into the model, the asymptotic dependence is only shared by sites that are close to one another.
+Figure \ref{fig:chi} shows how partitioning reduces the extremal dependence.
 
 \begin{figure}
-  %\includegraphics[width=\linewidth]{plots/sim-chi.pdf}
-  \caption{$\chi(h)$ for different settings. I need to replace the plot here.}
-  \label{fig:simchi}
+  \centering
+  \includegraphics[width=0.5\linewidth]{plots/chi-h.pdf}
+  \caption{$\chi(h)$ for $K = 1, 3, 5$, and $10$ partitions as a function of distance.}
+  \label{fig:chi}
 \end{figure}
@@
 \end{enumerate}
-In the first five designs, the $v_t(\bs)$ terms are generated using a \Matern covariance with smoothness parameter $\nu = 0.5$ and spatial range $\rho = 0.1$.
+In the first five designs, the $v_t(\bs)$ terms are generated using a \Matern covariance with smoothness parameter $\nu = 0.5$ and spatial range $\rho = 1$.
 For the covariance matrices in designs 1 -- 5, the proportion of the variance accounted for by the spatial variation is $\gamma = 0.9$ while the proportion of the variance accounted for by the nugget effect is $0.1$.
@@
 For designs four and five, $\lambda = 3$ was used, and the $z_t$ are generated as described in (\ref{eq:sitez}).
-In the sixth design, we generate from a spatial max-stable distribution \citep{Reich2012} with parameters $\mu = 1, \sigma=1, \xi=0.2$ and 144 spatial knots on a regular lattice in the square $[1, 9] \times [1, 9]$.
-\hl{I know I need more here about the max-stable method}
+In the sixth design, we generate from a spatial max-stable distribution \citep{Reich2012}.
+In this design, data have marginal distributions that follow a generalized extreme value distribution with parameters $\mu = 1, \sigma=1, \xi=0.2$.
+In this model, a random effect is used to induce spatial dependence using 144 spatial knots on a regular lattice in the square $[1, 9] \times [1, 9]$.
+For this setting, we set $\gamma = 0.5$.
 In all six designs, the mean $\bX \bbeta = 10$ is assumed to be constant across space.
@@
   \item Skew-$t$ marginal, $K=1$ knots, $T=-\infty$
-  \item $T$ marginal, $K=1$ knots, $T=q(0.80)$
+  \item Skew-$t$ marginal, $K=1$ knots, $T=q(0.80)$
   \item Skew-$t$ marginal, $K=5$ knots, $T=-\infty$
-  \item $T$ marginal, $K=5$ knots, $T=q(0.80)$
+  \item Skew-$t$ marginal, $K=5$ knots, $T=q(0.80)$
 \end{enumerate}
@@
 \subsection{Cross validation}\label{s:modelselect}
-Models were compared using cross validation with 100 sites used as training sites and 30 sites witheld for testing.
+Models were compared using cross validation with 100 sites used as training sites and 44 sites witheld for testing.
 The model was fit using the training set, and predictions were generated at the testing site locations.
@@
 \end{align*}
-For data that are relatively symmetric, the standard Gaussian method outperforms the other methods.
-However, when the data come from a skewed distribution (settings 3 -- 6), the partitioned and thresholded methods outperform the Gaussian method.
-Additionally, for both of the skew-$t$ settings ($K = 1$, $K = 5$), the skew-$t$ method with the appropriate number of partitions performs the best.
-Finally, the simulation results suggest that for data from a skew-$t$ distribution, thresholded methods provide some improvement over non-thresholded models for the extreme thresholds.
-With the exception of the Brier score for $c = q(0.995)$ in setting 6, the improvement over the Gaussian method is statistically significant using a Wilcoxon signed rank test ($\alpha = 0.05$).
+We analyzed the results for the simulation study using two-sided Wilcoxon signed rank tests.
+For the first five data settings, our model with no thresholding and $K = 1$ or $5$ partitions performs significantly better (p-value $<0.001$) than the Gaussian method.
+Additionally, for the $T$ and skew-$t$ settings with $K = 5$, the skew-$t$ method with $K=5$ partitions performs significantly better (p-value $<0.001$) than the skew-$t$ method with $K=1$ partition except for the Brier score at $T = q(0.995)$ for the $T$ setting with $K = 5$ (p-value $=0.036$).
+For the $T$ and skew-$t$ settings with $K = 1$, the Brier scores for the model with $K = 1$ partition are significantly better (p-value $<0.001$) for thesholds less than $q(0.95)$.
+However, as the threshold for the Brier score increases, the difference becomes less significant.
+Finally, in the max-stable data setting, our model with $K=1$ partition has significantly better Brier scores than the Gaussian method for thresholds $q(0.92)$ -- $q(0.95)$ (p-value $<0.05$) and $q(0.95)$ -- $q(0.995)$ (p-value $<0.001$).
 
@@
 \end{align}
-The data from July 10 are shown in Figure \ref{fig:ozone} along with a Q-Q plot of the residuals compared to a skew-$t$ distribution with 8 d.f. and $\alpha = 1$.
-\begin{center}
+The data from July 10 are shown in Figure \ref{fig:ozone} along with a Q-Q plot of the residuals compared to a skew-$t$ distribution with 10 d.f. and $\alpha = 1$.
 \begin{figure}
-  \includegraphics[width=0.5\linewidth]{plots/ozone-10jul.pdf}
-  \includegraphics[width=0.5\linewidth]{plots/qq-res.pdf}
+  \centering
+  \includegraphics[width=0.55\linewidth]{plots/ozone-10jul-us.pdf}
+  \includegraphics[width=0.44\linewidth]{plots/qq-res.pdf}
   \caption{Ozone values on 10 July 2005 (left) Q-Q plot of the residuals (right)}
@@
 \end{figure}
-\end{center}
-The $\widehat{\chi}(h)$ and $\widehat{\chi}(t)$ plots estimate the spatial and temporal dependence using sample data.
-For $\widehat{\chi}(h)$, we take the sample proportion of all pairs of sites within distance $h$ of one another that are above a threshold $T$.
-\hl{I know I need another sentence or two here about $\widehat{\chi}(t)$}
-indicate that the residuals still demonstrate spatial and temporal dependence (see Figure \ref{fig:chi-st})
-\begin{center}
+
+We explore spatial and temporal extremal dependence by considering $\chi_c = \Pr[Y(\bs) > c | Y(\bt) > c]$.
+To examine spatial dependence in high quantiles, we consider observations at all pairs of sites $\bs$ and $\bt$ that are distance $h$ apart where $h$ is separated into bins of size 0.25 km.
+Then conditioned on $Y(\bt) > c$, we take the sample proportion of $Y(\bs) > c$.
+Finally, $\widehat{\chi}_c(h)$ is averaged over all days at each of the three threshold quantiles.
+To examine temporal dependence in high quantiles, we consider observations at a single site that are taken lag-$t$ days apart.
+Then conditioned on $Y_n(\bs) > c$, we take the sample proportion of $Y_{n + t}(\bs) > c$.
+Finally, $\widehat{\chi}(t)$ is averaged over all sites at each of the three threshold quantiles.
+The $\widehat{\chi}_c(h)$ and $\widehat{\chi}_c(t)$ plots in Figure \ref{fig:chi-st} show the estimated spatial and temporal dependence of the residuals for the ozone data at three quantile levels $q(0.90), q(0.95)$, and $q(0.99)$.
+Both plots indicate that there is dependence in the high quantile levels beyond what we expect if the residuals were independent.
 \begin{figure}
-  \includegraphics[width=0.5\linewidth]{plots/chi-h-ozone.pdf}
-  \includegraphics[width=0.5\linewidth]{plots/chi-t-ozone.pdf}
-  \caption{$\widehat{\chi}(h)$ plot for the residuals.
-The vertical line indicates the distance after which observations no longer demonstrate dependence (left) $\widehat{\chi}(t)$ plot for the residuals (right)}
+  \centering
+  \includegraphics[width=0.49\linewidth]{plots/chi-h-ozone.pdf}
+  \includegraphics[width=0.49\linewidth]{plots/chi-t-ozone.pdf}
+  \caption{$\widehat{\chi}_c(h)$ plot for the residuals (left). $\widehat{\chi}_c(t)$ plot for the residuals (right).}
   \label{fig:chi-st}
 \end{figure}
-\end{center}
 
 \subsection{Model comparisons}
-We fit the model using Gaussian and skew-$t$ marginal distributions, $K=1, 5, 10, 15$ partitions, with $Y(\bs)$ censored at $T = 0, 50, 75, 90$ ppb as described in Section \ref{s:data}.
-We also include a max-stable analysis using the method by ????
+The data analysis was done in two stages.
+In the first stage, we fit the model using Gaussian and skew-$t$ marginal distributions with $K=1, 5, 10, 15$ partitions.
+We chose to censor $Y(\bs)$ at $T = 0, 50, 75, 85$ ppb in order to compare results from no, moderate, and high censoring.
+We also include a max-stable analysis using the hierarchical max-stable model of \citet{Reich2012}.
 All methods assume the mean function given in (\ref{eq:datamean}).
-For each model, Brier scores and quantile scores were were averaged over all sites and days to obtain a single quantile score for each dataset.
+For each model, Brier scores were were averaged over all sites and days to obtain a single Brier score for each dataset.
 At a particular threshold or quantile level, the model that fits the best is the one with the lowest score.
-Looking at the heatplot in Figure \ref{fig:heatplot}, we can see that the methods using no thresholding tend to outperform those methods that do use thresholding.
+Based on the results from the first stage, we conducted a second stage analysis with $K = 6, 7, 8, 9$.
 
 \subsection{Results}\label{s:results}
+The plot in Figure \ref{fig:bs-ozone} shows the relative Brier scores (see Section \ref{s:simresults}) for the first and second stages of the analysis.
 \begin{figure}
-  \includegraphics[width=0.5\linewidth]{plots/heatplot.pdf}
-  \includegraphics[width=0.5\linewidth]{plots/qscore-best.pdf}
-  \caption{Heatplot of quantile scores for the 99th quantile (left) Quantile scores of Gaussian, skew-$t$ (K=10, T=q(0)), and Max-stable (right)}
-  \label{fig:heatplot}
+  \includegraphics[width=0.5\linewidth]{plots/bs-ozone-1.pdf}
+  \includegraphics[width=0.5\linewidth]{plots/bs-ozone-2.pdf}
+  \caption{Relative Brier scores for stage 1 ozone analysis (left) and stage 2 ozone analysis (right)}
+  \label{fig:bs-ozone}
 \end{figure}
+As with the simulation results, the Brier scores from the thresholded models were higher than those from the non-thresholded models, so we do not include them here.
+Based upon these results, it would appear that our model using a partition with anywhere from $K = 5$ to $K = 10$ partitions gives Brier scores across all threshold levels that are consistently low.
 
-\section{Conclusions}\label{s:con}
+\section{Discussion}\label{s:con}
+Given the nature
+
+There are also questions that remain regarding this model
 

### 2015-01-05 pot/LaTeX/plots/plots.R

@@ -20,10 +20,10 @@ tau.beta.t  <- 8
 
 # covariate data
 s         <- cbind(runif(500), runif(500))
-ns        <- nrow(s)  
+ns        <- nrow(s)
 nt        <- 50
 nsets     <- 1
-nsettings <- length(mixprob.t) 
+nsettings <- length(mixprob.t)
 ntest     <- floor(ns / 2)
 
 x <- array(1, c(ns, nt, 3))
@@ -48,13 +48,13 @@ for (setting in 1:nsettings) {
                      mixprob=mixprob.t[setting], z.alpha=z.alpha.t[setting],
                      tau.alpha=tau.alpha.t, tau.beta=tau.beta.t,
                      nknots=nknots.t[setting])
-    
+
     y[, , set, setting]        <- data$y
     tau.t.setting[, , set]     <- data$tau
     z.t.setting[, , set]       <- data$z
     knots.t.setting[, , , set] <- data$knots
   }
-    
+
   tau.t[[setting]]   <- tau.t.setting
   z.t[[setting]]     <- z.t.setting
   knots.t[[setting]] <- knots.t.setting
@@ -127,7 +127,7 @@ for (thresh in 1:length(threshs)){
 par(mfrow=c(1, 2))
 xplot <- bins[-11]
 plot(xplot, exceed[, 1], type="b", ylim=c(0, 0.75), ylab="exceed", xlab="bin distance", pch=1, lty=3, main="chi-plot: t, 1 partition")
-for (line in 2:9) { 
+for (line in 2:9) {
 	lines(xplot, exceed[, line], lty=3)
 	points(xplot, exceed[, line], pch=line)
 }
@@ -160,7 +160,7 @@ for (thresh in 1:length(threshs)){
 
 xplot <- bins[-11]
 plot(xplot, exceed[, 1], type="b", ylim=c(0, 0.75), ylab="exceed", xlab="bin distance", pch=1, lty=3, main="chi-plot: t, 5 partitions")
-for (line in 2:9) { 
+for (line in 2:9) {
 	lines(xplot, exceed[, line], lty=3)
 	points(xplot, exceed[, line], pch=line)
 }
[…]

### 2015-01-05 pot/code/analysis/simstudy/results.R

@@ -203,7 +203,7 @@ for (i in 1:length(done.groups)){
   idx <- (i - 1) * 5 + seq(1:5)
   done.sets[idx] <- (done.groups[i] - 1) * 5 + seq(1:5)
 }
-nsettings <- dim(y)[4]
+nsettings <- 6
 nmethods <- 5
 obs <- c(rep(T, 100), rep(F, 44))
 
@@ -271,7 +271,7 @@ rm(quant.score.all, brier.score.all, beta.0.all, beta.1.all, beta.2.all,
 ns <- dim(y)[1]
 nt <- dim(y)[2]
 nsets <- 5
-nsettings <- dim(y)[4]
+nsettings <- 6
 nmethods <- 5
 
 # get single brier scores and quantile scores for each setting x method x quantile
@@ -315,14 +315,19 @@ for (i in 1:length(probs)) { for (j in 1:(nmethods-1)) { for (k in 1:nsettings) 
 round(paired.results, 4)
 
 # wilcoxon signed-rank test for Brier scores
-wilcox.results <- array(NA, dim=c(length(probs), (nmethods-1), nsettings))
-best.methods <- vector(mode="list")
+wilcox.results.gau <- array(NA, dim=c(length(probs), (nmethods-1), nsettings))
 for (i in 1:length(probs)) { for (k in 1:nsettings) {
-  ref.j <- best.brier[i, k]
-  cmp.j <- (1:5)[-ref.j]
-  for (j in 1:(nmethods-1)) {
-    wilcox.results[i, j, k] <- wilcox.test(brier.score[i, , ref.j, k], brier.score[i, , cmp.j[j], k], paired=T)$p.value
+  ref.j <- 1
+  for (j in 2:nmethods) {
+    wilcox.results.gau[i, (j-1), k] <- wilcox.test(brier.score[i, , ref.j, k], brier.score[i, , j, k],
+                                               paired=T, alternative="greater")$p.value
   }
+}}
+
+wilcox.results.5 <- matrix(NA, nrow=length(probs), ncol=nsettings)
+for (i in 1:length(probs)) { for (k in 1:nsettings) {
+  wilcox.results.5[i, k] <- wilcox.test(brier.score[i, , 2, k], brier.score[i, , 4, k],
+                                               paired=T)$p.value
 }}
 
 # get single brier scores and quantile scores for each setting x method x quantile
[…]

### 2015-01-05 pot/LaTeX/skewnormal.tex
manuscript updates

@@
 \section{Discussion}\label{s:con}
-Given the nature
+In this paper we propose a new approach for spatitemporal modeling of extreme values.
+The proposed model gives flexible tail behavior, demonstrates asymptotic dependence for observations at sites that are near to one another, and has computation on the order of Gaussian models for large space-time datasets.
+In the simulation study, we demonstrate that there are statistically significant improvements over a na\"{i}ve Gaussian approach.
+In both the simulation study, and the application to ozone data, we find that incorporating a partition in the model improves extreme prediction.
+We also find that censoring the data below a threshold does not improve the results, and in many cases makes them worse.
 
-There are also questions that remain regarding this model
+This model presents new avenues for future research.
+One possibility is the implementation of a different partition structure.
+We choose to define the random effects for a site by using an indicator function based on closeness to a knot.
+However, this indicator function could be replaced by kernel function that would allow for multiple knots to impact each site, with the weight of each knot to be determined by some characteristic such as distance.
+Another area that should be explored is the temporal dependence in the model.
+Instead of implementing a time series on the random effects, a three-dimensional covariance structure on the residuals could be implemented to address temporal dependence.
 

### 2015-01-06 prelim/oral_prelim.tex
moving files around for oral prelim

@@
     \end{itemize}
-    \item Bayesian hierarchical model
+    \item Bayesian hierarchical model (Reich and Shaby, 2012)
     \item Pairwise likelihood approach (Huser and Davison, 2014)
@@
 
-\begin{frame}{Thresholding data}
+\begin{frame}{Censoring data}
   \begin{itemize} \setlength{\itemsep}{0.5em}
-    \item We threshold the observed data at a high threshold $T$.
-    \item Thresholded data:
+    \item We censor the observed data at a high threshold $T$.
+    \item Censored data:
     \begin{align*}
-      Y_t^*(\bs) = \left\{ \begin{array}{ll}
-          Y_t(\bs) \quad & Y_t(\bs) > T\\
-          T \quad & Y_t(\bs) \le T
+      \tilde{Y}_t(\bs) = \left\{ \begin{array}{ll}
+          Y_t(\bs) \quad & \delta(\bs) = 1\\
+          T \quad & \delta(\bs) = 0
       \end{array}\right.
     \end{align*}
+    where $\delta(\bs) = I[Y(\bs) > T]$
     \item Allows tails of the distribution to speak for themselves.
@@
     \begin{align*}
-      \chi(\bh) = \lim_{c \rightarrow \infty}\Pr(Y(\bs) > c \mid Y(\bs + \bh) > c)
+      \chi(h) = \lim_{c \rightarrow \infty}\Pr(Y(\bs) > c \mid Y(\bt) > c)
     \end{align*}
-    \item If $ \chi(\bh) = 0$, then observations are asymptotically independent at distance $\bh$.
+    where $h = ||\bs - \bt||$
+    \item If $ \chi(h) = 0$, then observations are asymptotically independent at distance $\bh$.
     \item We expect $\lim_{\bh \rightarrow \infty}\chi(\bh)=0$.
@@
     \begin{figure}
-    \includegraphics[width=0.54\linewidth]{./plots/example-partition-1.pdf}
-    \includegraphics[width=0.54\linewidth]{./plots/example-partition-2.pdf}
+    \includegraphics[width=0.54\linewidth]{./plots/pot/example-partition-1.pdf}
+    \includegraphics[width=0.54\linewidth]{./plots/pot/example-partition-2.pdf}
     \caption{Two sample partitions (number is at partition center)}
@@
   \begin{figure}
-  \includegraphics[width=1\linewidth]{./plots/chi-plots.pdf}\\[-0.25in]
-  \includegraphics[width=0.2\linewidth]{./plots/chi-legend.pdf}
-  \caption{$\chi$ plots from simulated datasets}
+  \includegraphics[width=1\linewidth]{./plots/pot/chi-h.pdf}\\[-0.25in]
+  \caption{$\chi$ plot for different data settings}
   \end{figure}
@@
   \begin{figure}
-  \includegraphics[width=1\linewidth]{./plots/gauss-vs-t3.pdf}
+  \includegraphics[width=1\linewidth]{./plots/pot/gauss-vs-t3.pdf}
   \caption{Gaussian and $t$ with 3 partitions}
@@
   \begin{figure}
-  \includegraphics[width=1\linewidth]{./plots/gauss-vs-skew-t3.pdf}
+  \includegraphics[width=1\linewidth]{./plots/pot/gauss-vs-skew-t3.pdf}
   \caption{Gaussian and skew-$t$ with 3 partitions}
@@
 
+\begin{frame}{Simulation study}
+  \begin{itemize} \setlength{\itemsep}{0.5em}
+    \item 6 different data settings:
+    \begin{itemize}
+      \item Gaussian vs $t$ vs skew-$t$ marginal distribution
+        \item $K=1$ partition vs $K=5$ partitions
+    \end{itemize}
+  \end{itemize}
+\end{frame}
+
+\begin{frame}{Brier score results}
+\centering
+
+\end{frame}
+
 \begin{frame}{Data analysis}
@@
 	\begin{itemize} \setlength{\itemsep}{0.5em}
-	\item Data analysis uses
+	\item Ozone measurements
 	\begin{itemize}
 		\item max 8-hour ozone measurements
-		\item 85 sites
-		\item 92 days
+		\item data from 1089 sites
+		\item July 2005
+	\end{itemize}
+  \item We take a stratified sample of $n = 800$ sites:
+  \begin{itemize}
+    \item 271 from northeast
+    \item 96 from northwest
+    \item 269 from southeast
+    \item 164 from southwest
   \end{itemize}
@@
 	\begin{figure}
-    \includegraphics[width=1\linewidth]{./plots/ozone_station.pdf}
+    \includegraphics[width=1\linewidth]{./plots/pot/ozone_stations.pdf}
     \caption{Ozone monitoring station locations}
@@
   \begin{figure}
-    \includegraphics[width=1\linewidth]{./plots/ozone-day.pdf}
-    \caption{Max 8-hour ozone measurements at 85 sites in NC, SC, and GA for days 5 and 34}
+    \includegraphics[width=1\linewidth]{./plots/pot/ozone-10jul-us.pdf}
+    \caption{Max 8-hour ozone measurements on July 10, 2005}
    \end{figure}
@@
   \begin{figure}
-    \includegraphics[width=1\linewidth]{./plots/chi-plot-ozone-res.pdf}
+    \includegraphics[width=1\linewidth]{./plots/pot/chi-plot-ozone-res.pdf}
     \caption{$\widehat{\chi}$-plot for sample quantiles of ozone observations}
@@
 
-\begin{frame}{Quantile score for cross-validation}
-  \begin{itemize} \setlength{\itemsep}{0.5em}
-    \item The quantile score for the $\tau$th quantile is
-    \begin{align*}
-      2 \{ I[y < \widehat{q}(\tau)] - \tau\} (\widehat{q} - y)
-    \end{align*}
-    where:
-    \begin{itemize}
-      \item $y$ is a test set value
-      \item $\widehat{q}(\tau)$ is the estimated $\tau$th quantile
-    \end{itemize}
-  \end{itemize}
-\end{frame}
+% \begin{frame}{Quantile score for cross-validation}
+%   \begin{itemize} \setlength{\itemsep}{0.5em}
+%     \item The quantile score for the $\tau$th quantile is
+%     \begin{align*}
+%       2 \{ I[y < \widehat{q}(\tau)] - \tau\} (\widehat{q} - y)
+%     \end{align*}
+%     where:
+%     \begin{itemize}
+%       \item $y$ is a test set value
+%       \item $\widehat{q}(\tau)$ is the estimated $\tau$th quantile
+%     \end{itemize}
+%   \end{itemize}
+% \end{frame}
 
@@
 $t$		& 1 & 0 	& {\bf 31.008} & {\bf 13.898} & 7.229 & {\bf 3.405} & 0.879\\
-$t$		& 3 & 0	& 31.213 & 13.920 & {\bf 7.218} & 3.498 & 0.918\\
+$t$		& 5 & 0	& 31.213 & 13.920 & {\bf 7.218} & 3.498 & 0.918\\
 $t$		& 1 & 0.9	& 32.221 & 14.519 & 7.549 & 3.604 & 0.896\\
-$t$		& 3 & 0.9 	& 38.842 & 16.781 & 8.434 & 4.180 & 1.020\\
+$t$		& 5 & 0.9 	& 38.842 & 16.781 & 8.434 & 4.180 & 1.020\\
 skew-$t$	& 1 & 0	& 31.845 & 14.542 & 7.533 & 3.645 & {\bf 0.844}\\
@@
 
-\begin{frame}{Predicted 95th quantile}
-\centering
-\begin{figure}
-    \includegraphics[width=.5\linewidth]{./plots/quantile-95-gau.pdf}
-    \includegraphics[width=.5\linewidth]{./plots/quantile-95-t10.pdf}
-    \caption{Predicted 95th quantile using Gaussian and $t$}
-\end{figure}
-\end{frame}
+% % TODO: Update with new plots
+% \begin{frame}{Predicted 95th quantile}
+% \centering
+% \begin{figure}
+%     \includegraphics[width=.5\linewidth]{./plots/quantile-95-gau.pdf}
+%     \includegraphics[width=.5\linewidth]{./plots/quantile-95-t10.pdf}
+%     \caption{Predicted 95th quantile using Gaussian and $t$}
+% \end{figure}
+% \end{frame}
 
-\begin{frame}{Predicted 95th quantile}
-\centering
-\begin{figure}
-    \includegraphics[width=.5\linewidth]{./plots/quantile-95-t10.pdf}
-    \includegraphics[width=.5\linewidth]{./plots/quantile-95-t19.pdf}
-    \caption{Predicted 95th quantile using $t$ and $t$ thresholded at $T=0.9$}
-\end{figure}
-\end{frame}
+% \begin{frame}{Predicted 95th quantile}
+% \centering
+% \begin{figure}
+%     \includegraphics[width=.5\linewidth]{./plots/quantile-95-t10.pdf}
+%     \includegraphics[width=.5\linewidth]{./plots/quantile-95-t19.pdf}
+%     \caption{Predicted 95th quantile using $t$ and $t$ thresholded at $T=0.9$}
+% \end{figure}
+% \end{frame}
 
-\begin{frame}{Predicted 99th quantile}
-\centering
-\begin{figure}
-    \includegraphics[width=.5\linewidth]{./plots/quantile-99-gau.pdf}
-    \includegraphics[width=.5\linewidth]{./plots/quantile-99-t10.pdf}
-    \caption{Predicted 99th quantile using Gaussian and $t$}
-\end{figure}
-\end{frame}
+% \begin{frame}{Predicted 99th quantile}
+% \centering
+% \begin{figure}
+%     \includegraphics[width=.5\linewidth]{./plots/quantile-99-gau.pdf}
+%     \includegraphics[width=.5\linewidth]{./plots/quantile-99-t10.pdf}
+%     \caption{Predicted 99th quantile using Gaussian and $t$}
+% \end{figure}
+% \end{frame}
 
-\begin{frame}{Predicted 99th quantile}
-\centering
-\begin{figure}
-    \includegraphics[width=.5\linewidth]{./plots/quantile-99-t10.pdf}
-    \includegraphics[width=.5\linewidth]{./plots/quantile-99-t19.pdf}
-    \caption{Predicted 99th quantile using $t$ and $t$ thresholded at $T=0.9$}
-\end{figure}
-\end{frame}
+% \begin{frame}{Predicted 99th quantile}
+% \centering
+% \begin{figure}
+%     \includegraphics[width=.5\linewidth]{./plots/quantile-99-t10.pdf}
+%     \includegraphics[width=.5\linewidth]{./plots/quantile-99-t19.pdf}
+%     \caption{Predicted 99th quantile using $t$ and $t$ thresholded at $T=0.9$}
+% \end{figure}
+% \end{frame}
 
-\begin{frame}{Probability of exceedance}
-\centering
-\begin{figure}
-    \includegraphics[width=.5\linewidth]{./plots/p-exceed-std-gau.pdf}
-    \includegraphics[width=.5\linewidth]{./plots/p-exceed-std-t10.pdf}
-    \caption{Probability of exceeding the 75 ppb ozone standard using Gaussian and $t$}
-\end{figure}
-\end{frame}
+% \begin{frame}{Probability of exceedance}
+% \centering
+% \begin{figure}
+%     \includegraphics[width=.5\linewidth]{./plots/p-exceed-std-gau.pdf}
+%     \includegraphics[width=.5\linewidth]{./plots/p-exceed-std-t10.pdf}
+%     \caption{Probability of exceeding the 75 ppb ozone standard using Gaussian and $t$}
+% \end{figure}
+% \end{frame}
 
-\begin{frame}{Probability of exceedance}
-\centering
-\begin{figure}
-    \includegraphics[width=.5\linewidth]{./plots/p-exceed-std-t10.pdf}
-    \includegraphics[width=.5\linewidth]{./plots/p-exceed-std-t19.pdf}
-    \caption{Probability of exceeding the 75 ppb ozone standard using $t$ and $t$ thresholded at $T=0.9$}
-\end{figure}
-\end{frame}
-
-\begin{frame}{Simulation study}
-  \begin{itemize} \setlength{\itemsep}{0.5em}
-    \item 6 different data settings:
-    \begin{itemize}
-    	\item Gaussian vs $t$ vs skew-$t$ marginal distribution
-        \item $K=1$ partition vs $K=5$ partitions
-    \end{itemize}
-    \item Preliminary results are inconclusive
-  \end{itemize}
-\end{frame}
+% \begin{frame}{Probability of exceedance}
+% \centering
+% \begin{figure}
+%     \includegraphics[width=.5\linewidth]{./plots/p-exceed-std-t10.pdf}
+%     \includegraphics[width=.5\linewidth]{./plots/p-exceed-std-t19.pdf}
+%     \caption{Probability of exceeding the 75 ppb ozone standard using $t$ and $t$ thresholded at $T=0.9$}
+% \end{figure}
+% \end{frame}
 

### 2015-01-06 pot/code/R/mem_cpp.R

@@ -1,9 +1,9 @@
 library(inline)
 code <- '
   arma::mat d = Rcpp::as<arma::mat>(d_knots);
-  int ns = d.n_rows; int nknots = d.n_cols; 
+  int ns = d.n_rows; int nknots = d.n_cols;
   arma::uword index; double min;
-  arma::rowvec d_temp(nknots);  
+  arma::rowvec d_temp(nknots);
   NumericVector g(ns);
   for (int i = 0; i < ns; i++) {
     d_temp = d.row(i);

### 2015-01-06 pot/code/analysis/ozone/US-all/us-all-setup.R

@@ -20,6 +20,10 @@ image.plot(x, y, matrix(CMAQ[, 5], nx, ny), main="CMAQ output (ppb) - 2005-07-0
 points(S)       # Locations of monitoring stations
 lines(borders)  # Add state lines
 
+# for presentation
+plot(S, axes=F, ylab="", xlab="", ylim=c(-1600, 1300), xlim=c(-2300, 2400))
+lines(borders)
+
 #### start data preprocessing
 # only include sites from the eastern US
 # keep.these <- which(S[, 1] > 0)

### 2015-01-06 prelim/oral_prelim.tex
Oral prelim updates

@@
 
+\begin{frame}{Standard analysis - Block maxima}
+  \begin{itemize} \setlength{\itemsep}{0.5em}
+    \item Uses yearly maxima
+    \item Discards many observations
+    \item Models are fit using the generalized extreme value distribution with parameters $\mu, \sigma,$ and $\xi$
+    \begin{align*}
+      \Pr(Y < y) = \left\{  \begin{array}{ll}
+        \exp\left\{ -\left[ 1 + \xi \left( \frac{ y - \mu }{ \sigma } \right) \right]^{ -1 / \xi} \right\} & \quad \xi \neq 0 \\
+        \exp \left\{ -\exp \left( - \frac{ y - \mu }{ \sigma} \right) \right\} & \quad \xi = 0
+      \end{array}\right.
+    \end{align*}
+    \item Standardized distribution is unit Fr\'{e}chet or GEV(1, 1, 1)
+    \begin{align*}
+      \Pr(Z < z) = exp(-z^{-1})
+    \end{align*}
+  \end{itemize}
+\end{frame}
+
+\begin{frame}{Standard analysis - Peaks over threshold}
+  \begin{itemize}  \setlength{\itemsep}{0.5em}
+      \item Incorporates more data than block maxima
+      \item Select a threshold, $T$, and use the Generalized Pareto distribution (GPD) to model the exceedances
+      \item The generalized Parety distribution has three parameters $\mu, \sigma$, and $\xi$
+      \begin{align*}
+        P(Y < y) = \left\{ \begin{array}{ll}
+          1 - \left[1 - \xi \left( \frac{ y - \mu }{ \sigma } \right) \right]^{-1 / \xi} & \quad \xi \neq 0 \\
+          1 - \exp \left\{ \frac{ y - \mu }{ \sigma} \right\} & \quad \xi = 0
+        \end{array}\right.
+      \end{align*}
+      \item Temporal dependence may be an issue between observations (e.g. flood levels don't dissipate overnight)
+  \end{itemize}
+\end{frame}
+
 \begin{frame}{Introduction to extremes}
   \begin{itemize} \setlength{\itemsep}{0.5em}
-    \item Max-stable processes (Cooley et al., 2012):
+    \item For a spatial analysis, max-stable processes give an appropriate limiting distribution (Cooley et al., 2012):
     \begin{itemize}
@@
 
-\begin{frame}{Standard analysis - Block maxima}
-  \begin{itemize} \setlength{\itemsep}{0.5em}
-      \item Uses yearly maxima
-      \item Discards many observations
-      \item Models are fit using the generalized extreme value distribution
-    \item For a spatial analysis, max-stable processes give an appropriate limiting distribution
-  \end{itemize}
-\end{frame}
-
-\begin{frame}{Standard analysis - Peaks over threshold}
-  \begin{itemize}  \setlength{\itemsep}{0.5em}
-      \item Incorporates more data than block maxima
-      \item Select a threshold, $T$, and use the Generalized Pareto distribution (GPD) to model the exceedances
-      \item Temporal dependence may be an issue between observations (e.g. flood levels don't dissipate overnight)
-  \end{itemize}
-\end{frame}
-
 \begin{frame}{Multivariate representations}
@@
       \item Assume common standardized max-stable marginal, like unit-Fr\'{e}chet
-      \begin{align*}
-        \Pr(Z < z) = exp(-z^{-1})
-      \end{align*}
+
       \item The multivariate representation for the GEV is
@@
 
-\begin{frame}{Model objectives}
+\begin{frame}{Three principal contributions}
+  \begin{enumerate}[1] \setlength{\itemsep}{0.5em}
+    \item A spatio-temporal model with flexible tails, asymptotic spatial dependence, and computation on the order of Gaussian models for large space-time datasets
+    \item Predicting exceedances using a spatially dependent generalized extreme-value link function
+    \item A Bayesian hierarchical model to allow for non-stationary covariance in extreme value models.
+  \end{enumerate}
+\end{frame}
+
+\begin{frame}{Spatiotemporal modeling for extreme values}
   \begin{itemize} \setlength{\itemsep}{0.5em}
-    \item Our objective is to build a model that
+    \item Model to analysis spatiotemporal extreme values
+    \item Model objectives
     \begin{itemize}
@@
 
-\begin{frame}{Censoring data}
+\begin{frame}{Spatial skew-$t$ distribution}
+  \begin{itemize} \setlength{\itemsep}{0.5em}
+    \item Assume observed data $Y(\bs)$ come from a skew-$t$ (Zhang and El-Shaarawi, 2012)
+    \begin{align*}
+      Y(\bs) = \bX(\bs)\bbeta + \lambda |z| + v(\bs)
+    \end{align*}
+    where
+    \begin{itemize} \setlength{\itemsep}{0.25em}
+      \item $\lambda \in \calR$ controls the skewness
+      \item $z \sim N(0, \sigma^2)$ is a random effect
+      \item $v(\bs)$ is a Gaussian process with variance $\sigma^2$ and \Matern correlation
+      \item $\sigma^2 \sim \text{IG}(a, b)$
+    \end{itemize}
+  \end{itemize}
+\end{frame}
+
+\begin{frame}{Spatial skew-$t$ distribution}
+  \begin{itemize} \setlength{\itemsep}{0.5em}
+   \item \alert{Conditioned} on $z$ and $\sigma^2$, $Y(\bs)$ is a Gaussian spatial model
+    \item Can use standard geostatistical methods to fit this model
+    \item Predictions can be made through Kriging
+  \end{itemize}
+\end{frame}
+
+\begin{frame}{Spatial skew-$t$ distribution}
+  \begin{itemize} \setlength{\itemsep}{0.5em}
+    \item \alert{Marginalizing} over $z$ and $\sigma^2$ (via MCMC),
+    \begin{align*}
+      Y(\bs) \sim \text{skew-t}(\bX(\bs), \bOmega, \alpha, \text{df}=2a)
+    \end{align*}
+    where
+    \begin{itemize}
+      \item $\bX(\bs) \bbeta$ is the location
+      \item $\bOmega = \bomega \bar{\bOmega} \bomega$ is a correlation matrix
+      \item $\bomega = \text{diag} \left(\frac{1}{\sqrt{ab}}, \ldots, \frac{1}{ \sqrt{ab}}\right)$
+      \item $\bar{\bOmega} = (\bSigma + \lambda^2 \bOne \bOne^T)$
+      \item $\Sigma$ is a postive definite correlation matrix
+      \item $\alpha = \lambda( 1 + \lambda^2 \bOne^T \bSigma^{-1} \bOne)^{-1/2} \bOne^T \bSigma^{-1}$ controls the skewness
+    \end{itemize}
+  \end{itemize}
+\end{frame}
+
+\begin{frame}{Censoring data to focus on tail behavior}
   \begin{itemize} \setlength{\itemsep}{0.5em}
@@
 
-\begin{frame}{$\chi$ coefficient}
+\begin{frame}{$\chi$ statistic}
   \begin{itemize} \setlength{\itemsep}{0.5em}
-   \item The $\chi$ coefficient is a measure of extremal dependence
+   \item The $\chi$ statistic is a measure of extremal dependence
    \item Specifically, we focus on $\chi(\bh)$ for the upper tail given by
@@
     \item If $ \chi(h) = 0$, then observations are asymptotically independent at distance $\bh$.
-    \item We expect $\lim_{\bh \rightarrow \infty}\chi(\bh)=0$.
   \end{itemize}
@@
   \begin{itemize} \setlength{\itemsep}{0.5em}
-    \item Assume observed data $Y_t(\bs)$ come from a skew-$t$ (Zhang and El-Shaarawi, 2012)
-    \begin{align*}
-      Y_t(\bs) = X_t(\bs)\beta + \alpha z_t + v_t(\bs)
-    \end{align*}
-    where
-    \begin{itemize} \setlength{\itemsep}{0.25em}
-      \item $\alpha \in \calR$ controls the skewness
-      \item $z_t \iid N_{(0, \infty)}(0, \sigma^2_t)$ is a random effect
-      \item $v_t(\bs)$ is a Gaussian process with variance $\sigma^2_t$ and \Matern correlation
-      \item $\sigma_t^2 \iid \text{IG}(a, b)$
-    \end{itemize}
-  \end{itemize}
-\end{frame}
-
-\begin{frame}{Spatial skew-$t$ distribution}
-  \begin{itemize} \setlength{\itemsep}{0.5em}
-   \item \alert{Conditioned} on $z_t$ and $\sigma^2_t$, $Y_t(\bs)$ is a Gaussian spatial model
-    \item Can use standard geostatistical methods to fit this model
-    \item Predictions can be made through Kriging
-    \item \alert{Marginalizing} over $z_t$ and $\sigma^2_t$ (via MCMC),
-    \begin{align*}
-      Y_t(\bs) \sim \text{skew-t}(\mu, \Sigma^*, \alpha, \text{df}=2a)
-    \end{align*}
-    where
-    \begin{itemize}
-    	\item $\mu$ is the location
-	\item $a$, $b$ are the IG parameters for $\sigma^2_t$
-	\item $\Sigma^* = \frac{ b }{ a } \Sigma$ is a scale matrix, and $\Sigma$ is a \Matern covariance matrix
-	\item $\alpha \in \calR$ controls the skewness
-    \end{itemize}
-  \end{itemize}
-\end{frame}
-
-\begin{frame}{Spatial skew-$t$ distribution}
-  \begin{itemize} \setlength{\itemsep}{0.5em}
       \item Model properties
@@
     \item For this distribution, $\chi(\bh)$ shows asymptotic dependence that does not approach 0 as $\bh \rightarrow \infty$
-   \item This occurs because all observations (near and far) share the same $z_t$ and $\sigma_t^2$
-    \item We deal with this through a daily random partition (similar to Huser and Davison)
+   \item This occurs because all observations (near and far) share the same $z$ and $\sigma^2$
+    \item We deal with this through a daily random partition (similar to Kim et al., 2005)
   \end{itemize}
@@
 
-\begin{frame}{Daily random partition}
+\begin{frame}{Random partition}
   \begin{itemize} \setlength{\itemsep}{0.5em}
-    \item Daily random partition allows $z_t$ and $\sigma^2_t$ to vary by site
+    \item Daily random partition allows $z$ and $\sigma^2$ to vary by site
     \begin{align*}
-      Y_t(\bs) = X_t(\bs) \beta + \alpha z_t(\bs) + \sigma(\bs) v_t(\bs)
+      Y(\bs) = \bX(\bs) \bbeta + \lambda z(\bs) + \sigma(\bs) v(\bs)
     \end{align*}
-    \item Consider a set of daily knots $\bw_{tk} \sim$ Uniform that define a random daily partition
-    $P_{t1}, \ldots, P_{tK}$ such that
+    \item Consider a set of knots $\bw_{k} \sim$ Uniform that define a random partition
+    $P_{1}, \ldots, P_{K}$ such that
     \begin{align*}
-      P_{tk} = \{s : k = \argmin_\ell|| \bs - \bw_{t\ell}|| \}
+      P_{k} = \{s : k = \argmin_\ell|| \bs - \bw_{\ell}|| \}
     \end{align*}
-    \item For $\bs \in P_{tk}$
+    where $\bw = (w_1, w_2)$
+    \item For $\bs \in P_{k}$
     \begin{align*}
-      z_t(\bs) &= z_{tk}\\
-      \sigma^2_t(\bs) &= \sigma^2_{tk}
+      z(\bs) &= z_{k}\\
+      \sigma^2(\bs) &= \sigma^2_{k}
     \end{align*}
-    \item Within each partition $Y_t(\bs)$ has the same MV skew-t distribution as before
+    \item Within each partition $Y(\bs)$ has the same MV skew-t distribution as before
   \end{itemize}
@@
 
-\begin{frame}{Example daily partition}
+\begin{frame}{Example partition}
     \centering
@@
 
-\begin{frame}{Simulated $\widehat{\chi}(\bh)$ plots}
+\begin{frame}{$\chi(\bh)$ plot}
+  \vspace{-2em}
   \centering
   \begin{figure}
-  \includegraphics[width=1\linewidth]{./plots/pot/chi-h.pdf}\\[-0.25in]
+  \includegraphics[width=.7\linewidth]{./plots/pot/chi-h.pdf}\\[-0.4in]
   \caption{$\chi$ plot for different data settings}
@@
   \end{figure}
+\end{frame}
+
+\begin{frame}{Extension to space-time data}
+  \begin{itemize} \setlength{\itemsep}{0.5em}
+    \item We extend our previous model to be
+    \begin{align*}
+      Y_t(\bs) = \bX_t(\bs)^T \bbeta + \lambda \sigma_t(\bs) | z_t(\bs) | + \sigma_t(\bs) v_t(\bs)
+    \end{align*}
+    where $t = 1, \ldots, T$ denotes the day of each observation.
+    \item We incorporate an AR(1) time series on $\bw^*_{tk} = (w^*_{tk1}, w^*_{tk2})$, $z_{tk}$, and $\sigma^*_{tk}$ where
+    \begin{align*}
+      w^*_{tki} &= \Phi^{-1}\left[ \frac{w_{tki} - \min(\bs_i)}{ \text{range}(\bs_i)}\right] \quad i = 1, 2 \\
+      \sigma^{2*}_t(\bs) &= \Phi^{-1}\{ \text{IG}[\sigma^2_t(\bs)] \}
+    \end{align*}
+    are transformations to $\calR^2$
+  \end{itemize}
 \end{frame}
@@
     \end{itemize}
+    \item 5 different models:
+    \begin{itemize}
+      \item Gaussian vs skew-$t$ marginal distribution
+      \item $K = 1$ partition vs $K = 5$ partitions
+    \end{itemize}
+    \item Brier score used to determine model that gives best fit
   \end{itemize}
@@
 
-\begin{frame}{Brier score results}
+% \begin{frame}{Quantile score for cross-validation}
+%   \begin{itemize} \setlength{\itemsep}{0.5em}
+%     \item The quantile score for the $\tau$th quantile is
+%     \begin{align*}
+%       2 \{ I[y < \widehat{q}(\tau)] - \tau\} (\widehat{q} - y)
+%     \end{align*}
+%     where:
+%     \begin{itemize}
+%       \item $y$ is a test set value
+%       \item $\widehat{q}(\tau)$ is the estimated $\tau$th quantile
+%     \end{itemize}
+%   \end{itemize}
+% \end{frame}
+
+\begin{frame}{Brier score}
+  \begin{itemize} \setlength{\itemsep}{0.5em}
+  \item The Brier score for predicting exceedance of threshold $c$ is
+  \begin{align*}
+    [e(c) - P(c)]^2
+  \end{align*}
+  where
+  \begin{itemize}
+    \item $y$ is a test set value
+    \item $e(c) = I[y > c]$
+    \item $P(c)$ is the predicted probability of exceeding $c$
+  \end{itemize}
+  \item Relative Brier scores:
+  \begin{align*}
+    \text{BS}_\text{rel} = \frac{ \text{BS}_\text{method}}{ \text{BS}_\text{Gaussian}}
+  \end{align*}
+  \end{itemize}
+\end{frame}
+
+\begin{frame}{Simulation study results}
 \centering
+  \begin{figure}
+    \includegraphics[width=0.45\linewidth]{./plots/pot/bs-sim-gaus.pdf}
+    \includegraphics[width=0.45\linewidth]{./plots/pot/bs-sim-max.pdf} \\
+    \includegraphics[width=0.3\linewidth, trim=2in 3.25in 2in 3in]{./plots/pot/bs-sim-legend.pdf}
+    \caption{Relative Brier score results}
+  \end{figure}
+\end{frame}
 
+\begin{frame}{Simulation study results}
+\centering
+  \begin{figure}
+    \includegraphics[width=0.45\linewidth]{./plots/pot/bs-sim-t1.pdf}
+    \includegraphics[width=0.45\linewidth]{./plots/pot/bs-sim-st1.pdf} \\
+    \includegraphics[width=0.3\linewidth, trim=2in 3.25in 2in 3in]{./plots/pot/bs-sim-legend.pdf}
+    \caption{Relative Brier score results}
+  \end{figure}
+\end{frame}
+
+\begin{frame}{Simulation study results}
+\centering
+  \begin{figure}
+    \includegraphics[width=0.45\linewidth]{./plots/pot/bs-sim-t5.pdf}
+    \includegraphics[width=0.45\linewidth]{./plots/pot/bs-sim-st5.pdf} \\
+    \includegraphics[width=0.3\linewidth, trim=2in 3.25in 2in 3in]{./plots/pot/bs-sim-legend.pdf}
+    \caption{Relative Brier score results}
+  \end{figure}
+\end{frame}
+
+\begin{frame}{Simulation study results}
+  \begin{itemize} \setlength{\itemsep}{0.5em}
+    \item Key findings:
+    \begin{itemize}
+      \item Improvement over Gaussian methods when partitioning
+      \item Underestimating the number of knots has a detrimental impact
+      \item In all cases, non-thresholded models perform better than thresholded models
+    \end{itemize}
+  \end{itemize}
 \end{frame}
@@
   \begin{figure}
-    \includegraphics[width=1\linewidth]{./plots/pot/ozone-10jul-us.pdf}
+    \includegraphics[width=\linewidth, trim=0 1in 0 0in ]{./plots/pot/ozone-10jul-us.pdf}
     \caption{Max 8-hour ozone measurements on July 10, 2005}
@@
       \item Gaussian vs $t$ vs skew-$t$ marginal distribution
-      \item $K=1$ partition vs $K=3$ partitions
-      \item No thresholding vs thresholding at $T=0.90$ sample quantile
+      \item $K=1, 5, 6, 7, 8, 9, 10, 15$ partitions
+      \item Thresholding at $T=0, 50, 75$, and $85$ ppb
     \end{itemize}
     \item All methods use a \Matern or exponential covariance ($\nu = 0.5$)
-    \item Compare quantile and Brier scores using 5-fold cross validation (Gneiting and Raftery, 2007)
+    \item Compare quantile and Brier scores using two-fold cross validation (Gneiting and Raftery, 2007)
     \item Mean function modeled as
     \begin{align*}
-    	\beta_0 + \beta_1 \cdot \text{lat} + \beta_2 \cdot \text{long} +  \beta_3 \cdot \text{lat}^2 + \beta_4 \cdot \text{long}^2 + \beta_5 \cdot \text{lat} \cdot \text{long}
+    	\beta_0 + \beta_1 \cdot \text{CMAQ}
     \end{align*}
@@
 
-% \begin{frame}{Quantile score for cross-validation}
-%   \begin{itemize} \setlength{\itemsep}{0.5em}
-%     \item The quantile score for the $\tau$th quantile is
-%     \begin{align*}
-%       2 \{ I[y < \widehat{q}(\tau)] - \tau\} (\widehat{q} - y)
-%     \end{align*}
-%     where:
-%     \begin{itemize}
-%       \item $y$ is a test set value
-%       \item $\widehat{q}(\tau)$ is the estimated $\tau$th quantile
-%     \end{itemize}
-%   \end{itemize}
-% \end{frame}
-
-\begin{frame}{Brier score}
-  \begin{itemize} \setlength{\itemsep}{0.5em}
-	\item The Brier score for predicting exceedance of threshold $c$ is
-	\begin{align*}
-	  [e(c) - P(c)]^2
-	\end{align*}
-	where
-	\begin{itemize}
-		\item $y$ is a test set value
-		\item $e(c) = I[y > c]$
-		\item $P(c)$ is the predicted probability of exceeding $c$
-	\end{itemize}
-  \end{itemize}
+\begin{frame}{Two-fold cross-validation results}
+  \centering
+  \begin{figure}
+    \includegraphics[width=0.45\linewidth]{./plots/pot/bs-ozone-1.pdf}
+    \includegraphics[width=0.45\linewidth]{./plots/pot/bs-ozone-2.pdf} \\
+    \caption{Relative Brier score results}
+  \end{figure}
 \end{frame}
 
-\begin{frame}{Five-fold cross-validation results}
-  \begin{table}[htbp]
-    \small
-    \centering
-    \begin{tabular}{l|c|r|rrrrr}
-          \multicolumn{3}{c}{\ } & \multicolumn{5}{c}{$\tau$}\\
-           \hline
-  Marginal & $K$ & $T$  & 0.950 & 0.980 & 0.990 & 0.995 & 0.999\\
-  \hline
-Gaussian 	& 1 & 0 	& 39.820 & 17.539 & 9.167 & 4.720 & 1.057\\
-$t$		& 1 & 0 	& {\bf 31.008} & {\bf 13.898} & 7.229 & {\bf 3.405} & 0.879\\
-$t$		& 5 & 0	& 31.213 & 13.920 & {\bf 7.218} & 3.498 & 0.918\\
-$t$		& 1 & 0.9	& 32.221 & 14.519 & 7.549 & 3.604 & 0.896\\
-$t$		& 5 & 0.9 	& 38.842 & 16.781 & 8.434 & 4.180 & 1.020\\
-skew-$t$	& 1 & 0	& 31.845 & 14.542 & 7.533 & 3.645 & {\bf 0.844}\\
-skew-$t$	& 1 & 0.9	& 32.132 & 14.296 & 7.484 & 3.497 & 0.890\\
-skew-$t$	& 3 & 0	& 33.653 & 15.453 & 8.119 & 4.338 & 1.188\\
-skew-$t$	& 3 & 0.9 	& 32.157 & 14.727 & 7.794 & 3.825 & 0.917\\
-\hline
-    \end{tabular}
-    \caption{Brier score for predicting exceedance of $c = \hat{q}(\tau)$ from five-fold cross-validation ($\times 1000$)}
-  \end{table} \vspace{-1.5em}
+\begin{frame}{Two-fold cross-validation results}
+  \begin{itemize} \setlength{\itemsep}{0.5em}
+    \item Key findings:
     \begin{itemize}
-  	\item Quantile score results are similar
+      \item Partitioning improves performance across all high thresholds.
+      \item Models with anywhere from $K = 5$ to $K = 10$ partitions perform similarly
+      \item In all cases, non-thresholded models perform better than thresholded models
+    \end{itemize}
   \end{itemize}
@@
 
-\begin{frame}{Future Work}
+\begin{frame}{Discussion and future work}
   \begin{itemize} \setlength{\itemsep}{0.5em}
+    \item Improvement of model performance when using partitioned models
+    \item Thresholding makes results worse
+    \begin{itemize}
+      \item Possible numerical instability due to truncated normal distribution
+    \end{itemize}
     \item Different ways to incorporate the temporal dependence
@@
     	\item Three dimensional covariance model for $v_t(\bs)$ (e.g. Huser and Davison, 2014)
-    	\item Use a temporal structure for $z_t(\bs)$:
+    \end{itemize}
+    \item Different partition structure
     \begin{itemize}
-		\item AR(1)
-		\item Moving average
-		\item Association between $\bw_{t, k}$ and $\bw_{t+1, k}$
+      \item Distance weighting for each knot vs indicator functions
+    \end{itemize}
+    \item Knot selection
+    \begin{itemize}
+      \item Possible prior on the probability a knot is in the spatial domain
     \end{itemize}
   \end{itemize}
-     \item Comparison with extreme value analysis methods
+\end{frame}
+
+\begin{frame}{Spatial binary regression}
+  \begin{itemize} \setlength{\itemsep}{0.5em}
+    \item We observe $Y_i = I[Z(\bs_i) > T]$, an indicator variable that a continuous latent variable has exceeded a pre-specified threshold $T$.
+    \item We model $\Pr[Y_i = 1]$
+    \item Common examples:
+    \begin{itemize}
+      \item Logistic regression
+      \begin{align*}
+        \Pr[Y_i = 1] = \frac{ \exp(\bX(\bs_i) \bbeta) }{ 1 + \exp(\bX(\bs_i) \bbeta)}
+      \end{align*}
+      \item Probit regression
+      \begin{align*}
+        \Pr[Y_i = 1] = \Phi[\bX(\bs_i) \bbeta]
+      \end{align*}
+      where $\Phi$ is the standard normal distribution function
+    \end{itemize}
+  \end{itemize}
+\end{frame}
+
+\begin{frame}{Spatial binary regression}
+  \begin{itemize} \setlength{\itemsep}{0.5em}
+    \item Logistic and probit regression amount to modeling
+    \begin{align*}
+      \Pr[Y_i = 1] = g[\bX(\bs_i) \bbeta]
+    \end{align*}
+    where $g[\cdot]$ is a link function transforming from $\calR$ to $(0, 1)$.
+    \item Wang and Dey (2010): Generalized extreme value link function
+    \begin{align*}
+      g[\bx(\bs_i) \bbeta] = 1 - \exp\left[-(1 + \xi \bX_i \bbeta)^{-1 / \xi} \right]
+    \end{align*}
+  \end{itemize}
+\end{frame}
+
+\begin{frame}{Spatial binary regression}
+  \begin{itemize} \setlength{\itemsep}{0.5em}
+    \item Proposed method will
+    \begin{itemize}
+      \item use the GEV link function
+      \item use the hierarchical likelihood from Reich and Shaby (2012)
+    \end{itemize}
+    \item Model parameters fit using MCMC
+  \end{itemize}
+\end{frame}
+
+\begin{frame}{Spatial binary regression}
+  \begin{itemize} \setlength{\itemsep}{0.5em}
+    \item We fit parameters $\xi$ and $\beta$ in order to transform the data to GEV(1, 1, 1) marginal distributions.
+    \item Using the link function
+    \begin{align*}
+      p_i = 1 - \exp \left[-(1 + \xi \bX(\bs_i) \bbeta)^{-1/\xi}\right]
+    \end{align*}
+    we set the latent variable $z_i = -\frac{1}{\log(1 - p_i)}$
+    \item Then we evaluate the joint likelihood using $z_i$.
+  \end{itemize}
+\end{frame}
+
+\begin{frame}{Likelihood function}
+  \begin{itemize} \setlength{\itemsep}{0.5em}
+    \item We use a multivariate generalized extreme value distribution with asymetric Laplace deendence function given by
+    \footnotesize{
+    \begin{align*}
+      G(\bz) = \Pr[Z_1 < z_z, \ldots, Z_n < z_n] = \exp\left\{ - \sum_{l = 1}^{L} \left[ \sum_{i = 1}^n \left( \frac{ w_l(\bs_i) }{ z_i } \right)^{ 1 / \alpha} \right]^\alpha \right\}
+    \end{align*}
+    }
+    where
+    \begin{itemize}
+      \item $w_l$ is a weighting function subject to the constraint that $\sum_{l = 1}^L w_l = 1$.
+      \item $\alpha$ controls spatial dependence
+      \begin{itemize}
+        \item $\alpha = 0$ is strong dependence
+        \item $\alpha = 1$ is joint independence
+      \end{itemize}
+    \end{itemize}
+  \end{itemize}
+\end{frame}
+
+\begin{frame}{Weighting function}
+  \begin{itemize} \setlength{\itemsep}{0.5em}
+    \item We use the Gaussian weights proposed by Reich and Shaby (2012) given by
+    \footnotesize{
+    \begin{align*}
+      w_l(\bs_i) = \frac{ \exp\left[ -0.5 \left( \frac{ || \bs - \bv_l || }{ \rho} \right)^2 \right]}{ \sum_{l = 1}^L \exp\left[ -0.5 \left( \frac{ || \bs - \bv_l || }{ \rho} \right)^2 \right] }
+    \end{align*}
+    }
+    where
+    \begin{itemize}
+      \item $\bv_l$ are spatial knots
+      \item $\rho$ is a bandwidth term for the kernel function
+    \end{itemize}
+  \end{itemize}
+\end{frame}
+
+\begin{frame}{Joint likelihood}
+  \begin{itemize} \setlength{\itemsep}{0.5em}
+    \item Let $K_t = \sum_{i = 1}^n Y_{it}$ be the number of exceedances that occur on day $t$.
+    \item Rearrange the sites so
+    \begin{itemize}
+      \item $Y_1, \ldots, Y_K$ are the observations where $Y(\bs_i) = 1$
+      \item $Y_{K+1}, \ldots, Y_n$ are the observations where $Y(\bs_i) = 0$
+    \end{itemize}
+    \item Then for $K = 0, 1, 2$
+    {\scriptsize
+    \begin{align*}
+      \Pr(Y_1 = y_1, \ldots, Y_n = y_n) = \left\{ \begin{array}{ll}
+        G(\bz)  & K = 0\\
+        G(\bz_{(1)}) - G(\bz) & K = 1\\
+        G(\bz_{(12)}) - G(\bz_{(1)}) - G(\bz_{(2)}) + G(\bz) & K =2
+      \end{array}\right.
+    \end{align*}
+    }
+    where $G(\bz_{(1)} = \Pr(Z_2 < z_2, \ldots, Z_n < z_n)$.
+    \item $K > 2$ can be derived similarly
+  \end{itemize}
+\end{frame}
+
+% TODO: more about this model for sure
+\begin{frame}{Non-stationary covariance for extreme values}
+  \begin{itemize} \setlength{\itemsep}{0.5em}
+    \item Knot-specific bandwidth
+  \end{itemize}
+\end{frame}
+
+\begin{frame}{Joint likelihood}
+  \begin{itemize} \setlength{\itemsep}{0.5em}
+    \item For small $K$, we can evaluate the likelihood directly.
+    \item For large $K$, we use the hierarchical model of Reich and Shaby (2012).
+  \end{itemize}
+\end{frame}
+
+\begin{frame}{Thesis outline}
+  \begin{itemize} \setlength{\itemsep}{0.5em}
+    \item Chapter 1: Extreme value theory \alert{August 2015}
+    \item Chapter 2: Spatiotemporal model for extreme value analysis based on the skew-$t$ distribution \alert{February 2015}
+    \item Chapter 3: Spatial binary regression \alert{May / June 2015}
+    \item Chapter 4: Non-stationary covariance through knot-specific bandwidth \alert{August 2015}
   \end{itemize}

### 2015-01-07 prelim/oral_prelim.tex
updated prelim slides

@@
 
+\begin{frame}{Motivation}
+  \centering
+  \begin{figure}
+    \includegraphics[width=\linewidth, trim=0 1in 0 0in ]{./plots/pot/ozone-10jul-us.pdf}
+    \caption{Max 8-hour ozone measurements on July 10, 2005}
+   \end{figure}
+\end{frame}
+
+\begin{frame}{Defining extremes}
+\begin{columns}[c]
+\column{.45 \linewidth}
+  \begin{itemize} \setlength{\itemsep}{0.5em}
+    \item Key in extreme value analysis is to define extremes
+    \item Typically done in one of two ways
+    \begin{itemize}
+      \item Block maxima
+      \begin{itemize}
+        \item Red dots
+      \end{itemize}
+      \item Over threshold
+      \begin{itemize}
+        \item Values over threshold considered extreme
+      \end{itemize}
+    \end{itemize}
+  \end{itemize}
+
+  \column{.55\linewidth}
+  \begin{figure}
+    \includegraphics[width=1\linewidth]{./plots/define_extreme.pdf}
+    \caption{Monthly maximums recorded over a three years.}
+    \end{figure}
+\end{columns}
+\end{frame}
+
 \begin{frame}{Standard analysis - Block maxima}
   \begin{itemize} \setlength{\itemsep}{0.5em}
-    \item Uses yearly maxima
-    \item Discards many observations
-    \item Models are fit using the generalized extreme value distribution with parameters $\mu, \sigma,$ and $\xi$
+    \item Asymptotic result:
+    \begin{itemize}
+      \item Let $X_1, \ldots, X_n$ be i.i.d.
+      \item Consider $M_n = \max(X_1, \ldots, X_n)$
+      \item If there exist normalizing sequences $a_n > 0$ and $b_n \in \calR$ such that
       \begin{align*}
-      \Pr(Y < y) = \left\{  \begin{array}{ll}
-        \exp\left\{ -\left[ 1 + \xi \left( \frac{ y - \mu }{ \sigma } \right) \right]^{ -1 / \xi} \right\} & \quad \xi \neq 0 \\
+        a_n^{-1} (M_n - b_n) \converged G(z)
+      \end{align*}
+      then $G(z)$ follows a generalized extreme value distribution
+    \end{itemize}
+  \end{itemize}
+\end{frame}
+
+\begin{frame}{Standard analysis - Block maxima}
+  \begin{itemize} \setlength{\itemsep}{0.5em}
+    \item Generalized extreme value distribution has three parameters:
+    \begin{itemize}
+      \item $\mu \in \calR$ is a location parameter
+      \item $\sigma > 0$ is a scale parameter
+      \item $\xi \in \calR$ is a shape parameter
+      \begin{itemize}
+        \item Unbounded above if $\xi \ge 0$
+        \item Bouded above by $(\mu - \sigma) / \xi$ when $\xi < 0$
+      \end{itemize}
+    \end{itemize}
+    \item Challenge:
+    \begin{itemize}
+      \item Lose information by only considering maximum in a block
+    \end{itemize}
+  \end{itemize}
+\end{frame}
+
+\begin{frame}{Standard analysis - Block maxima}
+  \begin{itemize} \setlength{\itemsep}{0.5em}
+    \item Generalized extreme value distribution
+    \begin{align*}
+      G(y) = \Pr(Y < y) = \left\{  \begin{array}{ll}
+        \exp\left\{ -\left[ 1 + \xi \left( \frac{ y - \mu }{ \sigma } \right) \right]^{ -1 / \xi} \right\} & \quad \xi \neq 0 \\[0.5em]
         \exp \left\{ -\exp \left( - \frac{ y - \mu }{ \sigma} \right) \right\} & \quad \xi = 0
@@
   \begin{itemize} \setlength{\itemsep}{0.5em}
-      \item Incorporates more data than block maxima
-      \item Select a threshold, $T$, and use the Generalized Pareto distribution (GPD) to model the exceedances
-      \item The generalized Parety distribution has three parameters $\mu, \sigma$, and $\xi$
-      \begin{align*}
-        P(Y < y) = \left\{ \begin{array}{ll}
-          1 - \left[1 - \xi \left( \frac{ y - \mu }{ \sigma } \right) \right]^{-1 / \xi} & \quad \xi \neq 0 \\
-          1 - \exp \left\{ \frac{ y - \mu }{ \sigma} \right\} & \quad \xi = 0
-        \end{array}\right.
-      \end{align*}
-      \item Temporal dependence may be an issue between observations (e.g. flood levels don't dissipate overnight)
+    \item Generalized Pareto distribution has two parameters:
+    \begin{itemize}
+      \item $\sigma > 0$ is a scale parameter
+      \item $\xi \in \calR$ is a shape parameter
+      \begin{itemize}
+        \item Unbounded above if $\xi \ge 0$
+        \item Bouded above by $(\mu - \sigma) / \xi$ when $\xi < 0$
+      \end{itemize}
+    \end{itemize}
+    \item Challenges:
+    \begin{itemize}
+      \item Sensitive to threshold selection
+      \item Temporal dependence between observations (e.g. flood levels don't dissipate overnight)
+    \end{itemize}
   \end{itemize}
@@
 
-\begin{frame}{Introduction to extremes}
+\begin{frame}{Standard analysis - Peaks over threshold}
+  \begin{itemize}  \setlength{\itemsep}{0.5em}
+    \item Select a threshold, $T$, and use the Generalized Pareto distribution to model the exceedances
+    \begin{align*}
+      H(y) = P(Y < y) = \left\{ \begin{array}{ll}
+        1 - \left[1 - \xi \left( \frac{ y - T }{ \sigma } \right) \right]^{-1 / \xi} & \quad \xi \neq 0 \\[0.5em]
+        1 - \exp \left\{ \frac{ y - T }{ \sigma} \right\} & \quad \xi = 0
+      \end{array}\right.
+    \end{align*}
+    \item Related to GEV distribution through
+    \begin{align*}
+      H(y) = 1 + \log[G(y)], \quad y \ge T
+    \end{align*}
+  \end{itemize}
+\end{frame}
+
+\begin{frame}{Max-stable processes}
   \begin{itemize} \setlength{\itemsep}{0.5em}
@@
   \begin{itemize}
-    \item Multivariate distributions:
+    \item Marginally at each site, observations follow a generalized extreme value distribution
+    \item Finite collection of sites:
     \begin{itemize}
-      \item Assume common standardized max-stable marginal, like unit-Fr\'{e}chet
-
       \item The multivariate representation for the GEV is
@@
 
-\begin{frame}{Multivariate analysis}
+\begin{frame}{Multivariate GEV challenges}
+  \begin{itemize}
+    \item Only a few closed-form expressions for $V(\bz)$ exist (Stephenson, 2003)
+    \item Two common forms for $V(\bz)$:
+    \begin{itemize}
+      \item Symmetric logistic
+      \begin{align*}
+        V(\bz) = \left[\sum_{i = 1}^n \left( \frac{ 1 }{ z_i } \right)^{1/\alpha}\right]^\alpha
+      \end{align*}
+      \item Asymmetric logistic
+      \begin{align*}
+        V(\bz) = \sum_{l = 1}^L \left[\sum_{i = 1}^n \left(\frac{w_{il}}{z_i} \right)^{1 / \alpha_l} \right]^{\alpha_l}
+      \end{align*}
+      where $w_{il} \in [0, 1]$ and $\sum_{l = 1}^L w_{il} = 1$.
+    \end{itemize}
+  \end{itemize}
+\end{frame}
+
+\begin{frame}{Multivariate peaks over threshold}
+  \begin{itemize}
+    \item Not a lot of existing methods
+    \item Often use max-stable methods due to the relationship between GEV and GPD
+    \item Joint distribution function given by Falk et al. (2011)
+    \begin{align*}
+      H(\bz) = 1 - V(\bz)
+    \end{align*}
+    where $V(\bz)$ is defined as in the GEV
+  \end{itemize}
+\end{frame}
+
+\begin{frame}{Extremal dependence: $\chi$ statistic}
+  \begin{itemize} \setlength{\itemsep}{0.5em}
+   \item The $\chi$ statistic is a measure of extremal dependence in the tails
+   \item Specifically, we focus on $\chi(h)$ for the upper tail given by
+    \begin{align*}
+      \chi(h) = \lim_{c \rightarrow \infty}\Pr(Y(\bs) > c \mid Y(\bt) > c)
+    \end{align*}
+    where $h = ||\bs - \bt||$
+    \item If $ \chi(h) = 0$, then observations are asymptotically independent at distance $h$
+  \end{itemize}
+\end{frame}
+
+\begin{frame}{Existing challenges}
   \begin{itemize} \setlength{\itemsep}{0.5em}
@@
     \begin{itemize}
-      \item computationally challenging to work with
+      \item computationally challenging (Falk et al., 2011)
+      \begin{itemize}
+        \item Asymmetric logistic has $2^{n-1}(n + 2) - (2n + 1)$ free parameters
+      \end{itemize}
       \item joint distribution only available in low dimension
     \end{itemize}
+    \item Some recent approaches:
+    \begin{itemize}
       \item Bayesian hierarchical model (Reich and Shaby, 2012)
       \item Pairwise likelihood approach (Huser and Davison, 2014)
+    \end{itemize}
+    \item Many opportunities to explore new methods
   \end{itemize}
@@
     \item Model to analysis spatiotemporal extreme values
-    \item Model objectives
+    \item Model objectives:
     \begin{itemize}
-      \item has marginal distribution with a flexible tail
-      \item has asymptotic spatial dependence
-      \item has computation on the order of Gaussian models for large space-time datasets
+      \item Has marginal distribution with a flexible tail
+      \begin{itemize}
+        \item allow for asymmetric distributions
+        \item allow for heavy tails
+      \end{itemize}
+      \item Has asymptotic spatial dependence
+      \item Has computation on the order of Gaussian models for large space-time datasets
+    \end{itemize}
+  \end{itemize}
+\end{frame}
+
+\begin{frame}{Gaussian spatial model}
+  \begin{itemize} \setlength{\itemsep}{0.5em}
+    \item In geostatistics $Y(\bs)$ are often modeled using a Gaussian process with mean function $\mu(\bs)$ and covariance function $\rho(h)$.
+    \item Model properties:
+    \begin{itemize}
+      \item Nice computing properties (closed-form likelihood)
+      \item For a Gaussian spatial model $\lim_{c \rightarrow \infty} \chi(h) = 0$ regardless of the strength of the correlation in the bulk of the distribution
+      \item Tail is not flexible:
+      \begin{itemize}
+        \item light tailed
+        \item symmetric
+      \end{itemize}
     \end{itemize}
@@
       \item $\bX(\bs) \bbeta$ is the location
-      \item $\bOmega = \bomega \bar{\bOmega} \bomega$ is a correlation matrix
-      \item $\bomega = \text{diag} \left(\frac{1}{\sqrt{ab}}, \ldots, \frac{1}{ \sqrt{ab}}\right)$
+      \item $\bOmega = \frac{1}{ab} \bar{\bOmega}$ is a correlation matrix
       \item $\bar{\bOmega} = (\bSigma + \lambda^2 \bOne \bOne^T)$
@@
   \end{itemize}
+\end{frame}
+
+\begin{frame}{$\chi(h)$ plot}
+  \vspace{-2em}
+  \centering
+  \begin{figure}
+  \includegraphics[width=.65\linewidth]{./plots/pot/chi-h-t_and_gaus.pdf}\\[-0.2in]
+  \caption{$\chi$ plot for symmetric $t$, skew-$t$, and Gaussian}
+  \end{figure}
 \end{frame}
@@
 
-\begin{frame}{$\chi$ statistic}
-  \begin{itemize} \setlength{\itemsep}{0.5em}
-   \item The $\chi$ statistic is a measure of extremal dependence
-   \item Specifically, we focus on $\chi(\bh)$ for the upper tail given by
-    \begin{align*}
-      \chi(h) = \lim_{c \rightarrow \infty}\Pr(Y(\bs) > c \mid Y(\bt) > c)
-    \end{align*}
-    where $h = ||\bs - \bt||$
-    \item If $ \chi(h) = 0$, then observations are asymptotically independent at distance $\bh$.
-  \end{itemize}
-\end{frame}
-
-\begin{frame}{Gaussian spatial model}
-  \begin{itemize} \setlength{\itemsep}{0.5em}
-    \item In geostatistics $Y(\bs)$ are often modeled using a Gaussian process with mean function $\mu(\bs)$ and covariance function
-$\rho(\bh)$.
-    \item Model properties:
-    \begin{itemize}
-      \item Nice computing properties (closed-form likelihood)
-    \item For a Gaussian spatial model $\lim_{c \rightarrow \infty} \chi(\bh) = 0$ regardless of the strength of the correlation in the bulk of the distribution
-    \item Tail is not flexible (Gaussian is light tailed)
-    \end{itemize}
-    \end{itemize}
-\end{frame}
-
-\begin{frame}{Spatial skew-$t$ distribution}
+\begin{frame}{Spatial skew-$t$ distribution properties}
   \begin{itemize} \setlength{\itemsep}{0.5em}
@@
     \begin{itemize}
-    	\item Has flexible tail controlled by skewness $\alpha$ and degrees of freedom $2a$
-    	\item For a skew-$t$ distribution $\lim_{c \rightarrow \infty} \chi(\bh) > 0$ (Padoan, 2011)
+    	\item Has flexible tail:
+      \begin{itemize}
+        \item Skewness controlled by $\lambda$
+        \item Weight of tails controlled by degrees of freedom $2a$
+      \end{itemize}
+    	\item For a skew-$t$ distribution $\lim_{c \rightarrow \infty} \chi(h) > 0$ (Padoan, 2011)
     	\item Computation that is on the order of Gaussian computation
     \end{itemize}
-    \item For this distribution, $\chi(\bh)$ shows asymptotic dependence that does not approach 0 as $\bh \rightarrow \infty$
+    \item Challenge: $\chi(h) > 0$ as $h \rightarrow \infty$ (Padoan, 2011)
+    \begin{itemize}
       \item This occurs because all observations (near and far) share the same $z$ and $\sigma^2$
       \item We deal with this through a daily random partition (similar to Kim et al., 2005)
+    \end{itemize}
   \end{itemize}
@@
     \end{align*}
+  \end{itemize}
+\end{frame}
+
+\begin{frame}{Random partition}
+  \begin{itemize} \setlength{\itemsep}{0.5em}
     \item Within each partition $Y(\bs)$ has the same MV skew-t distribution as before
+    \item Across partitions $Y(\bs)$ are asymptotically independent, but still correlated through $v(\bs)$
   \end{itemize}
@@
 
-\begin{frame}{$\chi(\bh)$ plot}
+\begin{frame}{$\chi(h)$ plot}
   \vspace{-2em}
@@
 
-\begin{frame}{Sample simulated datasets}
-  \centering
-  \begin{figure}
-  \includegraphics[width=1\linewidth]{./plots/pot/gauss-vs-t3.pdf}
-  \caption{Gaussian and $t$ with 3 partitions}
-  \end{figure}
+\begin{frame}{Random partition skew-$t$ model}
+  \begin{itemize} \setlength{\itemsep}{0.5em}
+    \item This new model is called a random partition skew-$t$ model, and it has the properties we desire
+    \begin{itemize}
+      \item Marginal distribution with flexible tails
+      \begin{itemize}
+        \item $\lambda$ term allows for asymmetry
+        \item d.f. controls heavy vs light tails
+      \end{itemize}
+      \item Asymptotic spatial dependence for that decays with distance between sites via partition
+      \item Computation is on the order of Gaussian models for large space-time datasets
+    \end{itemize}
+  \end{itemize}
 \end{frame}
 
-\begin{frame}{Sample simulated datasets}
-  \centering
-  \begin{figure}
-  \includegraphics[width=1\linewidth]{./plots/pot/gauss-vs-skew-t3.pdf}
-  \caption{Gaussian and skew-$t$ with 3 partitions}
-  \end{figure}
-\end{frame}
-
-\begin{frame}{Extension to space-time data}
+\begin{frame}{Temporal dependence}
   \begin{itemize} \setlength{\itemsep}{0.5em}
-    \item We extend our previous model to be
+    \item Temporal dependence should be accounted for when using daily data
+    \item For multiple days of observations the model becomes
       \begin{align*}
@@
       where $t = 1, \ldots, T$ denotes the day of each observation.
-    \item We incorporate an AR(1) time series on $\bw^*_{tk} = (w^*_{tk1}, w^*_{tk2})$, $z_{tk}$, and $\sigma^*_{tk}$ where
-    \begin{align*}
-      w^*_{tki} &= \Phi^{-1}\left[ \frac{w_{tki} - \min(\bs_i)}{ \text{range}(\bs_i)}\right] \quad i = 1, 2 \\
-      \sigma^{2*}_t(\bs) &= \Phi^{-1}\{ \text{IG}[\sigma^2_t(\bs)] \}
-    \end{align*}
-    are transformations to $\calR^2$
+    \item Different ways to incorporate the temporal dependence
+    \begin{itemize}
+      \item Time series on $\bw_t$, $z_t(\bs)$, and $\sigma_t(\bs)$
+      \item Three dimensional covariance model for $v_t(\bs)$ (e.g. Huser and Davison, 2014)
+    \end{itemize}
+    \item Our model incorporates the time series because the $z_t(\bs)$ and $\sigma_t(\bs)$ terms dictate the tail behavior.
   \end{itemize}
 \end{frame}
+
+% \begin{frame}{Sample simulated datasets}
+%   \centering
+%   \begin{figure}
+%   \includegraphics[width=1\linewidth]{./plots/pot/gauss-vs-t3.pdf}
+%   \caption{Gaussian and $t$ with 3 partitions}
+%   \end{figure}
+% \end{frame}
+
+% \begin{frame}{Sample simulated datasets}
+%   \centering
+%   \begin{figure}
+%   \includegraphics[width=1\linewidth]{./plots/pot/gauss-vs-skew-t3.pdf}
+%   \caption{Gaussian and skew-$t$ with 3 partitions}
+%   \end{figure}
+% \end{frame}
 
@@
     \begin{itemize}
-      \item Gaussian vs $t$ vs skew-$t$ marginal distribution
-        \item $K=1$ partition vs $K=5$ partitions
+      \item Gaussian, $K = 1$ partition
+      \item $T$, $K = 1$ partition
+      \item $T$, $K = 5$ partitions
+      \item Skew-$t$, $K = 1$ partition
+      \item Skew-$t$, $K = 5$ partitions
+      \item Max-stable
+      \begin{itemize}
+        \item Marginally: GEV$(\mu=1, \sigma=1, \xi=0.2)$
+        \item Dependence function: asymmetric logistic with $\alpha = 0.5$
       \end{itemize}
+    \end{itemize}
+  \end{itemize}
+\end{frame}
+
+\begin{frame}{Simulation study}
+  \begin{itemize} \setlength{\itemsep}{0.5em}
     \item 5 different models:
     \begin{itemize}
-      \item Gaussian vs skew-$t$ marginal distribution
-      \item $K = 1$ partition vs $K = 5$ partitions
+      \item Gaussian
+      \item Skew-$t$ with $K = 1$ partition, no thresholding
+      \item Skew-$t$ with $K = 1$ partition, thresholding at $q(0.80)$
+      \item Skew-$t$ with $K = 5$ partitions, no thresholding
+      \item Skew-$t$ with $K = 5$ partitions, thresholding at $q(0.80)$
     \end{itemize}
-    \item Brier score used to determine model that gives best fit
   \end{itemize}
@@
   \begin{itemize} \setlength{\itemsep}{0.5em}
+    \item Brier score used to determine model that gives best fit (Gneiting and Raftery, 2007)
     \item The Brier score for predicting exceedance of threshold $c$ is
@@
    \end{figure}
-
 \end{frame}
@@
       \item $K=1, 5, 6, 7, 8, 9, 10, 15$ partitions
-      \item Thresholding at $T=0, 50, 75$, and $85$ ppb
+      \item Thresholding at
+      \begin{itemize}
+         \item $T = 0$
+         \item $T = 50$ppb, $q(0.48)$
+         \item $T = 75$ppb, $q(0.92)$
+         \item $T = 85$ppb, $q(0.97)$
       \end{itemize}
+    \end{itemize}
+    \item Compare Brier scores using two-fold cross validation
+  \end{itemize}
+\end{frame}
+
+\begin{frame}{Model comparisons}
+  \begin{itemize} \setlength{\itemsep}{0.5em}
     \item All methods use a \Matern or exponential covariance ($\nu = 0.5$)
-    \item Compare quantile and Brier scores using two-fold cross validation (Gneiting and Raftery, 2007)
+    \item Covariate data from the Environmental Protection Agency's Community Multiscale Air Quality (CMAQ) system.
     \item Mean function modeled as
     \begin{align*}
-    	\beta_0 + \beta_1 \cdot \text{CMAQ}
+    	\bX_t(\bs) \bbeta = \beta_0 + \beta_1 \cdot \text{CMAQ}_t (\bs)
     \end{align*}
@@
     \end{itemize}
-    \item Different ways to incorporate the temporal dependence
-    \begin{itemize}
-    	\item Three dimensional covariance model for $v_t(\bs)$ (e.g. Huser and Davison, 2014)
-    \end{itemize}
     \item Different partition structure
@@
 
-\begin{frame}{Spatial binary regression}
+\begin{frame}{Future work: temporal dependence}
   \begin{itemize} \setlength{\itemsep}{0.5em}
-    \item We observe $Y_i = I[Z(\bs_i) > T]$, an indicator variable that a continuous latent variable has exceeded a pre-specified threshold $T$.
+    \item We extend our previous model to be
+    \begin{align*}
+      Y_t(\bs) = \bX_t(\bs)^T \bbeta + \lambda \sigma_t(\bs) | z_t(\bs) | + \sigma_t(\bs) v_t(\bs)
+    \end{align*}
+    where $t = 1, \ldots, T$ denotes the day of each observation.
+    \item We incorporate an AR(1) time series on $\bw^*_{tk} = (w^*_{tk1}, w^*_{tk2})$, $z_{tk}$, and $\sigma^*_{tk}$ where
+    \begin{align*}
+      w^*_{tki} &= \Phi^{-1}\left[ \frac{w_{tki} - \min(\bs_i)}{ \text{range}(\bs_i)}\right] \quad i = 1, 2 \\
+      \sigma^{2*}_t(\bs) &= \Phi^{-1}\{ \text{IG}[\sigma^2_t(\bs)] \}
+    \end{align*}
+    are transformations to $\calR^2$
+  \end{itemize}
+\end{frame}
+
+\begin{frame}{Rare spatial binary regression}
+  \begin{itemize} \setlength{\itemsep}{0.5em}
+    \item Movivation:
+    \begin{itemize}
+      \item Want to incorporate spatial dependence when modeling rare events.
+      \item Examples:
+      \begin{itemize}
+        \item Diseased trees
+        \item Disease outbreak
+      \end{itemize}
+    \end{itemize}
+    \item We observe
+    \begin{align*}
+      Y_i = \left\{ \begin{array}{ll}
+        1 \quad & \text{event occurred}\\
+        0 \quad & \text{no event occurred}
+      \end{array} \right.
+    \end{align*}
     \item We model $\Pr[Y_i = 1]$
-    \item Common examples:
+  \end{itemize}
+\end{frame}
+
+\begin{frame}{Rare spatial binary regression}
+  \begin{itemize} \setlength{\itemsep}{0.5em}
+    \item Common examples with non-spatial analysis
     \begin{itemize}
@@
       \begin{align*}
-        \Pr[Y_i = 1] = \frac{ \exp(\bX(\bs_i) \bbeta) }{ 1 + \exp(\bX(\bs_i) \bbeta)}
+        \Pr[Y_i = 1] = \frac{ \exp(\bX_i \bbeta) }{ 1 + \exp(\bX_i \bbeta)}
       \end{align*}
@@
       \begin{align*}
-        \Pr[Y_i = 1] = \Phi[\bX(\bs_i) \bbeta]
+        \Pr[Y_i = 1] = \Phi[\bX_i \bbeta]
       \end{align*}
       where $\Phi$ is the standard normal distribution function
+      \item Wang and Dey (2010): Generalized extreme value link function
+      \begin{align*}
+        \Pr[Y_i = 1] = 1 - \exp\left[-(1 + \xi \bX_i \bbeta)^{-1 / \xi} \right]
+      \end{align*}
     \end{itemize}
@@
 
-\begin{frame}{Spatial binary regression}
-  \begin{itemize} \setlength{\itemsep}{0.5em}
-    \item Logistic and probit regression amount to modeling
-    \begin{align*}
-      \Pr[Y_i = 1] = g[\bX(\bs_i) \bbeta]
-    \end{align*}
-    where $g[\cdot]$ is a link function transforming from $\calR$ to $(0, 1)$.
-    \item Wang and Dey (2010): Generalized extreme value link function
-    \begin{align*}
-      g[\bx(\bs_i) \bbeta] = 1 - \exp\left[-(1 + \xi \bX_i \bbeta)^{-1 / \xi} \right]
-    \end{align*}
-  \end{itemize}
-\end{frame}
-
-\begin{frame}{Spatial binary regression}
+\begin{frame}{Rare spatial binary regression}
   \begin{itemize} \setlength{\itemsep}{0.5em}
@@
       \item use the GEV link function
-      \item use the hierarchical likelihood from Reich and Shaby (2012)
+      \item use the hierarchical method for spatially dependent extremes from Reich and Shaby (2012)
     \end{itemize}
@@
 
-\begin{frame}{Spatial binary regression}
+\begin{frame}{Rare spatial binary regression}
   \begin{itemize} \setlength{\itemsep}{0.5em}
@@
     \end{align*}
-    we set the latent variable $z_i = -\frac{1}{\log(1 - p_i)}$
-    \item Then we evaluate the joint likelihood using $z_i$.
+    \item Latent variable $z_i = -\frac{1}{\log(1 - p_i)}$ used to evaluate the likelihood.
   \end{itemize}
@@
   \begin{itemize} \setlength{\itemsep}{0.5em}
-    \item We use a multivariate generalized extreme value distribution with asymetric Laplace deendence function given by
+    \item We use a multivariate generalized extreme value distribution with asymmetric logistic dependence function (Reich and Shaby, 2012)
     \footnotesize{
@@
     \begin{align*}
-      w_l(\bs_i) = \frac{ \exp\left[ -0.5 \left( \frac{ || \bs - \bv_l || }{ \rho} \right)^2 \right]}{ \sum_{l = 1}^L \exp\left[ -0.5 \left( \frac{ || \bs - \bv_l || }{ \rho} \right)^2 \right] }
+      w_l(\bs_i) = \frac{ \exp\left[ -0.5 \left( \frac{ || \bs_i - \bv_l || }{ \rho} \right)^2 \right]}{ \sum_{l = 1}^L \exp\left[ -0.5 \left( \frac{ || \bs_i - \bv_l || }{ \rho} \right)^2 \right] }
     \end{align*}
@@
 
+\begin{frame}{Joint likelihood}
+  \begin{itemize} \setlength{\itemsep}{0.5em}
+    \item For small $K$, we can evaluate the likelihood directly.
+    \item For large $K$, we use the hierarchical model of Reich and Shaby (2012).
+  \end{itemize}
+\end{frame}
+
 % TODO: more about this model for sure
@@
     \item Knot-specific bandwidth
-  \end{itemize}
-\end{frame}
-
-\begin{frame}{Joint likelihood}
-  \begin{itemize} \setlength{\itemsep}{0.5em}
-    \item For small $K$, we can evaluate the likelihood directly.
-    \item For large $K$, we use the hierarchical model of Reich and Shaby (2012).
   \end{itemize}

### 2015-01-07 prelim/oral_prelim.tex
Oral prelim updates

@@
 \documentclass{beamer}
-\usepackage{beamerthemesplit}
+% \usepackage{beamerthemesplit}
+\usetheme{Madrid}
 % \usepackage{pstricks}
@@
 \usepackage{lineno, hyperref}
+\usepackage{stackrel}
 
@@
 
-\title[Spatial methods for extreme value analysis] % (optional, use only with long paper titles)
+\title[Spatial methods for EVA] % (optional, use only with long paper titles)
 {
@@
 }
-\author[Samuel A. Morris]{Samuel A. Morris}
-\institute[NCSU]{}
-\date{January 9, 2015}
+\author[S. Morris]{Samuel A. Morris}
+\institute[]{North Carolina State University}
+\date[]{January 9, 2015}
 
@@
   \begin{figure}
-    \includegraphics[width=\linewidth, trim=0 1in 0 0in ]{./plots/pot/ozone-10jul-us.pdf}
+    \includegraphics[width=\linewidth, trim=0 1in 0 1in ]{./plots/pot/ozone-10jul-us.pdf}
     \caption{Max 8-hour ozone measurements on July 10, 2005}
    \end{figure}
+\end{frame}
+
+\begin{frame}{Motivation}
+  \begin{itemize} \setlength{\itemsep}{0.5em}
+    \item Ozone compliance for Clean Air Act (EPA)
+    \begin{itemize}
+      \item Annual fourth-highest daily maximum 8-hour concentration, averaged over 3 years, not to exceed 75 ppb
+      \item Annual fourth-highest is the 99th percentile for the year
+    \end{itemize}
+  \end{itemize}
 \end{frame}
@@
   \begin{figure}
-    \includegraphics[width=1\linewidth]{./plots/define_extreme.pdf}
-    \caption{Monthly maximums recorded over a three years.}
+    \includegraphics[width=1\linewidth, trim=0 0.5in 0 1in]{./plots/define_extreme.pdf}
+    \caption{Monthly maximums recorded over a three years}
     \end{figure}
@@
 
-\begin{frame}{Standard analysis - Block maxima}
-  \begin{itemize} \setlength{\itemsep}{0.5em}
-    \item Asymptotic result:
+\begin{frame}{Standard non-spatial analysis: Block maxima}
+  \begin{itemize} \setlength{\itemsep}{0.5em}
+    \item Asymptotic result (Falk et al., 2011)
     \begin{itemize}
@@
 
-\begin{frame}{Standard analysis - Block maxima}
+\begin{frame}{Standard non-spatial analysis: Block maxima}
   \begin{itemize} \setlength{\itemsep}{0.5em}
@@
 
-\begin{frame}{Standard analysis - Peaks over threshold}
+\begin{frame}{Standard non-spatial analysis: Peaks over threshold}
+  \begin{itemize} \setlength{\itemsep}{0.5em}
+    \item Asymptotic result (Falk et al., 2011)
+    \begin{itemize}
+      \item Let $X_1, \ldots, X_n \sim F$
+      \item If there exist normalizing sequences $a_n > 0$ and $b_n \in \calR$ with
+      \begin{align*}
+        \lim_{n \rightarrow \infty} 1 - F(b_n) = 0 \qquad \text{and} \qquad \lim_{ n \rightarrow \infty} \frac{ 1 - F(b_{n+1})}{1 - F(b_n)} = 1
+      \end{align*}
+      where $b_n$ is a sequence of thresholds
+      \item If for any $x \ge 0$,
+      \begin{align*}
+        1 - \frac{ 1 - F(a_n x + b_n) }{ 1 - F(b_n) } \converged H(x),
+      \end{align*}
+      then $H(x)$ follows a generalized Pareto distribution
+    \end{itemize}
+  \end{itemize}
+\end{frame}
+
+\begin{frame}{Standard non-spatial analysis: Peaks over threshold}
   \begin{itemize} \setlength{\itemsep}{0.5em}
@@
         \item Unbounded above if $\xi \ge 0$
-        \item Bouded above by $(\mu - \sigma) / \xi$ when $\xi < 0$
+        \item Bounded above by $(\mu - \sigma) / \xi$ when $\xi < 0$
       \end{itemize}
@@
 
-\begin{frame}{Standard analysis - Peaks over threshold}
+\begin{frame}{Standard non-spatial analysis: Peaks over threshold}
   \begin{itemize}  \setlength{\itemsep}{0.5em}
@@
   \begin{itemize} \setlength{\itemsep}{0.5em}
-    \item For a spatial analysis, max-stable processes give an appropriate limiting distribution (Cooley et al., 2012):
-    \begin{itemize}
-      \item Consider a spatial process $x_t(\bs)$, $t = 1, \ldots, T$.
+    \item For a spatial analysis, max-stable processes give an appropriate limiting distribution (Cooley et al., 2012)
+    \begin{itemize}
+      \item Consider a spatial process $x_t(\bs)$, $t = 1, \ldots, T$
       \item Let $M_T(\bs) = \left\{ \bigvee_{t=1}^T x_t(\bs_1), \ldots, \bigvee_{t=1}^T x_t(\bs_n) \right\}$
@@
       \end{align*}
-      which has a non-degenerate distribution, then $Y(\bs)$ is a max-stable process.
+      which has a non-degenerate distribution, then $Y(\bs)$ is a max-stable process
     \end{itemize}
@@
     \item Marginally at each site, observations follow a generalized extreme value distribution
-    \item Finite collection of sites:
+    \item For a finite collection of sites
     \begin{itemize}
@@
   \begin{itemize} \setlength{\itemsep}{0.5em}
-   \item The $\chi$ statistic is a measure of extremal dependence in the tails
+    \item Correlation is the most common measure of dependence, but it is irrelevant for extreme value analysis
+    \item Extreme value analysis focuses on the $\chi$ statistic, a measure of extremal dependence
     \item Specifically, we focus on $\chi(h)$ for the upper tail given by
@@
     \begin{itemize}
-      \item computationally challenging (Falk et al., 2011)
+      \item Computationally challenging (Falk et al., 2011)
       \begin{itemize}
@@
       \end{itemize}
-      \item joint distribution only available in low dimension
-    \end{itemize}
-    \item Some recent approaches:
+      \item Joint distribution only available in low dimension
+    \end{itemize}
+    \item Some recent approaches
     \begin{itemize}
@@
     \item A spatio-temporal model with flexible tails, asymptotic spatial dependence, and computation on the order of Gaussian models for large space-time datasets
-    \item Predicting exceedances using a spatially dependent generalized extreme-value link function
-    \item A Bayesian hierarchical model to allow for non-stationary covariance in extreme value models.
+    \item Predicting rare binary events with a spatially dependent generalized extreme-value link function
+    \item A Bayesian hierarchical model to allow for non-stationary covariance in extreme value models
   \end{enumerate}
@@
     \item Model to analysis spatiotemporal extreme values
-    \item Model objectives:
+    \item Model objectives
     \begin{itemize}
@@
       \begin{itemize}
-        \item allow for asymmetric distributions
-        \item allow for heavy tails
+        \item Allow for asymmetric distributions
+        \item Allow for heavy tails
       \end{itemize}
@@
     \item In geostatistics $Y(\bs)$ are often modeled using a Gaussian process with mean function $\mu(\bs)$ and covariance function $\rho(h)$.
-    \item Model properties:
+    \item Model properties
     \begin{itemize}
@@
       \item For a Gaussian spatial model $\lim_{c \rightarrow \infty} \chi(h) = 0$ regardless of the strength of the correlation in the bulk of the distribution
-      \item Tail is not flexible:
-      \begin{itemize}
-        \item light tailed
-        \item symmetric
+      \item Tail is not flexible
+      \begin{itemize}
+        \item Light-tailed
+        \item Symmetric
       \end{itemize}
@@
       \item $\bar{\bOmega} = (\bSigma + \lambda^2 \bOne \bOne^T)$
-      \item $\Sigma$ is a postive definite correlation matrix
+      \item $\bSigma$ is a postive definite correlation matrix
       \item $\alpha = \lambda( 1 + \lambda^2 \bOne^T \bSigma^{-1} \bOne)^{-1/2} \bOne^T \bSigma^{-1}$ controls the skewness
@@
   \includegraphics[width=.65\linewidth]{./plots/pot/chi-h-t_and_gaus.pdf}\\[-0.2in]
-  \caption{$\chi$ plot for symmetric $t$, skew-$t$, and Gaussian}
-  \end{figure}
+  \caption{$\chi$ plot for skew-$t$, and Gaussian}
+  \end{figure}
+\end{frame}
+
+\begin{frame}{Extension of the skew-$t$ distribution}
+  \begin{itemize} \setlength{\itemsep}{0.5em}
+    \item Skew-$t$ distribution addresses two modeling concerns
+    \begin{itemize}
+      \item Extremal dependence
+      \item Reasonable computing
+    \end{itemize}
+    \item Our contribution is to extend the skew-$t$
+    \begin{itemize}
+      \item Censoring to focus on extreme observations
+      \item Partitioning to address long-range dependence
+    \end{itemize}
+  \end{itemize}
 \end{frame}
@@
   \begin{itemize} \setlength{\itemsep}{0.5em}
-    \item We censor the observed data at a high threshold $T$.
-    \item Censored data:
+    \item We censor the observed data at a high threshold $T$
+    \item Censored data
     \begin{align*}
@@
     where $\delta(\bs) = I[Y(\bs) > T]$
-    \item Allows tails of the distribution to speak for themselves.
+    \item Allows tails of the distribution to speak for themselves
   \end{itemize}
@@
     \begin{itemize}
-    	\item Has flexible tail:
+    	\item Has flexible tail
       \begin{itemize}
@@
       \item This occurs because all observations (near and far) share the same $z$ and $\sigma^2$
-      \item We deal with this through a daily random partition (similar to Kim et al., 2005)
+      \item We deal with this through a daily random partition (similar to Kim et al., 2005 for non-extreme modeling)
     \end{itemize}
@@
     \begin{figure}
-    \includegraphics[width=0.54\linewidth]{./plots/pot/example-partition-1.pdf}
-    \includegraphics[width=0.54\linewidth]{./plots/pot/example-partition-2.pdf}
+    \includegraphics[width=0.54\linewidth, trim=0 0 0 1in]{./plots/pot/example-partition-1.pdf}
+    \includegraphics[width=0.54\linewidth, trim=0 0 0 1in]{./plots/pot/example-partition-2.pdf}
     \caption{Two sample partitions (number is at partition center)}
@@
   \begin{figure}
-  \includegraphics[width=.7\linewidth]{./plots/pot/chi-h.pdf}\\[-0.4in]
+  \includegraphics[width=.65\linewidth]{./plots/pot/chi-h.pdf}\\[-0.2in]
   \caption{$\chi$ plot for different data settings}
@@
         \item $\lambda$ term allows for asymmetry
-        \item d.f. controls heavy vs light tails
-      \end{itemize}
-      \item Asymptotic spatial dependence for that decays with distance between sites via partition
+        \item Degrees of freedom control heavy vs light tails
+      \end{itemize}
+      \item Asymptotic spatial dependence for that decays with distance between sites through partitioning
       \item Computation is on the order of Gaussian models for large space-time datasets
     \end{itemize}
-  \end{itemize}
-\end{frame}
-
-\begin{frame}{Temporal dependence}
-  \begin{itemize} \setlength{\itemsep}{0.5em}
-    \item Temporal dependence should be accounted for when using daily data
-    \item For multiple days of observations the model becomes
-      \begin{align*}
-        Y_t(\bs) = \bX_t(\bs)^T \bbeta + \lambda \sigma_t(\bs) | z_t(\bs) | + \sigma_t(\bs) v_t(\bs)
-      \end{align*}
-      where $t = 1, \ldots, T$ denotes the day of each observation.
-    \item Different ways to incorporate the temporal dependence
-    \begin{itemize}
-      \item Time series on $\bw_t$, $z_t(\bs)$, and $\sigma_t(\bs)$
-      \item Three dimensional covariance model for $v_t(\bs)$ (e.g. Huser and Davison, 2014)
-    \end{itemize}
-    \item Our model incorporates the time series because the $z_t(\bs)$ and $\sigma_t(\bs)$ terms dictate the tail behavior.
   \end{itemize}
@@
   \begin{itemize} \setlength{\itemsep}{0.5em}
-    \item Three main steps:
+    \item Three main steps
     \begin{enumerate}[1.]
@@
       \end{itemize}
+    \end{itemize}
+  \end{itemize}
+\end{frame}
+
+\begin{frame}{Simulation study}
+  \begin{itemize} \setlength{\itemsep}{0.5em}
+    \item 50 dataset for each setting
+    \begin{itemize}
+      \item 144 sites in $[0, 10] \times [0, 10]$
+      \begin{itemize}
+        \item 100 training
+        \item 44 testing
+      \end{itemize}
+    \end{itemize}
+    \item Model parameters
+    \begin{itemize}
+      \item Spatial range: $\rho = 1$
+      \item Skew parameter: $\lambda = 3$
+      \item Degrees of freedom: 6 for $t$ distributions
     \end{itemize}
@@
   \begin{itemize} \setlength{\itemsep}{0.5em}
-    \item Key findings:
+    \item Key findings
     \begin{itemize}
       \item Improvement over Gaussian methods when partitioning
-      \item Underestimating the number of knots has a detrimental impact
+      \item Specifying too few knots has a detrimental impact
       \item In all cases, non-thresholded models perform better than thresholded models
@@
 	\end{itemize}
-  \item We take a stratified sample of $n = 800$ sites:
+  \item We take a stratified sample of $n = 800$ sites
   \begin{itemize}
@@
   \begin{figure}
-    \includegraphics[width=\linewidth, trim=0 1in 0 0in ]{./plots/pot/ozone-10jul-us.pdf}
+    \includegraphics[width=\linewidth, trim=0 1in 0 0.7in ]{./plots/pot/ozone-10jul-us.pdf}
     \caption{Max 8-hour ozone measurements on July 10, 2005}
@@
   \begin{figure}
-    \includegraphics[width=1\linewidth]{./plots/pot/chi-plot-ozone-res.pdf}
+    \includegraphics[width=1\linewidth, trim=0 0 0 1in]{./plots/pot/chi-plot-ozone-res.pdf}
     \caption{$\widehat{\chi}$-plot for sample quantiles of ozone observations}
@@
       \item $K=1, 5, 6, 7, 8, 9, 10, 15$ partitions
-      \item Thresholding at
+      \item 4 threshold levels
       \begin{itemize}
@@
   \begin{figure}
-    \includegraphics[width=0.45\linewidth]{./plots/pot/bs-ozone-1.pdf}
-    \includegraphics[width=0.45\linewidth]{./plots/pot/bs-ozone-2.pdf} \\
+    \includegraphics[width=0.47\linewidth]{./plots/pot/bs-ozone-1.pdf}
+    \includegraphics[width=0.47\linewidth]{./plots/pot/bs-ozone-2.pdf} \\
     \caption{Relative Brier score results}
@@
   \begin{itemize} \setlength{\itemsep}{0.5em}
-    \item Key findings:
-    \begin{itemize}
-      \item Partitioning improves performance across all high thresholds.
+    \item Key findings
+    \begin{itemize}
+      \item Partitioning improves performance across all high thresholds
       \item Models with anywhere from $K = 5$ to $K = 10$ partitions perform similarly
@@
 
-\begin{frame}{Discussion and future work}
+\begin{frame}{Discussion}
   \begin{itemize} \setlength{\itemsep}{0.5em}
@@
     \end{itemize}
+  \end{itemize}
+\end{frame}
+
+\begin{frame}{Future work: Temporal dependence}
+  \begin{itemize} \setlength{\itemsep}{0.5em}
+    \item Temporal dependence should be accounted for when using daily data
+    \item For multiple days of observations the model becomes
+      \begin{align*}
+        Y_t(\bs) = \bX_t(\bs)^T \bbeta + \lambda \sigma_t(\bs) | z_t(\bs) | + \sigma_t(\bs) v_t(\bs)
+      \end{align*}
+      where $t = 1, \ldots, T$ denotes the day of each observation
+    \item Different ways to incorporate the temporal dependence
+    \begin{itemize}
+      \item Time series on $\bw_t$, $z_t(\bs)$, and $\sigma_t(\bs)$
+      \item Three dimensional covariance model for $v_t(\bs)$ (e.g. Huser and Davison, 2014)
+    \end{itemize}
+  \end{itemize}
+\end{frame}
+
+\begin{frame}{Future work: Temporal dependence}
+  \begin{itemize} \setlength{\itemsep}{0.5em}
+    \item We choose the time series approach because the $z_t(\bs)$ and $\sigma_t(\bs)$ terms dictate the tail behavior
+    \item We incorporate an AR(1) time series on $\bw^*_{tk} = (w^*_{tk1}, w^*_{tk2})$, $z_{tk}$, and $\sigma^*_{tk}$ where
+    \begin{align*}
+      w^*_{tki} &= \Phi^{-1}\left[ \frac{w_{tki} - \min(\bs_i)}{ \text{range}(\bs_i)}\right] \quad i = 1, 2 \\[0.5em]
+      \sigma^{2*}_t(\bs) &= \Phi^{-1}\{ \text{IG}[\sigma^2_t(\bs)] \}
+    \end{align*}
+    are transformations to $\calR^2$
+  \end{itemize}
+\end{frame}
+
+\begin{frame}{Future work: Knots and their impact}
+  \begin{itemize} \setlength{\itemsep}{0.5em}
     \item Different partition structure
@@
 
-\begin{frame}{Future work: temporal dependence}
-  \begin{itemize} \setlength{\itemsep}{0.5em}
-    \item We extend our previous model to be
-    \begin{align*}
-      Y_t(\bs) = \bX_t(\bs)^T \bbeta + \lambda \sigma_t(\bs) | z_t(\bs) | + \sigma_t(\bs) v_t(\bs)
-    \end{align*}
-    where $t = 1, \ldots, T$ denotes the day of each observation.
-    \item We incorporate an AR(1) time series on $\bw^*_{tk} = (w^*_{tk1}, w^*_{tk2})$, $z_{tk}$, and $\sigma^*_{tk}$ where
-    \begin{align*}
-      w^*_{tki} &= \Phi^{-1}\left[ \frac{w_{tki} - \min(\bs_i)}{ \text{range}(\bs_i)}\right] \quad i = 1, 2 \\
-      \sigma^{2*}_t(\bs) &= \Phi^{-1}\{ \text{IG}[\sigma^2_t(\bs)] \}
-    \end{align*}
-    are transformations to $\calR^2$
-  \end{itemize}
-\end{frame}
-
 \begin{frame}{Rare spatial binary regression}
   \begin{itemize} \setlength{\itemsep}{0.5em}
-    \item Movivation:
-    \begin{itemize}
-      \item Want to incorporate spatial dependence when modeling rare events.
-      \item Examples:
+    \item Motivation
+    \begin{itemize}
+      \item Want to incorporate spatial dependence when modeling rare events
+      \item Examples
       \begin{itemize}
@@
       \begin{align*}
-        \Pr[Y_i = 1] = \Phi[\bX_i \bbeta]
+        \Pr[Y_i = 1] = \Phi(\bX_i \bbeta)
       \end{align*}
       where $\Phi$ is the standard normal distribution function
-      \item Wang and Dey (2010): Generalized extreme value link function
+      \item Cloglog regression
+      \begin{align*}
+        \Pr[Y_i = 1] = \exp[ -\exp(-\bX_i \bbeta)]
+      \end{align*}
+    \end{itemize}
+  \end{itemize}
+\end{frame}
+
+\begin{frame}{Rare spatial binary regression}
+  \begin{itemize} \setlength{\itemsep}{0.5em}
+    \item Generalized extreme value link function (Wang and Dey, 2010)
     \begin{align*}
@@
     \end{align*}
+    \item Link function allows for greater positive skew than existing methods
+    \begin{itemize}
+      \item When $\xi = 0$, the link is the Cloglog link
+      \item When $\xi > 0$, the link allows for greater positive skew than Cloglog link
     \end{itemize}
@@
     \begin{itemize}
-      \item use the GEV link function
-      \item use the hierarchical method for spatially dependent extremes from Reich and Shaby (2012)
+      \item Use the GEV link function
+      \item Use the hierarchical method for spatially dependent extremes from Reich and Shaby (2012)
     \end{itemize}
@@
   \begin{itemize} \setlength{\itemsep}{0.5em}
-    \item We fit parameters $\xi$ and $\beta$ in order to transform the data to GEV(1, 1, 1) marginal distributions.
+    \item We fit parameters $\xi$ and $\beta$ in order to transform the data to GEV(1, 1, 1) marginal distributions
     \item Using the link function
@@
     \end{align*}
-    \item Latent variable $z_i = -\frac{1}{\log(1 - p_i)}$ used to evaluate the likelihood.
+    \item We model $Y_i = I(Z_i > 0)$ where
+    \begin{itemize}
+      \item $z_i\sim$ Multivariate GEV (Reich and Shaby, 2012) is a latent variable
+    \end{itemize}
   \end{itemize}
@@
   \begin{itemize} \setlength{\itemsep}{0.5em}
-    \item We use a multivariate generalized extreme value distribution with asymmetric logistic dependence function (Reich and Shaby, 2012)
+    \item Asymmetric logistic dependence function (Reich and Shaby, 2012)
     \footnotesize{
@@
     \begin{itemize}
-      \item $w_l$ is a weighting function subject to the constraint that $\sum_{l = 1}^L w_l = 1$.
+      \item $w_l$ is a weighting function subject to the constraint that $\sum_{l = 1}^L w_l = 1$
       \item $\alpha$ controls spatial dependence
@@
   \end{itemize}
+\end{frame}
+
+\begin{frame}{Illustrating asymmetric logistic dependence in one-dimension}
+  \centering
+  \begin{figure}
+    \includegraphics[width=\linewidth, trim=0 0.5in 0 0]{./plots/max-stable.pdf}
+    \caption{Knot intensity $\times$ weight, red dots give intensity of random effects (left) Impact of $\alpha$ (right)}
+   \end{figure}
 \end{frame}
@@
     }
-    where $G(\bz_{(1)} = \Pr(Z_2 < z_2, \ldots, Z_n < z_n)$.
+    where $G(\bz_{(1)} = \Pr(Z_2 < z_2, \ldots, Z_n < z_n)$
     \item $K > 2$ can be derived similarly
@@
   \begin{itemize} \setlength{\itemsep}{0.5em}
-    \item For small $K$, we can evaluate the likelihood directly.
-    \item For large $K$, we use the hierarchical model of Reich and Shaby (2012).
+    \item For small $K$, we can evaluate the likelihood directly
+    \item For large $K$, we use the hierarchical model of Reich and Shaby (2012)
+  \end{itemize}
+\end{frame}
+
+\begin{frame}{Simulation study}
+  \begin{itemize} \setlength{\itemsep}{0.5em}
+    \item Data generated using logistic, Cloglog, and GEV links
+    \begin{itemize}
+      \item Exploring how sparseness of observations impacts prediction of events
+    \end{itemize}
+    \item Models fit using multivariate GEV, and Gaussian distribution
   \end{itemize}
@@
   \begin{itemize} \setlength{\itemsep}{0.5em}
-    \item Knot-specific bandwidth
+    \item Stationary covariance functions are a function of distance between two sites.
+    \begin{align*}
+      \rho(Y(\bs), Y(\bt)) = \rho(h)
+    \end{align*}
+    where $h = ||\bs - \bt||$
+    \item In extremes, stationary extremal dependence means
+    \begin{align*}
+      \chi(h) = \Pr[Y(\bs) > c | Y(\bt) > c]
+    \end{align*}
+    \item Sometimes the angle between
+    \begin{itemize}
+      \item Model does not allow for non-stationary covariance functions
+    \end{itemize}
+  \end{itemize}
+\end{frame}
+
+\begin{frame}{Non-stationary covariance for extreme values}
+  \begin{itemize} \setlength{\itemsep}{0.5em}
+    \item Semiparamtric approach using spectral density ratios (de Carvalho and Davison, 2014)
+    \begin{itemize}
+      \item Vector of observations can be transformed to psuedo-polar coordinates
+      \item Pairwise analysis
+    \end{itemize}
+    \item New approach extending Reich and Shaby (2012)
+    \begin{itemize}
+      \item Current model uses a single bandwidth term $\rho$ for all knots
+      \item Proposed idea is to implement a knot-specific $\rho$ to induce non-stationarity
+    \end{itemize}
   \end{itemize}

### 2015-01-08 prelim/oral_prelim.tex
prelim updates

@@
     \end{itemize}
-    \item Spatial methods borrow information across space to estimate spatial correlation and make predictions by Kriging at unknown locations
-    \item Want to explore similar methods for extremes
+    \item Estimating the probability of rare events is challenging because these events are, by definition, rare
+    \item Spatial extremes is promising because it borrows informaation across space
+    \item Spatial extremes is also useful for estimating probability of extremes at sites without data
   \end{itemize}
@@
 \begin{frame}{Motivation}
-  \begin{itemize} \setlength{\itemsep}{0.5em}
+  \begin{itemize} \setlength{\itemsep}{1em}
     \item Ozone compliance for Clean Air Act (EPA)
@@
       \item Annual fourth-highest is the 99th percentile for the year
+      \item Common objectives are
+      \begin{itemize}
+        \item To interpolate to unmonitored sites
+        \item Detect changes in extremes over time
+        \item Study factors that lead to extreme events
+      \end{itemize}
     \end{itemize}
@@
 \column{.45 \linewidth}
-  \begin{itemize} \setlength{\itemsep}{0.5em}
+  \begin{itemize} \setlength{\itemsep}{1em}
     \item Key in extreme value analysis is to define extremes
@@
     \begin{itemize}
-      \item Block maxima
-      \begin{itemize}
-        \item Red dots
-      \end{itemize}
-      \item Over threshold
-      \begin{itemize}
+      \item Block maxima (red dots)
       \item Values over threshold considered extreme
@@
   \end{itemize}
-  \end{itemize}
 
@@
     \includegraphics[width=1\linewidth, trim=0 0.5in 0 1in]{./plots/define_extreme.pdf}
-    \caption{Monthly maximums recorded over a three years}
+    \caption{Hypothetical monthly maximums recorded over a three years}
     \end{figure}
@@
 \begin{frame}{Standard non-spatial analysis: Block maxima}
-  \begin{itemize} \setlength{\itemsep}{0.5em}
-    \item Asymptotic result (Falk et al., 2011)
-    \begin{itemize}
+  \begin{itemize} \setlength{\itemsep}{1em}
     \item Let $X_1, \ldots, X_n$ be i.i.d.
-      \item Consider $M_n = \max(X_1, \ldots, X_n)$
+    \item Consider the block maximum $M_n = \max(X_1, \ldots, X_n)$
     \item If there exist normalizing sequences $a_n > 0$ and $b_n \in \calR$ such that
     \begin{align*}
-        a_n^{-1} (M_n - b_n) \converged G(z)
-      \end{align*}
-      then $G(z)$ follows a generalized extreme value distribution
-    \end{itemize}
+      \frac{ M_n - b_n }{ a_n } \converged G(z)
+    \end{align*}
+    then $G(z)$ follows a generalized extreme value distribution (GEV) (Falk et al., 2011)
+    \item This motivates the use of the GEV for block max data
   \end{itemize}
@@
 \begin{frame}{Standard non-spatial analysis: Block maxima}
-  \begin{itemize} \setlength{\itemsep}{0.5em}
-    \item Generalized extreme value distribution has three parameters:
-    \begin{itemize}
-      \item $\mu \in \calR$ is a location parameter
-      \item $\sigma > 0$ is a scale parameter
-      \item $\xi \in \calR$ is a shape parameter
-      \begin{itemize}
-        \item Unbounded above if $\xi \ge 0$
-        \item Bouded above by $(\mu - \sigma) / \xi$ when $\xi < 0$
-      \end{itemize}
-    \end{itemize}
-    \item Challenge:
-    \begin{itemize}
-      \item Lose information by only considering maximum in a block
-    \end{itemize}
-  \end{itemize}
-\end{frame}
-
-\begin{frame}{Standard non-spatial analysis: Block maxima}
-  \begin{itemize} \setlength{\itemsep}{0.5em}
-    \item Generalized extreme value distribution
+  \begin{itemize} \setlength{\itemsep}{1em}
+    \item GEV distribution
     \begin{align*}
@@
     \end{align*}
-    \item Standardized distribution is unit Fr\'{e}chet or GEV(1, 1, 1)
-    \begin{align*}
-      \Pr(Z < z) = exp(-z^{-1})
-    \end{align*}
-  \end{itemize}
-\end{frame}
-
-\begin{frame}{Standard non-spatial analysis: Peaks over threshold}
-  \begin{itemize} \setlength{\itemsep}{0.5em}
-    \item Asymptotic result (Falk et al., 2011)
-    \begin{itemize}
-      \item Let $X_1, \ldots, X_n \sim F$
-      \item If there exist normalizing sequences $a_n > 0$ and $b_n \in \calR$ with
-      \begin{align*}
-        \lim_{n \rightarrow \infty} 1 - F(b_n) = 0 \qquad \text{and} \qquad \lim_{ n \rightarrow \infty} \frac{ 1 - F(b_{n+1})}{1 - F(b_n)} = 1
-      \end{align*}
-      where $b_n$ is a sequence of thresholds
-      \item If for any $x \ge 0$,
-      \begin{align*}
-        1 - \frac{ 1 - F(a_n x + b_n) }{ 1 - F(b_n) } \converged H(x),
-      \end{align*}
-      then $H(x)$ follows a generalized Pareto distribution
-    \end{itemize}
-  \end{itemize}
-\end{frame}
-
-\begin{frame}{Standard non-spatial analysis: Peaks over threshold}
-  \begin{itemize} \setlength{\itemsep}{0.5em}
-    \item Generalized Pareto distribution has two parameters:
-    \begin{itemize}
+    where
+    \begin{itemize}
+      \item $\mu \in \calR$ is a location parameter
       \item $\sigma > 0$ is a scale parameter
@@
     \begin{itemize}
-      \item Sensitive to threshold selection
-      \item Temporal dependence between observations (e.g. flood levels don't dissipate overnight)
+      \item Lose information by only considering maximum in a block
+      \item Underlying data may not be i.i.d.
     \end{itemize}
@@
 \begin{frame}{Standard non-spatial analysis: Peaks over threshold}
-  \begin{itemize}  \setlength{\itemsep}{0.5em}
+  \begin{itemize} \setlength{\itemsep}{1em}
+    \item Let $X_1, \ldots, X_n \sim F$
+    \item If there exist normalizing sequences $a_t > 0$ and $b_t \in \calR$ such that for any $x \ge 0$, as $T \rightarrow \infty$
+    \begin{align*}
+      \Pr\left(\frac{X - b_t}{a_t} > t \mid X > T\right) \converged H(x),
+    \end{align*}
+    where $T$ is a thresholding value, then $H(x)$ follows a generalized Pareto distribution (GPD) (Balkema and de Haan, 1974)
+  \end{itemize}
+\end{frame}
+
+\begin{frame}{Standard non-spatial analysis: Peaks over threshold}
+  \begin{itemize}  \setlength{\itemsep}{1em}
     \item Select a threshold, $T$, and use the Generalized Pareto distribution to model the exceedances
@@
     \end{align*}
-    \item Related to GEV distribution through
+    where
+    \begin{itemize}
+      \item $\sigma > 0$ is a scale parameter
+      \item $\xi \in \calR$ is a shape parameter
+      \begin{itemize}
+        \item Unbounded above if $\xi \ge 0$
+        \item Bounded above by $(T - \sigma) / \xi$ when $\xi < 0$
+      \end{itemize}
+    \end{itemize}
+  \end{itemize}
+\end{frame}
+
+\begin{frame}{Standard non-spatial analysis: Peaks over threshold}
+  \begin{itemize} \setlength{\itemsep}{1em}
+    \item The GPD is related to GEV distribution through
     \begin{align*}
@@
     \end{align*}
-  \end{itemize}
-\end{frame}
-
-\begin{frame}{Max-stable processes}
-  \begin{itemize} \setlength{\itemsep}{0.5em}
-    \item For a spatial analysis, max-stable processes give an appropriate limiting distribution (Cooley et al., 2012)
-    \begin{itemize}
-      \item Consider a spatial process $x_t(\bs)$, $t = 1, \ldots, T$
-      \item Let $M_T(\bs) = \left\{ \bigvee_{t=1}^T x_t(\bs_1), \ldots, \bigvee_{t=1}^T x_t(\bs_n) \right\}$
-      \item If there exists normalizing sequences $a_T(\bs)$ and $b_T(\bs)$ such
-      that for all sites, $\bs_i, i = 1, \ldots, d$,
-      \begin{align*}
-        a_T^{-1}(\bs) \left\{ M_T(\bs) - b_T(\bs) \right\} \converged Y(\bs)
+    \item Challenges:
+    \begin{itemize}
+      \item Sensitive to threshold selection
+      \item Temporal dependence between observations (e.g. flood levels don't dissipate overnight)
+    \end{itemize}
+  \end{itemize}
+\end{frame}
+
+\begin{frame}{Max-stable processes for spatial data}
+  \begin{itemize} \setlength{\itemsep}{1em}
+    \item Consider i.i.d. spatial processes $x_j(\bs)$, $j = 1, \ldots, J$
+    \item Let $M_J(\bs) = \bigvee_{j=1}^J x_j(\bs_i)$ be the block max at site $\bs$
+    \item If there exists normalizing sequences $a_J(\bs)$ and $b_J(\bs)$ such that for all sites, $\bs_i, i = 1, \ldots, d$,
+    \begin{align*}
+      a_J^{-1}(\bs) \left\{ M_J(\bs) - b_J(\bs) \right\} \converged Y(\bs)
     \end{align*}
     which has a non-degenerate distribution, then $Y(\bs)$ is a max-stable process
-    \end{itemize}
+    \item Therefore, max-stable processes are the standard for block maxima
   \end{itemize}
@@
   \begin{itemize}
-    \item Marginally at each site, observations follow a generalized extreme value distribution
-    \item For a finite collection of sites
-    \begin{itemize}
-      \item The multivariate representation for the GEV is
+    \item Marginally at each site, observations follow a GEV distribution
+    \item For a finite collection of sites the multivariate representation for the GEV (mGEV) is
     \begin{align*}
       \Pr(\bZ \le \bz)  &= G^*(\bz) = \exp(-V(\bz))\\
-                V(\bs)    &= d \int_{\Delta_d} \bigvee_{i = 1}^d \frac{w_i}{z_i} H(\ddd w)
+            V(\bz)    &= d \int_{\Delta_d} \bigvee_{i = 1}^d \frac{w_i}{z_i} H(\ddd w)
     \end{align*}
@@
   \end{itemize}
-  \end{itemize}
 \end{frame}
@@
   \begin{itemize}
-    \item Not a lot of existing methods
+    \item Few existing methods
     \item Often use max-stable methods due to the relationship between GEV and GPD
@@
 \begin{frame}{Extremal dependence: $\chi$ statistic}
-  \begin{itemize} \setlength{\itemsep}{0.5em}
-    \item Correlation is the most common measure of dependence, but it is irrelevant for extreme value analysis
-    \item Extreme value analysis focuses on the $\chi$ statistic, a measure of extremal dependence
-    \item Specifically, we focus on $\chi(h)$ for the upper tail given by
-    \begin{align*}
-      \chi(h) = \lim_{c \rightarrow \infty}\Pr(Y(\bs) > c \mid Y(\bt) > c)
+  \begin{itemize} \setlength{\itemsep}{1em}
+    \item Correlation is the most common measure of dependence
+    \begin{itemize}
+      \item Focuses on the center and not tails
+      \item This makes it irrelevant for extreme value analysis
+    \end{itemize}
+    \item Extreme value analysis focuses on the $\chi$ statistic, a measure of extremal dependence given by
+    \begin{align*}
+      \chi(h) = \lim_{c \rightarrow \infty}\Pr[Y(\bs) > c \mid Y(\bt) > c]
     \end{align*}
@@
 \begin{frame}{Existing challenges}
-  \begin{itemize} \setlength{\itemsep}{0.5em}
+  \begin{itemize} \setlength{\itemsep}{1em}
     \item Multivariate max-stable and GPD models have nice features, but they are
     \begin{itemize}
-      \item Computationally challenging (Falk et al., 2011)
-      \begin{itemize}
-        \item Asymmetric logistic has $2^{n-1}(n + 2) - (2n + 1)$ free parameters
-      \end{itemize}
-      \item Joint distribution only available in low dimension
+      \item Computationally challenging (e.g,  the asymmetric logistic has $2^{n-1}(n + 2) - (2n + 1)$ free parameters)
+      \item Joint distribution only available in low dimensions
     \end{itemize}
@@
 \begin{frame}{Three principal contributions}
-  \begin{enumerate}[1] \setlength{\itemsep}{0.5em}
+  \begin{enumerate}[1] \setlength{\itemsep}{1em}
     \item A spatio-temporal model with flexible tails, asymptotic spatial dependence, and computation on the order of Gaussian models for large space-time datasets
@@
 \begin{frame}{Spatiotemporal modeling for extreme values}
-  \begin{itemize} \setlength{\itemsep}{0.5em}
-    \item Model to analysis spatiotemporal extreme values
-    \item Model objectives
-    \begin{itemize}
-      \item Has marginal distribution with a flexible tail
+  \begin{itemize} \setlength{\itemsep}{1em}
+    \item Model objectives:
+    \begin{itemize}
+      \item Marginal distribution with a flexible tail
       \begin{itemize}
@@
       \end{itemize}
-      \item Has asymptotic spatial dependence
-      \item Has computation on the order of Gaussian models for large space-time datasets
+      \item Asymptotic spatial dependence
+      \item Computation on the order of Gaussian models for large space-time datasets
     \end{itemize}
@@
 \begin{frame}{Gaussian spatial model}
-  \begin{itemize} \setlength{\itemsep}{0.5em}
+  \begin{itemize} \setlength{\itemsep}{1em}
     \item In geostatistics $Y(\bs)$ are often modeled using a Gaussian process with mean function $\mu(\bs)$ and covariance function $\rho(h)$.
@@
 \begin{frame}{Spatial skew-$t$ distribution}
-  \begin{itemize} \setlength{\itemsep}{0.5em}
-    \item Assume observed data $Y(\bs)$ come from a skew-$t$ (Zhang and El-Shaarawi, 2012)
+  \begin{itemize} \setlength{\itemsep}{1em}
+    \item A more flexible alternative is the spatial skew-$t$ process (Zhang and El-Shaarawi, 2012)
     \begin{align*}
@@
 \begin{frame}{Spatial skew-$t$ distribution}
-  \begin{itemize} \setlength{\itemsep}{0.5em}
+  \begin{itemize} \setlength{\itemsep}{1em}
     \item \alert{Conditioned} on $z$ and $\sigma^2$, $Y(\bs)$ is a Gaussian spatial model
@@
 \begin{frame}{Spatial skew-$t$ distribution}
-  \begin{itemize} \setlength{\itemsep}{0.5em}
+  \begin{itemize} \setlength{\itemsep}{1em}
     \item \alert{Marginalizing} over $z$ and $\sigma^2$ (via MCMC),
@@
 
+\begin{frame}{Spatial skew-$t$ distribution properties}
+  \begin{itemize} \setlength{\itemsep}{1em}
+    \item Model properties:
+    \begin{itemize}
+      \item Flexible tail
+      \begin{itemize}
+        \item Skewness controlled by $\lambda$
+        \item Weight of tails controlled by degrees of freedom $2a$
+      \end{itemize}
+      \item For a skew-$t$ distribution $\lim_{c \rightarrow \infty} \chi(h) > 0$ (Padoan, 2011)
+      \item Computation that is on the order of Gaussian computation
+    \end{itemize}
+    \item Challenge: $\chi(h) > 0$ as $h \rightarrow \infty$ (Padoan, 2011)
+    \begin{itemize}
+      \item This occurs because all observations (near and far) share the same $z$ and $\sigma^2$
+    \end{itemize}
+  \end{itemize}
+\end{frame}
+
 \begin{frame}{Extension of the skew-$t$ distribution}
-  \begin{itemize} \setlength{\itemsep}{0.5em}
+  \begin{itemize} \setlength{\itemsep}{1em}
     \item Skew-$t$ distribution addresses two modeling concerns
@@
 \begin{frame}{Censoring data to focus on tail behavior}
-  \begin{itemize} \setlength{\itemsep}{0.5em}
+  \begin{itemize} \setlength{\itemsep}{1em}
     \item We censor the observed data at a high threshold $T$
@@
 
-\begin{frame}{Spatial skew-$t$ distribution properties}
-  \begin{itemize} \setlength{\itemsep}{0.5em}
-      \item Model properties
-    \begin{itemize}
-    	\item Has flexible tail
-      \begin{itemize}
-        \item Skewness controlled by $\lambda$
-        \item Weight of tails controlled by degrees of freedom $2a$
-      \end{itemize}
-    	\item For a skew-$t$ distribution $\lim_{c \rightarrow \infty} \chi(h) > 0$ (Padoan, 2011)
-    	\item Computation that is on the order of Gaussian computation
-    \end{itemize}
-    \item Challenge: $\chi(h) > 0$ as $h \rightarrow \infty$ (Padoan, 2011)
-    \begin{itemize}
-      \item This occurs because all observations (near and far) share the same $z$ and $\sigma^2$
-      \item We deal with this through a daily random partition (similar to Kim et al., 2005 for non-extreme modeling)
-    \end{itemize}
-  \end{itemize}
-\end{frame}
-
 \begin{frame}{Random partition}
-  \begin{itemize} \setlength{\itemsep}{0.5em}
+  \begin{itemize} \setlength{\itemsep}{1em}
     \item Daily random partition allows $z$ and $\sigma^2$ to vary by site
@@
     \end{align*}
-    where $\bw = (w_1, w_2)$
+    where $\bw = (w_1, w_2)$ (similar to Kim et al., 2005 for non-extreme modeling)
     \item For $\bs \in P_{k}$
@@
     \end{align*}
-  \end{itemize}
-\end{frame}
-
-\begin{frame}{Random partition}
-  \begin{itemize} \setlength{\itemsep}{0.5em}
-    \item Within each partition $Y(\bs)$ has the same MV skew-t distribution as before
-    \item Across partitions $Y(\bs)$ are asymptotically independent, but still correlated through $v(\bs)$
   \end{itemize}
@@
 
+\begin{frame}{Random partition}
+  \begin{itemize} \setlength{\itemsep}{1em}
+    \item Within each partition $Y(\bs)$ has the same MV skew-t distribution as before
+    \item Across partitions $Y(\bs)$ are asymptotically independent, but still correlated through $v(\bs)$
+    \item New expression for $\chi(h)$
+    \begin{align*}
+      \chi(h) = \pi(h) \chi_{\text{skew-}t}(h)
+    \end{align*}
+    where
+    \begin{itemize}
+      \item $\pi(h)$ is the probability two sites are in the same partition
+    \end{itemize}
+  \end{itemize}
+\end{frame}
+
 \begin{frame}{$\chi(h)$ plot}
@@
 \begin{frame}{Random partition skew-$t$ model}
-  \begin{itemize} \setlength{\itemsep}{0.5em}
+  \begin{itemize} \setlength{\itemsep}{1em}
     \item This new model is called a random partition skew-$t$ model, and it has the properties we desire
@@
 \begin{frame}{MCMC details}
-  \begin{itemize} \setlength{\itemsep}{0.5em}
+  \begin{itemize} \setlength{\itemsep}{1em}
     \item Three main steps
@@
 \begin{frame}{Simulation study}
-  \begin{itemize} \setlength{\itemsep}{0.5em}
-    \item 6 different data settings:
+  \begin{itemize} \setlength{\itemsep}{1em}
+    \item 6 different data settings
     \begin{itemize}
       \item Gaussian, $K = 1$ partition
-      \item $T$, $K = 1$ partition
-      \item $T$, $K = 5$ partitions
+      \item Symmetric-$t$, $K = 1$ partition
+      \item Symmetric-$t$, $K = 5$ partitions
       \item Skew-$t$, $K = 1$ partition
@@
 \begin{frame}{Simulation study}
-  \begin{itemize} \setlength{\itemsep}{0.5em}
-    \item 50 dataset for each setting
+  \begin{itemize} \setlength{\itemsep}{1em}
+    \item 50 datasets for each setting
     \begin{itemize}
@@
 \begin{frame}{Simulation study}
-  \begin{itemize} \setlength{\itemsep}{0.5em}
-    \item 5 different models:
+  \begin{itemize} \setlength{\itemsep}{1em}
+    \item 5 different models fit to each data set
     \begin{itemize}
@@
 % \begin{frame}{Quantile score for cross-validation}
-%   \begin{itemize} \setlength{\itemsep}{0.5em}
+%   \begin{itemize} \setlength{\itemsep}{1em}
 %     \item The quantile score for the $\tau$th quantile is
@@
 \begin{frame}{Brier score}
-  \begin{itemize} \setlength{\itemsep}{0.5em}
+  \begin{itemize} \setlength{\itemsep}{1em}
     \item Brier score used to determine model that gives best fit (Gneiting and Raftery, 2007)
@@
 \begin{frame}{Simulation study results}
-  \begin{itemize} \setlength{\itemsep}{0.5em}
+  \begin{itemize} \setlength{\itemsep}{1em}
     \item Key findings
@@
 \column{.45 \linewidth}
-	\begin{itemize} \setlength{\itemsep}{0.5em}
+	\begin{itemize} \setlength{\itemsep}{1em}
 	\item Ozone measurements
@@
 		\item max 8-hour ozone measurements
-		\item data from 1089 sites
+		\item daily data from 1089 sites
 		\item July 2005
@@
   \end{itemize}
+  \item Conduct two-fold cross-validation on 800 sites
 	\end{itemize}
@@
 
-\begin{frame}{Exploratory data analysis}
-	\centering
-  \begin{figure}
-    \includegraphics[width=1\linewidth, trim=0 0 0 1in]{./plots/pot/chi-plot-ozone-res.pdf}
-    \caption{$\widehat{\chi}$-plot for sample quantiles of ozone observations}
-  \end{figure}
-\end{frame}
+% \begin{frame}{Exploratory data analysis}
+% 	\centering
+%   \begin{figure}
+%     \includegraphics[width=1\linewidth, trim=0 0 0 1in]{./plots/pot/chi-plot-ozone-res.pdf}
+%     \caption{$\widehat{\chi}$-plot for sample quantiles of ozone observations}
+%   \end{figure}
+% \end{frame}
 
 \begin{frame}{Model comparisons}
-  \begin{itemize} \setlength{\itemsep}{0.5em}
+  \begin{itemize} \setlength{\itemsep}{1em}
     \item 9 different analysis methods incorporating
@@
 \begin{frame}{Model comparisons}
-  \begin{itemize} \setlength{\itemsep}{0.5em}
-    \item All methods use a \Matern or exponential covariance ($\nu = 0.5$)
-    \item Covariate data from the Environmental Protection Agency's Community Multiscale Air Quality (CMAQ) system.
+  \begin{itemize} \setlength{\itemsep}{1em}
+    \item The Community Multiscale Air Quality (CMAQ) system provides fine-resolution simulated values for multiple air pollutants
+    \item We use the tropospheric ozone output from the corresponding days in the CMAQ model as a covariate
     \item Mean function modeled as
@@
     \end{align*}
-   \end{itemize}
-\end{frame}
-
-\begin{frame}{Two-fold cross-validation results}
+    \item All methods use a \Matern covariance
+   \end{itemize}
+\end{frame}
+
+\begin{frame}{Cross-validation results}
   \centering
   \begin{figure}
-    \includegraphics[width=0.47\linewidth]{./plots/pot/bs-ozone-1.pdf}
-    \includegraphics[width=0.47\linewidth]{./plots/pot/bs-ozone-2.pdf} \\
-    \caption{Relative Brier score results}
-  \end{figure}
-\end{frame}
-
-\begin{frame}{Two-fold cross-validation results}
-  \begin{itemize} \setlength{\itemsep}{0.5em}
+    \includegraphics[width=0.6\linewidth, trim=0 0 0 1in]{./plots/pot/bs-ozone-1.pdf}
+    % \includegraphics[width=0.47\linewidth]{./plots/pot/bs-ozone-2.pdf} \\
+    \caption{Relative Brier score results ($K = 6$ -- $9$ are similar to $K = 5$ and $K = 10$)}
+  \end{figure}
+\end{frame}
+
+\begin{frame}{Cross-validation results}
+  \begin{itemize} \setlength{\itemsep}{1em}
     \item Key findings
@@
 \begin{frame}{Discussion}
-  \begin{itemize} \setlength{\itemsep}{0.5em}
+  \begin{itemize} \setlength{\itemsep}{1em}
     \item Improvement of model performance when using partitioned models
@@
 
+\begin{frame}{Future work: Knots and their impact}
+  \begin{itemize} \setlength{\itemsep}{1em}
+    \item Different partition structure
+    \begin{itemize}
+      \item Distance weighting for each knot vs indicator functions
+    \end{itemize}
+    \item Knot selection
+    \begin{itemize}
+      \item Possible prior on the probability a knot is in the spatial domain
+    \end{itemize}
+  \end{itemize}
+\end{frame}
+
 \begin{frame}{Future work: Temporal dependence}
-  \begin{itemize} \setlength{\itemsep}{0.5em}
+  \begin{itemize} \setlength{\itemsep}{1em}
     \item Temporal dependence should be accounted for when using daily data
@@
       \end{align*}
-      where $t = 1, \ldots, T$ denotes the day of each observation
+      where $t$ denotes the day of each observation
     \item Different ways to incorporate the temporal dependence
@@
 \begin{frame}{Future work: Temporal dependence}
-  \begin{itemize} \setlength{\itemsep}{0.5em}
+  \begin{itemize} \setlength{\itemsep}{1em}
     \item We choose the time series approach because the $z_t(\bs)$ and $\sigma_t(\bs)$ terms dictate the tail behavior
@@
 
-\begin{frame}{Future work: Knots and their impact}
-  \begin{itemize} \setlength{\itemsep}{0.5em}
-    \item Different partition structure
-    \begin{itemize}
-      \item Distance weighting for each knot vs indicator functions
-    \end{itemize}
-    \item Knot selection
-    \begin{itemize}
-      \item Possible prior on the probability a knot is in the spatial domain
-    \end{itemize}
-  \end{itemize}
-\end{frame}
-
-\begin{frame}{Rare spatial binary regression}
-  \begin{itemize} \setlength{\itemsep}{0.5em}
+\begin{frame}{Rare binary regression}
+  \begin{itemize} \setlength{\itemsep}{1em}
     \item Motivation
     \begin{itemize}
-      \item Want to incorporate spatial dependence when modeling rare events
-      \item Examples
-      \begin{itemize}
-        \item Diseased trees
-        \item Disease outbreak
-      \end{itemize}
+      \item Want to incorporate spatial dependence when modeling rare events (e.g. Diseased trees, Disease outbreak, Crimes)
     \end{itemize}
@@
 
-\begin{frame}{Rare spatial binary regression}
-  \begin{itemize} \setlength{\itemsep}{0.5em}
+\begin{frame}{Rare binary regression}
+  \begin{itemize} \setlength{\itemsep}{1em}
     \item Common examples with non-spatial analysis
@@
       \begin{align*}
-        \Pr[Y_i = 1] = \exp[ -\exp(-\bX_i \bbeta)]
-      \end{align*}
-    \end{itemize}
-  \end{itemize}
-\end{frame}
-
-\begin{frame}{Rare spatial binary regression}
-  \begin{itemize} \setlength{\itemsep}{0.5em}
+        \Pr[Y_i = 1] = 1 - \exp[ -\exp(\bX_i \bbeta)]
+      \end{align*}
+    \end{itemize}
+  \end{itemize}
+\end{frame}
+
+\begin{frame}{Rare binary regression}
+  \begin{itemize} \setlength{\itemsep}{1em}
     \item Generalized extreme value link function (Wang and Dey, 2010)
     \begin{align*}
-      \Pr[Y_i = 1] = 1 - \exp\left[-(1 + \xi \bX_i \bbeta)^{-1 / \xi} \right]
+      \Pr[Y_i = 1] = 1 - \exp\left[-(1 - \xi \bX_i \bbeta)^{-1 / \xi} \right]
     \end{align*}
@@
 \begin{frame}{Rare spatial binary regression}
-  \begin{itemize} \setlength{\itemsep}{0.5em}
+  \begin{itemize} \setlength{\itemsep}{1em}
+    \item We propose to develop a spatial model
+    \item Objectives are to borrow strength across sites to estimate covariate effects and spatial prediction
     \item Proposed method will
@@
 \begin{frame}{Rare spatial binary regression}
-  \begin{itemize} \setlength{\itemsep}{0.5em}
-    \item We fit parameters $\xi$ and $\beta$ in order to transform the data to GEV(1, 1, 1) marginal distributions
-    \item Using the link function
-    \begin{align*}
-      p_i = 1 - \exp \left[-(1 + \xi \bX(\bs_i) \bbeta)^{-1/\xi}\right]
-    \end{align*}
-    \item We model $Y_i = I(Z_i > 0)$ where
-    \begin{itemize}
-      \item $z_i\sim$ Multivariate GEV (Reich and Shaby, 2012) is a latent variable
+  \begin{itemize} \setlength{\itemsep}{1em}
+    \item We model $Y_i = I(Z_i > 0)$ where $z_i\sim$ multivariate GEV (mGEV)  is a latent variable
+    \item Hierarchical model for mGEV (Reich and Shaby, 2012)
+    \begin{itemize}
+      \item $Z(\bs) = U(\bs) \theta(\bs)$
+      \begin{itemize}
+        \item $U(\bs) \iid$ GEV(1, $\alpha$, $\alpha$) is a nugget effect
+        \item $\theta(\bs) = [\sum_{l = 1}^L A_l w_l(\bs)^{1/\alpha}]^\alpha$ is the spatial process
+        \item $\alpha \in (0, 1)$ controls strength of nugget relative to spatial dependence
+        \item $A_l \iid$ Positive Stable($\alpha$) is an random effect representing the intensity
+        \item $w_l(\bs)$ gives the weight of the intensity of the $l$th random effect on site $\bs$
+      \end{itemize}
     \end{itemize}
@@
 \begin{frame}{Likelihood function}
-  \begin{itemize} \setlength{\itemsep}{0.5em}
-    \item Asymmetric logistic dependence function (Reich and Shaby, 2012)
+  \begin{itemize} \setlength{\itemsep}{1em}
+    \item After marginalizing out the $A_l$ terms, we have the asymmetric logistic dependence function (Reich and Shaby, 2012)
     \footnotesize{
@@
 \begin{frame}{Weighting function}
-  \begin{itemize} \setlength{\itemsep}{0.5em}
+  \begin{itemize} \setlength{\itemsep}{1em}
     \item We use the Gaussian weights proposed by Reich and Shaby (2012) given by
@@
     \includegraphics[width=\linewidth, trim=0 0.5in 0 0]{./plots/max-stable.pdf}
-    \caption{Knot intensity $\times$ weight, red dots give intensity of random effects (left) Impact of $\alpha$ (right)}
+    \caption{Knot intensity $\times$ weight ($\rho = 0.5$), red dots give intensity of random effects (left) Impact of $\alpha$ (right).}
    \end{figure}
@@
 \begin{frame}{Joint likelihood}
-  \begin{itemize} \setlength{\itemsep}{0.5em}
+  \begin{itemize} \setlength{\itemsep}{1em}
     \item Let $K_t = \sum_{i = 1}^n Y_{it}$ be the number of exceedances that occur on day $t$.
@@
     }
-    where $G(\bz_{(1)} = \Pr(Z_2 < z_2, \ldots, Z_n < z_n)$
+    where $G\left(\bz_{(1)}\right) = \Pr(Z_2 < z_2, \ldots, Z_n < z_n)$
     \item $K > 2$ can be derived similarly
@@
 \begin{frame}{Joint likelihood}
-  \begin{itemize} \setlength{\itemsep}{0.5em}
+  \begin{itemize} \setlength{\itemsep}{1em}
     \item For small $K$, we can evaluate the likelihood directly
@@
 
-\begin{frame}{Simulation study}
-  \begin{itemize} \setlength{\itemsep}{0.5em}
+\begin{frame}{Future simulation study and data application}
+  \begin{itemize} \setlength{\itemsep}{1em}
+    \item Simulation study
+    \begin{itemize}
       \item Data generated using logistic, Cloglog, and GEV links
       \begin{itemize}
-      \item Exploring how sparseness of observations impacts prediction of events
-    \end{itemize}
-    \item Models fit using multivariate GEV, and Gaussian distribution
+        \item Exploring how rarity of event impacts prediction
+      \end{itemize}
+      \item Models fit using
+      \begin{itemize}
+        \item mGEV
+        \item Random effects Gaussian distribution
+      \end{itemize}
+    \end{itemize}
+    \item Data application: Modeling crime data
+    \begin{itemize}
+      \item Homicides, car theft, vandalism
+    \end{itemize}
   \end{itemize}
@@
 \begin{frame}{Non-stationary covariance for extreme values}
-  \begin{itemize} \setlength{\itemsep}{0.5em}
+  \begin{itemize} \setlength{\itemsep}{1em}
     \item Stationary covariance functions are a function of distance between two sites.
     \begin{align*}
-      \rho(Y(\bs), Y(\bt)) = \rho(h)
+      \rho[Y(\bs), Y(\bt)] = \rho(h)
     \end{align*}
     where $h = ||\bs - \bt||$
+    \item This assumes the covariance is the same everywhere, e.g. east vs west, mountains vs desert
+    \item Misspecifying the covariance can impair spatial prediction and statistical inference
+  \end{itemize}
+\end{frame}
+
+\begin{frame}{Non-stationary covariance for extreme values}
+  \begin{itemize} \setlength{\itemsep}{1em}
     \item In extremes, stationary extremal dependence means
@@
     \end{align*}
-    \item Sometimes the angle between
-    \begin{itemize}
-      \item Model does not allow for non-stationary covariance functions
-    \end{itemize}
-  \end{itemize}
-\end{frame}
-
-\begin{frame}{Non-stationary covariance for extreme values}
-  \begin{itemize} \setlength{\itemsep}{0.5em}
+    \item Currently, there are no methods to model non-stationarity in spatial extremes
     \item Semiparamtric approach using spectral density ratios (de Carvalho and Davison, 2014)
@@
 \begin{frame}{Thesis outline}
-  \begin{itemize} \setlength{\itemsep}{0.5em}
-    \item Chapter 1: Extreme value theory \alert{August 2015}
+  \begin{itemize} \setlength{\itemsep}{1em}
+    \item Chapter 1: Review of extreme value theory \alert{August 2015}
     \item Chapter 2: Spatiotemporal model for extreme value analysis based on the skew-$t$ distribution \alert{February 2015}
-    \item Chapter 3: Spatial binary regression \alert{May / June 2015}
+    \item Chapter 3: Rare spatial binary regression \alert{May / June 2015}
     \item Chapter 4: Non-stationary covariance through knot-specific bandwidth \alert{August 2015}
@@
 \begin{frame}{Questions}
-  \begin{itemize} \setlength{\itemsep}{0.5em}
+  \begin{itemize} \setlength{\itemsep}{1em}
     \item Questions?
@@
 
-\begin{frame}{References}
-  \begin{itemize} \setlength{\itemsep}{0.5em}
-    \item Demarta, S. and McNeil, A. J. (2007) The $t$ copula and related copulas. {\it International Statistical Review}, {\bf 73}, 111--129.
-    \item Huser, R. and Davison, A. C. (2014) Space-time modelling of extreme events. {\it Journal of the Royal Statistical Society: Series B (Statistical Methodology)}, {\bf 76}, 439--461.
-    \item Padoan, S. A. (2011) Multivariate extreme models based on underlying skew-$t$ and skew-normal distributions. {\it Journal of Multivariate Analysis}, {\bf 102}, 977--991.
-    \item Zhang, H. and El-Shaarawi, A. (2010) On spatial skew-Gaussian processes and applications. {\it Environmetrics}, {\bf 21}, 33--47.
-  \end{itemize}
-\end{frame}
+% \begin{frame}{References}
+%   \begin{itemize} \setlength{\itemsep}{1em}
+%     \item Demarta, S. and McNeil, A. J. (2007) The $t$ copula and related copulas. {\it International Statistical Review}, {\bf 73}, 111--129.
+%     \item Huser, R. and Davison, A. C. (2014) Space-time modelling of extreme events. {\it Journal of the Royal Statistical Society: Series B (Statistical Methodology)}, {\bf 76}, 439--461.
+%     \item Padoan, S. A. (2011) Multivariate extreme models based on underlying skew-$t$ and skew-normal distributions. {\it Journal of Multivariate Analysis}, {\bf 102}, 977--991.
+%     \item Zhang, H. and El-Shaarawi, A. (2010) On spatial skew-Gaussian processes and applications. {\it Environmetrics}, {\bf 21}, 33--47.
+%   \end{itemize}
+% \end{frame}
 

### 2015-01-08 prelim/oral_prelim.tex
updates to prelim

@@
 \usepackage{stackrel}
+\usepackage{changepage}
 
@@
 \usepackage[latin1]{inputenc}
+\usepackage[authoryear]{natbib}
 \usepackage{multirow}
@@
       \item e.g. When constructing river levees, engineers need to be able to estimate a 100-year or 1000-year flood levels
-      \item e.g. Probability of exceeding a certain threshold level
+      \item e.g. Probability of ambient air pollution exceeding a certain threshold level
     \end{itemize}
@@
 \begin{frame}{Motivation}
-  \begin{itemize} \setlength{\itemsep}{1em}
-    \item Ozone compliance for Clean Air Act (EPA)
-    \begin{itemize}
+\begin{adjustwidth}{1em}{0em}
+  Ozone compliance for Clean Air Act (EPA) \vspace{1em}
+  \begin{itemize} \setlength{\itemsep}{1em}
     \item Annual fourth-highest daily maximum 8-hour concentration, averaged over 3 years, not to exceed 75 ppb
     \item Annual fourth-highest is the 99th percentile for the year
-      \item Common objectives are
-      \begin{itemize}
+    \item Common objectives are \vspace{0.5em}
+    \begin{itemize} \setlength{\itemsep}{0.5em}
       \item To interpolate to unmonitored sites
       \item Detect changes in extremes over time
-        \item Study factors that lead to extreme events
-      \end{itemize}
-    \end{itemize}
-  \end{itemize}
+      \item Study meterological conditions that lead to extreme events
+    \end{itemize}
+  \end{itemize}
+\end{adjustwidth}
 \end{frame}
@@
     \includegraphics[width=1\linewidth, trim=0 0.5in 0 1in]{./plots/define_extreme.pdf}
-    \caption{Hypothetical monthly maximums recorded over a three years}
+    \caption{Hypothetical monthly data}
     \end{figure}
@@
 
-\begin{frame}{Standard non-spatial analysis: Block maxima}
+\begin{frame}{Non-spatial analysis: Block maxima}
   \begin{itemize} \setlength{\itemsep}{1em}
@@
     then $G(z)$ follows a generalized extreme value distribution (GEV) (Falk et al., 2011)
-    \item This motivates the use of the GEV for block max data
-  \end{itemize}
-\end{frame}
-
-\begin{frame}{Standard non-spatial analysis: Block maxima}
+    \item This motivates the use of the GEV for block maximum data
+  \end{itemize}
+\end{frame}
+
+\begin{frame}{Non-spatial analysis: Block maxima}
   \begin{itemize} \setlength{\itemsep}{1em}
@@
     where
-    \begin{itemize}
+    \begin{itemize} \setlength{\itemsep}{0.25em}
       \item $\mu \in \calR$ is a location parameter
@@
 
-\begin{frame}{Standard non-spatial analysis: Peaks over threshold}
+\begin{frame}{Non-spatial analysis: Peaks over threshold}
   \begin{itemize} \setlength{\itemsep}{1em}
@@
     \begin{align*}
-      \Pr\left(\frac{X - b_t}{a_t} > t \mid X > T\right) \converged H(x),
+      \Pr\left(\frac{X - b_t}{a_t} > x \mid X > T\right) \converged H(x),
     \end{align*}
@@
 
-\begin{frame}{Standard non-spatial analysis: Peaks over threshold}
-  \begin{itemize}  \setlength{\itemsep}{1em}
-    \item Select a threshold, $T$, and use the Generalized Pareto distribution to model the exceedances
+\begin{frame}{Non-spatial analysis: Peaks over threshold}
+\begin{adjustwidth}{1em}{0em}
+  Select a threshold, $T$, and use the GPD to model the exceedances
   \begin{align*}
@@
   where
-    \begin{itemize}
+  \begin{itemize} \setlength{\itemsep}{0.25em}
     \item $\sigma > 0$ is a scale parameter
@@
   \end{itemize}
-  \end{itemize}
-\end{frame}
-
-\begin{frame}{Standard non-spatial analysis: Peaks over threshold}
+\end{adjustwidth}
+\end{frame}
+
+\begin{frame}{Non-spatial analysis: Peaks over threshold}
   \begin{itemize} \setlength{\itemsep}{1em}
@@
     \end{align*}
-    \item Challenges:
-    \begin{itemize}
+    \item Challenges: \vspace{0.5em}
+    \begin{itemize} \setlength{\itemsep}{0.5em}
       \item Sensitive to threshold selection
@@
     \item Consider i.i.d. spatial processes $x_j(\bs)$, $j = 1, \ldots, J$
-    \item Let $M_J(\bs) = \bigvee_{j=1}^J x_j(\bs_i)$ be the block max at site $\bs$
+    \item Let $M_J(\bs) = \bigvee_{j=1}^J x_j(\bs_i)$ be the block maximum at site $\bs$
     \item If there exists normalizing sequences $a_J(\bs)$ and $b_J(\bs)$ such that for all sites, $\bs_i, i = 1, \ldots, d$,
@@
     which has a non-degenerate distribution, then $Y(\bs)$ is a max-stable process
-    \item Therefore, max-stable processes are the standard for block maxima
+    \item Therefore, max-stable processes are the standard model for block maxima
   \end{itemize}
@@
 \begin{frame}{Multivariate representations}
-  \begin{itemize}
+  \begin{itemize} \setlength{\itemsep}{1em}
     \item Marginally at each site, observations follow a GEV distribution
@@
     where
-    \begin{itemize}
+    \begin{itemize} \setlength{\itemsep}{0.25em}
       \item $\Delta_d = \{ \bw \in \calR^d_+ \mid w_1 + \cdots + w_d = 1\}$
       \item $H$ is a probability measure on $\Delta_d$
-      \item $\int_{\Delta_d}w_i H(\ddd w) = 1 / d$ for $i = 1, \ldots, d$.
+      \item $\int_{\Delta_d}w_i H(\ddd w) = 1 / d$ for $i = 1, \ldots, d$
     \end{itemize}
@@
 \begin{frame}{Multivariate GEV challenges}
-  \begin{itemize}
+  \begin{itemize} \setlength{\itemsep}{1em}
     \item Only a few closed-form expressions for $V(\bz)$ exist (Stephenson, 2003)
@@
       \end{align*}
-      where $w_{il} \in [0, 1]$ and $\sum_{l = 1}^L w_{il} = 1$.
+      where $w_{il} \in [0, 1]$ and $\sum_{l = 1}^L w_{il} = 1$
     \end{itemize}
@@
 \begin{frame}{Multivariate peaks over threshold}
-  \begin{itemize}
+  \begin{itemize} \setlength{\itemsep}{1em}
     \item Few existing methods
@@
       \item Computationally challenging (e.g,  the asymmetric logistic has $2^{n-1}(n + 2) - (2n + 1)$ free parameters)
-      \item Joint distribution only available in low dimensions
+      \item Joint density only available in low dimensions
     \end{itemize}
@@
 \begin{frame}{Three principal contributions}
-  \begin{enumerate}[1] \setlength{\itemsep}{1em}
+\begin{adjustwidth}{1em}{0em}
+  \begin{enumerate}[1.] \setlength{\itemsep}{1.5em}
     \item A spatio-temporal model with flexible tails, asymptotic spatial dependence, and computation on the order of Gaussian models for large space-time datasets
@@
   \end{enumerate}
+\end{adjustwidth}
 \end{frame}
@@
 \begin{frame}{Spatiotemporal modeling for extreme values}
-  \begin{itemize} \setlength{\itemsep}{1em}
-    \item Model objectives:
-    \begin{itemize}
-      \item Marginal distribution with a flexible tail
+\begin{adjustwidth}{1em}{0em}
+  Model objectives: \vspace{1em}
+  \begin{itemize} \setlength{\itemsep}{1em} \setlength{\itemindent}{1em}
+    \item Marginal distribution at each site with a flexible tail
     \begin{itemize}
@@
   \end{itemize}
-  \end{itemize}
+\end{adjustwidth}
 \end{frame}
@@
   \begin{itemize} \setlength{\itemsep}{1em}
-    \item In geostatistics $Y(\bs)$ are often modeled using a Gaussian process with mean function $\mu(\bs)$ and covariance function $\rho(h)$.
+    \item In geostatistics, $Y(\bs)$ are often modeled using a Gaussian process with mean function $\mu(\bs)$ and covariance function $\rho(h)$.
     \item Model properties
@@
       \item Nice computing properties (closed-form likelihood)
-      \item For a Gaussian spatial model $\lim_{c \rightarrow \infty} \chi(h) = 0$ regardless of the strength of the correlation in the bulk of the distribution
+      \item For a Gaussian spatial model $\chi(h) = 0$ regardless of the strength of the correlation in the bulk of the distribution
       \item Tail is not flexible
@@
 \begin{frame}{Spatial skew-$t$ distribution}
-  \begin{itemize} \setlength{\itemsep}{1em}
-    \item A more flexible alternative is the spatial skew-$t$ process (Zhang and El-Shaarawi, 2012)
+\begin{adjustwidth}{1em}{0em}
+  A more flexible alternative is the spatial skew-$t$ process (Zhang and El-Shaarawi, 2012)
   \begin{align*}
@@
     \item \alert{Conditioned} on $z$ and $\sigma^2$, $Y(\bs)$ is a Gaussian spatial model
-    \item Can use standard geostatistical methods to fit this model
+    \item Standard geostatistical methods apply
     \item Predictions can be made through Kriging
@@
 \begin{frame}{Spatial skew-$t$ distribution}
-  \begin{itemize} \setlength{\itemsep}{1em}
-    \item \alert{Marginalizing} over $z$ and $\sigma^2$ (via MCMC),
+\begin{adjustwidth}{1em}{0em}
+  \alert{Marginalizing} over $z$ and $\sigma^2$ (via MCMC),
   \begin{align*}
@@
   where
-    \begin{itemize}
+  \begin{itemize} \setlength{\itemsep}{0.25em}
     \item $\bX(\bs) \bbeta$ is the location
@@
   \begin{itemize} \setlength{\itemsep}{1em}
-    \item Model properties:
-    \begin{itemize}
+    \item Model properties: \vspace{0.5em}
+    \begin{itemize} \setlength{\itemsep}{0.5em}
       \item Flexible tail
@@
     \end{itemize}
-    \item Challenge: $\chi(h) > 0$ as $h \rightarrow \infty$ (Padoan, 2011)
-    \begin{itemize}
-      \item This occurs because all observations (near and far) share the same $z$ and $\sigma^2$
-    \end{itemize}
+    \item Challenge: Long-range dependence occurs because all observations (near and far) share the same $z$ and $\sigma^2$ (Padoan, 2011)
+    \begin{align*}
+      \lim_{h \rightarrow \infty} \chi(h) > 0
+    \end{align*}
   \end{itemize}
@@
   \begin{itemize} \setlength{\itemsep}{1em}
-    \item Skew-$t$ distribution addresses two modeling concerns
-    \begin{itemize}
+    \item Skew-$t$ distribution addresses two modeling concerns \vspace{0.5em}
+    \begin{itemize} \setlength{\itemsep}{0.5em}
       \item Extremal dependence
@@
     \end{itemize}
-    \item Our contribution is to extend the skew-$t$
-    \begin{itemize}
+    \item Our contribution is to extend the skew-$t$ \vspace{0.5em}
+    \begin{itemize} \setlength{\itemsep}{0.5em}
       \item Censoring to focus on extreme observations
@@
   \begin{itemize} \setlength{\itemsep}{1em}
-    \item Within each partition $Y(\bs)$ has the same MV skew-t distribution as before
+    \item Within each partition, $Y(\bs)$ has the same MV skew-t distribution as before
     \item Across partitions $Y(\bs)$ are asymptotically independent, but still correlated through $v(\bs)$
@@
     \end{align*}
-    where
-    \begin{itemize}
-      \item $\pi(h)$ is the probability two sites are in the same partition
-    \end{itemize}
+    where $\pi(h)$ is the probability two sites are in the same partition
+  \end{itemize}
+\end{frame}
+
+\begin{frame}{Proof that $\lim_{h \rightarrow \infty} \chi(h) = 0$}
+  \begin{itemize} \setlength{\itemsep}{1em}
+    \item Let $A$ be the area between two sites $\bs$ and $\bt$
+    \item Let $N(A)$ be the number of knots in $A$
+    \item Assume that $N(A) \sim \text{HPP}[\mu(A)]$, where \vspace{0.25em}
+    \begin{itemize} \setlength{\itemsep}{0.25em}
+      \item HPP$[\mu]$ is a homogeneous poisson process with intensity $\mu$
+      \item $\mu(\cdot)$ is an intensity measure defined on $A$
+      \item $\lim_{h \rightarrow \infty} \mu(A) = \infty$
+    \end{itemize}
+    \begin{align*}
+      \Pr[N(A) = k] = \frac{ \mu(A)^k \exp\{-\mu(A)\}}{k!}
+    \end{align*}
+  \end{itemize}
+\end{frame}
+
+\begin{frame}{Proof that $\lim_{h \rightarrow \infty} \chi(h) = 0$}
+  \begin{itemize} \setlength{\itemsep}{1em}
+    \item For finite $k$, $\lim_{h \rightarrow \infty} P[N(A) = k] = 0$
+    \item As $N(A)$ increases, $\pi(h)$ decreases because partition is defined by closest knots, so
+    \begin{align*}
+      \lim_{h \rightarrow \infty} \chi(h) = \lim_{h \rightarrow \infty} \pi(h) \chi_{\text{skew-}t} (h) = 0
+    \end{align*}
   \end{itemize}
@@
 \begin{frame}{Random partition skew-$t$ model}
-  \begin{itemize} \setlength{\itemsep}{1em}
-    \item This new model is called a random partition skew-$t$ model, and it has the properties we desire
-    \begin{itemize}
-      \item Marginal distribution with flexible tails
-      \begin{itemize}
+\begin{adjustwidth}{1em}{0em}
+  This new model is called a random partition skew-$t$ model, and it has all the properties we desire \vspace{1em}
+  \begin{itemize} \setlength{\itemsep}{1em}
+    \item Marginal distribution with flexible tails \vspace{0.5em}
+    \begin{itemize} \setlength{\itemsep}{0.5em}
       \item $\lambda$ term allows for asymmetry
@@
   \begin{itemize} \setlength{\itemsep}{1em}
-    \item Three main steps
-    \begin{enumerate}[1.]
+    \item Three main steps \vspace{0.5em}
+    \begin{enumerate}[1.] \setlength{\itemsep}{0.5em}
       \item Impute censored data below $T$
@@
 \begin{frame}{Simulation study}
-  \begin{itemize} \setlength{\itemsep}{1em}
-    \item 6 different data settings
-    \begin{itemize}
+\begin{adjustwidth}{1em}{0em}
+  6 different data settings \vspace{0.5em}
+  \begin{enumerate}[1.] \setlength{\itemsep}{0.5em}
     \item Gaussian, $K = 1$ partition
@@
     \item Skew-$t$, $K = 5$ partitions
-      \item Max-stable
-      \begin{itemize}
+    \item Max-stable \vspace{0.25em}
+    \begin{itemize} \setlength{\itemsep}{0.25em}
       \item Marginally: GEV$(\mu=1, \sigma=1, \xi=0.2)$
@@
     \end{itemize}
-    \end{itemize}
-  \end{itemize}
+  \end{enumerate}
+\end{adjustwidth}
 \end{frame}
@@
   \begin{itemize} \setlength{\itemsep}{1em}
-    \item 50 datasets for each setting
-    \begin{itemize}
-      \item 144 sites in $[0, 10] \times [0, 10]$
-      \begin{itemize}
+    \item 50 datasets for each setting \vspace{0.5em}
+    \begin{itemize} \setlength{\itemsep}{0.5em}
+      \item 144 sites in $[0, 10] \times [0, 10]$ \vspace{0.25em}
+      \begin{itemize} \setlength{\itemsep}{0.25em}
         \item 100 training
@@
     \end{itemize}
-    \item Model parameters
-    \begin{itemize}
+    \item Model parameters \vspace{0.5em}
+    \begin{itemize} \setlength{\itemsep}{0.5em}
       \item Spatial range: $\rho = 1$
@@
 \begin{frame}{Simulation study}
-  \begin{itemize} \setlength{\itemsep}{1em}
-    \item 5 different models fit to each data set
-    \begin{itemize}
+\begin{adjustwidth}{1em}{0em}
+  5 different models fit to each data set \vspace{0.5em}
+  \begin{enumerate}[1.] \setlength{\itemsep}{0.5em}
     \item Gaussian
@@
     \item Skew-$t$ with $K = 5$ partitions, thresholding at $q(0.80)$
-    \end{itemize}
-  \end{itemize}
+  \end{enumerate}
+\end{adjustwidth}
 \end{frame}
@@
   \begin{itemize} \setlength{\itemsep}{1em}
-    \item Brier score used to determine model that gives best fit (Gneiting and Raftery, 2007)
+    \item Brier score (Gneiting and Raftery, 2007) used to compare fits
     \item The Brier score for predicting exceedance of threshold $c$ is
@@
     \end{align*}
-    where
-    \begin{itemize}
+    where \vspace{0.25em}
+    \begin{itemize} \setlength{\itemsep}{0.25em}
       \item $y$ is a test set value
@@
 \begin{frame}{Simulation study results}
-  \begin{itemize} \setlength{\itemsep}{1em}
-    \item Key findings
-    \begin{itemize}
+\begin{adjustwidth}{1em}{0em}
+  Key findings \vspace{1em}
+  \begin{itemize} \setlength{\itemsep}{1em}
     \item Improvement over Gaussian methods when partitioning
@@
   \begin{itemize} \setlength{\itemsep}{1em}
-    \item 9 different analysis methods incorporating
-    \begin{itemize}
+    \item 9 different analysis methods incorporating \vspace{0.5em}
+    \begin{itemize} \setlength{\itemsep}{0.5em}
       \item Gaussian vs $t$ vs skew-$t$ marginal distribution
       \item $K=1, 5, 6, 7, 8, 9, 10, 15$ partitions
-      \item 4 threshold levels
-      \begin{itemize}
+      \item 4 threshold levels \vspace{0.25em}
+      \begin{itemize} \setlength{\itemsep}{0.25em}
          \item $T = 0$
@@
     % \includegraphics[width=0.47\linewidth]{./plots/pot/bs-ozone-2.pdf} \\
-    \caption{Relative Brier score results ($K = 6$ -- $9$ are similar to $K = 5$ and $K = 10$)}
+    \caption{Relative Brier score results ($K = 6, \ldots, 9$ are similar to $K = 5, 10$)}
   \end{figure}
@@
 \begin{frame}{Cross-validation results}
-  \begin{itemize} \setlength{\itemsep}{1em}
-    \item Key findings
-    \begin{itemize}
+\begin{adjustwidth}{1em}{0em}
+  Key findings \vspace{1em}
+  \begin{itemize} \setlength{\itemsep}{1em}
     \item Partitioning improves performance across all high thresholds
@@
     \item Improvement of model performance when using partitioned models
-    \item Thresholding makes results worse
-    \begin{itemize}
+    \item Thresholding makes results worse \vspace{0.5em}
+    \begin{itemize} \setlength{\itemsep}{0.5em}
       \item Possible numerical instability due to truncated normal distribution
@@
   \begin{itemize} \setlength{\itemsep}{1em}
-    \item Different partition structure
-    \begin{itemize}
+    \item Different partition structure \vspace{0.5em}
+    \begin{itemize} \setlength{\itemsep}{0.5em}
       \item Distance weighting for each knot vs indicator functions
     \end{itemize}
-    \item Knot selection
-    \begin{itemize}
+    \item Knot selection \vspace{0.5em}
+    \begin{itemize} \setlength{\itemsep}{0.5em}
       \item Possible prior on the probability a knot is in the spatial domain
@@
       where $t$ denotes the day of each observation
-    \item Different ways to incorporate the temporal dependence
-    \begin{itemize}
+    \item Different ways to incorporate the temporal dependence \vspace{0.5em}
+    \begin{itemize} \setlength{\itemsep}{0.5em}
       \item Time series on $\bw_t$, $z_t(\bs)$, and $\sigma_t(\bs)$
@@
   \begin{itemize} \setlength{\itemsep}{1em}
-    \item Motivation
-    \begin{itemize}
+    \item Motivation \vspace{0.5em}
+    \begin{itemize} \setlength{\itemsep}{0.5em}
       \item Want to incorporate spatial dependence when modeling rare events (e.g. Diseased trees, Disease outbreak, Crimes)
@@
 \begin{frame}{Rare binary regression}
-  \begin{itemize} \setlength{\itemsep}{1em}
-    \item Common examples with non-spatial analysis
-    \begin{itemize}
+\begin{adjustwidth}{1em}{0em}
+  Common examples with non-spatial analysis \vspace{1em}
+  \begin{itemize} \setlength{\itemsep}{1em}
     \item Logistic regression
@@
     \end{align*}
-    \item Link function allows for greater positive skew than existing methods
-    \begin{itemize}
+    \item Link function allows for greater positive skew than existing methods \vspace{0.5em}
+    \begin{itemize} \setlength{\itemsep}{0.5em}
       \item When $\xi = 0$, the link is the Cloglog link
@@
     \item We propose to develop a spatial model
-    \item Objectives are to borrow strength across sites to estimate covariate effects and spatial prediction
-    \item Proposed method will
-    \begin{itemize}
+    \item Objectives are spatial prediction and to borrow strength across sites to estimate covariate effects
+    \item Proposed method will \vspace{0.5em}
+    \begin{itemize} \setlength{\itemsep}{0.5em}
       \item Use the GEV link function
@@
   \begin{itemize} \setlength{\itemsep}{1em}
-    \item We model $Y_i = I(Z_i > 0)$ where $z_i\sim$ multivariate GEV (mGEV)  is a latent variable
+    \item We model $Y_i = I(Z_i > 0)$ where $Z_i\sim$ multivariate GEV (mGEV)  is a latent variable
     \item Hierarchical model for mGEV (Reich and Shaby, 2012)
-    \begin{itemize}
-      \item $Z(\bs) = U(\bs) \theta(\bs)$
-      \begin{itemize}
+    \begin{align*}
+      Z(\bs) = U(\bs) \theta(\bs)
+    \end{align*}
+    where
+    \begin{itemize} \setlength{\itemsep}{0.25em}
       \item $U(\bs) \iid$ GEV(1, $\alpha$, $\alpha$) is a nugget effect
@@
       \item $\alpha \in (0, 1)$ controls strength of nugget relative to spatial dependence
-        \item $A_l \iid$ Positive Stable($\alpha$) is an random effect representing the intensity
+      \item $A_l \iid$ Positive Stable($\alpha$) is a random effect representing the intensity
       \item $w_l(\bs)$ gives the weight of the intensity of the $l$th random effect on site $\bs$
-      \end{itemize}
     \end{itemize}
@@
     where
-    \begin{itemize}
+    \begin{itemize} \setlength{\itemsep}{0.25em}
       \item $w_l$ is a weighting function subject to the constraint that $\sum_{l = 1}^L w_l = 1$
@@
 \begin{frame}{Weighting function}
-  \begin{itemize} \setlength{\itemsep}{1em}
-    \item We use the Gaussian weights proposed by Reich and Shaby (2012) given by
+\begin{adjustwidth}{1em}{0em}
+  We use the Gaussian weights proposed by Reich and Shaby (2012) given by
   \footnotesize{
@@
   where
-    \begin{itemize}
+  \begin{itemize} \setlength{\itemsep}{0.25em}
     \item $\bv_l$ are spatial knots
@@
   \end{itemize}
-  \end{itemize}
-\end{frame}
-
-\begin{frame}{Illustrating asymmetric logistic dependence in one-dimension}
+\end{adjustwidth}
+\end{frame}
+
+\begin{frame}{Illustrating asymmetric logistic dependence}
   \centering
@@
     \end{itemize}
-    \item Then for $K = 0, 1, 2$
+    \item For small $K$, we can evaluate the likelihood directly
+    \item For large $K$, we use the hierarchical model of Reich and Shaby (2012)
+  \end{itemize}
+\end{frame}
+
+\begin{frame}{Joint likelihood: $K$ small}
+  \begin{itemize} \setlength{\itemsep}{1em}
+    \item For $K = 0, 1, 2$
     {\scriptsize
@@
 
-\begin{frame}{Joint likelihood}
-  \begin{itemize} \setlength{\itemsep}{1em}
-    \item For small $K$, we can evaluate the likelihood directly
-    \item For large $K$, we use the hierarchical model of Reich and Shaby (2012)
-  \end{itemize}
+\begin{frame}{Joint likelihood: $K$ large}
+\begin{adjustwidth}{1em}{0em}
+  Hierarchical model: If $Z(\bs) \sim $ mGEV with marginal distribution GEV($\mu, \sigma, \xi$), then
+  \begin{align*}
+    Z(\bs) \mid A_1, \ldots, A_L &\ind \text{GEV}[\mu^*, \sigma^*, \xi^*] \\
+    A_l &\iid \text{Positive Stable}(\alpha)
+  \end{align*}
+  where
+  \begin{itemize} \setlength{\itemsep}{0.25em}
+    \item $\mu^* = \mu + \frac{ \sigma }{ \xi }[\theta(\bs)^\xi - 1]$
+    \item $\sigma^* = \alpha \sigma \theta(\bs)^\xi$
+    \item $\xi^* = \alpha \xi$
+  \end{itemize}
+\end{adjustwidth}
 \end{frame}
@@
   \begin{itemize} \setlength{\itemsep}{1em}
-    \item Simulation study
-    \begin{itemize}
-      \item Data generated using logistic, Cloglog, and GEV links
+    \item Simulation study \vspace{0.5em}
+    \begin{itemize} \setlength{\itemsep}{0.5em}
+      \item Data generated using logistic, Cloglog, and GEV links \vspace{0.25em}
       \begin{itemize}
@@
       \end{itemize}
-      \item Models fit using
-      \begin{itemize}
+      \item Models fit using \vspace{0.25em}
+      \begin{itemize} \setlength{\itemsep}{0.25em}
         \item mGEV
@@
     \end{itemize}
-    \item Data application: Modeling crime data
+    \item Data application: Modeling crime data \vspace{0.5em}
     \begin{itemize}
@@
 \begin{frame}{Non-stationary covariance for extreme values}
-  \begin{itemize} \setlength{\itemsep}{1em}
+  \begin{itemize} \setlength{\itemsep}{0.5em}
     \item In extremes, stationary extremal dependence means
@@
     \item Semiparamtric approach using spectral density ratios (de Carvalho and Davison, 2014)
-    \begin{itemize}
+    \begin{itemize} \setlength{\itemsep}{0.25em}
       \item Vector of observations can be transformed to psuedo-polar coordinates
@@
     \item New approach extending Reich and Shaby (2012)
-    \begin{itemize}
+    \begin{itemize} \setlength{\itemsep}{0.25em}
       \item Current model uses a single bandwidth term $\rho$ for all knots
@@
 
-% \begin{frame}{References}
-%   \begin{itemize} \setlength{\itemsep}{1em}
-%     \item Demarta, S. and McNeil, A. J. (2007) The $t$ copula and related copulas. {\it International Statistical Review}, {\bf 73}, 111--129.
-%     \item Huser, R. and Davison, A. C. (2014) Space-time modelling of extreme events. {\it Journal of the Royal Statistical Society: Series B (Statistical Methodology)}, {\bf 76}, 439--461.
-%     \item Padoan, S. A. (2011) Multivariate extreme models based on underlying skew-$t$ and skew-normal distributions. {\it Journal of Multivariate Analysis}, {\bf 102}, 977--991.
-%     \item Zhang, H. and El-Shaarawi, A. (2010) On spatial skew-Gaussian processes and applications. {\it Environmetrics}, {\bf 21}, 33--47.
-%   \end{itemize}
-% \end{frame}
+\begin{frame}[allowframebreaks]{References}
+  \begin{itemize} \setlength{\itemsep}{1em}
+    \item Demarta, S. and McNeil, A. J. (2007) The $t$ copula and related copulas. {\it International Statistical Review}, {\bf 73}, 111--129.
+    \item Huser, R. and Davison, A. C. (2014) Space-time modelling of extreme events. {\it Journal of the Royal Statistical Society: Series B (Statistical Methodology)}, {\bf 76}, 439--461.
+    \item Padoan, S. A. (2011) Multivariate extreme models based on underlying skew-$t$ and skew-normal distributions. {\it Journal of Multivariate Analysis}, {\bf 102}, 977--991.
+    \item Zhang, H. and El-Shaarawi, A. (2010) On spatial skew-Gaussian processes and applications. {\it Environmetrics}, {\bf 21}, 33--47.
+  \end{itemize}
+\end{frame}
 

### 2015-01-09 prelim/oral_prelim.tex
revisions and plots for prelim

@@
     \item Estimating the probability of rare events is challenging because these events are, by definition, rare
-    \item Spatial extremes is promising because it borrows informaation across space
+    \item Spatial extremes is promising because it borrows information across space
     \item Spatial extremes is also useful for estimating probability of extremes at sites without data
@@
       \item Detect changes in extremes over time
-      \item Study meterological conditions that lead to extreme events
+      \item Study meteorological conditions that lead to extreme events
     \end{itemize}
@@
 \begin{frame}{Non-spatial analysis: Block maxima}
+\begin{adjustwidth}{1em}{0em}
+  Fisher-Tippett-Gnedenko \vspace{1em}
   \begin{itemize} \setlength{\itemsep}{1em}
@@
     \end{align*}
-    then $G(z)$ follows a generalized extreme value distribution (GEV) (Falk et al., 2011)
+    then $G(z)$ follows a generalized extreme value distribution (GEV) (Gnedenko, 1943)
     \item This motivates the use of the GEV for block maximum data
   \end{itemize}
+\end{adjustwidth}
 \end{frame}
@@
 \begin{frame}{Non-spatial analysis: Peaks over threshold}
-  \begin{itemize} \setlength{\itemsep}{1em}
-    \item Let $X_1, \ldots, X_n \sim F$
-    \item If there exist normalizing sequences $a_t > 0$ and $b_t \in \calR$ such that for any $x \ge 0$, as $T \rightarrow \infty$
-    \begin{align*}
-      \Pr\left(\frac{X - b_t}{a_t} > x \mid X > T\right) \converged H(x),
+\begin{adjustwidth}{1em}{0em}
+  Pickands-Balkema-de Haan theorem \vspace{1em}
+  \begin{itemize} \setlength{\itemsep}{1em}
+    \item Let $X_1, \ldots, X_n \iid F$
+    \item If there exist normalizing sequences $a_T > 0$ and $b_T \in \calR$ such that for any $x \ge 0$, as $T \rightarrow \infty$
+    \begin{align*}
+      \Pr\left(\frac{X - b_T}{a_T} > x \mid X > T\right) \converged H(x),
     \end{align*}
@@
   \end{itemize}
+\end{adjustwidth}
 \end{frame}
@@
     \begin{align*}
-      a_J^{-1}(\bs) \left\{ M_J(\bs) - b_J(\bs) \right\} \converged Y(\bs)
-    \end{align*}
-    which has a non-degenerate distribution, then $Y(\bs)$ is a max-stable process
+      \frac{M_J(\bs) - b_J(\bs)}{a_J(\bs)} \converged G(\bs)
+    \end{align*}
+    then $G(\bs)$ is a max-stable process (Smith, 1990)
     \item Therefore, max-stable processes are the standard model for block maxima
@@
     \begin{align*}
-      \Pr(\bZ \le \bz)  &= G^*(\bz) = \exp(-V(\bz))\\
+      \Pr(\bZ \le \bz)  &= G^*(\bz) = \exp[-V(\bz)]\\
             V(\bz)    &= d \int_{\Delta_d} \bigvee_{i = 1}^d \frac{w_i}{z_i} H(\ddd w)
@@
     \begin{itemize} \setlength{\itemsep}{0.25em}
+      \item $V(\bz)$ is called the exponent measure
       \item $\Delta_d = \{ \bw \in \calR^d_+ \mid w_1 + \cdots + w_d = 1\}$
@@
   \begin{itemize} \setlength{\itemsep}{1em}
-    \item Only a few closed-form expressions for $V(\bz)$ exist (Stephenson, 2003)
-    \item Two common forms for $V(\bz)$:
-    \begin{itemize}
-      \item Symmetric logistic
+    \item Only a few closed-form expressions for $V(\bz)$ exist
+    \item Two common forms for $V(\bz)$
+    \begin{itemize}
+      \item Symmetric logistic (Gumbel, 1960)
       \begin{align*}
@@
       \end{align*}
-      \item Asymmetric logistic
+      \item Asymmetric logistic (Coles and Tawn, 1991)
       \begin{align*}
@@
     \end{itemize}
-    \item Extreme value analysis focuses on the $\chi$ statistic, a measure of extremal dependence given by
+    \item Extreme value analysis focuses on the $\chi$ statistic (Coles et al., 1999), a measure of extremal dependence given by
     \begin{align*}
@@
     \item A spatio-temporal model with flexible tails, asymptotic spatial dependence, and computation on the order of Gaussian models for large space-time datasets
-    \item Predicting rare binary events with a spatially dependent generalized extreme-value link function
-    \item A Bayesian hierarchical model to allow for non-stationary covariance in extreme value models
+    \item Predicting rare binary events with a spatially dependent generalized extreme value link function
+    \item A Bayesian hierarchical model to allow for non-stationary dependence in extreme value models
   \end{enumerate}
@@
   \begin{itemize} \setlength{\itemsep}{1em}
-    \item In geostatistics, $Y(\bs)$ are often modeled using a Gaussian process with mean function $\mu(\bs)$ and covariance function $\rho(h)$.
+    \item In geostatistics, $Y(\bs)$ are often modeled using a Gaussian process with mean function $\mu(\bs)$ and covariance function $\rho(h)$
     \item Model properties
@@
     \item $z \sim N(0, \sigma^2)$ is a random effect
-    \item $v(\bs)$ is a Gaussian process with variance $\sigma^2$ and \Matern correlation
+    \item $v(\bs)$ is a Gaussian process with variance $\sigma^2$, \Matern correlation, and $\gamma$ is the proportion of spatial variation in the \Matern correlation
     \item $\sigma^2 \sim \text{IG}(a, b)$
@@
   \begin{align*}
-    Y(\bs) \sim \text{skew-t}(\bX(\bs), \bOmega, \alpha, \text{df}=2a)
+    Y(\bs) \sim \text{skew-t}(\bX(\bs) \bbeta, \bOmega, \alpha, \text{df}=2a)
   \end{align*}
@@
     \item $\bar{\bOmega} = (\bSigma + \lambda^2 \bOne \bOne^T)$
-    \item $\bSigma$ is a postive definite correlation matrix
+    \item $\bSigma$ is a positive definite correlation matrix
     \item $\alpha = \lambda( 1 + \lambda^2 \bOne^T \bSigma^{-1} \bOne)^{-1/2} \bOne^T \bSigma^{-1}$ controls the skewness
@@
       \end{itemize}
-      \item For a skew-$t$ distribution $\lim_{c \rightarrow \infty} \chi(h) > 0$ (Padoan, 2011)
       \item Computation that is on the order of Gaussian computation
     \end{itemize}
-    \item Challenge: Long-range dependence occurs because all observations (near and far) share the same $z$ and $\sigma^2$ (Padoan, 2011)
-    \begin{align*}
-      \lim_{h \rightarrow \infty} \chi(h) > 0
-    \end{align*}
+    \item Challenge: For a skew-$t$ distribution $\lim_{h \rightarrow \infty} \chi(h) > 0$ (Padoan, 2011) \vspace{-1em}
+    \begin{itemize}
+      \item Long-range dependence occurs because all observations (near and far) share the same $z$ and $\sigma^2$
+    \end{itemize}
   \end{itemize}
@@
     \begin{itemize} \setlength{\itemsep}{0.25em}
-      \item HPP$[\mu]$ is a homogeneous poisson process with intensity $\mu$
-      \item $\mu(\cdot)$ is an intensity measure defined on $A$
+      \item HPP$[\mu(A)]$ is a homogeneous Poisson process with intensity measure $\mu(A)$ defined on $A$
       \item $\lim_{h \rightarrow \infty} \mu(A) = \infty$
@@
       \item Impute censored data below $T$
-      \item Update parameters with standard random walk Metropolis Hastings or Gibbs sampling
+      \item Update parameters with Metropolis Hastings or Gibbs sampling
       \item Make spatial predictions
@@
       \item Degrees of freedom: 6 for $t$ distributions
+      \item Proportion of spatial variation: $\gamma = 0.9$
     \end{itemize}
@@
     \item Brier score (Gneiting and Raftery, 2007) used to compare fits
+    \begin{itemize} \setlength{\itemsep}{0.5em}
+      \item Lower scores indicate better fits
+    \end{itemize}
     \item The Brier score for predicting exceedance of threshold $c$ is
@@
   \begin{itemize} \setlength{\itemsep}{1em}
-    \item 9 different analysis methods incorporating \vspace{0.5em}
-    \begin{itemize} \setlength{\itemsep}{0.5em}
-      \item Gaussian vs $t$ vs skew-$t$ marginal distribution
+    \item Many different analysis methods incorporating \vspace{0.5em}
+    \begin{itemize} \setlength{\itemsep}{0.5em}
+      \item Gaussian, symmetric-$t$, skew-$t$, and max-stable marginal distributions
       \item $K=1, 5, 6, 7, 8, 9, 10, 15$ partitions
-      \item 4 threshold levels \vspace{0.25em}
+      \item 4 threshold levels for $t$ marginals \vspace{0.25em}
       \begin{itemize} \setlength{\itemsep}{0.25em}
@@
       \end{itemize}
-    \end{itemize}
-    \item Compare Brier scores using two-fold cross validation
+      \item Thresholded at $T = 75$ for max-stable
+    \end{itemize}
+    \item Compare Brier scores from two-fold cross validation
   \end{itemize}
@@
     \caption{Relative Brier score results ($K = 6, \ldots, 9$ are similar to $K = 5, 10$)}
+  \end{figure}
+\end{frame}
+
+\begin{frame}{Probability of exceedance}
+  \centering
+  \begin{figure}
+    \includegraphics[width=0.33\linewidth, trim=0.2in 0 0.2in 1in]{./plots/pot/ozone-p2-gaus.pdf}
+    \includegraphics[width=0.33\linewidth, trim=0.2in 0 0.2in 1in]{./plots/pot/ozone-p2-t1.pdf}
+    \includegraphics[width=0.33\linewidth, trim=0.2in 0 0.2in 1in]{./plots/pot/ozone-p2-t6.pdf}
+    \caption{Probability of exceeding 75ppb at least two days for Gaussian (left), skew-$t$ with 1 partition (center), skew-$t$ with 6 partitions (right)}
+  \end{figure}
+\end{frame}
+
+\begin{frame}{Quantile plots}
+  \centering
+  \begin{figure}
+    \includegraphics[width=0.33\linewidth, trim=0.2in 0 0.2in 1in]{./plots/pot/ozone-q95-gaus.pdf}
+    \includegraphics[width=0.33\linewidth, trim=0.2in 0 0.2in 1in]{./plots/pot/ozone-q95-t1.pdf}
+    \includegraphics[width=0.33\linewidth, trim=0.2in 0 0.2in 1in]{./plots/pot/ozone-q95-t6.pdf}
+    \caption{$\widehat{q}(0.95)$ for Gaussian (left), skew-$t$ with 1 partition (center), skew-$t$ with 6 partitions (right)}
+  \end{figure}
+\end{frame}
+
+\begin{frame}{Quantile plots}
+  \centering
+  \begin{figure}
+    \includegraphics[width=0.33\linewidth, trim=0.2in 0 0.2in 1in]{./plots/pot/ozone-q99-gaus.pdf}
+    \includegraphics[width=0.33\linewidth, trim=0.2in 0 0.2in 1in]{./plots/pot/ozone-q99-t1.pdf}
+    \includegraphics[width=0.33\linewidth, trim=0.2in 0 0.2in 1in]{./plots/pot/ozone-q99-t6.pdf}
+    \caption{$\widehat{q}(0.99)$ for Gaussian (left), skew-$t$ with 1 partition (center), skew-$t$ with 6 partitions (right)}
   \end{figure}
@@
     where $\Phi$ is the standard normal distribution function
-    \item Cloglog regression
-    \begin{align*}
-      \Pr[Y_i = 1] = 1 - \exp[ -\exp(\bX_i \bbeta)]
-    \end{align*}
+    % \item Cloglog regression
+    % \begin{align*}
+    %   \Pr[Y_i = 1] = 1 - \exp[ -\exp(\bX_i \bbeta)]
+    % \end{align*}
   \end{itemize}
@@
       \item $\theta(\bs) = [\sum_{l = 1}^L A_l w_l(\bs)^{1/\alpha}]^\alpha$ is the spatial process
-      \item $\alpha \in (0, 1)$ controls strength of nugget relative to spatial dependence
       \item $A_l \iid$ Positive Stable($\alpha$) is a random effect representing the intensity
       \item $w_l(\bs)$ gives the weight of the intensity of the $l$th random effect on site $\bs$
+      \item $\alpha \in (0, 1)$ controls strength of nugget relative to spatial dependence
     \end{itemize}
@@
 % TODO: more about this model for sure
-\begin{frame}{Non-stationary covariance for extreme values}
+\begin{frame}{Non-stationary dependence for extreme values}
   \begin{itemize} \setlength{\itemsep}{1em}
@@
 
-\begin{frame}{Non-stationary covariance for extreme values}
+\begin{frame}{Non-stationary dependence for extreme values}
   \begin{itemize} \setlength{\itemsep}{0.5em}
@@
     \item Currently, there are no methods to model non-stationarity in spatial extremes
-    \item Semiparamtric approach using spectral density ratios (de Carvalho and Davison, 2014)
-    \begin{itemize} \setlength{\itemsep}{0.25em}
-      \item Vector of observations can be transformed to psuedo-polar coordinates
+    \item Semiparametric approach using spectral density ratios (de Carvalho and Davison, 2014)
+    \begin{itemize} \setlength{\itemsep}{0.25em}
+      \item Vector of observations can be transformed to pseudo-polar coordinates
       \item Pairwise analysis
@@
     \item Chapter 3: Rare spatial binary regression \alert{May / June 2015}
-    \item Chapter 4: Non-stationary covariance through knot-specific bandwidth \alert{August 2015}
+    \item Chapter 4: Non-stationary extremal dependence through knot-specific bandwidth \alert{August 2015}
   \end{itemize}
@@
   \begin{itemize} \setlength{\itemsep}{1em}
-    \item Demarta, S. and McNeil, A. J. (2007) The $t$ copula and related copulas. {\it International Statistical Review}, {\bf 73}, 111--129.
-    \item Huser, R. and Davison, A. C. (2014) Space-time modelling of extreme events. {\it Journal of the Royal Statistical Society: Series B (Statistical Methodology)}, {\bf 76}, 439--461.
-    \item Padoan, S. A. (2011) Multivariate extreme models based on underlying skew-$t$ and skew-normal distributions. {\it Journal of Multivariate Analysis}, {\bf 102}, 977--991.
-    \item Zhang, H. and El-Shaarawi, A. (2010) On spatial skew-Gaussian processes and applications. {\it Environmetrics}, {\bf 21}, 33--47.
+    \item Balkema, A. A. and de Haan, L. (1974). Residual life time at great age, {\it Annals of Probability}, {\bf 2}, 792--804.
+    \item de Carvalho, M. and Davison, A. C. (2014). Spectral density ratio models for multivariate extremes, {\it Journal of the American Statistical Association}, {\bf 109}, 764--776.
+    \item Coles, S. G., Heffernan, J., and Tawn, J. A. (1999). Dependence measures for extreme value analysis, {\it Extremes}, {\bf 2}, 339--365.
+    \item Coles S. G. and Tawn, J. A. (1991). Modelling extreme multivariate events. {\it Journal of the Royal Statistical Society: Series B (Methodological)}, {\bf 53}, 337--392.
+    \item Demarta, S. and McNeil, A. J. (2007). The $t$ copula and related copulas. {\it International Statistical Review}, {\bf 73}, 111--129.
+    \item Falk. M., H\"{u}sler, J., and Reiss, R. D. (2010). {\it Laws of Small Numbers: Extremes and Rare Events}. Springer Basel.
+    \item Kim, H.-M., Mallik, B.
+K., and Holmes, C.
+C.
+(2005).
+Analyzing nonstationary spatial data using piecewise Gaussian processes.
+{\it Journal of the American Statistical Association}, {\bf 100}, 653--668.
+    \item Gnedenko, B. (1943). Sur la distribution limite du terme maximum d'une s\'{e}rie al\'{e}atoire. {\it Annals of Mathematics}, {\bf 44}, 423--453.
+    \item Gneiting, T. and Raftery, A. (2007). Strictly proper scoring rules, prediction, and estimation. {\it Journal of the American Statistical Association}, {\bf 102}, 359--378.
+    \item Gumbel, E. J. (1960). Multivariate extremal distributions. {\it Bulletin de l'Institut International de Statistique}, {\bf 37}, 471--475
+    \item Huser, R. and Davison, A. C. (2014). Space-time modelling of extreme events. {\it Journal of the Royal Statistical Society: Series B (Statistical Methodology)}, {\bf 76}, 439--461.
+    \item Padoan, S. A. (2011). Multivariate extreme models based on underlying skew-$t$ and skew-normal distributions. {\it Journal of Multivariate Analysis}, {\bf 102}, 977--991.
+    \item Reich, B. J. and Shaby, B. A. (2012). A hierarchical max-stable spatial model for extreme precipitation. {\it Annals of Applied Statistics}, {\bf 6}, 1430--1451.
+    \item Smith R. L. (1990). Max-stable processes and spatial extremes, unpublished manuscript.
+    \item Tawn, J. A. (1990). Modelling multivariate extreme value distributions. {\it Biometrika}, {\bf 77}, 245--253.
+    \item Zhang, H. and El-Shaarawi, A. (2010). On spatial skew-Gaussian processes and applications. {\it Environmetrics}, {\bf 21}, 33--47.
   \end{itemize}

### 2015-01-09 pot/code/analysis/ozone/US-all/us-all-results.R

@@ -18,7 +18,7 @@ phi.w <- array(NA, dim=c(5000, nsets, 24))
 phi.tau <- array(NA, dim=c(5000, nsets, 24))
 
 load("us-all-results.RData")
-done <- c(43:50)
+done <- 26
 quant.score <- savelist[[1]]
 brier.score <- savelist[[2]]
 beta.0 <- savelist[[3]]
@@ -36,12 +36,13 @@ for (i in 1:50) {
       val.idx <- cv.lst[[d]]
       validate <- Y[val.idx, ]
       pred.d <- fit.d$yp[, , ]
-      if (i == 26) {
+      if (i == 26 && d == 1) {
         validate <- t(validate)
         pred.d <- fit.d$yp[(25001:30000), , ]
-      }
+
       quant.score[, d, i] <- QuantScore(pred.d, probs, validate)
       brier.score[, d, i] <- BrierScore(pred.d, thresholds, validate)
+    }
       if (i != 26) {
         beta.0[, d, i] <- fit.d$beta[, 1]
         beta.1[, d, i] <- fit.d$beta[, 2]
@@ -76,12 +77,12 @@ brier.score.mean <- matrix(NA, 50, length(thresholds))
 quant.score.se <- matrix(NA, 50, length(probs))
 brier.score.se <- matrix(NA, 50, length(thresholds))
 
-done <- c(1:25, 27:50)
+done <- c(1:50)
 for (i in 1:50) {
   if (i %in% done) {
-    quant.score.mean[i, ] <- apply(quant.score[, , i], 1, mean)
+    quant.score.mean[i, ] <- apply(quant.score[, , i], 1, mean, na.rm=T)
     quant.score.se[i, ] <- apply(quant.score[, , i], 1, sd) / sqrt(2)
-    brier.score.mean[i, ] <- apply(brier.score[, , i], 1, mean)
+    brier.score.mean[i, ] <- apply(brier.score[, , i], 1, mean, na.rm=T)
     brier.score.se[i, ] <- apply(brier.score[, , i], 1, sd) / sqrt(2)
   }
 }
[…]

### 2015-01-12 prelim/oral_prelim.tex
update to test out truncating below

@@
 \begin{adjustwidth}{1em}{0em}
-  Fisher-Tippett-Gnedenko \vspace{1em}
+  Fisher-Tippett-Gnedenko theorem \vspace{1em}
   \begin{itemize} \setlength{\itemsep}{1em}
@@
     \item Marginally at each site, observations follow a GEV distribution
-    \item For a finite collection of sites the multivariate representation for the GEV (mGEV) is
+    \item For a finite collection of sites the representation for the multivariate GEV (mGEV) is
     \begin{align*}
@@
   \begin{align*}
-    Y(\bs) \sim \text{skew-t}(\bX(\bs) \bbeta, \bOmega, \alpha, \text{df}=2a)
+    Y(\bs) \sim \text{skew-t}[\bX(\bs) \bbeta, \bOmega, \alpha, \text{df}=2a]
   \end{align*}
@@
   \begin{itemize} \setlength{\itemsep}{1em}
-    \item Skew-$t$ distribution addresses two modeling concerns \vspace{0.5em}
-    \begin{itemize} \setlength{\itemsep}{0.5em}
+    \item Skew-$t$ distribution addresses three modeling concerns \vspace{0.5em}
+    \begin{itemize} \setlength{\itemsep}{0.5em}
+      \item Flexible tails
       \item Extremal dependence
@@
       \end{itemize}
+      \item 50 days
     \end{itemize}
@@
       \item Skew parameter: $\lambda = 3$
-      \item Degrees of freedom: 6 for $t$ distributions
+      \item Degrees of freedom: $2a = 6$
       \item Proportion of spatial variation: $\gamma = 0.9$
@@
 
-\begin{frame}{Simulation study results}
-\centering
-  \begin{figure}
-    \includegraphics[width=0.45\linewidth]{./plots/pot/bs-sim-t1.pdf}
-    \includegraphics[width=0.45\linewidth]{./plots/pot/bs-sim-st1.pdf} \\
-    \includegraphics[width=0.3\linewidth, trim=2in 3.25in 2in 3in]{./plots/pot/bs-sim-legend.pdf}
-    \caption{Relative Brier score results}
-  \end{figure}
-\end{frame}
+% \begin{frame}{Simulation study results}
+% \centering
+%   \begin{figure}
+%     \includegraphics[width=0.45\linewidth]{./plots/pot/bs-sim-t1.pdf}
+%     \includegraphics[width=0.45\linewidth]{./plots/pot/bs-sim-st1.pdf} \\
+%     \includegraphics[width=0.3\linewidth, trim=2in 3.25in 2in 3in]{./plots/pot/bs-sim-legend.pdf}
+%     \caption{Relative Brier score results}
+%   \end{figure}
+% \end{frame}
 
@@
          \item $T = 0$
-         \item $T = 50$ppb, $q(0.48)$
-         \item $T = 75$ppb, $q(0.92)$
-         \item $T = 85$ppb, $q(0.97)$
+         \item $T = 50$ ppb, $q(0.48)$
+         \item $T = 75$ ppb, $q(0.92)$
+         \item $T = 85$ ppb, $q(0.97)$
       \end{itemize}
@@
 
-\begin{frame}{Probability of exceedance}
+\begin{frame}{Probability of exceedance (2 days)}
   \centering
@@
 
-\begin{frame}{Quantile plots}
+\begin{frame}{Quantile plots: $\widehat{q}(0.95)$}
   \centering
@@
 
-\begin{frame}{Quantile plots}
+\begin{frame}{Quantile plots: $\widehat{q}(0.99)$}
   \centering
@@
   \begin{itemize} \setlength{\itemsep}{1em}
-    \item We model $Y_i = I(Z_i > 0)$ where $Z_i\sim$ multivariate GEV (mGEV)  is a latent variable
+    \item We model $Y_i = I(Z_i > 0)$ where $Z_i\sim$ mGEV is a latent variable
     \item Hierarchical model for mGEV (Reich and Shaby, 2012)
@@
     \end{itemize}
-    \item Data application: Modeling crime data \vspace{0.5em}
-    \begin{itemize}
-      \item Homicides, car theft, vandalism
-    \end{itemize}
+    \item Data application: Modeling ozone exceedances \vspace{0.5em}
   \end{itemize}

### 2015-01-12 pot/code/analysis/ozone/US-all/us-all-setup.R

@@ -1,6 +1,6 @@
 library(fields)
-# library(geoR)
 library(mvtnorm)
+library(sn)
 
 rm(list=ls())
 source('../../../R/mcmc.R', chdir=T)
@@ -288,7 +288,71 @@ xplot <- (xplot - mean(xplot))
 plot(xplot, res.qq, xlab="Theoretical Quantile", ylab="Observed Quantile", main="Q-Q plot: Skew-t with 10 d.f. and alpha = 1")
 abline(0, 1)
 
+# ML - Skew-t fit and diagnostics
+library(fields)
+library(mvtnorm)
+library(sn)
 
+rm(list=ls())
+source('../../../R/mcmc.R', chdir=T)
+source('../../../R/auxfunctions.R')
+
+# Setup from Brian
+load("../ozone_data.RData")
+S <- cbind(x[s[, 1]], y[s[, 2]])  # expands the grid of x, y locs where we have CMAQ
+
+# Exclude if site is missing more than 50% of its days
+excl  <- which(rowMeans(is.na(Y)) > 0.50)
+index <- index[-excl]
+Y     <- Y[-excl, ]
+S     <- S[-excl, ]
+cmaq  <- CMAQ[index, ]
+cmaq  <- (cmaq - mean(cmaq)) / sd(cmaq)  # center and scale CMAQ data
+
+# add in intercept
+nt <- ncol(Y)
+ns <- nrow(Y)
+X  <- array(1, dim=c(nrow(cmaq), nt, 2))
+for (t in 1:nt) {
+  X[, t, 2] <- cmaq[, t]
+}
+
+S <- S / 1000
+x <- x / 1000
+y <- y / 1000
+
+# put all covariate and observations into one long sting
+y.lm <- Y[, 1]
+x.lm <- X[, 1, ]
+for (t in 2:nt) {
+  y.lm <- c(y.lm, Y[, t])
+  x.lm <- rbind(x.lm, X[, t, c(1, 2)])
+}
+
+ozone.stlm  <- selm(y.lm ~ x.lm[, 2], family="ST")  # automatically includes int
+ozone.lm    <- lm(y.lm ~ x.lm[, 2])
+ozone.res   <- residuals(ozone.stlm)
+ozone.dp    <- ozone.stlm@param$dp
+ozone.omega <- ozone.dp[3]
+ozone.alpha <- ozone.dp[4]
+ozone.nu    <- ozone.dp[5]
+
+probs <- seq(0.000001, 0.999999, length=length(ozone.res))
[…]

### 2015-02-17 LaTeX/MVN.tex
change to file organization

--- 2014-12-11 pot/LaTeX/skewnormal.tex
@@
 
-\section{Introduction}\label{s:intro}
-In most climatological applications, researchers are interested in learning about the average behavior of different climate variables (e.g. ozone, temperature, rainfall).
-However, averages do not help regulators prepare for the unusual events that only happen once every 100 years.
-For example, it is important to have an idea of how much rain will come in a 100-year floor in order to construct strong enough river levees to protect lands from flooding.
-
-Unlike multivariate normal distributions, it is challenging to model multivariate extreme value distributions (e.g.
-generalized extreme value and generalized Pareto distribution) because few closed-form expressions exist for the density in more than two-dimensions \citep{Coles1991}.
-Given this limitation, pairwise composite likelihoods have been used when modeling dependent extremes \citep{Padoan2010,Blanchet2011,Huser2013}.
-
-One way around the multi-dimensional limitation of multivariate extreme value distributions is to use skew elliptical distributions to model dependent extreme values \citep{Genton2004,Zhang2010,Padoan2011}.
-Due to their flexibility, the skew-normal and skew-$t$ distribution offer a flexible way to handle non-symmetric data within a framework of multivariate normal and multivariate t-distributions.
-As with the spatial Gaussian process, the skew-normal distribution is also asymptotically independent; however, the skew-$t$ does demonstrate asymptotic dependence \citep{Padoan2011}.
-Although asymptotic dependence is desirable between sites that are near one another, one drawback to the skew-$t$ is that sites remain asymptotically dependent even at far distances.
-
-In this paper, we present a model that has marginal distributions with flexible tails, demonstrates asymptotic dependence for observations at sites that are near to one another, and has computation on the order of Gaussian models for large space-time datasets.
-Specifically, our contribution is to incorporate thresholding and random spatial partitions using a multivariate skew-$t$ distribution.
-The advantage of using a thresholded model as opposed to a non-thresholded model is that is allows for the tails of the distribution to inform the predictions in the tails \citep{DuMouchel1983}.
-The random spatial partition alleviates the long-range spatial dependence seen by the skew-$t$.
-
-The paper is organized as follows.
-Section \ref{s:skewt} is a brief review of the spatial skew-$t$ process.
-In section \ref{s:temporal}, we build upon the traditional skew-$t$ by incorporating censoring to focus on tails, partitioning to remove long-range asymptotic dependence, and extending the model to space-time data.
-The computing is described in section \ref{s:comp}.
-In section \ref{s:simstudy}, we present a simulation study that examines the predictive capabilities of this model compared with a na{\"i}ve Gaussian method.
-We then compare our method to Gaussian and max-stable methods with a data analysis of ozone measurements from the eastern US in section \ref{s:analysis}.
-The final section provides brief discussion and direction for future research.
-
-\section{Spatial skew processes}\label{s:spatialskew}
-Many types of data demonstrate some level of skewness and therefore should be modeled with distributions that allow for asymmetry.
-The skew-elliptical family of distributions provides models that are mathematically tractable while introducing a slant parameter to account for asymmetric data \citep{Genton2004}.
-A brief review of the additive process by which a skew-$t$ process is created is given here.
-
-\subsection{Skew-$t$ process} \label{s:skewt}
-Let $Y(\bs)$ be the observation at spatial location $\bs = (s_1, s_2)$.
-The spatial skew-$t$ process can be written
-\begin{align}
-  Y(\bs) = \bX(\bs)^T \bbeta + \lambda \sigma |z| + \sigma v(\bs)
-\end{align}
-where $\bX(\bs)$ is a set of spatial covariates at site $\bs$, $\bbeta$ is a set of regression parameters, $\lambda \in \calR$ is a parameter controlling skew, $z \sim N(0, 1)$, $\sigma^2 \sim \text{IG}(a, b)$ is an inverse gamma random variable, and $v(\bs)$ is a spatial Gaussian process with mean zero and variance one.
-A common spatial correlation function is the \Matern with
-\begin{align}
-  \text{cor}(v(\bs), v(\bt)) = \gamma I(\bs = \bt) + (1 - \gamma)  \frac{ 1 }{ \Gamma(\nu) 2^{ \nu - 1}} \left( \sqrt{2\nu} \frac{ h }{ \rho } \right)^{\nu} K_{\nu} \left( \sqrt{2\nu} \frac{ h }{ \rho } \right) \label{eq:matern}
-\end{align}
-where $\rho$ is the spatial range, $\nu$ is the smoothness, $\gamma$ is the proportion of variance accounted for by the spatial variation, $K_\nu$ is a modified Bessel function of the second kind, and $h = || \bs - \bt ||$.
-
-Let $\bY = [Y(\bs_1), \ldots, Y(\bs_n)]^T$ be a set of observations at a finite collection of locations $\bs_1, \ldots, \bs_n$.
-After marginalizing over both $z$ and $\sigma$,
-\begin{align}
-  \bY \sim \text{ST}_n(\bX \bbeta, \bOmega, \balpha, 2a), \label{eq:fullmodel}
-\end{align}
-that is, $\bY$ follows an $n$-dimensional skew-$t$ process with location $\bX \bbeta$, correlation matrix $\bOmega$, slant parameters $\balpha$ and degrees of freedom $2a$, where $\bX = [\bX(\bs_1)^T, \ldots, \bX(\bs_n)^T]$, $\bOmega = \bomega \bar{\bOmega} \bomega$, $\bomega = \text{diag}\left(\frac{1}{ \sqrt{ab}}, \ldots, \frac{1}{\sqrt{ab}}\right)$, $\bar{\bOmega} = (\bSigma + \lambda^2 \bOne \bOne^T$), $\bSigma$ is a positive definite correlation matrix, $\balpha = \lambda (1 + \lambda^2 \bOne^T \bSigma^{-1} \bOne)^{-1/2} \bOne^T \bSigma^{-1}$ is a vector of slant parameters.
-This process is desirable because of its flexible tail that is controlled by the skewness parameter $\lambda$ and degrees of freedom $2a$.
-Furthermore, the marginal distributions at each location also follow a univariate skew-$t$ distribution \citep{Azzalini2013}.
-
-\subsection{Extremal dependence}
-One measure of extremal dependence is the $\chi$ statistic \citep{Padoan2011}.
-The $\chi$ statistic for the upper tail is given by $\lim_{c \rightarrow \infty} \Pr(Y(\bs_1) > c | Y(\bs_2) > c)$.
-For a stationary spatial process, we can write the $\chi$ coefficient as
-\begin{align}
-  \chi(h) = \lim_{c \rightarrow \infty} \Pr(Y(\bs) > c | Y(\bt) > c).
-\end{align}
-If $\chi(h) = 0$, then observations are asymptotically independent at distance $h$.
-For Gaussian processes, $\chi(h) = 0$ regardless of the distance, so they are not suitable for modeling spatially-dependent extremes.
-Unlike the Gaussian process, the skew-$t$ process is asymptotically dependent.
-However, one problem with the spatial skew-$t$ process is that $\lim_{h \rightarrow \infty} \chi(h) > 0$.
-This occurs because all observations, both near and far, share the same $z$ and $\sigma$ terms.
-Therefore, the skew-$t$ process is not ideal for spatial analysis of large geographic regions where we expect only local spatial dependence.
-The explicit expression for $\chi(h)$ \citep{Padoan2011} and a proof of this are given in Appendix \ref{a:skewt}.
-
-\section{Spatiotemporal skew-$t$ model for extremes}\label{s:spatial}
-In this section, we propose extensions to the skew-$t$ process to model spatial extremes over a large geographic region by introducing censoring to focus on tail behavior and a random partition similar to \citet{Kim2005} to remove long-range asymptotic dependence.
-
-\subsection{Censoring to focus on the tails}
-To avoid bias in estimating tail parameters, we model censored data.
-Let
-\beq\label{Yt}
-  \tilde{Y}(\bs) = \left\{ \begin{array}{ll}
-      Y(\bs) \quad & \delta(\bs) = 1 \\
-      T & \delta(\bs) = 0
-  \end{array} \right.
-\eeq
-be the censored observation at site $\bs$ where $Y(\bs)$ is the uncensored observation, $\delta(\bs) = I[Y(\bs) > T]$, and $T$ is a pre-specified threshold value.
-Then, assuming the uncensored data $Y(\bs)$ are observations from a skew-$t$ process, we update values censored below the threshold using standard Bayesian missing data methods as described in Section \ref{s:comp}.
-
-\subsection{Partitioning to remove long-range asymptotic dependence}\label{s:part}
-We handle the problem of long-range asymptotic dependence with a random partition model.
-As discussed in Section \ref{s:spatialskew}, the source of long-range dependence is the shared $z$ and $\sigma$.
-Therefore, to alleviate this dependence, we allow $z$ and $\sigma$ to vary by site.
-Then, the model becomes
-\begin{align} \label{eq:partition}
-  Y(\bs) &= \bX(\bs)^T \bbeta + \lambda \sigma(\bs) |z(\bs)| + \sigma(\bs) v(\bs).
-\end{align}
-Let $\bw = (w_1, w_2)$ be the location of a spatial knot.
-To model spatial variation, consider a set of spatial knots $\bw_1, \ldots, \bw_K$ from a homogeneous Poisson process with intensity $\mu$ over spatial domain $\calD \in \calR^2$.
-The knots define a random daily partition of $\calD$ by subregions $P_{1}, \ldots, P_{K}$ defined as
-\begin{align} \label{eq:subregions}
-  P_{k} = \{ \bs : k = \argmin_\ell || \bs - \bw_{\ell} || \}.
-\end{align}
-All $z(\bs)$ and $\sigma(\bs)$ for sites in subregion $k$ are assigned common values
-\begin{align}
-  z(\bs) &= z_{k} \label{eq:sitez}\\
-  \sigma(\bs) &= \sigma_{k} \label{eq:sitesig},
-\end{align}
-and the $z_k$ and $\sigma^2_k$ are distributed as $z_k \iid N(0, 1)$ and $\sigma^2 \iid$ IG$(a, b)$.
-So, within each partition, $Y(\bs)$ follows the spatial skew-$t$ process defined in Section \ref{s:spatialskew}.
-Across partitions, the $Y(\bs)$ remain correlated via the correlation function for $v(\bs)$ because it spans the partitions.
-
-When incorporating the random daily partition, conditional on knots $\bw_1, \ldots, \bw_K$, the $\chi$ statistic for two sites $\bs$ and $\bt$ in partitions $k_s$ and $k_t$ respectively is
-\begin{align}
-  \chi(h) &= I(k_s = k_t) \chi_{\text{skew-}t}(h) + I(k_s \neq k_t) \chi_{\text{Gaus}}(h) \nonumber \\
-         &= I(k_s = k_t) \chi_{\text{skew-}t}(h)
-\end{align}
-where $\chi_{\text{skew-}t}(h)$ is the $\chi$ statistic for a skew-$t$ process, $\chi_{\text{Gaus}}(h)$ is the $\chi$ statistic for a Gaussian process, and $h = ||\bs - \bt||$.
-Therefore, sites in different subregions are asymptotically independent.
-Marginally, over the knots $\bw_1, \ldots, \bw_K$, $\chi(h) = \pi(h) \chi_{\text{skew}-t}(h)$, where $\pi(h) = \Pr(k_s = k_t)$ is the probability that two sites separated by distance $h$ are in the same partition.
-So, to show that $\lim_{h \rightarrow \infty} \chi(h) = 0$, we need only know that $\lim_{h \rightarrow \infty} \pi(h) = 0$.
-A proof of this is given in Appendix \ref{a:proofsamepartition}.
-
-Figure \ref{fig:simchi} shows $\chi(h)$ for varying data settings when $\bs \in [0, 1] \times [0, 1]$.
-The plot clearly shows that without partitioning, in a skew-$t$ process, all sites remain asymptotically dependent.
-However, upon incorporating partitioning into the model, the asymptotic dependence is only shared by sites that are close to one another.
-
-\begin{figure}
-  %\includegraphics[width=\linewidth]{plots/sim-chi.pdf}
-  \caption{$\chi(h)$ for different settings. I need to replace the plot here.}
-  \label{fig:simchi}
-\end{figure}
-
-\subsection{Extension to space-time data} \label{s:temporal}
-When using daily measurements, the assumption of temporal independence is inappropriate.
-There are several places where temporal dependence could be incorporated in the model, including the residual $v_t(\bs)$.
-However, we choose to allow for temporal dependence in the $\bw$, $z$, and $\sigma$ terms because these terms dictate the tail behavior which is our primary focus.
-In this section, we extend (\ref{eq:partition}) to the spatiotemporal setting.
-Let
-\begin{align} \label{eq:spatiotemp}
-  Y_t(\bs) = \bX_t(\bs)^T \bbeta + \lambda \sigma_t(\bs) |z_t(\bs)| + \sigma_t(\bs) v_t(\bs),
-\end{align}
-where $t \in \{1, \ldots, T\}$ denotes the day of each observation.
-Let \hbox{$\bw_{tk} = (w_{tk1}, w_{tk2})$} be a spatial knot on day $t$, and let $w_{t1}, \ldots, w_{tK}$ be a collection of spatial knots on day $t$.
-As in section \ref{s:part}, these knots define a daily partition $P_{t1}, \ldots, P_{tK}$, and for $\bs \in P_{tk}$,
-\begin{align}
-  z_t(\bs) = z_{tk}\\
-  \sigma_{t}(\bs) = \sigma_{tk}.
-\end{align}
-
-We use an AR(1) time series model for $w_{tk}$, $z_{tk}$, and $\sigma_{tk}$.
-The time series model must be specified after a transformation to preserve the skew-$t$ process at each time point.
-We first transform the spatial knots from $\calD$ to $\calR^2$ as follows.
-Let
-\begin{align}
-  w^*_{tki} = \Phi^{-1}\left[ \frac{ (w_{tki} - \min(\bs_i))}{ \text{range}(\bs_i) } \right], \quad i = 1, 2
-\end{align}
-where $\Phi$ is a univariate standard normal density function, and $\bs_i = [s_{1i}, \ldots, s_{ni}]$.
-Then $\bw^*_{tk} \in \calR^2$.
-We use a copula on the $\sigma^2_t(\bs)$ to ensure that the marginal distributions of $\sigma^2_t(\bs)$ are inverse gamma.
-Let
-\begin{align}
-  \sigma^{2*}_t(\bs) =\Phi^{-1}\left\{ \text{IG}[\sigma^2_t(\bs)] \right\}
-\end{align}
-where IG is the distribution function for an IG$(a, b)$ random variable.
-The AR(1) process for each tail parameter is $\bw^*_{1k} \sim N_w(0, 1)$, $z_{1k} \sim N(0, \sigma^2_{1k})$, $\sigma^{2*}_{1k} \sim N(0, 1)$, and for $t > 1$ the time series is modeled as
-\begin{align}
-  \bw^*_{tk} | \bw^*_{t-1, k} &\sim N_2\left[\phi_w \bw^*_{t-1, k}, (1 - \phi_w^2) \right] \\
-  z_{tk} | z_{t-1, k} &\sim N \left[\phi_z z_{t-1, k}, \sigma^2_{tk} (1 - \phi_z^2)\right] \\
-  \sigma^{2*}_{tk} | \sigma^{2*}_{t-1, k} &\sim N \left[\phi_\sigma \sigma^{2*}_{t-1, k}, (1 - \phi_\sigma^2) \right]
-\end{align}
-where $|\phi_w| < 1$, $0 < \phi_z < 1$, and $|\phi_\sigma| < 1$.
-We restrict $\phi_z$ to be positive because in the model, we assume $Y(\bs)$ is a function of $|z_t(\bs)|$.
-These are stationary time series models with marginal distributions $\bw^*_{k} \sim N_2(0, 1)$, $z_{k} \sim N(0, \sigma^2_{k})$, $\sigma^{2*}_{k} \sim N(0, 1)$.
-After transformation back to the original space, $\bw_{tk} \sim \text{Unif}(\calD)$, $\sigma^2_{tk} \sim \text{IG}(a, b)$.
-For each day, the model is identical to the spatial-only model in (\ref{eq:partition}) by construction.
-
-\subsection{Hierarchical model}\label{s:hier}
-Conditioned on $z_{tk}(\bs)$, $\sigma^2_{tk}(\bs)$, and $P_{tk}$, the marginal distributions are Gaussian and the joint distribution multivariate Gaussian.
-However, we do not fix the partitions, they are treated as unknown and updated in the MCMC.
-We model this with a Bayesian hierarchical model as follows.
-Let $\bw_{t1}, \ldots, \bw_{tK}$ be a set of daily spatial knots in a spatial domain of interest, $\calD$, and $P_{tk}$ as defined in (\ref{eq:subregions}).
-
-Then
-\begin{align}
-   Y_t(\bs) \mid z_{t}(\bs), \sigma^2(\bs), P_{tk}, \alpha, \beta, \Theta &= \bX_t(\bs)^T \beta + \lambda |z_t(\bs)| + \sigma_t(\bs) v_t(\bs) \label{eq:hier}\\
-   z_t(\bs) &= z_{tk} \text{ if } \bs \in P_{tk}\\
-   \sigma^2_{t}(\bs) &= \sigma^2_{tk} \text{ if } \bs \in P_{tk}\\
-   \lambda &\sim N(0, 10)\\
-   v_t(\bs) \mid \Theta &\sim \Matern(0, \Sigma)\\
-   z_{tk} \mid z_{t-1, k}, \sigma^2_{tk} &\sim N(\phi_z z_{t-1, k}, \sigma^2_{tk} (1 - \phi_z^2))\\
-   \sigma^{2*}_{tk} \mid \sigma^{2*}_{t-1, k} &\sim N(\phi_\sigma \sigma^{2*}_{t-1, k}, (1 - \phi_\sigma^2))\\
-   \bw^*_{tk} \mid \bw^*_{t-1, k} &\sim N_2(\phi_w \bw^*_{t-1, k}, (1 - \phi_w^2))
-\end{align}
-where $\Theta = \{\rho, \nu, \gamma\}$, and $\Sigma$ is a \Matern covariance matrix as described in Section \ref{s:skewt}.
-
-\section{Computation}\label{s:comp}
-First, we impute values below the threshold.
-Then, we update $\Theta$ using Metropolis Hastings or Gibbs sampling when appropriate.
-Finally, we make spatial predictions using conditional multivariate normal results and the fact that the distribution of $Y_t(\bs) \mid \Theta, z(\bs)$ is the usual multivariate normal distribution with a \Matern spatial covariance structure.
-
-We can use Gibbs sampling to update $Y_t(\bs)$ for censored observations that are below the threshold $T$.
-After conditioning on $\lambda$, $z_t(\bs)$ and non-censored observations, $Y_t(\bs)$ has truncated normal full conditionals.
-So we sample $Y_t(\bs) \sim N_{(-\infty, T)}(\bX \bbeta, \bSigma)$.
-After imputing the censored observations, we update the model parameters.
-To update the model parameters, we use standard Gibbs updates for parameters when possible.
-In the case Gibbs sampling is not possible, parameters are updated using a random-walk Metropolis Hastings algorithm.
-See Appendices A.1 and A.2 for details regarding the MCMC.
-The final step of the computation is to use Bayesian Kriging to generate a predictive distribution for $Y_t(\bs^*)$ at prediction location $\bs^*$.
-This step is similar to the imputation for censored observations except that the full conditionals are no longer truncated at $T$.
-
-\section{Simulation study}\label{s:simstudy}
-In this section, we conduct a simulation study to investigate how the number of partitions and the level of thresholding impact the accuracy of predictions made by the model.
-
-\subsection{Design}\label{s:simdesign}
-For all simulation designs, we generate data from the model in Section \ref{s:part} using $n_s=144$ sites and $n_t=50$ independent days.
-The sites are generated Uniform$([0, 10] \times [0, 10])$.
-We generate data from 6 different simulation designs:
-\begin{enumerate} \setlength{\itemsep}{-0.5em}
-  \item Gaussian marginal, $K=1$ knot
-  \item $T$ marginal, $K=1$ knot
-  \item $T$ marginal, $K=5$ knots
-  \item Skew-$t$ marginal, $K=1$ knots
-  \item Skew-$t$ marginal, $K=5$ knots
-  \item Max-stable.
-\end{enumerate}
-In the first five designs, the $v_t(\bs)$ terms are generated using a \Matern covariance with smoothness parameter $\nu = 0.5$ and spatial range $\rho = 0.1$.
-For the covariance matrices in designs 1 -- 5, the proportion of the variance accounted for by the spatial variation is $\gamma = 0.9$ while the proportion of the variance accounted for by the nugget effect is $0.1$.
-In the first design, $\sigma^2 = 2$ is used for all days.
-For designs 2 -- 4, $\sigma^2_{tk} \iid \text{IG}(3, 8)$
-For designs 1 -- 3, we set $\lambda = 0$.
-For designs four and five, $\lambda = 3$ was used, and the $z_t$ are generated as described in (\ref{eq:sitez}).
-In the sixth design, we generate from a spatial max-stable distribution \citep{Reich2012} with parameters $\mu = 1, \sigma=1, \xi=0.2$ and 144 spatial knots on a regular lattice in the square $[1, 9] \times [1, 9]$.
-\hl{I know I need more here about the max-stable method}
-In all six designs, the mean $\bX \bbeta = 10$ is assumed to be constant across space.
-
-$M = 50$ data sets are generated for each design.
-For each data set we fit the data using
-\begin{enumerate} \setlength{\itemsep}{-0.5em}
-  \item Gaussian marginal, $K=1$ knots
-  \item Skew-$t$ marginal, $K=1$ knots, $T=-\infty$
-  \item $T$ marginal, $K=1$ knots, $T=q(0.80)$
-  \item Skew-$t$ marginal, $K=5$ knots, $T=-\infty$
-  \item $T$ marginal, $K=5$ knots, $T=q(0.80)$
-\end{enumerate}
-where $q(0.80)$ is the 80th sample quantile of the data.
-The design matrix $\bX$ includes an the intercept with a prior of $\beta \sim \text{N}(0, 10)$.
-The spatial covariance parameters have priors $\log(\nu) \sim \text{N}(-1.2, 1)$, $\gamma \sim \text{U}(0, 1)$, $\log(\rho) \sim \text{N}(-2, 1)$.
-The skewness parameter has prior $\lambda \sim \text{N}(0, 2)$.
-The residual variance terms have priors $\sigma^2_t(\bs) \sim \text{IG}(0.1, 0.1)$.
-The knots have priors $\bw \sim \text{Unif}(\calD)$.
-We do not fit the data using the max-stable methods from \citet{Reich2012} because of the time it takes.
-
-\subsection{Cross validation}\label{s:modelselect}
-Models were compared using cross validation with 100 sites used as training sites and 30 sites witheld for testing.
-The model was fit using the training set, and predictions were generated at the testing site locations.
-Because one of the primary goals of this model is to predict extreme events, we use Brier scores to select the model that best fits the data \citep{Gneiting2007}.
-The Brier score for predicting exceedance of a threshold $c$ is given by $[e(c) - P(c)]^2$ where $e(c) = I[y>c]$ is an indicator function indicating that a test set value, $y$, has exceeded the threshold, $c$, and $P(c)$ is the predicted probability of exceeding $c$.
-We average the Brier scores over all test sites and days.
-For the Brier score, a lower score indicates a better fit.
-
-\subsection{Results}\label{s:simresults}
-We compared the Brier scores for exceeding 11 different thresholds for each dataset.
-The thresholds used for the Brier scores are extreme quantiles from the simulated data.
-Figure \ref{fig:simbrierscores} gives the Brier score relative to the Brier score for the Gaussian method calculated as
+MVN results:
+The distribution of $\bx_1$ conditional on $\bx_2 = a$ is $N( \mu^*, \Sigma^*)$ where
 \begin{align*}
-  \text{BS}_{\text{rel}} = \frac{\text{BS}_{\text{method}}}{\text{BS}_{\text{Gaussian}}}.
+  \mu^* &= \mu_1 + \Sigma_{12} \Sigma_{22}^{-1} (a - \mu_2)\\
+  \Sigma^* &= \Sigma_{11} - \Sigma_{12} \Sigma_{22}^{-1} \Sigma_{21}
 \end{align*}
-For data that are relatively symmetric, the standard Gaussian method outperforms the other methods.
-However, when the data come from a skewed distribution (settings 3 -- 6), the partitioned and thresholded methods outperform the Gaussian method.
-Additionally, for both of the skew-$t$ settings ($K = 1$, $K = 5$), the skew-$t$ method with the appropriate number of partitions performs the best.
-Finally, the simulation results suggest that for data from a skew-$t$ distribution, thresholded methods provide some improvement over non-thresholded models for the extreme thresholds.
-With the exception of the Brier score for $c = q(0.995)$ in setting 6, the improvement over the Gaussian method is statistically significant using a Wilcoxon signed rank test ($\alpha = 0.05$).
-
-\begin{figure}
-  \includegraphics[width=\linewidth]{plots/bsplots-mean.pdf}
-  \caption{Brier scores relative to the Gaussian method for simulation study results. A ratio lower than 1 indicates that the method outperforms relative to the Gaussian method.}
-  \label{fig:simbrierscores}
-\end{figure}
-
-\section{Data analysis}\label{s:analysis}
-To illustrate this method, we consider the daily maximum 8-hour ozone measurements for July 2005 at 735 Air Quality System (AQS) monitoring sites in the eastern United States as the response (see Figure \ref{fig:ozone}).
-For each site, we also have covariate information containing the estimated ozone from the Community Multi-scale Air Quality (CMAQ) modeling system.
-Initially, we fit a linear regression assuming a mean function of
-\begin{align}
-  \bX \bbeta = \beta_0 + \beta_1 \cdot \text{CMAQ}_t(\bs). \label{eq:datamean}
-\end{align}
-The data from July 10 are shown in Figure \ref{fig:ozone} along with a Q-Q plot of the residuals compared to a skew-$t$ distribution with 8 d.f. and $\alpha = 1$.
-\begin{center}
-\begin{figure}
-  \includegraphics[width=0.5\linewidth]{plots/ozone-10jul.pdf}
-  \includegraphics[width=0.5\linewidth]{plots/qq-res.pdf}
-  \caption{Ozone values on 10 July 2005 (left) Q-Q plot of the residuals (right)}
-  \label{fig:ozone}
-\end{figure}
-\end{center}
-The $\widehat{\chi}(h)$ and $\widehat{\chi}(t)$ plots estimate the spatial and temporal dependence using sample data.
-For $\widehat{\chi}(h)$, we take the sample proportion of all pairs of sites within distance $h$ of one another that are above a threshold $T$.
-\hl{I know I need another sentence or two here about $\widehat{\chi}(t)$}
-indicate that the residuals still demonstrate spatial and temporal dependence (see Figure \ref{fig:chi-st})
-\begin{center}
-\begin{figure}
-  \includegraphics[width=0.5\linewidth]{plots/chi-h-ozone.pdf}
-  \includegraphics[width=0.5\linewidth]{plots/chi-t-ozone.pdf}
-  \caption{$\widehat{\chi}(h)$ plot for the residuals.
-The vertical line indicates the distance after which observations no longer demonstrate dependence (left) $\widehat{\chi}(t)$ plot for the residuals (right)}
-  \label{fig:chi-st}
-\end{figure}
-\end{center}
-
-\subsection{Model comparisons}
-We fit the model using Gaussian and skew-$t$ marginal distributions, $K=1, 5, 10, 15$ partitions, with $Y(\bs)$ censored at $T = 0, 50, 75, 90$ ppb as described in Section \ref{s:data}.
-We also include a max-stable analysis using the method by ????
-All methods assume the mean function given in (\ref{eq:datamean}).
-For each model, Brier scores and quantile scores were were averaged over all sites and days to obtain a single quantile score for each dataset.
-At a particular threshold or quantile level, the model that fits the best is the one with the lowest score.
-Looking at the heatplot in Figure \ref{fig:heatplot}, we can see that the methods using no thresholding tend to outperform those methods that do use thresholding.
-
-\subsection{Results}\label{s:results}
-\begin{figure}
-  \includegraphics[width=0.5\linewidth]{plots/heatplot.pdf}
-  \includegraphics[width=0.5\linewidth]{plots/qscore-best.pdf}
-  \caption{Heatplot of quantile scores for the 99th quantile (left) Quantile scores of Gaussian, skew-$t$ (K=10, T=q(0)), and Max-stable (right)}
-  \label{fig:heatplot}
-\end{figure}
-
-\section{Conclusions}\label{s:con}
-
-\section*{Acknowledgments}
-
-\appendix
-\section{Appendices}
-\subsection{MCMC details} \label{a:mcmc}
-The MCMC sampling for the model \ref{s:hier} is done using {\tt R} (http://www.r-project.org).
-Whenever possible, we select conjugate priors (see Appendix \ref{a:posterior}); however, for some of the parameters, no conjugate prior distributions exist.
-When no conjugate prior distribution exists, we use a random walk Metropolis Hastings update step.
-In each Metropolis Hastings update, we tune the algorithm to give acceptance rates near 0.40.
-
-\subsubsection*{Spatial knot locations}
-For each day, we update the spatial knot locations, $\bw_1, \ldots, \bw_K$, using a Metropolis Hastings block update.
-Because the spatial domain is bounded, we generate candidate knots using the transformed knots $\bw^*_1, \ldots, \bw^*_K$ (see section \ref{s:temporal}) and a random walk bivariate Gaussian candidate distribution
-\begin{align*}
-	{\bw^*_k}^{(c)} \sim \text{N}({\bw^*_k}^{(r - 1)}, s^2 I_2)
-\end{align*}
-where ${\bw^*_k}^{(r - 1)}$ is the location for the transformed knot at MCMC iteration $r - 1$, $s$ is a tuning parameter, and $I_2$ is an identity matrix.
-After candidates have been generated for all $K$ knots, the acceptance ratio is
-\begin{align*}
-  R = \left\{ \frac{ l[ Y_t(\bs | \bw_1^{(c)}, \ldots, \bw_K^{(c)}, \ldots)] }{l[ Y_t(\bs | \bw_1^{(r - 1)}, \ldots, \bw_K^{(r - 1)}, \ldots)]} \right\} \times \left\{ \frac{ \prod_{k = 1}^{K}\phi(\bw_k^{(c)})}{ \prod_{k = 1}^{K}\phi(\bw_k^{(r - 1)})} \right\} \times \left\{ \frac{ \prod_{k = 1}^{K} p({\bw^*_k}^{(c)})}{ \prod_{k = 1}^{K} p({\bw^*_k}^{(r - 1)})}\right\}
-\end{align*}
-where $l$ is the likelihood given in (\ref{eq:hier}), and $p(\cdot)$ is the prior either taken from the time series given in (\ref{s:temporal}) or assumed to be uniform over $\calD$.
-The candidate knots are accepted with probability $\min\{R, 1\}$.
-
-\subsubsection*{Spatial random effects}
-If there is temporal dependence amongst the observations, then we update $z_{tk}$ using a Metropolis Hastings update.
-First, we generate a candidate using a random walk Gaussian candidate distribution
-\begin{align*}
-  z_{tk}^{(c)} \sim \text{N}(z_{tk}^{(r - 1)}, s^2)
-\end{align*}
-where $z_{tk}^{(r-1)}$ is the value at MCMC iteration $r - 1$, and $s$ is a tuning parameter.
-The acceptance ratio is
-\begin{align*}
-  R = \left\{ \frac{ l[Y_t(\bs) | z_{tk}^{(c)}, \ldots] }{ l[Y_t(\bs) | z_{tk}^{(r - 1)}]} \right\} \times \left\{ \frac{ p[ z_{tk}^{(c)} ] }{ p[ z_{tk}^{(r - 1)}]}\right\}.
-\end{align*}
-The candidate is accepted with probability $\min\{R, 1\}$.
-
-\subsubsection*{Variance terms}
-When there is more than one site in a partition, then we update $\sigma^2_{tk}$ using a Metropolis Hastings update.
-First, we generate a candidate for $\sigma^2_{tk}$ using an IG$(a^*/s, b^*/s)$ candidate distribution in an independence Metropolis Hastings update where $a^* = (n_{tk} + 1) / 2 + a$, $b^* = [Y_{tk}^T \Sigma^{-1}_{tk} Y_{tk} + z_{tk}^2] / 2 + b$, $n_{tk}$ is the number of sites in partition $k$ on day $t$, and $Y_{tk}$ and $\Sigma^{-1}_{tk}$ are the observations and precision matrix for partition $k$ on day $t$.
-The acceptance ratio is
-\begin{align*}
-  R = \left\{
-    \frac{ l[Y_t(\bs) | {\sigma^2_{tk}}^{(c)}, \ldots] }
-         { l[Y_t(\bs) | {\sigma^2_{tk}}^{(r - 1)}]}
-    \right\} \times \left\{
-    \frac{ l[z_{tk} | {\sigma^2_{tk}}^{(c)}, \ldots] }
-         { l[z_{tk} | {\sigma^2_{tk}}^{(r - 1)}, \ldots] }
-    \right\} \times \left\{
-    \frac{ p[ {\sigma^2_{tk}}^{(c)} ] }
-         { p[ {\sigma^2_{tk}}^{(r - 1)}] }
-    \right\} \times \left\{
-    \frac{ c[ {\sigma^2_{tk}}^{(r - 1)}] }
-         { c[ {\sigma^2_{tk}}^{(c)}]}
-    \right\}
-\end{align*}
-where $p[\cdot]$ is the prior either taken from the time series given in (\ref{s:temporal}) or assumed to be IG$(a, b)$, and $c[\cdot]$ is the candidate distribution.
-The candidate is accepted with probability $\min\{R, 1\}$.
-
-\subsubsection*{Spatial covariance parameters}
-We update the three spatial covariance parameters, $\log(\rho)$, $\log(\nu)$, $\gamma$, using a Metropolis Hastings block update step.
-First, we generate a candidate using a random walk Gaussian candidate distribution
-\begin{align*}
-	\log(\rho)^{(c)} \sim \text{N}(\log(\rho)^{(r - 1)}, s^2)
-\end{align*}
-where $\log(\rho)^{(r-1)}$ is the value at MCMC iteration $r - 1$, and $s$ is a tuning parameter.
-Candidates are generated for $\log(\nu)$ and $\gamma$ in a similar fashion.
-The acceptance ratio is
-\begin{align*}
-	R = \left\{ \frac{ \prod_{t = 1}^{T} l[Y_t(\bs) | \rho^{(c)}, \nu^{(c)}, \gamma^{(c)}, \ldots] }{\prod_{t = 1}^{T} l[Y_t(\bs) | \rho^{(r-1)}, \nu^{(r-1)}, \gamma^{(r-1)}, \ldots] } \right\} \times \left\{ \frac{ p[\rho^{(c)}] }{ p[\rho^{(r - 1)] } } \right\} \times \left\{ \frac{ p[\nu^{(c)}] }{ p[\nu^{(r - 1)}] } \right\} \times \left\{ \frac{ p[ \gamma^{(c)} ] }{ p[\gamma^{(r - 1)} ] } \right\}.
-\end{align*}
-All three candidates are accepted with probability $\min\{R, 1\}$.
-
-\subsection{Posterior distributions} \label{a:posterior}
-\input{zpost.tex}
-\input{betapost.tex}
-\input{sigpost.tex}
-\input{lambdapost.tex}
-
-\subsection{Proof that $\lim_{h \rightarrow \infty} \pi(h) = 0$} \label{a:proofsamepartition}
-Let $N(A)$ be the number of knots in $A$, the area between sites $\bs_1$ and $\bs_2$.
-Consider a spatial Poisson process with intensity $\mu(A)$.
-So,
-\begin{align*}
-  P[ N(A) = k] = \frac{ \mu(A)^k \exp\{ -\mu(A)\}}{k!}.
-\end{align*}
-Then for any finite $k$, $\lim_{h \rightarrow \infty} P[N(A) = k] = 0$ because $\lim_{h \rightarrow \infty} \mu(A) = \infty$.
-With each additional knot in $A$, the chance that $\bs_1$ and $\bs_2$ will be be in the same partition will decrease, because partition membership is defined by the closest knot to a site.
-Therefore, $\lim_{h \rightarrow \infty} \pi(h) = 0$.
-
-% \subsection{Half-normal distribution}
-% Let $u = |z|$ where $Z \sim N(\mu, \sigma^2)$.
-% Specifically, we consider the case where $\mu = 0$. Then $U$ follows a half-normal distribution which we denote as $U \sim HN(0, 1)$, and the density is given by
-% \begin{align}
-%   f_U(u) = \frac{ \sqrt{2} }{ \sqrt{\pi \sigma^2} } \exp \left( - \frac{ u^2 }{ 2 \sigma^2 } \right) I(u > 0)
-% \end{align}
-% When $\mu = 0$, the half-normal distribution is also equivalent to a $N_{(0, \infty)}(0, \sigma^2)$ where $N_{(a, b)}(\mu, \sigma^2)$ represents a normal distribution with mean $\mu$ and standard deviation $\sigma$ that has been truncated below at $a$ and above at $b$.
-
-\subsection{Skew-$t$ distribution} \label{a:skewt}
-\subsubsection*{Univariate extended skew-$t$ distribution}
-We say that $Y$ follow a univariate extended skew-$t$ distribution with location $\xi \in \calR$, scale $\omega > 0$, skew parameter $\alpha \in \calR$, extended parameter $\tau \in \calR$, and degrees of freedom $\nu$ if has distribution function
-\begin{align}
-  f_{\text{EST}}(y) = \omega^{-1} \frac{ f_T (z; \nu) }{ F_T(\tau / \sqrt{ 1 + \alpha^2}; \nu)} F_T\left[ (\alpha z + \tau) \sqrt{ \frac{ \nu + 1 }{ \nu + z^2}}; 0, 1, \nu + 1 \right]
-\end{align}
-where $f_T(t; \nu)$ is a univariate Student's $t$ with $\nu$ degrees of freedom, $F_T(t; \nu) = P(T < t)$, and \hbox{$z = (y - \xi) / \omega$}.
-In the case that $\tau = 0$, then $Y$ follows a univariate skew-$t$ distribution.
-
-\subsection*{Multivariate skew-$t$ distribution}
-If $\bZ \sim \text{ST}_d(0, \bar{\bOmega}, \balpha, \eta)$ is a $d$-dimensional skew-$t$ distribution, and $\bY = \xi + \bomega \bZ$, where $\bomega = \text{diag}(\omega_1, \ldots, \omega_d)$, then the density of $Y$ at $y$ is
-\begin{align}
-  f_y(\by) = det(\bomega)^{-1} f_z(\bz)
-\end{align}
-where
-\begin{align}
-  f_z(\bz) &= 2 t_d(\bz; \bar{\bOmega}, \eta) T \left[ \balpha^T \bz \sqrt{ \frac{\eta + d}{\nu + Q(\bz)} }; \eta + d\right] \\
-  \bz &= \bomega^{-1}(\by - \xi)
-\end{align}
-where $t_d(\bz; \bar{\bOmega}, \eta)$ is a $d$-dimensional Student's $t$-distribution with scale matrix $\bar{\bOmega}$ and degrees of freedom $\eta$, $Q(z) = \bz^T \bar{\Omega}^{-1}\bz$ and $T(\cdot; \eta)$ denotes the univariate Student's $t$ distribution function with $\eta$ degrees of freedom \citep{Azzalini2013}.
-
-\subsubsection*{Extremal dependence}
-For a bivariate skew-$t$ random variable $\bY = [Y(\bs), Y(\bt)]^T$, the $\chi(h)$ statistic \citep{Padoan2011} is given by
-\begin{align}
-  \chi(h) = \bar{F}_{\text{EST}}\left\{ \frac{[x_1^{1 / \eta} - \varrho(h)] \sqrt{\eta + 1} }{\sqrt{1 - \varrho(h)^2}}; 0, 1, \alpha_1, \tau_1, \eta + 1 \right\} + \bar{F}_{\text{EST}}\left\{ \frac{ [x_2^{1 / \eta} - \varrho(h)] \sqrt{\eta + 1} }{ \sqrt{1 - \varrho(h)^2} }; 0, 1, \alpha_2, \tau_2, \eta + 1 \right\},
-\end{align}
-where $\bar{F}_{\text{EST}}$ is the univariate survival extended skew-$t$ function with zero location and unit scale, \hbox{$\varrho(h) = \text{cor}(y_1, y_2)$}, $\alpha_j = \alpha_i \sqrt{1 - \varrho^2}$, $\tau_j = \sqrt{\eta + 1}(\alpha_j + \alpha_i \varrho)$, and $x_j = F_T(\bar{\alpha}_i \sqrt{\eta + 1}; 0, 1, \eta) / F_T(\bar{\alpha}_j \sqrt{\eta + 1}; 0, 1, \eta)$ with $j = 1, 2$ and $i = 2, 1$ and where $\bar{\alpha}_j = (\alpha_j + \alpha_i \varrho) / \sqrt{ 1 + \alpha_i^2 [1 - \varrho(h)^2]}$.
-
-\subsubsection*{Proof that $\lim_{h \rightarrow \infty} \chi(h) > 0$}
-Consider the bivariate distribution of $\bY = [Y(\bs), Y(\bt)]^T$, with $\varrho(h)$ given by (\ref{eq:matern}).
-So, $\lim_{h \rightarrow \infty} \varrho(h) = 0$.
-Then
-\begin{align}
-  \lim_{h \rightarrow \infty} \chi(h) = \bar{F}_{\text{EST}}\left\{ \sqrt{\eta + 1}; 0, 1, \alpha_1, \tau_1, \eta + 1 \right\} + \bar{F}_{\text{EST}}\left\{ \sqrt{\eta + 1}; 0, 1, \alpha_2, \tau_2, \eta + 1 \right\}.
-\end{align}
-Because the extended skew-$t$ distribution is not bounded above, for all $\bar{F}_{\text{EST}}(x) = 1 - F_{\text{EST}} > 0$ for all $x < \infty$.
-Therefore, for a skew-$t$ distribution, $\lim_{h \rightarrow \infty} \chi(h) > 0$.
-
-\bibliographystyle{rss}
-\bibliography{skewnormal}
 

### 2015-02-17 binary/code/densities.R

@@ -1,232 +1,0 @@
-###############################################################################
-###### pmf eval                                                          ######
-###############################################################################
-
-pspatbin <- function(alpha.f, FAC.f, xi.f, beta.f, X.f, y.f){
-	
-	#### A single day
-	#### To do: Rewrite function to take all observations at once
-	n <- length(y.f)
-	K <- sum(y.f)
-	
-	#### Identify the indices that exceed the threshold
-	y.exceed <- which(y.f == 1)
-	
-	#### Create inclusion/exclusion matrix
-	#### Starting with a matrix that is only 1
-	incl.excl <- matrix(1, n, 2^K)
-	
-	#### Now change entries for proper inclusion/exclusion
-	incl.excl[y.exceed,] <- create.comb(K)	
-	z <- (1 + xi.f * X.f %*% beta.f)^(1/xi.f)			#should be n x 1
-	inner <- exp(1/alpha.f * ( log(FAC.f) - log(z) ) )	#should be n x L
-	
-	
-}
-
-######################################################
-#### Positive Stable Density                      ####
-#### From Reich and Shaby (2011)                  ####
-######################################################
-
-h1<-function(logs,u,alpha,log=T){
-    s<-exp(logs)
-    psi<-pi*u
-    c<-(sin(alpha*psi)/sin(psi))^(1/(1-alpha))
-    c<-c*sin((1-alpha)*psi)/sin(alpha*psi)
-  
-    logd<-log(alpha)-log(1-alpha)-(1/(1-alpha))*logs+
-          log(c)-c*(1/s^(alpha/(1-alpha)))+logs
-        
-	if(!log){
-		logd<-exp(logd)
-  	}
-  
-  return(logd)
-}
-
-###############################################################################
-###### cdf eval                                                          ######
-###############################################################################
-
-
-
-dtnorm<-function(y,mn,sd=.25,fudge=0){
-	upper<-pnorm(1-fudge,mn,sd)
-	lower<-pnorm(fudge,mn,sd)
-	l<-dnorm(y,mn,sd,log=T)-log(upper-lower)
-
-	return(l)
-}
-
-
-###############################################################################
[…]

### 2015-02-17 binary/code/mcmc.R

@@ -1,621 +1,0 @@
-#########|#########|#########|#########|#########|#########|#########|#########
-######################################################
-########      THE MAIN MCMC FUNCTION       ###########
-######################################################
-
-
-Bayes_GPD <-function(
-	y,								#data
-	S,								#locations
-	knots=NULL,						#knots
-    X=NULL,							#covariates
-    Sp=NULL, 						#Prediction locations
-    Xp=NULL,						#Prediction covariates
-    thresh=100,						#Threshold value
-    init.beta=c(0,.1, -2.94),		#Initial GPD parameters - logscale, shape, pat
-    init.alpha=.5,					#Initial alpha
-    init.range=1,					#Initial range
-    init.bw=1,						#Initial bandwidth
-    init.logs=NULL,
-    vary=c(F,F,F),					#Varying GPD parameters by location (logsig, xi, pat)
-    pri.mn.range=-2,				#Range
-    pri.sd.range=1,
-    pri.mn.bw=0,					#Bandwidth
-    pri.sd.bw=1,
-    pri.var.a=0.1,					#a and b (gamma parameters for var)
-    pri.var.b=0.1,
-    pri.alpha.a=1,					#a and b (beta parameters for alpha)
-    pri.alpha.b=1,
-    #pri.mn.Phi.alpha=0,
-    #pri.sd.Phi.alpha=0.25,
-    pri.sd.beta=10,					#beta sd
-    pri.mn.gev=c(0,0,-2.5),			    #mean for gev
-    pri.sd.gev=c(1,0.25,0.5),		    #sd for gev params
-    keep.samples=T,					#Retain samples
-    nthreads = 2,
-    iters=200,burn=50,update=50, nthin=1){
-
-    start.time <- proc.time()
-
-    if(is.null(knots)){
-    	knots<-S
-    }
-    if(is.null(X)){
-    	X<-cbind(1,S)
-    }
-
-	y[y <= thresh] <- thresh		#Everything below the threshold = threshold
-
-    #### Some bookkeeping values
-    n <- nrow(y)					#Number of sites
-    nt <- ncol(y)					#Number of days
-    nF <- nrow(knots)				#Number of knots
-    p <- ncol(X)					#Number of covariates
-
-    d<-rdist(S,S)					#Distance between locations
-    diag(d)<-0
[…]

### 2015-02-17 binary/code/plots.R

@@ -1,127 +1,0 @@
-###############################################################################
-#### Basic setup                                                           ####
-###############################################################################
-#setwd('~/svn-repos/dissertation/extreme-value/code')
-#setwd('S:/Documents/Dissertation/extreme-value/code')
-
-library(fields)
-library(emulator)
-library(evd)
-library(POT)
-library(geoR)
-
-load(file="./simdata.RData")
-
-#### Model parameters                
-alpha <- c(.3, .5, .7)  # Spatial correlation 
-                            # (0 = spatial dep, 1 = spatial ind)     
-logrange <- 0         				    # Kernel bandwidth 
-
-for(set in 1:nsets){
-	for(day in 1:nt){
-		file1 <- paste("dataplots/sim1day",day,".pdf", sep="")
-		file2 <- paste("dataplots/sim2day",day,".pdf", sep="")
-		file3 <- paste("dataplots/sim3day",day,".pdf", sep="")
-		dev.new(width=10, height=5)
-		par(mfrow=c(1, 2))
-		z <- matrix(y.1[,day,set], length(s2), length(s2))
-		image.plot(s1, s2, z, 
-			main=bquote(paste(
-				alpha == .(alpha[1]), ", ", xi==.(xi), " I(Z > Thresh)")
-			)
-		)
-		z <- matrix(latent.1[,day,set], length(s2), length(s2))
-		image.plot(s1, s2, z, 
-			main=bquote(paste(
-				alpha == .(alpha[1]), ", ", xi==.(xi), " Latent Process")
-			)
-		)
-		dev.print(file=file1, device=pdf)
-		
-		z <- matrix(y.2[,day,set], length(s2), length(s2))
-		image.plot(s1, s2, z, 
-			main=bquote(paste(
-				alpha == .(alpha[2]), ", ", xi==.(xi), " I(Z > Thresh)")
-			)
-		)
-		z <- matrix(latent.2[,day,set], length(s2), length(s2))
-		image.plot(s1, s2, z, 
-			main=bquote(paste(
-				alpha == .(alpha[2]), ", ", xi==.(xi), " Latent Process")
-			)
-		)
-		dev.print(file=file2, device=pdf)
[…]

### 2015-02-17 binary/code/setup.R

@@ -1,127 +1,0 @@
-###############################################################################
-#### Basic setup                                                           ####
-###############################################################################
-#setwd('~/svn-repos/dissertation/extreme-value/code')
-#setwd('S:/Documents/Dissertation/extreme-value/code')
-
-library(fields)
-library(emulator)
-library(evd)
-library(POT)
-library(geoR)
-
-#### mcmc.R - MCMC algorithm
-source('./mcmc.R')
-
-#### densities.R - Required density calculations for mcmc
-source('./densities.R')
-
-#### Spatial GP code
-source('./spatial_gp.R')
-
-###############################################################################
-#### Data settings                                                         #### 
-###############################################################################
-
-#### Model parameters                
-
-## Latent variable
-alpha <- c(.3, .5, .7)  # Spatial correlation 
-                      # (0 = spatial dep, 1 = spatial ind)     
-logrange <- 0       # Kernel bandwidth 
-
-## Regression
-xi <- .1
-nt <- 10       				    # Number of days 
-
-#### Two options for beta terms
-beta.1 <- c(3.458, 0, 0)  # Intercept, Lat, Long 5%
-beta.2 <- c(1.993, 0, 0)  # Intercept, Lat, Long 15%
-
-#### Three different grid sizes for checking timing
-s1 <- 1:6       				# Location x
-s2 <- 1:6      				    # Location y
-S.1 <- expand.grid(s1, s2)
-knots.1 <- S.1      			# Knots at grid points  
-n.1 <- nrow(S.1)  
-X.hm.1 <- as.matrix(cbind(1, S.1), n.1, 3)
-
-s1 <- 1:10       				# Location x
-s2 <- 1:10      			    # Location y
-S.2 <- expand.grid(s1, s2)
[…]

### 2015-02-17 thresh/code/simstudy/results-fine.R

@@ -1,242 +1,0 @@
-rm(list=ls())
-load("simdata.RData")
-n <- dim(y)[1]
-nsets <- dim(y)[2]
-nsettings <- dim(y)[3]
-nthreshs <- length(thresh.probs)
-
-# source("../../R/auxfunctions.R")	# Included for easy access if we need to change score functions
-
-iters <- 6000; burn <- 2000
-usable <- (burn + 1):iters
-quant.grid <- seq(0.95, 1, 0.0001)
-quant.grid <- quant.grid[-length(quant.grid)] # remove the last one for integration
-grid.count <- length(quant.grid) 
-grid.length <- (1 - U) / grid.count # used for integration
-quant.scores <- brier.scores <- list()
-
-for (setting in 1:nsettings) {
-  dataset <- paste("thresh-", setting, ".RData", sep="")
-  cat("setting", dataset, "started \n")
-  load(dataset)
-  quant.scores[[setting]] <- brier.scores[[setting]] <- list()
-  
-  for (set in 1:nsets) {
-    # storage
-    # quant.scores.set <- matrix(NA, nrow=grid.count, nthreshs)
-    brier.scores.set <- matrix(NA, nrow=grid.count, nthreshs)
-    
-    for (thresh in 1:nthreshs) {
-	  validate <- y[, set, setting]
-	  pred.thresh <- t(preds[[set]][usable, , thresh])
-	  # quant.scores.set[, thresh] <- QuantScore(preds=pred.thresh, probs=quant.grid, 
-	                                           # validate=validate)
-	  brier.scores.set[, thresh] <- BrierScore(preds=pred.thresh, probs=quant.grid, 
-	                                           validate=validate)
-    }
-
-    # quant.scores[[setting]][[set]] <- quant.scores.set
-    brier.scores[[setting]][[set]] <- brier.scores.set
-    
-    save(
-      # quant.scores, 
-      brier.scores,
-      quant.grid,
-      file="scores-fine.RData"
-    )
-    cat("    set", set, "finished \n")
-  }
-  
-  cat("setting", dataset, "finished \n")
-}
-
-
-save(
-    # quant.scores, 
-    brier.scores,
-    quant.grid,
-    file="scores-fine.RData"
-)
-
-rm(list=ls())
[…]

### 2015-02-17 thresh/code/simstudy/setting1-low.R

@@ -1,47 +1,0 @@
-library(evd)
-rm(list=ls())
-load(file='simdata.RData')
-
-setting <- 1
-
-n <- dim(y)[1]
-nsets <- dim(y)[2]
-iters <- 6000; burn <- 2000
-thresh.probs <- seq(0.50, 0.95, 0.05)
-
-nthreshs <- length(thresh.probs)
-outputfile <- paste("thresh-low", setting, ".RData", sep="")
-
-# want to store all of the predicted datasets - list that is nsets long
-preds <- beta.prob <- beta.scale <- beta.shape <- list()
-
-for (set in 1:nsets) {
-  cat("Dataset", set, "started \n")
-  tic <- proc.time()
-  y.set <- y[, set, setting]
-  
-  threshs <- quantile(y.set, probs=thresh.probs)
-  
-  preds[[set]] <- array(NA, dim=c(iters, n, nthreshs))
-  beta.prob[[set]] <- beta.scale[[set]] <- beta.shape[[set]] <- array(NA, dim=c(iters, ncol(X), nthreshs))
-  
-  for (t in 1:nthreshs){
-    fit <- Bayes_GPD(y=y.set, X.prob=X, X.sig=X, X.xi=X, 
-                     Xp.prob=X, Xp.sig=X, Xp.xi=X,
-                     thresh=threshs[t], iters=iters, burn=burn, update=1000,
-                     iterplot=F)
-  
-    beta.prob[[set]][, , t] <- fit$beta.prob
-    beta.scale[[set]][, , t] <- fit$beta.sig
-    beta.shape[[set]][, , t] <- fit$beta.xi
-    preds[[set]][, , t] <- fit$yp
-  	cat("  thresh", t, "finished \n")
-  	
-  	save(beta.prob, beta.scale, beta.shape, preds,
-  	     file=outputfile)
-  }
-  
-  elap.time.set <- (proc.time() - tic)[3]
-  avg.time.set  <- elap.time.set / set
-  cat("Dataset", set, "finished ", round(avg.time.set, 2), "per dataset \n")
-}

### 2015-02-17 thresh/code/simstudy/setting2.R

@@ -1,45 +1,0 @@
-library(evd)
-rm(list=ls())
-load(file='simdata.RData')
-
-setting <- 2
-
-n <- dim(y)[1]
-nsets <- dim(y)[2]
-iters <- 6000; burn <- 2000
-nthreshs <- length(thresh.probs)
-outputfile <- paste("thresh-", setting, ".RData", sep="")
-
-# want to store all of the predicted datasets - list that is nsets long
-preds <- beta.prob <- beta.scale <- beta.shape <- list()
-
-for (set in 1:nsets) {
-  cat("Dataset", set, "started \n")
-  tic <- proc.time()
-  y.set <- y[, set, setting]
-  
-  threshs <- quantile(y.set, probs=thresh.probs)
-  
-  preds[[set]] <- array(NA, dim=c(iters, n, nthreshs))
-  beta.prob[[set]] <- beta.scale[[set]] <- beta.shape[[set]] <- array(NA, dim=c(iters, ncol(X), nthreshs))
-  
-  for (t in 1:nthreshs){
-    fit <- Bayes_GPD(y=y.set, X.prob=X, X.sig=X, X.xi=X, 
-                     Xp.prob=X, Xp.sig=X, Xp.xi=X,
-                     thresh=threshs[t], iters=iters, burn=burn, update=1000,
-                     iterplot=F)
-  
-    beta.prob[[set]][, , t] <- fit$beta.prob
-    beta.scale[[set]][, , t] <- fit$beta.sig
-    beta.shape[[set]][, , t] <- fit$beta.xi
-    preds[[set]][, , t] <- fit$yp
-  	cat("  thresh", t, "finished \n")
-  	
-  	save(beta.prob, beta.scale, beta.shape, preds,
-  	     file=outputfile)
-  }
-  
-  elap.time.set <- (proc.time() - tic)[3]
-  avg.time.set  <- elap.time.set / set
-  cat("Dataset", set, "finished ", round(avg.time.set, 2), "per dataset \n")
-}

### 2015-02-17 thresh/code/simstudy/setting3-low.R

@@ -1,47 +1,0 @@
-library(evd)
-rm(list=ls())
-load(file='simdata.RData')
-
-setting <- 3
-
-n <- dim(y)[1]
-nsets <- dim(y)[2]
-iters <- 6000; burn <- 2000
-thresh.probs <- seq(0.50, 0.95, 0.05)
-
-nthreshs <- length(thresh.probs)
-outputfile <- paste("thresh-low", setting, ".RData", sep="")
-
-# want to store all of the predicted datasets - list that is nsets long
-preds <- beta.prob <- beta.scale <- beta.shape <- list()
-
-for (set in 1:nsets) {
-  cat("Dataset", set, "started \n")
-  tic <- proc.time()
-  y.set <- y[, set, setting]
-  
-  threshs <- quantile(y.set, probs=thresh.probs)
-  
-  preds[[set]] <- array(NA, dim=c(iters, n, nthreshs))
-  beta.prob[[set]] <- beta.scale[[set]] <- beta.shape[[set]] <- array(NA, dim=c(iters, ncol(X), nthreshs))
-  
-  for (t in 1:nthreshs){
-    fit <- Bayes_GPD(y=y.set, X.prob=X, X.sig=X, X.xi=X, 
-                     Xp.prob=X, Xp.sig=X, Xp.xi=X,
-                     thresh=threshs[t], iters=iters, burn=burn, update=1000,
-                     iterplot=F)
-  
-    beta.prob[[set]][, , t] <- fit$beta.prob
-    beta.scale[[set]][, , t] <- fit$beta.sig
-    beta.shape[[set]][, , t] <- fit$beta.xi
-    preds[[set]][, , t] <- fit$yp
-  	cat("  thresh", t, "finished \n")
-  	
-  	save(beta.prob, beta.scale, beta.shape, preds,
-  	     file=outputfile)
-  }
-  
-  elap.time.set <- (proc.time() - tic)[3]
-  avg.time.set  <- elap.time.set / set
-  cat("Dataset", set, "finished ", round(avg.time.set, 2), "per dataset \n")
-}

### 2015-03-30 LaTeX/skewnormal.tex
new plots for manuscript

--- 2015-01-05 pot/LaTeX/skewnormal.tex
@@
 Then $\bw^*_{tk} \in \calR^2$.
-We use a copula on the $\sigma^2_t(\bs)$ to ensure that the marginal distributions of $\sigma^2_t(\bs)$ are inverse gamma.
+We use a copula on $\sigma^2_t(\bs)$ to ensure that the marginal distributions of $\sigma^2_t(\bs)$ are inverse gamma.
 Let
@@
 where IG is the distribution function for an IG$(a, b)$ random variable.
-The AR(1) process for each tail parameter is $\bw^*_{1k} \sim N_w(0, 1)$, $z_{1k} \sim N(0, \sigma^2_{1k})$, $\sigma^{2*}_{1k} \sim N(0, 1)$, and for $t > 1$ the time series is modeled as
+We also use a copula on $z_{t}(\bs)$ to ensure that the marginal distributions of $z_t(\bs)$ are half-normal.
+Let
+\begin{align}
+  z^*_t(\bs) = \Phi^{-1}\left\{ \text{HN}[\sigma^2_t(\bs)] \right\}
+\end{align}
+The AR(1) process for each tail parameter is $\bw^*_{1k} \sim N_w(0, 1)$, $z^*_{1k} \sim N(0, \sigma^2_{1k})$, $\sigma^{2*}_{1k} \sim N(0, 1)$, and for $t > 1$ the time series is modeled as
 \begin{align}
   \bw^*_{tk} | \bw^*_{t-1, k} &\sim N_2\left[\phi_w \bw^*_{t-1, k}, (1 - \phi_w^2) \right] \\
-  z_{tk} | z_{t-1, k} &\sim N \left[\phi_z z_{t-1, k}, \sigma^2_{tk} (1 - \phi_z^2)\right] \\
+  z^*_{tk} | z^*_{t-1, k} &\sim N \left[\phi_z z^*_{t-1, k}, \sigma^2_{tk} (1 - \phi_z^2)\right] \\
   \sigma^{2*}_{tk} | \sigma^{2*}_{t-1, k} &\sim N \left[\phi_\sigma \sigma^{2*}_{t-1, k}, (1 - \phi_\sigma^2) \right]
 \end{align}
-where $|\phi_w| < 1$, $0 < \phi_z < 1$, and $|\phi_\sigma| < 1$.
-We restrict $\phi_z$ to be positive because in the model, we assume $Y(\bs)$ is a function of $|z_t(\bs)|$.
-These are stationary time series models with marginal distributions $\bw^*_{k} \sim N_2(0, 1)$, $z_{k} \sim N(0, \sigma^2_{k})$, $\sigma^{2*}_{k} \sim N(0, 1)$.
-After transformation back to the original space, $\bw_{tk} \sim \text{Unif}(\calD)$, $\sigma^2_{tk} \sim \text{IG}(a, b)$.
+where $|\phi_w|$, $|\phi_z|$, $|\phi_\sigma| < 1$.
+These are stationary time series models with marginal distributions $\bw^*_{k} \sim N_2(0, 1)$, $z^*_{k} \sim N(0, \sigma^2_{k})$, $\sigma^{2*}_{k} \sim N(0, 1)$.
+After transformation back to the original space, $\bw_{tk} \sim \text{Unif}(\calD)$, $z_{tk} \sim HN(0, \sigma^2_{tk})$ $\sigma^2_{tk} \sim \text{IG}(a, b)$.
 For each day, the model is identical to the spatial-only model in (\ref{eq:partition}) by construction.
@@
    \sigma^2_{t}(\bs) &= \sigma^2_{tk} \text{ if } \bs \in P_{tk}\\
-   \lambda &\sim N(0, 10)\\
+   \lambda &= \lambda_1 \lambda_2\\
+   \lambda_1 &= \left\{ \begin{array}{ll}
+      +1 \quad & \text{w.p.} 0.5\\
+      -1 \quad & \text{w.p.} 0.5
+   \end{array}\right.\\
+   \lambda^2_2 & \sim IG(a, b)\\
    v_t(\bs) \mid \Theta &\sim \Matern(0, \Sigma)\\
-   z_{tk} \mid z_{t-1, k}, \sigma^2_{tk} &\sim N(\phi_z z_{t-1, k}, \sigma^2_{tk} (1 - \phi_z^2))\\
+   z^*_{tk} \mid z^*_{t-1, k}, \sigma^2_{tk} &\sim N(\phi_z z^*_{t-1, k}, \sigma^2_{tk} (1 - \phi_z^2))\\
    \sigma^{2*}_{tk} \mid \sigma^{2*}_{t-1, k} &\sim N(\phi_\sigma \sigma^{2*}_{t-1, k}, (1 - \phi_\sigma^2))\\
@@
 where $\Theta = \{\rho, \nu, \gamma\}$, and $\Sigma$ is a \Matern covariance matrix as described in Section \ref{s:skewt}.
+We parameterize $\lambda = \lambda_1 \lambda_2$ to help with convergence in the MCMC.
 
@@
 The sites are generated Uniform$([0, 10] \times [0, 10])$.
-We generate data from 6 different simulation designs:
+We generate data from 7 different simulation designs:
 \begin{enumerate} \setlength{\itemsep}{-0.5em}
@@
   \item Skew-$t$ marginal, $K=5$ knots
-  \item Max-stable.
+  \item Max-stable
+  \item Transformation below $T = q(0.80)$
 \end{enumerate}
@@
 For this setting, we set $\gamma = 0.5$.
-In all six designs, the mean $\bX \bbeta = 10$ is assumed to be constant across space.
+In the final design, we generate $\tilde{y}$ using the setting from design 4, and then consider the data
+\begin{align}
+  y = \left\{ \begin{array}{lc}
+    \tilde{y}, \quad & \tilde{y} > T \\[0.5em]
+    T \exp\{\tilde{y} - T\}, \quad & \tilde{y} \le T
+  \end{array}\right.
+\end{align}
+where $T = q(0.80)$ is the 80th sample quantile of the data.
+In all seven designs, the mean $\bX \bbeta = 10$ is assumed to be constant across space.
 
@@
   \item Skew-$t$ marginal, $K=1$ knots, $T=-\infty$
-  \item Skew-$t$ marginal, $K=1$ knots, $T=q(0.80)$
+  \item Symmetric-$t$ marginal, $K=1$ knots, $T=q(0.80)$
   \item Skew-$t$ marginal, $K=5$ knots, $T=-\infty$
-  \item Skew-$t$ marginal, $K=5$ knots, $T=q(0.80)$
+  \item Symmetric-$t$ marginal, $K=5$ knots, $T=q(0.80)$
 \end{enumerate}
@@
 The design matrix $\bX$ includes an the intercept with a prior of $\beta \sim \text{N}(0, 10)$.
-The spatial covariance parameters have priors $\log(\nu) \sim \text{N}(-1.2, 1)$, $\gamma \sim \text{U}(0, 1)$, $\log(\rho) \sim \text{N}(-2, 1)$.
+The spatial covariance parameters have priors $\log(\nu) \sim \text{N}(-1.2, 1)$, $\gamma \sim \text{Unif}(0, 1)$, $\rho \sim \text{Unif}(15)$.
 The skewness parameter has prior $\lambda \sim \text{N}(0, 2)$.
@@
 Figure \ref{fig:simbrierscores} gives the Brier score relative to the Brier score for the Gaussian method calculated as
-\begin{align*}
+\begin{align}
   \text{BS}_{\text{rel}} = \frac{\text{BS}_{\text{method}}}{\text{BS}_{\text{Gaussian}}}.
-\end{align*}
-We analyzed the results for the simulation study using two-sided Wilcoxon signed rank tests.
-For the first five data settings, our model with no thresholding and $K = 1$ or $5$ partitions performs significantly better (p-value $<0.001$) than the Gaussian method.
-Additionally, for the $T$ and skew-$t$ settings with $K = 5$, the skew-$t$ method with $K=5$ partitions performs significantly better (p-value $<0.001$) than the skew-$t$ method with $K=1$ partition except for the Brier score at $T = q(0.995)$ for the $T$ setting with $K = 5$ (p-value $=0.036$).
-For the $T$ and skew-$t$ settings with $K = 1$, the Brier scores for the model with $K = 1$ partition are significantly better (p-value $<0.001$) for thesholds less than $q(0.95)$.
-However, as the threshold for the Brier score increases, the difference becomes less significant.
-Finally, in the max-stable data setting, our model with $K=1$ partition has significantly better Brier scores than the Gaussian method for thresholds $q(0.92)$ -- $q(0.95)$ (p-value $<0.05$) and $q(0.95)$ -- $q(0.995)$ (p-value $<0.001$).
+\end{align}
+We analyzed the results for the simulation study using two-sided Wilcoxon signed rank tests at a significance level of $\alpha = 0.001$.
+When the data come from a Gaussian distribution, the Gaussian method performs significantly better than the other methods for quantiles below $q(0.95)$.
+However, as we move further out in the tails, the difference in Brier scores is no longer significant beyond $q(0.95)$ when compared to method two and beyond $q(0.99)$ for methods three through five.
+When the data are generated using method two, our methods perform significantly better than the Gaussian method in all cases except for $q(0.90)$ with method five.
+For data setting three, method four performs significantly better at all quantile levels, and method five performs significantly better for quantiles more extreme than $q(0.93)$.
+In this setting, methods two and three do not show significant improvement over the Gaussian method until $q(0.995)$.
+In setting four, method two shows a significant improvement over the Gaussian method at all quantile levels.
+Methods three and four show improvement over the Gaussian method at quantile levels below $q(0.94)$.
+The Gaussian method performs significantly better than method five for all quantile levels in data setting four.
+For setting five, methods three and four perform significantly better than the Gaussian method at all quantile levels.
+Method one performs better than the Gaussian method at quantiles higher than $q(0.99)$, and method two performs better than the Gaussian method at quantiles below $q(0.98)$.
+For setting six, the Gaussian method performs significantly better than all methods for quantiles $q(0.90)$ -- $q(0.94)$.
+However, for quantiles $q(0.99)$ and $q(0.995)$, methods two and three outperform the Gaussian method.
+Finally, in setting seven, methods three demonstrate a significant improvement over the Gaussian method for all quantiles, and method two demonstrates a significant improvement of the Gaussian method for all quantiles except $q(0.94)$.
+Method four performs significantly better than the Gaussian method for $q(0.90)$, $q(0.92)$, and $q(0.95)$ -- $q(0.99)$.
+Finally, method five for setting seven shows no signifcant difference in Brier scores from the Gaussian method.
 
@@
   \centering
-  \includegraphics[width=0.55\linewidth]{plots/ozone-10jul-us.pdf}
-  \includegraphics[width=0.44\linewidth]{plots/qq-res.pdf}
+  \includegraphics[width=0.56\linewidth]{plots/ozone-10jul-us.pdf}
+  \includegraphics[width=0.41\linewidth]{plots/qq-res.pdf}
   \caption{Ozone values on 10 July 2005 (left) Q-Q plot of the residuals (right)}
@@
 \subsection{Model comparisons}
-The data analysis was done in two stages.
-In the first stage, we fit the model using Gaussian and skew-$t$ marginal distributions with $K=1, 5, 10, 15$ partitions.
-We chose to censor $Y(\bs)$ at $T = 0, 50, 75, 85$ ppb in order to compare results from no, moderate, and high censoring.
-We also include a max-stable analysis using the hierarchical max-stable model of \citet{Reich2012}.
+We fit the model using Gaussian and skew-$t$ marginal distributions with $K=1, 5, 6, 7, 8, 9, 10, 15$ partitions.
+We choose to censor $Y(\bs)$ at $T = 0$, $\widehat{q}(0.42) = 50$, $\widehat{q}(0.92)=75$ ppb in order to compare results from no, moderate, and high censoring.
+We also compare models with no time series to models that include the time series.
+Finally, as a comparison to max-stable methods, we fit the model using the hierarchical max-stable model of \citet{Reich2012}.
 All methods assume the mean function given in (\ref{eq:datamean}).
@@
 At a particular threshold or quantile level, the model that fits the best is the one with the lowest score.
-Based on the results from the first stage, we conducted a second stage analysis with $K = 6, 7, 8, 9$.
+We then compute the relative Brier scores to compare each model.
 
 \subsection{Results}\label{s:results}
-The plot in Figure \ref{fig:bs-ozone} shows the relative Brier scores (see Section \ref{s:simresults}) for the first and second stages of the analysis.
+The plots in Figure \ref{fig:bs-ozone} shows the relative Brier scores (see Section \ref{s:simresults}) for time-series and non-time-series models, using $K = $1, 7, and 15 knots at thresholds $T = $0, 50, and 75 ppb.
+Both plots have similar features suggesting that most settings do reasonably well.
+In particular, for all extreme quantiles, selecting a moderate number of knots (e.g. $K = 5, \ldots, 10$) tends to give the best results.
 \begin{figure}
-  \includegraphics[width=0.5\linewidth]{plots/bs-ozone-1.pdf}
-  \includegraphics[width=0.5\linewidth]{plots/bs-ozone-2.pdf}
-  \caption{Relative Brier scores for stage 1 ozone analysis (left) and stage 2 ozone analysis (right)}
+  \includegraphics[width=\linewidth]{plots/bs-ozone.pdf}
+  \caption{Relative Brier scores for time-series models (left) and non-time-series models (right). Relative brier score for the max-stable model is between 1.07 and 1.15}
   \label{fig:bs-ozone}
 \end{figure}
-As with the simulation results, the Brier scores from the thresholded models were higher than those from the non-thresholded models, so we do not include them here.
-Based upon these results, it would appear that our model using a partition with anywhere from $K = 5$ to $K = 10$ partitions gives Brier scores across all threshold levels that are consistently low.
+Table \ref{tbl:ozoneresults} shows the best two models for selected extreme quantiles.
+The results demonstrate the importance of accounting for the temporal dependence when making extreme predictions.
+They also show the importance of thresholding for predictions further out in the tails of the data.
+This is further illustrated by the plots in Figure \ref{fig:ozoneq99}.
+
+\begin{table}
+  \small
+  \caption{Top two performing models for ozone analysis at extreme quantiles with Relative Brier score}
+  \label{tbl:ozoneresults}
+  \centering
+  \begin{tabular}{|l|l l l c|l l l c|}
+    \cline{2-9}
+    \multicolumn{1}{c|}{}  & \multicolumn{4}{c|}{1st} & \multicolumn{4}{c|}{2nd} \\
+    \hline
+    $q(0.90)$  & No time series & $K=7$  & $T=0$  & BS: 0.980 &
+                 No time series & $K=9$  & $T=0$  & BS: 0.980 \\
+    $q(0.95)$  & No time series & $K=15$ & $T=50$ & BS: 0.970 &
+                 No time series & $K=9$  & $T=50$ & BS: 0.970\\
+    $q(0.98)$  & No time series & $K=5$  & $T=50$ & BS: 0.945 &
+                 No time series & $K=10$ & $T=50$ & BS: 0.946\\
+    $q(0.99)$  & Time series    & $K=10$ & $T=75$ & BS: 0.912 &
+                 Time series    & $K=6$  & $T=75$ & BS: 0.913\\
+    $q(0.995)$ & Time series    & $K=6$  & $T=75$ & BS: 0.917 &
+                 Time series    & $K=10$ & $T=75$ & BS: 0.918\\
+    \hline
+  \end{tabular}
+\end{table}
+
+\begin{figure}
+  \includegraphics[width=\linewidth]{plots/q99-ozone.pdf}
+  \caption{(a) -- (d) give give the posterior predictive $\widehat{q}(0.99)$ for the month of July under four different models, (e) gives the difference between $\widehat{q}(0.99)$ in plots (d) and (a), (f) gives the difference between $\widehat{q}(0.99)$ in plots (d) and (b).}
+  \label{fig:ozoneq99}
+\end{figure}
 
@@
 In this paper we propose a new approach for spatitemporal modeling of extreme values.
-The proposed model gives flexible tail behavior, demonstrates asymptotic dependence for observations at sites that are near to one another, and has computation on the order of Gaussian models for large space-time datasets.
-In the simulation study, we demonstrate that there are statistically significant improvements over a na\"{i}ve Gaussian approach.
+The proposed model gives flexible tail behavior, demonstrates asymptotic dependence for observations at sites that are near to one another, \hl{and has computation on the order of Gaussian models for large space-time datasets -- Do I need to add anything about the computation in the paper?}.
+In the simulation study, we demonstrate that this model shows statistically significant improvements over a na\"{i}ve Gaussian approach.
 In both the simulation study, and the application to ozone data, we find that incorporating a partition in the model improves extreme prediction.
-We also find that censoring the data below a threshold does not improve the results, and in many cases makes them worse.
+Furthermore the results suggest that thresholding can improve performance when predicting in the extreme tails of the data.
 
@@
 \subsubsection*{Spatial random effects}
+If there is no temporal dependence amongst the observations, we use a Gibbs update for $z_{tk}$, and the posterior distribution is given in \ref{a:posterior}.
 If there is temporal dependence amongst the observations, then we update $z_{tk}$ using a Metropolis Hastings update.
-First, we generate a candidate using a random walk Gaussian candidate distribution
+Because this model uses $|z_{tk}|$, we generate candidate random effects using the $z^*_{tk}$ (see Section \ref{s:temporal}) and a random walk Gaussian candidate distribution
 \begin{align*}
-  z_{tk}^{(c)} \sim \text{N}(z_{tk}^{(r - 1)}, s^2)
+  {z^*_{tk}}^{(c)} \sim \text{N}({z^*_{tk}}^{(r - 1)}, s^2)
 \end{align*}
-where $z_{tk}^{(r-1)}$ is the value at MCMC iteration $r - 1$, and $s$ is a tuning parameter.
+where ${z^*_{tk}}^{(r-1)}$ is the value at MCMC iteration $r - 1$, and $s$ is a tuning parameter.
 The acceptance ratio is
 \begin{align*}
-  R = \left\{ \frac{ l[Y_t(\bs) | z_{tk}^{(c)}, \ldots] }{ l[Y_t(\bs) | z_{tk}^{(r - 1)}]} \right\} \times \left\{ \frac{ p[ z_{tk}^{(c)} ] }{ p[ z_{tk}^{(r - 1)}]}\right\}.
+  R = \left\{ \frac{ l[Y_t(\bs) | z_{tk}^{(c)}, \ldots] }{ l[Y_t(\bs) | z_{tk}^{(r - 1)}]} \right\} \times \left\{ \frac{ p[ z_{tk}^{(c)} ] }{ p[ z_{tk}^{(r - 1)}]}\right\}
 \end{align*}
+where $p[\cdot]$ is the prior taken from the time series given in Section \ref{s:temporal}.
 The candidate is accepted with probability $\min\{R, 1\}$.
@@
 \end{align*}
-where $p[\cdot]$ is the prior either taken from the time series given in (\ref{s:temporal}) or assumed to be IG$(a, b)$, and $c[\cdot]$ is the candidate distribution.
+where $p[\cdot]$ is the prior either taken from the time series given in Section \ref{s:temporal} or assumed to be IG$(a, b)$, and $c[\cdot]$ is the candidate distribution.
 The candidate is accepted with probability $\min\{R, 1\}$.

### 2015-03-30 code/analysis/ozone/US-all/us-all-results.R

@@ -98,6 +98,29 @@ for (i in 1:73) {
   qs.mean.ref.gau[i, ] <- quant.score.mean[(i + 1), ] / quant.score.mean[1, ]
 }
 
+# find top two for selected quantiles
+score.compare <- bs.mean.ref.gau[, c(1, 6, 9:11)]
+idx <- which(score.compare[, 1] == min(score.compare[, 1], na.rm=T))
+score.compare[idx, 1]  # 33
+idx <- which(score.compare[, 2] == min(score.compare[, 2], na.rm=T))
+score.compare[idx, 2]
+idx <- which(score.compare[, 3] == min(score.compare[, 3], na.rm=T))
+score.compare[idx, 3]
+idx <- which(score.compare[, 4] == min(score.compare[, 4], na.rm=T))
+score.compare[idx, 4]
+idx <- which(score.compare[, 5] == min(score.compare[, 5], na.rm=T))
+score.compare[idx, 5]
+
+idx <- order(score.compare[, 1])[2]
+score.compare[idx, 1]
+idx <- order(score.compare[, 2])[2]
+score.compare[idx, 2]
+idx <- order(score.compare[, 3])[2]
+score.compare[idx, 3]
+idx <- order(score.compare[, 4])[2]
+score.compare[idx, 4]
+idx <- order(score.compare[, 5])[2]
+score.compare[idx, 5]
 # three main plots (keep max-stable in all for now)
 #   time series vs no time-series
 #   3 different threshold levels
@@ -273,7 +296,7 @@ par(mfrow=c(1, 2), mar=c(5.1, 5.1, 4.1, 2.1))
 plot(x.plot, y.plot[[1]], type="b", lty=1, ylim=c(0.9, 1),
      bg=bg[1], col=col[1], pch=pch[1], # main="Time Series Models",
      ylab="Relative Brier Score", xlab="Threshold quantile",
-     cex=1.5, cex.lab=2, cex.axis=2, cex.main=2)
+     cex=1.5, cex.lab=1.5, cex.axis=1.5, cex.main=2)
 for(i in 2:9) {
   lines(x.plot, y.plot[[i]], type="b", bg=bg[i], col=col[i], pch=pch[i],
         lty=lty[i], cex=1.5)
[…]

### 2015-03-31 LaTeX/skewnormal.tex
double spacing
updates to plots and tables

@@
 \usepackage{color}
+\usepackage{setspace}
 
@@
 \pdfpagewidth 8.5in
-\linespread{1.0}
+\linespread{2.0}
 \input{mycommands.sty}
@@
   \includegraphics[width=0.5\linewidth]{plots/chi-h.pdf}
-  \caption{$\chi(h)$ for $K = 1, 3, 5$, and $10$ partitions as a function of distance.}
+  \caption{$\chi(h)$ for $K = 1, 3, 5$, and $10$ knots as a function of distance.}
   \label{fig:chi}
@@
   \item Gaussian marginal, $K=1$ knot
-  \item $T$ marginal, $K=1$ knot
-  \item $T$ marginal, $K=5$ knots
+  \item Symmetric-$t$ marginal, $K=1$ knot
+  \item Symmetric-$t$ marginal, $K=5$ knots
   \item Skew-$t$ marginal, $K=1$ knots
@@
 \subsection{Results}\label{s:simresults}
-We compared the Brier scores for exceeding 11 different thresholds for each dataset.
-The thresholds used for the Brier scores are extreme quantiles from the simulated data.
+We compared the Brier scores for exceeding 4 different thresholds for each dataset.
+The thresholds used for the Brier scores are extreme quantiles from the simulated data for $q(0.90)$, $q(0.95)$, $q(0.98)$, and $q(0.99)$.
 Figure \ref{fig:simbrierscores} gives the Brier score relative to the Brier score for the Gaussian method calculated as
@@
 \end{align}
-We analyzed the results for the simulation study using two-sided Wilcoxon signed rank tests at a significance level of $\alpha = 0.001$.
-When the data come from a Gaussian distribution, the Gaussian method performs significantly better than the other methods for quantiles below $q(0.95)$.
-However, as we move further out in the tails, the difference in Brier scores is no longer significant beyond $q(0.95)$ when compared to method two and beyond $q(0.99)$ for methods three through five.
-When the data are generated using method two, our methods perform significantly better than the Gaussian method in all cases except for $q(0.90)$ with method five.
-For data setting three, method four performs significantly better at all quantile levels, and method five performs significantly better for quantiles more extreme than $q(0.93)$.
-In this setting, methods two and three do not show significant improvement over the Gaussian method until $q(0.995)$.
-In setting four, method two shows a significant improvement over the Gaussian method at all quantile levels.
-Methods three and four show improvement over the Gaussian method at quantile levels below $q(0.94)$.
-The Gaussian method performs significantly better than method five for all quantile levels in data setting four.
-For setting five, methods three and four perform significantly better than the Gaussian method at all quantile levels.
-Method one performs better than the Gaussian method at quantiles higher than $q(0.99)$, and method two performs better than the Gaussian method at quantiles below $q(0.98)$.
-For setting six, the Gaussian method performs significantly better than all methods for quantiles $q(0.90)$ -- $q(0.94)$.
-However, for quantiles $q(0.99)$ and $q(0.995)$, methods two and three outperform the Gaussian method.
-Finally, in setting seven, methods three demonstrate a significant improvement over the Gaussian method for all quantiles, and method two demonstrates a significant improvement of the Gaussian method for all quantiles except $q(0.94)$.
-Method four performs significantly better than the Gaussian method for $q(0.90)$, $q(0.92)$, and $q(0.95)$ -- $q(0.99)$.
-Finally, method five for setting seven shows no signifcant difference in Brier scores from the Gaussian method.
+We analyzed the results for the simulation study using a Friedman test at $\alpha = 0.05$.
+If the Friedman test came back with a significant results, we conducted a Wilcoxon-Nemenyi-McDonald-Thompson test to see which methods had different results.
+The full results for the Wilcoxon-Nemenyi-McDonald-Thompson tests are given in Appendix \ref{a:pdiffs}.
+Generally, the results of the study show that when the data come from a Gaussian process, our methods at least perform comparably to the Gaussian method.
+The results also show that our method performs better when the number of knots is the same as used in the data generation.
+
+In the Gaussian setting, the performance of the single-partition skew-$t$ method is comparable to the Gaussian method.
+Looking at Brier scores for more extreme quantiles, either skew-$t$ methods have results that are comparable to the Gaussian method.
+For data settings with symmetric-$t$ and skew-$t$ marginals (settings 2 -- 5), we find significant improvement over the Gaussian method.
+Furthermore in these data settings, we find the best performance occurs when the number of knots used in the method matches the number of knots used for data generation.
+The non-thresholded methods tend to outperform the thresholded methods, but this is not surprising given that the data are generated directly from the model used in the method.
+In setting 6, we see that for low-extreme quantiles, the Gaussian method performs better, for more extreme quantiles, the single-partition method, both thresholded and non-thresholded, perform significantly better.
+Finally, for setting 7, although the thresholded version of the single-partition model tends to perform the best across all of the extreme quantiles, the difference between the thresholded and non-thresholded methods is no longer significant in the more extreme quantiles.
 
@@
 In both the simulation study, and the application to ozone data, we find that incorporating a partition in the model improves extreme prediction.
-Furthermore the results suggest that thresholding can improve performance when predicting in the extreme tails of the data.
+Furthermore the results from the data analysis suggest that thresholding can improve performance when predicting in the extreme tails of the data.
 
@@
 Instead of implementing a time series on the random effects, a three-dimensional covariance structure on the residuals could be implemented to address temporal dependence.
+Finally, we acknowledge that by specifying the number of knots, we may be underestimating the uncertainty in the model.
+This could be incoporated by treating the number of knots as a model parameter instead of fixing it to be a specific value.
 
@@
 
+\subsection{Simulation study pairwise difference results} \label{a:pdiffs}
+The following tables show the methods that have significantly different Brier scores when using a Wilcoxon-Nemenyi-McDonald-Thompson test.
+In each column, different letters signify that the methods have significantly different Brier scores.
+For example, there is significant evidence to suggest that method 1 and method 4 have  different Brier scores at $q(0.90)$, whereas there is not significant evidence to suggest that method 1 and method 2 have different Brier scores at $q(0.90)$.
+In each table group A represents the group with the lowest Brier scores.
+Groups are significant with a familywise error rate of $\alpha = 0.05$.
+
+\begin{table}[htbp]
+  \centering
+  \caption{Setting 1: Gaussian marginal, $K = 1$ knot}
+  \label{tbl:gaussim}
+  \begin{tabular}{|l|ccc|ccc|ccc|cc|}
+    \cline{2-12}
+    \multicolumn{1}{c}{} & \multicolumn{3}{|c}{$q(0.90)$} & \multicolumn{3}{|c}{$q(0.95)$} & \multicolumn{3}{|c}{$q(0.98)$} & \multicolumn{2}{|c|}{$q(0.99)$} \\
+    \hline
+    Method 1 & A &   &   & A &   &   & A &   &   & A &   \\
+    \hline
+    Method 2 & A & B &   & A & B &   & A &   &   & A &   \\
+    \hline
+    Method 3 &   &   & C &   &   & C &   &   & C &   & B \\
+    \hline
+    Method 4 &   & B &   &   & B &   &   & B &   &   & B \\
+    \hline
+    Method 5 &   &   & C &   &   & C &   &   & C &   & B \\
+    \hline
+  \end{tabular}
+\end{table}
+
+\begin{table}[htbp]
+  \centering
+  \caption{Setting 2: Symmetric-$t$ marginal, $K = 1$ knot}
+  \label{tbl:t1k1sim}
+  \begin{tabular}{|l|cc|cccc|cccc|ccc|}
+    \cline{2-14}
+    \multicolumn{1}{c}{} & \multicolumn{2}{|c}{$q(0.90)$} & \multicolumn{4}{|c}{$q(0.95)$} & \multicolumn{4}{|c}{$q(0.98)$} & \multicolumn{3}{|c|}{$q(0.99)$} \\
+    \hline
+    Method 1 &   & B &   &   &   & D &   &   &   & D &   &   & C \\
+    \hline
+    Method 2 & A &   & A &   &   &   & A &   &   &   & A &   &   \\
+    \hline
+    Method 3 &   & B &   & B &   &   &   & B & C &   &   & B &   \\
+    \hline
+    Method 4 & A &   &   &   & C &   &   & B &   &   &   & B &   \\
+    \hline
+    Method 5 &   & B &   &   &   & D &   &   & C & D &   & B & C \\
+    \hline
+  \end{tabular}
+\end{table}
+
+\begin{table}[htbp]
+  \centering
+  \caption{Setting 3: Symmetric-$t$ marginal, $K = 5$ knots}
+  \label{tbl:gaussim}
+  \begin{tabular}{|l|cc|cc|cc|cc|}
+    \cline{2-9}
+    \multicolumn{1}{c}{} & \multicolumn{2}{|c}{$q(0.90)$} & \multicolumn{2}{|c}{$q(0.95)$} & \multicolumn{2}{|c}{$q(0.98)$} & \multicolumn{2}{|c|}{$q(0.99)$} \\
+    \hline
+    Method 1 &   & B &   & B &   & B &   & B \\
+    \hline
+    Method 2 &   & B &   & B &   & B &   & B \\
+    \hline
+    Method 3 &   & B &   & B &   & B & A & B \\
+    \hline
+    Method 4 & A &   & A &   & A &   & A &   \\
+    \hline
+    Method 5 &   & B &   & B & A & B & A & B \\
+    \hline
+  \end{tabular}
+\end{table}
+
+\begin{table}[htbp]
+  \centering
+  \caption{Setting 4: Skew-$t$ marginal, $K = 1$ knot}
+  \label{tbl:gaussim}
+  \begin{tabular}{|l|ccc|ccc|ccc|ccc|}
+    \cline{2-13}
+    \multicolumn{1}{c}{} & \multicolumn{3}{|c}{$q(0.90)$} & \multicolumn{3}{|c}{$q(0.95)$} & \multicolumn{3}{|c}{$q(0.98)$} & \multicolumn{3}{|c|}{$q(0.99)$} \\
+    \hline
+    Method 1 &   & B &   &   & B &   &   & B &   &   & B &   \\
+    \hline
+    Method 2 & A &   &   & A &   &   & A &   &   & A &   &   \\
+    \hline
+    Method 3 &   & B &   &   & B &   &   & B &   & A & B &   \\
+    \hline
+    Method 4 &   & B &   &   & B &   &   & B &   &   & B &   \\
+    \hline
+    Method 5 &   &   & C &   &   & C &   &   & C &   &   & C \\
+    \hline
+  \end{tabular}
+\end{table}
+
+\begin{table}[htbp]
+  \centering
+  \caption{Setting 5: Skew-$t$ marginal, $K = 5$ knots}
+  \label{tbl:gaussim}
+  \begin{tabular}{|l|cc|cc|cc|ccc|}
+    \cline{2-10}
+    \multicolumn{1}{c}{} & \multicolumn{2}{|c}{$q(0.90)$} & \multicolumn{2}{|c}{$q(0.95)$} & \multicolumn{2}{|c}{$q(0.98)$} & \multicolumn{3}{|c|}{$q(0.99)$} \\
+    \hline
+    Method 1 &   & B &   & B &   & B &   &   & C \\
+    \hline
+    Method 2 &   & B &   & B &   & B &   & B & C \\
+    \hline
+    Method 3 & A &   & A &   &   & B &   &   & C \\
+    \hline
+    Method 4 & A &   & A &   & A &   & A & B &   \\
+    \hline
+    Method 5 & A &   & A &   & A &   & A &   &   \\
+    \hline
+  \end{tabular}
+\end{table}
+
+\begin{table}[htbp]
+  \centering
+  \caption{Setting 6: Max-stable}
+  \label{tbl:gaussim}
+  \begin{tabular}{|l|cccc|cccc|ccc|ccc|}
+    \cline{2-15}
+    \multicolumn{1}{c}{} & \multicolumn{4}{|c}{$q(0.90)$} & \multicolumn{4}{|c}{$q(0.95)$} & \multicolumn{3}{|c}{$q(0.98)$} & \multicolumn{3}{|c|}{$q(0.99)$} \\
+    \hline
+    Method 1 & A &   &   &   & A &   &   &   & A & B &   &   &   & C \\
+    \hline
+    Method 2 &   & B &   &   & A & B &   &   & A &   &   & A & B &   \\
+    \hline
+    Method 3 &   &   & C &   &   & B & C &   & A & B &   & A &   &   \\
+    \hline
+    Method 4 &   &   &   & D &   &   &   & D &   &   & C &   &   & C \\
+    \hline
+    Method 5 &   &   & C & D &   &   & C &   &   & B &   &   & B & C \\
+    \hline
+  \end{tabular}
+\end{table}
+
+\begin{table}[htbp]
+  \centering
+  \caption{Setting 7: Transformation below $T = q(0.80)$}
+  \label{tbl:gaussim}
+  \begin{tabular}{|l|ccc|cc|ccc|ccc|}
+    \cline{2-12}
+    \multicolumn{1}{c}{} & \multicolumn{3}{|c}{$q(0.90)$} & \multicolumn{2}{|c}{$q(0.95)$} & \multicolumn{3}{|c}{$q(0.98)$} & \multicolumn{3}{|c|}{$q(0.99)$} \\
+    \hline
+    Method 1 &   &   & C &   & B &   &   & C &   &   & C \\
+    \hline
+    Method 2 &   & B &   &   & B & A & B &   & A & B &   \\
+    \hline
+    Method 3 & A &   &   & A &   & A &   &   & A &   &   \\
+    \hline
+    Method 4 &   & B & C &   & B &   & B &   &   & B & C \\
+    \hline
+    Method 5 &   & B & C &   & B &   &   & C &   &   & C \\
+    \hline
+  \end{tabular}
+\end{table}
+
 \bibliographystyle{rss}

### 2015-04-03 LaTeX/skewnormal.tex
Edits to the manuscript

@@
 \begin{center}
-{\Large {\bf A new spatial model for points above a threshold}}\\
+{\Large {\bf A spatial-skew model for extreme values}}\\
 \today
@@
 The advantage of using a thresholded model as opposed to a non-thresholded model is that is allows for the tails of the distribution to inform the predictions in the tails \citep{DuMouchel1983}.
-The random spatial partition alleviates the long-range spatial dependence seen by the skew-$t$.
+The random spatial partitions are similar to the method used by \citet{Kim2005} for non-stationary Gaussian data.
+The partition alleviates the long-range spatial dependence present in the skew-$t$ distribution.
 
 The paper is organized as follows.
-Section \ref{s:skewt} is a brief review of the spatial skew-$t$ process.
-In section \ref{s:temporal}, we build upon the traditional skew-$t$ by incorporating censoring to focus on tails, partitioning to remove long-range asymptotic dependence, and extending the model to space-time data.
-The computing is described in section \ref{s:comp}.
-In section \ref{s:simstudy}, we present a simulation study that examines the predictive capabilities of this model compared with a na{\"i}ve Gaussian method.
-We then compare our method to Gaussian and max-stable methods with a data analysis of ozone measurements from the eastern US in section \ref{s:analysis}.
+Section \ref{s:spatialskew} is a brief review of the spatial skew-$t$ process.
+In Section \ref{s:temporal}, we build upon the traditional skew-$t$ by incorporating censoring to focus on tails, partitioning to remove long-range asymptotic dependence, and extending the model to space-time data.
+The computing is described in Section \ref{s:comp}.
+In Section \ref{s:simstudy}, we present a simulation study that examines the predictive capabilities of this model compared with a na{\"i}ve Gaussian method.
+We then compare our method to Gaussian and max-stable methods with a data analysis of ozone measurements from throughout the US in section \ref{s:analysis}.
 The final section provides brief discussion and direction for future research.
@@
 \end{align}
-where $\bX(\bs)$ is a set of spatial covariates at site $\bs$, $\bbeta$ is a set of regression parameters, $\lambda \in \calR$ is a parameter controlling skew, $z \sim N(0, 1)$, $\sigma^2 \sim \text{IG}(a, b)$ is an inverse gamma random variable, and $v(\bs)$ is a spatial Gaussian process with mean zero and variance one.
-A common spatial correlation function is the \Matern with
-\begin{align}
-  \text{cor}(v(\bs), v(\bt)) = \gamma I(\bs = \bt) + (1 - \gamma)  \frac{ 1 }{ \Gamma(\nu) 2^{ \nu - 1}} \left( \sqrt{2\nu} \frac{ h }{ \rho } \right)^{\nu} K_{\nu} \left( \sqrt{2\nu} \frac{ h }{ \rho } \right) \label{eq:matern}
-\end{align}
-where $\rho$ is the spatial range, $\nu$ is the smoothness, $\gamma$ is the proportion of variance accounted for by the spatial variation, $K_\nu$ is a modified Bessel function of the second kind, and $h = || \bs - \bt ||$.
+where $\bX(\bs)$ is a set of spatial covariates at site $\bs$, $\bbeta$ is the vector of regression parameters, $\lambda \in \calR$ is a parameter controlling skew, $z \sim N(0, 1)$, $\sigma^2 \sim \text{IG}(a, b)$ is an inverse gamma random variable, and $v(\bs)$ is a spatial Gaussian process with mean zero and variance one.
 
@@
 \end{align}
-that is, $\bY$ follows an $n$-dimensional skew-$t$ process with location $\bX \bbeta$, correlation matrix $\bOmega$, slant parameters $\balpha$ and degrees of freedom $2a$, where $\bX = [\bX(\bs_1)^T, \ldots, \bX(\bs_n)^T]$, $\bOmega = \bomega \bar{\bOmega} \bomega$, $\bomega = \text{diag}\left(\frac{1}{ \sqrt{ab}}, \ldots, \frac{1}{\sqrt{ab}}\right)$, $\bar{\bOmega} = (\bSigma + \lambda^2 \bOne \bOne^T$), $\bSigma$ is a positive definite correlation matrix, $\balpha = \lambda (1 + \lambda^2 \bOne^T \bSigma^{-1} \bOne)^{-1/2} \bOne^T \bSigma^{-1}$ is a vector of slant parameters.
+that is, $\bY$ follows an $n$-dimensional skew-$t$ distribution with location $\bX \bbeta$, correlation matrix $\bOmega$, slant parameters $\balpha$ and degrees of freedom $2a$, where $\bX = [\bX(\bs_1)^T, \ldots, \bX(\bs_n)^T]$, $\bOmega = \bomega \bar{\bOmega} \bomega$, $\bomega = \text{diag}\left(\frac{1}{ \sqrt{ab}}, \ldots, \frac{1}{\sqrt{ab}}\right)$, $\bar{\bOmega} = (\bSigma + \lambda^2 \bOne \bOne^T$), $\bSigma$ is a positive definite correlation matrix, $\balpha = \lambda (1 + \lambda^2 \bOne^T \bSigma^{-1} \bOne)^{-1/2} \bOne^T \bSigma^{-1}$ is a vector of slant parameters.
+Although $\bSigma$ can be any positive definite correlation matrix, we choose to use the stationary isotropic \Matern correlation with
+\begin{align}
+  \text{cor}[v(\bs), v(\bt)] = \gamma I(\bs = \bt) + (1 - \gamma)  \frac{ 1 }{ \Gamma(\nu) 2^{ \nu - 1}} \left( \sqrt{2\nu} \frac{ h }{ \rho } \right)^{\nu} K_{\nu} \left( \sqrt{2\nu} \frac{ h }{ \rho } \right) \label{eq:matern}
+\end{align}
+where $\rho$ is the spatial range, $\nu$ is the smoothness, $\gamma$ is the proportion of variance accounted for by the spatial variation, $K_\nu$ is a modified Bessel function of the second kind, and $h = || \bs - \bt ||$.
 This process is desirable because of its flexible tail that is controlled by the skewness parameter $\lambda$ and degrees of freedom $2a$.
@@
 \subsection{Extremal dependence}
+Our interest lies in spatial dependence in the tail of the skew-$t$ process.
 One measure of extremal dependence is the $\chi$ statistic \citep{Padoan2011}.
-The $\chi$ statistic for the upper tail is given by $\lim_{c \rightarrow \infty} \Pr(Y(\bs_1) > c | Y(\bs_2) > c)$.
-For a stationary spatial process, we can write the $\chi$ coefficient as
+For a stationary and isotropic spatial process, the $\chi$ coefficient for two location $\bs$ and $\bt$ separated by distance $h = || \bs - \bt ||$ is
 \begin{align}
@@
 This occurs because all observations, both near and far, share the same $z$ and $\sigma$ terms.
-Therefore, the skew-$t$ process is not ideal for spatial analysis of large geographic regions where we expect only local spatial dependence.
+Therefore, this long-range dependence feature of the skew-$t$ process is not ideal for spatial analysis of large geographic regions where we expect only local spatial dependence.
 The explicit expression for $\chi(h)$ \citep{Padoan2011} and a proof of this are given in Appendix \ref{a:skewt}.
@@
 \section{Spatiotemporal skew-$t$ model for extremes}\label{s:spatial}
-In this section, we propose extensions to the skew-$t$ process to model spatial extremes over a large geographic region by introducing censoring to focus on tail behavior and a random partition similar to \citet{Kim2005} to remove long-range asymptotic dependence.
+In this section, we propose extensions to the skew-$t$ process to model spatial extremes over a large geographic region by introducing censoring to focus on tail behavior and a random partition to remove long-range asymptotic dependence.
+For notational convenice, we introduce the model for a single replication, and then extend this model to the spatiotemporal setting in Section \ref{s:temporal}.
 
@@
 Therefore, to alleviate this dependence, we allow $z$ and $\sigma$ to vary by site.
-Then, the model becomes
+The model becomes
 \begin{align} \label{eq:partition}
@@
 To model spatial variation, consider a set of spatial knots $\bw_1, \ldots, \bw_K$ from a homogeneous Poisson process with intensity $\mu$ over spatial domain $\calD \in \calR^2$.
-The knots define a random daily partition of $\calD$ by subregions $P_{1}, \ldots, P_{K}$ defined as
+The knots define a random partition of $\calD$ by subregions $P_{1}, \ldots, P_{K}$ defined as
 \begin{align} \label{eq:subregions}
@@
 So, within each partition, $Y(\bs)$ follows the spatial skew-$t$ process defined in Section \ref{s:spatialskew}.
-Across partitions, the $Y(\bs)$ remain correlated via the correlation function for $v(\bs)$ because it spans the partitions.
+Across partitions, the $Y(\bs)$ remain correlated via the correlation function for $v(\bs)$ because $v(\bs)$ spans all partitions.
 
@@
 \end{align}
-where $\chi_{\text{skew-}t}(h)$ is the $\chi$ statistic for a skew-$t$ process, $\chi_{\text{Gaus}}(h)$ is the $\chi$ statistic for a Gaussian process, and $h = ||\bs - \bt||$.
-Therefore, sites in different subregions are asymptotically independent.
+where $I(\cdot)$ is an indicator function, $\chi_{\text{skew-}t}(h)$ is the $\chi$ statistic for a skew-$t$ process, $\chi_{\text{Gaus}}(h)$ is the $\chi$ statistic for a Gaussian process, and $h = ||\bs - \bt||$.
+Therefore, sites in different subregions are asymptotically independent because $\chi_{\text{Gaus}}(h) = 0$ for all $h$.
 Marginally, over the knots $\bw_1, \ldots, \bw_K$, $\chi(h) = \pi(h) \chi_{\text{skew}-t}(h)$, where $\pi(h) = \Pr(k_s = k_t)$ is the probability that two sites separated by distance $h$ are in the same partition.
-So, to show that $\lim_{h \rightarrow \infty} \chi(h) = 0$, we need only know that $\lim_{h \rightarrow \infty} \pi(h) = 0$.
-A proof of this is given in Appendix \ref{a:proofsamepartition}.
-Figure \ref{fig:chi} shows how partitioning reduces the extremal dependence.
+So, to show that $\lim_{h \rightarrow \infty} \chi(h) = 0$, we need only know that $\lim_{h \rightarrow \infty} \pi(h) = 0$. A proof of this is given in Appendix \ref{a:proofsamepartition}.
+
+In Figure \ref{fig:chi}, we give $\chi(h)$ for $K = 1, 3, 5, 10$ partitions for a skew-$t$ distribution with $\alpha = 10$, and $\nu = 3$ of freedom.
+To estimate $\pi(h)$, we generate 500 sites uniformly over the unit-square.
+We then randomly generate 400 different sets of partitions using $K = 3$, $5$, and $10$.
+For each set of knots, we take $\pi(h)$ to be the proportion of sites in the same partition that are separated by distance $h$.
 
@@
 \begin{align}
-  z_t(\bs) = z_{tk}\\
-  \sigma_{t}(\bs) = \sigma_{tk}.
+  z_t(\bs) = z_{tk}\quad \text{and} \quad \sigma_{t}(\bs) = \sigma_{tk}.
 \end{align}
@@
 The time series model must be specified after a transformation to preserve the skew-$t$ process at each time point.
+For each time-varying parameter, we transform to obtain a standard normal marginal distribution, place a Gaussian prior with autocorrelation of the transformed parameter, and then transform back to obtain the marginal distribution required to preserve the skew-$t$ process.
 We first transform the spatial knots from $\calD$ to $\calR^2$ as follows.
@@
 \begin{align}
-  w^*_{tki} = \Phi^{-1}\left[ \frac{ (w_{tki} - \min(\bs_i))}{ \text{range}(\bs_i) } \right], \quad i = 1, 2
+  w^*_{tki} = \Phi^{-1}\left[ \frac{ w_{tki} - \min(\bs_i)}{ \max(\bs_i) - \min(\bs_i) } \right], \quad i = 1, 2
 \end{align}
 where $\Phi$ is a univariate standard normal density function, and $\bs_i = [s_{1i}, \ldots, s_{ni}]$.
-Then $\bw^*_{tk} \in \calR^2$.
+Then the transformed knots $\bw^*_{tk} \in \calR^2$.
 We use a copula on $\sigma^2_t(\bs)$ to ensure that the marginal distributions of $\sigma^2_t(\bs)$ are inverse gamma.
@@
 \end{align}
-where IG is the distribution function for an IG$(a, b)$ random variable.
+where IG is the distribution function of an inverse gamma random variable.
 We also use a copula on $z_{t}(\bs)$ to ensure that the marginal distributions of $z_t(\bs)$ are half-normal.
@@
 
-\subsection{Hierarchical model}\label{s:hier}
-Conditioned on $z_{tk}(\bs)$, $\sigma^2_{tk}(\bs)$, and $P_{tk}$, the marginal distributions are Gaussian and the joint distribution multivariate Gaussian.
-However, we do not fix the partitions, they are treated as unknown and updated in the MCMC.
-We model this with a Bayesian hierarchical model as follows.
-Let $\bw_{t1}, \ldots, \bw_{tK}$ be a set of daily spatial knots in a spatial domain of interest, $\calD$, and $P_{tk}$ as defined in (\ref{eq:subregions}).
-
-Then
-\begin{align}
-   Y_t(\bs) \mid z_{t}(\bs), \sigma^2(\bs), P_{tk}, \alpha, \beta, \Theta &= \bX_t(\bs)^T \beta + \lambda |z_t(\bs)| + \sigma_t(\bs) v_t(\bs) \label{eq:hier}\\
-   z_t(\bs) &= z_{tk} \text{ if } \bs \in P_{tk}\\
-   \sigma^2_{t}(\bs) &= \sigma^2_{tk} \text{ if } \bs \in P_{tk}\\
-   \lambda &= \lambda_1 \lambda_2\\
-   \lambda_1 &= \left\{ \begin{array}{ll}
-      +1 \quad & \text{w.p.} 0.5\\
-      -1 \quad & \text{w.p.} 0.5
-   \end{array}\right.\\
-   \lambda^2_2 & \sim IG(a, b)\\
-   v_t(\bs) \mid \Theta &\sim \Matern(0, \Sigma)\\
-   z^*_{tk} \mid z^*_{t-1, k}, \sigma^2_{tk} &\sim N(\phi_z z^*_{t-1, k}, \sigma^2_{tk} (1 - \phi_z^2))\\
-   \sigma^{2*}_{tk} \mid \sigma^{2*}_{t-1, k} &\sim N(\phi_\sigma \sigma^{2*}_{t-1, k}, (1 - \phi_\sigma^2))\\
-   \bw^*_{tk} \mid \bw^*_{t-1, k} &\sim N_2(\phi_w \bw^*_{t-1, k}, (1 - \phi_w^2))
-\end{align}
-where $\Theta = \{\rho, \nu, \gamma\}$, and $\Sigma$ is a \Matern covariance matrix as described in Section \ref{s:skewt}.
-We parameterize $\lambda = \lambda_1 \lambda_2$ to help with convergence in the MCMC.
-
 \section{Computation}\label{s:comp}
@@
 
+\subsection{Hierarchical model}\label{s:hier}
+Conditioned on $z_{tk}(\bs)$, $\sigma^2_{tk}(\bs)$, and $P_{tk}$, the marginal distributions are Gaussian and the joint distribution multivariate Gaussian.
+However, we do not fix the partitions, they are treated as unknown and updated in the MCMC.
+We model this with a Bayesian hierarchical model as follows.
+Let $\bw_{t1}, \ldots, \bw_{tK}$ be a set of daily spatial knots in a spatial domain of interest, $\calD$, and $P_{tk}$ as defined in (\ref{eq:subregions}).
+Then
+\begin{align}
+   Y_t(\bs) \mid z_{t}(\bs), \sigma^2(\bs), P_{tk}, \alpha, \beta, \Theta &= \bX_t(\bs)^T \beta + \lambda |z_t(\bs)| + \sigma_t(\bs) v_t(\bs) \label{eq:hier}\\
+   z_t(\bs) &= z_{tk} \text{ if } \bs \in P_{tk}\nonumber\\
+   \sigma^2_{t}(\bs) &= \sigma^2_{tk} \text{ if } \bs \in P_{tk}\nonumber\\
+   \lambda &= \lambda_1 \lambda_2\nonumber\\
+   \lambda_1 &= \left\{ \begin{array}{ll}
+      +1 \quad & \text{w.p.} 0.5\\
+      -1 \quad & \text{w.p.} 0.5
+   \end{array}\right.\nonumber\\
+   \lambda^2_2 & \sim IG(a, b)\nonumber\\
+   v_t(\bs) \mid \Theta &\sim \Matern(0, \Sigma)\nonumber\\
+   z^*_{tk} \mid z^*_{t-1, k}, \sigma^2_{tk} &\sim N(\phi_z z^*_{t-1, k}, \sigma^2_{tk} (1 - \phi_z^2))\nonumber\\
+   \sigma^{2*}_{tk} \mid \sigma^{2*}_{t-1, k} &\sim N(\phi_\sigma \sigma^{2*}_{t-1, k}, (1 - \phi_\sigma^2))\nonumber\\
+   \bw^*_{tk} \mid \bw^*_{t-1, k} &\sim N_2(\phi_w \bw^*_{t-1, k}, (1 - \phi_w^2)) \nonumber
+\end{align}
+where $\Theta = \{\rho, \nu, \gamma\}$, and $\Sigma$ is a \Matern covariance matrix as described in Section \ref{s:skewt}.
+We parameterize $\lambda = \lambda_1 \lambda_2$ to help with convergence in the MCMC.
+
 \section{Simulation study}\label{s:simstudy}
@@
   \item Gaussian marginal, $K=1$ knot
-  \item Symmetric-$t$ marginal, $K=1$ knot
-  \item Symmetric-$t$ marginal, $K=5$ knots
+  % \item Symmetric-$t$ marginal, $K=1$ knot
+  % \item Symmetric-$t$ marginal, $K=5$ knots
   \item Skew-$t$ marginal, $K=1$ knots
@@
 \end{enumerate}
-In the first five designs, the $v_t(\bs)$ terms are generated using a \Matern covariance with smoothness parameter $\nu = 0.5$ and spatial range $\rho = 1$.
-For the covariance matrices in designs 1 -- 5, the proportion of the variance accounted for by the spatial variation is $\gamma = 0.9$ while the proportion of the variance accounted for by the nugget effect is $0.1$.
-In the first design, $\sigma^2 = 2$ is used for all days.
-For designs 2 -- 4, $\sigma^2_{tk} \iid \text{IG}(3, 8)$
-For designs 1 -- 3, we set $\lambda = 0$.
-For designs four and five, $\lambda = 3$ was used, and the $z_t$ are generated as described in (\ref{eq:sitez}).
-In the sixth design, we generate from a spatial max-stable distribution \citep{Reich2012}.
+In the first three designs, the $v_t(\bs)$ terms are generated using a \Matern covariance with smoothness parameter $\nu = 0.5$ and spatial range $\rho = 1$.
+For the covariance matrices in designs 1 -- 3, the proportion of the variance accounted for by the spatial variation is $\gamma = 0.9$ while the proportion of the variance accounted for by the nugget effect is $0.1$.
+In the first design, $\sigma^2 = 2$ is used for all days which results in a Gaussian distribution.
+For designs 2 and 3, $\sigma^2_{tk} \iid \text{IG}(3, 8)$
+For designs 1, we set $\lambda = 0$.
+For designs 2 and 3, $\lambda = 3$ was used as to simulate moderate skewness, and the $z_t$ are generated as described in (\ref{eq:sitez}).
+In the fourth design, we generate from a spatial max-stable distribution \citep{Reich2012}.
 In this design, data have marginal distributions that follow a generalized extreme value distribution with parameters $\mu = 1, \sigma=1, \xi=0.2$.
@@
 For this setting, we set $\gamma = 0.5$.
-In the final design, we generate $\tilde{y}$ using the setting from design 4, and then consider the data
+In the final design, we generate $\tilde{y}$ using the setting from design two, and then consider the data
 \begin{align}
@@
 where $T = q(0.80)$ is the 80th sample quantile of the data.
-In all seven designs, the mean $\bX \bbeta = 10$ is assumed to be constant across space.
+In all five designs, the mean $\bX^T \bbeta = 10$ is assumed to be constant across space.
 
 $M = 50$ data sets are generated for each design.
-For each data set we fit the data using
+For each data set we fit the data using five models
 \begin{enumerate} \setlength{\itemsep}{-0.5em}
@@
 The knots have priors $\bw \sim \text{Unif}(\calD)$.
-We do not fit the data using the max-stable methods from \citet{Reich2012} because of the time it takes.
+We do not include skewness for the thresholded models because it cannot be identified with only one tail worth of data, and we do not fit the data using the max-stable methods from \citet{Reich2012} because of the computational constraints.
+Each chain of the mcmc ran for 20000 iterations with a burn-in period of 10000 iterations.
+Parameters appear to converge properly; however, in the models with multiple partitions (i.e.models 4 and 5) it is hard to assess the convergence of $\bw$, $z(\bs)$, and $\sigma^2(\bs)$ because of partition label switching throughout the MCMC.
 
@@
 The full results for the Wilcoxon-Nemenyi-McDonald-Thompson tests are given in Appendix \ref{a:pdiffs}.
-Generally, the results of the study show that when the data come from a Gaussian process, our methods at least perform comparably to the Gaussian method.
-The results also show that our method performs better when the number of knots is the same as used in the data generation.
 
-In the Gaussian setting, the performance of the single-partition skew-$t$ method is comparable to the Gaussian method.
-Looking at Brier scores for more extreme quantiles, either skew-$t$ methods have results that are comparable to the Gaussian method.
-For data settings with symmetric-$t$ and skew-$t$ marginals (settings 2 -- 5), we find significant improvement over the Gaussian method.
+Figure \ref{fig:simbrierscores} shows that when the data come from a Gaussian process, our methods perform comparably to the Gaussian method.
+For data settings with skew-$t$ marginals (settings 2 -- 5), we find significant improvement over the Gaussian method.
 Furthermore in these data settings, we find the best performance occurs when the number of knots used in the method matches the number of knots used for data generation.
 The non-thresholded methods tend to outperform the thresholded methods, but this is not surprising given that the data are generated directly from the model used in the method.
-In setting 6, we see that for low-extreme quantiles, the Gaussian method performs better, for more extreme quantiles, the single-partition method, both thresholded and non-thresholded, perform significantly better.
-Finally, for setting 7, although the thresholded version of the single-partition model tends to perform the best across all of the extreme quantiles, the difference between the thresholded and non-thresholded methods is no longer significant in the more extreme quantiles.
+For the max-stable data, we see that for low-extreme quantiles, the Gaussian method performs better, for more extreme quantiles, the single-partition method, both thresholded and non-thresholded, perform significantly better than the Gaussian.
+Finally, for setting 5, although the thresholded version of the single-partition model tends to perform the best across all of the extreme quantiles, the difference between the thresholded and non-thresholded methods is no longer significant in the more extreme quantiles.
 
@@
   \includegraphics[width=\linewidth]{plots/bsplots-mean.pdf}
-  \caption{Brier scores relative to the Gaussian method for simulation study results. A ratio lower than 1 indicates that the method outperforms relative to the Gaussian method.}
+  \caption{Brier scores relative to the Gaussian method for simulation study results. A ratio lower than 1 indicates that the method outperforms the Gaussian method.}
   \label{fig:simbrierscores}
@@
 \end{figure}
+Exploratory data analysis indicates that there is dependence in the high quantile levels of the residuals beyond what we expect if the residuals were independent.
 
-We explore spatial and temporal extremal dependence by considering $\chi_c = \Pr[Y(\bs) > c | Y(\bt) > c]$.
-To examine spatial dependence in high quantiles, we consider observations at all pairs of sites $\bs$ and $\bt$ that are distance $h$ apart where $h$ is separated into bins of size 0.25 km.
-Then conditioned on $Y(\bt) > c$, we take the sample proportion of $Y(\bs) > c$.
-Finally, $\widehat{\chi}_c(h)$ is averaged over all days at each of the three threshold quantiles.
-To examine temporal dependence in high quantiles, we consider observations at a single site that are taken lag-$t$ days apart.
-Then conditioned on $Y_n(\bs) > c$, we take the sample proportion of $Y_{n + t}(\bs) > c$.
-Finally, $\widehat{\chi}(t)$ is averaged over all sites at each of the three threshold quantiles.
-The $\widehat{\chi}_c(h)$ and $\widehat{\chi}_c(t)$ plots in Figure \ref{fig:chi-st} show the estimated spatial and temporal dependence of the residuals for the ozone data at three quantile levels $q(0.90), q(0.95)$, and $q(0.99)$.
-Both plots indicate that there is dependence in the high quantile levels beyond what we expect if the residuals were independent.
-\begin{figure}
-  \centering
-  \includegraphics[width=0.49\linewidth]{plots/chi-h-ozone.pdf}
-  \includegraphics[width=0.49\linewidth]{plots/chi-t-ozone.pdf}
-  \caption{$\widehat{\chi}_c(h)$ plot for the residuals (left). $\widehat{\chi}_c(t)$ plot for the residuals (right).}
-  \label{fig:chi-st}
-\end{figure}
+% We explore spatial and temporal extremal dependence by considering $\chi_c = \Pr[Y(\bs) > c | Y(\bt) > c]$.
+% To examine spatial dependence in high quantiles, we consider observations at all pairs of sites $\bs$ and $\bt$ that are distance $h$ apart where $h$ is separated into bins of size 0.25 km.
+% Then conditioned on $Y(\bt) > c$, we take the sample proportion of $Y(\bs) > c$.
+% Finally, $\widehat{\chi}_c(h)$ is averaged over all days at each of the three threshold quantiles.
+% To examine temporal dependence in high quantiles, we consider observations at a single site that are taken lag-$t$ days apart.
+% Then conditioned on $Y_n(\bs) > c$, we take the sample proportion of $Y_{n + t}(\bs) > c$.
+% Finally, $\widehat{\chi}(t)$ is averaged over all sites at each of the three threshold quantiles.
+% The $\widehat{\chi}_c(h)$ and $\widehat{\chi}_c(t)$ plots in Figure \ref{fig:chi-st} show the estimated spatial and temporal dependence of the residuals for the ozone data at three quantile levels $q(0.90), q(0.95)$, and $q(0.99)$.
+
+% \begin{figure}
+%   \centering
+%   \includegraphics[width=0.49\linewidth]{plots/chi-h-ozone.pdf}
+%   \includegraphics[width=0.49\linewidth]{plots/chi-t-ozone.pdf}
+%   \caption{$\widehat{\chi}_c(h)$ plot for the residuals (left). $\widehat{\chi}_c(t)$ plot for the residuals (right).}
+%   \label{fig:chi-st}
+% \end{figure}
 
@@
 We fit the model using Gaussian and skew-$t$ marginal distributions with $K=1, 5, 6, 7, 8, 9, 10, 15$ partitions.
-We choose to censor $Y(\bs)$ at $T = 0$, $\widehat{q}(0.42) = 50$, $\widehat{q}(0.92)=75$ ppb in order to compare results from no, moderate, and high censoring.
+We choose to censor $Y(\bs)$ at $T = $0, 50 (0.42 sample quantile), and 75 (0.92 sample quantile) ppb in order to compare results from no, moderate, and high censoring.
+The upper thresholde of 75 ppb was used because the current air quality standard is based on exceedance of 75 ppb.
 We also compare models with no time series to models that include the time series.
@@
 All methods assume the mean function given in (\ref{eq:datamean}).
+\hl{number of iterations, burnin, convergence}
 For each model, Brier scores were were averaged over all sites and days to obtain a single Brier score for each dataset.
 At a particular threshold or quantile level, the model that fits the best is the one with the lowest score.
-We then compute the relative Brier scores to compare each model.
+We then compute the relative (to Gaussian) Brier scores to compare each model.
 
 \subsection{Results}\label{s:results}
-The plots in Figure \ref{fig:bs-ozone} shows the relative Brier scores (see Section \ref{s:simresults}) for time-series and non-time-series models, using $K = $1, 7, and 15 knots at thresholds $T = $0, 50, and 75 ppb.
+\hl{Include more here. 1: mention something about the max-stable results (pull from figure caption), 2: discuss general take home points}
+The results demonstrate the importance of accounting for the temporal dependence when making extreme predictions.
+They also show the importance of thresholding for predictions further out in the tails of the data.
+This is further illustrated by the plots in Figure \ref{fig:ozoneq99}.
+
+The plots in Figure \ref{fig:bs-ozone} shows the relative Brier scores (see Section \ref{s:simresults}) for time-series and non-time-series models, using $K = $ 1, 7, and 15 knots at thresholds $T = $ 0, 50, and 75 ppb.
 Both plots have similar features suggesting that most settings do reasonably well.
 In particular, for all extreme quantiles, selecting a moderate number of knots (e.g. $K = 5, \ldots, 10$) tends to give the best results.
+
+Table \ref{tbl:ozoneresults} shows the best two models for selected extreme quantiles.
+
 \begin{figure}
@@
 \end{figure}
-Table \ref{tbl:ozoneresults} shows the best two models for selected extreme quantiles.
-The results demonstrate the importance of accounting for the temporal dependence when making extreme predictions.
-They also show the importance of thresholding for predictions further out in the tails of the data.
-This is further illustrated by the plots in Figure \ref{fig:ozoneq99}.
-
 \begin{table}
@@
 In this paper we propose a new approach for spatitemporal modeling of extreme values.
-The proposed model gives flexible tail behavior, demonstrates asymptotic dependence for observations at sites that are near to one another, \hl{and has computation on the order of Gaussian models for large space-time datasets -- Do I need to add anything about the computation in the paper?}.
+The proposed model gives flexible tail behavior, demonstrates asymptotic dependence for observations at sites that are near to one another, and has computation on the order of Gaussian models for large space-time datasets.
 In the simulation study, we demonstrate that this model shows statistically significant improvements over a na\"{i}ve Gaussian approach.
@@
   \centering
-  \caption{Setting 1: Gaussian marginal, $K = 1$ knot}
+  \caption{Setting 1 -- Gaussian marginal, $K = 1$ knot}
   \label{tbl:gaussim}
@@
 
-\begin{table}[htbp]
-  \centering
-  \caption{Setting 2: Symmetric-$t$ marginal, $K = 1$ knot}
-  \label{tbl:t1k1sim}
-  \begin{tabular}{|l|cc|cccc|cccc|ccc|}
-    \cline{2-14}
-    \multicolumn{1}{c}{} & \multicolumn{2}{|c}{$q(0.90)$} & \multicolumn{4}{|c}{$q(0.95)$} & \multicolumn{4}{|c}{$q(0.98)$} & \multicolumn{3}{|c|}{$q(0.99)$} \\
-    \hline
-    Method 1 &   & B &   &   &   & D &   &   &   & D &   &   & C \\
-    \hline
-    Method 2 & A &   & A &   &   &   & A &   &   &   & A &   &   \\
-    \hline
-    Method 3 &   & B &   & B &   &   &   & B & C &   &   & B &   \\
-    \hline
-    Method 4 & A &   &   &   & C &   &   & B &   &   &   & B &   \\
-    \hline
-    Method 5 &   & B &   &   &   & D &   &   & C & D &   & B & C \\
-    \hline
-  \end{tabular}
-\end{table}
+% \begin{table}[htbp]
+%   \centering
+%   \caption{Setting 2: Symmetric-$t$ marginal, $K = 1$ knot}
+%   \label{tbl:t1k1sim}
+%   \begin{tabular}{|l|cc|cccc|cccc|ccc|}
+%     \cline{2-14}
+%     \multicolumn{1}{c}{} & \multicolumn{2}{|c}{$q(0.90)$} & \multicolumn{4}{|c}{$q(0.95)$} & \multicolumn{4}{|c}{$q(0.98)$} & \multicolumn{3}{|c|}{$q(0.99)$} \\
+%     \hline
+%     Method 1 &   & B &   &   &   & D &   &   &   & D &   &   & C \\
+%     \hline
+%     Method 2 & A &   & A &   &   &   & A &   &   &   & A &   &   \\
+%     \hline
+%     Method 3 &   & B &   & B &   &   &   & B & C &   &   & B &   \\
+%     \hline
+%     Method 4 & A &   &   &   & C &   &   & B &   &   &   & B &   \\
+%     \hline
+%     Method 5 &   & B &   &   &   & D &   &   & C & D &   & B & C \\
+%     \hline
+%   \end{tabular}
+% \end{table}
+
+% \begin{table}[htbp]
+%   \centering
+%   \caption{Setting 3: Symmetric-$t$ marginal, $K = 5$ knots}
+%   \label{tbl:gaussim}
+%   \begin{tabular}{|l|cc|cc|cc|cc|}
+%     \cline{2-9}
+%     \multicolumn{1}{c}{} & \multicolumn{2}{|c}{$q(0.90)$} & \multicolumn{2}{|c}{$q(0.95)$} & \multicolumn{2}{|c}{$q(0.98)$} & \multicolumn{2}{|c|}{$q(0.99)$} \\
+%     \hline
+%     Method 1 &   & B &   & B &   & B &   & B \\
+%     \hline
+%     Method 2 &   & B &   & B &   & B &   & B \\
+%     \hline
+%     Method 3 &   & B &   & B &   & B & A & B \\
+%     \hline
+%     Method 4 & A &   & A &   & A &   & A &   \\
+%     \hline
+%     Method 5 &   & B &   & B & A & B & A & B \\
+%     \hline
+%   \end{tabular}
+% \end{table}
 
@@
   \centering
-  \caption{Setting 3: Symmetric-$t$ marginal, $K = 5$ knots}
-  \label{tbl:gaussim}
-  \begin{tabular}{|l|cc|cc|cc|cc|}
-    \cline{2-9}
-    \multicolumn{1}{c}{} & \multicolumn{2}{|c}{$q(0.90)$} & \multicolumn{2}{|c}{$q(0.95)$} & \multicolumn{2}{|c}{$q(0.98)$} & \multicolumn{2}{|c|}{$q(0.99)$} \\
-    \hline
-    Method 1 &   & B &   & B &   & B &   & B \\
-    \hline
-    Method 2 &   & B &   & B &   & B &   & B \\
-    \hline
-    Method 3 &   & B &   & B &   & B & A & B \\
-    \hline
-    Method 4 & A &   & A &   & A &   & A &   \\
-    \hline
-    Method 5 &   & B &   & B & A & B & A & B \\
-    \hline
-  \end{tabular}
-\end{table}
-
-\begin{table}[htbp]
-  \centering
-  \caption{Setting 4: Skew-$t$ marginal, $K = 1$ knot}
+  \caption{Setting 2 -- Skew-$t$ marginal, $K = 1$ knot}
   \label{tbl:gaussim}
@@
   \centering
-  \caption{Setting 5: Skew-$t$ marginal, $K = 5$ knots}
+  \caption{Setting 3 -- Skew-$t$ marginal, $K = 5$ knots}
   \label{tbl:gaussim}
@@
   \centering
-  \caption{Setting 6: Max-stable}
+  \caption{Setting 4 -- Max-stable}
   \label{tbl:gaussim}
@@
   \centering
-  \caption{Setting 7: Transformation below $T = q(0.80)$}
+  \caption{Setting 5 -- Transformation below $T = q(0.80)$}
   \label{tbl:gaussim}

### 2015-04-03 LaTeX/skewnormal.tex
updates to the manuscript

@@
 One measure of extremal dependence is the $\chi$ statistic \citep{Padoan2011}.
-For a stationary and isotropic spatial process, the $\chi$ coefficient for two location $\bs$ and $\bt$ separated by distance $h = || \bs - \bt ||$ is
+For a stationary and isotropic spatial process, the $\chi$ statistic for two location $\bs$ and $\bt$ separated by distance $h = || \bs - \bt ||$ is
 \begin{align}
@@
 Unlike the Gaussian process, the skew-$t$ process is asymptotically dependent.
-However, one problem with the spatial skew-$t$ process is that $\lim_{h \rightarrow \infty} \chi(h) > 0$.
+However, one problem with the spatial skew-$t$ process is that $\lim_{h \rightarrow \infty} \chi(h) > 0$ (see Appendix \ref{a:skewt} for a proof).
 This occurs because all observations, both near and far, share the same $z$ and $\sigma$ terms.
 Therefore, this long-range dependence feature of the skew-$t$ process is not ideal for spatial analysis of large geographic regions where we expect only local spatial dependence.
-The explicit expression for $\chi(h)$ \citep{Padoan2011} and a proof of this are given in Appendix \ref{a:skewt}.
+The explicit expression for $\chi(h)$ is given in Appendix \ref{a:skewt}.
 
@@
 \beq\label{Yt}
-  \tilde{Y}(\bs) = \left\{ \begin{array}{ll}
+  \widetilde{Y}(\bs) = \left\{ \begin{array}{ll}
       Y(\bs) \quad & \delta(\bs) = 1 \\
@@
 \subsection{Partitioning to remove long-range asymptotic dependence}\label{s:part}
-We handle the problem of long-range asymptotic dependence with a random partition model.
+We handle the problem of long-range asymptotic dependence with a random partition.
 As discussed in Section \ref{s:spatialskew}, the source of long-range dependence is the shared $z$ and $\sigma$.
@@
 \begin{align}
-  z(\bs) &= z_{k} \label{eq:sitez}\\
-  \sigma(\bs) &= \sigma_{k} \label{eq:sitesig},
+  z(\bs) = z_{k} \quad \text{and} \quad \sigma(\bs) = \sigma_{k} \label{eq:sitezsig}
 \end{align}
-and the $z_k$ and $\sigma^2_k$ are distributed as $z_k \iid N(0, 1)$ and $\sigma^2 \iid$ IG$(a, b)$.
+and the $z_k$ and $\sigma^2_k$ are distributed as $z_k \iid N(0, 1)$ and $\sigma^2 \iid$ IG$(a, b)$ where IG is the distribution function of an inverse gamma random variable.
 So, within each partition, $Y(\bs)$ follows the spatial skew-$t$ process defined in Section \ref{s:spatialskew}.
@@
 
-When incorporating the random daily partition, conditional on knots $\bw_1, \ldots, \bw_K$, the $\chi$ statistic for two sites $\bs$ and $\bt$ in partitions $k_s$ and $k_t$ respectively is
+When incorporating the random partition, conditional on knots $\bw_1, \ldots, \bw_K$, the $\chi$ statistic for two sites $\bs$ and $\bt$ in partitions $k_s$ and $k_t$ respectively is
 \begin{align}
@@
 
-In Figure \ref{fig:chi}, we give $\chi(h)$ for $K = 1, 3, 5, 10$ partitions for a skew-$t$ distribution with $\alpha = 10$, and $\nu = 3$ of freedom.
+In Figure \ref{fig:chi}, we give $\chi(h)$ for $K = 1, 3, 5, 10$ partitions for a skew-$t$ distribution with $\alpha = 10$, and 3 degrees of freedom.
 To estimate $\pi(h)$, we generate 500 sites uniformly over the unit-square.
@@
 For each set of knots, we take $\pi(h)$ to be the proportion of sites in the same partition that are separated by distance $h$.
+This plot demonstrates how partitioning helps to reduce extremal dependence as $h$ increases.
 
@@
 The time series model must be specified after a transformation to preserve the skew-$t$ process at each time point.
-For each time-varying parameter, we transform to obtain a standard normal marginal distribution, place a Gaussian prior with autocorrelation of the transformed parameter, and then transform back to obtain the marginal distribution required to preserve the skew-$t$ process.
+For each time-varying parameter, we transform to obtain a standard normal marginal distribution, place a Gaussian prior with autocorrelation on the transformed parameter, and then transform back to obtain the marginal distribution required to preserve the skew-$t$ process.
 We first transform the spatial knots from $\calD$ to $\calR^2$ as follows.
@@
 \end{align}
-where IG is the distribution function of an inverse gamma random variable.
+where IG is defined as before.
 We also use a copula on $z_{t}(\bs)$ to ensure that the marginal distributions of $z_t(\bs)$ are half-normal.
@@
 \begin{align}
-  z^*_t(\bs) = \Phi^{-1}\left\{ \text{HN}[\sigma^2_t(\bs)] \right\}
+  z^*_t(\bs) = \Phi^{-1}\left\{ \text{HN}[z_t(\bs)] \right\}
 \end{align}
+where HN is the distribution function of a half-normal random variable.
 The AR(1) process for each tail parameter is $\bw^*_{1k} \sim N_w(0, 1)$, $z^*_{1k} \sim N(0, \sigma^2_{1k})$, $\sigma^{2*}_{1k} \sim N(0, 1)$, and for $t > 1$ the time series is modeled as
@@
 where $|\phi_w|$, $|\phi_z|$, $|\phi_\sigma| < 1$.
-These are stationary time series models with marginal distributions $\bw^*_{k} \sim N_2(0, 1)$, $z^*_{k} \sim N(0, \sigma^2_{k})$, $\sigma^{2*}_{k} \sim N(0, 1)$.
+These are stationary time series models with marginal distributions \hbox{$\bw^*_{k} \sim N_2(0, 1)$}, \hbox{$z^*_{k} \sim N(0, \sigma^2_{k})$}, and \hbox{$\sigma^{2*}_{k} \sim N(0, 1)$}.
 After transformation back to the original space, $\bw_{tk} \sim \text{Unif}(\calD)$, $z_{tk} \sim HN(0, \sigma^2_{tk})$ $\sigma^2_{tk} \sim \text{IG}(a, b)$.
@@
 First, we impute values below the threshold.
-Then, we update $\Theta$ using Metropolis Hastings or Gibbs sampling when appropriate.
+Then, we update model parameters, $\Theta$, using Metropolis Hastings or Gibbs sampling when appropriate.
 Finally, we make spatial predictions using conditional multivariate normal results and the fact that the distribution of $Y_t(\bs) \mid \Theta, z(\bs)$ is the usual multivariate normal distribution with a \Matern spatial covariance structure.
@@
 After conditioning on $\lambda$, $z_t(\bs)$ and non-censored observations, $Y_t(\bs)$ has truncated normal full conditionals.
-So we sample $Y_t(\bs) \sim N_{(-\infty, T)}(\bX \bbeta, \bSigma)$.
+So we sample $Y_t(\bs) \sim N_{(-\infty, T)}(\bX_t^T(\bs) \beta + \lambda | z_t(\bs)|, \bSigma)$.
 After imputing the censored observations, we update the model parameters.
@@
 \begin{align}
-   Y_t(\bs) \mid z_{t}(\bs), \sigma^2(\bs), P_{tk}, \alpha, \beta, \Theta &= \bX_t(\bs)^T \beta + \lambda |z_t(\bs)| + \sigma_t(\bs) v_t(\bs) \label{eq:hier}\\
+   Y_t(\bs) \mid z_{t}(\bs), \sigma_t^2(\bs), P_{tk}, \Theta &= \bX_t(\bs)^T \beta + \lambda |z_t(\bs)| + \sigma_t(\bs) v_t(\bs) \label{eq:hier}\\
    z_t(\bs) &= z_{tk} \text{ if } \bs \in P_{tk}\nonumber\\
@@
    \lambda_1 &= \left\{ \begin{array}{ll}
-      +1 \quad & \text{w.p.} 0.5\\
-      -1 \quad & \text{w.p.} 0.5
+      +1 \quad & \text{w.p. } 0.5\\
+      -1 \quad & \text{w.p. } 0.5
    \end{array}\right.\nonumber\\
@@
 \end{align}
-where $\Theta = \{\rho, \nu, \gamma\}$, and $\Sigma$ is a \Matern covariance matrix as described in Section \ref{s:skewt}.
+where $\Theta = \{\rho, \nu, \gamma\, \lambda, \beta\}$, and $\Sigma$ is a \Matern covariance matrix as described in Section \ref{s:skewt}.
 We parameterize $\lambda = \lambda_1 \lambda_2$ to help with convergence in the MCMC.
@@
 The sites are generated Uniform$([0, 10] \times [0, 10])$.
-We generate data from 7 different simulation designs:
+We generate data from 5 different simulation designs:
 \begin{enumerate} \setlength{\itemsep}{-0.5em}
@@
 In the first design, $\sigma^2 = 2$ is used for all days which results in a Gaussian distribution.
-For designs 2 and 3, $\sigma^2_{tk} \iid \text{IG}(3, 8)$
+For designs 2 and 3, $\sigma^2_{tk} \iid \text{IG}(3, 8)$ to give a $t$ distribution with 6 degrees of freedom.
 For designs 1, we set $\lambda = 0$.
-For designs 2 and 3, $\lambda = 3$ was used as to simulate moderate skewness, and the $z_t$ are generated as described in (\ref{eq:sitez}).
+For designs 2 and 3, $\lambda = 3$ was used as to simulate moderate skewness, and the $z_t$ are generated as described in (\ref{eq:sitezsig}).
 In the fourth design, we generate from a spatial max-stable distribution \citep{Reich2012}.
@@
 where $q(0.80)$ is the 80th sample quantile of the data.
-The design matrix $\bX$ includes an the intercept with a prior of $\beta \sim \text{N}(0, 10)$.
+The design matrix $\bX$ includes an intercept with a first-order spatial trend with priors of $\beta_\text{int}$, $\beta_\text{lat}$, $\beta_\text{long},  \iid \text{N}(0, 10)$.
 The spatial covariance parameters have priors $\log(\nu) \sim \text{N}(-1.2, 1)$, $\gamma \sim \text{Unif}(0, 1)$, $\rho \sim \text{Unif}(15)$.
-The skewness parameter has prior $\lambda \sim \text{N}(0, 2)$.
+The skewness parameter has prior $\lambda_2 \sim \text{IG}(0.1, 0.1)$.
 The residual variance terms have priors $\sigma^2_t(\bs) \sim \text{IG}(0.1, 0.1)$.
@@
 Figure \ref{fig:simbrierscores} shows that when the data come from a Gaussian process, our methods perform comparably to the Gaussian method.
-For data settings with skew-$t$ marginals (settings 2 -- 5), we find significant improvement over the Gaussian method.
+For data settings with skew-$t$ marginals (settings 2 -- 3), we find significant improvement over the Gaussian method.
 Furthermore in these data settings, we find the best performance occurs when the number of knots used in the method matches the number of knots used for data generation.
@@
 \begin{align}
-  \bX \bbeta = \beta_0 + \beta_1 \cdot \text{CMAQ}_t(\bs). \label{eq:datamean}
+  \bX^T_t(\bs) \bbeta = \beta_0 + \beta_1 \cdot \text{CMAQ}_t(\bs). \label{eq:datamean}
 \end{align}
@@
 \end{figure}
-Exploratory data analysis indicates that there is dependence in the high quantile levels of the residuals beyond what we expect if the residuals were independent.
+Exploratory data analysis indicates that there is dependence in the high quantile levels of the residuals beyond what we expect in the case if independence.
 
@@
 We fit the model using Gaussian and skew-$t$ marginal distributions with $K=1, 5, 6, 7, 8, 9, 10, 15$ partitions.
-We choose to censor $Y(\bs)$ at $T = $0, 50 (0.42 sample quantile), and 75 (0.92 sample quantile) ppb in order to compare results from no, moderate, and high censoring.
-The upper thresholde of 75 ppb was used because the current air quality standard is based on exceedance of 75 ppb.
+We choose to censor $Y(\bs)$ at $T =$ 0, 50 (0.42 sample quantile), and 75 (0.92 sample quantile) ppb in order to compare results from no, moderate, and high censoring.
+The upper threshold of 75 ppb was used because the current air quality standard is based on exceedance of 75 ppb.
+As with the simuliation study, for models with a threshold of $T = 75$, we use a symmetric-$t$ marginal distribution.
 We also compare models with no time series to models that include the time series.
@@
 All methods assume the mean function given in (\ref{eq:datamean}).
-\hl{number of iterations, burnin, convergence}
+Each chain of the mcmc ran for 30000 iterations with a burn-in period of 25000 iterations.
+Parameters appear to converge properly; however, as before, for models with multiple partitions it is hard to assess the convergence of $\bw$, $z(\bs)$, and $\sigma^2(\bs)$ because of partition label switching throughout the MCMC.
 For each model, Brier scores were were averaged over all sites and days to obtain a single Brier score for each dataset.
 At a particular threshold or quantile level, the model that fits the best is the one with the lowest score.
-We then compute the relative (to Gaussian) Brier scores to compare each model.
+We then compute the relative (to Gaussian) Brier scores (see Section \ref{s:simresults}) to compare each model.
 
 \subsection{Results}\label{s:results}
-\hl{Include more here. 1: mention something about the max-stable results (pull from figure caption), 2: discuss general take home points}
-The results demonstrate the importance of accounting for the temporal dependence when making extreme predictions.
-They also show the importance of thresholding for predictions further out in the tails of the data.
-This is further illustrated by the plots in Figure \ref{fig:ozoneq99}.
-
-The plots in Figure \ref{fig:bs-ozone} shows the relative Brier scores (see Section \ref{s:simresults}) for time-series and non-time-series models, using $K = $ 1, 7, and 15 knots at thresholds $T = $ 0, 50, and 75 ppb.
+The results suggest that the skew-$t$, thresholded, partitioned, and time series models all give an improvement in predictions over the Gaussian model, whereas the max-stable method results in relative Brier scores between 1.07 and 1.15 indicating poorer performance than the Gaussian model.
+The plots in Figure \ref{fig:bs-ozone} show the relative Brier scores for time-series and non-time-series models, using $K = $ 1, 7, and 15 knots at thresholds $T = $ 0, 50, and 75 ppb.
+Most of the models perform similarly across all the Brier scores; however, for single-partition models without thresholding, performance tends to diminish in the extreme quantiles.
+The results also suggest that thresholding improves performance for estimates in the extreme quantiles.
 Both plots have similar features suggesting that most settings do reasonably well.
 In particular, for all extreme quantiles, selecting a moderate number of knots (e.g. $K = 5, \ldots, 10$) tends to give the best results.
+Table \ref{tbl:ozoneresults} shows the best two models for selected extreme quantiles.
 
-Table \ref{tbl:ozoneresults} shows the best two models for selected extreme quantiles.
+We illustrate the predictive capability of our model in Figure \ref{fig:ozoneq99} by plotting the 99th quantile of the posterior predictive density for July in South Carolina and Georgia.
+\hl{Include more here about the maps}.
 
@@
   \caption{Setting 2 -- Skew-$t$ marginal, $K = 1$ knot}
-  \label{tbl:gaussim}
+  \label{tbl:st1sim}
   \begin{tabular}{|l|ccc|ccc|ccc|ccc|}
@@
   \caption{Setting 3 -- Skew-$t$ marginal, $K = 5$ knots}
-  \label{tbl:gaussim}
+  \label{tbl:st5sim}
   \begin{tabular}{|l|cc|cc|cc|ccc|}
@@
   \caption{Setting 4 -- Max-stable}
-  \label{tbl:gaussim}
+  \label{tbl:mssim}
   \begin{tabular}{|l|cccc|cccc|ccc|ccc|}
@@
   \caption{Setting 5 -- Transformation below $T = q(0.80)$}
-  \label{tbl:gaussim}
+  \label{tbl:transsim}
   \begin{tabular}{|l|ccc|cc|ccc|ccc|}

### 2015-04-06 LaTeX/skewnormal.tex
updated plots for manuscript

@@
 We do not include skewness for the thresholded models because it cannot be identified with only one tail worth of data, and we do not fit the data using the max-stable methods from \citet{Reich2012} because of the computational constraints.
-Each chain of the mcmc ran for 20000 iterations with a burn-in period of 10000 iterations.
+Each chain of the MCMC ran for 20000 iterations with a burn-in period of 10000 iterations.
 Parameters appear to converge properly; however, in the models with multiple partitions (i.e.models 4 and 5) it is hard to assess the convergence of $\bw$, $z(\bs)$, and $\sigma^2(\bs)$ because of partition label switching throughout the MCMC.
@@
 \section{Data analysis}\label{s:analysis}
-To illustrate this method, we consider the daily maximum 8-hour ozone measurements for July 2005 at 735 Air Quality System (AQS) monitoring sites in the eastern United States as the response (see Figure \ref{fig:ozone}).
+To illustrate this method, we consider the daily maximum 8-hour ozone measurements for July 2005 at 1089 Air Quality System (AQS) monitoring sites in the United States as the response (see Figure \ref{fig:ozone}).
 For each site, we also have covariate information containing the estimated ozone from the Community Multi-scale Air Quality (CMAQ) modeling system.
@@
 All methods assume the mean function given in (\ref{eq:datamean}).
-Each chain of the mcmc ran for 30000 iterations with a burn-in period of 25000 iterations.
+To ensure that the max-stable method runs in a reasonable amount of time, we take a stratified sample of the sites to get 800 sites and consider this our new dataset.
+We conduct two-fold cross validation using 400 training sites and 400 validation sites as described in Section \ref{s:modelselect}
+
+Each chain of the MCMC ran for 30000 iterations with a burn-in period of 25000 iterations.
 Parameters appear to converge properly; however, as before, for models with multiple partitions it is hard to assess the convergence of $\bw$, $z(\bs)$, and $\sigma^2(\bs)$ because of partition label switching throughout the MCMC.
@@
 We illustrate the predictive capability of our model in Figure \ref{fig:ozoneq99} by plotting the 99th quantile of the posterior predictive density for July in South Carolina and Georgia.
-\hl{Include more here about the maps}.
+We fit the model using four methods, two reference and two that performed better.
+These four methods are
+\begin{enumerate}[1.]
+  \item Gaussian (reference)
+  \item Skew-$t$, $K =$ 1 knot, $T = $ 0, no time series (reference)
+  \item Skew-$t$, $K =$ 5 knots, $T = $ 50, no time series (comparison)
+  \item Symmetric-$t$, $K =$ 10 knots, $T = $ 75, time series (comparison)
+\end{enumerate}
 

### 2015-04-06 LaTeX/skewnormal.tex
updated manuscript

@@
 \usepackage{color}
+\usepackage{enumitem}
 \usepackage{setspace}
@@
 These four methods are
-\begin{enumerate}[1.]
+\begin{enumerate}\setlength{\itemsep}{-0.5em}
   \item Gaussian (reference)
@@
   \item Skew-$t$, $K =$ 5 knots, $T = $ 50, no time series (comparison)
-  \item Symmetric-$t$, $K =$ 10 knots, $T = $ 75, time series (comparison)
+  \item Symmetric-$t$, $K =$ 10 knots, $T = $ 75, time series (comparison).
 \end{enumerate}
+In the bottom two plots, we plot the differences between method 4 and methods 1 and 2.
+The most noticeable differences between the reference methods and the comparison methods is that the comparison methods tend to give higher estimates of the 99th quantile along the I-85 corridor between Charlotte and Atlanta.
 

### 2015-05-15 LaTeX/skewnormal.tex
change manuscript to point to library file.
Major revisions to manuscript introduction

@@
 In most climatological applications, researchers are interested in learning about the average behavior of different climate variables (e.g. ozone, temperature, rainfall).
-However, averages do not help regulators prepare for the unusual events that only happen once every 100 years.
-For example, it is important to have an idea of how much rain will come in a 100-year floor in order to construct strong enough river levees to protect lands from flooding.
+However, averages are not adequate to prepare for the unusual events that only happen once every 100 years.
+For example, it is important to have an idea of how much rain will come in a 100-year flood in order to construct strong enough river levees to protect lands from flooding.
+To accurately estimate the risk associated with such rare events, it is imperative to use statistical models designed for extreme values.
+Given that many of these applications demonstrate some level of spatial and temporal dependence, there is a growing interest in spatio-temporal methods for extreme values.
 
-Unlike multivariate normal distributions, it is challenging to model multivariate extreme value distributions (e.g.
-generalized extreme value and generalized Pareto distribution) because few closed-form expressions exist for the density in more than two-dimensions \citep{Coles1991}.
-Given this limitation, pairwise composite likelihoods have been used when modeling dependent extremes \citep{Padoan2010,Blanchet2011,Huser2013}.
+Traditionally, spatial methods for extreme values analysis are conducted from one of two perspectives.
+The first of these is based on the convergence of the maximums of independent stochastic processes to a max-stable process \citep{deHaan2006}.
+Finite dimensional realizations of a max-stable process follow a generalized extreme value distribution \citep{Cooley2012}.
+One drawback to a block-maxima approach is that information is lost by discarding all but the most extreme observations in a block.
+Furthermore, in using multivariate block-maxima methods, the observations in the vector of block maxima rarely occur simultaneously \citep{Coles2001}.
+The other perspective incorporates a peaks-over-threshold approach.
+This methodology requires selecting a threshold, beyond which, the data are assumed to be considered extreme.
+Recently, threshold methods have become a popular choice through censored pairwise composite likelihoods \citep{Wadsworth2012,Thibaud2013,Thibaud2013a,Huser2014}.
 
-One way around the multi-dimensional limitation of multivariate extreme value distributions is to use skew elliptical distributions to model dependent extreme values \citep{Genton2004,Zhang2010,Padoan2011}.
-Due to their flexibility, the skew-normal and skew-$t$ distribution offer a flexible way to handle non-symmetric data within a framework of multivariate normal and multivariate t-distributions.
-As with the spatial Gaussian process, the skew-normal distribution is also asymptotically independent; however, the skew-$t$ does demonstrate asymptotic dependence \citep{Padoan2011}.
-Although asymptotic dependence is desirable between sites that are near one another, one drawback to the skew-$t$ is that sites remain asymptotically dependent even at far distances.
+Unlike multivariate normal distributions, multivariate extreme value distributions are challenging to use directly because closed-form expressions for the density in more than two dimensions are complicated \citep{Coles1991}.
+One exception to this is the Brown-Resnick process, a process for which it is possible to conduct inference in $d$-dimensions \citep{Wadsworth2013,Engelke2014}.
+Extreme value distributions based on the skew-$t$ distribution also permit modeling of extreme values in higher dimensions \citep{Padoan2011}.
+This limitation has led to the development of pairwise composite likelihood methods to model spatially dependent extremes \citep{Padoan2010,Blanchet2011,Huser2013}.
+Bayesian methods have also been proposed using hierarchical models with positive stable random effects \citep{Reich2012}.
 
-In this paper, we present a model that has marginal distributions with flexible tails, demonstrates asymptotic dependence for observations at sites that are near to one another, and has computation on the order of Gaussian models for large space-time datasets.
+Despite these challenges, the spatial modeling of extremes is important because it provides a mechanism by which we can borrow information about extreme events across space.
+There are two primary goals when modeling spatial extremes.
+The first of these goals is to understand the marginal behavior at sites, and the second is to describe the asymptotic dependence between sites.
+In many cases, these are done separately; however, with the development of pairwise composite likelihoods and Bayesian methods, it is resonable to model both simultaneously.
+One challenge to estimating the asymptotic dependence is that the observed extremal dependence may not be the same as the limiting dependence \citep{Davison2012b}.
+Furthermore, in a spatial setting, it may be desirable to have varying degrees of asymptotic dependence based upon the distance between two sites \citep{Wadsworth2012}.
+
+One way around the multi-dimensional limitations of multivariate extreme value distributions is to use skew elliptical distributions to model dependent extreme values \citep{Genton2004,Zhang2010,Padoan2011}.
+For example, the skew-normal and skew-$t$ distribution offer a flexible way to handle non-symmetric data within a framework of multivariate normal and multivariate $t$ distributions.
+As with multivariate Gaussian distributions, the multivariate skew-normal distribution demonstrates asymptotic independence.
+Conversely both the multivariate $t$ and skew-$t$ distributions demonstrate asymptotic dependence \citep{Padoan2011}.
+Additionally, the limiting distribution of the maxima of skew-$t$ random vectors is the extremal skew-$t$ distribution \citep{Padoan2011} of which the extremal-$t$ \citep{Opitz2013} is a special case.
+
+In many cases, the goal of analyzing extreme values is to understand the behavior of very rare events; however, we take a slightly different approach that is motivated by assessing ozone compliance.
+A site is said to be in compliance if the fourth highest daily maximum 8-hour concentration averaged over 3 years does not exceed 75 parts per billion.
+Because this value typically occurs during the summer months, the fourth highest daily maximum is roughly equivalent to modeling the 95th quantile for ozone for the summer or 99th quantile for the year.
+These high quantiles are extreme in the sense that they are unusual for a given year, but they are not extreme in the sense that we are always guaranteed to observe this quantile level every year at each site by definition.
+
+In this paper, we present a new spatio-temporal model based on the skew-$t$ process that is informed by two concepts from extreme value analysis.
 Specifically, our contribution is to incorporate thresholding and random spatial partitions using a multivariate skew-$t$ distribution.
 The advantage of using a thresholded model as opposed to a non-thresholded model is that is allows for the tails of the distribution to inform the predictions in the tails \citep{DuMouchel1983}.
-The random spatial partitions are similar to the method used by \citet{Kim2005} for non-stationary Gaussian data.
-The partition alleviates the long-range spatial dependence present in the skew-$t$ distribution.
+The random spatial partitions are similar to the method used by \citet{Kim2005} for non-stationary Gaussian data, and they alleviate the asymptotic spatial dependence present in the skew-$t$ distribution for sites that are far apart.
+Finally, our model allows for inference and predictions using the full likelihood with computing on the order of Gaussian models for large space-time datasets.
 
@@
 \bibliographystyle{rss}
-\bibliography{skewnormal}
+\bibliography{library}
 

### 2015-05-31 LaTeX/skewnormal.tex
single space bibliography
updates to skewnormal from CO

--- 2015-04-06 LaTeX/skewnormal.tex
@@
 \begin{center}
-{\Large {\bf A spatial-skew model for extreme values}}\\
+{\Large {\bf A spatial-skew model for threshold exceedances}}\\
 \today
@@
 \section{Introduction}\label{s:intro}
-In most climatological applications, researchers are interested in learning about the average behavior of different climate variables (e.g. ozone, temperature, rainfall).
-However, averages do not help regulators prepare for the unusual events that only happen once every 100 years.
-For example, it is important to have an idea of how much rain will come in a 100-year floor in order to construct strong enough river levees to protect lands from flooding.
+In many climatological applications, researchers are interested in learning about the average behavior of different climate variables (e.g. ozone, temperature, rainfall).
+Our study is motivated by an air pollution application where the focus is not on the average behavior, but instead the behavior over a fixed threshold determined by government regulation.
+More specifically, we consider consider the case of compliance for ozone.
+A site is said to be in compliance if the fourth highest daily maximum 8-hour concentration averaged over 3 years does not exceed 75 parts per billion (ppb).
 
-Unlike multivariate normal distributions, it is challenging to model multivariate extreme value distributions (e.g.
-generalized extreme value and generalized Pareto distribution) because few closed-form expressions exist for the density in more than two-dimensions \citep{Coles1991}.
-Given this limitation, pairwise composite likelihoods have been used when modeling dependent extremes \citep{Padoan2010,Blanchet2011,Huser2013}.
+Traditional geostatistical modeling is based upon the assumption that observations come from a Gaussian process, a process that is fully defined by its mean and covariance functions.
+In the limit of the Gaussian distribution, all observations are independent regardless of the strength of the correlation in the bulk of the data.
+Furthermore, the Gaussian distribution is light-tailed and symmetric.
+Therefore, it is inappropriate to use standard geostatistical methods when trying to describe dependence in the tails of the distribution.
 
-One way around the multi-dimensional limitation of multivariate extreme value distributions is to use skew elliptical distributions to model dependent extreme values \citep{Genton2004,Zhang2010,Padoan2011}.
-Due to their flexibility, the skew-normal and skew-$t$ distribution offer a flexible way to handle non-symmetric data within a framework of multivariate normal and multivariate t-distributions.
-As with the spatial Gaussian process, the skew-normal distribution is also asymptotically independent; however, the skew-$t$ does demonstrate asymptotic dependence \citep{Padoan2011}.
-Although asymptotic dependence is desirable between sites that are near one another, one drawback to the skew-$t$ is that sites remain asymptotically dependent even at far distances.
+Threshold modeling is popular in the field of extreme value statistics where extreme events are naturally defined in terms of exceedances over a high threshold.
+\citet{Davison1990} considered modeling threshold exceedances of univariate time series by the generalized Pareto distribution.
+Bivariate threshold models for extreme value distributions were considered by \citet{Ledford1996} who introduced a censored approach that provides a way to deal with different types of exceedances of a bivariate threshold in terms of only one or both components.
+These threshold models were extended to spatial models for extremes by \citet{Wadsworth2012} and \citet{Thibaud2013} who fit various models to spatial extremes using a censored pairwise likelihood \citep{Padoan2010} based on the approach of \citet{Ledford1996}.
+\citet{Huser2014} further extended this to spate-time modeling.
+\citet{Engelke2014}, \citet{Wadsworth2014}, and \citet{Thibaud2013a} introduced more efficient inference for threshold exceedances of extremal spatial processes with full likelihood methods.
+The previous approaches to threshold modeling are motivated by extreme value theory and assume the threshold is high enough to assume extremal models are valid for the data, and for extrapolation beyond the range of observed values.
+Moreover, these approaches are computationally intensive and limited to rather small datasets.
+Our application with ozone data does not fit into this framework because we do not focus on exceedances of a very high threshold, but on exceedances of a fixed threshold.
 
-In this paper, we present a model that has marginal distributions with flexible tails, demonstrates asymptotic dependence for observations at sites that are near to one another, and has computation on the order of Gaussian models for large space-time datasets.
-Specifically, our contribution is to incorporate thresholding and random spatial partitions using a multivariate skew-$t$ distribution.
-The advantage of using a thresholded model as opposed to a non-thresholded model is that is allows for the tails of the distribution to inform the predictions in the tails \citep{DuMouchel1983}.
-The random spatial partitions are similar to the method used by \citet{Kim2005} for non-stationary Gaussian data.
-The partition alleviates the long-range spatial dependence present in the skew-$t$ distribution.
+Instead, we propose a new spatiotemporal threshold exceedance model based on the skew-$t$ process \citep{Padoan2011}.
+Our model is a threshold exceedance model for the multivariate skew-$t$ distribution that uses imputation for values below a fixed threshold.
+We use a skew-$t$ distribution because of its flexibility to model asymmetry and heavy-tailed data with the aim of predicting the probability of exceeding a high fixed threshold at an unobserved location.
+
+In a spatial setting, the multivariate skew-$t$ distribution demonstrates asymptotic dependence between observations at all sites regardless of the distance between the sites.
+In order to address this concern, we introduce a random spatial partition similar to the method used by \citet{Kim2005} for non-stationary Gaussian data.
+This partition alleviates the asymptotic spatial dependence present in the skew-$t$ distribution for sites that are far apart.
+Finally, our model allows for inference and predictions using the full likelihood with computing on the order of Gaussian models for large space-time datasets.
 
@@
 
+% These models describe dependence using the probability that two observations jointly exceed an extreme value, which in the limit of the distribution is referred to as asymptotic dependence.
+% When the spatial domain is small, it may be reasonable to assume that all sites are asymptotically dependent; however for large spatial domains, it becomes more challenging to justify this assumption.
+% This concern has been addressed by \citet{Wadsworth2012} in a spatial only setting and \citet{Huser2014} in a space-time setting.
+
+% The main goal of our application is to estimate the marginal probability a site's ground-level ozone measurement will exceed 75 ppb.
+% For July 2005, 75 ppb is approximately the 92nd sample quantile.
+% However, if we look marginally at each site, 75 ppb is lower than the 90th quantile for almost 255 sites, and is below the median at 29 sites.
+% The traditional threshold exceedance models arising from max-stable processes incorporate thresholding because the inference is only valid for very high values.
+% Because 75 ppb represents such a wide range of marginal quantiles, these models based on max-stable methods may not be appropriate.
+
+% Traditionally, spatial methods for extreme values analysis are conducted from one of two perspectives.
+% The first of these is based on the convergence of the maximums of independent stochastic processes to a max-stable process \citep{deHaan2006}.
+% Finite dimensional realizations of a max-stable process follow a generalized extreme value distribution \citep{Cooley2012}.
+% One drawback to a block-maxima approach is that information is lost by discarding all but the most extreme observations in a block.
+% Furthermore, in using multivariate block-maxima methods, the observations in the vector of block maxima rarely occur simultaneously \citep{Coles2001}.
+% The other perspective incorporates a peaks-over-threshold approach.
+
+% The use of multivarate models for threshold exceedances require evaluation of the underlying joint density.
+% Finite dimensional realizations of max-stable processes are challenging to use because closed-form expressions for the density in more than two dimensions are complicated \citep{Coles1991}.
+% One way to circumvent this challenge is to use pairwise composite likelihood methods \citep{Padoan2010}.
+% When using pairwise composite likelihoods, threshold methods can then be implemented as long as the bivariate distribution can be evaluated \citep{Wadsworth2012,Thibaud2013,Huser2014}.
+% Hierarchical models using Bayesian methods have also been proposed.
+% For example, \citet{Reich2012} present a Bayesian hierarchical model, when conditioned on positive stable spatial random effects, the observations are independent.
+
+% In certain settings, the multidimensional distributions are available.
+% For example, \citet{Wadsworth2014} discuss a censored Poisson process that exploits the fact that the spectral functions for Brown-Resnick processes are log-Gaussian random fields.
+% Additionally, \citep{Engelke2014}, show that upon standardizing marginals to the Gumbel distribution and conditioned on a fixed location exceeding a high threshold, the incremental distribution asymptotically forms a Gaussian process.
+% In addition to Brown-Resnick processes, skew elliptical distributions can be used for multivariate modeling with dependent extreme values \citep{Genton2004,Zhang2010,Padoan2011}.
+% For example, the skew-normal and skew-$t$ distribution offer a flexible way to handle non-symmetric data within a framework of multivariate normal and multivariate $t$ distributions.
+% As with multivariate Gaussian distributions, the multivariate skew-normal distribution demonstrates asymptotic independence.
+Conversely both the multivariate $t$ and skew-$t$ distributions demonstrate asymptotic dependence \citep{Padoan2011}.
+% Additionally, the limiting distribution of the maxima of skew-$t$ random vectors is the extremal skew-$t$ distribution \citep{Padoan2011} of which the extremal-$t$ \citep{Opitz2013} is a special case.
+
+% Despite this challenge, spatial modeling is important because it provides a mechanism by which we can borrow information about extreme events across space.
+% There are two primary goals for spatial methods.
+% The first of these goals is to understand the marginal behavior at sites, and the second is to describe the dependence in the data.
+% In many cases, these are done separately; however, with the development of pairwise composite likelihoods and Bayesian methods, it is resonable to model both simultaneously.
+% One challenge to estimating the dependence in the tails of the distribution is that the dependence that present in the data may not be the same as the dependence in the limit of the underlying distribution \citep{Davison2012b}.
+% Furthermore, in a spatial setting, it may be desirable to allow for varying degrees of tail dependence based upon the distance between two sites \citep{Wadsworth2012}.
+
 \section{Spatial skew processes}\label{s:spatialskew}
@@
 \subsection{Skew-$t$ process} \label{s:skewt}
-Let $Y(\bs)$ be the observation at spatial location $\bs = (s_1, s_2)$.
+Let $Y(\bs)$ be the observation at spatial location $\bs = (s_1, s_2)$ in a spatial domain of interest $\calD \in \calR^2$.
 The spatial skew-$t$ process can be written
@@
 \end{align}
-where $\bX(\bs)$ is a set of spatial covariates at site $\bs$, $\bbeta$ is the vector of regression parameters, $\lambda \in \calR$ is a parameter controlling skew, $z \sim N(0, 1)$, $\sigma^2 \sim \text{IG}(a, b)$ is an inverse gamma random variable, and $v(\bs)$ is a spatial Gaussian process with mean zero and variance one.
+where $\bX(\bs)$ is a set of spatial covariates at site $\bs$, $\bbeta$ is the vector of regression parameters, $\lambda \in \calR$ is a parameter controlling skew, $z \sim N(0, 1)$, $\sigma^2 \sim \text{IG}(a, b)$ is an inverse gamma random variable, and $v(\bs)$ is a spatial Gaussian process with mean zero, variance one, and a positive definite correlation function.
 
-Let $\bY = [Y(\bs_1), \ldots, Y(\bs_n)]^T$ be a set of observations at a finite collection of locations $\bs_1, \ldots, \bs_n$.
+For a finite collection of locations $\bs_1, \ldots, \bs_n$, consider observations $\bY = [Y(\bs_1), \ldots, Y(\bs_n)]^T$.
 After marginalizing over both $z$ and $\sigma$,
@@
 \end{align}
-that is, $\bY$ follows an $n$-dimensional skew-$t$ distribution with location $\bX \bbeta$, correlation matrix $\bOmega$, slant parameters $\balpha$ and degrees of freedom $2a$, where $\bX = [\bX(\bs_1)^T, \ldots, \bX(\bs_n)^T]$, $\bOmega = \bomega \bar{\bOmega} \bomega$, $\bomega = \text{diag}\left(\frac{1}{ \sqrt{ab}}, \ldots, \frac{1}{\sqrt{ab}}\right)$, $\bar{\bOmega} = (\bSigma + \lambda^2 \bOne \bOne^T$), $\bSigma$ is a positive definite correlation matrix, $\balpha = \lambda (1 + \lambda^2 \bOne^T \bSigma^{-1} \bOne)^{-1/2} \bOne^T \bSigma^{-1}$ is a vector of slant parameters.
-Although $\bSigma$ can be any positive definite correlation matrix, we choose to use the stationary isotropic \Matern correlation with
+that is, $\bY$ follows an $n$-dimensional skew-$t$ distribution with location $\bX \bbeta$, correlation matrix $\bOmega$, slant parameters $\balpha$ and degrees of freedom $2a$, where $\bX = [\bX(\bs_1)^T, \ldots, \bX(\bs_n)^T]$, $\bOmega = \bomega \bar{\bOmega} \bomega$, $\bomega = \text{diag}\left(\frac{1}{ \sqrt{ab}}, \ldots, \frac{1}{\sqrt{ab}}\right)$, $\bar{\bOmega} = (\bSigma + \lambda^2 \bOne \bOne^T$), $\bSigma$ is the positive definite correlation matrix of $[v(\bs_1), \ldots, v(\bs_n)]$, $\balpha = \lambda (1 + \lambda^2 \bOne^T \bSigma^{-1} \bOne)^{-1/2} \bOne^T \bSigma^{-1}$ is a vector of slant parameters.
+Although $\bSigma$ can be any positive definite correlation function, we choose to use the stationary isotropic \Matern correlation with
 \begin{align}
@@
 This process is desirable because of its flexible tail that is controlled by the skewness parameter $\lambda$ and degrees of freedom $2a$.
-Furthermore, the marginal distributions at each location also follow a univariate skew-$t$ distribution \citep{Azzalini2013}.
+Furthermore, the marginal distributions at each location also follow a univariate skew-$t$ distribution \citep{Azzalini2014}.
 
@@
 Our interest lies in spatial dependence in the tail of the skew-$t$ process.
-One measure of extremal dependence is the $\chi$ statistic \citep{Padoan2011}.
-For a stationary and isotropic spatial process, the $\chi$ statistic for two location $\bs$ and $\bt$ separated by distance $h = || \bs - \bt ||$ is
+One measure of extremal dependence is the $\chi$ statistic \citep{Coles1999}.
+For a stationary and isotropic spatial process, the $\chi$ statistic for two location $\bs$ and $\bt$ separated by distance $h = || \bs - \bt ||$ with identical marginal distributions is
 \begin{align}
-  \chi(h) = \lim_{c \rightarrow \infty} \Pr[Y(\bs) > c | Y(\bt) > c].
+  \chi(h) = \lim_{c \rightarrow c^*} \Pr[Y(\bs) > c | Y(\bt) > c]
 \end{align}
+where $c^*$ is the upper limit of the support of $Y$.
 If $\chi(h) = 0$, then observations are asymptotically independent at distance $h$.
-For Gaussian processes, $\chi(h) = 0$ regardless of the distance, so they are not suitable for modeling spatially-dependent extremes.
-Unlike the Gaussian process, the skew-$t$ process is asymptotically dependent.
-However, one problem with the spatial skew-$t$ process is that $\lim_{h \rightarrow \infty} \chi(h) > 0$ (see Appendix \ref{a:skewt} for a proof).
+For Gaussian processes, $\chi(h) = 0$ regardless of the distance, so they are not suitable for modeling asymptotically dependent extremes.
+Unlike the Gaussian process, the skew-$t$ process is asymptotically dependent (see Appendix \ref{a:skewt}).
+However, one problem with the spatial skew-$t$ process is that $\lim_{h \rightarrow \infty} \chi(h) > 0$.
 This occurs because all observations, both near and far, share the same $z$ and $\sigma$ terms.
@@
 \subsection{Censoring to focus on the tails}
-To avoid bias in estimating tail parameters, we model censored data.
+Because one of our goals is to model the dependence of the distribution in the tails of the data, we choose to censor values below threshold.
 Let
@@
 \subsection{Partitioning to remove long-range asymptotic dependence}\label{s:part}
+The motivation for the partition is that for a large spatial domain, it may not be reasonable to assume sites that are far apart demonstrate asymptotic dependence.
+Modeling different levels of asymptotic dependence was discussed by \citet{Wadsworth2012} with a hybrid spatial dependence model.
+\citet{Huser2014} also allow for asymptotic dependence across both space and time with a partition structure represented by random discs moving across the space for a random duration with a random velocity and random radius.
 We handle the problem of long-range asymptotic dependence with a random partition.
@@
 \subsection{Extension to space-time data} \label{s:temporal}
-When using daily measurements, the assumption of temporal independence is inappropriate.
+When using daily measurements, the assumption of temporal independence is often inappropriate.
 There are several places where temporal dependence could be incorporated in the model, including the residual $v_t(\bs)$.
@@
 \end{align}
+We allow the partition structure to vary from day to day in order to account for sharp spikes in ozone that may not be present every day (e.g. a forest fire).
 
@@
 \section{Computation}\label{s:comp}
-First, we impute values below the threshold.
+First, we impute values below the threshold conditional on observations above the threshold.
+This is feasible for large datasets with our model because for a single day, conditional on the model parameters, we only need to draw from a truncated multivariate normal distribution.
 Then, we update model parameters, $\Theta$, using Metropolis Hastings or Gibbs sampling when appropriate.
@@
 Let $\bw_{t1}, \ldots, \bw_{tK}$ be a set of daily spatial knots in a spatial domain of interest, $\calD$, and $P_{tk}$ as defined in (\ref{eq:subregions}).
+In practice, we fix $K$ at many different levels, and assess the impact of fit as described in \ref{s:modelselect}.
 Then
@@
   \item Symmetric-$t$ marginal, $K=5$ knots, $T=q(0.80)$
+  \item A max-stable model based on \citet{Reich2012} thresholded at $T = q(0.80)$
 \end{enumerate}
@@
 The knots have priors $\bw \sim \text{Unif}(\calD)$.
-We do not include skewness for the thresholded models because it cannot be identified with only one tail worth of data, and we do not fit the data using the max-stable methods from \citet{Reich2012} because of the computational constraints.
+We tried also fitting the skew-$t$ marginals for the thresholded models, but it is very challenging for the MCMC to properly identify the skewness parameter with only one tail worth of data.
 Each chain of the MCMC ran for 20000 iterations with a burn-in period of 10000 iterations.
@@
 The model was fit using the training set, and predictions were generated at the testing site locations.
-Because one of the primary goals of this model is to predict extreme events, we use Brier scores to select the model that best fits the data \citep{Gneiting2007}.
+Because one of the primary goals of this model is to predict exceedances over a fixed threshold, we use Brier scores to select the model that best fits the data \citep{Gneiting2007}.
 The Brier score for predicting exceedance of a threshold $c$ is given by $[e(c) - P(c)]^2$ where $e(c) = I[y>c]$ is an indicator function indicating that a test set value, $y$, has exceeded the threshold, $c$, and $P(c)$ is the predicted probability of exceeding $c$.
@@
 \section{Data analysis}\label{s:analysis}
-To illustrate this method, we consider the daily maximum 8-hour ozone measurements for July 2005 at 1089 Air Quality System (AQS) monitoring sites in the United States as the response (see Figure \ref{fig:ozone}).
+To illustrate this method, we consider the daily maximum 8-hour ozone measurements for July 1 - 31, 2005 at 1089 Air Quality System (AQS) monitoring sites in the United States as the response (see Figure \ref{fig:ozone}).
 For each site, we also have covariate information containing the estimated ozone from the Community Multi-scale Air Quality (CMAQ) modeling system.
@@
 \end{figure}
-Exploratory data analysis indicates that there is dependence in the high quantile levels of the residuals beyond what we expect in the case if independence.
+Exploratory data analysis indicates that there is dependence in the high quantile levels of the residuals beyond what we expect in the case of independence.
 
@@
 We also compare models with no time series to models that include the time series.
-Finally, as a comparison to max-stable methods, we fit the model using the hierarchical max-stable model of \citet{Reich2012}.
+Finally, as a comparison to max-stable methods, we fit the model using the hierarchical max-stable model of \citet{Reich2012} with the data thresholded at $T = 75$.
 All methods assume the mean function given in (\ref{eq:datamean}).
@@
 Parameters appear to converge properly; however, as before, for models with multiple partitions it is hard to assess the convergence of $\bw$, $z(\bs)$, and $\sigma^2(\bs)$ because of partition label switching throughout the MCMC.
-For each model, Brier scores were were averaged over all sites and days to obtain a single Brier score for each dataset.
+For each model, Brier scores were averaged over all sites and days to obtain a single Brier score for each dataset.
 At a particular threshold or quantile level, the model that fits the best is the one with the lowest score.
@@
 \section{Discussion}\label{s:con}
-In this paper we propose a new approach for spatitemporal modeling of extreme values.
+In this paper we propose a new threshold exceedance approach for spatiotemporal modeling based on the skew-$t$ process.
 The proposed model gives flexible tail behavior, demonstrates asymptotic dependence for observations at sites that are near to one another, and has computation on the order of Gaussian models for large space-time datasets.
@@
 \end{align}
-where $t_d(\bz; \bar{\bOmega}, \eta)$ is a $d$-dimensional Student's $t$-distribution with scale matrix $\bar{\bOmega}$ and degrees of freedom $\eta$, $Q(z) = \bz^T \bar{\Omega}^{-1}\bz$ and $T(\cdot; \eta)$ denotes the univariate Student's $t$ distribution function with $\eta$ degrees of freedom \citep{Azzalini2013}.
+where $t_d(\bz; \bar{\bOmega}, \eta)$ is a $d$-dimensional Student's $t$-distribution with scale matrix $\bar{\bOmega}$ and degrees of freedom $\eta$, $Q(z) = \bz^T \bar{\Omega}^{-1}\bz$ and $T(\cdot; \eta)$ denotes the univariate Student's $t$ distribution function with $\eta$ degrees of freedom \citep{Azzalini2014}.
 
@@
 
+\begin{singlespace}
 \bibliographystyle{rss}
-\bibliography{skewnormal}
+\bibliography{library}
+\end{singlespace}
 

### 2015-05-31 markdown/exploring_model_fit.Rmd
updates to skewnormal from CO

---
title: "Exploring Model Fits"
author: "Sam Morris"
date: "19 May 2015"
output: pdf_document
---

```{r, echo=FALSE, include=FALSE}
load("./code/analysis/simstudy/simdata.RData")
library(evd)  # to load in chiplot
```

The goal of this document is to try to explore different ways to think about dependence in the data.
In talking with Emeric and Dan, it's not clear to me that we have the best way to assess model fit.
Brier scores are nice in that they give a single number, and similar to other decision making criterion, you can decide on the best model simply by looking at which model has the lowest Brier score.
Furthermore, the Brier score is the proper scoring rule that can be used to assess how well you do at predicting the occurence of an event.

One thing I've been thinking about is trying to figure out how well we do at recreating the time series for a specific site. 
One of the main goals of our model is to predict at unobserved locations.
Therefore, it would be desirable to know how well we do at estimation of the time series at an unobserved location.
One plot that might be interesting is the posterior distribution at a particular site for all of the time points.
If we overlay the observed values, you could see how we do marginally for 3 or 4 sites.
You could also find some way to differentiate the values that exceed the 95th quantile (for the ozone data, the top 2 days; simulated data, the top 3 days).

### 2015-05-31 code/analysis/simstudy/changenumbering.R

@@ -1,0 +1,72 @@
+# a is for method 5
+# b is for method 4
+# c is for methods 1 - 3
+#   c: fit.1 is method 1
+#   c: fit.2 is method 2
+#   c: fit.3 is method 3
+setwd("/Volumes/sam-ext/spatial-skew-t/code/analysis/simstudy/results")
+
+for (setting in 7) {
+  for (g in 2:10) {
+    oldfile <- paste(setting, "-a-", g, ".RData", sep="")
+    load(oldfile)
+    fit.2 <- fit.1
+    
+    for (i in 1:5) {
+      # filename should be setting-method-dataset.RData
+      filename <- paste(setting, "-5-", (g - 1) * 5 + i, ".RData", sep="")
+      cat(filename, ": start ")
+      fit.1 <- fit.2[[i]]
+      save(fit.1, file=filename)
+      cat("end \n")
+    }
+  }
+}
+
+for (setting in 1:7) {
+  for (g in 1:10) {
+    oldfile <- paste(setting, "-b-", g, ".RData", sep="")
+    load(oldfile)
+    fit.2 <- fit.1
+    
+    for (i in 1:5) {
+      # filename should be setting-method-dataset.RData
+      filename <- paste(setting, "-4-", (g - 1) * 5 + i, ".RData", sep="")
+      cat(filename, ": start ")
+      fit.1 <- fit.2[[i]]
+      save(fit.1, file=filename)
+      cat("end \n")
+    }
+  }
+}
+
+for (setting in 1:7) {
+  for (g in 1:10) {
+    oldfile <- paste(setting, "-c-", g, ".RData", sep="")
+    load(oldfile)
+    fit.4 <- fit.1
+    
+    for (i in 1:5) {
+      # filename should be setting-method-dataset.RData
+      filename <- paste(setting, "-3-", (g - 1) * 5 + i, ".RData", sep="")
+      cat(filename, ": start ")
+      fit.1 <- fit.3[[i]]
+      save(fit.1, file=filename)
+      cat("end \n")
+      
+      # filename should be setting-method-dataset.RData
+      filename <- paste(setting, "-2-", (g - 1) * 5 + i, ".RData", sep="")
+      cat(filename, ": start ")
+      fit.1 <- fit.2[[i]]
+      save(fit.1, file=filename)
+      cat("end \n")
+      
+      # filename should be setting-method-dataset.RData
[…]

### 2015-06-10 LaTeX/skewnormal.tex
addition of new plot for brier score performance by marginal q-1(75)

@@
 
+\hl{NEED TO ADD STUFF HERE ACKNOWLEDGING THAT MARGINALS ARE DIFFERENT}
+Also, add an explanation about the purpose of the plot in \ref{fig:bs-marginal}
+\begin{figure}
+  \centering
+  \includegraphics[width=0.5\linewidth]{plots/bs-ozone-marginal.pdf}
+  \caption{Comparison of Brier score performance by marginal quantile representing 75ppb}
+  \label{fig:bs-marginal}
+\end{figure}
+
 \begin{figure}

### 2015-07-01 LaTeX/skewnormal.tex
updates to draft

@@
 \begin{center}
-{\Large {\bf A spatial-skew model for threshold exceedances}}\\
+{\Large {\bf A spatial skew-$t$ model for threshold exceedances}}\\
 \today
@@
 
-For a finite collection of locations $\bs_1, \ldots, \bs_n$, consider observations $\bY = [Y(\bs_1), \ldots, Y(\bs_n)]^T$.
+For a finite collection of locations $\bs_1, \ldots, \bs_n$, denote the vector of observations $\bY = [Y(\bs_1), \ldots, Y(\bs_n)]^T$.
 After marginalizing over both $z$ and $\sigma$,
@@
 that is, $\bY$ follows an $n$-dimensional skew-$t$ distribution with location $\bX \bbeta$, correlation matrix $\bOmega$, slant parameters $\balpha$ and degrees of freedom $2a$, where $\bX = [\bX(\bs_1)^T, \ldots, \bX(\bs_n)^T]$, $\bOmega = \bomega \bar{\bOmega} \bomega$, $\bomega = \text{diag}\left(\frac{1}{ \sqrt{ab}}, \ldots, \frac{1}{\sqrt{ab}}\right)$, $\bar{\bOmega} = (\bSigma + \lambda^2 \bOne \bOne^T$), $\bSigma$ is the positive definite correlation matrix of $[v(\bs_1), \ldots, v(\bs_n)]$, $\balpha = \lambda (1 + \lambda^2 \bOne^T \bSigma^{-1} \bOne)^{-1/2} \bOne^T \bSigma^{-1}$ is a vector of slant parameters.
-Although $\bSigma$ can be any positive definite correlation function, we choose to use the stationary isotropic \Matern correlation with
+Although any positive definite correlation function could be used, we choose to use the stationary isotropic \Matern correlation with
 \begin{align}
@@
 One measure of extremal dependence is the $\chi$ statistic \citep{Coles1999}.
-For a stationary and isotropic spatial process, the $\chi$ statistic for two location $\bs$ and $\bt$ separated by distance $h = || \bs - \bt ||$ with identical marginal distributions is
+For a stationary and isotropic spatial process, the $\chi$ statistic for locations $\bs$ and $\bt$ separated by distance $h = || \bs - \bt ||$ with identical marginal distributions is
 \begin{align}
@@
 If $\chi(h) = 0$, then observations are asymptotically independent at distance $h$.
-For Gaussian processes, $\chi(h) = 0$ regardless of the distance, so they are not suitable for modeling asymptotically dependent extremes.
-Unlike the Gaussian process, the skew-$t$ process is asymptotically dependent (see Appendix \ref{a:skewt}).
+For Gaussian processes, $\chi(h) = 0$ regardless of the distance $h$, so they are not suitable for modeling asymptotically dependent extremes.
+Unlike the Gaussian process, the skew-$t$ process is asymptotically dependent (the explicit expression for $\chi(h)$ is given in Appendix \ref{a:skewt}).
 However, one problem with the spatial skew-$t$ process is that $\lim_{h \rightarrow \infty} \chi(h) > 0$.
@@
 Therefore, this long-range dependence feature of the skew-$t$ process is not ideal for spatial analysis of large geographic regions where we expect only local spatial dependence.
-The explicit expression for $\chi(h)$ is given in Appendix \ref{a:skewt}.
 
@@
 \end{align}
-where $I(\cdot)$ is an indicator function, $\chi_{\text{skew-}t}(h)$ is the $\chi$ statistic for a skew-$t$ process, $\chi_{\text{Gaus}}(h)$ is the $\chi$ statistic for a Gaussian process, and $h = ||\bs - \bt||$.
+where $I(\cdot)$ is an indicator function, $\chi_{\text{skew-}t}(h)$ is the $\chi$ statistic for a skew-$t$ process given in (\ref{eq:chiskew-t}), $\chi_{\text{Gaus}}(h)$ is the $\chi$ statistic for a Gaussian process, and $h = ||\bs - \bt||$.
 Therefore, sites in different subregions are asymptotically independent because $\chi_{\text{Gaus}}(h) = 0$ for all $h$.
@@
 When using daily measurements, the assumption of temporal independence is often inappropriate.
-There are several places where temporal dependence could be incorporated in the model, including the residual $v_t(\bs)$.
+In this section, we extend (\ref{eq:partition}) to the spatiotemporal setting.
+There are several places where temporal dependence could be incorporated in the model, including the residuals $v_t(\bs)$.
 However, we choose to allow for temporal dependence in the $\bw$, $z$, and $\sigma$ terms because these terms dictate the tail behavior which is our primary focus.
-In this section, we extend (\ref{eq:partition}) to the spatiotemporal setting.
 Let
@@
 Let \hbox{$\bw_{tk} = (w_{tk1}, w_{tk2})$} be a spatial knot on day $t$, and let $w_{t1}, \ldots, w_{tK}$ be a collection of spatial knots on day $t$.
-As in section \ref{s:part}, these knots define a daily partition $P_{t1}, \ldots, P_{tK}$, and for $\bs \in P_{tk}$,
+As in Section \ref{s:part}, these knots define a daily partition $P_{t1}, \ldots, P_{tK}$, and for $\bs \in P_{tk}$,
 \begin{align}
@@
 
-\section{Computation}\label{s:comp}
-First, we impute values below the threshold conditional on observations above the threshold.
-This is feasible for large datasets with our model because for a single day, conditional on the model parameters, we only need to draw from a truncated multivariate normal distribution.
-Then, we update model parameters, $\Theta$, using Metropolis Hastings or Gibbs sampling when appropriate.
-Finally, we make spatial predictions using conditional multivariate normal results and the fact that the distribution of $Y_t(\bs) \mid \Theta, z(\bs)$ is the usual multivariate normal distribution with a \Matern spatial covariance structure.
-
-We can use Gibbs sampling to update $Y_t(\bs)$ for censored observations that are below the threshold $T$.
-After conditioning on $\lambda$, $z_t(\bs)$ and non-censored observations, $Y_t(\bs)$ has truncated normal full conditionals.
-So we sample $Y_t(\bs) \sim N_{(-\infty, T)}(\bX_t^T(\bs) \beta + \lambda | z_t(\bs)|, \bSigma)$.
-After imputing the censored observations, we update the model parameters.
-To update the model parameters, we use standard Gibbs updates for parameters when possible.
-In the case Gibbs sampling is not possible, parameters are updated using a random-walk Metropolis Hastings algorithm.
-See Appendices A.1 and A.2 for details regarding the MCMC.
-The final step of the computation is to use Bayesian Kriging to generate a predictive distribution for $Y_t(\bs^*)$ at prediction location $\bs^*$.
-This step is similar to the imputation for censored observations except that the full conditionals are no longer truncated at $T$.
-
-\subsection{Hierarchical model}\label{s:hier}
+\section{Hierarchical model}\label{s:hier}
 Conditioned on $z_{tk}(\bs)$, $\sigma^2_{tk}(\bs)$, and $P_{tk}$, the marginal distributions are Gaussian and the joint distribution multivariate Gaussian.
@@
 
+\subsection{Computation}\label{s:comp}
+We use Markov chain Monte Carlo methods to explore the posterior.
+At each MCMC iteration, we first impute values below the threshold conditional on observations above the threshold.
+This is feasible for large datasets with our model because for a single day, conditional on the model parameters, we only need to draw from a truncated multivariate normal distribution.
+We can use Gibbs sampling to update $Y_t(\bs)$ for censored observations that are below the threshold $T$.
+After conditioning on $\lambda$, $z_t(\bs)$ and non-censored observations, $Y_t(\bs)$ has truncated normal full conditionals.
+So we sample $Y_t(\bs) \sim N_{(-\infty, T)}(\bX_t^T(\bs) \beta + \lambda | z_t(\bs)|, \bSigma)$.
+
+Then, we update model parameters, $\Theta$, using Metropolis Hastings or Gibbs sampling when appropriate.
+The final step of the computation is to use Bayesian Kriging to generate a predictive distribution for $Y_t(\bs^*)$ at prediction location $\bs^*$.
+This step is similar to the imputation for censored observations except that the full conditionals are no longer truncated at $T$.
+See Appendices A.1 and A.2 for details regarding the MCMC.
+
 \section{Simulation study}\label{s:simstudy}
@@
 For designs 2 and 3, $\sigma^2_{tk} \iid \text{IG}(3, 8)$ to give a $t$ distribution with 6 degrees of freedom.
-For designs 1, we set $\lambda = 0$.
+For design 1, we set $\lambda = 0$.
 For designs 2 and 3, $\lambda = 3$ was used as to simulate moderate skewness, and the $z_t$ are generated as described in (\ref{eq:sitezsig}).
@@
 For this setting, we set $\gamma = 0.5$.
-In the final design, we generate $\tilde{y}$ using the setting from design two, and then consider the data
+In the final design, we generate $\tilde{y}$ using the setting from design 2, and then transform the data
 \begin{align}
@@
 where $T = q(0.80)$ is the 80th sample quantile of the data.
-In all five designs, the mean $\bX^T \bbeta = 10$ is assumed to be constant across space.
+The final design is included to explore the importance of using a threshold exceedance model when the distribution for the bulk of the data is misspecified.
+In all six designs, the mean $\bX^T \bbeta = 10$ is assumed to be constant across space.
 
@@
 The knots have priors $\bw \sim \text{Unif}(\calD)$.
-We tried also fitting the skew-$t$ marginals for the thresholded models, but it is very challenging for the MCMC to properly identify the skewness parameter with only one tail worth of data.
-Each chain of the MCMC ran for 20000 iterations with a burn-in period of 10000 iterations.
+We tried also fitting the skew-$t$ marginals for the thresholded models, but it is very challenging for the MCMC to properly identify the skewness parameter with a censored left tail.
+Each chain of the MCMC ran for 20,000 iterations with a burn-in period of 10,000 iterations.
 Parameters appear to converge properly; however, in the models with multiple partitions (i.e.models 4 and 5) it is hard to assess the convergence of $\bw$, $z(\bs)$, and $\sigma^2(\bs)$ because of partition label switching throughout the MCMC.
@@
 \end{align}
-We analyzed the results for the simulation study using a Friedman test at $\alpha = 0.05$.
-If the Friedman test came back with a significant results, we conducted a Wilcoxon-Nemenyi-McDonald-Thompson test to see which methods had different results.
+We analyzed the results for the simulation study using a Friedman test at $\alpha = 0.05$ to see if at least one method had a significantly different Brier score.
+For Friedman tests that came back with a significant p-value, we conducted a Wilcoxon-Nemenyi-McDonald-Thompson test to see which of the methods had different results.
 The full results for the Wilcoxon-Nemenyi-McDonald-Thompson tests are given in Appendix \ref{a:pdiffs}.
@@
 Furthermore in these data settings, we find the best performance occurs when the number of knots used in the method matches the number of knots used for data generation.
-The non-thresholded methods tend to outperform the thresholded methods, but this is not surprising given that the data are generated directly from the model used in the method.
+The non-thresholded methods tend to outperform the thresholded methods, but this is not surprising given that in most cases, the data are generated directly from the model used in the method.
 For the max-stable data, we see that for low-extreme quantiles, the Gaussian method performs better, for more extreme quantiles, the single-partition method, both thresholded and non-thresholded, perform significantly better than the Gaussian.
 Finally, for setting 5, although the thresholded version of the single-partition model tends to perform the best across all of the extreme quantiles, the difference between the thresholded and non-thresholded methods is no longer significant in the more extreme quantiles.
+In summary, our method provides more flexibility for data that demonstrate some level of asymmetry or heavy tails, while still performing comparably to Gaussian methods when the data are symmetric and have light tails.
 
@@
 For a bivariate skew-$t$ random variable $\bY = [Y(\bs), Y(\bt)]^T$, the $\chi(h)$ statistic \citep{Padoan2011} is given by
-\begin{align}
+\begin{align} \label{eq:chiskew-t}
   \chi(h) = \bar{F}_{\text{EST}}\left\{ \frac{[x_1^{1 / \eta} - \varrho(h)] \sqrt{\eta + 1} }{\sqrt{1 - \varrho(h)^2}}; 0, 1, \alpha_1, \tau_1, \eta + 1 \right\} + \bar{F}_{\text{EST}}\left\{ \frac{ [x_2^{1 / \eta} - \varrho(h)] \sqrt{\eta + 1} }{ \sqrt{1 - \varrho(h)^2} }; 0, 1, \alpha_2, \tau_2, \eta + 1 \right\},

### 2015-07-22 LaTeX/skewnormal.tex
updates to results with max-stable stuff
update to results in appendix
updates to the manuscript with results

@@
 The sites are generated Uniform$([0, 10] \times [0, 10])$.
-We generate data from 5 different simulation designs:
+We generate data from 4 different simulation designs:
 \begin{enumerate} \setlength{\itemsep}{-0.5em}
@@
   \item Max-stable
-  \item Transformation below $T = q(0.80)$
+  % \item Transformation below $T = q(0.80)$
 \end{enumerate}
@@
 For designs 2 and 3, $\lambda = 3$ was used as to simulate moderate skewness, and the $z_t$ are generated as described in (\ref{eq:sitezsig}).
+In designs 1 -- 3, the mean $\bX^T \bbeta = 10$ is assumed to be constant across space.
 In the fourth design, we generate from a spatial max-stable distribution \citep{Reich2012}.
@@
 For this setting, we set $\gamma = 0.5$.
-In the final design, we generate $\tilde{y}$ using the setting from design 2, and then transform the data
-\begin{align}
-  y = \left\{ \begin{array}{lc}
-    \tilde{y}, \quad & \tilde{y} > T \\[0.5em]
-    T \exp\{\tilde{y} - T\}, \quad & \tilde{y} \le T
-  \end{array}\right.
-\end{align}
-where $T = q(0.80)$ is the 80th sample quantile of the data.
-The final design is included to explore the importance of using a threshold exceedance model when the distribution for the bulk of the data is misspecified.
-In all six designs, the mean $\bX^T \bbeta = 10$ is assumed to be constant across space.
+% In the final design, we generate $\tilde{y}$ using the setting from design 2, and then transform the data
+% \begin{align}
+%   y = \left\{ \begin{array}{lc}
+%     \tilde{y}, \quad & \tilde{y} > T \\[0.5em]
+%     T \exp\{\tilde{y} - T\}, \quad & \tilde{y} \le T
+%   \end{array}\right.
+% \end{align}
+% where $T = q(0.80)$ is the 80th sample quantile of the data.
+% The final design is included to explore the importance of using a threshold exceedance model when the distribution for the bulk of the data is misspecified.
 
 $M = 50$ data sets are generated for each design.
-For each data set we fit the data using five models
+For each data set we fit the data using six models
 \begin{enumerate} \setlength{\itemsep}{-0.5em}
@@
 
-Figure \ref{fig:simbrierscores} shows that when the data come from a Gaussian process, our methods perform comparably to the Gaussian method.
-For data settings with skew-$t$ marginals (settings 2 -- 3), we find significant improvement over the Gaussian method.
-Furthermore in these data settings, we find the best performance occurs when the number of knots used in the method matches the number of knots used for data generation.
+A plot of the relative Brier scores is given in Figure \ref{fig:simbrierscores}.
+We find that when the data are generated from a Gaussian process, our method performs comparably to a Gaussian approach.
+In general, when the underlying process is not Gaussian, our method results in an improvement over both the max-stable and Gaussian methods.
+One exception to this is the case when the generative process is max-stable.
+In this case, the max-stable method outperforms our method; however, for predictions further out in the tail, the differences between the max-stable method and our method diminish.
 The non-thresholded methods tend to outperform the thresholded methods, but this is not surprising given that in most cases, the data are generated directly from the model used in the method.
-For the max-stable data, we see that for low-extreme quantiles, the Gaussian method performs better, for more extreme quantiles, the single-partition method, both thresholded and non-thresholded, perform significantly better than the Gaussian.
-Finally, for setting 5, although the thresholded version of the single-partition model tends to perform the best across all of the extreme quantiles, the difference between the thresholded and non-thresholded methods is no longer significant in the more extreme quantiles.
+% Finally, for setting 5, although the thresholded version of the single-partition model tends to perform the best across all of the extreme quantiles, the difference between the thresholded and non-thresholded methods is no longer significant in the more extreme quantiles.
 In summary, our method provides more flexibility for data that demonstrate some level of asymmetry or heavy tails, while still performing comparably to Gaussian methods when the data are symmetric and have light tails.
@@
 \section{Data analysis}\label{s:analysis}
-To illustrate this method, we consider the daily maximum 8-hour ozone measurements for July 1 - 31, 2005 at 1089 Air Quality System (AQS) monitoring sites in the United States as the response (see Figure \ref{fig:ozone}).
+To illustrate this method, we consider 31 daily observations of maximum 8-hour ozone measurements for July 2005 at 1089 Air Quality System (AQS) monitoring sites in the United States as the response (see Figure \ref{fig:ozone}).
 For each site, we also have covariate information containing the estimated ozone from the Community Multi-scale Air Quality (CMAQ) modeling system.
@@
 \begin{align}
-  \bX^T_t(\bs) \bbeta = \beta_0 + \beta_1 \cdot \text{CMAQ}_t(\bs). \label{eq:datamean}
+  \E[Y_i(\bs)] = \beta_0 + \beta_1 \cdot \text{CMAQ}_t(\bs). \label{eq:datamean}
 \end{align}
@@
 \end{figure}
-Exploratory data analysis indicates that there is dependence in the high quantile levels of the residuals beyond what we expect in the case of independence.
+Standard exploratory data analysis techniques for extremal dependence are very challenging with only 31 days worth of data because it is difficult to estimate extreme quantiles at each site to obtain empirical estimates of $\chi$.
+Despite the fact that there is only one month of data, we can get some sense of extremal dependence between sites by looking at joint occurences of high quantiles.
+Figure \ref{fig:bivariateozone} illustrates the differing levels of dependence in extreme quantiles for a pair of sites that are close to one another versus a pair of sites that are far apart.
+\begin{figure}
+  \label{fig:bivariateozone}
+  \centering
+  \includegraphics[width=\linewidth]{plots/daily-quantiles-ozone.pdf}
+  \caption{Daily quantiles for sites that are close (left) and far apart (right)}
+\end{figure}
 
@@
   \label{tbl:gaussim}
-  \begin{tabular}{|l|ccc|ccc|ccc|cc|}
-    \cline{2-12}
-    \multicolumn{1}{c}{} & \multicolumn{3}{|c}{$q(0.90)$} & \multicolumn{3}{|c}{$q(0.95)$} & \multicolumn{3}{|c}{$q(0.98)$} & \multicolumn{2}{|c|}{$q(0.99)$} \\
+  \begin{tabular}{|l|ccc|ccc|cccc|ccc|}
+    \cline{2-14}
+    \multicolumn{1}{c}{} & \multicolumn{3}{|c}{$q(0.90)$} & \multicolumn{3}{|c}{$q(0.95)$} & \multicolumn{4}{|c}{$q(0.98)$} & \multicolumn{3}{|c|}{$q(0.99)$} \\
     \hline
-    Method 1 & A &   &   & A &   &   & A &   &   & A &   \\
+    Method 1 & A &   &   & A &   &   & A &   &   &   & A & B &   \\
     \hline
-    Method 2 & A & B &   & A & B &   & A &   &   & A &   \\
+    Method 2 & A &   &   & A &   &   & A &   &   &   & A &   &   \\
     \hline
-    Method 3 &   &   & C &   &   & C &   &   & C &   & B \\
+    Method 3 &   & B &   &   & B &   &   &   & C &   &   & B &   \\
     \hline
-    Method 4 &   & B &   &   & B &   &   & B &   &   & B \\
+    Method 4 & A &   &   & A &   &   & A & B &   &   & A & B &   \\
     \hline
-    Method 5 &   &   & C &   &   & C &   &   & C &   & B \\
+    Method 5 &   & B &   &   & B &   &   & B & C &   & A & B &   \\
+    \hline
+    Method 6 &   &   & C &   &   & C &   &   &   & D &   &   & C \\
     \hline
@@
   \label{tbl:st1sim}
-  \begin{tabular}{|l|ccc|ccc|ccc|ccc|}
-    \cline{2-13}
-    \multicolumn{1}{c}{} & \multicolumn{3}{|c}{$q(0.90)$} & \multicolumn{3}{|c}{$q(0.95)$} & \multicolumn{3}{|c}{$q(0.98)$} & \multicolumn{3}{|c|}{$q(0.99)$} \\
+  \begin{tabular}{|l|ccccc|cccc|cccc|ccc|}
+    \cline{2-17}
+    \multicolumn{1}{c}{} & \multicolumn{5}{|c}{$q(0.90)$} & \multicolumn{4}{|c}{$q(0.95)$} & \multicolumn{4}{|c}{$q(0.98)$} & \multicolumn{3}{|c|}{$q(0.99)$} \\
     \hline
-    Method 1 &   & B &   &   & B &   &   & B &   &   & B &   \\
+    Method 1 &   &   & C &   &   &   & B &   &   &   & B & C &   &   & B &   \\
     \hline
-    Method 2 & A &   &   & A &   &   & A &   &   & A &   &   \\
+    Method 2 & A &   &   &   &   & A &   &   &   & A &   &   &   & A &   &   \\
     \hline
-    Method 3 &   & B &   &   & B &   &   & B &   & A & B &   \\
+    Method 3 &   & B & C &   &   & A & B &   &   & A & B &   &   & A & B &   \\
     \hline
-    Method 4 &   & B &   &   & B &   &   & B &   &   & B &   \\
+    Method 4 & A & B &   &   &   &   & B &   &   &   & B &   &   & A &   &   \\
     \hline
-    Method 5 &   &   & C &   &   & C &   &   & C &   &   & C \\
+    Method 5 &   &   &   & D &   &   &   & C &   &   &   & C &   &   & B &   \\
+    \hline
+    Method 6 &   &   &   &   & E &   &   &   & D &   &   &   & D &   &   & C \\
     \hline
@@
   \label{tbl:st5sim}
-  \begin{tabular}{|l|cc|cc|cc|ccc|}
-    \cline{2-10}
-    \multicolumn{1}{c}{} & \multicolumn{2}{|c}{$q(0.90)$} & \multicolumn{2}{|c}{$q(0.95)$} & \multicolumn{2}{|c}{$q(0.98)$} & \multicolumn{3}{|c|}{$q(0.99)$} \\
+  \begin{tabular}{|l|ccc|cccc|ccc|ccc|}
+    \cline{2-14}
+    \multicolumn{1}{c}{} & \multicolumn{3}{|c}{$q(0.90)$} & \multicolumn{4}{|c}{$q(0.95)$} & \multicolumn{3}{|c}{$q(0.98)$} & \multicolumn{3}{|c|}{$q(0.99)$} \\
     \hline
-    Method 1 &   & B &   & B &   & B &   &   & C \\
+    Method 1 &   & B &   &   &   & C &   &   & B &   &   & B &   \\
     \hline
-    Method 2 &   & B &   & B &   & B &   & B & C \\
+    Method 2 &   & B &   &   &   & C &   &   & B &   &   & B &   \\
     \hline
-    Method 3 & A &   & A &   &   & B &   &   & C \\
+    Method 3 & A &   &   &   & B &   &   &   & B &   &   & B &   \\
     \hline
-    Method 4 & A &   & A &   & A &   & A & B &   \\
+    Method 4 & A &   &   & A &   &   &   & A &   &   & A &   &   \\
     \hline
-    Method 5 & A &   & A &   & A &   & A &   &   \\
+    Method 5 & A &   &   & A &   &   &   & A &   &   & A &   &   \\
+    \hline
+    Method 6 &   &   & C &   &   &   & D &   &   & C &   &   & C \\
     \hline
@@
     \hline
-    Method 1 & A &   &   &   & A &   &   &   & A & B &   &   &   & C \\
+    Method 1 & A & B &   &   &   & B &   &   &   & B &   &   &   & C \\
     \hline
-    Method 2 &   & B &   &   & A & B &   &   & A &   &   & A & B &   \\
+    Method 2 &   & B &   &   &   & B & C &   &   & B &   &   & B & C \\
     \hline
-    Method 3 &   &   & C &   &   & B & C &   & A & B &   & A &   &   \\
+    Method 3 &   &   & C & D &   &   & C &   &   & B &   &   & B &   \\
     \hline
@@
     \hline
-    Method 5 &   &   & C & D &   &   & C &   &   & B &   &   & B & C \\
+    Method 5 &   &   & C &   &   &   & C &   &   & B &   &   & B & C \\
+    \hline
+    Method 6 & A &   &   &   & A &   &   &   & A &   &   & A &   &   \\
     \hline
@@
   \label{tbl:transsim}
-  \begin{tabular}{|l|ccc|cc|ccc|ccc|}
-    \cline{2-12}
-    \multicolumn{1}{c}{} & \multicolumn{3}{|c}{$q(0.90)$} & \multicolumn{2}{|c}{$q(0.95)$} & \multicolumn{3}{|c}{$q(0.98)$} & \multicolumn{3}{|c|}{$q(0.99)$} \\
+  \begin{tabular}{|l|cccc|ccc|cccc|cccc|}
+    \cline{2-16}
+    \multicolumn{1}{c}{} & \multicolumn{4}{|c}{$q(0.90)$} & \multicolumn{3}{|c}{$q(0.95)$} & \multicolumn{4}{|c}{$q(0.98)$} & \multicolumn{4}{|c|}{$q(0.99)$} \\
     \hline
-    Method 1 &   &   & C &   & B &   &   & C &   &   & C \\
+    Method 1 &   &   & C &   &   & B &   &   &   & C &   &   &   & C &   \\
     \hline
-    Method 2 &   & B &   &   & B & A & B &   & A & B &   \\
+    Method 2 &   & B &   &   &   & B &   &   & B &   &   & A & B &   &   \\
     \hline
-    Method 3 & A &   &   & A &   & A &   &   & A &   &   \\
+    Method 3 & A &   &   &   & A &   &   & A &   &   &   & A &   &   &   \\
     \hline
-    Method 4 &   & B & C &   & B &   & B &   &   & B & C \\
+    Method 4 &   & B & C &   &   & B &   &   & B &   &   &   & B & C &   \\
     \hline
-    Method 5 &   & B & C &   & B &   &   & C &   &   & C \\
+    Method 5 &   & B &   &   &   & B &   &   & B & C &   &   &   & C &   \\
+    \hline
+    Method 6 &   &   &   & D &   &   & C &   &   &   & D &   &   &   & D \\
     \hline

### 2015-07-22 LaTeX/skewnormal.tex
more manuscript updates

@@
 Standard exploratory data analysis techniques for extremal dependence are very challenging with only 31 days worth of data because it is difficult to estimate extreme quantiles at each site to obtain empirical estimates of $\chi$.
-Despite the fact that there is only one month of data, we can get some sense of extremal dependence between sites by looking at joint occurences of high quantiles.
-Figure \ref{fig:bivariateozone} illustrates the differing levels of dependence in extreme quantiles for a pair of sites that are close to one another versus a pair of sites that are far apart.
+Despite the fact that there is only one month of data, we can get some sense of extremal dependence between sites by looking at joint occurences of high sample quantiles.
+Figure \ref{fig:bivariateozone} suggests suggests there is more agreement between sites that are close to one another than sites that are far from one another.
 \begin{figure}

### 2015-07-23 LaTeX/skewnormal.tex
updates to plots and manuscript

@@
 Many types of data demonstrate some level of skewness and therefore should be modeled with distributions that allow for asymmetry.
-The skew-elliptical family of distributions provides models that are mathematically tractable while introducing a slant parameter to account for asymmetric data \citep{Genton2004}.
+The skew-elliptical family of distributions provides models that are mathematically tractable while introducing a slant parameter to account for asymmetric data \citep{Azzalini2014}.
 A brief review of the additive process by which a skew-$t$ process is created is given here.
@@
 \end{align}
-that is, $\bY$ follows an $n$-dimensional skew-$t$ distribution with location $\bX \bbeta$, correlation matrix $\bOmega$, slant parameters $\balpha$ and degrees of freedom $2a$, where $\bX = [\bX(\bs_1)^T, \ldots, \bX(\bs_n)^T]$, $\bOmega = \bomega \bar{\bOmega} \bomega$, $\bomega = \text{diag}\left(\frac{1}{ \sqrt{ab}}, \ldots, \frac{1}{\sqrt{ab}}\right)$, $\bar{\bOmega} = (\bSigma + \lambda^2 \bOne \bOne^T$), $\bSigma$ is the positive definite correlation matrix of $[v(\bs_1), \ldots, v(\bs_n)]$, $\balpha = \lambda (1 + \lambda^2 \bOne^T \bSigma^{-1} \bOne)^{-1/2} \bOne^T \bSigma^{-1}$ is a vector of slant parameters.
+that is, $\bY$ follows an $n$-dimensional skew-$t$ distribution with location $\bX \bbeta$, correlation matrix $\bOmega$, slant parameters $\balpha$ and degrees of freedom $2a$, where $\bX = [\bX(\bs_1)^T, \ldots, \bX(\bs_n)^T]$, $\bOmega = \bomega \bar{\bOmega} \bomega$, $\bomega = \text{diag}\left(\frac{1}{ \sqrt{ab}}, \ldots, \frac{1}{\sqrt{ab}}\right)$, $\bar{\bOmega} = (\bSigma + \lambda^2 \bOne \bOne^T$), $\balpha = \lambda (1 + \lambda^2 \bOne^T \bSigma^{-1} \bOne)^{-1/2} \bOne^T \bSigma^{-1}$ is a vector of slant parameters, and $\bSigma$ is the positive definite correlation matrix of $[v(\bs_1), \ldots, v(\bs_n)]$.
 Although any positive definite correlation function could be used, we choose to use the stationary isotropic \Matern correlation with
@@
 \begin{align}
-  \E[Y_i(\bs)] = \beta_0 + \beta_1 \cdot \text{CMAQ}_t(\bs). \label{eq:datamean}
+  \text{E}[Y_i(\bs)] = \beta_0 + \beta_1 \cdot \text{CMAQ}_t(\bs). \label{eq:datamean}
 \end{align}
@@
   \includegraphics[width=0.41\linewidth]{plots/qq-res.pdf}
-  \caption{Ozone values on 10 July 2005 (left) Q-Q plot of the residuals (right)}
+  \caption{Ozone values on 10 July 2005 (left) Q-Q plot of the residuals for a skew-$t$ distribution with 10 d.f. and $\alpha = 1$ (right)}
   \label{fig:ozone}
 \end{figure}
+
 Standard exploratory data analysis techniques for extremal dependence are very challenging with only 31 days worth of data because it is difficult to estimate extreme quantiles at each site to obtain empirical estimates of $\chi$.
@@
 
-Each chain of the MCMC ran for 30000 iterations with a burn-in period of 25000 iterations.
+Each chain of the MCMC ran for 30,000 iterations with a burn-in period of 25,000 iterations.
 Parameters appear to converge properly; however, as before, for models with multiple partitions it is hard to assess the convergence of $\bw$, $z(\bs)$, and $\sigma^2(\bs)$ because of partition label switching throughout the MCMC.
@@
 
-We illustrate the predictive capability of our model in Figure \ref{fig:ozoneq99} by plotting the 99th quantile of the posterior predictive density for July in South Carolina and Georgia.
-We fit the model using four methods, two reference and two that performed better.
-These four methods are
+We illustrate the predictive capability of our model in Figure \ref{fig:ozoneq99} by plotting the 99th quantile for South Carolina and Georgia, a subset of the spatial domain, in order to study local features.
+The four methods used are
 \begin{enumerate}\setlength{\itemsep}{-0.5em}
-  \item Gaussian (reference)
-  \item Skew-$t$, $K =$ 1 knot, $T = $ 0, no time series (reference)
-  \item Skew-$t$, $K =$ 5 knots, $T = $ 50, no time series (comparison)
-  \item Symmetric-$t$, $K =$ 10 knots, $T = $ 75, time series (comparison).
+  \item Gaussian
+  \item Skew-$t$, $K =$ 1 knot, $T = $ 0, no time series
+  \item Skew-$t$, $K =$ 5 knots, $T = $ 50, no time series
+  \item Symmetric-$t$, $K =$ 10 knots, $T = $ 75, time series.
 \end{enumerate}
@@
 
-\hl{NEED TO ADD STUFF HERE ACKNOWLEDGING THAT MARGINALS ARE DIFFERENT}
-Also, add an explanation about the purpose of the plot in \ref{fig:bs-marginal}
 \begin{figure}
   \centering
-  \includegraphics[width=0.5\linewidth]{plots/bs-ozone-marginal.pdf}
-  \caption{Comparison of Brier score performance by marginal quantile representing 75ppb}
-  \label{fig:bs-marginal}
-\end{figure}
-
-\begin{figure}
-  \includegraphics[width=\linewidth]{plots/q99-ozone.pdf}
-  \caption{(a) -- (d) give give the posterior predictive $\widehat{q}(0.99)$ for the month of July under four different models, (e) gives the difference between $\widehat{q}(0.99)$ in plots (d) and (a), (f) gives the difference between $\widehat{q}(0.99)$ in plots (d) and (b).}
+  \includegraphics[width=0.8\linewidth]{plots/q99-ozone.pdf}
+  \caption{Panels (a) -- (d) give the posterior predictive $\widehat{q}(0.99)$ for the month of July under four different models, panel (e) gives the difference between $\widehat{q}(0.99)$ in panels (d) and (a), panel (f) gives the difference between $\widehat{q}(0.99)$ in panels (d) and (b).}
   \label{fig:ozoneq99}
@@
 \subsection{Proof that $\lim_{h \rightarrow \infty} \pi(h) = 0$} \label{a:proofsamepartition}
-Let $N(A)$ be the number of knots in $A$, the area between sites $\bs_1$ and $\bs_2$.
-Consider a spatial Poisson process with intensity $\mu(A)$.
-So,
+Consider a homogeneous spatial Poisson process with intensity $\mu$.
+Define $A$ as the circle with center \hbox{$(\bs_1 + \bs_2) / 2$} and radius $h / 2$.
+Then $\bs_1$ and $\bs_2$ are in different partitions almost surely if two or more points are in $A$.
+Let $N(A)$ be the number of points in $A$, and let
 \begin{align*}
-  P[ N(A) = k] = \frac{ \mu(A)^k \exp\{ -\mu(A)\}}{k!}.
+  \mu(A) = \mu |A| = \mu \pi \left(\frac{h}{2}\right)^2 = \lambda h^2.
 \end{align*}
-Then for any finite $k$, $\lim_{h \rightarrow \infty} P[N(A) = k] = 0$ because $\lim_{h \rightarrow \infty} \mu(A) = \infty$.
-With each additional knot in $A$, the chance that $\bs_1$ and $\bs_2$ will be be in the same partition will decrease, because partition membership is defined by the closest knot to a site.
-Therefore, $\lim_{h \rightarrow \infty} \pi(h) = 0$.
+Then
+\begin{align*}
+  P[N(A) \ge 2] &= 1 - P[N(A) = 0] - P[N(A) = 1]\\
+                &= 1 - \exp\{-\lambda h^2\} - \lambda h^2 \exp\{-\lambda h^2\} \\
+                &= 1 - (1 + \lambda h^2) \exp\{-\lambda h^2\}
+\end{align*}
+which goes to one as $h \rightarrow \infty$.
+
+% Let $N(A)$ be the number of knots in $A$.
+% So,
+% \begin{align*}
+%   P[ N(A) = k] = \frac{ \mu(A)^k \exp\{ -\mu(A)\}}{k!}.
+% \end{align*}
+% Then for any finite $k$, $\lim_{h \rightarrow \infty} P[N(A) = k] = 0$ because $\lim_{h \rightarrow \infty} \mu(A) = \infty$.
+% With each additional knot in $A$, the chance that $\bs_1$ and $\bs_2$ will be be in the same partition will decrease, because partition membership is defined by the closest knot to a site.
+% Therefore, $\lim_{h \rightarrow \infty} \pi(h) = 0$.
 
@@
 \end{align}
-where $\bar{F}_{\text{EST}}$ is the univariate survival extended skew-$t$ function with zero location and unit scale, \hbox{$\varrho(h) = \text{cor}(y_1, y_2)$}, $\alpha_j = \alpha_i \sqrt{1 - \varrho^2}$, $\tau_j = \sqrt{\eta + 1}(\alpha_j + \alpha_i \varrho)$, and $x_j = F_T(\bar{\alpha}_i \sqrt{\eta + 1}; 0, 1, \eta) / F_T(\bar{\alpha}_j \sqrt{\eta + 1}; 0, 1, \eta)$ with $j = 1, 2$ and $i = 2, 1$ and where $\bar{\alpha}_j = (\alpha_j + \alpha_i \varrho) / \sqrt{ 1 + \alpha_i^2 [1 - \varrho(h)^2]}$.
+where $\bar{F}_{\text{EST}}$ is the univariate survival extended skew-$t$ function with zero location and unit scale, \hbox{$\varrho(h) = \text{cor}[y(\bs), y(\bt)]$}, $\alpha_j = \alpha_i \sqrt{1 - \varrho^2}$, $\tau_j = \sqrt{\eta + 1}(\alpha_j + \alpha_i \varrho)$, and $x_j = F_T(\bar{\alpha}_i \sqrt{\eta + 1}; 0, 1, \eta) / F_T(\bar{\alpha}_j \sqrt{\eta + 1}; 0, 1, \eta)$ with $j = 1, 2$ and $i = 2, 1$ and where $\bar{\alpha}_j = (\alpha_j + \alpha_i \varrho) / \sqrt{ 1 + \alpha_i^2 [1 - \varrho(h)^2]}$.
 

### 2015-07-23 code/analysis/ozone/US-all/us-all-results.R

@@ -829,7 +829,7 @@ library(fields)
 library(SpatialTools)
 
 # probability at least one day exceeds
-quartz(width=12, height=8)
+quartz(width=11, height=8)
 par(mfrow=c(2, 3))
 quilt.plot(x=S.p[, 1], y=S.p[, 2], matrix(set.1.p.atleast1), nx=nx, ny=ny,
            yaxt="n", xaxt="n")
@@ -927,8 +927,8 @@ quilt.plot(x=S.p[, 1], y=S.p[, 2], matrix(set.71.95), nx=nx, ny=ny,
 lines(borders/1000)
 
 # 99th quantiles
-quartz(width=12, height=12)
-par(mfrow=c(3, 2))
+quartz(width=6, height=9)
+par(mfrow=c(3, 2), mar = c(3, 2, 3, 2))
 quilt.plot(x=S.p[, 1], y=S.p[, 2], matrix(set.1.99), nx=nx, ny=ny,
            yaxt="n", xaxt="n", zlim=c(60, 120),
            main="(a) Gaussian", cex=1.5)
@@ -1264,35 +1264,41 @@ plot(fit[[2]]$beta[, 1], type="l")
 plot(fit[[2]]$beta[, 2], type="l")
 
 # get an idea of two sites that are close to one another vs far apart
+rm(list=ls())
+library(fields)
 load("us-all-setup.RData")
 source("../../../R/auxfunctions.R")
 settings <- read.csv("settings.csv")
 no.na <- which(rowSums(is.na(Y)) == 0)
 indices <- 1:nrow(S)
+
+# remove the NAs
 S <- S[no.na, ]
 indices <- indices[no.na]
+
+# find the sites that are close to one another
 d <- rdist(S) 
 diag(d) <- 0
 
+close <- which(d < 0.012 & d > 0, arr.ind = TRUE)
+close.these <- duplicated(rowSums(close))
+close <- close[!close.these, ]
 far <- which(d > 4.46, arr.ind = TRUE)
+far.these <- duplicated(rowSums(far))
+far <- far[!far.these, ]
 
 quant.close.1 <- quant.close.2 <- quant.far.1 <- quant.far.2 <- rep(NA, ncol(Y))
-ids.1 <- ids.2 <- fill.count <- rep(NA, nrow(close))
 for (i in 1:nrow(close)) {
[…]

### 2015-07-24 LaTeX/skewnormal.tex
addition of abstract and authors

@@
 \begin{document}\linenumbers
-
 \begin{center}
 {\Large {\bf A spatial skew-$t$ model for threshold exceedances}}\\
+
+{\large Samuel A Morris\footnote[1]{North Carolina State University}, Brian J Reich\footnotemark[1]{}, Emeric Thibaud\footnote[2]{Colorado State University}, and Dan Cooley\footnotemark[2]{}}
+
 \today
@@
 
+\begin{abstract}
+\singlespacing
+To assess the compliance of air quality regulations, the Environmental Protection Agency (EPA) must know if a site exceeds a pre-specified threshold.
+In the case of ozone, the threshold for compliance is fixed at 75 parts per billion, which is high, but not extreme at all locations.
+We present a new method based on the spatial skew-$t$ process.
+Our method incorporates a random partition to permit long-distance asymptotic independence while allowing for sites that are near one another to be asymptotically dependent, and we incorporate thresholding to allow the tails of the data to speak for themselves.
+We also introduce a transformed AR(1) time-series to allow for temporal dependence.
+Finally, our model allows for high-dimensional Bayesian inference that is comparable in speed to traditional geostatistical methods for large datasets.
+We apply our method to an ozone analysis for July 2005, and find that our model does demonstrate some improvement over both Gaussian and max-stable methods in terms of predicting exceedances over a fixed threshold.
+
+{\bf Key words}: Skew-$t$, random partition,
+\end{abstract}
+
+\newpage
 \section{Introduction}\label{s:intro}
@@
 These threshold models were extended to spatial models for extremes by \citet{Wadsworth2012} and \citet{Thibaud2013} who fit various models to spatial extremes using a censored pairwise likelihood \citep{Padoan2010} based on the approach of \citet{Ledford1996}.
-\citet{Huser2014} further extended this to spate-time modeling.
+\citet{Huser2014} further extended this to space-time modeling.
 \citet{Engelke2014}, \citet{Wadsworth2014}, and \citet{Thibaud2013a} introduced more efficient inference for threshold exceedances of extremal spatial processes with full likelihood methods.
-The previous approaches to threshold modeling are motivated by extreme value theory and assume the threshold is high enough to assume extremal models are valid for the data, and for extrapolation beyond the range of observed values.
+The previous approaches to threshold modeling are motivated by extreme value theory and assume the threshold is high enough that extremal models are valid for the data, and for extrapolation beyond the range of observed values.
 Moreover, these approaches are computationally intensive and limited to rather small datasets.
@@
 The computing is described in Section \ref{s:comp}.
-In Section \ref{s:simstudy}, we present a simulation study that examines the predictive capabilities of this model compared with a na{\"i}ve Gaussian method.
+In Section \ref{s:simstudy}, we present a simulation study that examines the predictive capabilities of this model compared Gaussian amd max-stable methods.
 We then compare our method to Gaussian and max-stable methods with a data analysis of ozone measurements from throughout the US in section \ref{s:analysis}.
@@
 \end{align}
-that is, $\bY$ follows an $n$-dimensional skew-$t$ distribution with location $\bX \bbeta$, correlation matrix $\bOmega$, slant parameters $\balpha$ and degrees of freedom $2a$, where $\bX = [\bX(\bs_1)^T, \ldots, \bX(\bs_n)^T]$, $\bOmega = \bomega \bar{\bOmega} \bomega$, $\bomega = \text{diag}\left(\frac{1}{ \sqrt{ab}}, \ldots, \frac{1}{\sqrt{ab}}\right)$, $\bar{\bOmega} = (\bSigma + \lambda^2 \bOne \bOne^T$), $\balpha = \lambda (1 + \lambda^2 \bOne^T \bSigma^{-1} \bOne)^{-1/2} \bOne^T \bSigma^{-1}$ is a vector of slant parameters, and $\bSigma$ is the positive definite correlation matrix of $[v(\bs_1), \ldots, v(\bs_n)]$.
+that is, $\bY$ follows an $n$-dimensional skew-$t$ distribution with location $\bX \bbeta$, correlation matrix $\bOmega$, slant parameters $\balpha$ and degrees of freedom $2a$, where $\bX = [\bX(\bs_1)^T, \ldots, \bX(\bs_n)^T]$, $\bOmega = \bomega \bar{\bOmega} \bomega$, $\bomega = \text{diag}\left(\frac{1}{ \sqrt{ab}}, \ldots, \frac{1}{\sqrt{ab}}\right)$, $\bar{\bOmega} = (\bSigma + \lambda^2 \bOne \bOne^T$), $\balpha = \lambda (1 + \lambda^2 \bOne^T \bSigma^{-1} \bOne)^{-1/2} \bOne^T \bSigma^{-1}$, and $\bSigma$ is the positive definite correlation matrix of $[v(\bs_1), \ldots, v(\bs_n)]$.
 Although any positive definite correlation function could be used, we choose to use the stationary isotropic \Matern correlation with
@@
 In this section, we propose extensions to the skew-$t$ process to model spatial extremes over a large geographic region by introducing censoring to focus on tail behavior and a random partition to remove long-range asymptotic dependence.
-For notational convenice, we introduce the model for a single replication, and then extend this model to the spatiotemporal setting in Section \ref{s:temporal}.
+For notational convenience, we introduce the model for a single replication, and then extend this model to the spatiotemporal setting in Section \ref{s:temporal}.
 
 \subsection{Censoring to focus on the tails}
-Because one of our goals is to model the dependence of the distribution in the tails of the data, we choose to censor values below threshold.
+Because one of our goals is to model the dependence of the distribution in the tails of the data, we choose to censor values below a threshold.
 Let
@@
 Modeling different levels of asymptotic dependence was discussed by \citet{Wadsworth2012} with a hybrid spatial dependence model.
-\citet{Huser2014} also allow for asymptotic dependence across both space and time with a partition structure represented by random discs moving across the space for a random duration with a random velocity and random radius.
+\citet{Huser2014} also allow for varying asymptotic dependence across both space and time with a partition structure represented by random discs moving across the space for a random duration with a random velocity and random radius.
 We handle the problem of long-range asymptotic dependence with a random partition.
@@
 Therefore, sites in different subregions are asymptotically independent because $\chi_{\text{Gaus}}(h) = 0$ for all $h$.
-Marginally, over the knots $\bw_1, \ldots, \bw_K$, $\chi(h) = \pi(h) \chi_{\text{skew}-t}(h)$, where $\pi(h) = \Pr(k_s = k_t)$ is the probability that two sites separated by distance $h$ are in the same partition.
+Marginally, over the knots, $\chi(h) = \pi(h) \chi_{\text{skew}-t}(h)$, where $\pi(h) = \Pr(k_s = k_t)$ is the probability that two sites separated by distance $h$ are in the same partition.
 So, to show that $\lim_{h \rightarrow \infty} \chi(h) = 0$, we need only know that $\lim_{h \rightarrow \infty} \pi(h) = 0$. A proof of this is given in Appendix \ref{a:proofsamepartition}.
@@
 \end{align}
-We allow the partition structure to vary from day to day in order to account for sharp spikes in ozone that may not be present every day (e.g. a forest fire).
+We allow the partition structure to vary from day to day in order to account for sharp spikes in a response that may not be present every day (e.g. the impact of a forest fire on ozone levels).
 
@@
 The time series model must be specified after a transformation to preserve the skew-$t$ process at each time point.
-For each time-varying parameter, we transform to obtain a standard normal marginal distribution, place a Gaussian prior with autocorrelation on the transformed parameter, and then transform back to obtain the marginal distribution required to preserve the skew-$t$ process.
+For each time-varying parameter, we transform to obtain a standard normal marginal distribution, place a Gaussian prior with autocorrelation on the transformed parameter, and then transform back to the appropriate marginal distribution for the skew-$t$ process.
 We first transform the spatial knots from $\calD$ to $\calR^2$ as follows.
@@
 \subsection{Cross validation}\label{s:modelselect}
-Models were compared using cross validation with 100 sites used as training sites and 44 sites witheld for testing.
+Models were compared using cross validation with 100 sites used as training sites and 44 sites withheld for testing.
 The model was fit using the training set, and predictions were generated at the testing site locations.
@@
 One exception to this is the case when the generative process is max-stable.
-In this case, the max-stable method outperforms our method; however, for predictions further out in the tail, the differences between the max-stable method and our method diminish.
+In this case, the max-stable method outperforms our method; however, for predictions at high quantile levels, the differences between the max-stable method and our method decrease.
 The non-thresholded methods tend to outperform the thresholded methods, but this is not surprising given that in most cases, the data are generated directly from the model used in the method.
@@
 Standard exploratory data analysis techniques for extremal dependence are very challenging with only 31 days worth of data because it is difficult to estimate extreme quantiles at each site to obtain empirical estimates of $\chi$.
-Despite the fact that there is only one month of data, we can get some sense of extremal dependence between sites by looking at joint occurences of high sample quantiles.
-Figure \ref{fig:bivariateozone} suggests suggests there is more agreement between sites that are close to one another than sites that are far from one another.
+Despite the fact that there is only one month of data, we can get some sense of extremal dependence between sites by looking at joint occurrences of high sample quantiles.
+For example, Figure \ref{fig:bivariateozone} suggests there is more agreement between sites that are close to one another than sites that are far from one another.
 \begin{figure}
-  \label{fig:bivariateozone}
   \centering
@@
   \caption{Daily quantiles for sites that are close (left) and far apart (right)}
+  \label{fig:bivariateozone}
 \end{figure}
+Another aspect that distinguishes our approach from more traditional extremes analyses, is how the threshold is selected.
+Traditionally, a threshold is chosen based on the assumption that beyond the threshold, the data follow an extremal model.
+In our example, a threshold of 75 ppb which is $q(0.92)$ for all observations, but marginally it represents anywhere from $q(0.06)$ to $q(1)$.
 
@@
 The upper threshold of 75 ppb was used because the current air quality standard is based on exceedance of 75 ppb.
-As with the simuliation study, for models with a threshold of $T = 75$, we use a symmetric-$t$ marginal distribution.
+As with the simulation study, for models with a threshold of $T = 75$, we use a symmetric-$t$ marginal distribution.
 We also compare models with no time series to models that include the time series.
@@
 All methods assume the mean function given in (\ref{eq:datamean}).
-To ensure that the max-stable method runs in a reasonable amount of time, we take a stratified sample of the sites to get 800 sites and consider this our new dataset.
+To ensure that the max-stable method runs in a reasonable amount of time, we take a stratified sample of 800 sites and consider this our new dataset.
 We conduct two-fold cross validation using 400 training sites and 400 validation sites as described in Section \ref{s:modelselect}
@@
 The proposed model gives flexible tail behavior, demonstrates asymptotic dependence for observations at sites that are near to one another, and has computation on the order of Gaussian models for large space-time datasets.
-In the simulation study, we demonstrate that this model shows statistically significant improvements over a na\"{i}ve Gaussian approach.
-In both the simulation study, and the application to ozone data, we find that incorporating a partition in the model improves extreme prediction.
+In the simulation study, we demonstrate that this model shows statistically significant improvements over a na\"{i}ve Gaussian approach and in most cases, a max-stable approach.
+In both the simulation study, and the application to ozone data, we find that incorporating a partition in the model can improve extreme predictions.
 Furthermore the results from the data analysis suggest that thresholding can improve performance when predicting in the extreme tails of the data.
@@
 Finally, we acknowledge that by specifying the number of knots, we may be underestimating the uncertainty in the model.
-This could be incoporated by treating the number of knots as a model parameter instead of fixing it to be a specific value.
+This could be incorporated by treating the number of knots as a model parameter instead of fixing it to be a specific value.
 

### 2015-08-06 LaTeX/skewnormal.tex
updates to manuscript from Emeric and Dan

@@
 \begin{center}
-{\Large {\bf A spatial skew-$t$ model for threshold exceedances}}\\
+{\Large {\bf A space-time skew-$t$ model for threshold exceedances}}\\
 
-{\large Samuel A Morris\footnote[1]{North Carolina State University}, Brian J Reich\footnotemark[1]{}, Emeric Thibaud\footnote[2]{Colorado State University}, and Dan Cooley\footnotemark[2]{}}
+{\large Samuel A Morris\footnote[1]{North Carolina State University}, Brian J Reich\footnotemark[1]{}, Emeric Thibaud\footnote[2]{Colorado State University}, and Daniel Cooley\footnotemark[2]{}}
 
@@
 Finally, our model allows for high-dimensional Bayesian inference that is comparable in speed to traditional geostatistical methods for large datasets.
-We apply our method to an ozone analysis for July 2005, and find that our model does demonstrate some improvement over both Gaussian and max-stable methods in terms of predicting exceedances over a fixed threshold.
+We apply our method to an ozone analysis for July 2005, and find that our model improves over both Gaussian and max-stable methods in terms of predicting exceedances over a fixed threshold.
 
-{\bf Key words}: Skew-$t$, random partition,
+{\bf Key words}: Skew-$t$, random partition, MCMC, extreme value analysis, spatio-temporal
 \end{abstract}
@@
 Our study is motivated by an air pollution application where the focus is not on the average behavior, but instead the behavior over a fixed threshold determined by government regulation.
-More specifically, we consider consider the case of compliance for ozone.
+More specifically, we consider the case of compliance for ozone.
 A site is said to be in compliance if the fourth highest daily maximum 8-hour concentration averaged over 3 years does not exceed 75 parts per billion (ppb).
+Figure \ref{fig:ozone-10jul} shows the ozone levels from July 10, 2005, at 1089 stations across the United States.
+We see a large area above the compliance level in the midwest covering Ohio, Indiana, Illinois, and parts of the surrounding states.
+\begin{figure}
+  \centering
+  \includegraphics[width=\linewidth]{plots/ozone-10jul-us.pdf}
+  \caption{Ozone values on July 10, 2005}
+  \label{fig:ozone-10jul}
+\end{figure}
 
-Traditional geostatistical modeling is based upon the assumption that observations come from a Gaussian process, a process that is fully defined by its mean and covariance functions.
-In the limit of the Gaussian distribution, all observations are independent regardless of the strength of the correlation in the bulk of the data.
-Furthermore, the Gaussian distribution is light-tailed and symmetric.
-Therefore, it is inappropriate to use standard geostatistical methods when trying to describe dependence in the tails of the distribution.
+A spatial model for threshold exceedances warrants special consideration and standard spatial methods are likely to perform poorly.
+First, because we are interested only in high values, we want to ``let the tail speak for itself''.
+That is, if we fit a model to the entire data set, low-to-moderate values would influence the fit of the overall model.
+As there are more of these values, they can unduly influence the distribution at the higher levels about which we are interested.
+Our inference method will only use data which exceed a threshold, and will impute data below the threshold, thereby tailoring the fit to the levels of interest.
+Second, likelihood-based spatial modeling typically assumes a Gaussian process, which is appropriate when mean behavior is of interest.
+However, the Gaussian distribution is light-tailed and symmetric, and therefore may be inappropriate for modeling data which does not share this tail behavior.
+Third, we aim to capture the dependence structure when ozone is at high levels, and dependence at these levels may not be well-represented by covariances which focus again on mean behavior.
+Asymptotic dependence/independence (see Section \ref{s:extdep}) are notions which describe how two random variables' dependence behaves as one looks further into the joint tail.
+The Gaussian distribution always exhibits asymptotic independence, except in the case of perfect dependence, thus is an inappropriate model for data which exhibits asymptotic dependence.
+To allow for more flexibility in the marginal tail and to allow for asymptotic dependence, the skew-$t$ distribution forms the basis for our model.
 
-Threshold modeling is popular in the field of extreme value statistics where extreme events are naturally defined in terms of exceedances over a high threshold.
+Our approach differs from threshold modeling approaches based on extreme value distributions.
+There has been extensive work on threshold modeling in the field of extreme value statistics where extreme events are naturally defined in terms of exceedances over a high threshold.
 \citet{Davison1990} considered modeling threshold exceedances of univariate time series by the generalized Pareto distribution.
@@
 \citet{Huser2014} further extended this to space-time modeling.
-\citet{Engelke2014}, \citet{Wadsworth2014}, and \citet{Thibaud2013a} introduced more efficient inference for threshold exceedances of extremal spatial processes with full likelihood methods.
+\citet{Thibaud2013a}, \citet{Engelke2014}, and \citet{Wadsworth2014}, introduced more efficient inference for threshold exceedances of extremal spatial processes with full likelihood methods.
 The previous approaches to threshold modeling are motivated by extreme value theory and assume the threshold is high enough that extremal models are valid for the data, and for extrapolation beyond the range of observed values.
 Moreover, these approaches are computationally intensive and limited to rather small datasets.
+For example, \citet{Wadsworth2014} present a simulation study with observations at 16 sites on a regular grid, and \citet{Engelke2014} analyze a dataset with observations at 35 meteorological stations.
 Our application with ozone data does not fit into this framework because we do not focus on exceedances of a very high threshold, but on exceedances of a fixed threshold.
+Furthermore, in our application, we have observations at over 1,000 ozone monitoring locations.
 
-Instead, we propose a new spatiotemporal threshold exceedance model based on the skew-$t$ process \citep{Padoan2011}.
-Our model is a threshold exceedance model for the multivariate skew-$t$ distribution that uses imputation for values below a fixed threshold.
+We propose a new spatiotemporal threshold exceedance model based on the skew-$t$ process \citep{Padoan2011}.
+Our model is a threshold exceedance model for the multivariate skew-$t$ distribution for a fixed threshold.
+In this setting, we describe the threshold as fixed because it is specified in advance by regulatory compliance.
+This differs from the more traditional extremes literature where a threshold is selected to be the value beyond which an extremal model is appropriate for the data.
 We use a skew-$t$ distribution because of its flexibility to model asymmetry and heavy-tailed data with the aim of predicting the probability of exceeding a high fixed threshold at an unobserved location.
 
+Our model allows for inference and predictions using the full likelihood with computing on the order of Gaussian models for large space-time datasets.
+This allows us to use Bayesian methods to impute data below the threshold as well as make predictions at unobserved locations.
 In a spatial setting, the multivariate skew-$t$ distribution demonstrates asymptotic dependence between observations at all sites regardless of the distance between the sites.
@@
 This partition alleviates the asymptotic spatial dependence present in the skew-$t$ distribution for sites that are far apart.
-Finally, our model allows for inference and predictions using the full likelihood with computing on the order of Gaussian models for large space-time datasets.
 
@@
 Section \ref{s:spatialskew} is a brief review of the spatial skew-$t$ process.
-In Section \ref{s:temporal}, we build upon the traditional skew-$t$ by incorporating censoring to focus on tails, partitioning to remove long-range asymptotic dependence, and extending the model to space-time data.
-The computing is described in Section \ref{s:comp}.
+In Section \ref{s:spatial}, we build upon the traditional skew-$t$ process by incorporating censoring to focus on tails, partitioning to remove long-range asymptotic dependence, and extending the model to space-time data.
+The computing is described in Section \ref{s:hier}.
 In Section \ref{s:simstudy}, we present a simulation study that examines the predictive capabilities of this model compared Gaussian amd max-stable methods.
-We then compare our method to Gaussian and max-stable methods with a data analysis of ozone measurements from throughout the US in section \ref{s:analysis}.
+We then compare our method to Gaussian and max-stable methods with a data analysis of ozone measurements from 800 sites throughout the US in Section \ref{s:analysis}.
 The final section provides brief discussion and direction for future research.
@@
 \subsection{Skew-$t$ process} \label{s:skewt}
-Let $Y(\bs)$ be the observation at spatial location $\bs = (s_1, s_2)$ in a spatial domain of interest $\calD \in \calR^2$.
+Let $Y(\bs)$ be the observation at spatial location $\bs$ in a spatial domain of interest $\calD \in \calR^2$.
 The spatial skew-$t$ process can be written
@@
 \end{align}
-where $\bX(\bs)$ is a set of spatial covariates at site $\bs$, $\bbeta$ is the vector of regression parameters, $\lambda \in \calR$ is a parameter controlling skew, $z \sim N(0, 1)$, $\sigma^2 \sim \text{IG}(a, b)$ is an inverse gamma random variable, and $v(\bs)$ is a spatial Gaussian process with mean zero, variance one, and a positive definite correlation function.
+where $\bX(\bs)$ is a set of spatial covariates at site $\bs$, $\bbeta$ is the vector of regression parameters, $\lambda \in \calR$ is a parameter controlling skew, $z \sim N(0, 1)$, $\sigma^2 \sim \text{IG}(a, b)$ is random scale parameter, IG is the distribution function of an inverse gamma random variable, and $v(\bs)$ is a spatial Gaussian process with mean zero, variance one, and a positive definite correlation function.
 
@@
 that is, $\bY$ follows an $n$-dimensional skew-$t$ distribution with location $\bX \bbeta$, correlation matrix $\bOmega$, slant parameters $\balpha$ and degrees of freedom $2a$, where $\bX = [\bX(\bs_1)^T, \ldots, \bX(\bs_n)^T]$, $\bOmega = \bomega \bar{\bOmega} \bomega$, $\bomega = \text{diag}\left(\frac{1}{ \sqrt{ab}}, \ldots, \frac{1}{\sqrt{ab}}\right)$, $\bar{\bOmega} = (\bSigma + \lambda^2 \bOne \bOne^T$), $\balpha = \lambda (1 + \lambda^2 \bOne^T \bSigma^{-1} \bOne)^{-1/2} \bOne^T \bSigma^{-1}$, and $\bSigma$ is the positive definite correlation matrix of $[v(\bs_1), \ldots, v(\bs_n)]$.
-Although any positive definite correlation function could be used, we choose to use the stationary isotropic \Matern correlation with
-\begin{align}
-  \text{cor}[v(\bs), v(\bt)] = \gamma I(\bs = \bt) + (1 - \gamma)  \frac{ 1 }{ \Gamma(\nu) 2^{ \nu - 1}} \left( \sqrt{2\nu} \frac{ h }{ \rho } \right)^{\nu} K_{\nu} \left( \sqrt{2\nu} \frac{ h }{ \rho } \right) \label{eq:matern}
-\end{align}
-where $\rho$ is the spatial range, $\nu$ is the smoothness, $\gamma$ is the proportion of variance accounted for by the spatial variation, $K_\nu$ is a modified Bessel function of the second kind, and $h = || \bs - \bt ||$.
 This process is desirable because of its flexible tail that is controlled by the skewness parameter $\lambda$ and degrees of freedom $2a$.
@@
 
-\subsection{Extremal dependence}
+Although any positive definite correlation function could be used, we choose to use the stationary isotropic \Matern correlation with
+\begin{align}
+  \text{cor}[v(\bs_1), v(\bs_2)] = \gamma I(\bs_1 = \bs_2) + (1 - \gamma)  \frac{ 1 }{ \Gamma(\nu) 2^{ \nu - 1}} \left( \sqrt{2\nu} \frac{ h }{ \rho } \right)^{\nu} K_{\nu} \left( \sqrt{2\nu} \frac{ h }{ \rho } \right) \label{eq:matern}
+\end{align}
+where $\rho$ is the spatial range, $\nu$ is the smoothness, $\gamma$ is the proportion of variance accounted for by the spatial variation, $K_\nu$ is a modified Bessel function of the second kind, and $h = || \bs_1 - \bs_2 ||$.
+
+\subsection{Extremal dependence}\label{s:extdep}
 Our interest lies in spatial dependence in the tail of the skew-$t$ process.
@@
 Therefore, this long-range dependence feature of the skew-$t$ process is not ideal for spatial analysis of large geographic regions where we expect only local spatial dependence.
+We propose a solution to this in Section \ref{s:part}.
 
-\section{Spatiotemporal skew-$t$ model for extremes}\label{s:spatial}
+\section{Spatiotemporal skew-$t$ model for threshold exceedances}\label{s:spatial}
 In this section, we propose extensions to the skew-$t$ process to model spatial extremes over a large geographic region by introducing censoring to focus on tail behavior and a random partition to remove long-range asymptotic dependence.
@@
 \subsection{Censoring to focus on the tails}
-Because one of our goals is to model the dependence of the distribution in the tails of the data, we choose to censor values below a threshold.
+As mentioned previously, we propose to use a censored approach because we are interested in high values and do not want the low-to-moderate values to influence the fit of the overall model.
+The censored observations below the threshold give information on the marginal probabilities to exceed the threshold and on the dependence, but their values are not used to fit the model.
 Let
@@
 be the censored observation at site $\bs$ where $Y(\bs)$ is the uncensored observation, $\delta(\bs) = I[Y(\bs) > T]$, and $T$ is a pre-specified threshold value.
-Then, assuming the uncensored data $Y(\bs)$ are observations from a skew-$t$ process, we update values censored below the threshold using standard Bayesian missing data methods as described in Section \ref{s:comp}.
+We impute the censored values as a step in the MCMC algorithm used to fit the model described in Section \ref{s:comp}.
 
@@
 \end{align}
-Let $\bw = (w_1, w_2)$ be the location of a spatial knot.
 To model spatial variation, consider a set of spatial knots $\bw_1, \ldots, \bw_K$ from a homogeneous Poisson process with intensity $\mu$ over spatial domain $\calD \in \calR^2$.
@@
 \end{align}
-and the $z_k$ and $\sigma^2_k$ are distributed as $z_k \iid N(0, 1)$ and $\sigma^2 \iid$ IG$(a, b)$ where IG is the distribution function of an inverse gamma random variable.
+and the $z_k$ and $\sigma^2_k$ are distributed as $z_k \iid N(0, 1)$ and $\sigma^2_k \iid$ IG$(a, b)$.
 So, within each partition, $Y(\bs)$ follows the spatial skew-$t$ process defined in Section \ref{s:spatialskew}.
@@
 
-When incorporating the random partition, conditional on knots $\bw_1, \ldots, \bw_K$, the $\chi$ statistic for two sites $\bs$ and $\bt$ in partitions $k_s$ and $k_t$ respectively is
+The partitioning model remove long-range dependence. Conditional on knots $\bw_1, \ldots, \bw_K$, the $\chi$ statistic for two sites $\bs$ and $\bt$ in partitions $k_s$ and $k_t$ respectively is
 \begin{align}
@@
 Marginally, over the knots, $\chi(h) = \pi(h) \chi_{\text{skew}-t}(h)$, where $\pi(h) = \Pr(k_s = k_t)$ is the probability that two sites separated by distance $h$ are in the same partition.
-So, to show that $\lim_{h \rightarrow \infty} \chi(h) = 0$, we need only know that $\lim_{h \rightarrow \infty} \pi(h) = 0$. A proof of this is given in Appendix \ref{a:proofsamepartition}.
-
+In Appendix \ref{a:proofsamepartition}, we show that $\lim_{h \rightarrow \infty} \pi(h) = 0$, implying $\lim_{h \rightarrow \infty} \chi(h) = 0$.
 In Figure \ref{fig:chi}, we give $\chi(h)$ for $K = 1, 3, 5, 10$ partitions for a skew-$t$ distribution with $\alpha = 10$, and 3 degrees of freedom.
-To estimate $\pi(h)$, we generate 500 sites uniformly over the unit-square.
-We then randomly generate 400 different sets of partitions using $K = 3$, $5$, and $10$.
-For each set of knots, we take $\pi(h)$ to be the proportion of sites in the same partition that are separated by distance $h$.
-This plot demonstrates how partitioning helps to reduce extremal dependence as $h$ increases.
+We estimate $\pi(h)$ through simulation.
+% To estimate $\pi(h)$, we generate 500 sites uniformly over the unit-square.
+% We then randomly generate 400 different sets of partitions using $K = 3$, $5$, and $10$.
+% For each set of knots, we take $\pi(h)$ to be the proportion of sites in the same partition that are separated by distance $h$.
+% This plot demonstrates how partitioning helps to reduce extremal dependence as $h$ increases.
 
@@
   \includegraphics[width=0.5\linewidth]{plots/chi-h.pdf}
-  \caption{$\chi(h)$ for $K = 1, 3, 5$, and $10$ knots as a function of distance.}
+  \caption{$\chi$, as a function of distance, for $K = 1, 3, 5$, and $10$ knots.}
   \label{fig:chi}
@@
 Conditioned on $z_{tk}(\bs)$, $\sigma^2_{tk}(\bs)$, and $P_{tk}$, the marginal distributions are Gaussian and the joint distribution multivariate Gaussian.
-However, we do not fix the partitions, they are treated as unknown and updated in the MCMC.
+However, we do not fix the partitions, they are treated as unknown and updated in the MCMC algorithm.
 We model this with a Bayesian hierarchical model as follows.
 Let $\bw_{t1}, \ldots, \bw_{tK}$ be a set of daily spatial knots in a spatial domain of interest, $\calD$, and $P_{tk}$ as defined in (\ref{eq:subregions}).
-In practice, we fix $K$ at many different levels, and assess the impact of fit as described in \ref{s:modelselect}.
+In practice, we fix $K$ at different levels, and assess its impact on prediction as described in \ref{s:modelselect}.
 Then
@@
 \subsection{Computation}\label{s:comp}
-We use Markov chain Monte Carlo methods to explore the posterior.
+We use MCMC methods to explore the posterior.
 At each MCMC iteration, we first impute values below the threshold conditional on observations above the threshold.
@@
 
-Then, we update model parameters, $\Theta$, using Metropolis Hastings or Gibbs sampling when appropriate.
+Then, we update model parameters, $\Theta$, using a Metropolis-Hastings algorithm with Gibbs sampling when needed.
 The final step of the computation is to use Bayesian Kriging to generate a predictive distribution for $Y_t(\bs^*)$ at prediction location $\bs^*$.
 This step is similar to the imputation for censored observations except that the full conditionals are no longer truncated at $T$.
-See Appendices A.1 and A.2 for details regarding the MCMC.
+See Appendices A.1 and A.2 for details regarding the MCMC algorithm.
 
 \section{Simulation study}\label{s:simstudy}
-In this section, we conduct a simulation study to investigate how the number of partitions and the level of thresholding impact the accuracy of predictions made by the model.
+In this section, we present the results from a simulation study to investigate how the number of partitions and the level of thresholding impact the accuracy of predictions made by the model.
 
 \subsection{Design}\label{s:simdesign}
-For all simulation designs, we generate data from the model in Section \ref{s:part} using $n_s=144$ sites and $n_t=50$ independent days.
-The sites are generated Uniform$([0, 10] \times [0, 10])$.
-We generate data from 4 different simulation designs:
+For all simulation designs, we generated data from model (\ref{eq:partition})) in Section \ref{s:part} using $n_s=144$ sites and $n_t=50$ independent days.
+The sites were generated Uniform$([0, 10] \times [0, 10])$.
+We generated data from 4 different simulation designs:
 \begin{enumerate} \setlength{\itemsep}{-0.5em}
@@
 \end{enumerate}
-In the first three designs, the $v_t(\bs)$ terms are generated using a \Matern covariance with smoothness parameter $\nu = 0.5$ and spatial range $\rho = 1$.
-For the covariance matrices in designs 1 -- 3, the proportion of the variance accounted for by the spatial variation is $\gamma = 0.9$ while the proportion of the variance accounted for by the nugget effect is $0.1$.
-In the first design, $\sigma^2 = 2$ is used for all days which results in a Gaussian distribution.
+In the first three designs, the $v_t(\bs)$ terms were generated using a \Matern covariance with smoothness parameter $\nu = 0.5$ and spatial range $\rho = 1$.
+For the covariance matrices in designs 1 -- 3, the proportion of the variance accounted for by the spatial variation was $\gamma = 0.9$ while the proportion of the variance accounted for by the nugget effect was $0.1$.
+In the first design, $\sigma^2 = 2$ was used for all days which results in a Gaussian distribution.
 For designs 2 and 3, $\sigma^2_{tk} \iid \text{IG}(3, 8)$ to give a $t$ distribution with 6 degrees of freedom.
 For design 1, we set $\lambda = 0$.
-For designs 2 and 3, $\lambda = 3$ was used as to simulate moderate skewness, and the $z_t$ are generated as described in (\ref{eq:sitezsig}).
-In designs 1 -- 3, the mean $\bX^T \bbeta = 10$ is assumed to be constant across space.
-In the fourth design, we generate from a spatial max-stable distribution \citep{Reich2012}.
-In this design, data have marginal distributions that follow a generalized extreme value distribution with parameters $\mu = 1, \sigma=1, \xi=0.2$.
-In this model, a random effect is used to induce spatial dependence using 144 spatial knots on a regular lattice in the square $[1, 9] \times [1, 9]$.
+For designs 2 and 3, $\lambda = 3$ was used as to simulate moderate skewness, and the $z_t$ were generated as described in (\ref{eq:sitezsig}).
+In designs 1 -- 3, the mean $\bX^T \bbeta = 10$ was assumed to be constant across space.
+In the fourth design, we generated from a spatial max-stable distribution \citep{Reich2012}.
+In this design, data have marginal distributions that follow a generalized extreme value distribution with location parameter $1$, scale parameter $1$, and shape parameter $0.2$.
+In this model, a random effect was used to induce spatial dependence using 144 spatial knots on a regular lattice in the square $[1, 9] \times [1, 9]$.
 For this setting, we set $\gamma = 0.5$.
@@
 The skewness parameter has prior $\lambda_2 \sim \text{IG}(0.1, 0.1)$.
-The residual variance terms have priors $\sigma^2_t(\bs) \sim \text{IG}(0.1, 0.1)$.
+The residual variance terms have priors $\sigma^2_t(\bs) \sim \text{IG}(a, b)$, where $a$ has a Gamma$(0.1, 0.1)$ prior and $b$ has a discrete uniform prior on a mesh from $0.1$ to $10$ with spacing of $0.1$.
 The knots have priors $\bw \sim \text{Unif}(\calD)$.
@@
 Each chain of the MCMC ran for 20,000 iterations with a burn-in period of 10,000 iterations.
-Parameters appear to converge properly; however, in the models with multiple partitions (i.e.models 4 and 5) it is hard to assess the convergence of $\bw$, $z(\bs)$, and $\sigma^2(\bs)$ because of partition label switching throughout the MCMC.
+Parameters appear to converge properly; however, in the models with multiple partitions (i.e.
+models 4 and 5) it is hard to assess the convergence of $\bw$, $z(\bs)$, and $\sigma^2(\bs)$ because of partition label switching throughout the MCMC.
 
 \subsection{Cross validation}\label{s:modelselect}
-Models were compared using cross validation with 100 sites used as training sites and 44 sites withheld for testing.
-The model was fit using the training set, and predictions were generated at the testing site locations.
-Because one of the primary goals of this model is to predict exceedances over a fixed threshold, we use Brier scores to select the model that best fits the data \citep{Gneiting2007}.
+Models were compared using cross validation, with 100 sites used as training sites to fit the models, and 44 sites withheld for testing the predictions.
+Because one of the primary goals of this model is to predict exceedances over a fixed threshold, we use Brier scores to compare the models \citep{Gneiting2007}.
 The Brier score for predicting exceedance of a threshold $c$ is given by $[e(c) - P(c)]^2$ where $e(c) = I[y>c]$ is an indicator function indicating that a test set value, $y$, has exceeded the threshold, $c$, and $P(c)$ is the predicted probability of exceeding $c$.
@@
 
-A plot of the relative Brier scores is given in Figure \ref{fig:simbrierscores}.
-We find that when the data are generated from a Gaussian process, our method performs comparably to a Gaussian approach.
+The results show that when the data are generated from a Gaussian process, our method performs comparably to a Gaussian approach.
 In general, when the underlying process is not Gaussian, our method results in an improvement over both the max-stable and Gaussian methods.
@@
 \section{Data analysis}\label{s:analysis}
-To illustrate this method, we consider 31 daily observations of maximum 8-hour ozone measurements for July 2005 at 1089 Air Quality System (AQS) monitoring sites in the United States as the response (see Figure \ref{fig:ozone}).
+We consider daily observations of maximum 8-hour ozone measurements for the 31 days of July 2005 at 1,089 Air Quality System (AQS) monitoring sites in the United States as the response (see Figure \ref{fig:ozone-10jul}).
 For each site, we also have covariate information containing the estimated ozone from the Community Multi-scale Air Quality (CMAQ) modeling system.
@@
 \end{align}
-The data from July 10 are shown in Figure \ref{fig:ozone} along with a Q-Q plot of the residuals compared to a skew-$t$ distribution with 10 d.f. and $\alpha = 1$.
+Figure \ref{fig:ozone-qq} shows a Q-Q plot of the residuals compared to a skew-$t$ distribution with 10 d.f. and $\alpha = 1$.
 \begin{figure}
   \centering
-  \includegraphics[width=0.56\linewidth]{plots/ozone-10jul-us.pdf}
-  \includegraphics[width=0.41\linewidth]{plots/qq-res.pdf}
-  \caption{Ozone values on 10 July 2005 (left) Q-Q plot of the residuals for a skew-$t$ distribution with 10 d.f. and $\alpha = 1$ (right)}
-  \label{fig:ozone}
+  \includegraphics[width=\linewidth]{plots/qq-res.pdf}
+  \caption{Q-Q plot of the residuals for a skew-$t$ distribution with 10 d.f. and $\alpha = 1$ (right)}
+  \label{fig:ozone-qq}
 \end{figure}
@@
   \includegraphics[width=\linewidth]{plots/daily-quantiles-ozone.pdf}
-  \caption{Daily quantiles for sites that are close (left) and far apart (right)}
+  \caption{Daily quantiles for two monitoring locations near Columbus, OH (left) and daily quantiles for a monitoring location in Los Angeles, CA and Columbus, OH (right)}
   \label{fig:bivariateozone}
@@
 Another aspect that distinguishes our approach from more traditional extremes analyses, is how the threshold is selected.
-Traditionally, a threshold is chosen based on the assumption that beyond the threshold, the data follow an extremal model.
-In our example, a threshold of 75 ppb which is $q(0.92)$ for all observations, but marginally it represents anywhere from $q(0.06)$ to $q(1)$.
+In our example, a threshold of 75 ppb which corresponds to $q(92)$ for all observations, but marginally it represents anywhere from $q(0.06)$ to $q(1)$.
 
@@
 We fit the model using Gaussian and skew-$t$ marginal distributions with $K=1, 5, 6, 7, 8, 9, 10, 15$ partitions.
-We choose to censor $Y(\bs)$ at $T =$ 0, 50 (0.42 sample quantile), and 75 (0.92 sample quantile) ppb in order to compare results from no, moderate, and high censoring.
+We choose to censor $Y(\bs)$ at $T = 0$, $T = 50$ (0.42 sample quantile), and $T = 75$ (0.92 sample quantile) ppb in order to compare results from no, moderate, and high censoring.
 The upper threshold of 75 ppb was used because the current air quality standard is based on exceedance of 75 ppb.
@@
 All methods assume the mean function given in (\ref{eq:datamean}).
-To ensure that the max-stable method runs in a reasonable amount of time, we take a stratified sample of 800 sites and consider this our new dataset.
+To ensure that the max-stable method runs in a reasonable amount of time, we use a stratified sub-sample of 800 sites.
 We conduct two-fold cross validation using 400 training sites and 400 validation sites as described in Section \ref{s:modelselect}
@@
 \section*{Acknowledgments}
+The authors' work was partially supported by grants from the Department of the Interior (14-1-04-9), National Institutes of Health (R21ES022795-01A1), the US Environmental Protection Agency (R835228), the National Science Foundation (1107046), and the Research Network for Statistical Methods for Atmospheric and Oceanic Sciences (STATMOS).
+The authors also received high-performance computing support from Yellowstone (ark:/85065/d7wd3xhc) provided by NCAR's Computational and Information Systems Laboratory, sponsored by the National Science Foundation.
 
@@
 Whenever possible, we select conjugate priors (see Appendix \ref{a:posterior}); however, for some of the parameters, no conjugate prior distributions exist.
-When no conjugate prior distribution exists, we use a random walk Metropolis Hastings update step.
-In each Metropolis Hastings update, we tune the algorithm to give acceptance rates near 0.40.
+When no conjugate prior distribution exists, we use a random walk Metropolis-Hastings update step.
+In each Metropolis-Hastings update, we tune the algorithm to give acceptance rates near 0.40.
 
 \subsubsection*{Spatial knot locations}
-For each day, we update the spatial knot locations, $\bw_1, \ldots, \bw_K$, using a Metropolis Hastings block update.
+For each day, we update the spatial knot locations, $\bw_1, \ldots, \bw_K$, using a Metropolis-Hastings block update.
 Because the spatial domain is bounded, we generate candidate knots using the transformed knots $\bw^*_1, \ldots, \bw^*_K$ (see section \ref{s:temporal}) and a random walk bivariate Gaussian candidate distribution
@@
 If there is no temporal dependence amongst the observations, we use a Gibbs update for $z_{tk}$, and the posterior distribution is given in \ref{a:posterior}.
-If there is temporal dependence amongst the observations, then we update $z_{tk}$ using a Metropolis Hastings update.
+If there is temporal dependence amongst the observations, then we update $z_{tk}$ using a Metropolis-Hastings update.
 Because this model uses $|z_{tk}|$, we generate candidate random effects using the $z^*_{tk}$ (see Section \ref{s:temporal}) and a random walk Gaussian candidate distribution
@@
 \subsubsection*{Variance terms}
-When there is more than one site in a partition, then we update $\sigma^2_{tk}$ using a Metropolis Hastings update.
-First, we generate a candidate for $\sigma^2_{tk}$ using an IG$(a^*/s, b^*/s)$ candidate distribution in an independence Metropolis Hastings update where $a^* = (n_{tk} + 1) / 2 + a$, $b^* = [Y_{tk}^T \Sigma^{-1}_{tk} Y_{tk} + z_{tk}^2] / 2 + b$, $n_{tk}$ is the number of sites in partition $k$ on day $t$, and $Y_{tk}$ and $\Sigma^{-1}_{tk}$ are the observations and precision matrix for partition $k$ on day $t$.
+When there is more than one site in a partition, then we update $\sigma^2_{tk}$ using a Metropolis-Hastings update.
+First, we generate a candidate for $\sigma^2_{tk}$ using an IG$(a^*/s, b^*/s)$ candidate distribution in an independence Metropolis-Hastings update where $a^* = (n_{tk} + 1) / 2 + a$, $b^* = [Y_{tk}^T \Sigma^{-1}_{tk} Y_{tk} + z_{tk}^2] / 2 + b$, $n_{tk}$ is the number of sites in partition $k$ on day $t$, and $Y_{tk}$ and $\Sigma^{-1}_{tk}$ are the observations and precision matrix for partition $k$ on day $t$.
 The acceptance ratio is
@@
 \subsubsection*{Spatial covariance parameters}
-We update the three spatial covariance parameters, $\log(\rho)$, $\log(\nu)$, $\gamma$, using a Metropolis Hastings block update step.
+We update the three spatial covariance parameters, $\log(\rho)$, $\log(\nu)$, $\gamma$, using a Metropolis-Hastings block update step.
 First, we generate a candidate using a random walk Gaussian candidate distribution

### 2015-08-06 code/analysis/ozone/US-all/us-all-setup.R

@@ -10,6 +10,10 @@ source('../../../R/auxfunctions.R')
 load("../ozone_data.RData")
 S <- cbind(x[s[, 1]], y[s[, 2]])  # expands the grid of x, y locs where we have CMAQ
 
+# S[582, ] is near Lincoln, NE  (40.8106N, 96.6803W)
+# S[1069, ] is near Topeka, KS  (39.0558N, 95.6894W)
+# distance from rdist is around 216 and actual distance between cities is 212.6km
+
 # Exclude if site is missing more than 50% of its days
 excl   <- which(rowMeans(is.na(Y)) > 0.50)
 index  <- index[-excl]

### 2015-09-02 LaTeX/skewnormal.tex
few minor edits, almost ready to submit.

@@
 \section{Introduction}\label{s:intro}
-In many climatological applications, researchers are interested in learning about the average behavior of different climate variables (e.g. ozone, temperature, rainfall).
+% In many climatological applications, researchers are interested in learning about the average behavior of different climate variables (e.g. ozone, temperature, rainfall).
+Epidemiological studies have linked air quality to public health concerns regarding morbidity and mortality \citep{Samet2000}.
+As a results, the Environmental Protection Agency (EPA) has developed a set of standards to help reduce air pollution thereby improving air quality.
 Our study is motivated by an air pollution application where the focus is not on the average behavior, but instead the behavior over a fixed threshold determined by government regulation.
@@
 where $\rho$ is the spatial range, $\nu$ is the smoothness, $\gamma$ is the proportion of variance accounted for by the spatial variation, $K_\nu$ is a modified Bessel function of the second kind, and $h = || \bs_1 - \bs_2 ||$.
+We use this parameterization for spatial correlation because the $\gamma$ parameter permits the inclusion of a nugget effect to account for non-spatial variability due to issues like measurement error.
 
@@
 In this model, a random effect was used to induce spatial dependence using 144 spatial knots on a regular lattice in the square $[1, 9] \times [1, 9]$.
-For this setting, we set $\gamma = 0.5$.
+For this setting, $\gamma_{\text{RS}} \in (0, 1)$, is similar to $\gamma$ in that it controls the relative strength of the nugget effect.
+We set $\gamma_{\text{RS}} = 0.5$, which represents moderate spatial dependence.
 % In the final design, we generate $\tilde{y}$ using the setting from design 2, and then transform the data
@@
   \centering
-  \includegraphics[width=\linewidth]{plots/qq-res.pdf}
+  \includegraphics[width=0.5\linewidth]{plots/qq-res.pdf}
   \caption{Q-Q plot of the residuals for a skew-$t$ distribution with 10 d.f. and $\alpha = 1$ (right)}
@@
 Whenever possible, we select conjugate priors (see Appendix \ref{a:posterior}); however, for some of the parameters, no conjugate prior distributions exist.
-When no conjugate prior distribution exists, we use a random walk Metropolis-Hastings update step.
-In each Metropolis-Hastings update, we tune the algorithm to give acceptance rates near 0.40.
+For these parameters, we use a random walk Metropolis-Hastings update step.
+In each Metropolis-Hastings update, we tune the algorithm during the burn-in period to give acceptance rates near 0.40.
 
@@
 \subsection{Posterior distributions} \label{a:posterior}
-\input{zpost.tex}
-\input{betapost.tex}
-\input{sigpost.tex}
-\input{lambdapost.tex}
+
+\subsubsection*{Conditional posterior of $z_{tk} \mid \ldots $}\label{s:mvcondu}
+If knots are independent over days, then the conditional posterior distribution of $|z_{tk}|$ is conjugate.
+For simplicity, drop the subscript $t$, let $\tilde{z}_{tk} = |z_{tk}|$, and define
+\begin{align*}
+R(\bs) = \left\{
+    \begin{array}{ll}
+        Y(\bs) - X(\bs) \beta &s \in P_l\\[1em]
+        Y(\bs) - X(\bs) \beta - \lambda \tilde{z}(\bs) \qquad & s \notin P_l
+    \end{array}
+\right.
+\end{align*}
+Let
+\begin{align*}
+    R_1 &= \text{the vector of } R(\bs) \text{ for } s \in P_l \\
+    R_2 &= \text{the vector of } R(\bs) \text{ for } s \notin P_l \\
+    \Omega &= \Sigma^{-1}.
+\end{align*}
+Then
+\begin{align*}
+    \pi(z_l | \ldots) &\propto \exp \left\{ -\frac{ 1 }{ 2 } \left[
+        \left( \begin{array}{c}
+            R_1 - \lambda \tilde{z}_l \bOne\\
+            R_2
+        \end{array} \right)^T
+        \left( \begin{array}{cc}
+            \Omega_{11} & \Omega_{12}\\
+            \Omega_{21} & \Omega_{22}
+        \end{array} \right)
+        \left( \begin{array}{c}
+            R_1 - \lambda \tilde{z}_l \bOne\\
+            R_2
+        \end{array} \right)
+        +  \frac{ {\tilde{z}_l}^2 }{ \sigma_l^2 }\right]
+    \right\} I(z_l > 0) \\
+        &\propto \exp \left\{ -\frac{ 1 }{ 2 } \left[ \Lambda_l {\tilde{z}_l}^2 - 2 \mu_l \tilde{z}_l \right] \right\}
+\end{align*}
+where
+\begin{align*}
+    \mu_l &= \lambda ( R_1^T \Omega_{11} + R_2^T \Omega_{21} )\bOne\\
+    \Lambda_l &= \lambda^2 \bOne^T \Omega_{11} \bOne + \frac{ 1 }{ \sigma^2_l }.
+\end{align*}
+Then $\tilde{Z}_l | \ldots \sim N_{(0, \infty)} (\Lambda_l^{-1} \mu_l, \Lambda_l^{-1})$
+
+\subsection*{Conditional posterior of $\beta \mid \ldots$}\label{s:betapost}
+Let $\beta \sim \mbox{N}_{p}(0, \Lambda_0)$ where $\Lambda_0$ is a precision matrix. Then
+\begin{align*}
+    \pi(\beta \mid \ldots) & \propto \exp \left\{ - \frac{ 1 }{ 2 } \beta^T \Lambda_0 \beta - \frac{ 1 }{ 2 } \sum_{t = 1 }^T [\bY_t - X_t \beta - \lambda |z_t|]^T \Omega [\bY_t - X_t \beta - \lambda |z_t|] \right\}\\
+     & \propto \exp \left\{ -\frac{ 1 }{ 2 } \left[ \beta^T \Lambda_\beta \beta  - 2 \sum_{ t = 1 }^T [ \beta^T X_t^T \Omega (\bY_t - \lambda |z_t| )] \right] \right\}\\
+     & \propto \mbox{N} ( \Lambda_\beta^{-1} \mu_\beta , \Lambda_\beta^{-1})
+\end{align*}
+where
+\begin{align*}
+    \mu_\beta &= \sum_{t=1}^{T} \left[ X_t^T \Omega (\bY_t - \lambda |z_t|) \right]\\
+    \Lambda_\beta &= \Lambda_0 + \sum_{ t = 1 }^{ T} X_t^T \Omega X_t.
+\end{align*}
+
+\subsection*{Conditional posterior of $\sigma^2 \mid \ldots$}\label{s:sigpost}
+In the case where $L = 1$ and temporal dependence is negligible, then $\sigma^2$ has a conjugate posterior distribution.
+Let $\sigma_t^2 \iid \mbox{IG}(\alpha_0, \beta_0)$. For simplicity, drop the subscript $t$. Then
+\begin{align*}
+    \pi(\sigma^2 \mid \ldots) & \propto (\sigma^2)^{ -\alpha_0 - 1 / 2 - n / 2 - 1} \exp \left\{ -\frac{\beta_0}{\sigma^2} - \frac{ |z|^2 }{2 \sigma^2} - \frac{ (\bY - \bmu)^T \Sigma^{-1} (\bY - \bmu) }{2 \sigma^2} \right\} \\
+    & \propto (\sigma^2)^{ -\alpha_0 - 1 / 2 - n / 2 - 1} \exp \left\{ - \frac{ 1 }{ \sigma^2 } \left[\beta_0 + \frac{ |z|^2 }{ 2 } + \frac{ 1 }{ 2 } (\bY - \bmu)^T \Sigma^{-1} (\bY - \mu) \right] \right\} \\
+    & \propto \mbox{IG} (\alpha^*, \beta^*)
+\end{align*}
+where
+\begin{align*}
+    \alpha^* &= \alpha_0 + \frac{1}{2} + \frac{n}{2} \\
+    \beta^* &= \beta_0 + \frac{ |z|^2 }{ 2 } + \frac{ 1 }{ 2 }(\bY - \bmu)^T \Sigma^{-1} (\bY - \bmu).
+\end{align*}
+In the case that $L > 1$, a random walk Metropolis Hastings step will be used to update $\sigma^2_{lt}$.
+
+\subsection*{Conditional posterior of $\lambda \mid \ldots$}\label{s:lambdapost}
+For convergence purposes we model $\lambda = \lambda_1 \lambda_2$ where
+\begin{align}
+  \lambda_1 &= \left\{ \begin{array}{ll}
+      +1 \quad & \text{w.p.} 0.5\\
+      -1 \quad & \text{w.p.} 0.5
+   \end{array}\right.\\
+   \lambda^2_2 & \sim IG(\alpha_\lambda, \beta_\lambda).\\
+\end{align}
+Then
+\begin{align*}
+  \pi(\lambda_2 \mid \ldots) &\propto {\lambda_2^2}^{(-\alpha_\lambda - 1)} \exp\left\{ - \frac{ \beta_\lambda }{ \lambda_2^2} \right\} \prod_{t = 1}^{T} \prod_{k = 1}^K \frac{1} {\lambda_2} \exp\left\{ - \frac{ z^2_{tk} }{ 2 \lambda_2^2  \sigma_{tk})^2} \right\} \\
+   & \propto {\lambda_2^2}^{(-\alpha_\lambda - kt - 1)} \exp \left\{ -\frac{1}{\lambda^2_2} \left[ \beta_\lambda + \frac{ z^2 }{2 \sigma^2_{tk}} \right] \right\}
+\end{align*}
+Then $\lambda_2 \mid \ldots \sim IG \left(\alpha_\lambda + kt, \beta_\lambda + \frac{ z^2 }{2 \sigma^2_{tk}} \right)$
 
@@
 \subsection{Skew-$t$ distribution} \label{a:skewt}
-\subsubsection*{Univariate extended skew-$t$ distribution}
-We say that $Y$ follow a univariate extended skew-$t$ distribution with location $\xi \in \calR$, scale $\omega > 0$, skew parameter $\alpha \in \calR$, extended parameter $\tau \in \calR$, and degrees of freedom $\nu$ if has distribution function
+\subsubsection*{Univariate skew-$t$ distribution}
+We say that $Y$ follows a univariate extended skew-$t$ distribution with location $\xi \in \calR$, scale $\omega > 0$, skew parameter $\alpha \in \calR$, and degrees of freedom $\nu$ if has distribution function
 \begin{align}
-  f_{\text{EST}}(y) = \omega^{-1} \frac{ f_T (z; \nu) }{ F_T(\tau / \sqrt{ 1 + \alpha^2}; \nu)} F_T\left[ (\alpha z + \tau) \sqrt{ \frac{ \nu + 1 }{ \nu + z^2}}; 0, 1, \nu + 1 \right]
+  f_{\text{EST}}(y) = 2 f_T (z; \nu) F_T\left[ \alpha z \sqrt{ \frac{ \nu + 1 }{ \nu + z^2}}; \nu + 1 \right]
 \end{align}
 where $f_T(t; \nu)$ is a univariate Student's $t$ with $\nu$ degrees of freedom, $F_T(t; \nu) = P(T < t)$, and \hbox{$z = (y - \xi) / \omega$}.
-In the case that $\tau = 0$, then $Y$ follows a univariate skew-$t$ distribution.
 
@@
 \end{align}
-Because the extended skew-$t$ distribution is not bounded above, for all $\bar{F}_{\text{EST}}(x) = 1 - F_{\text{EST}} > 0$ for all $x < \infty$.
+Because the extended skew-$t$ distribution is not bounded above, for all $\bar{F}_{\text{EST}}(x) = 1 - F_{\text{EST} (x)} > 0$ for all $x < \infty$.
 Therefore, for a skew-$t$ distribution, $\lim_{h \rightarrow \infty} \chi(h) > 0$.

### 2015-09-09 LaTeX/skewnormal.tex
commenting out figure 4
Reduce file size for qq plot and minor updates to manuscript from Brian

@@
 Epidemiological studies have linked air quality to public health concerns regarding morbidity and mortality \citep{Samet2000}.
-As a results, the Environmental Protection Agency (EPA) has developed a set of standards to help reduce air pollution thereby improving air quality.
+As a result, the Environmental Protection Agency (EPA) has developed a set of standards to help reduce air pollution thereby improving air quality.
 Our study is motivated by an air pollution application where the focus is not on the average behavior, but instead the behavior over a fixed threshold determined by government regulation.
@@
   \centering
-  \includegraphics[width=\linewidth]{plots/ozone-10jul-us.pdf}
+  \includegraphics[width=0.75\linewidth]{plots/ozone-10jul-us.pdf}
   \caption{Ozone values on July 10, 2005}
@@
 Third, we aim to capture the dependence structure when ozone is at high levels, and dependence at these levels may not be well-represented by covariances which focus again on mean behavior.
-Asymptotic dependence/independence (see Section \ref{s:extdep}) are notions which describe how two random variables' dependence behaves as one looks further into the joint tail.
+Asymptotic dependence/independence (see Section \ref{s:extdep}) are notions which describe how two random variables' probability of simultaneous exceedance of an extreme threshold behaves as one increases the threshold.
 The Gaussian distribution always exhibits asymptotic independence, except in the case of perfect dependence, thus is an inappropriate model for data which exhibits asymptotic dependence.

### 2015-09-09 LaTeX/skewnormal.tex
move appendices into supplemental document

--- 2015-08-06 LaTeX/skewnormal.tex
@@
 \usepackage{setspace}
+
+% cross-referencing appendices
+\usepackage{xr}
+\externaldocument{supplemental}
 
@@
 \section{Introduction}\label{s:intro}
-In many climatological applications, researchers are interested in learning about the average behavior of different climate variables (e.g. ozone, temperature, rainfall).
+% In many climatological applications, researchers are interested in learning about the average behavior of different climate variables (e.g. ozone, temperature, rainfall).
+Epidemiological studies have linked air quality to public health concerns regarding morbidity and mortality \citep{Samet2000}.
+As a result, the Environmental Protection Agency (EPA) has developed a set of standards to help reduce air pollution thereby improving air quality.
 Our study is motivated by an air pollution application where the focus is not on the average behavior, but instead the behavior over a fixed threshold determined by government regulation.
@@
 We see a large area above the compliance level in the midwest covering Ohio, Indiana, Illinois, and parts of the surrounding states.
-\begin{figure}
+
+\begin{figure}[h!tbp]
   \centering
-  \includegraphics[width=\linewidth]{plots/ozone-10jul-us.pdf}
-  \caption{Ozone values on July 10, 2005}
+  \includegraphics[width=0.75\linewidth]{plots/ozone-10jul-us.pdf}
+  \caption{Ozone values (ppb) on July 10, 2005}
   \label{fig:ozone-10jul}
@@
 In order to address this concern, we introduce a random spatial partition similar to the method used by \citet{Kim2005} for non-stationary Gaussian data.
-This partition alleviates the asymptotic spatial dependence present in the skew-$t$ distribution for sites that are far apart.
+This partitioning alleviates the asymptotic spatial dependence present in the skew-$t$ distribution for sites that are far apart.
 
@@
 \section{Spatial skew processes}\label{s:spatialskew}
-Many types of data demonstrate some level of skewness and therefore should be modeled with distributions that allow for asymmetry.
 The skew-elliptical family of distributions provides models that are mathematically tractable while introducing a slant parameter to account for asymmetric data \citep{Azzalini2014}.
@@
 As mentioned previously, we propose to use a censored approach because we are interested in high values and do not want the low-to-moderate values to influence the fit of the overall model.
-The censored observations below the threshold give information on the marginal probabilities to exceed the threshold and on the dependence, but their values are not used to fit the model.
+% The censored observations below the threshold give information on the marginal probabilities to exceed the threshold and on the dependence, but their values are not used to fit the model.
 Let
@@
 In Figure \ref{fig:chi}, we give $\chi(h)$ for $K = 1, 3, 5, 10$ partitions for a skew-$t$ distribution with $\alpha = 10$, and 3 degrees of freedom.
-We estimate $\pi(h)$ through simulation.
+In Figure \ref{fig:chi} we estimate $\pi(h)$ through simulation.
 % To estimate $\pi(h)$, we generate 500 sites uniformly over the unit-square.
@@
   \includegraphics[width=0.5\linewidth]{plots/chi-h.pdf}
-  \caption{$\chi$, as a function of distance, for $K = 1, 3, 5$, and $10$ knots.}
+  \caption{Extremal dependence measure $\chi(h)$, as a function of distance, $h$, for $K = 1, 3, 5$, and $10$ knots.}
   \label{fig:chi}
@@
 \end{align}
-where $\Phi$ is a univariate standard normal density function, and $\bs_i = [s_{1i}, \ldots, s_{ni}]$.
+where $\Phi$ is a univariate standard normal density function and $\bs_i = [s_{1i}, \ldots, s_{ni}]$.
 Then the transformed knots $\bw^*_{tk} \in \calR^2$.
@@
 \section{Simulation study}\label{s:simstudy}
-In this section, we present the results from a simulation study to investigate how the number of partitions and the level of thresholding impact the accuracy of predictions made by the model.
+In this section, we present the results from a simulation study to investigate how the number of partitions and the level of thresholding impact the accuracy of predictions made by the model and to compare with Gaussian and max-stable methods.
 
 \subsection{Design}\label{s:simdesign}
-For all simulation designs, we generated data from model (\ref{eq:partition})) in Section \ref{s:part} using $n_s=144$ sites and $n_t=50$ independent days.
+For all simulation designs, we generated data from model (\ref{eq:partition}) in Section \ref{s:part} using $n_s=144$ sites and $n_t=50$ independent days.
 The sites were generated Uniform$([0, 10] \times [0, 10])$.
@@
 
-\appendix
-\section{Appendices}
-\subsection{MCMC details} \label{a:mcmc}
-The MCMC sampling for the model \ref{s:hier} is done using {\tt R} (http://www.r-project.org).
-Whenever possible, we select conjugate priors (see Appendix \ref{a:posterior}); however, for some of the parameters, no conjugate prior distributions exist.
-When no conjugate prior distribution exists, we use a random walk Metropolis-Hastings update step.
-In each Metropolis-Hastings update, we tune the algorithm to give acceptance rates near 0.40.
-
-\subsubsection*{Spatial knot locations}
-For each day, we update the spatial knot locations, $\bw_1, \ldots, \bw_K$, using a Metropolis-Hastings block update.
-Because the spatial domain is bounded, we generate candidate knots using the transformed knots $\bw^*_1, \ldots, \bw^*_K$ (see section \ref{s:temporal}) and a random walk bivariate Gaussian candidate distribution
-\begin{align*}
-	{\bw^*_k}^{(c)} \sim \text{N}({\bw^*_k}^{(r - 1)}, s^2 I_2)
-\end{align*}
-where ${\bw^*_k}^{(r - 1)}$ is the location for the transformed knot at MCMC iteration $r - 1$, $s$ is a tuning parameter, and $I_2$ is an identity matrix.
-After candidates have been generated for all $K$ knots, the acceptance ratio is
-\begin{align*}
-  R = \left\{ \frac{ l[ Y_t(\bs | \bw_1^{(c)}, \ldots, \bw_K^{(c)}, \ldots)] }{l[ Y_t(\bs | \bw_1^{(r - 1)}, \ldots, \bw_K^{(r - 1)}, \ldots)]} \right\} \times \left\{ \frac{ \prod_{k = 1}^{K}\phi(\bw_k^{(c)})}{ \prod_{k = 1}^{K}\phi(\bw_k^{(r - 1)})} \right\} \times \left\{ \frac{ \prod_{k = 1}^{K} p({\bw^*_k}^{(c)})}{ \prod_{k = 1}^{K} p({\bw^*_k}^{(r - 1)})}\right\}
-\end{align*}
-where $l$ is the likelihood given in (\ref{eq:hier}), and $p(\cdot)$ is the prior either taken from the time series given in (\ref{s:temporal}) or assumed to be uniform over $\calD$.
-The candidate knots are accepted with probability $\min\{R, 1\}$.
-
-\subsubsection*{Spatial random effects}
-If there is no temporal dependence amongst the observations, we use a Gibbs update for $z_{tk}$, and the posterior distribution is given in \ref{a:posterior}.
-If there is temporal dependence amongst the observations, then we update $z_{tk}$ using a Metropolis-Hastings update.
-Because this model uses $|z_{tk}|$, we generate candidate random effects using the $z^*_{tk}$ (see Section \ref{s:temporal}) and a random walk Gaussian candidate distribution
-\begin{align*}
-  {z^*_{tk}}^{(c)} \sim \text{N}({z^*_{tk}}^{(r - 1)}, s^2)
-\end{align*}
-where ${z^*_{tk}}^{(r-1)}$ is the value at MCMC iteration $r - 1$, and $s$ is a tuning parameter.
-The acceptance ratio is
-\begin{align*}
-  R = \left\{ \frac{ l[Y_t(\bs) | z_{tk}^{(c)}, \ldots] }{ l[Y_t(\bs) | z_{tk}^{(r - 1)}]} \right\} \times \left\{ \frac{ p[ z_{tk}^{(c)} ] }{ p[ z_{tk}^{(r - 1)}]}\right\}
-\end{align*}
-where $p[\cdot]$ is the prior taken from the time series given in Section \ref{s:temporal}.
-The candidate is accepted with probability $\min\{R, 1\}$.
-
-\subsubsection*{Variance terms}
-When there is more than one site in a partition, then we update $\sigma^2_{tk}$ using a Metropolis-Hastings update.
-First, we generate a candidate for $\sigma^2_{tk}$ using an IG$(a^*/s, b^*/s)$ candidate distribution in an independence Metropolis-Hastings update where $a^* = (n_{tk} + 1) / 2 + a$, $b^* = [Y_{tk}^T \Sigma^{-1}_{tk} Y_{tk} + z_{tk}^2] / 2 + b$, $n_{tk}$ is the number of sites in partition $k$ on day $t$, and $Y_{tk}$ and $\Sigma^{-1}_{tk}$ are the observations and precision matrix for partition $k$ on day $t$.
-The acceptance ratio is
-\begin{align*}
-  R = \left\{
-    \frac{ l[Y_t(\bs) | {\sigma^2_{tk}}^{(c)}, \ldots] }
-         { l[Y_t(\bs) | {\sigma^2_{tk}}^{(r - 1)}]}
-    \right\} \times \left\{
-    \frac{ l[z_{tk} | {\sigma^2_{tk}}^{(c)}, \ldots] }
-         { l[z_{tk} | {\sigma^2_{tk}}^{(r - 1)}, \ldots] }
-    \right\} \times \left\{
-    \frac{ p[ {\sigma^2_{tk}}^{(c)} ] }
-         { p[ {\sigma^2_{tk}}^{(r - 1)}] }
-    \right\} \times \left\{
-    \frac{ c[ {\sigma^2_{tk}}^{(r - 1)}] }
-         { c[ {\sigma^2_{tk}}^{(c)}]}
-    \right\}
-\end{align*}
-where $p[\cdot]$ is the prior either taken from the time series given in Section \ref{s:temporal} or assumed to be IG$(a, b)$, and $c[\cdot]$ is the candidate distribution.
-The candidate is accepted with probability $\min\{R, 1\}$.
-
-\subsubsection*{Spatial covariance parameters}
-We update the three spatial covariance parameters, $\log(\rho)$, $\log(\nu)$, $\gamma$, using a Metropolis-Hastings block update step.
-First, we generate a candidate using a random walk Gaussian candidate distribution
-\begin{align*}
-	\log(\rho)^{(c)} \sim \text{N}(\log(\rho)^{(r - 1)}, s^2)
-\end{align*}
-where $\log(\rho)^{(r-1)}$ is the value at MCMC iteration $r - 1$, and $s$ is a tuning parameter.
-Candidates are generated for $\log(\nu)$ and $\gamma$ in a similar fashion.
-The acceptance ratio is
-\begin{align*}
-	R = \left\{ \frac{ \prod_{t = 1}^{T} l[Y_t(\bs) | \rho^{(c)}, \nu^{(c)}, \gamma^{(c)}, \ldots] }{\prod_{t = 1}^{T} l[Y_t(\bs) | \rho^{(r-1)}, \nu^{(r-1)}, \gamma^{(r-1)}, \ldots] } \right\} \times \left\{ \frac{ p[\rho^{(c)}] }{ p[\rho^{(r - 1)] } } \right\} \times \left\{ \frac{ p[\nu^{(c)}] }{ p[\nu^{(r - 1)}] } \right\} \times \left\{ \frac{ p[ \gamma^{(c)} ] }{ p[\gamma^{(r - 1)} ] } \right\}.
-\end{align*}
-All three candidates are accepted with probability $\min\{R, 1\}$.
-
-\subsection{Posterior distributions} \label{a:posterior}
-\input{zpost.tex}
-\input{betapost.tex}
-\input{sigpost.tex}
-\input{lambdapost.tex}
-
-\subsection{Proof that $\lim_{h \rightarrow \infty} \pi(h) = 0$} \label{a:proofsamepartition}
-Consider a homogeneous spatial Poisson process with intensity $\mu$.
-Define $A$ as the circle with center \hbox{$(\bs_1 + \bs_2) / 2$} and radius $h / 2$.
-Then $\bs_1$ and $\bs_2$ are in different partitions almost surely if two or more points are in $A$.
-Let $N(A)$ be the number of points in $A$, and let
-\begin{align*}
-  \mu(A) = \mu |A| = \mu \pi \left(\frac{h}{2}\right)^2 = \lambda h^2.
-\end{align*}
-Then
-\begin{align*}
-  P[N(A) \ge 2] &= 1 - P[N(A) = 0] - P[N(A) = 1]\\
-                &= 1 - \exp\{-\lambda h^2\} - \lambda h^2 \exp\{-\lambda h^2\} \\
-                &= 1 - (1 + \lambda h^2) \exp\{-\lambda h^2\}
-\end{align*}
-which goes to one as $h \rightarrow \infty$.
-
-% Let $N(A)$ be the number of knots in $A$.
-% So,
-% \begin{align*}
-%   P[ N(A) = k] = \frac{ \mu(A)^k \exp\{ -\mu(A)\}}{k!}.
-% \end{align*}
-% Then for any finite $k$, $\lim_{h \rightarrow \infty} P[N(A) = k] = 0$ because $\lim_{h \rightarrow \infty} \mu(A) = \infty$.
-% With each additional knot in $A$, the chance that $\bs_1$ and $\bs_2$ will be be in the same partition will decrease, because partition membership is defined by the closest knot to a site.
-% Therefore, $\lim_{h \rightarrow \infty} \pi(h) = 0$.
-
-% \subsection{Half-normal distribution}
-% Let $u = |z|$ where $Z \sim N(\mu, \sigma^2)$.
-% Specifically, we consider the case where $\mu = 0$. Then $U$ follows a half-normal distribution which we denote as $U \sim HN(0, 1)$, and the density is given by
-% \begin{align}
-%   f_U(u) = \frac{ \sqrt{2} }{ \sqrt{\pi \sigma^2} } \exp \left( - \frac{ u^2 }{ 2 \sigma^2 } \right) I(u > 0)
-% \end{align}
-% When $\mu = 0$, the half-normal distribution is also equivalent to a $N_{(0, \infty)}(0, \sigma^2)$ where $N_{(a, b)}(\mu, \sigma^2)$ represents a normal distribution with mean $\mu$ and standard deviation $\sigma$ that has been truncated below at $a$ and above at $b$.
-
-\subsection{Skew-$t$ distribution} \label{a:skewt}
-\subsubsection*{Univariate extended skew-$t$ distribution}
-We say that $Y$ follow a univariate extended skew-$t$ distribution with location $\xi \in \calR$, scale $\omega > 0$, skew parameter $\alpha \in \calR$, extended parameter $\tau \in \calR$, and degrees of freedom $\nu$ if has distribution function
-\begin{align}
-  f_{\text{EST}}(y) = \omega^{-1} \frac{ f_T (z; \nu) }{ F_T(\tau / \sqrt{ 1 + \alpha^2}; \nu)} F_T\left[ (\alpha z + \tau) \sqrt{ \frac{ \nu + 1 }{ \nu + z^2}}; 0, 1, \nu + 1 \right]
-\end{align}
-where $f_T(t; \nu)$ is a univariate Student's $t$ with $\nu$ degrees of freedom, $F_T(t; \nu) = P(T < t)$, and \hbox{$z = (y - \xi) / \omega$}.
-In the case that $\tau = 0$, then $Y$ follows a univariate skew-$t$ distribution.
-
-\subsection*{Multivariate skew-$t$ distribution}
-If $\bZ \sim \text{ST}_d(0, \bar{\bOmega}, \balpha, \eta)$ is a $d$-dimensional skew-$t$ distribution, and $\bY = \xi + \bomega \bZ$, where $\bomega = \text{diag}(\omega_1, \ldots, \omega_d)$, then the density of $Y$ at $y$ is
-\begin{align}
-  f_y(\by) = det(\bomega)^{-1} f_z(\bz)
-\end{align}
-where
-\begin{align}
-  f_z(\bz) &= 2 t_d(\bz; \bar{\bOmega}, \eta) T \left[ \balpha^T \bz \sqrt{ \frac{\eta + d}{\nu + Q(\bz)} }; \eta + d\right] \\
-  \bz &= \bomega^{-1}(\by - \xi)
-\end{align}
-where $t_d(\bz; \bar{\bOmega}, \eta)$ is a $d$-dimensional Student's $t$-distribution with scale matrix $\bar{\bOmega}$ and degrees of freedom $\eta$, $Q(z) = \bz^T \bar{\Omega}^{-1}\bz$ and $T(\cdot; \eta)$ denotes the univariate Student's $t$ distribution function with $\eta$ degrees of freedom \citep{Azzalini2014}.
-
-\subsubsection*{Extremal dependence}
-For a bivariate skew-$t$ random variable $\bY = [Y(\bs), Y(\bt)]^T$, the $\chi(h)$ statistic \citep{Padoan2011} is given by
-\begin{align} \label{eq:chiskew-t}
-  \chi(h) = \bar{F}_{\text{EST}}\left\{ \frac{[x_1^{1 / \eta} - \varrho(h)] \sqrt{\eta + 1} }{\sqrt{1 - \varrho(h)^2}}; 0, 1, \alpha_1, \tau_1, \eta + 1 \right\} + \bar{F}_{\text{EST}}\left\{ \frac{ [x_2^{1 / \eta} - \varrho(h)] \sqrt{\eta + 1} }{ \sqrt{1 - \varrho(h)^2} }; 0, 1, \alpha_2, \tau_2, \eta + 1 \right\},
-\end{align}
-where $\bar{F}_{\text{EST}}$ is the univariate survival extended skew-$t$ function with zero location and unit scale, \hbox{$\varrho(h) = \text{cor}[y(\bs), y(\bt)]$}, $\alpha_j = \alpha_i \sqrt{1 - \varrho^2}$, $\tau_j = \sqrt{\eta + 1}(\alpha_j + \alpha_i \varrho)$, and $x_j = F_T(\bar{\alpha}_i \sqrt{\eta + 1}; 0, 1, \eta) / F_T(\bar{\alpha}_j \sqrt{\eta + 1}; 0, 1, \eta)$ with $j = 1, 2$ and $i = 2, 1$ and where $\bar{\alpha}_j = (\alpha_j + \alpha_i \varrho) / \sqrt{ 1 + \alpha_i^2 [1 - \varrho(h)^2]}$.
-
-\subsubsection*{Proof that $\lim_{h \rightarrow \infty} \chi(h) > 0$}
-Consider the bivariate distribution of $\bY = [Y(\bs), Y(\bt)]^T$, with $\varrho(h)$ given by (\ref{eq:matern}).
-So, $\lim_{h \rightarrow \infty} \varrho(h) = 0$.
-Then
-\begin{align}
-  \lim_{h \rightarrow \infty} \chi(h) = \bar{F}_{\text{EST}}\left\{ \sqrt{\eta + 1}; 0, 1, \alpha_1, \tau_1, \eta + 1 \right\} + \bar{F}_{\text{EST}}\left\{ \sqrt{\eta + 1}; 0, 1, \alpha_2, \tau_2, \eta + 1 \right\}.
-\end{align}
-Because the extended skew-$t$ distribution is not bounded above, for all $\bar{F}_{\text{EST}}(x) = 1 - F_{\text{EST}} > 0$ for all $x < \infty$.
-Therefore, for a skew-$t$ distribution, $\lim_{h \rightarrow \infty} \chi(h) > 0$.
-
-\subsection{Simulation study pairwise difference results} \label{a:pdiffs}
-The following tables show the methods that have significantly different Brier scores when using a Wilcoxon-Nemenyi-McDonald-Thompson test.
-In each column, different letters signify that the methods have significantly different Brier scores.
-For example, there is significant evidence to suggest that method 1 and method 4 have different Brier scores at $q(0.90)$, whereas there is not significant evidence to suggest that method 1 and method 2 have different Brier scores at $q(0.90)$.
-In each table group A represents the group with the lowest Brier scores.
-Groups are significant with a familywise error rate of $\alpha = 0.05$.
-
-\begin{table}[htbp]
-  \centering
-  \caption{Setting 1 -- Gaussian marginal, $K = 1$ knot}
-  \label{tbl:gaussim}
-  \begin{tabular}{|l|ccc|ccc|cccc|ccc|}
-    \cline{2-14}
-    \multicolumn{1}{c}{} & \multicolumn{3}{|c}{$q(0.90)$} & \multicolumn{3}{|c}{$q(0.95)$} & \multicolumn{4}{|c}{$q(0.98)$} & \multicolumn{3}{|c|}{$q(0.99)$} \\
-    \hline
-    Method 1 & A &   &   & A &   &   & A &   &   &   & A & B &   \\
-    \hline
-    Method 2 & A &   &   & A &   &   & A &   &   &   & A &   &   \\
-    \hline
-    Method 3 &   & B &   &   & B &   &   &   & C &   &   & B &   \\
-    \hline
-    Method 4 & A &   &   & A &   &   & A & B &   &   & A & B &   \\
-    \hline
-    Method 5 &   & B &   &   & B &   &   & B & C &   & A & B &   \\
-    \hline
-    Method 6 &   &   & C &   &   & C &   &   &   & D &   &   & C \\
-    \hline
-  \end{tabular}
-\end{table}
-
-% \begin{table}[htbp]
-%   \centering
-%   \caption{Setting 2: Symmetric-$t$ marginal, $K = 1$ knot}
-%   \label{tbl:t1k1sim}
-%   \begin{tabular}{|l|cc|cccc|cccc|ccc|}
-%     \cline{2-14}
-%     \multicolumn{1}{c}{} & \multicolumn{2}{|c}{$q(0.90)$} & \multicolumn{4}{|c}{$q(0.95)$} & \multicolumn{4}{|c}{$q(0.98)$} & \multicolumn{3}{|c|}{$q(0.99)$} \\
-%     \hline
-%     Method 1 &   & B &   &   &   & D &   &   &   & D &   &   & C \\
-%     \hline
-%     Method 2 & A &   & A &   &   &   & A &   &   &   & A &   &   \\
-%     \hline
-%     Method 3 &   & B &   & B &   &   &   & B & C &   &   & B &   \\
-%     \hline
-%     Method 4 & A &   &   &   & C &   &   & B &   &   &   & B &   \\
-%     \hline
-%     Method 5 &   & B &   &   &   & D &   &   & C & D &   & B & C \\
-%     \hline
-%   \end{tabular}
-% \end{table}
-
-% \begin{table}[htbp]
-%   \centering
-%   \caption{Setting 3: Symmetric-$t$ marginal, $K = 5$ knots}
-%   \label{tbl:gaussim}
-%   \begin{tabular}{|l|cc|cc|cc|cc|}
-%     \cline{2-9}
-%     \multicolumn{1}{c}{} & \multicolumn{2}{|c}{$q(0.90)$} & \multicolumn{2}{|c}{$q(0.95)$} & \multicolumn{2}{|c}{$q(0.98)$} & \multicolumn{2}{|c|}{$q(0.99)$} \\
-%     \hline
-%     Method 1 &   & B &   & B &   & B &   & B \\
-%     \hline
-%     Method 2 &   & B &   & B &   & B &   & B \\
-%     \hline
-%     Method 3 &   & B &   & B &   & B & A & B \\
-%     \hline
-%     Method 4 & A &   & A &   & A &   & A &   \\
-%     \hline
-%     Method 5 &   & B &   & B & A & B & A & B \\
-%     \hline
-%   \end{tabular}
-% \end{table}
-
-\begin{table}[htbp]
-  \centering
-  \caption{Setting 2 -- Skew-$t$ marginal, $K = 1$ knot}
-  \label{tbl:st1sim}
-  \begin{tabular}{|l|ccccc|cccc|cccc|ccc|}
-    \cline{2-17}
-    \multicolumn{1}{c}{} & \multicolumn{5}{|c}{$q(0.90)$} & \multicolumn{4}{|c}{$q(0.95)$} & \multicolumn{4}{|c}{$q(0.98)$} & \multicolumn{3}{|c|}{$q(0.99)$} \\
-    \hline
-    Method 1 &   &   & C &   &   &   & B &   &   &   & B & C &   &   & B &   \\
-    \hline
-    Method 2 & A &   &   &   &   & A &   &   &   & A &   &   &   & A &   &   \\
-    \hline
-    Method 3 &   & B & C &   &   & A & B &   &   & A & B &   &   & A & B &   \\
-    \hline
-    Method 4 & A & B &   &   &   &   & B &   &   &   & B &   &   & A &   &   \\
-    \hline
-    Method 5 &   &   &   & D &   &   &   & C &   &   &   & C &   &   & B &   \\
-    \hline
-    Method 6 &   &   &   &   & E &   &   &   & D &   &   &   & D &   &   & C \\
-    \hline
-  \end{tabular}
-\end{table}
-
-\begin{table}[htbp]
-  \centering
-  \caption{Setting 3 -- Skew-$t$ marginal, $K = 5$ knots}
-  \label{tbl:st5sim}
-  \begin{tabular}{|l|ccc|cccc|ccc|ccc|}
-    \cline{2-14}
-    \multicolumn{1}{c}{} & \multicolumn{3}{|c}{$q(0.90)$} & \multicolumn{4}{|c}{$q(0.95)$} & \multicolumn{3}{|c}{$q(0.98)$} & \multicolumn{3}{|c|}{$q(0.99)$} \\
-    \hline
-    Method 1 &   & B &   &   &   & C &   &   & B &   &   & B &   \\
-    \hline
-    Method 2 &   & B &   &   &   & C &   &   & B &   &   & B &   \\
-    \hline
-    Method 3 & A &   &   &   & B &   &   &   & B &   &   & B &   \\
-    \hline
-    Method 4 & A &   &   & A &   &   &   & A &   &   & A &   &   \\
-    \hline
-    Method 5 & A &   &   & A &   &   &   & A &   &   & A &   &   \\
-    \hline
-    Method 6 &   &   & C &   &   &   & D &   &   & C &   &   & C \\
-    \hline
-  \end{tabular}
-\end{table}
-
-\begin{table}[htbp]
-  \centering
-  \caption{Setting 4 -- Max-stable}
-  \label{tbl:mssim}
-  \begin{tabular}{|l|cccc|cccc|ccc|ccc|}
-    \cline{2-15}
-    \multicolumn{1}{c}{} & \multicolumn{4}{|c}{$q(0.90)$} & \multicolumn{4}{|c}{$q(0.95)$} & \multicolumn{3}{|c}{$q(0.98)$} & \multicolumn{3}{|c|}{$q(0.99)$} \\
-    \hline
-    Method 1 & A & B &   &   &   & B &   &   &   & B &   &   &   & C \\
-    \hline
-    Method 2 &   & B &   &   &   & B & C &   &   & B &   &   & B & C \\
-    \hline
-    Method 3 &   &   & C & D &   &   & C &   &   & B &   &   & B &   \\
-    \hline
-    Method 4 &   &   &   & D &   &   &   & D &   &   & C &   &   & C \\
-    \hline
-    Method 5 &   &   & C &   &   &   & C &   &   & B &   &   & B & C \\
-    \hline
-    Method 6 & A &   &   &   & A &   &   &   & A &   &   & A &   &   \\
-    \hline
-  \end{tabular}
-\end{table}
-
-\begin{table}[htbp]
-  \centering
-  \caption{Setting 5 -- Transformation below $T = q(0.80)$}
-  \label{tbl:transsim}
-  \begin{tabular}{|l|cccc|ccc|cccc|cccc|}
-    \cline{2-16}
-    \multicolumn{1}{c}{} & \multicolumn{4}{|c}{$q(0.90)$} & \multicolumn{3}{|c}{$q(0.95)$} & \multicolumn{4}{|c}{$q(0.98)$} & \multicolumn{4}{|c|}{$q(0.99)$} \\
-    \hline
-    Method 1 &   &   & C &   &   & B &   &   &   & C &   &   &   & C &   \\
-    \hline
-    Method 2 &   & B &   &   &   & B &   &   & B &   &   & A & B &   &   \\
-    \hline
-    Method 3 & A &   &   &   & A &   &   & A &   &   &   & A &   &   &   \\
-    \hline
-    Method 4 &   & B & C &   &   & B &   &   & B &   &   &   & B & C &   \\
-    \hline
-    Method 5 &   & B &   &   &   & B &   &   & B & C &   &   &   & C &   \\
-    \hline
-    Method 6 &   &   &   & D &   &   & C &   &   &   & D &   &   &   & D \\
-    \hline
-  \end{tabular}
-\end{table}
-
 \begin{singlespace}

### 2015-09-09 LaTeX/supplemental.tex
move appendices into supplemental document

\documentclass[11pt]{article}
\usepackage{amssymb, amsthm, amsmath}
\usepackage{bm}
\usepackage{graphicx}
\usepackage[authoryear]{natbib}
\usepackage{bm}
\usepackage{verbatim}
\usepackage{lineno}
\usepackage{times}
\usepackage{soul}
\usepackage{color}
\usepackage{enumitem}
\usepackage{setspace}

% cross-referencing appendices
\usepackage{xr}
\externaldocument{skewnormal}

\usepackage[left=1in,top=1in,right=1in]{geometry}
\pdfpageheight 11in
\pdfpagewidth 8.5in
\linespread{2.0}
\input{mycommands.sty}

\begin{document}\linenumbers
\begin{center}
{\Large {\bf Supplemental material for A space-time skew-$t$ model for threshold exceedances}}\\

{\large Samuel A Morris\footnote[1]{North Carolina State University}, Brian J Reich\footnotemark[1]{}, Emeric Thibaud\footnote[2]{Colorado State University}, and Daniel Cooley\footnotemark[2]{}}

\today
\end{center}

\appendix
\section{Appendices}
\subsection{MCMC details} \label{a:mcmc}
The MCMC sampling for the model \ref{s:hier} is done using {\tt R} (http://www.r-project.org). Whenever possible, we select conjugate priors (see Appendix \ref{a:posterior}); however, for some of the parameters, no conjugate prior distributions exist.
For these parameters, we use a random walk Metropolis-Hastings update step.
In each Metropolis-Hastings update, we tune the algorithm during the burn-in period to give acceptance rates near 0.40.

\subsubsection*{Spatial knot locations}
For each day, we update the spatial knot locations, $\bw_1, \ldots, \bw_K$, using a Metropolis-Hastings block update.
Because the spatial domain is bounded, we generate candidate knots using the transformed knots $\bw^*_1, \ldots, \bw^*_K$ (see section \ref{s:temporal}) and a random walk bivariate Gaussian candidate distribution
\begin{align*}
	{\bw^*_k}^{(c)} \sim \text{N}({\bw^*_k}^{(r - 1)}, s^2 I_2)
\end{align*}
where ${\bw^*_k}^{(r - 1)}$ is the location for the transformed knot at MCMC iteration $r - 1$, $s$ is a tuning parameter, and $I_2$ is an identity matrix.
After candidates have been generated for all $K$ knots, the acceptance ratio is
\begin{align*}
  R = \left\{ \frac{ l[ Y_t(\bs | \bw_1^{(c)}, \ldots, \bw_K^{(c)}, \ldots)] }{l[ Y_t(\bs | \bw_1^{(r - 1)}, \ldots, \bw_K^{(r - 1)}, \ldots)]} \right\} \times \left\{ \frac{ \prod_{k = 1}^{K}\phi(\bw_k^{(c)})}{ \prod_{k = 1}^{K}\phi(\bw_k^{(r - 1)})} \right\} \times \left\{ \frac{ \prod_{k = 1}^{K} p({\bw^*_k}^{(c)})}{ \prod_{k = 1}^{K} p({\bw^*_k}^{(r - 1)})}\right\}
\end{align*}
where $l$ is the likelihood given in (\ref{eq:hier}), and $p(\cdot)$ is the prior either taken from the time series given in (\ref{s:temporal}) or assumed to be uniform over $\calD$.
The candidate knots are accepted with probability $\min\{R, 1\}$.

\subsubsection*{Spatial random effects}
If there is no temporal dependence amongst the observations, we use a Gibbs update for $z_{tk}$, and the posterior distribution is given in \ref{a:posterior}.
If there is temporal dependence amongst the observations, then we update $z_{tk}$ using a Metropolis-Hastings update.
Because this model uses $|z_{tk}|$, we generate candidate random effects using the $z^*_{tk}$ (see Section \ref{s:temporal}) and a random walk Gaussian candidate distribution
\begin{align*}
  {z^*_{tk}}^{(c)} \sim \text{N}({z^*_{tk}}^{(r - 1)}, s^2)
\end{align*}
where ${z^*_{tk}}^{(r-1)}$ is the value at MCMC iteration $r - 1$, and $s$ is a tuning parameter.
The acceptance ratio is
\begin{align*}
  R = \left\{ \frac{ l[Y_t(\bs) | z_{tk}^{(c)}, \ldots] }{ l[Y_t(\bs) | z_{tk}^{(r - 1)}]} \right\} \times \left\{ \frac{ p[ z_{tk}^{(c)} ] }{ p[ z_{tk}^{(r - 1)}]}\right\}
\end{align*}
where $p[\cdot]$ is the prior taken from the time series given in Section \ref{s:temporal}.
The candidate is accepted with probability $\min\{R, 1\}$.

\subsubsection*{Variance terms}
When there is more than one site in a partition, then we update $\sigma^2_{tk}$ using a Metropolis-Hastings update.
First, we generate a candidate for $\sigma^2_{tk}$ using an IG$(a^*/s, b^*/s)$ candidate distribution in an independence Metropolis-Hastings update where $a^* = (n_{tk} + 1) / 2 + a$, $b^* = [Y_{tk}^T \Sigma^{-1}_{tk} Y_{tk} + z_{tk}^2] / 2 + b$, $n_{tk}$ is the number of sites in partition $k$ on day $t$, and $Y_{tk}$ and $\Sigma^{-1}_{tk}$ are the observations and precision matrix for partition $k$ on day $t$.
The acceptance ratio is
\begin{align*}
  R = \left\{
    \frac{ l[Y_t(\bs) | {\sigma^2_{tk}}^{(c)}, \ldots] }
         { l[Y_t(\bs) | {\sigma^2_{tk}}^{(r - 1)}]}
    \right\} \times \left\{
    \frac{ l[z_{tk} | {\sigma^2_{tk}}^{(c)}, \ldots] }
         { l[z_{tk} | {\sigma^2_{tk}}^{(r - 1)}, \ldots] }
    \right\} \times \left\{
    \frac{ p[ {\sigma^2_{tk}}^{(c)} ] }
         { p[ {\sigma^2_{tk}}^{(r - 1)}] }
    \right\} \times \left\{
    \frac{ c[ {\sigma^2_{tk}}^{(r - 1)}] }
         { c[ {\sigma^2_{tk}}^{(c)}]}
    \right\}
\end{align*}
where $p[\cdot]$ is the prior either taken from the time series given in Section \ref{s:temporal} or assumed to be IG$(a, b)$, and $c[\cdot]$ is the candidate distribution.
The candidate is accepted with probability $\min\{R, 1\}$.

\subsubsection*{Spatial covariance parameters}
We update the three spatial covariance parameters, $\log(\rho)$, $\log(\nu)$, $\gamma$, using a Metropolis-Hastings block update step.
First, we generate a candidate using a random walk Gaussian candidate distribution
\begin{align*}
	\log(\rho)^{(c)} \sim \text{N}(\log(\rho)^{(r - 1)}, s^2)
\end{align*}
where $\log(\rho)^{(r-1)}$ is the value at MCMC iteration $r - 1$, and $s$ is a tuning parameter.
Candidates are generated for $\log(\nu)$ and $\gamma$ in a similar fashion.
The acceptance ratio is
\begin{align*}
	R = \left\{ \frac{ \prod_{t = 1}^{T} l[Y_t(\bs) | \rho^{(c)}, \nu^{(c)}, \gamma^{(c)}, \ldots] }{\prod_{t = 1}^{T} l[Y_t(\bs) | \rho^{(r-1)}, \nu^{(r-1)}, \gamma^{(r-1)}, \ldots] } \right\} \times \left\{ \frac{ p[\rho^{(c)}] }{ p[\rho^{(r - 1)] } } \right\} \times \left\{ \frac{ p[\nu^{(c)}] }{ p[\nu^{(r - 1)}] } \right\} \times \left\{ \frac{ p[ \gamma^{(c)} ] }{ p[\gamma^{(r - 1)} ] } \right\}.
\end{align*}
All three candidates are accepted with probability $\min\{R, 1\}$.

\subsection{Posterior distributions} \label{a:posterior}

\subsubsection*{Conditional posterior of $z_{tk} \mid \ldots $}\label{s:mvcondu}
If knots are independent over days, then the conditional posterior distribution of $|z_{tk}|$ is conjugate.
For simplicity, drop the subscript $t$, let $\tilde{z}_{tk} = |z_{tk}|$, and define
\begin{align*}
R(\bs) = \left\{
    \begin{array}{ll}
        Y(\bs) - X(\bs) \beta &s \in P_l\\[1em]
        Y(\bs) - X(\bs) \beta - \lambda \tilde{z}(\bs) \qquad & s \notin P_l
    \end{array}
\right.
\end{align*}
Let
\begin{align*}
    R_1 &= \text{the vector of } R(\bs) \text{ for } s \in P_l \\
    R_2 &= \text{the vector of } R(\bs) \text{ for } s \notin P_l \\
    \Omega &= \Sigma^{-1}.
\end{align*}
Then
\begin{align*}
    \pi(z_l | \ldots) &\propto \exp \left\{ -\frac{ 1 }{ 2 } \left[
        \left( \begin{array}{c}
            R_1 - \lambda \tilde{z}_l \bOne\\
            R_2
        \end{array} \right)^T
        \left( \begin{array}{cc}
            \Omega_{11} & \Omega_{12}\\
            \Omega_{21} & \Omega_{22}
        \end{array} \right)
        \left( \begin{array}{c}
            R_1 - \lambda \tilde{z}_l \bOne\\
            R_2
        \end{array} \right)
        +  \frac{ {\tilde{z}_l}^2 }{ \sigma_l^2 }\right]
    \right\} I(z_l > 0) \\
        &\propto \exp \left\{ -\frac{ 1 }{ 2 } \left[ \Lambda_l {\tilde{z}_l}^2 - 2 \mu_l \tilde{z}_l \right] \right\}
\end{align*}
where
\begin{align*}
    \mu_l &= \lambda ( R_1^T \Omega_{11} + R_2^T \Omega_{21} )\bOne\\
    \Lambda_l &= \lambda^2 \bOne^T \Omega_{11} \bOne + \frac{ 1 }{ \sigma^2_l }.
\end{align*}
Then $\tilde{Z}_l | \ldots \sim N_{(0, \infty)} (\Lambda_l^{-1} \mu_l, \Lambda_l^{-1})$

\subsection*{Conditional posterior of $\beta \mid \ldots$}\label{s:betapost}
Let $\beta \sim \mbox{N}_{p}(0, \Lambda_0)$ where $\Lambda_0$ is a precision matrix. Then
\begin{align*}
    \pi(\beta \mid \ldots) & \propto \exp \left\{ - \frac{ 1 }{ 2 } \beta^T \Lambda_0 \beta - \frac{ 1 }{ 2 } \sum_{t = 1 }^T [\bY_t - X_t \beta - \lambda |z_t|]^T \Omega [\bY_t - X_t \beta - \lambda |z_t|] \right\}\\
     & \propto \exp \left\{ -\frac{ 1 }{ 2 } \left[ \beta^T \Lambda_\beta \beta  - 2 \sum_{ t = 1 }^T [ \beta^T X_t^T \Omega (\bY_t - \lambda |z_t| )] \right] \right\}\\
     & \propto \mbox{N} ( \Lambda_\beta^{-1} \mu_\beta , \Lambda_\beta^{-1})
\end{align*}
where
\begin{align*}
    \mu_\beta &= \sum_{t=1}^{T} \left[ X_t^T \Omega (\bY_t - \lambda |z_t|) \right]\\
    \Lambda_\beta &= \Lambda_0 + \sum_{ t = 1 }^{ T} X_t^T \Omega X_t.
\end{align*}

\subsection*{Conditional posterior of $\sigma^2 \mid \ldots$}\label{s:sigpost}
In the case where $L = 1$ and temporal dependence is negligible, then $\sigma^2$ has a conjugate posterior distribution.
Let $\sigma_t^2 \iid \mbox{IG}(\alpha_0, \beta_0)$. For simplicity, drop the subscript $t$. Then
\begin{align*}
    \pi(\sigma^2 \mid \ldots) & \propto (\sigma^2)^{ -\alpha_0 - 1 / 2 - n / 2 - 1} \exp \left\{ -\frac{\beta_0}{\sigma^2} - \frac{ |z|^2 }{2 \sigma^2} - \frac{ (\bY - \bmu)^T \Sigma^{-1} (\bY - \bmu) }{2 \sigma^2} \right\} \\
    & \propto (\sigma^2)^{ -\alpha_0 - 1 / 2 - n / 2 - 1} \exp \left\{ - \frac{ 1 }{ \sigma^2 } \left[\beta_0 + \frac{ |z|^2 }{ 2 } + \frac{ 1 }{ 2 } (\bY - \bmu)^T \Sigma^{-1} (\bY - \mu) \right] \right\} \\
    & \propto \mbox{IG} (\alpha^*, \beta^*)
\end{align*}
where
\begin{align*}
    \alpha^* &= \alpha_0 + \frac{1}{2} + \frac{n}{2} \\
    \beta^* &= \beta_0 + \frac{ |z|^2 }{ 2 } + \frac{ 1 }{ 2 }(\bY - \bmu)^T \Sigma^{-1} (\bY - \bmu).
\end{align*}
In the case that $L > 1$, a random walk Metropolis Hastings step will be used to update $\sigma^2_{lt}$.

\subsection*{Conditional posterior of $\lambda \mid \ldots$}\label{s:lambdapost}
For convergence purposes we model $\lambda = \lambda_1 \lambda_2$ where
\begin{align}
  \lambda_1 &= \left\{ \begin{array}{ll}
      +1 \quad & \text{w.p.} 0.5\\
      -1 \quad & \text{w.p.} 0.5
   \end{array}\right.\\
   \lambda^2_2 & \sim IG(\alpha_\lambda, \beta_\lambda).\\
\end{align}
Then
\begin{align*}
  \pi(\lambda_2 \mid \ldots) &\propto {\lambda_2^2}^{(-\alpha_\lambda - 1)} \exp\left\{ - \frac{ \beta_\lambda }{ \lambda_2^2} \right\} \prod_{t = 1}^{T} \prod_{k = 1}^K \frac{1} {\lambda_2} \exp\left\{ - \frac{ z^2_{tk} }{ 2 \lambda_2^2  \sigma_{tk})^2} \right\} \\
   & \propto {\lambda_2^2}^{(-\alpha_\lambda - kt - 1)} \exp \left\{ -\frac{1}{\lambda^2_2} \left[ \beta_\lambda + \frac{ z^2 }{2 \sigma^2_{tk}} \right] \right\}
\end{align*}
Then $\lambda_2 \mid \ldots \sim IG \left(\alpha_\lambda + kt, \beta_\lambda + \frac{ z^2 }{2 \sigma^2_{tk}} \right)$

\subsection{Proof that $\lim_{h \rightarrow \infty} \pi(h) = 0$} \label{a:proofsamepartition}
Consider a homogeneous spatial Poisson process with intensity $\mu$.
Define $A$ as the circle with center \hbox{$(\bs_1 + \bs_2) / 2$} and radius $h / 2$.
Then $\bs_1$ and $\bs_2$ are in different partitions almost surely if two or more points are in $A$.
Let $N(A)$ be the number of points in $A$, and let
\begin{align*}
  \mu(A) = \mu |A| = \mu \pi \left(\frac{h}{2}\right)^2 = \lambda h^2.
\end{align*}
Then
\begin{align*}
  P[N(A) \ge 2] &= 1 - P[N(A) = 0] - P[N(A) = 1]\\
                &= 1 - \exp\{-\lambda h^2\} - \lambda h^2 \exp\{-\lambda h^2\} \\
                &= 1 - (1 + \lambda h^2) \exp\{-\lambda h^2\}
\end{align*}
which goes to one as $h \rightarrow \infty$.

% Let $N(A)$ be the number of knots in $A$.
% So,
% \begin{align*}
%   P[ N(A) = k] = \frac{ \mu(A)^k \exp\{ -\mu(A)\}}{k!}.
% \end{align*}
% Then for any finite $k$, $\lim_{h \rightarrow \infty} P[N(A) = k] = 0$ because $\lim_{h \rightarrow \infty} \mu(A) = \infty$.
% With each additional knot in $A$, the chance that $\bs_1$ and $\bs_2$ will be be in the same partition will decrease, because partition membership is defined by the closest knot to a site.
% Therefore, $\lim_{h \rightarrow \infty} \pi(h) = 0$.

% \subsection{Half-normal distribution}
% Let $u = |z|$ where $Z \sim N(\mu, \sigma^2)$.
% Specifically, we consider the case where $\mu = 0$. Then $U$ follows a half-normal distribution which we denote as $U \sim HN(0, 1)$, and the density is given by
% \begin{align}
%   f_U(u) = \frac{ \sqrt{2} }{ \sqrt{\pi \sigma^2} } \exp \left( - \frac{ u^2 }{ 2 \sigma^2 } \right) I(u > 0)
% \end{align}
% When $\mu = 0$, the half-normal distribution is also equivalent to a $N_{(0, \infty)}(0, \sigma^2)$ where $N_{(a, b)}(\mu, \sigma^2)$ represents a normal distribution with mean $\mu$ and standard deviation $\sigma$ that has been truncated below at $a$ and above at $b$.

\subsection{Skew-$t$ distribution} \label{a:skewt}
\subsubsection*{Univariate skew-$t$ distribution}
We say that $Y$ follows a univariate extended skew-$t$ distribution with location $\xi \in \calR$, scale $\omega > 0$, skew parameter $\alpha \in \calR$, and degrees of freedom $\nu$ if has distribution function
\begin{align}
  f_{\text{EST}}(y) = 2 f_T (z; \nu) F_T\left[ \alpha z \sqrt{ \frac{ \nu + 1 }{ \nu + z^2}}; \nu + 1 \right]
\end{align}
where $f_T(t; \nu)$ is a univariate Student's $t$ with $\nu$ degrees of freedom, $F_T(t; \nu) = P(T < t)$, and \hbox{$z = (y - \xi) / \omega$}.

\subsection*{Multivariate skew-$t$ distribution}
If $\bZ \sim \text{ST}_d(0, \bar{\bOmega}, \balpha, \eta)$ is a $d$-dimensional skew-$t$ distribution, and $\bY = \xi + \bomega \bZ$, where $\bomega = \text{diag}(\omega_1, \ldots, \omega_d)$, then the density of $Y$ at $y$ is
\begin{align}
  f_y(\by) = det(\bomega)^{-1} f_z(\bz)
\end{align}
where
\begin{align}
  f_z(\bz) &= 2 t_d(\bz; \bar{\bOmega}, \eta) T \left[ \balpha^T \bz \sqrt{ \frac{\eta + d}{\nu + Q(\bz)} }; \eta + d\right] \\
  \bz &= \bomega^{-1}(\by - \xi)
\end{align}
where $t_d(\bz; \bar{\bOmega}, \eta)$ is a $d$-dimensional Student's $t$-distribution with scale matrix $\bar{\bOmega}$ and degrees of freedom $\eta$, $Q(z) = \bz^T \bar{\Omega}^{-1}\bz$ and $T(\cdot; \eta)$ denotes the univariate Student's $t$ distribution function with $\eta$ degrees of freedom \citep{Azzalini2014}.

\subsubsection*{Extremal dependence}
For a bivariate skew-$t$ random variable $\bY = [Y(\bs), Y(\bt)]^T$, the $\chi(h)$ statistic \citep{Padoan2011} is given by
\begin{align} \label{eq:chiskew-t}
  \chi(h) = \bar{F}_{\text{EST}}\left\{ \frac{[x_1^{1 / \eta} - \varrho(h)] \sqrt{\eta + 1} }{\sqrt{1 - \varrho(h)^2}}; 0, 1, \alpha_1, \tau_1, \eta + 1 \right\} + \bar{F}_{\text{EST}}\left\{ \frac{ [x_2^{1 / \eta} - \varrho(h)] \sqrt{\eta + 1} }{ \sqrt{1 - \varrho(h)^2} }; 0, 1, \alpha_2, \tau_2, \eta + 1 \right\},
\end{align}
where $\bar{F}_{\text{EST}}$ is the univariate survival extended skew-$t$ function with zero location and unit scale, \hbox{$\varrho(h) = \text{cor}[y(\bs), y(\bt)]$}, $\alpha_j = \alpha_i \sqrt{1 - \varrho^2}$, $\tau_j = \sqrt{\eta + 1}(\alpha_j + \alpha_i \varrho)$, and $x_j = F_T(\bar{\alpha}_i \sqrt{\eta + 1}; 0, 1, \eta) / F_T(\bar{\alpha}_j \sqrt{\eta + 1}; 0, 1, \eta)$ with $j = 1, 2$ and $i = 2, 1$ and where $\bar{\alpha}_j = (\alpha_j + \alpha_i \varrho) / \sqrt{ 1 + \alpha_i^2 [1 - \varrho(h)^2]}$.

\subsubsection*{Proof that $\lim_{h \rightarrow \infty} \chi(h) > 0$}
Consider the bivariate distribution of $\bY = [Y(\bs), Y(\bt)]^T$, with $\varrho(h)$ given by (\ref{eq:matern}).
So, $\lim_{h \rightarrow \infty} \varrho(h) = 0$.
Then
\begin{align}
  \lim_{h \rightarrow \infty} \chi(h) = \bar{F}_{\text{EST}}\left\{ \sqrt{\eta + 1}; 0, 1, \alpha_1, \tau_1, \eta + 1 \right\} + \bar{F}_{\text{EST}}\left\{ \sqrt{\eta + 1}; 0, 1, \alpha_2, \tau_2, \eta + 1 \right\}.
\end{align}
Because the extended skew-$t$ distribution is not bounded above, for all $\bar{F}_{\text{EST}}(x) = 1 - F_{\text{EST} (x)} > 0$ for all $x < \infty$.
Therefore, for a skew-$t$ distribution, $\lim_{h \rightarrow \infty} \chi(h) > 0$.

\subsection{Simulation study pairwise difference results} \label{a:pdiffs}
The following tables show the methods that have significantly different Brier scores when using a Wilcoxon-Nemenyi-McDonald-Thompson test.
In each column, different letters signify that the methods have significantly different Brier scores.
For example, there is significant evidence to suggest that method 1 and method 4 have different Brier scores at $q(0.90)$, whereas there is not significant evidence to suggest that method 1 and method 2 have different Brier scores at $q(0.90)$.
In each table group A represents the group with the lowest Brier scores.
Groups are significant with a familywise error rate of $\alpha = 0.05$.

\begin{table}[htbp]
  \centering
  \caption{Setting 1 -- Gaussian marginal, $K = 1$ knot}
  \label{tbl:gaussim}
  \begin{tabular}{|l|ccc|ccc|cccc|ccc|}
    \cline{2-14}
    \multicolumn{1}{c}{} & \multicolumn{3}{|c}{$q(0.90)$} & \multicolumn{3}{|c}{$q(0.95)$} & \multicolumn{4}{|c}{$q(0.98)$} & \multicolumn{3}{|c|}{$q(0.99)$} \\
    \hline
    Method 1 & A &   &   & A &   &   & A &   &   &   & A & B &   \\
    \hline
    Method 2 & A &   &   & A &   &   & A &   &   &   & A &   &   \\
    \hline
    Method 3 &   & B &   &   & B &   &   &   & C &   &   & B &   \\
    \hline
    Method 4 & A &   &   & A &   &   & A & B &   &   & A & B &   \\
    \hline
    Method 5 &   & B &   &   & B &   &   & B & C &   & A & B &   \\
    \hline
    Method 6 &   &   & C &   &   & C &   &   &   & D &   &   & C \\
    \hline
  \end{tabular}
\end{table}

% \begin{table}[htbp]
%   \centering
%   \caption{Setting 2: Symmetric-$t$ marginal, $K = 1$ knot}
%   \label{tbl:t1k1sim}
%   \begin{tabular}{|l|cc|cccc|cccc|ccc|}
%     \cline{2-14}
%     \multicolumn{1}{c}{} & \multicolumn{2}{|c}{$q(0.90)$} & \multicolumn{4}{|c}{$q(0.95)$} & \multicolumn{4}{|c}{$q(0.98)$} & \multicolumn{3}{|c|}{$q(0.99)$} \\
%     \hline
%     Method 1 &   & B &   &   &   & D &   &   &   & D &   &   & C \\
%     \hline
%     Method 2 & A &   & A &   &   &   & A &   &   &   & A &   &   \\
%     \hline
%     Method 3 &   & B &   & B &   &   &   & B & C &   &   & B &   \\
%     \hline
%     Method 4 & A &   &   &   & C &   &   & B &   &   &   & B &   \\
%     \hline
%     Method 5 &   & B &   &   &   & D &   &   & C & D &   & B & C \\
%     \hline
%   \end{tabular}
% \end{table}

% \begin{table}[htbp]
%   \centering
%   \caption{Setting 3: Symmetric-$t$ marginal, $K = 5$ knots}
%   \label{tbl:gaussim}
%   \begin{tabular}{|l|cc|cc|cc|cc|}
%     \cline{2-9}
%     \multicolumn{1}{c}{} & \multicolumn{2}{|c}{$q(0.90)$} & \multicolumn{2}{|c}{$q(0.95)$} & \multicolumn{2}{|c}{$q(0.98)$} & \multicolumn{2}{|c|}{$q(0.99)$} \\
%     \hline
%     Method 1 &   & B &   & B &   & B &   & B \\
%     \hline
%     Method 2 &   & B &   & B &   & B &   & B \\
%     \hline
%     Method 3 &   & B &   & B &   & B & A & B \\
%     \hline
%     Method 4 & A &   & A &   & A &   & A &   \\
%     \hline
%     Method 5 &   & B &   & B & A & B & A & B \\
%     \hline
%   \end{tabular}
% \end{table}

\begin{table}[htbp]
  \centering
  \caption{Setting 2 -- Skew-$t$ marginal, $K = 1$ knot}
  \label{tbl:st1sim}
  \begin{tabular}{|l|ccccc|cccc|cccc|ccc|}
    \cline{2-17}
    \multicolumn{1}{c}{} & \multicolumn{5}{|c}{$q(0.90)$} & \multicolumn{4}{|c}{$q(0.95)$} & \multicolumn{4}{|c}{$q(0.98)$} & \multicolumn{3}{|c|}{$q(0.99)$} \\
    \hline
    Method 1 &   &   & C &   &   &   & B &   &   &   & B & C &   &   & B &   \\
    \hline
    Method 2 & A &   &   &   &   & A &   &   &   & A &   &   &   & A &   &   \\
    \hline
    Method 3 &   & B & C &   &   & A & B &   &   & A & B &   &   & A & B &   \\
    \hline
    Method 4 & A & B &   &   &   &   & B &   &   &   & B &   &   & A &   &   \\
    \hline
    Method 5 &   &   &   & D &   &   &   & C &   &   &   & C &   &   & B &   \\
    \hline
    Method 6 &   &   &   &   & E &   &   &   & D &   &   &   & D &   &   & C \\
    \hline
  \end{tabular}
\end{table}

\begin{table}[htbp]
  \centering
  \caption{Setting 3 -- Skew-$t$ marginal, $K = 5$ knots}
  \label{tbl:st5sim}
  \begin{tabular}{|l|ccc|cccc|ccc|ccc|}
    \cline{2-14}
    \multicolumn{1}{c}{} & \multicolumn{3}{|c}{$q(0.90)$} & \multicolumn{4}{|c}{$q(0.95)$} & \multicolumn{3}{|c}{$q(0.98)$} & \multicolumn{3}{|c|}{$q(0.99)$} \\
    \hline
    Method 1 &   & B &   &   &   & C &   &   & B &   &   & B &   \\
    \hline
    Method 2 &   & B &   &   &   & C &   &   & B &   &   & B &   \\
    \hline
    Method 3 & A &   &   &   & B &   &   &   & B &   &   & B &   \\
    \hline
    Method 4 & A &   &   & A &   &   &   & A &   &   & A &   &   \\
    \hline
    Method 5 & A &   &   & A &   &   &   & A &   &   & A &   &   \\
    \hline
    Method 6 &   &   & C &   &   &   & D &   &   & C &   &   & C \\
    \hline
  \end{tabular}
\end{table}

\begin{table}[htbp]
  \centering
  \caption{Setting 4 -- Max-stable}
  \label{tbl:mssim}
  \begin{tabular}{|l|cccc|cccc|ccc|ccc|}
    \cline{2-15}
    \multicolumn{1}{c}{} & \multicolumn{4}{|c}{$q(0.90)$} & \multicolumn{4}{|c}{$q(0.95)$} & \multicolumn{3}{|c}{$q(0.98)$} & \multicolumn{3}{|c|}{$q(0.99)$} \\
    \hline
    Method 1 & A & B &   &   &   & B &   &   &   & B &   &   &   & C \\
    \hline
    Method 2 &   & B &   &   &   & B & C &   &   & B &   &   & B & C \\
    \hline
    Method 3 &   &   & C & D &   &   & C &   &   & B &   &   & B &   \\
    \hline
    Method 4 &   &   &   & D &   &   &   & D &   &   & C &   &   & C \\
    \hline
    Method 5 &   &   & C &   &   &   & C &   &   & B &   &   & B & C \\
    \hline
    Method 6 & A &   &   &   & A &   &   &   & A &   &   & A &   &   \\
    \hline
  \end{tabular}
\end{table}

\begin{table}[htbp]
  \centering
  \caption{Setting 5 -- Transformation below $T = q(0.80)$}
  \label{tbl:transsim}
  \begin{tabular}{|l|cccc|ccc|cccc|cccc|}
    \cline{2-16}
    \multicolumn{1}{c}{} & \multicolumn{4}{|c}{$q(0.90)$} & \multicolumn{3}{|c}{$q(0.95)$} & \multicolumn{4}{|c}{$q(0.98)$} & \multicolumn{4}{|c|}{$q(0.99)$} \\
    \hline
    Method 1 &   &   & C &   &   & B &   &   &   & C &   &   &   & C &   \\
    \hline
    Method 2 &   & B &   &   &   & B &   &   & B &   &   & A & B &   &   \\
    \hline
    Method 3 & A &   &   &   & A &   &   & A &   &   &   & A &   &   &   \\
    \hline
    Method 4 &   & B & C &   &   & B &   &   & B &   &   &   & B & C &   \\
    \hline
    Method 5 &   & B &   &   &   & B &   &   & B & C &   &   &   & C &   \\
    \hline
    Method 6 &   &   &   & D &   &   & C &   &   &   & D &   &   &   & D \\
    \hline
  \end{tabular}
\end{table}

% \begin{singlespace}
% \bibliographystyle{rss}
% \bibliography{library}
% \end{singlespace}

\end{document}

### 2015-09-17 LaTeX/skewnormal.tex
setup biometrics style

@@
-\documentclass[11pt]{article}
-\usepackage{amssymb, amsthm, amsmath}
+% \documentclass[11pt]{article}
+\documentclass[useAMS,usenatbib,referee]{biom}
+% \usepackage{amssymb, amsthm, amsmath}
+\usepackage{amssymb, amsmath}
 \usepackage{bm}
 \usepackage{graphicx}
-\usepackage[authoryear]{natbib}
+% \usepackage[authoryear]{natbib}
+\usepackage{natbib}
 \usepackage{bm}
@@
 
-\usepackage[left=1in,top=1in,right=1in]{geometry}
-\pdfpageheight 11in
-\pdfpagewidth 8.5in
-\linespread{2.0}
+% \usepackage[left=1in,top=1in,right=1in]{geometry}
+% \pdfpageheight 11in
+% \pdfpagewidth 8.5in
+% \linespread{2.0}
 \input{mycommands.sty}
 
-\begin{document}\linenumbers
-\begin{center}
-{\Large {\bf A space-time skew-$t$ model for threshold exceedances}}\\
+\title[A Space-time Skew-$t$ Model for Threshold Exceedances]{A space-time skew-$t$ model for threshold exceedances}
+\author
+{Samuel A Morris$^{1,*}$\email{samorris@ncsu.edu},
+Brian J Reich$^{1}$,
+Emeric Thibaud$^{2}$, and
+Daniel Cooley$^{2}$\\
+$^{1}$Department of Statistics, North Carolina State University, Raleigh, North Carolina, U.S.A. \\
+$^{2}$Department of Statistics, Colorado State University, Fort Collins, Colorado, U.S.A.}
 
-{\large Samuel A Morris\footnote[1]{North Carolina State University}, Brian J Reich\footnotemark[1]{}, Emeric Thibaud\footnote[2]{Colorado State University}, and Daniel Cooley\footnotemark[2]{}}
+\begin{document} %\linenumbers
+% \begin{center}
+% {\Large {\bf A space-time skew-$t$ model for threshold exceedances}}\\
 
-\today
-\end{center}
+% {\large Samuel A Morris\footnote[1]{North Carolina State University}, Brian J Reich\footnotemark[1]{}, Emeric Thibaud\footnote[2]{Colorado State University}, and Daniel Cooley\footnotemark[2]{}}
+
+% \today
+% \end{center}
+
+\pagerange{\pageref{firstpage}--\pageref{lastpage}}
+
+\label{firstpage}
 
 \begin{abstract}
-\singlespacing
 To assess the compliance of air quality regulations, the Environmental Protection Agency (EPA) must know if a site exceeds a pre-specified threshold.
@@
 We apply our method to an ozone analysis for July 2005, and find that our model improves over both Gaussian and max-stable methods in terms of predicting exceedances over a fixed threshold.
-
-{\bf Key words}: Skew-$t$, random partition, MCMC, extreme value analysis, spatio-temporal
 \end{abstract}
 
-\newpage
+\begin{keywords}
+Skew-$t$, random partition, MCMC, extreme value analysis, spatio-temporal
+\end{keywords}
+
+\maketitle
+
 \section{Introduction}\label{s:intro}
@@
 For Gaussian processes, $\chi(h) = 0$ regardless of the distance $h$, so they are not suitable for modeling asymptotically dependent extremes.
-Unlike the Gaussian process, the skew-$t$ process is asymptotically dependent (the explicit expression for $\chi(h)$ is given in Appendix \ref{a:skewt}).
+Unlike the Gaussian process, the skew-$t$ process is asymptotically dependent (the explicit expression for $\chi(h)$ is given in \ref{a:skewt}).
 However, one problem with the spatial skew-$t$ process is that $\lim_{h \rightarrow \infty} \chi(h) > 0$.
@@
 Marginally, over the knots, $\chi(h) = \pi(h) \chi_{\text{skew}-t}(h)$, where $\pi(h) = \Pr(k_s = k_t)$ is the probability that two sites separated by distance $h$ are in the same partition.
-In Appendix \ref{a:proofsamepartition}, we show that $\lim_{h \rightarrow \infty} \pi(h) = 0$, implying $\lim_{h \rightarrow \infty} \chi(h) = 0$.
+In \ref{a:proofsamepartition}, we show that $\lim_{h \rightarrow \infty} \pi(h) = 0$, implying $\lim_{h \rightarrow \infty} \chi(h) = 0$.
 In Figure \ref{fig:chi}, we give $\chi(h)$ for $K = 1, 3, 5, 10$ partitions for a skew-$t$ distribution with $\alpha = 10$, and 3 degrees of freedom.
@@
 For Friedman tests that came back with a significant p-value, we conducted a Wilcoxon-Nemenyi-McDonald-Thompson test to see which of the methods had different results.
-The full results for the Wilcoxon-Nemenyi-McDonald-Thompson tests are given in Appendix \ref{a:pdiffs}.
+The full results for the Wilcoxon-Nemenyi-McDonald-Thompson tests are given in \ref{a:pdiffs}.
 
@@
 
+\backmatter
+
 \section*{Acknowledgments}
@@
 
-\begin{singlespace}
-\bibliographystyle{rss}
+\section*{Supplementary Materials}
+
+\bibliographystyle{biom}
 \bibliography{library}
-\end{singlespace}
+
+\label{lastpage}
 

### 2015-09-17 LaTeX/supplemental.tex
setup biometrics style

@@
-\documentclass[11pt]{article}
-\usepackage{amssymb, amsthm, amsmath}
+% \documentclass[11pt]{article}
+\documentclass[useAMS,usenatbib,referee]{biomweb}
+% \usepackage{amssymb, amsthm, amsmath}
+\usepackage{amssymb, amsmath}
 \usepackage{bm}
 \usepackage{graphicx}
-\usepackage[authoryear]{natbib}
+% \usepackage[authoryear]{natbib}
 \usepackage{bm}
@@
 \usepackage{setspace}
+\usepackage{appendix}
+% \usepackage{caption}
+% \captionsetup[table]{name=Web Table}
 
@@
 
-\usepackage[left=1in,top=1in,right=1in]{geometry}
-\pdfpageheight 11in
-\pdfpagewidth 8.5in
-\linespread{2.0}
+% \usepackage[left=1in,top=1in,right=1in]{geometry}
+% \pdfpageheight 11in
+% \pdfpagewidth 8.5in
+% \linespread{2.0}
 \input{mycommands.sty}
 
-\begin{document}\linenumbers
-\begin{center}
-{\Large {\bf Supplemental material for A space-time skew-$t$ model for threshold exceedances}}\\
+\title[Web-based Supplementary Materials for A Space-time Skew-$t$ Model for Threshold Exceedances]{Web-based Supplementary Materials for A Space-time Skew-$t$ Model for Threshold Exceedances by Morris, Reich, Thibaud, and Cooley}
+\author
+{Samuel A Morris$^{1,*}$\email{samorris@ncsu.edu},
+Brian J Reich$^{1}$,
+Emeric Thibaud$^{2}$, and
+Daniel Cooley$^{2}$\\
+$^{1}$Department of Statistics, North Carolina State University, Raleigh, North Carolina, U.S.A. \\
+$^{2}$Department of Statistics, Colorado State University, Fort Collins, Colorado, U.S.A.}
 
-{\large Samuel A Morris\footnote[1]{North Carolina State University}, Brian J Reich\footnotemark[1]{}, Emeric Thibaud\footnote[2]{Colorado State University}, and Daniel Cooley\footnotemark[2]{}}
+\begin{document} %\linenumbers
+\maketitle
+% \begin{center}
+% {\Large {\bf Supplemental material for A space-time skew-$t$ model for threshold exceedances}}\\
 
-\today
-\end{center}
+% {\large Samuel A Morris\footnote[1]{North Carolina State University}, Brian J Reich\footnotemark[1]{}, Emeric Thibaud\footnote[2]{Colorado State University}, and Daniel Cooley\footnotemark[2]{}}
 
-\appendix
-\section{Appendices}
-\subsection{MCMC details} \label{a:mcmc}
+% \today
+% \end{center}
+
+% \appendix
+\renewcommand{\thesection}{Web Appendix \Alph{section}}
+% \renewcommand{\tablename}{Web Table}
+% \renewcommand{\figurename}{Web Figure}
+% \renewcommand{\thetable}{Web Table \arabic{table}}
+
+\section{MCMC details} \label{a:mcmc}
 The MCMC sampling for the model \ref{s:hier} is done using {\tt R} (http://www.r-project.org).
@@
 
-\subsubsection*{Spatial knot locations}
+\subsection*{Spatial knot locations}
 For each day, we update the spatial knot locations, $\bw_1, \ldots, \bw_K$, using a Metropolis-Hastings block update.
@@
 
-\subsubsection*{Spatial random effects}
+\subsection*{Spatial random effects}
 If there is no temporal dependence amongst the observations, we use a Gibbs update for $z_{tk}$, and the posterior distribution is given in \ref{a:posterior}.
@@
 
-\subsubsection*{Variance terms}
+\subsection*{Variance terms}
 When there is more than one site in a partition, then we update $\sigma^2_{tk}$ using a Metropolis-Hastings update.
@@
 
-\subsubsection*{Spatial covariance parameters}
+\subsection*{Spatial covariance parameters}
 We update the three spatial covariance parameters, $\log(\rho)$, $\log(\nu)$, $\gamma$, using a Metropolis-Hastings block update step.
@@
 
-\subsection{Posterior distributions} \label{a:posterior}
+\section{Posterior distributions} \label{a:posterior}
 
-\subsubsection*{Conditional posterior of $z_{tk} \mid \ldots $}\label{s:mvcondu}
+\subsection*{Conditional posterior of $z_{tk} \mid \ldots $}\label{s:mvcondu}
 If knots are independent over days, then the conditional posterior distribution of $|z_{tk}|$ is conjugate.
@@
 
-\subsection{Proof that $\lim_{h \rightarrow \infty} \pi(h) = 0$} \label{a:proofsamepartition}
+\section{Proof that $\lim_{h \rightarrow \infty} \pi(h) = 0$} \label{a:proofsamepartition}
 Consider a homogeneous spatial Poisson process with intensity $\mu$.
@@
 
-\subsection{Skew-$t$ distribution} \label{a:skewt}
-\subsubsection*{Univariate skew-$t$ distribution}
+\section{Skew-$t$ distribution} \label{a:skewt}
+\subsection*{Univariate skew-$t$ distribution}
 We say that $Y$ follows a univariate extended skew-$t$ distribution with location $\xi \in \calR$, scale $\omega > 0$, skew parameter $\alpha \in \calR$, and degrees of freedom $\nu$ if has distribution function
@@
 
-\subsubsection*{Extremal dependence}
+\subsection*{Extremal dependence}
 For a bivariate skew-$t$ random variable $\bY = [Y(\bs), Y(\bt)]^T$, the $\chi(h)$ statistic \citep{Padoan2011} is given by
@@
 
-\subsubsection*{Proof that $\lim_{h \rightarrow \infty} \chi(h) > 0$}
+\subsection*{Proof that $\lim_{h \rightarrow \infty} \chi(h) > 0$}
 Consider the bivariate distribution of $\bY = [Y(\bs), Y(\bt)]^T$, with $\varrho(h)$ given by (\ref{eq:matern}).
@@
 
-\subsection{Simulation study pairwise difference results} \label{a:pdiffs}
+\section{Simulation study pairwise difference results} \label{a:pdiffs}
 The following tables show the methods that have significantly different Brier scores when using a Wilcoxon-Nemenyi-McDonald-Thompson test.
@@
 % \begin{singlespace}
-% \bibliographystyle{rss}
-% \bibliography{library}
+\bibliographystyle{biom}
+\bibliography{library}
 % \end{singlespace}

### 2015-11-04 LaTeX/student-competition/skewnormal.tex
updates for student competition

--- 2015-09-09 LaTeX/skewnormal.tex
@@
 \documentclass[11pt]{article}
-\usepackage{amssymb, amsthm, amsmath}
+% \documentclass[useAMS,usenatbib,referee]{biom}
+% \usepackage{amssymb, amsthm, amsmath}
+\usepackage{amssymb, amsmath}
 \usepackage{bm}
 \usepackage{graphicx}
-\usepackage[authoryear]{natbib}
+% \usepackage[authoryear]{natbib}
+\usepackage{natbib}
 \usepackage{bm}
@@
 \usepackage{enumitem}
+% \usepackage{enumerate}
 \usepackage{setspace}
@@
 \linespread{2.0}
-\input{mycommands.sty}
+\input{../mycommands.sty}
 
-\begin{document}\linenumbers
+\begin{document} %\linenumbers
+
+% \title[A Space-time Skew-$t$ Model for Threshold Exceedances]{A space-time skew-$t$ model for threshold exceedances}
+% \author
+% {Samuel A Morris$^{1,*}$\email{samorris@ncsu.edu},
+% Brian J Reich$^{1}$,
+% Emeric Thibaud$^{2}$, and
+% Daniel Cooley$^{2}$\\
+% $^{1}$Department of Statistics, North Carolina State University, Raleigh, North Carolina, U.S.A. \\
+% $^{2}$Department of Statistics, Colorado State University, Fort Collins, Colorado, U.S.A.}
+
 \begin{center}
@@
 
+\label{firstpage}
+
 \begin{abstract}
-\singlespacing
 To assess the compliance of air quality regulations, the Environmental Protection Agency (EPA) must know if a site exceeds a pre-specified threshold.
@@
 
-{\bf Key words}: Skew-$t$, random partition, MCMC, extreme value analysis, spatio-temporal
+Keywords: Extreme value analysis, MCMC, Random partition, Skew-$t$, Spatio-temporal
 \end{abstract}
 
-\newpage
+% \maketitle
+
 \section{Introduction}\label{s:intro}
@@
   \centering
-  \includegraphics[width=0.75\linewidth]{plots/ozone-10jul-us.pdf}
+  \includegraphics[width=0.75\linewidth]{../plots/ozone-10jul-us.pdf}
   \caption{Ozone values (ppb) on July 10, 2005}
@@
   \centering
-  \includegraphics[width=0.5\linewidth]{plots/chi-h.pdf}
+  \includegraphics[width=0.5\linewidth]{../plots/chi-h.pdf}
   \caption{Extremal dependence measure $\chi(h)$, as a function of distance, $h$, for $K = 1, 3, 5$, and $10$ knots.}
@@
 We generated data from 4 different simulation designs:
-\begin{enumerate} \setlength{\itemsep}{-0.5em}
+\begin{enumerate}[leftmargin=3em]
   \item Gaussian marginal, $K=1$ knot
@@
 For each data set we fit the data using six models
-\begin{enumerate} \setlength{\itemsep}{-0.5em}
+\begin{enumerate}[leftmargin=3em]
   \item Gaussian marginal, $K=1$ knots
@@
 \begin{figure}
-  \includegraphics[width=\linewidth]{plots/bsplots-mean.pdf}
+  \includegraphics[width=\linewidth]{../plots/bsplots-mean.pdf}
   \caption{Brier scores relative to the Gaussian method for simulation study results. A ratio lower than 1 indicates that the method outperforms the Gaussian method.}
@@
   \centering
-  \includegraphics[width=0.5\linewidth]{plots/qq-res.pdf}
+  \includegraphics[width=0.5\linewidth]{../plots/qq-res.pdf}
   \caption{Q-Q plot of the residuals for a skew-$t$ distribution with 10 d.f. and $\alpha = 1$ (right)}
@@
   \centering
-  \includegraphics[width=\linewidth]{plots/daily-quantiles-ozone.pdf}
+  \includegraphics[width=\linewidth]{../plots/daily-quantiles-ozone.pdf}
   \caption{Daily quantiles for two monitoring locations near Columbus, OH (left) and daily quantiles for a monitoring location in Los Angeles, CA and Columbus, OH (right)}
@@
 The four methods used are
-\begin{enumerate}\setlength{\itemsep}{-0.5em}
+\begin{enumerate}[leftmargin=3em]
   \item Gaussian
@@
 \begin{figure}
-  \includegraphics[width=\linewidth]{plots/bs-ozone.pdf}
+  \includegraphics[width=\linewidth]{../plots/bs-ozone.pdf}
   \caption{Relative Brier scores for time-series models (left) and non-time-series models (right). Relative brier score for the max-stable model is between 1.07 and 1.15}
@@
   \centering
-  \includegraphics[width=0.8\linewidth]{plots/q99-ozone.pdf}
+  \includegraphics[width=0.8\linewidth]{../plots/q99-ozone.pdf}
   \caption{Panels (a) -- (d) give the posterior predictive $\widehat{q}(0.99)$ for the month of July under four different models, panel (e) gives the difference between $\widehat{q}(0.99)$ in panels (d) and (a), panel (f) gives the difference between $\widehat{q}(0.99)$ in panels (d) and (b).}
@@
 
+% \backmatter
+
 \section*{Acknowledgments}
@@
 
-\begin{singlespace}
+% \section*{Supplementary Materials}
+% Web Appendices A--E are available with this paper at the Biometrics website on Wiley Online Library. \vspace*{-8pt}
+
+\bibliography{library}
 \bibliographystyle{rss}
-\bibliography{library}
-\end{singlespace}
+
+\label{lastpage}
 

### 2016-01-24 LaTeX/student-competition/skewnormal.tex
cover letter for student paper competition

@@
 For Gaussian processes, $\chi(h) = 0$ regardless of the distance $h$, so they are not suitable for modeling asymptotically dependent extremes.
-Unlike the Gaussian process, the skew-$t$ process is asymptotically dependent (the explicit expression for $\chi(h)$ is given in \ref{a:skewt}).
+Unlike the Gaussian process, the skew-$t$ process is asymptotically dependent \citep{Padoan2011}.
 However, one problem with the spatial skew-$t$ process is that $\lim_{h \rightarrow \infty} \chi(h) > 0$.
@@
 \end{align}
-where $I(\cdot)$ is an indicator function, $\chi_{\text{skew-}t}(h)$ is the $\chi$ statistic for a skew-$t$ process given in (\ref{eq:chiskew-t}), $\chi_{\text{Gaus}}(h)$ is the $\chi$ statistic for a Gaussian process, and $h = ||\bs - \bt||$.
+where $I(\cdot)$ is an indicator function, $\chi_{\text{skew-}t}(h) > 0$ is the $\chi$ statistic for a skew-$t$ process given in \citet{Padoan2011}, $\chi_{\text{Gaus}}(h)$ is the $\chi$ statistic for a Gaussian process, and $h = ||\bs - \bt||$.
 Therefore, sites in different subregions are asymptotically independent because $\chi_{\text{Gaus}}(h) = 0$ for all $h$.
 Marginally, over the knots, $\chi(h) = \pi(h) \chi_{\text{skew}-t}(h)$, where $\pi(h) = \Pr(k_s = k_t)$ is the probability that two sites separated by distance $h$ are in the same partition.
-In \ref{a:proofsamepartition}, we show that $\lim_{h \rightarrow \infty} \pi(h) = 0$, implying $\lim_{h \rightarrow \infty} \chi(h) = 0$.
+It is straightforward to show that $\lim_{h \rightarrow \infty} \pi(h) = 0$, implying $\lim_{h \rightarrow \infty} \chi(h) = 0$.
 In Figure \ref{fig:chi}, we give $\chi(h)$ for $K = 1, 3, 5, 10$ partitions for a skew-$t$ distribution with $\alpha = 10$, and 3 degrees of freedom.
@@
 For Friedman tests that came back with a significant p-value, we conducted a Wilcoxon-Nemenyi-McDonald-Thompson test to see which of the methods had different results.
-The full results for the Wilcoxon-Nemenyi-McDonald-Thompson tests are given in \ref{a:pdiffs}.
+% The full results for the Wilcoxon-Nemenyi-McDonald-Thompson tests are given in \ref{a:pdiffs}.
 
@@
 % Web Appendices A--E are available with this paper at the Biometrics website on Wiley Online Library. \vspace*{-8pt}
+% \renewcommand{\thesection}{Appendix \Alph{section}}
+% \setcounter{section}{0}
+% \section{Proof that $\lim_{h \rightarrow \infty} \pi(h) = 0$} \label{a:proofsamepartition}
+% Consider a homogeneous spatial Poisson process with intensity $\mu$.
+% Define $A$ as the circle with center \hbox{$(\bs_1 + \bs_2) / 2$} and radius $h / 2$.
+% Then $\bs_1$ and $\bs_2$ are in different partitions almost surely if two or more points are in $A$.
+% Let $N(A)$ be the number of points in $A$, and let
+% \begin{align*}
+%   \mu(A) = \mu |A| = \mu \pi \left(\frac{h}{2}\right)^2 = \lambda h^2.
+% \end{align*}
+% Then
+% \begin{align*}
+%   P[N(A) \ge 2] &= 1 - P[N(A) = 0] - P[N(A) = 1]\\
+%                 &= 1 - \exp\{-\lambda h^2\} - \lambda h^2 \exp\{-\lambda h^2\} \\
+%                 &= 1 - (1 + \lambda h^2) \exp\{-\lambda h^2\}
+% \end{align*}
+% which goes to one as $h \rightarrow \infty$.
+
+% \section{Simulation study pairwise difference results} \label{a:pdiffs}
+% The following tables show the methods that have significantly different Brier scores when using a Wilcoxon-Nemenyi-McDonald-Thompson test.
+% In each column, different letters signify that the methods have significantly different Brier scores.
+% For example, there is significant evidence to suggest that method 1 and method 4 have different Brier scores at $q(0.90)$, whereas there is not significant evidence to suggest that method 1 and method 2 have different Brier scores at $q(0.90)$.
+% In each table group A represents the group with the lowest Brier scores.
+% Groups are significant with a familywise error rate of $\alpha = 0.05$.
+
+% \begin{table}[htbp]
+%   \centering
+%   \caption{Setting 1 -- Gaussian marginal, $K = 1$ knot}
+%   \label{tbl:gaussim}
+%   \begin{tabular}{|l|ccc|ccc|cccc|ccc|}
+%     \cline{2-14}
+%     \multicolumn{1}{c}{} & \multicolumn{3}{|c}{$q(0.90)$} & \multicolumn{3}{|c}{$q(0.95)$} & \multicolumn{4}{|c}{$q(0.98)$} & \multicolumn{3}{|c|}{$q(0.99)$} \\
+%     \hline
+%     Method 1 & A &   &   & A &   &   & A &   &   &   & A & B &   \\
+%     \hline
+%     Method 2 & A &   &   & A &   &   & A &   &   &   & A &   &   \\
+%     \hline
+%     Method 3 &   & B &   &   & B &   &   &   & C &   &   & B &   \\
+%     \hline
+%     Method 4 & A &   &   & A &   &   & A & B &   &   & A & B &   \\
+%     \hline
+%     Method 5 &   & B &   &   & B &   &   & B & C &   & A & B &   \\
+%     \hline
+%     Method 6 &   &   & C &   &   & C &   &   &   & D &   &   & C \\
+%     \hline
+%   \end{tabular}
+% \end{table}
+
+% % \begin{table}[htbp]
+% %   \centering
+% %   \caption{Setting 2: Symmetric-$t$ marginal, $K = 1$ knot}
+% %   \label{tbl:t1k1sim}
+% %   \begin{tabular}{|l|cc|cccc|cccc|ccc|}
+% %     \cline{2-14}
+% %     \multicolumn{1}{c}{} & \multicolumn{2}{|c}{$q(0.90)$} & \multicolumn{4}{|c}{$q(0.95)$} & \multicolumn{4}{|c}{$q(0.98)$} & \multicolumn{3}{|c|}{$q(0.99)$} \\
+% %     \hline
+% %     Method 1 &   & B &   &   &   & D &   &   &   & D &   &   & C \\
+% %     \hline
+% %     Method 2 & A &   & A &   &   &   & A &   &   &   & A &   &   \\
+% %     \hline
+% %     Method 3 &   & B &   & B &   &   &   & B & C &   &   & B &   \\
+% %     \hline
+% %     Method 4 & A &   &   &   & C &   &   & B &   &   &   & B &   \\
+% %     \hline
+% %     Method 5 &   & B &   &   &   & D &   &   & C & D &   & B & C \\
+% %     \hline
+% %   \end{tabular}
+% % \end{table}
+
+% % \begin{table}[htbp]
+% %   \centering
+% %   \caption{Setting 3: Symmetric-$t$ marginal, $K = 5$ knots}
+% %   \label{tbl:gaussim}
+% %   \begin{tabular}{|l|cc|cc|cc|cc|}
+% %     \cline{2-9}
+% %     \multicolumn{1}{c}{} & \multicolumn{2}{|c}{$q(0.90)$} & \multicolumn{2}{|c}{$q(0.95)$} & \multicolumn{2}{|c}{$q(0.98)$} & \multicolumn{2}{|c|}{$q(0.99)$} \\
+% %     \hline
+% %     Method 1 &   & B &   & B &   & B &   & B \\
+% %     \hline
+% %     Method 2 &   & B &   & B &   & B &   & B \\
+% %     \hline
+% %     Method 3 &   & B &   & B &   & B & A & B \\
+% %     \hline
+% %     Method 4 & A &   & A &   & A &   & A &   \\
+% %     \hline
+% %     Method 5 &   & B &   & B & A & B & A & B \\
+% %     \hline
+% %   \end{tabular}
+% % \end{table}
+
+% \begin{table}[htbp]
+%   \centering
+%   \caption{Setting 2 -- Skew-$t$ marginal, $K = 1$ knot}
+%   \label{tbl:st1sim}
+%   \begin{tabular}{|l|ccccc|cccc|cccc|ccc|}
+%     \cline{2-17}
+%     \multicolumn{1}{c}{} & \multicolumn{5}{|c}{$q(0.90)$} & \multicolumn{4}{|c}{$q(0.95)$} & \multicolumn{4}{|c}{$q(0.98)$} & \multicolumn{3}{|c|}{$q(0.99)$} \\
+%     \hline
+%     Method 1 &   &   & C &   &   &   & B &   &   &   & B & C &   &   & B &   \\
+%     \hline
+%     Method 2 & A &   &   &   &   & A &   &   &   & A &   &   &   & A &   &   \\
+%     \hline
+%     Method 3 &   & B & C &   &   & A & B &   &   & A & B &   &   & A & B &   \\
+%     \hline
+%     Method 4 & A & B &   &   &   &   & B &   &   &   & B &   &   & A &   &   \\
+%     \hline
+%     Method 5 &   &   &   & D &   &   &   & C &   &   &   & C &   &   & B &   \\
+%     \hline
+%     Method 6 &   &   &   &   & E &   &   &   & D &   &   &   & D &   &   & C \\
+%     \hline
+%   \end{tabular}
+% \end{table}
+
+% \begin{table}[htbp]
+%   \centering
+%   \caption{Setting 3 -- Skew-$t$ marginal, $K = 5$ knots}
+%   \label{tbl:st5sim}
+%   \begin{tabular}{|l|ccc|cccc|ccc|ccc|}
+%     \cline{2-14}
+%     \multicolumn{1}{c}{} & \multicolumn{3}{|c}{$q(0.90)$} & \multicolumn{4}{|c}{$q(0.95)$} & \multicolumn{3}{|c}{$q(0.98)$} & \multicolumn{3}{|c|}{$q(0.99)$} \\
+%     \hline
+%     Method 1 &   & B &   &   &   & C &   &   & B &   &   & B &   \\
+%     \hline
+%     Method 2 &   & B &   &   &   & C &   &   & B &   &   & B &   \\
+%     \hline
+%     Method 3 & A &   &   &   & B &   &   &   & B &   &   & B &   \\
+%     \hline
+%     Method 4 & A &   &   & A &   &   &   & A &   &   & A &   &   \\
+%     \hline
+%     Method 5 & A &   &   & A &   &   &   & A &   &   & A &   &   \\
+%     \hline
+%     Method 6 &   &   & C &   &   &   & D &   &   & C &   &   & C \\
+%     \hline
+%   \end{tabular}
+% \end{table}
+
+% \begin{table}[htbp]
+%   \centering
+%   \caption{Setting 4 -- Max-stable}
+%   \label{tbl:mssim}
+%   \begin{tabular}{|l|cccc|cccc|ccc|ccc|}
+%     \cline{2-15}
+%     \multicolumn{1}{c}{} & \multicolumn{4}{|c}{$q(0.90)$} & \multicolumn{4}{|c}{$q(0.95)$} & \multicolumn{3}{|c}{$q(0.98)$} & \multicolumn{3}{|c|}{$q(0.99)$} \\
+%     \hline
+%     Method 1 & A & B &   &   &   & B &   &   &   & B &   &   &   & C \\
+%     \hline
+%     Method 2 &   & B &   &   &   & B & C &   &   & B &   &   & B & C \\
+%     \hline
+%     Method 3 &   &   & C & D &   &   & C &   &   & B &   &   & B &   \\
+%     \hline
+%     Method 4 &   &   &   & D &   &   &   & D &   &   & C &   &   & C \\
+%     \hline
+%     Method 5 &   &   & C &   &   &   & C &   &   & B &   &   & B & C \\
+%     \hline
+%     Method 6 & A &   &   &   & A &   &   &   & A &   &   & A &   &   \\
+%     \hline
+%   \end{tabular}
+% \end{table}
+
+% \begin{table}[htbp]
+%   \centering
+%   \caption{Setting 5 -- Transformation below $T = q(0.80)$}
+%   \label{tbl:transsim}
+%   \begin{tabular}{|l|cccc|ccc|cccc|cccc|}
+%     \cline{2-16}
+%     \multicolumn{1}{c}{} & \multicolumn{4}{|c}{$q(0.90)$} & \multicolumn{3}{|c}{$q(0.95)$} & \multicolumn{4}{|c}{$q(0.98)$} & \multicolumn{4}{|c|}{$q(0.99)$} \\
+%     \hline
+%     Method 1 &   &   & C &   &   & B &   &   &   & C &   &   &   & C &   \\
+%     \hline
+%     Method 2 &   & B &   &   &   & B &   &   & B &   &   & A & B &   &   \\
+%     \hline
+%     Method 3 & A &   &   &   & A &   &   & A &   &   &   & A &   &   &   \\
+%     \hline
+%     Method 4 &   & B & C &   &   & B &   &   & B &   &   &   & B & C &   \\
+%     \hline
+%     Method 5 &   & B &   &   &   & B &   &   & B & C &   &   &   & C &   \\
+%     \hline
+%     Method 6 &   &   &   & D &   &   & C &   &   &   & D &   &   &   & D \\
+%     \hline
+%   \end{tabular}
+% \end{table}
 

### 2016-03-04 LaTeX/supplemental.tex
changes to enumerate to make it fit in the margins better
circles addin

@@
 Define $A$ as the circle with center \hbox{$(\bs_1 + \bs_2) / 2$} and radius $h / 2$.
-Then $\bs_1$ and $\bs_2$ are in different partitions almost surely if two or more points are in $A$.
-Let $N(A)$ be the number of points in $A$, and let
-\begin{align*}
-  \mu(A) = \mu |A| = \mu \pi \left(\frac{h}{2}\right)^2 = \lambda h^2.
-\end{align*}
+Without loss of generality, shift $A$ so that its center is at the point $(0, 0)$, and rotate $A$ so that $\bs_1$ and $bs_2$ are on the $x$ axis.
+Partition $A$ into four equal sized partitions, $A_1, A_2, A_3, A_4$ corresponding to the four quadrants of the Cartesian plane.
+The probability of
+
+%Then $\bs_1$ and $\bs_2$ are in different partitions almost surely if two or more points are in $A$.
+%Let $N(A)$ be the number of points in $A$, and let
+%\begin{align*}
+%  \mu(A) = \lambda_{PP} |A| = \lambda_{PP} \pi \left(\frac{h}{2}\right)^2 = \lambda_{PP}^* h^2.
+%\end{align*}
+
 Then

### 2016-05-16 LaTeX/skewt.tex
edits to the manuscript for revision

--- 2015-09-17 LaTeX/skewnormal.tex
@@
 \begin{keywords}
-Skew-$t$, random partition, MCMC, extreme value analysis, spatio-temporal
+Extreme value analysis, MCMC, Random partition, Skew-$t$, Spatio-temporal
 \end{keywords}
@@
 \section*{Supplementary Materials}
+Web Appendices A--E are available with this paper at the Biometrics website on Wiley Online Library. \vspace*{-8pt}
 

### 2016-05-16 LaTeX/supplemental.tex
edits to the manuscript for revision

@@
 \usepackage{xr}
-\externaldocument{skewnormal}
+\externaldocument{skewt}
 
@@
 \section{Proof that $\lim_{h \rightarrow \infty} \pi(h) = 0$} \label{a:proofsamepartition}
-Consider a homogeneous spatial Poisson process with intensity $\mu$.
-Define $A$ as the circle with center \hbox{$(\bs_1 + \bs_2) / 2$} and radius $h / 2$.
-Without loss of generality, shift $A$ so that its center is at the point $(0, 0)$, and rotate $A$ so that $\bs_1$ and $bs_2$ are on the $x$ axis.
-Partition $A$ into four equal sized partitions, $A_1, A_2, A_3, A_4$ corresponding to the four quadrants of the Cartesian plane.
-The probability of
+Let $c$ be the midpoint of $\bs_1$ and $\bs_2$.
+Define $A$ as the circle centered at $c$ with radius $h / 2$ where $h = ||\bs_1 - \bs_2||$ is the distance between sites $\bs_1$ and $\bs_2$.
+Consider a homogeneous spatial Poisson process over $A$ with intensity given by
+\begin{align*}
+  \mu(A) = \lambda_{PP} |A| = \lambda_{PP} \pi \left(\frac{h}{2}\right)^2 = \lambda_{PPA}^* h^2.
+\end{align*}
+Consider a partition of $A$ into four regions, $B_1$, $B_2$, $R_1$, $R_2$ as seen in Web Figure \ref{fig:hpp}.
+\begin{figure}
+  \includegraphics[width=\linewidth]{plots/circles}
+  \caption{Illustration of the partition of $A$.}
+  \label{fig:hpp}
+\end{figure}
+Let $N_j$ be the number of knots in $B_j, j = 1, 2$.
+Then
+\begin{align}
+  P(\bs_1 \in P_i, \bs_2 \in P_{j \neq i}) \ge P(N_1 > 0, N_2 > 0)
+\end{align}
+since knots in both $B_1$ and $B_2$ is sufficient, but not necessary, to ensure that $\bs_1$ and $\bs_2$ are in different partition sets.
+By definition of a Poission process, $N_1$ and $N_2$ are independent and thus $P(N_1 > 0, N_2 > 0) = P(N_1 > 0)^2$, and the intensity measure over $B_1$ is given by
+\begin{align}
+  \mu(B_1) &= \lambda_{PP} |B_1| = \lambda_{PP} \frac{h^2}{4} \left(\frac{2 \pi}{3} - \frac{\sqrt{3}}{2} \right) \nonumber \\
+       &= \lambda^*_{PPB1} h^2.
+\end{align}
+So,
+\begin{align}
+  P(\bs_1 \in P_i, \bs_2 \in P_{j \neq i}) >= P(N_1 > 0)^2 = [1 - P(N_1 = 0)]^2 = [1 - \exp\left(-\lambda^*_{PPB1} h^2\right)]^2
+\end{align}
+which goes to 1 as $h$ goes to infinity.
 
@@
 
-Then
-\begin{align*}
-  P[N(A) \ge 2] &= 1 - P[N(A) = 0] - P[N(A) = 1]\\
-                &= 1 - \exp\{-\lambda h^2\} - \lambda h^2 \exp\{-\lambda h^2\} \\
-                &= 1 - (1 + \lambda h^2) \exp\{-\lambda h^2\}
-\end{align*}
-which goes to one as $h \rightarrow \infty$.
+%
+%
+%Then
+%\begin{align*}
+%  P[N(A) \ge 2] &= 1 - P[N(A) = 0] - P[N(A) = 1]\\
+%                &= 1 - \exp\{-\lambda h^2\} - \lambda h^2 \exp\{-\lambda h^2\} \\
+%                &= 1 - (1 + \lambda h^2) \exp\{-\lambda h^2\}
+%\end{align*}
+%which goes to one as $h \rightarrow \infty$.
 

### 2016-05-24 markdown/star-partition.R
changes to some of the manuscript preparing for revision

@@ -1,0 +1,32 @@
+library(fields)
+library(colorspace)
+s1 <- s2 <- seq(0.001, 0.999, length = 150)
+s <- as.matrix(expand.grid(s1, s2))
+
+nknots <- 8
+w1 <- runif(8, 0, 1)
+w2 <- runif(8, 0, 1)
+
+w <- cbind(w1, w2)
+
+s.star <- qnorm(s)
+w.star <- qnorm(w)
+
+d.star <- rdist(s.star, w.star)
+d <- rdist(s, w)
+
+get.min <- function(vec) {
+  return (which(vec == min(vec)))
+}
+
+part.star <- apply(d.star, 1, get.min)
+part      <- apply(d, 1, get.min)
+
+plot(s)
+plot(s, pch = 19, col = rainbow_hcl(8)[part])
+points(w, pch = 20, col = 1)
+
+par(mfrow = c(1, 2))
+plot(s.star, pch = 19, col = rainbow_hcl(8)[part])
+plot(s.star, pch = 19, col = rainbow_hcl(8)[part.star])
+

### 2016-05-25 LaTeX/skewt-rev.tex
changes to some of the manuscript preparing for revision
manuscript updates

--- 2016-05-16 LaTeX/skewt.tex
@@
 \usepackage{xr}
-\externaldocument{supplemental}
+\externaldocument{supplemental-rev}
 
@@
 
-Our model allows for inference and predictions using the full likelihood with computing on the order of Gaussian models for large space-time datasets.
-This allows us to use Bayesian methods to impute data below the threshold as well as make predictions at unobserved locations.
+Our model allows for inference and predictions using the full likelihood with computing on the order of Gaussian models.
+This allows us to use Bayesian methods to impute data below the threshold as well as to make predictions at unobserved locations.
 In a spatial setting, the multivariate skew-$t$ distribution demonstrates asymptotic dependence between observations at all sites regardless of the distance between the sites.
@@
 The computing is described in Section \ref{s:hier}.
-In Section \ref{s:simstudy}, we present a simulation study that examines the predictive capabilities of this model compared Gaussian amd max-stable methods.
+In Section \ref{s:simstudy}, we present a simulation study that examines the predictive capabilities of this model compared to Gaussian and max-stable methods.
 We then compare our method to Gaussian and max-stable methods with a data analysis of ozone measurements from 800 sites throughout the US in Section \ref{s:analysis}.
@@
 \subsection{Skew-$t$ process} \label{s:skewt}
-Let $Y(\bs)$ be the observation at spatial location $\bs$ in a spatial domain of interest $\calD \in \calR^2$.
-The spatial skew-$t$ process can be written
+% Let $Y(\bs)$ be the observation at spatial location $\bs$ in a spatial domain of interest $\calD \in \calR^2$.
+% The spatial skew-$t$ process can be written
+% \begin{align}
+%   Y(\bs) = \bX(\bs)^T \bbeta + \lambda \sigma |z| + \sigma v(\bs)
+% \end{align}
+% where $\bX(\bs)$ is a set of spatial covariates at site $\bs$, $\bbeta$ is the vector of regression parameters, $\lambda \in \calR$ is a parameter controlling skew, $z \sim N(0, 1)$, $\sigma^2 \sim \text{IG}(a, b)$ is random scale parameter, IG is the distribution function of an inverse gamma random variable, and $v(\bs)$ is a spatial Gaussian process with mean zero, variance one, and a positive definite correlation function.
+
+% For a finite collection of locations $\bs_1, \ldots, \bs_n$, denote the vector of observations $\bY = [Y(\bs_1), \ldots, Y(\bs_n)]^T$.
+% After marginalizing over both $z$ and $\sigma$,
+% \begin{align}
+%   \bY \sim \text{ST}_n(\bX \bbeta, \bOmega, \balpha, 2a), \label{eq:fullmodel}
+% \end{align}
+% that is, $\bY$ follows an $n$-dimensional skew-$t$ distribution with location $\bX \bbeta$, correlation matrix $\bOmega$, slant parameters $\balpha$ and degrees of freedom $2a$, where $\bX = [\bX(\bs_1)^T, \ldots, \bX(\bs_n)^T]$, $\bOmega = \bomega \bar{\bOmega} \bomega$, $\bomega = \text{diag}\left(\frac{1}{ \sqrt{ab}}, \ldots, \frac{1}{\sqrt{ab}}\right)$, $\bar{\bOmega} = (\bSigma + \lambda^2 \bOne \bOne^T$), $\balpha = \lambda (1 + \lambda^2 \bOne^T \bSigma^{-1} \bOne)^{-1/2} \bOne^T \bSigma^{-1}$, and $\bSigma$ is the positive definite correlation matrix of $[v(\bs_1), \ldots, v(\bs_n)]$.
+% This process is desirable because of its flexible tail that is controlled by the skewness parameter $\lambda$ and degrees of freedom $2a$.
+% Furthermore, the marginal distributions at each location also follow a univariate skew-$t$ distribution \citep{Azzalini2014}.
+Various forms of multivariate skew-normal and skew-$t$ distributions have been proposed in the literature.
+In this section we introduce a popular version \citep{Azzalini2014,Beranger2016}, and demonstrate how the connection between this parameterization and one that we propose to use.
+\citet{Azzalini2014} and \citet{Beranger2016} define a skew-normal process as
+\begin{align}\label{eq:origskewnorm}
+  \tilde{X}(\bs) = \tilde{\lambda} |z| + (1 - \tilde{\lambda}^2)^{1 / 2} v(\bs), \quad \bs \in \calD
+\end{align}
+where $\tilde{\lambda} \in (-1, 1)$, $z \sim N(0, 1)$, $v(\bs)$ is a Gaussian process with mean zero, variance one, and spatial correlation function cor$(h)$ where $h = || \bs_1 - \bs_2 || \in \calR$ is the distance between sites $\bs_1$ and $\bs_2$, and $\calD \in \calR^2$ is the spatial domain of interest.
+To extend the skew-normal process to a skew-$t$ process \citet{Azzalini2003} take
+\begin{align} \label{eq:originalskewt}
+  \tilde{Y}(\bs) = \xi + V X(\bs)
+\end{align}
+where $V^{-2} \sim$ Gamma$(\nu / 2, \nu / 2)$.
+Now, let $\xi = \bX(\bs)^\prime \bbeta$ where $\bX(\bs)^\prime$ is a set of $p$ spatial covariates at site $\bs$, $\bbeta$ is the $p$-vector of regression parameters, $\sigma^{-2} = \frac{a}{b} V^{-2}$ where $a = \nu / 2$ and $b = \frac{\nu}{2 (1 - \tilde{\lambda}^2)} > 0$, and let $\lambda = \tilde{\lambda} / \sqrt{1 - \tilde{\lambda}^2} \in \calR$.
+Then we can express (\ref{eq:originalskewt}) as
 \begin{align}
-  Y(\bs) = \bX(\bs)^T \bbeta + \lambda \sigma |z| + \sigma v(\bs)
+  Y(\bs) = \bX(\bs)^\prime \bbeta + \lambda \sigma |z| + \sigma v(\bs)
 \end{align}
-where $\bX(\bs)$ is a set of spatial covariates at site $\bs$, $\bbeta$ is the vector of regression parameters, $\lambda \in \calR$ is a parameter controlling skew, $z \sim N(0, 1)$, $\sigma^2 \sim \text{IG}(a, b)$ is random scale parameter, IG is the distribution function of an inverse gamma random variable, and $v(\bs)$ is a spatial Gaussian process with mean zero, variance one, and a positive definite correlation function.
-
-For a finite collection of locations $\bs_1, \ldots, \bs_n$, denote the vector of observations $\bY = [Y(\bs_1), \ldots, Y(\bs_n)]^T$.
-After marginalizing over both $z$ and $\sigma$,
-\begin{align}
-  \bY \sim \text{ST}_n(\bX \bbeta, \bOmega, \balpha, 2a), \label{eq:fullmodel}
-\end{align}
-that is, $\bY$ follows an $n$-dimensional skew-$t$ distribution with location $\bX \bbeta$, correlation matrix $\bOmega$, slant parameters $\balpha$ and degrees of freedom $2a$, where $\bX = [\bX(\bs_1)^T, \ldots, \bX(\bs_n)^T]$, $\bOmega = \bomega \bar{\bOmega} \bomega$, $\bomega = \text{diag}\left(\frac{1}{ \sqrt{ab}}, \ldots, \frac{1}{\sqrt{ab}}\right)$, $\bar{\bOmega} = (\bSigma + \lambda^2 \bOne \bOne^T$), $\balpha = \lambda (1 + \lambda^2 \bOne^T \bSigma^{-1} \bOne)^{-1/2} \bOne^T \bSigma^{-1}$, and $\bSigma$ is the positive definite correlation matrix of $[v(\bs_1), \ldots, v(\bs_n)]$.
-This process is desirable because of its flexible tail that is controlled by the skewness parameter $\lambda$ and degrees of freedom $2a$.
-Furthermore, the marginal distributions at each location also follow a univariate skew-$t$ distribution \citep{Azzalini2014}.
+where $\sigma^2 \sim $IG$(a, b)$.
 
@@
 \begin{align}
-  \text{cor}[v(\bs_1), v(\bs_2)] = \gamma I(\bs_1 = \bs_2) + (1 - \gamma)  \frac{ 1 }{ \Gamma(\nu) 2^{ \nu - 1}} \left( \sqrt{2\nu} \frac{ h }{ \rho } \right)^{\nu} K_{\nu} \left( \sqrt{2\nu} \frac{ h }{ \rho } \right) \label{eq:matern}
+  \text{cor}(h) = (1 - \gamma) I(h = 0) + \gamma \frac{ 1 }{ \Gamma(\nu) 2^{ \nu - 1}} \left( \sqrt{2\nu} \frac{ h }{ \rho } \right)^{\nu} K_{\nu} \left( \sqrt{2\nu} \frac{ h }{ \rho } \right) \label{eq:matern}
 \end{align}
-where $\rho$ is the spatial range, $\nu$ is the smoothness, $\gamma$ is the proportion of variance accounted for by the spatial variation, $K_\nu$ is a modified Bessel function of the second kind, and $h = || \bs_1 - \bs_2 ||$.
+where $\rho$ is the spatial range, $\nu$ is the smoothness, $\gamma$ is the proportion of variance accounted for by the spatial variation, $K_\nu$ is a modified Bessel function of the second kind, and $h = || \bs_1 - \bs_2 ||$ is the distance between sites $\bs_1$ and $\bs_2$.
 We use this parameterization for spatial correlation because the $\gamma$ parameter permits the inclusion of a nugget effect to account for non-spatial variability due to issues like measurement error.
@@
 One measure of extremal dependence is the $\chi$ statistic \citep{Coles1999}.
-For a stationary and isotropic spatial process, the $\chi$ statistic for locations $\bs$ and $\bt$ separated by distance $h = || \bs - \bt ||$ with identical marginal distributions is
+For a stationary and isotropic spatial process, the $\chi$ statistic for locations $\bs_1$ and $\bs_2$ separated by distance $h = || \bs_1 - \bs_2 ||$ with identical marginal distributions is
 \begin{align}
-  \chi(h) = \lim_{c \rightarrow c^*} \Pr[Y(\bs) > c | Y(\bt) > c]
+  \chi(h) = \lim_{c \rightarrow c^*} \Pr[Y(\bs_1) > c | Y(\bs_2) > c]
 \end{align}
 where $c^*$ is the upper limit of the support of $Y$.
+For the skew-$t$ distribution, $c^* = \infty$.
 If $\chi(h) = 0$, then observations are asymptotically independent at distance $h$.
@@
 \end{align}
+In other words, $P_1$ is composed of all sites for which the closest knot is $\bw_1$.
 All $z(\bs)$ and $\sigma(\bs)$ for sites in subregion $k$ are assigned common values
@@
 
-The partitioning model remove long-range dependence. Conditional on knots $\bw_1, \ldots, \bw_K$, the $\chi$ statistic for two sites $\bs$ and $\bt$ in partitions $k_s$ and $k_t$ respectively is
+The partitioning model remove long-range dependence. Conditional on knots $\bw_1, \ldots, \bw_K$, the $\chi$ statistic for two sites $\bs_1$ and $\bs_2$ in partitions $k_1$ and $k_2$ respectively is
 \begin{align}
-  \chi(h) &= I(k_s = k_t) \chi_{\text{skew-}t}(h) + I(k_s \neq k_t) \chi_{\text{Gaus}}(h) \nonumber \\
-         &= I(k_s = k_t) \chi_{\text{skew-}t}(h)
+  \chi(h) &= I(k_1 = k_2) \chi_{\text{skew-}t}(h) + I(k_1 \neq k_2) \chi_{\text{Gaus}}(h) \nonumber \\
+         &= I(k_1 = k_2) \chi_{\text{skew-}t}(h)
 \end{align}
@@
 Therefore, sites in different subregions are asymptotically independent because $\chi_{\text{Gaus}}(h) = 0$ for all $h$.
-Marginally, over the knots, $\chi(h) = \pi(h) \chi_{\text{skew}-t}(h)$, where $\pi(h) = \Pr(k_s = k_t)$ is the probability that two sites separated by distance $h$ are in the same partition.
+Marginally, over the knots, $\chi(h) = \pi(h) \chi_{\text{skew}-t}(h)$, where $\pi(h) = \Pr(k_1 = k_2)$ is the probability that two sites separated by distance $h$ are in the same partition.
 In \ref{a:proofsamepartition}, we show that $\lim_{h \rightarrow \infty} \pi(h) = 0$, implying $\lim_{h \rightarrow \infty} \chi(h) = 0$.
@@
 where $t \in \{1, \ldots, T\}$ denotes the day of each observation.
-Let \hbox{$\bw_{tk} = (w_{tk1}, w_{tk2})$} be a spatial knot on day $t$, and let $w_{t1}, \ldots, w_{tK}$ be a collection of spatial knots on day $t$.
+Let \hbox{$\bw_{tk} = (w_{tk1}, w_{tk2})$} be a spatial knot on day $t$, and let $\bw_{t1}, \ldots, \bw_{tK}$ be a collection of spatial knots on day $t$.
 As in Section \ref{s:part}, these knots define a daily partition $P_{t1}, \ldots, P_{tK}$, and for $\bs \in P_{tk}$,
@@
 
-We use an AR(1) time series model for $w_{tk}$, $z_{tk}$, and $\sigma_{tk}$.
+We use an AR(1) time series model for $\bw_{tk}$, $z_{tk}$, and $\sigma_{tk}$.
 The time series model must be specified after a transformation to preserve the skew-$t$ process at each time point.
@@
 Then the transformed knots $\bw^*_{tk} \in \calR^2$.
-We use a copula on $\sigma^2_t(\bs)$ to ensure that the marginal distributions of $\sigma^2_t(\bs)$ are inverse gamma.
+We use a transformation of a Gaussian random variable on $\sigma^2_t(\bs)$ to ensure that the marginal distributions of $\sigma^2_t(\bs)$ are inverse gamma.
 Let
@@
 where IG is defined as before.
-We also use a copula on $z_{t}(\bs)$ to ensure that the marginal distributions of $z_t(\bs)$ are half-normal.
+We also use a transformation of a Gaussian random variable on $z_{t}(\bs)$ to ensure that the marginal distributions of $z_t(\bs)$ are half-normal.
 Let
@@
 where HN is the distribution function of a half-normal random variable.
-The AR(1) process for each tail parameter is $\bw^*_{1k} \sim N_w(0, 1)$, $z^*_{1k} \sim N(0, \sigma^2_{1k})$, $\sigma^{2*}_{1k} \sim N(0, 1)$, and for $t > 1$ the time series is modeled as
+The AR(1) process for each tail parameter is $\bw^*_{1k} \sim N_w(0, 1)$, $z^*_{1k} \sim N(0, 1)$, $\sigma^{2*}_{1k} \sim N(0, 1)$, and for $t > 1$ the time series is modeled as
 \begin{align}
   \bw^*_{tk} | \bw^*_{t-1, k} &\sim N_2\left[\phi_w \bw^*_{t-1, k}, (1 - \phi_w^2) \right] \\
-  z^*_{tk} | z^*_{t-1, k} &\sim N \left[\phi_z z^*_{t-1, k}, \sigma^2_{tk} (1 - \phi_z^2)\right] \\
+  z^*_{tk} | z^*_{t-1, k} &\sim N \left[\phi_z z^*_{t-1, k}, (1 - \phi_z^2)\right] \\
   \sigma^{2*}_{tk} | \sigma^{2*}_{t-1, k} &\sim N \left[\phi_\sigma \sigma^{2*}_{t-1, k}, (1 - \phi_\sigma^2) \right]
@@
 where $|\phi_w|$, $|\phi_z|$, $|\phi_\sigma| < 1$.
-These are stationary time series models with marginal distributions \hbox{$\bw^*_{k} \sim N_2(0, 1)$}, \hbox{$z^*_{k} \sim N(0, \sigma^2_{k})$}, and \hbox{$\sigma^{2*}_{k} \sim N(0, 1)$}.
-After transformation back to the original space, $\bw_{tk} \sim \text{Unif}(\calD)$, $z_{tk} \sim HN(0, \sigma^2_{tk})$ $\sigma^2_{tk} \sim \text{IG}(a, b)$.
+These are stationary time series models with marginal distributions \hbox{$\bw^*_{k} \sim N_2(0, 1)$}, \hbox{$z^*_{k} \sim N(0, 1)$}, and \hbox{$\sigma^{2*}_{k} \sim N(0, 1)$}.
+After transformation back to the original space, $\bw_{tk} \sim \text{Unif}(\calD)$, $z_{tk} \sim HN(0, 1)$. and $\sigma^2_{tk} \sim \text{IG}(a, b)$.
 For each day, the model is identical to the spatial-only model in (\ref{eq:partition}) by construction.
@@
 \section{Hierarchical model}\label{s:hier}
-Conditioned on $z_{tk}(\bs)$, $\sigma^2_{tk}(\bs)$, and $P_{tk}$, the marginal distributions are Gaussian and the joint distribution multivariate Gaussian.
+Conditioned on $z_{tk}(\bs)$, $\sigma^2_{tk}(\bs)$, and $P_{tk}$, the marginal distributions of $Y(\bs)$ are Gaussian and the joint distribution of $Y(\bs)$ is multivariate Gaussian.
 However, we do not fix the partitions, they are treated as unknown and updated in the MCMC algorithm.
@@
 \begin{align}
-   Y_t(\bs) \mid z_{t}(\bs), \sigma_t^2(\bs), P_{tk}, \Theta &= \bX_t(\bs)^T \beta + \lambda |z_t(\bs)| + \sigma_t(\bs) v_t(\bs) \label{eq:hier}\\
+   Y_t(\bs) \mid z_{t}(\bs), \sigma_t^2(\bs), P_{tk}, \Theta &= \bX_t(\bs)^T \beta + \lambda \sigma_t(\bs) |z_t(\bs)| + \sigma_t(\bs) v_t(\bs) \label{eq:hier}\\
    z_t(\bs) &= z_{tk} \text{ if } \bs \in P_{tk}\nonumber\\
    \sigma^2_{t}(\bs) &= \sigma^2_{tk} \text{ if } \bs \in P_{tk}\nonumber\\
-   \lambda &= \lambda_1 \lambda_2\nonumber\\
-   \lambda_1 &= \left\{ \begin{array}{ll}
-      +1 \quad & \text{w.p. } 0.5\\
-      -1 \quad & \text{w.p. } 0.5
-   \end{array}\right.\nonumber\\
-   \lambda^2_2 & \sim IG(a, b)\nonumber\\
+   \lambda & \sim N(0, 20)\nonumber\\
+   % \lambda_1 &= \left\{ \begin{array}{ll}
+   %    +1 \quad & \text{w.p. } 0.5\\
+   %    -1 \quad & \text{w.p. } 0.5
+   % \end{array}\right.\nonumber\\
+   % \lambda^2_2 & \sim IG(a, b)\nonumber\\
    v_t(\bs) \mid \Theta &\sim \Matern(0, \Sigma)\nonumber\\
-   z^*_{tk} \mid z^*_{t-1, k}, \sigma^2_{tk} &\sim N(\phi_z z^*_{t-1, k}, \sigma^2_{tk} (1 - \phi_z^2))\nonumber\\
+   z^*_{tk} \mid z^*_{t-1, k} &\sim N(\phi_z z^*_{t-1, k}, (1 - \phi_z^2))\nonumber\\
    \sigma^{2*}_{tk} \mid \sigma^{2*}_{t-1, k} &\sim N(\phi_\sigma \sigma^{2*}_{t-1, k}, (1 - \phi_\sigma^2))\nonumber\\
@@
 where $\Theta = \{\rho, \nu, \gamma\, \lambda, \beta\}$, and $\Sigma$ is a \Matern covariance matrix as described in Section \ref{s:skewt}.
-We parameterize $\lambda = \lambda_1 \lambda_2$ to help with convergence in the MCMC.
 
@@
   % \item Transformation below $T = q(0.80)$
+  \item A Brown-Resnick process with range 1 and smoothness 0.5
 \end{enumerate}
@@
 We analyzed the results for the simulation study using a Friedman test at $\alpha = 0.05$ to see if at least one method had a significantly different Brier score.
-For Friedman tests that came back with a significant p-value, we conducted a Wilcoxon-Nemenyi-McDonald-Thompson test to see which of the methods had different results.
+For Friedman tests that came back with a significant p-value, we conducted a Wilcoxon-Nemenyi-McDonald-Thompson \citep{Hollander2014} test to see which of the methods had different results.
 The full results for the Wilcoxon-Nemenyi-McDonald-Thompson tests are given in \ref{a:pdiffs}.
@@
 For each site, we also have covariate information containing the estimated ozone from the Community Multi-scale Air Quality (CMAQ) modeling system.
-Initially, we fit a linear regression assuming a mean function of
-\begin{align}
-  \text{E}[Y_i(\bs)] = \beta_0 + \beta_1 \cdot \text{CMAQ}_t(\bs). \label{eq:datamean}
-\end{align}
+Initially, we fit a linear regression with $\bX_t(\bs) = [1, \text{CMAQ}_t(\bs)]^\prime$. % assuming a mean function of
+% \begin{align}
+%   \text{E}[Y_i(\bs)] = \beta_0 + \beta_1 \cdot \text{CMAQ}_t(\bs). \label{eq:datamean}
+% \end{align}
 Figure \ref{fig:ozone-qq} shows a Q-Q plot of the residuals compared to a skew-$t$ distribution with 10 d.f. and $\alpha = 1$.

### 2016-05-27 LaTeX/response.tex
updates to manuscript and some small bugfixes in code

\documentclass[11pt]{article}
\usepackage[letterpaper, margin=1in]{geometry}
% \documentclass[useAMS,usenatbib,referee]{biomweb}
\usepackage{amssymb, amsthm, amsmath}
\usepackage{amssymb, amsmath}
\usepackage{bm}
\usepackage{graphicx}
\usepackage[authoryear]{natbib}
\usepackage{bm}
\usepackage{verbatim}
\usepackage{lineno}
\usepackage{soul}
\usepackage{color}
% \usepackage{enumitem}
\usepackage{enumerate}
\usepackage{setspace}
\usepackage{appendix}
% \usepackage{caption}
% \captionsetup[table]{name=Web Table}

% cross-referencing appendices
\usepackage{xr}
\externaldocument{skewt_rev}

% \usepackage[left=1in,top=1in,right=1in]{geometry}
% \pdfpageheight 11in
% \pdfpagewidth 8.5in
% \linespread{2.0}
\input{mycommands.sty}
\begin{document}

\begin{center}
Point-by-Point response to Reviewer Comments\\
Biometrics MS \#150926M\\
\today
\end{center}

We wish to thank the associate editor and two reviewers for their thorough reviews and constructive suggestions.
To address the concerns raised in this review we have carefully edited the manuscript to clarify issues regarding the skew-$t$ parameterization and other notational issues.
We have also added some new simulation results to explore more complicated max-stable processes, and a discussion of the properties of the spatiotemporal extension of our model.
We hope you agree that these changes have improved the manuscript.

\section*{Response to the comments of the Associate Editor}
\subsection*{Major comments}
\begin{enumerate}[1.]
\item The model in (1) has a non-standard skew distribution notation. As far as I could check, the parameter $\sigma^2$ should have an inverse gamma with parameters $a = b$, and not arbitrary and different $a$ and $b$. The equality is important in the derivations and it is one of the reasons we get $2a$ degrees of freedom. Indeed, some reserachers adopt $a/2$ and $a/2$ (rather than $a$ and $b$) such that they end up with a degrees of freedom. Can you provide a web-supplement proof that your more general model with $a \neq b$ is a skew-$t$ distribution with $2a$ degrees of freedom, or provide a reference for this result?

\item In the simulations, you adopted a prior distribution for the $a$ and $b$ parameters in $\sigma_t^2 (\bs) \sim IG(a, b)$ as $a \sim $Gamma$(0.1, 0.1)$ and $b$ with a discrete distribution over a mesh from 0.1 to 10 with spacing $0.1$. The parameter $\sigma_t^2(\bs)$ is clearly a crucial parameter as it controls the scale. Why this choice? Trial and error? What have you used in the real dataset? Is this supposed to be a default choice?

\item A very confusing issue for me was the adoption of the new parametrization $\lambda = \lambda_1 \lambda_2$. Typically, users adopt a prior such as $\lambda \sim N(0,V)$ with a large and known value for $V$. One consequence is that the posterior distribution of $\lambda$ has non negligible probability mass around 0, making the skew model irrelevant. When you adopted the parametrization $\lambda = \lambda_1 \lambda_2$, you ``solved'' this issue by making zero an impossible value under the prior. The parameter $\lambda_1$ is equal to $−1$ or $+1$ and it basically selects the semi-axis where the slant parameter will live. The parameter $\lambda_2 \sim IG(a, b)$ and hence $\lambda_2 > 0$ and bounded away from zero, in practice. The value $\lambda = 0$ is not supported by the prior. Under your parametrization, you are artificially avoiding your model to become a simple gaussian process.

What, exactly, is the convergence problem you had with the λ parameter before adopting this reparametrization?
It affected only λ?

Incidentally, you used the same notation $IG(a,b)$ for two different purposes: for $\lambda_2$ in page 10, and for $\sigma^2$ in page 4. Please, keep them different as there are a lot of things going on in your model.

\item In page 6, section 3.1, you say that you ``impute the censored values as a step in the MCMC algorithm''. Why do you need these pseudo-values, specially considering that you do have the real below threshold observations? Is it because you need them follwoing to your tail model (the skew-$t$ model)? But why do you need all the uncensored pseudo data?

Anyway, you should make clear that trading a focus only on the extreme tail data you loose efficiency by decreasing the number of (real) data available. Censoring on purpose seems a rather drastic technique.

\item The temporal dependence structure is induced by putting AR(1) models on some of the skew-$t$ parameters.
It is not clear what sort of temporal dependence that induces on the data.
\end{enumerate}

\subsection*{Minor comments}
\begin{enumerate}[1.]
  \item Page 3, line 11: ``\ldots with computing on the order of Gaussian models ''

  \item Page 3, line 13: \ldots as well as TO make predictions \ldots

  \item Page 3, line -2: \ldots compared TO Gaussian AND max-stable \ldots

  \item Page 8, line -1: a partition in the $\bw^∗$ space induces a partition in the $\bw^∗$. The property (7) is used in the $\bw^∗$ space. Is this property still respected in the $\bw$ space since the transformation is non-linear and acts independently in each coordinate?

  \item Page 9, line -5: ``the marginal distributions are Gaussian'': which distributions, those of $Y(\bs)$?

  \item The hierarchical model in page 10 has $\sigma_t(\bs)$ multiplying only $v_t(\bs)$ and leaving out the term $|z_t(\bs)|$.
  I think this is a typo.

  \item In page 12, line -1: you mention the partition label switching problem appearing during the MCMC. This was not consequential for the inference? Can you expand on this?

  \item There are several papers using skew distributions in the spatial context. You should refer to the most relevant ones in your related work section.

\end{enumerate}

\newpage
\section*{Response to the comments of the Reviewer 1}

\subsection*{Major comments}
\begin{enumerate}[1.]
  \item The overall modeling goal is never fully clear. In the abstract you write ``We present a new method based on the spatial skew-t process.'' But a new method for what? Reading the paper there are hints that you the aim of the modeling procedure might be to predict the probability of exceeding 75ppb (e.g. first line of the abstract), but this is then taken as one of your modeling thresholds (p.3 l.5--6; p. 15 l.5). In a new version of the paper, the modeling aims should be stated clearly upfront so that the results can be assessed agains these aims.

  \item There is a lot of confusion about the ``threshold'', and what this means in various contexts. Sometimes you discuss this as fixed and specified by regulation (e.g. p.3 l.5) but other times you refer to the role of censoring as not allowing low-moderate values influence the fit (e.g. Section 3.1 or the stated goal of modeling ``spatial extremes'' in the first line of Section 3) even though you state that the constant threshold is not extreme everywhere (e.g. in the Abstract, and p.15 l.3 where in fact the highest threshold you use in the application is the 0.06 quantile at some sites, and so hardly the tail!). The role of the threshold needs to be much more clearly explained in the context of the application (fixed thresholds of interest), and modeling the extremes.

  \item Section 3.2: you introduce a Poisson process as a device for allowing long range asymptotic independence. In fact, after its first mention on p.7 l.1 you never refer to a Poisson process again, but instead use a fixed number of knots, i.e. a uniform distribution. Amazingly this is not even commented on! The purpose of the Poisson process seems to be for the proof in Web Appendix C, so if spatial long-range independence relies on this feature (which, assuming a constant rate μ would eventually guarantee an incease in the number of points, as opposed to taking a fixed number of draws from a uniform distribution) then the discrepancies between these two assumptions ought to be thoroughly discussed.

  The proof in Web Appendix C also needs some greater attention to detail. It looks to me as if using defintion (7) and for a fixed radius $h$ then $\bs_1$ and $\bs_2$ need not be in different partitions almost surely, although this may be the case as $h \rightarrow \infty$. Perhaps I am wrong, but either way it would be helpful to have the full proof spelt out to avoid any confusion.

  \item Section 3.3: As you note, there are several places where you could introduce temporal dependence. Are you able to analyze any theoretical properties of the suggested form, or even how this matches the data?

  \item Section 5: The simulation study needs more details. For example, when fitting datasets (1)--(4) to model (6) (max-stable), were the margins of this estimated or assumed fixed? Also you fit trend terms in some models when the mean is constant, and do not comment on the rationale for this or what happens to the estimates of these parameters (are $\beta_\text{lat}$ and $\beta_\text{long}$ close to 0 and $\beta_\text{int}$ close to 10?)

  Furthermore the range of models simulated from is rather limited; for (4) you could simulate from a range of different max-stable processes (even if you don’t fit these, it might be interesting to see how the skew-t model handles them).

  It would be nice to see alternative model evaluations to the Brier score, or at least perhaps gain some sense of whether there is strong spatial variation in the score. If the model is much better at predicting in some locations than others, then one may want to be wary of using it in some areas\ldots

  \item Section 6: why do you only use July data? Is this a computational restriction, due to availability of data, or due to particular interest in that month? As you point out, if you used more data then you would be able to say more about the extremal dependence, so this point is rather pertinent. Indeed, although Figure 6 shows that two particular sites close together are relatively strongly dependent, we see nothing about the general picture of $\chi(h)$. I agree this would be difficult to assess with so few data, but it’s not clear whether you have more data available to help assess this (even if you didn’t fit the model to all of them).
\end{enumerate}

\subsection*{Minor comments}
\begin{enumerate}[1.]
  \item p.5l.6 $\gamma \rightarrow 1 - \gamma$.

  \item p.8 l.8 should $w_{t,1}$ etc. be bold?

  \item p. 8 final line: the phrase ``use a copula'' could be omitted as it seems superfluous and may introduce unnecessary confusion. Giving adequate details on the transformation and dependence structure is sufficient.

  \item p. 12 final line: you fit the correct model in some of these cases, so is it still difficult to tell whether the parameters converge even in this situation?

  \item p.13 l.13 and l.15 please give references for these tests

  \item p.14 eq.(20): Do you specify the mean of $Y$, or use the covariate specification given in (1)? These don’t look to be the same if $\lambda \neq 0$.
\end{enumerate}

\newpage
\section*{Response to the comments of the Reviewer 2}

\subsection*{Major comments}
\begin{enumerate}[1.]
  \item Throughout the paper I see an inappropriate use of terminologies which gives an idea that the paper is not carefully written. Examples are:
  \begin{enumerate}[i.]
    \item Page 4 at the beginning of Section 2.1 you write that $Y(s)$ is an observation at the spatial location $s$. Then you use the same symbol to define a random field (spatial random process) in equation (1). I think you mean that $y(si), i = 1, 2, \ldots$ is an observation while $Y(s; ω) \equiv Y (s)$ is a random function for $s$ varying on a certain set. A distinction between observations and random objects which represent the phenomena that you are interested on, must be done.

    \item Page 5, equation (3). You used the $h = ||\bs_1 − \bs_2||$ (above equation (4)) to denote the distance between two specific fixed points and then $h = ||\bs−\bt||$ to denote the distance between two generic locations. The latter (more general) should also be used in equation (3) to avoid confusion;

    \item Page 7, below equation (7). The way you defined the partition is not clear to me. Suppose that we consider two spatial knots. Then, according to your definitions we have the two partitions:
    \begin{align*}
      P_k = \{s \in \calD : k = \argmin_i ||s−w_i||\}, \quad \text{with} \quad k= 1,2.
    \end{align*}
    The index of the partition is also used to denote the value of the radius. Does this mean that we also consider all the values $s ∈ \calD$ such that for a given $w_k$ we have that $\min ||s − w_k|| \le k$? In this simple example, is $\calD$ formed by two disjoint sets?

    \item Second, you said: ``all $z(s)$ and $\sigma(s)$ for sites in the sub-region k are assigned common values z(s) = zk and σ(s) = σk''. I guess you mean that for all $s \in P_k$, with $k = 1, 2, \ldots, K$, the functions $z(s)$ and $σ(s)$ are equal to the constants $z_k$ and $\sigma_k$, respectively.

    \item Finally, from the process definition in (6) $Y(s)$, the partitions in (7) and the elicitation of the random function in (8), why should we come to the conclusion that for locations that belong to different partitions, e.g. $s_i \in P_k$ and $s_j \in P_l$ with $P_k \neq P_l$, their distribution is not skew-$t$ anymore? You have only specified that within the same partition the distribution is skew-$t$, but what happens in the contrary case? In addition, from (9) it seems that for locations that belong to different partitions the distribution of the process is Gaussian. Is this correct? A clear explanation of this part is needed, adding the required steps.

  \end{enumerate}

  \item Noting differences in notation between Beranger et. al (2015), Azzalini (2013), and Azzalini and Capitanio (2003), the reviewer asks
  \begin{enumerate}[i.]
    \item Regardless of the different parametrization, i.e. you used $\sigma^2 \sim IG(a, b)$ instead of $V ∼ G(\nu/2, \nu/2)$, the formula should be equivalent. It seems that there is a missing term in your formula, if not you should explain why there is such a difference and refer to where you got that expression. My suggestion is that you should explain well, with the two steps (1) and (2) above, the process construction;

    \item Second, in your equation (1), $Y(s)$ is a stochastic process, a random function defined on $\calD$. Mathematically speaking, what is $\bX(s)^\top$? You only say that it is ``a set of spatial covariates at site $s$'', but this is not enough. I guess that $\bX(s)^\top$ should be a vector-valued or a matrix-valued function (a vector or matrix of functions). But if so, how should the product with the vector $\beta$ be interpreted?

  \end{enumerate}

  \item The section on the spatio-temporal extension is hard to follow and the technical details are not immediate to verify, in the present form. The presentation should be simplified. Is it not much easier that you define first the AR(1) time series and then you define appropriate transformations that link these to the functional parameters of the skew-$t$ process and you clearly show that the finite dimensional distribution of the resulting process including time and space is skew-$t$?

  \item Section 5 Simulation study. The current session should be shortened. Furthermore the section should be divided into two parts. Before showing the performances of your method in predicting exceedances, you should show by means of a simulation study the ability of your approach when estimating the model parameters. Since that you are considering also replications over time it should be possible to estimate reasonably well the slant parameters and the degrees of freedom. It may be the case that you have to consider larger samples. Have you tried to consider alternative parameterizations? See e.g. Arellano-Valle and Azzalini (2008). You should investigate more seriously this aspect.

\end{enumerate}

\subsection*{Minor comments}
\begin{enumerate}[1.]
  \item Page 4, first paragraph of Section 2. The references Beranger et al. (2015), Azzalini (2013) page 129 and Azzalini and Capitanio (2003), should be added for a clearer explanation of the additive representation;

  \item Page 4, equation (1). You used the symbol ``$T$'' at least 3 times, e.g. in order to denote: transposed operation, a threshold (at page 6), a time index (at page 8). Different symbols should be used to avoid confusion;

  \item Page 4 below equation (1). Specify what the size of $\bX(s)^\top$ and of $\beta$ are. Are you sure that $\lambda \in \mathbb{R}$? Is it actually $\lambda \in (−1,1)$?;

  \item Page 4 below equation (1). A symbol for the correlation function of the process $v(s)$ should be used, e.g $\rho(h)$, in order to justify the presence of the matrix $\Sigma$ which is obtained from $\rho(h)$;

  \item Page 4 below equation (2). The size of the vector and matrices should be specified. Furthermore, specify that matrix $\Sigma$ is obtained from $\rho(h)$ and this part should be jointed to equation (3) for brevity. Finally $\Omega$ is the dispersion matrix of $Y$;

  \item Page 5, equation (3). What does $|| \cdot ||$ mean?

  \item Page 5, equation (4). Since you are specifically referring to the skew-$t$ process, why don’t you simply say that $c^∗ = +\infty$? You should also confirm once more that for all given $\bs \in \calD$, $Y(\bs)$ is a skew-$t$ random variable, to avoid confusion;

  \item Page 8, two rows below equation (10). The elements of the sequence $w_{t1}, \ldots , w_{tK}$ should be bolded.

  \item Page 11, Section 5, 8 lines from the bottom. You wrote ``the $v_t(s)$ terms were generated using\ldots ''. This is incorrect, $v_t(s)$ is not an observation. According to (10), $v_t(s)$ is a stochastic process.

  \item Appendix C, you use $\mu(\cdot)$ to denote the intensity measure of the Poisson point process but then $\mu$ is also a constant in the definition of the intensity measure. You should use different symbols.
\end{enumerate}

\end{document}

### 2016-05-27 LaTeX/skewt_rev.tex
updates to manuscript and some small bugfixes in code

--- 2016-05-25 LaTeX/skewt-rev.tex
@@
 \usepackage{xr}
-\externaldocument{supplemental-rev}
+\externaldocument{supplemental_rev}
 
@@
 \subsection{Skew-$t$ process} \label{s:skewt}
-% Let $Y(\bs)$ be the observation at spatial location $\bs$ in a spatial domain of interest $\calD \in \calR^2$.
-% The spatial skew-$t$ process can be written
-% \begin{align}
-%   Y(\bs) = \bX(\bs)^T \bbeta + \lambda \sigma |z| + \sigma v(\bs)
-% \end{align}
-% where $\bX(\bs)$ is a set of spatial covariates at site $\bs$, $\bbeta$ is the vector of regression parameters, $\lambda \in \calR$ is a parameter controlling skew, $z \sim N(0, 1)$, $\sigma^2 \sim \text{IG}(a, b)$ is random scale parameter, IG is the distribution function of an inverse gamma random variable, and $v(\bs)$ is a spatial Gaussian process with mean zero, variance one, and a positive definite correlation function.
-
-% For a finite collection of locations $\bs_1, \ldots, \bs_n$, denote the vector of observations $\bY = [Y(\bs_1), \ldots, Y(\bs_n)]^T$.
-% After marginalizing over both $z$ and $\sigma$,
-% \begin{align}
-%   \bY \sim \text{ST}_n(\bX \bbeta, \bOmega, \balpha, 2a), \label{eq:fullmodel}
-% \end{align}
-% that is, $\bY$ follows an $n$-dimensional skew-$t$ distribution with location $\bX \bbeta$, correlation matrix $\bOmega$, slant parameters $\balpha$ and degrees of freedom $2a$, where $\bX = [\bX(\bs_1)^T, \ldots, \bX(\bs_n)^T]$, $\bOmega = \bomega \bar{\bOmega} \bomega$, $\bomega = \text{diag}\left(\frac{1}{ \sqrt{ab}}, \ldots, \frac{1}{\sqrt{ab}}\right)$, $\bar{\bOmega} = (\bSigma + \lambda^2 \bOne \bOne^T$), $\balpha = \lambda (1 + \lambda^2 \bOne^T \bSigma^{-1} \bOne)^{-1/2} \bOne^T \bSigma^{-1}$, and $\bSigma$ is the positive definite correlation matrix of $[v(\bs_1), \ldots, v(\bs_n)]$.
-% This process is desirable because of its flexible tail that is controlled by the skewness parameter $\lambda$ and degrees of freedom $2a$.
-% Furthermore, the marginal distributions at each location also follow a univariate skew-$t$ distribution \citep{Azzalini2014}.
-Various forms of multivariate skew-normal and skew-$t$ distributions have been proposed in the literature.
-In this section we introduce a popular version \citep{Azzalini2014,Beranger2016}, and demonstrate how the connection between this parameterization and one that we propose to use.
-\citet{Azzalini2014} and \citet{Beranger2016} define a skew-normal process as
-\begin{align}\label{eq:origskewnorm}
-  \tilde{X}(\bs) = \tilde{\lambda} |z| + (1 - \tilde{\lambda}^2)^{1 / 2} v(\bs), \quad \bs \in \calD
+Let $Y(\bs)$ be the observation at spatial location $\bs$ in a spatial domain of interest $\calD \in \calR^2$.
+The spatial skew-$t$ process can be written
+\begin{align}
+  Y(\bs) = \bX(\bs)^\prime \bbeta + \lambda \sigma |z| + \sigma v(\bs) \label{eq:fullmodel}
 \end{align}
-where $\tilde{\lambda} \in (-1, 1)$, $z \sim N(0, 1)$, $v(\bs)$ is a Gaussian process with mean zero, variance one, and spatial correlation function cor$(h)$ where $h = || \bs_1 - \bs_2 || \in \calR$ is the distance between sites $\bs_1$ and $\bs_2$, and $\calD \in \calR^2$ is the spatial domain of interest.
-To extend the skew-normal process to a skew-$t$ process \citet{Azzalini2003} take
-\begin{align} \label{eq:originalskewt}
-  \tilde{Y}(\bs) = \xi + V X(\bs)
-\end{align}
-where $V^{-2} \sim$ Gamma$(\nu / 2, \nu / 2)$.
-Now, let $\xi = \bX(\bs)^\prime \bbeta$ where $\bX(\bs)^\prime$ is a set of $p$ spatial covariates at site $\bs$, $\bbeta$ is the $p$-vector of regression parameters, $\sigma^{-2} = \frac{a}{b} V^{-2}$ where $a = \nu / 2$ and $b = \frac{\nu}{2 (1 - \tilde{\lambda}^2)} > 0$, and let $\lambda = \tilde{\lambda} / \sqrt{1 - \tilde{\lambda}^2} \in \calR$.
-Then we can express (\ref{eq:originalskewt}) as
-\begin{align}
-  Y(\bs) = \bX(\bs)^\prime \bbeta + \lambda \sigma |z| + \sigma v(\bs)
-\end{align}
-where $\sigma^2 \sim $IG$(a, b)$.
-
+where $\bX(\bs)$ is a set of spatial covariates at site $\bs$, $\bbeta$ is the vector of regression parameters, $\lambda \in \calR$ is a parameter controlling skew, $z \sim N(0, b)$, $\sigma^2 \sim \text{IG}(a / 2, a / 2)$ is random scale parameter, IG is the distribution function of an inverse gamma random variable, and $v(\bs)$ is a spatial Gaussian process with mean zero, variance $b^2$, and a positive definite correlation function $\rho(\bs_1, \bs_2)$.
 Although any positive definite correlation function could be used, we choose to use the stationary isotropic \Matern correlation with
 \begin{align}
-  \text{cor}(h) = (1 - \gamma) I(h = 0) + \gamma \frac{ 1 }{ \Gamma(\nu) 2^{ \nu - 1}} \left( \sqrt{2\nu} \frac{ h }{ \rho } \right)^{\nu} K_{\nu} \left( \sqrt{2\nu} \frac{ h }{ \rho } \right) \label{eq:matern}
+  \rho(h) = (1 - \gamma) I(h = 0) + \gamma \frac{ 1 }{ \Gamma(\nu) 2^{ \nu - 1}} \left( \sqrt{2\nu} \frac{ h }{ \varphi } \right)^{\nu} K_{\nu} \left( \sqrt{2\nu} \frac{ h }{ \varphi } \right) \label{eq:matern}
 \end{align}
-where $\rho$ is the spatial range, $\nu$ is the smoothness, $\gamma$ is the proportion of variance accounted for by the spatial variation, $K_\nu$ is a modified Bessel function of the second kind, and $h = || \bs_1 - \bs_2 ||$ is the distance between sites $\bs_1$ and $\bs_2$.
-We use this parameterization for spatial correlation because the $\gamma$ parameter permits the inclusion of a nugget effect to account for non-spatial variability due to issues like measurement error.
+where $\varphi$ is the spatial range, $\nu$ is the smoothness, $\gamma$ is the proportion of variance accounted for by the spatial variation, $K_\nu$ is a modified Bessel function of the second kind, and $h = || \bs_1 - \bs_2 ||$ is the distance between sites $\bs_1$ and $\bs_2$.
+
+For a finite collection of locations $\bs_1, \ldots, \bs_n$, denote the vector of observations $\bY = [Y(\bs_1), \ldots, Y(\bs_n)]^\prime$.
+After marginalizing over both $z$ and $\sigma$,
+\begin{align}
+  \bY \sim \text{ST}_n(\bX \bbeta, \bOmega, \balpha, 2a),
+\end{align}
+that is, $\bY$ follows an $n$-dimensional skew-$t$ distribution with location $\bX \bbeta$, variance matrix $\bOmega$, slant parameters $\balpha$ and degrees of freedom $2a$, where $\bX = [\bX(\bs_1)^\prime, \ldots, \bX(\bs_n)^\prime]$, $\bOmega = \bomega \bar{\bOmega} \bomega$, $\bomega = \text{diag}\left(\frac{1}{ \sqrt{ab}}, \ldots, \frac{1}{\sqrt{ab}}\right)$, $\bar{\bOmega} = (\bSigma + \lambda^2 \bOne \bOne^\prime$), $\balpha = \lambda (1 + \lambda^2 \bOne^\prime \bSigma^{-1} \bOne)^{-1/2} \bOne^\prime \bSigma^{-1}$, and $\bSigma$ is the positive definite correlation matrix which is obtained from $\rho(h)$.
+This process is desirable because of its flexible tail that is controlled by the skewness parameter $\lambda$ and degrees of freedom $2a$.
+Furthermore, the marginal distributions at each location also follow a univariate skew-$t$ distribution \citep{Azzalini2014}.
+
+% We use this parameterization for spatial correlation because the $\gamma$ parameter permits the inclusion of a nugget effect to account for non-spatial variability due to issues like measurement error.
 
@@
 \begin{align} \label{eq:partition}
-  Y(\bs) &= \bX(\bs)^T \bbeta + \lambda \sigma(\bs) |z(\bs)| + \sigma(\bs) v(\bs).
+  Y(\bs) &= \bX(\bs)^\prime \bbeta + \lambda \sigma(\bs) |z(\bs)| + \sigma(\bs) v(\bs).
 \end{align}
@@
 \begin{align} \label{eq:spatiotemp}
-  Y_t(\bs) = \bX_t(\bs)^T \bbeta + \lambda \sigma_t(\bs) |z_t(\bs)| + \sigma_t(\bs) v_t(\bs),
+  Y_t(\bs) = \bX_t(\bs)^\prime \bbeta + \lambda \sigma_t(\bs) |z_t(\bs)| + \sigma_t(\bs) v_t(\bs),
 \end{align}
@@
 \begin{align}
-   Y_t(\bs) \mid z_{t}(\bs), \sigma_t^2(\bs), P_{tk}, \Theta &= \bX_t(\bs)^T \beta + \lambda \sigma_t(\bs) |z_t(\bs)| + \sigma_t(\bs) v_t(\bs) \label{eq:hier}\\
+   Y_t(\bs) \mid z_{t}(\bs), \sigma_t^2(\bs), P_{tk}, \Theta &= \bX_t(\bs)^\prime \beta + \lambda \sigma_t(\bs) |z_t(\bs)| + \sigma_t(\bs) v_t(\bs) \label{eq:hier}\\
    z_t(\bs) &= z_{tk} \text{ if } \bs \in P_{tk}\nonumber\\
@@
 \end{align}
-where $\Theta = \{\rho, \nu, \gamma\, \lambda, \beta\}$, and $\Sigma$ is a \Matern covariance matrix as described in Section \ref{s:skewt}.
+where $\Theta = \{\varphi, \nu, \gamma\, \lambda, \beta\}$, and $\Sigma$ is a \Matern covariance matrix as described in Section \ref{s:skewt}.
 
@@
 After conditioning on $\lambda$, $z_t(\bs)$ and non-censored observations, $Y_t(\bs)$ has truncated normal full conditionals.
-So we sample $Y_t(\bs) \sim N_{(-\infty, T)}(\bX_t^T(\bs) \beta + \lambda | z_t(\bs)|, \bSigma)$.
+So we sample $Y_t(\bs) \sim N_{(-\infty, T)}(\bX_t^\prime(\bs) \beta + \lambda | z_t(\bs)|, \bSigma)$.
 
@@
 \end{enumerate}
-In the first three designs, the $v_t(\bs)$ terms were generated using a \Matern covariance with smoothness parameter $\nu = 0.5$ and spatial range $\rho = 1$.
+In the first three designs, the $v_t(\bs)$ terms were generated using a \Matern covariance with smoothness parameter $\nu = 0.5$ and spatial range $\varphi = 1$.
 For the covariance matrices in designs 1 -- 3, the proportion of the variance accounted for by the spatial variation was $\gamma = 0.9$ while the proportion of the variance accounted for by the nugget effect was $0.1$.
@@
 For designs 2 and 3, $\lambda = 3$ was used as to simulate moderate skewness, and the $z_t$ were generated as described in (\ref{eq:sitezsig}).
-In designs 1 -- 3, the mean $\bX^T \bbeta = 10$ was assumed to be constant across space.
+In designs 1 -- 3, the mean $\bX^\prime \bbeta = 10$ was assumed to be constant across space.
 In the fourth design, we generated from a spatial max-stable distribution \citep{Reich2012}.
@@
 The design matrix $\bX$ includes an intercept with a first-order spatial trend with priors of $\beta_\text{int}$, $\beta_\text{lat}$, $\beta_\text{long},  \iid \text{N}(0, 10)$.
-The spatial covariance parameters have priors $\log(\nu) \sim \text{N}(-1.2, 1)$, $\gamma \sim \text{Unif}(0, 1)$, $\rho \sim \text{Unif}(15)$.
+The spatial covariance parameters have priors $\log(\nu) \sim \text{N}(-1.2, 1)$, $\gamma \sim \text{Unif}(0, 1)$, $\varphi \sim \text{Unif}(15)$.
 The skewness parameter has prior $\lambda_2 \sim \text{IG}(0.1, 0.1)$.

### 2016-05-27 LaTeX/supplemental_rev.tex
updates to manuscript and some small bugfixes in code

--- 2016-05-16 LaTeX/supplemental.tex
@@
 \usepackage{xr}
-\externaldocument{skewt}
+\externaldocument{skewt_rev}
 
@@
 \section{MCMC details} \label{a:mcmc}
-The MCMC sampling for the model \ref{s:hier} is done using {\tt R} (http://www.r-project.org).
+The MCMC sampling for the model in Section \ref{s:hier} is done using {\tt R} (http://www.r-project.org).
 Whenever possible, we select conjugate priors (see Appendix \ref{a:posterior}); however, for some of the parameters, no conjugate prior distributions exist.
@@
 
+\section{Comparisons with other parameterizations} \label{a:otherparams}
+Various forms of multivariate skew-normal and skew-$t$ distributions have been proposed in the literature.
+In this section, we make a connection between our parameterization in (\ref{eq:fullmodel}) of the main text and another popular version.
+\citet{Azzalini2014} and \citet{Beranger2016} define a skew-normal process as
+\begin{align}
+  \tilde{X}(\bs) = \tilde{\lambda}|z| + (1 - \tilde{\lambda}^2)^{1 / 2} v(\bs)
+\end{align}
+where $\tilde{\lambda} \in (-1, 1)$, $z ~ N(0, 1)$, $v(\bs)$ is a Gaussian process with mean zero, variance one, and spatial correlation function $\rho$.
+To extend this to the skew-$t$ distribution, \citet{Azzalini2003} take $\tilde{Y}(\bs) = V\tilde{X}(\bs)$ where $V^{-2} \sim $ Gamma$(\nu / 2, \nu / 2)$.
+Returning to the proposed parameterization, let $\tau^{-2} = \frac{b}{a}\sigma^{-2} \sim$ Gamma$(a, a)$ so that (\ref{eq:fullmodel}) becomes
+\begin{align}
+  Y(\bs) = \tau \left[ \lambda \left(\frac{a}{b}\right)^{1 / 2} |z| + \left(\frac{a}{b}\right)^{1 / 2} v(\bs) \right].
+\end{align}
+Clearly setting $a = \nu / 2 > 0$, $b = \frac{\nu}{2(1 - \tilde{\lambda}^2)} > 0$, and $\lambda = \tilde{\lambda} / (1 - \tilde{\lambda}^2)^{1 / 2} \in (-\infty, \infty)$ resolves the difference in parameterizations.
+We note that our parameterization has three parameters $(a, b, \lambda)$ compared to the to parameters of the alternative parameterization $(\nu, \tilde{\lambda})$.
+However, if we introduce an overall scale parameter $c > 0$ into the alternative parameterization so that $\tilde{Y}(\bs) = c V \tilde{X}(\bs)$, then the two models remain equivalent by setting $a = \nu / 2$, $b = \frac{\nu}{2 c^2 (1 - \tilde{\lambda}^2)}$, and $\lambda = \tilde{\lambda} / (1 - \tilde{\lambda}^2)^{1 / 2}$.
+
 \section{Simulation study pairwise difference results} \label{a:pdiffs}

### 2016-05-30 LaTeX/response.tex
minor edits to the manuscript revisions
made a few minor tweaks to the responses. Added in a few new commands
to mycommands to make referencing things a little easier (e.g. \aref,
\eref, \fref)

@@
 \externaldocument{skewt_rev}
+\externaldocument{supplemental_rev}
 
@@
 Indeed, some reserachers adopt $a/2$ and $a/2$ (rather than $a$ and $b$) such that they end up with a degrees of freedom.
-Can you provide a web-supplement proof that your more general model with $a \neq b$ is a skew-$t$ distribution with $2a$ degrees of freedom, or provide a reference for this result?
+Can you provide a web-supplement proof that your more general model with $a \neq b$ is a skew-$t$ distribution with $2a$ degrees of freedom, or provide a reference for this result?\\
+
+\begin{response}
+  \hl{DONE:}
+  We have assumed that the errors $v(\bs)$ have unit variance, and $z \sim N(0, 1)$, so we require separate $a$ and $b$ so that $a$ can control the shape/degrees of freedom while $b$ controls the scale.
+  Viewed this way, if $a = b$, then the degrees of freedom would determine the scale which is clearly too restrictive.
+  To help clarify this, we have slightly modified the parameterization so that the degrees of freedom are $a$, and added additional clarification of the role of $b$ in Section \ref{s:skewt}.
+  We have also added \ref{a:otherparams} to demonstrate our more general model with $a \neq b$ is a skew-$t$ distribution with $a$ degrees of freedom.
+\end{response}
 
@@
 What have you used in the real dataset?
-Is this supposed to be a default choice?
+Is this supposed to be a default choice?\\
+
+\begin{response}
+  \hl{DONE:}
+  There was actually a typo in the manuscript as submitted.
+  The priors for $a$ and $b$ were inadvertently switched, so $a$ should have had a discrete distribution over a mesh from 0.1 to 10 with spacing $0.1$, and $b \sim $Gamma$(0.1, 0.1)$.
+  With the reparameterization given in the response to Major Comment 1, this actually means that the prior for $a$ is now a discrete distribution over a mesh from 0.2 to 20 with spacing $0.2$.
+  As we now clarify in the specification of these priors, $a$ is the scale-free shape parameter, and so this prior should be uninformative in all application.
+  In contrast, $b$ is a precision parameter, so it's prior should be adapted to the scale of the response.
+  In the real data analysis, we use the same discrete prior as in the simulation study, and find that for all settings in the cross validation, the posterior mean of $a$ is less than 3.
+\end{response}
 
@@
 
-What, exactly, is the convergence problem you had with the λ parameter before adopting this reparametrization?
-It affected only λ?
+What, exactly, is the convergence problem you had with the $\lambda$ parameter before adopting this reparametrization?
+It affected only $\lambda$?
 
 Incidentally, you used the same notation $IG(a,b)$ for two different purposes: for $\lambda_2$ in page 10, and for $\sigma^2$ in page 4.
-Please, keep them different as there are a lot of things going on in your model.
+Please, keep them different as there are a lot of things going on in your model.\\
+
+\begin{response}
+  \hl{DONE:}
+  There was a typo in the manuscript submission, the prior on $\lambda_1$ should have been a discrete prior on $\{-1, 0, 1\}$ with probability of $1 / 3$ for each option.
+  This was originally done to help with a computing issue when imputing the values below the threshold.
+  The computing issue was fixed prior to the original submission, but we elected to use the $\lambda$ parameterization as originally presented.
+  We have opted to use the parameterization in your suggestion where $\lambda \sim N(0, \sigma^2_\lambda)$.
+\end{response}
 
@@
 Anyway, you should make clear that trading a focus only on the extreme tail data you loose efficiency by decreasing the number of (real) data available.
-Censoring on purpose seems a rather drastic technique.
+Censoring on purpose seems a rather drastic technique.\\
+
+\begin{response}
+  \hl{TODO:}
+  This is a key argument in extreme value analysis which we realize we inadvertently took for granted.
+  This is a rendition of the classic bias variance trade-off.
+  A low threshold reduces variance by increasing the amount of data used to estimate parameters but risks inducing bias if the model is misspecified
+  On the other hand, a high threshold mitigates potential bias because only the tails must be modeled.
+  Therefore, by using a high threshold fewer precarious assumptions are required, but this increases variance by discarding data.
+  For our fairly flexible skew-$t$ process, we found that a moderate threshold optimized this balance for both simulated and ozone data.
+  Section \ref{s:censoring} now elaborates on this trade-off.
+\end{response}
 
 \item The temporal dependence structure is induced by putting AR(1) models on some of the skew-$t$ parameters.
-It is not clear what sort of temporal dependence that induces on the data.
+It is not clear what sort of temporal dependence that induces on the data.\\
+
+\begin{response}
+  \hl{TODO:}
+  Using simulated data, we have evidence to suggest that the model exhibits extremal dependence provided that the site $\bs$ is in the same partition for day $t$ and $t + 1$.
+  However, we are unable to give a formal proof that our temporal construction maintains extremal dependence.
+\end{response}
 \end{enumerate}
@@
 \subsection*{Minor comments}
+\hl{Unless otherwise stated, these are done}
 \begin{enumerate}[1.]
-  \item Page 3, line 11: ``\ldots with computing on the order of Gaussian models ''
+  \item Page 3, line 11: ``\ldots with computing on the order of Gaussian models ''\\
 
-  \item Page 3, line 13: \ldots as well as TO make predictions \ldots
+  \begin{response}
+    This has been corrected
+  \end{response}
 
-  \item Page 3, line -2: \ldots compared TO Gaussian AND max-stable \ldots
+  \item Page 3, line 13: \ldots as well as TO make predictions \ldots \\
+
+  \begin{response}
+    This has been corrected
+  \end{response}
+
+  \item Page 3, line -2: \ldots compared TO Gaussian AND max-stable \ldots \\
+
+  \begin{response}
+    This has been corrected
+  \end{response}
 
@@
 Is this property still respected in the $\bw$ space since the transformation is non-linear and acts independently in each coordinate?
+\\
 
-  \item Page 9, line -5: ``the marginal distributions are Gaussian'': which distributions, those of $Y(\bs)$?
+  \begin{response}
+    When we create the partition, it is created in the space of the original data (i.e. using $\bw$).
+    As you mention, because the transformation is non-linear and acts independently in each coordinate, the partition created using $\bw$ will be similar to, but not the same as the partition created using $\bw^*$.
+    We have added clarification in Section \ref{s:temporal} to clarify that the partition is created in the $\bw$ space as opposed to the $\bw^*$ space.
+    \hl{Is this something we should be concerned about?}
+  \end{response}
 
-  \item The hierarchical model in page 10 has $\sigma_t(\bs)$ multiplying only $v_t(\bs)$ and leaving out the term $|z_t(\bs)|$.
-  I think this is a typo.
+  \item Page 9, line -5: ``the marginal distributions are Gaussian'': which distributions, those of $Y(\bs)$? \\
 
-  \item In page 12, line -1: you mention the partition label switching problem appearing during the MCMC. This was not consequential for the inference? Can you expand on this?
+  \begin{response}
+    Yes, we now clarify that we are referring to the distribution of $Y(\bs)$.
+  \end{response}
 
-  \item There are several papers using skew distributions in the spatial context. You should refer to the most relevant ones in your related work section.
+  \item The hierarchical model in page 10 has $\sigma_t(\bs)$ multiplying only $v_t(\bs)$ and leaving out the term $|z_t(\bs)|$. I think this is a typo. \\
+
+  \begin{response}
+    This typo has been corrected.
+  \end{response}
+
+  \item In page 12, line -1: you mention the partition label switching problem appearing during the MCMC. This was not consequential for the inference? Can you expand on this? \\
+
+  \begin{response}
+    \hl{TODO:}
+    In a spatial analysis, the primary objectives are to make spatial predictions and to estimate extreme probabilities.
+    Both of these objectives are met by marginalizing over partition-specific parameters and are thus unaffected by label switching.
+    We have added a comment to this effect in the MCMC description, and caution that label switching might be more concerning in other analyses.
+  \end{response}
+
+  \item There are several papers using skew distributions in the spatial context. You should refer to the most relevant ones in your related work section. \\
+
+  \begin{response}
+    \hl{TODO:}
+    We have added several references given by Reviewer 2 to the introduction.
+  \end{response}
 
@@
 15 l.5).
-In a new version of the paper, the modeling aims should be stated clearly upfront so that the results can be assessed agains these aims.
+In a new version of the paper, the modeling aims should be stated clearly upfront so that the results can be assessed against these aims.
+\\
+
+  \begin{response}
+    \hl{TODO:}
+    As we now describe at the end of the first paragraph of the introduction, two objectives for our data analysis are spatial prediction and mapping the probability of an extreme event.
+    In the extremes context, spatial prediction means using the values of monitored stations to determine whether an unmonitored site experienced an extreme event.
+    Mapping the probability of extreme events could either be the marginal probability of an event, such as ozone exceeding 75ppb, or a map of an upper quantile, such as the map of the 99th percentile in Figure \ref{fig:ozoneq99}.
+  \end{response}
 
@@
 The role of the threshold needs to be much more clearly explained in the context of the application (fixed thresholds of interest), and modeling the extremes.
+\\
+
+  \begin{response}
+    \hl{TODO:}
+    Yes, we can see how this could lead to confusion.
+    In the revision, we now refer to the ``threshold'' as the value we select in the model-fitting process, i.e. $T$ in (\ref{eq:Yt}).
+    We also bolstered the description of the trade-off involved in selecting $T$ (see AE comment 4) below (\ref{eq:Yt}).
+    As mentioned in the response to the previous comment, we also want to compute probabilities of extreme events (e.g. ozone exceeding 75ppb).
+    We now refer to this as the probability above a high \emph{level}, $L$.
+    That is, we might use the data above threshold $T = 50$ppb in order to estimate the probability of the response exceeding level $L = 75$ppb.
+    Of course, this is only reasonable if $L \ge T$, and the user is free to select the $T$ that gives the best estimate of this probability.
+  \end{response}
 
@@
 Amazingly this is not even commented on!
-The purpose of the Poisson process seems to be for the proof in Web Appendix C, so if spatial long-range independence relies on this feature (which, assuming a constant rate μ would eventually guarantee an incease in the number of points, as opposed to taking a fixed number of draws from a uniform distribution) then the discrepancies between these two assumptions ought to be thoroughly discussed.
+The purpose of the Poisson process seems to be for the proof in Web Appendix C, so if spatial long-range independence relies on this feature (which, assuming a constant rate $\mu$ would eventually guarantee an incease in the number of points, as opposed to taking a fixed number of draws from a uniform distribution) then the discrepancies between these two assumptions ought to be thoroughly discussed.
+\\
+
+  \begin{response}
+    \hl{DONE:}
+    Thank you for pointing out this important difference.
+    Below ($\ref{eq:chiskewt}$), we now clarify that $\pi(h)$ goes to zero as $h$ increases assuming that the knots follow a homogeneous Poisson process.
+    In the model fitting procedure above (\ref{eq:hier})
+  \end{response}
 
@@
 Perhaps I am wrong, but either way it would be helpful to have the full proof spelt out to avoid any confusion.
+\\
+
+  \begin{response}
+    \hl{DONE:}
+    We have provided a more complete proof that $\lim_{h \rightarrow \infty} \pi(h) = 0$ in \ref{a:proofsamepartition}.
+  \end{response}
 
@@
 Are you able to analyze any theoretical properties of the suggested form, or even how this matches the data?
+\\
+
+  \begin{response}
+    \hl{TODO:}
+    Yes, as noted in the response to AE (major comment 5), we have evidence using simulated data that our temporal construction maintains extremal dependence.
+    This now appears \hl{where it will appear}.
+    Thank you both for the suggestion!
+  \end{response}
 
@@
   Furthermore the range of models simulated from is rather limited; for (4) you could simulate from a range of different max-stable processes (even if you don’t fit these, it might be interesting to see how the skew-t model handles them).
+\\
+
+  \begin{response}
+    \hl{Partly done:}
+    We have added clarification for these points in the description of the simulation study.
+    The parameters in the marginal distribution for all methods were estimated in a fully-Bayesian manner using MCMC methods.
+    The constant mean function was chosen somewhat arbitrarily, but permits a common threshold to apply to all sites.
+    Finally, we have also added a 5th design to the simulation study using a Brown-Resnick process with range 1 and smoothness 0.5, and find that the skew-$t$ method performs similarly to the other setting.
+    For this new setting, we find that our model performs similarly to the case in setting (4) with some improvement over the max-stable method for lower quantiles (0.90 -- 0.98), and the max-stable method tends to perform better at more extreme quantiles.
+  \end{response}
 
   It would be nice to see alternative model evaluations to the Brier score, or at least perhaps gain some sense of whether there is strong spatial variation in the score.
-If the model is much better at predicting in some locations than others, then one may want to be wary of using it in some areas\ldots
+If the model is much better at predicting in some locations than others, then one may want to be wary of using it in some areas\ldots \\
+
+  \begin{response}
+    \hl{TODO:}
+    We agree that maps of Brier scores would be informative.
+    This information could conceptually inform future monitor sites.
+    We have now added a map of Brier scores (see Web Figure), which shows \hl{XXXX}.
+  \end{response}
 
@@
 I agree this would be difficult to assess with so few data, but it’s not clear whether you have more data available to help assess this (even if you didn’t fit the model to all of them).
+\\
+
+  \begin{response}
+    \hl{TODO:}
+    Ozone exhibits strong seasonal patterns, so analyzing a longer series would require additional parameters.
+    Additional monitoring data would be easy enough to gather because it is available online.
+    The real limitation in our case is the CMAQ data, which we only have for this time period.
+  \end{response}
 \end{enumerate}
@@
 \subsection*{Minor comments}
+\hl{Unless otherwise stated, these are done}
+
 \begin{enumerate}[1.]
-  \item p.5l.6 $\gamma \rightarrow 1 - \gamma$.
+  \item p.5l.6 $\gamma \rightarrow 1 - \gamma$. \\
 
-  \item p.8 l.8 should $w_{t,1}$ etc. be bold?
+  \begin{response}
+    This typo has been corrected.
+  \end{response}
+
+  \item p.8 l.8 should $w_{t,1}$ etc. be bold? \\
+
+  \begin{response}
+    This typo has been corrected.
+  \end{response}
 
@@
 Giving adequate details on the transformation and dependence structure is sufficient.
+\\
 
-  \item p. 12 final line: you fit the correct model in some of these cases, so is it still difficult to tell whether the parameters converge even in this situation?
+  \begin{response}
+    We have replaced ``use a copula'' with ``use a transformation of a Gaussian random variable'' above (\ref{eq:sigmastar}) and (\ref{eq:zstar}).
+  \end{response}
 
-  \item p.13 l.13 and l.15 please give references for these tests
+  \item p. 12 final line: you fit the correct model in some of these cases, so is it still difficult to tell whether the parameters converge even in this situation? \\
 
-  \item p.14 eq.(20): Do you specify the mean of $Y$, or use the covariate specification given in (1)? These don’t look to be the same if $\lambda \neq 0$.
+  \begin{response}
+    \hl{I'm going to try to add in something about whether or not the parameters converge for skew-$t$ with 1 knot}
+    As with any Bayesian mixture model, the labels of the partitions are arbitrary.
+    That is, we could swap the labels of knots 1 and 2, and this would not affect the value of the posterior.
+    It would, however, change the values of $z_1$ and $z_2$ making it difficult to study convergence for these parameters.
+    However, since the quantities of interest (spatial predictions and probability of extreme events) are defined marginally over the partitioning, label-switching does not affect convergence for these quantities.
+  \end{response}
+
+  \item p.13 l.13 and l.15 please give references for these tests \\
+
+  \begin{response}
+    These have been added
+  \end{response}
+
+  \item p.14 eq.(20): Do you specify the mean of $Y$, or use the covariate specification given in (1)? These don’t look to be the same if $\lambda \neq 0$. \\
+
+  \begin{response}
+    Yes, technically the mean of $Y(\bs)$ depends on the shape parameter.
+    To match this equation with (\ref{eq:fullmodel}), we now simply specify that $\bX_t(\bs) = [1, \text{CMAQ}_t(\bs)]^\top$
+  \end{response}
 \end{enumerate}
@@
 \section*{Response to the comments of the Reviewer 2}
+\hl{I've addressed some of these in the updated draft, but not all of them.}
 
@@
 A distinction between observations and random objects which represent the phenomena that you are interested on, must be done.
+\\
+
+    \begin{response}
+      \hl{TODO:}
+      Conceptually we can imagine that $Y(\bs)$ is a random field that is only observed at $n$ locations as described in and around (\ref{eq:stazzalini}).
+      This notation could be problematic if the data consisted of multiple measurements at some sites that gave different values, but this is not our case and so we elect to retain this streamlined notation which is used, for example, throughout Stein’s 1999 textbook.
+    \end{response}
 
@@
 You used the $h = ||\bs_1 − \bs_2||$ (above equation (4)) to denote the distance between two specific fixed points and then $h = ||\bs−\bt||$ to denote the distance between two generic locations.
-The latter (more general) should also be used in equation (3) to avoid confusion;
+The latter (more general) should also be used in equation (3) to avoid confusion; \\
+
+    \begin{response}
+      \hl{DONE:}
+      Thank you for pointing this out.
+      We now use $\bs_1$ and $\bs_2$ in (\ref{eq:matern}), (\ref{eq:chih}), and (\ref{eq:chiskewt}) as opposed to $\bs$ and $\bt$ to avoid confusion between space and time.
+    \end{response}
 
@@
 In this simple example, is $\calD$ formed by two disjoint sets?
+\\
 
-    \item Second, you said: ``all $z(s)$ and $\sigma(s)$ for sites in the sub-region k are assigned common values z(s) = zk and σ(s) = σk''.
-I guess you mean that for all $s \in P_k$, with $k = 1, 2, \ldots, K$, the functions $z(s)$ and $σ(s)$ are equal to the constants $z_k$ and $\sigma_k$, respectively.
+    \begin{response}
+      \hl{DONE:}
+      Yes, $P_k$ is not the set of points on the circle with center $\bw_k$ and radius $k$, it is the set of points that are closer to $\bw_k$ than any of the other knots.
+      Therefore by definition $P_1, \ldots, P_k$ partition the spatial domain.
+      This is clarified after (\ref{eq:subregions}).
+    \end{response}
+
+    \item Second, you said: ``all $z(s)$ and $\sigma(s)$ for sites in the sub-region k are assigned common values $z(s)$ = $z_k$ and $\sigma(s)$ = $\sigma_k$''.
+I guess you mean that for all $s \in P_k$, with $k = 1, 2, \ldots, K$, the functions $z(s)$ and $\sigma(s)$ are equal to the constants $z_k$ and $\sigma_k$, respectively.
+\\
+
+    \begin{response}
+      \hl{DONE:}
+      Yes, we now use your phrasing.
+    \end{response}
 
@@
 A clear explanation of this part is needed, adding the required steps.
+\\
+
+    \begin{response}
+      \hl{TODO: Emeric?}
+      Setting aside the skewness parameter, the joint distribution of observations in different partition sets is that of $Y_1 = \sigma_1 e_1$ and $Y_2 = \sigma_2 e_2$ where $\sigma_1^2$ and $\sigma_2^2$ are independent inverse gamma and $(e_1, e_2)$ is bivariate normal.
+      If $\bs_1$ and $\bs_2$ are fixed, then $(Y_1, Y_2)$ is bivariate normal, and if $\sigma_1 = \sigma_2$ is random then $(Y_1, Y_2)$ is bivariate $t$ marginal over $\sigma_1$.
+      However, if $\sigma_1$ and $\sigma_2$ are different and unknown, then $\bY$ is neither Gaussian nor $t$.
+      From our perspective, however, the important feature of this distribution is that it has asymptotic independence, which is sufficient to break the long-range dependence of the usual spatial-$t$ process.
+    \end{response}
 
@@
 It seems that there is a missing term in your formula, if not you should explain why there is such a difference and refer to where you got that expression.
-My suggestion is that you should explain well, with the two steps (1) and (2) above, the process construction;
+My suggestion is that you should explain well, with the two steps (1) and (2) above, the process construction; \\
+
+    \begin{response}
+      \hl{DONE:}
+      As discussed in the response to the Associate Editor's Major Comment 1, given the confusion shared by all reviewers, we have added clarification about the role of $b$ as a scale parameter and provided \ref{a:otherparams} which shows the connection between our parameterization and the more standard parameterization.
+    \end{response}
 
@@
 But if so, how should the product with the vector $\beta$ be interpreted?
+\\
+
+    \begin{response}
+      \hl{TODO:}
+      $\bX(\bs)$ is the usual covariate vector as in multiple linear regression.
+      This is now made explicit below (\ref{eq:fullmodel}).
+      Then the coefficient has the linear regression interpretation that an increase of one in the $j$th element of $\bX$ shifts the distribution of $Y$ by the $j$th element of $\bbeta$.
+    \end{response}
 
@@
 Is it not much easier that you define first the AR(1) time series and then you define appropriate transformations that link these to the functional parameters of the skew-$t$ process and you clearly show that the finite dimensional distribution of the resulting process including time and space is skew-$t$?
+\\
+
+  \begin{response}
+    \hl{TODO:}
+    We have heeded your advice and swapped the transformation equations (now equations (\ref{eq:wstar}) -- (\ref{eq:sigmastar})) and the latent time series definitions (now equations (\ref{eq:ts-wstar}) -- (\ref{eq:ts-sigmastar})).
+  \end{response}
 
@@
 You should investigate more seriously this aspect.
+\\
+
+  \begin{response}
+    \hl{TODO:}
+  \end{response}
 
@@
 \subsection*{Minor comments}
+\hl{Unless otherwise stated, these are done}
+
 \begin{enumerate}[1.]
@@
 The references Beranger et al.
-(2015), Azzalini (2013) page 129 and Azzalini and Capitanio (2003), should be added for a clearer explanation of the additive representation;
+(2015), Azzalini (2013) page 129 and Azzalini and Capitanio (2003), should be added for a clearer explanation of the additive representation; \\
+
+  \begin{response}
+    We have added these references to clarify the representation.
+  \end{response}
 
@@
 in order to denote: transposed operation, a threshold (at page 6), a time index (at page 8).
-Different symbols should be used to avoid confusion;
+Different symbols should be used to avoid confusion; \\
 
-  \item Page 4 below equation (1). Specify what the size of $\bX(s)^\top$ and of $\beta$ are. Are you sure that $\lambda \in \mathbb{R}$? Is it actually $\lambda \in (−1,1)$?;
+  \begin{response}
+    \hl{Partially done. I got the $^\top$ for transposes, but I still need to change $T \rightarrow m$ for time points.}
+    The revision uses $^\top$ for transposes and $n_t$ for the number of time points.
+  \end{response}
+
+  \item Page 4 below equation (1). Specify what the size of $\bX(s)^\top$ and of $\beta$ are. Are you sure that $\lambda \in \mathbb{R}$? Is it actually $\lambda \in (−1,1)$?; \\
+
+    \begin{response}
+      \hl{TODO:}
+      Below (\ref{eq:fullmodel}) we clarify that $\bX(\bs)$ is $n_s \times p$ and $\beta$ is a $p$-vector.
+      Yes, we use the same $\lambda$ as given in equation (5.20) of Azzalini and Capitanio (2014, p. 129), so $\lambda \in \calR$.
+    \end{response}
 
   \item Page 4 below equation (1).
-A symbol for the correlation function of the process $v(s)$ should be used, e.g $\rho(h)$, in order to justify the presence of the matrix $\Sigma$ which is obtained from $\rho(h)$;
+A symbol for the correlation function of the process $v(s)$ should be used, e.g $\rho(h)$, in order to justify the presence of the matrix $\Sigma$ which is obtained from $\rho(h)$; \\
+
+    \begin{response}
+      At the end of the sentence following (\ref{eq:fullmodel}), we now define correlation function $\rho(\bs_1, \bs_2)$ and indicate the $\rho$ we use to populate the correlation matrix in (\ref{eq:matern}).
+    \end{response}
 
@@
 Furthermore, specify that matrix $\Sigma$ is obtained from $\rho(h)$ and this part should be jointed to equation (3) for brevity.
-Finally $\Omega$ is the dispersion matrix of $Y$;
+Finally $\Omega$ is the dispersion matrix of $Y$; \\
 
-  \item Page 5, equation (3). What does $|| \cdot ||$ mean?
+  \begin{response}
+    \hl{Partially done: Still need to specify the size of the matrices and vectors}
+    Based on the Minor Comment 4, we have now moved the $\rho(h)$ specification to (\ref{eq:matern}), and we now indicate after (\ref{eq:stazzalini}) that $\bSigma$ is obtained from $\rho(h)$.
+  \end{response}
+
+  \item Page 5, equation (3). What does $|| \cdot ||$ mean? \\
+
+  \begin{response}
+    As we now clarify below (\ref{eq:matern}), this is the distance between locations $\bs_1$ and $\bs_2$.
+  \end{response}
 
@@
 Since you are specifically referring to the skew-$t$ process, why don’t you simply say that $c^∗ = +\infty$?
-You should also confirm once more that for all given $\bs \in \calD$, $Y(\bs)$ is a skew-$t$ random variable, to avoid confusion;
+You should also confirm once more that for all given $\bs \in \calD$, $Y(\bs)$ is a skew-$t$ random variable, to avoid confusion; \\
 
-  \item Page 8, two rows below equation (10). The elements of the sequence $w_{t1}, \ldots , w_{tK}$ should be bolded.
+  \begin{response}
+    Here we provide a general definition which allows for a bounding density.
+    We now immediately clarify below (\ref{eq:chih}) that for the skew-$t$ distribution, $c^* = \infty$.
+  \end{response}
+
+  \item Page 8, two rows below equation (10). The elements of the sequence $w_{t1}, \ldots , w_{tK}$ should be bolded. \\
+
+  \begin{response}
+    This has been fixed
+  \end{response}
 
@@
 According to (10), $v_t(s)$ is a stochastic process.
+\\
+
+  \begin{response}
+    \hl{Not sure what to do about this one.}
+  \end{response}
 
@@
 You should use different symbols.
+\\
+
+  \begin{response}
+    We now say throughout the proof in \ref{a:proofsamepartition} that the point process is a homogeneous Poisson process with intensity $\mu(\cdot)$.
+  \end{response}
 \end{enumerate}

### 2016-06-06 LaTeX/response.tex
manuscript changes for revisions.

@@
 \begin{response}
-  \hl{DONE:}
   We have assumed that the errors $v(\bs)$ have unit variance, and $z \sim N(0, 1)$, so we require separate $a$ and $b$ so that $a$ can control the shape/degrees of freedom while $b$ controls the scale.
   Viewed this way, if $a = b$, then the degrees of freedom would determine the scale which is clearly too restrictive.
-  To help clarify this, we have slightly modified the parameterization so that the degrees of freedom are $a$, and added additional clarification of the role of $b$ in Section \ref{s:skewt}.
-  We have also added \ref{a:otherparams} to demonstrate our more general model with $a \neq b$ is a skew-$t$ distribution with $a$ degrees of freedom.
+  To help clarify this, we have slightly modified the parameterization so that the degrees of freedom are $a$, and added additional clarification of the role of $b$ in \sref{sts:skewt}.
+  We have also added \aref{sta:otherparams} to demonstrate our more general model with $a \neq b$ is a skew-$t$ distribution with $a$ degrees of freedom.
 \end{response}
@@
 \begin{response}
-  \hl{DONE:}
-  There was actually a typo in the manuscript as submitted.
-  The priors for $a$ and $b$ were inadvertently switched, so $a$ should have had a discrete distribution over a mesh from 0.1 to 10 with spacing $0.1$, and $b \sim $Gamma$(0.1, 0.1)$.
+  In the original submission, the priors for $a$ and $b$ were inadvertently switched, so $a$ should have had a discrete distribution over a mesh from 0.1 to 10 with spacing $0.1$, and $b \sim $Gamma$(0.1, 0.1)$.
   With the reparameterization given in the response to Major Comment 1, this actually means that the prior for $a$ is now a discrete distribution over a mesh from 0.2 to 20 with spacing $0.2$.
   As we now clarify in the specification of these priors, $a$ is the scale-free shape parameter, and so this prior should be uninformative in all application.
-  In contrast, $b$ is a precision parameter, so it's prior should be adapted to the scale of the response.
+  In contrast, $b$ is a precision parameter, so its prior should be adapted to the scale of the response.
   In the real data analysis, we use the same discrete prior as in the simulation study, and find that for all settings in the cross validation, the posterior mean of $a$ is less than 3.
@@
 \begin{response}
-  \hl{DONE:}
-  There was a typo in the manuscript submission, the prior on $\lambda_1$ should have been a discrete prior on $\{-1, 0, 1\}$ with probability of $1 / 3$ for each option.
-  This was originally done to help with a computing issue when imputing the values below the threshold.
-  The computing issue was fixed prior to the original submission, but we elected to use the $\lambda$ parameterization as originally presented.
+  We agree this was unnecessarily confusing.
   We have opted to use the parameterization in your suggestion where $\lambda \sim N(0, \sigma^2_\lambda)$.
@@
 \begin{response}
-  \hl{TODO:}
   This is a key argument in extreme value analysis which we realize we inadvertently took for granted.
   This is a rendition of the classic bias variance trade-off.
-  A low threshold reduces variance by increasing the amount of data used to estimate parameters but risks inducing bias if the model is misspecified
+  A low threshold reduces variance by increasing the amount of data used to estimate parameters but risks inducing bias if the model is misspecified.
   On the other hand, a high threshold mitigates potential bias because only the tails must be modeled.
@@
   For our fairly flexible skew-$t$ process, we found that a moderate threshold optimized this balance for both simulated and ozone data.
-  Section \ref{s:censoring} now elaborates on this trade-off.
+  \sref{sts:censoring} now elaborates on this trade-off.
 \end{response}
@@
 \subsection*{Minor comments}
-\hl{Unless otherwise stated, these are done}
 \begin{enumerate}[1.]
@@
     As you mention, because the transformation is non-linear and acts independently in each coordinate, the partition created using $\bw$ will be similar to, but not the same as the partition created using $\bw^*$.
-    We have added clarification in Section \ref{s:temporal} to clarify that the partition is created in the $\bw$ space as opposed to the $\bw^*$ space.
+    We have added clarification in Section \sref{sts:temporal} to clarify that the partition is created in the $\bw$ space as opposed to the $\bw^*$ space.
     \hl{Is this something we should be concerned about?}
@@
   \begin{response}
-    \hl{TODO:}
     In a spatial analysis, the primary objectives are to make spatial predictions and to estimate extreme probabilities.
     Both of these objectives are met by marginalizing over partition-specific parameters and are thus unaffected by label switching.
-    We have added a comment to this effect in the MCMC description, and caution that label switching might be more concerning in other analyses.
+    We have added a comment to this effect in the \sref{sts:simdesign}.
   \end{response}
@@
   \begin{response}
-    \hl{TODO:}
     As we now describe at the end of the first paragraph of the introduction, two objectives for our data analysis are spatial prediction and mapping the probability of an extreme event.
     In the extremes context, spatial prediction means using the values of monitored stations to determine whether an unmonitored site experienced an extreme event.
-    Mapping the probability of extreme events could either be the marginal probability of an event, such as ozone exceeding 75ppb, or a map of an upper quantile, such as the map of the 99th percentile in Figure \ref{fig:ozoneq99}.
+    Mapping the probability of extreme events could either be the marginal probability of an event, such as ozone exceeding 75ppb, or a map of an upper quantile, such as the map of the 99th percentile in Figure \fref{stfig:ozoneq99}.
   \end{response}
@@
     Yes, we can see how this could lead to confusion.
-    In the revision, we now refer to the ``threshold'' as the value we select in the model-fitting process, i.e. $T$ in (\ref{eq:Yt}).
-    We also bolstered the description of the trade-off involved in selecting $T$ (see AE comment 4) below (\ref{eq:Yt}).
+    In the revision, we now refer to the ``threshold'' as the value we select in the model-fitting process, i.e. $T$ in (\eref{steq:Yt}).
+    We also bolstered the description of the trade-off involved in selecting $T$ (see AE comment 4) below (\eref{steq:Yt}).
     As mentioned in the response to the previous comment, we also want to compute probabilities of extreme events (e.g. ozone exceeding 75ppb).
@@
   \begin{response}
-    \hl{DONE:}
     Thank you for pointing out this important difference.
-    Below ($\ref{eq:chiskewt}$), we now clarify that $\pi(h)$ goes to zero as $h$ increases assuming that the knots follow a homogeneous Poisson process.
-    In the model fitting procedure above (\ref{eq:hier})
+    Below \eref{steq:chiskewt}, we now clarify that $\pi(h)$ goes to zero as $h$ increases assuming that the knots follow a homogeneous Poisson process.
+    We also state (end of \sref{sts:part}) that in practice, the number of knots is fixed and selected by cross-validation.
   \end{response}
@@
   \begin{response}
-    \hl{DONE:}
-    We have provided a more complete proof that $\lim_{h \rightarrow \infty} \pi(h) = 0$ in \ref{a:proofsamepartition}.
+    We have provided a more complete proof that $\lim_{h \rightarrow \infty} \pi(h) = 0$ in \aref{sta:proofsamepartition}.
   \end{response}
@@
     The constant mean function was chosen somewhat arbitrarily, but permits a common threshold to apply to all sites.
-    Finally, we have also added a 5th design to the simulation study using a Brown-Resnick process with range 1 and smoothness 0.5, and find that the skew-$t$ method performs similarly to the other setting.
+    Finally, we have also added a 5th design to the simulation study using a Brown-Resnick process to generate data, and find that the skew-$t$ method performs similarly to the other setting.
     For this new setting, we find that our model performs similarly to the case in setting (4) with some improvement over the max-stable method for lower quantiles (0.90 -- 0.98), and the max-stable method tends to perform better at more extreme quantiles.
@@
   \begin{response}
-    \hl{TODO:}
     Ozone exhibits strong seasonal patterns, so analyzing a longer series would require additional parameters.
@@
 \subsection*{Minor comments}
-\hl{Unless otherwise stated, these are done}
 
@@
   \begin{response}
-    We have replaced ``use a copula'' with ``use a transformation of a Gaussian random variable'' above (\ref{eq:sigmastar}) and (\ref{eq:zstar}).
+    We have replaced ``use a copula'' with ``use a transformation of a Gaussian random variable'' above \eref{steq:sigmastar} and \eref{steq:zstar}.
   \end{response}
@@
     Yes, technically the mean of $Y(\bs)$ depends on the shape parameter.
-    To match this equation with (\ref{eq:fullmodel}), we now simply specify that $\bX_t(\bs) = [1, \text{CMAQ}_t(\bs)]^\top$
+    To match this equation with \eref{steq:fullmodel}, we now simply specify that $\bX_t(\bs) = [1, \text{CMAQ}_t(\bs)]^\top$
   \end{response}
@@
 \section*{Response to the comments of the Reviewer 2}
-\hl{I've addressed some of these in the updated draft, but not all of them.}
 
@@
 Then you use the same symbol to define a random field (spatial random process) in equation (1).
-I think you mean that $y(si), i = 1, 2, \ldots$ is an observation while $Y(s; ω) \equiv Y (s)$ is a random function for $s$ varying on a certain set.
+I think you mean that $y(s_i), i = 1, 2, \ldots$ is an observation while $Y(s; \omega) \equiv Y (s)$ is a random function for $s$ varying on a certain set.
 A distinction between observations and random objects which represent the phenomena that you are interested on, must be done.
@@
     \begin{response}
-      \hl{TODO:}
-      Conceptually we can imagine that $Y(\bs)$ is a random field that is only observed at $n$ locations as described in and around (\ref{eq:stazzalini}).
-      This notation could be problematic if the data consisted of multiple measurements at some sites that gave different values, but this is not our case and so we elect to retain this streamlined notation which is used, for example, throughout Stein’s 1999 textbook.
+      Conceptually we can imagine that $Y(\bs)$ is a random field that is only observed at $n$ locations as described in and around \eref{steq:stazzalini}.
+      This notation could be problematic if the data consisted of multiple measurements at some sites that gave different values, but this is not our case and so we elect to retain this streamlined notation which is used, for example, throughout \citet{Gelfand2010}.
     \end{response}
@@
     \item Page 5, equation (3).
-You used the $h = ||\bs_1 − \bs_2||$ (above equation (4)) to denote the distance between two specific fixed points and then $h = ||\bs−\bt||$ to denote the distance between two generic locations.
+You used the $h = ||\bs_1 - \bs_2||$ (above equation (4)) to denote the distance between two specific fixed points and then $h = ||\bs - \bt||$ to denote the distance between two generic locations.
 The latter (more general) should also be used in equation (3) to avoid confusion; \\
@@
     \begin{response}
-      \hl{DONE:}
       Thank you for pointing this out.
-      We now use $\bs_1$ and $\bs_2$ in (\ref{eq:matern}), (\ref{eq:chih}), and (\ref{eq:chiskewt}) as opposed to $\bs$ and $\bt$ to avoid confusion between space and time.
+      We now use $\bs_1$ and $\bs_2$ in \eref{steq:matern}, \eref{steq:chih}, and \eref{steq:chiskewt} as opposed to $\bs$ and $\bt$ to avoid confusion between space and time.
     \end{response}
@@
     The index of the partition is also used to denote the value of the radius.
-Does this mean that we also consider all the values $s ∈ \calD$ such that for a given $w_k$ we have that $\min ||s − w_k|| \le k$?
+Does this mean that we also consider all the values $s \in \calD$ such that for a given $w_k$ we have that $\min ||s - w_k|| \le k$?
 In this simple example, is $\calD$ formed by two disjoint sets?
@@
     \begin{response}
-      \hl{DONE:}
       Yes, $P_k$ is not the set of points on the circle with center $\bw_k$ and radius $k$, it is the set of points that are closer to $\bw_k$ than any of the other knots.
       Therefore by definition $P_1, \ldots, P_k$ partition the spatial domain.
-      This is clarified after (\ref{eq:subregions}).
+      This is clarified after \eref{steq:subregions}.
     \end{response}
@@
     \begin{response}
-      \hl{DONE:}
       Yes, we now use your phrasing.
@@
     \begin{response}
-      \hl{TODO: Emeric?}
       Setting aside the skewness parameter, the joint distribution of observations in different partition sets is that of $Y_1 = \sigma_1 e_1$ and $Y_2 = \sigma_2 e_2$ where $\sigma_1^2$ and $\sigma_2^2$ are independent inverse gamma and $(e_1, e_2)$ is bivariate normal.
-      If $\bs_1$ and $\bs_2$ are fixed, then $(Y_1, Y_2)$ is bivariate normal, and if $\sigma_1 = \sigma_2$ is random then $(Y_1, Y_2)$ is bivariate $t$ marginal over $\sigma_1$.
+      If $\sigma_1$ and $\sigma_2$ are fixed, then $(Y_1, Y_2)$ is bivariate normal, and if $\sigma_1 = \sigma_2$ is random then $(Y_1, Y_2)$ is bivariate $t$ marginal over $\sigma_1$.
       However, if $\sigma_1$ and $\sigma_2$ are different and unknown, then $\bY$ is neither Gaussian nor $t$.
@@
     \item Regardless of the different parametrization, i.e.
-you used $\sigma^2 \sim IG(a, b)$ instead of $V ∼ G(\nu/2, \nu/2)$, the formula should be equivalent.
+you used $\sigma^2 \sim IG(a, b)$ instead of $IG(\nu/2, \nu/2)$, the formula should be equivalent.
 It seems that there is a missing term in your formula, if not you should explain why there is such a difference and refer to where you got that expression.
@@
     \begin{response}
-      \hl{DONE:}
-      As discussed in the response to the Associate Editor's Major Comment 1, given the confusion shared by all reviewers, we have added clarification about the role of $b$ as a scale parameter and provided \ref{a:otherparams} which shows the connection between our parameterization and the more standard parameterization.
+      As discussed in the response to the Associate Editor's Major Comment 1, given the confusion shared by all reviewers, we have added clarification about the role of $b$ as a scale parameter and provided \aref{sta:otherparams} which shows the connection between our parameterization and the more standard parameterization.
     \end{response}
@@
     \begin{response}
-      \hl{TODO:}
       $\bX(\bs)$ is the usual covariate vector as in multiple linear regression.
-      This is now made explicit below (\ref{eq:fullmodel}).
+      This is now made explicit below \eref{steq:fullmodel}.
       Then the coefficient has the linear regression interpretation that an increase of one in the $j$th element of $\bX$ shifts the distribution of $Y$ by the $j$th element of $\bbeta$.
@@
   \begin{response}
-    \hl{TODO:}
-    We have heeded your advice and swapped the transformation equations (now equations (\ref{eq:wstar}) -- (\ref{eq:sigmastar})) and the latent time series definitions (now equations (\ref{eq:ts-wstar}) -- (\ref{eq:ts-sigmastar})).
+    We have heeded your advice and swapped the transformation equations (now equations \eref{steq:wstar} -- \eref{steq:sigmastar}) and the latent time series definitions (now equations \eref{steq:ts-wstar} -- \eref{steq:ts-sigmastar}).
   \end{response}
@@
 See e.g.
-Arellano-Valle and Azzalini (2008).
+\citet{ArellanoValle2008}.
 You should investigate more seriously this aspect.
@@
   \begin{response}
-    \hl{TODO:}
+    As mentioned in \sref{sts:simdesign}, some of the designs for the simulation study match the methods used to fit the datasets (e.g. design 2 and method 2, design 3 and method 4).
+    We have provided additional clarification in \sref{sts:simdesign} that for these models, we do appear to correctly estimate the skewness parameters and degrees of freedom.
+    We have also added clarification that the difficulty in assessing convergence is only relevant for models with multiple partitions (i.e.
+models 4 and 5) specifically for the parameters that change across partitions.
+    Because the parameters in our model do appear to properly converge when the design matches the method, we did not consider a centered parameterization as in \citet{ArellanoValle2008}.
   \end{response}
@@
   \item Page 4, first paragraph of Section 2.
-The references Beranger et al.
-(2015), Azzalini (2013) page 129 and Azzalini and Capitanio (2003), should be added for a clearer explanation of the additive representation; \\
+The references \citet{Beranger2016}, \citet{Azzalini2014} page 129 and \citet{Azzalini2003}, should be added for a clearer explanation of the additive representation; \\
 
@@
 
-  \item Page 4 below equation (1). Specify what the size of $\bX(s)^\top$ and of $\beta$ are. Are you sure that $\lambda \in \mathbb{R}$? Is it actually $\lambda \in (−1,1)$?; \\
+  \item Page 4 below equation (1). Specify what the size of $\bX(s)^\top$ and of $\beta$ are. Are you sure that $\lambda \in \mathbb{R}$? Is it actually $\lambda \in (-1,1)$?; \\
 
     \begin{response}
-      \hl{TODO:}
-      Below (\ref{eq:fullmodel}) we clarify that $\bX(\bs)$ is $n_s \times p$ and $\beta$ is a $p$-vector.
-      Yes, we use the same $\lambda$ as given in equation (5.20) of Azzalini and Capitanio (2014, p. 129), so $\lambda \in \calR$.
+      Below \eref{steq:fullmodel} we clarify that $X(\bs)$ and $\beta$ are $p$-vectors.
+      Yes, we use the same $\lambda$ as given in equation (5.20) of \citet[p. 129]{Azzalini2014}, so $\lambda \in \calR$.
     \end{response}
@@
     \begin{response}
-      At the end of the sentence following (\ref{eq:fullmodel}), we now define correlation function $\rho(\bs_1, \bs_2)$ and indicate the $\rho$ we use to populate the correlation matrix in (\ref{eq:matern}).
+      At the end of the sentence following \eref{steq:fullmodel}, we now define correlation function $\rho(\bs_1, \bs_2)$ and indicate the $\rho$ we use to populate the correlation matrix in \eref{steq:matern}.
     \end{response}
@@
   \begin{response}
-    \hl{Partially done: Still need to specify the size of the matrices and vectors}
-    Based on the Minor Comment 4, we have now moved the $\rho(h)$ specification to (\ref{eq:matern}), and we now indicate after (\ref{eq:stazzalini}) that $\bSigma$ is obtained from $\rho(h)$.
+    We have given the sizes of the vectors and matrices below equation \eref{steq:stazzalini} (previously equation (2)).
+Based on the Minor Comment 4, we have now moved the $\rho(h)$ specification to \eref{steq:matern}, and we now indicate after \eref{steq:stazzalini} that $\bSigma$ is obtained from $\rho(h)$.
   \end{response}
@@
   \begin{response}
-    As we now clarify below (\ref{eq:matern}), this is the distance between locations $\bs_1$ and $\bs_2$.
+    As we now clarify below \eref{steq:matern}, this is the Euclidean distance between locations $\bs_1$ and $\bs_2$.
   \end{response}
@@
   \item Page 5, equation (4).
-Since you are specifically referring to the skew-$t$ process, why don’t you simply say that $c^∗ = +\infty$?
+Since you are specifically referring to the skew-$t$ process, why don't you simply say that $c^∗ = +\infty$?
 You should also confirm once more that for all given $\bs \in \calD$, $Y(\bs)$ is a skew-$t$ random variable, to avoid confusion; \\
@@
     Here we provide a general definition which allows for a bounding density.
-    We now immediately clarify below (\ref{eq:chih}) that for the skew-$t$ distribution, $c^* = \infty$.
+    We now immediately clarify below \eref{steq:chih} that for the skew-$t$ distribution, $c^* = \infty$.
   \end{response}
@@
   \begin{response}
-    \hl{Not sure what to do about this one.}
+    Yes, $v_t(\bs)$ is a stochastic process, and simulating data from the model requires generating $v$ at the data points from a multivariate normal distribution.
   \end{response}
@@
   \begin{response}
-    We now say throughout the proof in \ref{a:proofsamepartition} that the point process is a homogeneous Poisson process with intensity $\mu(\cdot)$.
+    We now say throughout the proof in \aref{sta:proofsamepartition} that the point process is a homogeneous Poisson process with intensity $\mu(\cdot)$.
   \end{response}
@@
 
+\bibliographystyle{biom}
+\bibliography{library}
+
 \end{document}

### 2016-06-06 LaTeX/skewt_rev.tex
manuscript changes for revisions.

@@
 
-\pagerange{\pageref{firstpage}--\pageref{lastpage}}
+\pagerange{\pageref{stfirstpage}--\pageref{stlastpage}}
 
-\label{firstpage}
+\label{stfirstpage}
 
@@
 
-\section{Introduction}\label{s:intro}
+\section{Introduction}\label{sts:intro}
 % In many climatological applications, researchers are interested in learning about the average behavior of different climate variables (e.g. ozone, temperature, rainfall).
@@
 A site is said to be in compliance if the fourth highest daily maximum 8-hour concentration averaged over 3 years does not exceed 75 parts per billion (ppb).
-Figure \ref{fig:ozone-10jul} shows the ozone levels from July 10, 2005, at 1089 stations across the United States.
+\fref{stfig:ozone-10jul} shows the ozone levels from July 10, 2005, at 1089 stations across the United States.
 We see a large area above the compliance level in the midwest covering Ohio, Indiana, Illinois, and parts of the surrounding states.
+We analyze these data with the goals of spatial prediction for unmonitored locations and to map the probability of extreme events.
 
@@
   \caption{Ozone values (ppb) on July 10, 2005}
-  \label{fig:ozone-10jul}
+  \label{stfig:ozone-10jul}
 \end{figure}
@@
 Third, we aim to capture the dependence structure when ozone is at high levels, and dependence at these levels may not be well-represented by covariances which focus again on mean behavior.
-Asymptotic dependence/independence (see Section \ref{s:extdep}) are notions which describe how two random variables' probability of simultaneous exceedance of an extreme threshold behaves as one increases the threshold.
+Asymptotic dependence/independence (see \sref{sts:extdep}) are notions which describe how two random variables' probability of simultaneous exceedance of an extreme threshold behaves as one increases the threshold.
 The Gaussian distribution always exhibits asymptotic independence, except in the case of perfect dependence, thus is an inappropriate model for data which exhibits asymptotic dependence.
@@
 We use a skew-$t$ distribution because of its flexibility to model asymmetry and heavy-tailed data with the aim of predicting the probability of exceeding a high fixed threshold at an unobserved location.
+\hl{Need to add in more references here to skew-$t$ used in a spatial setting.}
 
@@
 The paper is organized as follows.
-Section \ref{s:spatialskew} is a brief review of the spatial skew-$t$ process.
-In Section \ref{s:spatial}, we build upon the traditional skew-$t$ process by incorporating censoring to focus on tails, partitioning to remove long-range asymptotic dependence, and extending the model to space-time data.
-The computing is described in Section \ref{s:hier}.
-In Section \ref{s:simstudy}, we present a simulation study that examines the predictive capabilities of this model compared to Gaussian and max-stable methods.
-We then compare our method to Gaussian and max-stable methods with a data analysis of ozone measurements from 800 sites throughout the US in Section \ref{s:analysis}.
+\sref{sts:spatialskew} is a brief review of the spatial skew-$t$ process.
+In \sref{sts:spatial}, we build upon the traditional skew-$t$ process by incorporating censoring to focus on tails, partitioning to remove long-range asymptotic dependence, and extending the model to space-time data.
+The computing is described in \sref{sts:hier}.
+In \sref{sts:simstudy}, we present a simulation study that examines the predictive capabilities of this model compared to Gaussian and max-stable methods.
+We then compare our method to Gaussian and max-stable methods with a data analysis of ozone measurements from 800 sites throughout the US in \sref{sts:analysis}.
 The final section provides brief discussion and direction for future research.
@@
 
-\section{Spatial skew processes}\label{s:spatialskew}
-The skew-elliptical family of distributions provides models that are mathematically tractable while introducing a slant parameter to account for asymmetric data \citep{Azzalini2014}.
-A brief review of the additive process by which a skew-$t$ process is created is given here.
+\section{Spatial skew processes}\label{sts:spatialskew}
+The skew-elliptical family of distributions provides models that are mathematically tractable while introducing a slant parameter to account for asymmetric data.
+A brief review of the additive process (\citealp{Azzalini2003}; \citealp[p. 129]{Azzalini2014}; \citealp{Beranger2016}) by which a skew-$t$ process is created is given here.
 
-\subsection{Skew-$t$ process} \label{s:skewt}
-Let $Y(\bs)$ be the observation at spatial location $\bs$ in a spatial domain of interest $\calD \in \calR^2$.
+\subsection{Skew-$t$ process} \label{sts:skewt}
+Let $Y(\bs)$ be a spatial process defined for spatial location $\bs$ in a spatial domain of interest $\calD \in \calR^2$.
 The spatial skew-$t$ process can be written
 \begin{align}
-  Y(\bs) = \bX(\bs)^\prime \bbeta + \lambda \sigma |z| + \sigma v(\bs) \label{eq:fullmodel}
+  Y(\bs) = \bX(\bs)^\top \bbeta + \lambda \sigma |z| + \sigma v(\bs) \label{steq:fullmodel}
 \end{align}
-where $\bX(\bs)$ is a set of spatial covariates at site $\bs$, $\bbeta$ is the vector of regression parameters, $\lambda \in \calR$ is a parameter controlling skew, $z \sim N(0, b)$, $\sigma^2 \sim \text{IG}(a / 2, a / 2)$ is random scale parameter, IG is the distribution function of an inverse gamma random variable, and $v(\bs)$ is a spatial Gaussian process with mean zero, variance $b^2$, and a positive definite correlation function $\rho(\bs_1, \bs_2)$.
+$\bX(\bs)$ is the observed covariate vector at site $\bs$, $\bbeta$ is the $p$-vector of regression parameters, $\lambda \in \calR$ is a parameter controlling skew, $z \sim N(0, 1)$, $\sigma^2 \sim \text{IG}(a / 2, b)$ is random scale parameter, IG is the distribution function of an inverse gamma random variable, $a$ is the degrees of freedom, $b$ controls the precision of the process, and $v(\bs)$ is a spatial Gaussian process with mean zero, variance one, and a positive definite correlation function $\text{Cor} \left[Y(\bs_1), Y(\bs_2)\right] = \rho(\bs_1, \bs_2)$.
 Although any positive definite correlation function could be used, we choose to use the stationary isotropic \Matern correlation with
 \begin{align}
-  \rho(h) = (1 - \gamma) I(h = 0) + \gamma \frac{ 1 }{ \Gamma(\nu) 2^{ \nu - 1}} \left( \sqrt{2\nu} \frac{ h }{ \varphi } \right)^{\nu} K_{\nu} \left( \sqrt{2\nu} \frac{ h }{ \varphi } \right) \label{eq:matern}
+  \rho(h) = (1 - \gamma) I(h = 0) + \gamma \frac{ 1 }{ \Gamma(\nu) 2^{ \nu - 1}} \left( \sqrt{2\nu} \frac{ h }{ \varphi } \right)^{\nu} K_{\nu} \left( \sqrt{2\nu} \frac{ h }{ \varphi } \right) \label{steq:matern}
 \end{align}
-where $\varphi$ is the spatial range, $\nu$ is the smoothness, $\gamma$ is the proportion of variance accounted for by the spatial variation, $K_\nu$ is a modified Bessel function of the second kind, and $h = || \bs_1 - \bs_2 ||$ is the distance between sites $\bs_1$ and $\bs_2$.
+where $\varphi$ is the spatial range, $\nu$ is the smoothness, $\gamma$ is the proportion of variance accounted for by the spatial variation, $K_\nu$ is a modified Bessel function of the second kind, and $h = || \bs_1 - \bs_2 ||$ is the Euclidean distance between sites $\bs_1$ and $\bs_2$.
 
-For a finite collection of locations $\bs_1, \ldots, \bs_n$, denote the vector of observations $\bY = [Y(\bs_1), \ldots, Y(\bs_n)]^\prime$.
-After marginalizing over both $z$ and $\sigma$,
-\begin{align}
-  \bY \sim \text{ST}_n(\bX \bbeta, \bOmega, \balpha, 2a),
+For a finite collection of locations $\bs_1, \ldots, \bs_n$, denote the vector of observations $\bY = [Y(\bs_1), \ldots, Y(\bs_n)]^\top$.
+After marginalizing over both $z$ and $\sigma$, using the notation from \citet[p. 176]{Azzalini2014},
+\begin{align} \label{steq:stazzalini}
+  \bY \sim \text{ST}_n(\bX \bbeta, \bOmega, \balpha, a),
 \end{align}
-that is, $\bY$ follows an $n$-dimensional skew-$t$ distribution with location $\bX \bbeta$, variance matrix $\bOmega$, slant parameters $\balpha$ and degrees of freedom $2a$, where $\bX = [\bX(\bs_1)^\prime, \ldots, \bX(\bs_n)^\prime]$, $\bOmega = \bomega \bar{\bOmega} \bomega$, $\bomega = \text{diag}\left(\frac{1}{ \sqrt{ab}}, \ldots, \frac{1}{\sqrt{ab}}\right)$, $\bar{\bOmega} = (\bSigma + \lambda^2 \bOne \bOne^\prime$), $\balpha = \lambda (1 + \lambda^2 \bOne^\prime \bSigma^{-1} \bOne)^{-1/2} \bOne^\prime \bSigma^{-1}$, and $\bSigma$ is the positive definite correlation matrix which is obtained from $\rho(h)$.
-This process is desirable because of its flexible tail that is controlled by the skewness parameter $\lambda$ and degrees of freedom $2a$.
+that is, $\bY$ follows an $n$-dimensional skew-$t$ distribution with location $\bX \bbeta \in \calR^n$, variance matrix $\bOmega_{n \times n}$, $\balpha \in \calR^n$ slant parameters, degrees of freedom $a$, where $\bX_{n \times p} = [\bX(\bs_1)^\top, \ldots, \bX(\bs_n)^\top]$,
+$\bX(\bs_i), i = 1, \ldots, n$, are observed covariates at each site,
+$\bOmega_{n \times n} = \bomega \bar{\bOmega} \bomega$,
+$\bar{\bOmega}_{n \times n} = \frac{1}{1 + \lambda^2}(\bSigma + \lambda^2 \bOne \bOne^\top)$,
+$\bomega_{n \times n} = \text{diag}\left(\sqrt{\frac{a(1 + \lambda^2)}{ b }}, \ldots, \sqrt{\frac{a(1 + \lambda^2)}{ b }} \right)$,
+$\balpha = \lambda (1 + \lambda^2)^{1 / 2} (1 + \lambda^2 \bOne^\top \bSigma^{-1} \bOne)^{-1/2} \bSigma^{-1} \bOne$, and $\bSigma_{n \times n}$ is the positive definite correlation matrix which is obtained from $\rho(h)$.
+This process is desirable because of its flexible tail that is controlled by the skewness parameter $\lambda$ and degrees of freedom $a$.
 Furthermore, the marginal distributions at each location also follow a univariate skew-$t$ distribution \citep{Azzalini2014}.
+For comparisons with other parameterizations, please see \aref{sta:otherparams}.
 
@@
 
-\subsection{Extremal dependence}\label{s:extdep}
+\subsection{Extremal dependence}\label{sts:extdep}
 Our interest lies in spatial dependence in the tail of the skew-$t$ process.
@@
 For a stationary and isotropic spatial process, the $\chi$ statistic for locations $\bs_1$ and $\bs_2$ separated by distance $h = || \bs_1 - \bs_2 ||$ with identical marginal distributions is
-\begin{align}
+\begin{align} \label{steq:chih}
   \chi(h) = \lim_{c \rightarrow c^*} \Pr[Y(\bs_1) > c | Y(\bs_2) > c]
@@
 For Gaussian processes, $\chi(h) = 0$ regardless of the distance $h$, so they are not suitable for modeling asymptotically dependent extremes.
-Unlike the Gaussian process, the skew-$t$ process is asymptotically dependent (the explicit expression for $\chi(h)$ is given in \ref{a:skewt}).
+Unlike the Gaussian process, the skew-$t$ process is asymptotically dependent (the explicit expression for $\chi(h)$ is given in \aref{sta:skewt}).
 However, one problem with the spatial skew-$t$ process is that $\lim_{h \rightarrow \infty} \chi(h) > 0$.
@@
 Therefore, this long-range dependence feature of the skew-$t$ process is not ideal for spatial analysis of large geographic regions where we expect only local spatial dependence.
-We propose a solution to this in Section \ref{s:part}.
+We propose a solution to this in \sref{sts:part}.
 
-\section{Spatiotemporal skew-$t$ model for threshold exceedances}\label{s:spatial}
+\section{Spatiotemporal skew-$t$ model for threshold exceedances}\label{sts:spatial}
 In this section, we propose extensions to the skew-$t$ process to model spatial extremes over a large geographic region by introducing censoring to focus on tail behavior and a random partition to remove long-range asymptotic dependence.
-For notational convenience, we introduce the model for a single replication, and then extend this model to the spatiotemporal setting in Section \ref{s:temporal}.
+For notational convenience, we introduce the model for a single replication, and then extend this model to the spatiotemporal setting in \sref{sts:temporal}.
 
-\subsection{Censoring to focus on the tails}
+\subsection{Censoring to focus on the tails} \label{sts:censoring}
 As mentioned previously, we propose to use a censored approach because we are interested in high values and do not want the low-to-moderate values to influence the fit of the overall model.
@@
 Let
-\beq\label{Yt}
+\beq \label{steq:Yt}
   \widetilde{Y}(\bs) = \left\{ \begin{array}{ll}
@@
 be the censored observation at site $\bs$ where $Y(\bs)$ is the uncensored observation, $\delta(\bs) = I[Y(\bs) > T]$, and $T$ is a pre-specified threshold value.
-We impute the censored values as a step in the MCMC algorithm used to fit the model described in Section \ref{s:comp}.
+Selecting $T$ too small may lead to bias in estimating the tail parameters; selecting $T$ to large increases variance.
+We impute the censored values as a step in the MCMC algorithm used to fit the model described in \sref{sts:comp}, and use cross-validation to select $T$.
 
-\subsection{Partitioning to remove long-range asymptotic dependence}\label{s:part}
+\subsection{Partitioning to remove long-range asymptotic dependence}\label{sts:part}
 The motivation for the partition is that for a large spatial domain, it may not be reasonable to assume sites that are far apart demonstrate asymptotic dependence.
@@
 We handle the problem of long-range asymptotic dependence with a random partition.
-As discussed in Section \ref{s:spatialskew}, the source of long-range dependence is the shared $z$ and $\sigma$.
+As discussed in \sref{sts:spatialskew}, the source of long-range dependence is the shared $z$ and $\sigma$.
 Therefore, to alleviate this dependence, we allow $z$ and $\sigma$ to vary by site.
 The model becomes
-\begin{align} \label{eq:partition}
-  Y(\bs) &= \bX(\bs)^\prime \bbeta + \lambda \sigma(\bs) |z(\bs)| + \sigma(\bs) v(\bs).
+\begin{align} \label{steq:partition}
+  Y(\bs) &= \bX(\bs)^\top \bbeta + \lambda \sigma(\bs) |z(\bs)| + \sigma(\bs) v(\bs).
 \end{align}
-To model spatial variation, consider a set of spatial knots $\bw_1, \ldots, \bw_K$ from a homogeneous Poisson process with intensity $\mu$ over spatial domain $\calD \in \calR^2$.
+To model spatial variation, consider a set of spatial knots $\bw_1, \ldots, \bw_K$ from a homogeneous Poisson process with intensity $\mu_{PP}$ over spatial domain $\calD \in \calR^2$.
 The knots define a random partition of $\calD$ by subregions $P_{1}, \ldots, P_{K}$ defined as
-\begin{align} \label{eq:subregions}
+\begin{align} \label{steq:subregions}
   P_{k} = \{ \bs : k = \argmin_\ell || \bs - \bw_{\ell} || \}.
 \end{align}
-In other words, $P_1$ is composed of all sites for which the closest knot is $\bw_1$.
-All $z(\bs)$ and $\sigma(\bs)$ for sites in subregion $k$ are assigned common values
-\begin{align}
-  z(\bs) = z_{k} \quad \text{and} \quad \sigma(\bs) = \sigma_{k} \label{eq:sitezsig}
-\end{align}
-and the $z_k$ and $\sigma^2_k$ are distributed as $z_k \iid N(0, 1)$ and $\sigma^2_k \iid$ IG$(a, b)$.
-So, within each partition, $Y(\bs)$ follows the spatial skew-$t$ process defined in Section \ref{s:spatialskew}.
+In other words, $P_k$ is composed of all sites for which the closest knot is $\bw_k$.
+For all $\bs \in P_k$, with $k = 1, 2, \ldots, K$, the functions $z(\bs)$ and $\sigma(\bs)$ are equal to the constants $z_k$ and $\sigma_k$ respectively, and the $z_k$ and $\sigma^2_k$ are distributed as $z_k \iid N(0, 1)$ and $\sigma^2_k \iid$ IG$(a, b)$.
+So, within each partition, $Y(\bs)$ follows the spatial skew-$t$ process defined in \sref{sts:spatialskew}.
 Across partitions, the $Y(\bs)$ remain correlated via the correlation function for $v(\bs)$ because $v(\bs)$ spans all partitions.
+However, the bivariate distribution for sites in different partitions is neither Gaussian nor skew-$t$ and does not have asymptotic dependence.
 
@@
   \chi(h) &= I(k_1 = k_2) \chi_{\text{skew-}t}(h) + I(k_1 \neq k_2) \chi_{\text{Gaus}}(h) \nonumber \\
-         &= I(k_1 = k_2) \chi_{\text{skew-}t}(h)
+         &= I(k_1 = k_2) \chi_{\text{skew-}t}(h) \label{steq:chiskewt}
 \end{align}
-where $I(\cdot)$ is an indicator function, $\chi_{\text{skew-}t}(h)$ is the $\chi$ statistic for a skew-$t$ process given in (\ref{eq:chiskew-t}), $\chi_{\text{Gaus}}(h)$ is the $\chi$ statistic for a Gaussian process, and $h = ||\bs - \bt||$.
+where $I(\cdot)$ is an indicator function, $\chi_{\text{skew-}t}(h)$ is the $\chi$ statistic for a skew-$t$ process given in \eref{steq:chiskew-t}, $\chi_{\text{Gaus}}(h)$ is the $\chi$ statistic for a Gaussian process, and $h = ||\bs_1 - \bs_2||$.
 Therefore, sites in different subregions are asymptotically independent because $\chi_{\text{Gaus}}(h) = 0$ for all $h$.
 Marginally, over the knots, $\chi(h) = \pi(h) \chi_{\text{skew}-t}(h)$, where $\pi(h) = \Pr(k_1 = k_2)$ is the probability that two sites separated by distance $h$ are in the same partition.
-In \ref{a:proofsamepartition}, we show that $\lim_{h \rightarrow \infty} \pi(h) = 0$, implying $\lim_{h \rightarrow \infty} \chi(h) = 0$.
-In Figure \ref{fig:chi}, we give $\chi(h)$ for $K = 1, 3, 5, 10$ partitions for a skew-$t$ distribution with $\alpha = 10$, and 3 degrees of freedom.
-In Figure \ref{fig:chi} we estimate $\pi(h)$ through simulation.
+In \aref{sta:proofsamepartition}, we show that $\lim_{h \rightarrow \infty} \pi(h) = 0$, assuming that the knots follow a homogeneous Poisson process.
+This implies that $\lim_{h \rightarrow \infty} \chi(h) = 0$.
+In practice we fix $K$ at a finite value and use a uniform distribution for the knots $\bw_1, \ldots, \bw_K$.
+In \fref{stfig:chi}, we give $\chi(h)$ for $K = 1, 3, 5, 10$ partitions for a skew-$t$ distribution with $\alpha = 10$, and 3 degrees of freedom.
+In \fref{stfig:chi} we estimate $\pi(h)$ through simulation.
 % To estimate $\pi(h)$, we generate 500 sites uniformly over the unit-square.
@@
   \caption{Extremal dependence measure $\chi(h)$, as a function of distance, $h$, for $K = 1, 3, 5$, and $10$ knots.}
-  \label{fig:chi}
+  \label{stfig:chi}
 \end{figure}
 
-\subsection{Extension to space-time data} \label{s:temporal}
+\subsection{Extension to space-time data} \label{sts:temporal}
 When using daily measurements, the assumption of temporal independence is often inappropriate.
-In this section, we extend (\ref{eq:partition}) to the spatiotemporal setting.
+In this section, we extend \eref{steq:partition} to the spatiotemporal setting.
 There are several places where temporal dependence could be incorporated in the model, including the residuals $v_t(\bs)$.
@@
 Let
-\begin{align} \label{eq:spatiotemp}
-  Y_t(\bs) = \bX_t(\bs)^\prime \bbeta + \lambda \sigma_t(\bs) |z_t(\bs)| + \sigma_t(\bs) v_t(\bs),
+\begin{align} \label{steq:spatiotemp}
+  Y_t(\bs) = \bX_t(\bs)^\top \bbeta + \lambda \sigma_t(\bs) |z_t(\bs)| + \sigma_t(\bs) v_t(\bs),
 \end{align}
@@
 Let \hbox{$\bw_{tk} = (w_{tk1}, w_{tk2})$} be a spatial knot on day $t$, and let $\bw_{t1}, \ldots, \bw_{tK}$ be a collection of spatial knots on day $t$.
-As in Section \ref{s:part}, these knots define a daily partition $P_{t1}, \ldots, P_{tK}$, and for $\bs \in P_{tk}$,
+As in \sref{sts:part}, these knots define a daily partition $P_{t1}, \ldots, P_{tK}$, and for $\bs \in P_{tk}$,
 \begin{align}
@@
 Let
-\begin{align}
+\begin{align} \label{steq:wstar}
   w^*_{tki} = \Phi^{-1}\left[ \frac{ w_{tki} - \min(\bs_i)}{ \max(\bs_i) - \min(\bs_i) } \right], \quad i = 1, 2
@@
 Then the transformed knots $\bw^*_{tk} \in \calR^2$.
-We use a transformation of a Gaussian random variable on $\sigma^2_t(\bs)$ to ensure that the marginal distributions of $\sigma^2_t(\bs)$ are inverse gamma.
+We use a transformation of a Gaussian random variable on $z_{t}(\bs)$ to ensure that the marginal distributions of $z_t(\bs)$ are half-normal.
 Let
-\begin{align}
+\begin{align} \label{steq:zstar}
+  z^*_t(\bs) = \Phi^{-1}\left\{ \text{HN}[z_t(\bs)] \right\}
+\end{align}
+where HN is the distribution function of a half-normal random variable.
+We also use a transformation of a Gaussian random variable on $\sigma^2_t(\bs)$ to ensure that the marginal distributions of $\sigma^2_t(\bs)$ are inverse gamma.
+Let
+\begin{align} \label{steq:sigmastar}
   \sigma^{2*}_t(\bs) =\Phi^{-1}\left\{ \text{IG}[\sigma^2_t(\bs)] \right\}
@@
 where IG is defined as before.
-We also use a transformation of a Gaussian random variable on $z_{t}(\bs)$ to ensure that the marginal distributions of $z_t(\bs)$ are half-normal.
-Let
-\begin{align}
-  z^*_t(\bs) = \Phi^{-1}\left\{ \text{HN}[z_t(\bs)] \right\}
-\end{align}
-where HN is the distribution function of a half-normal random variable.
 The AR(1) process for each tail parameter is $\bw^*_{1k} \sim N_w(0, 1)$, $z^*_{1k} \sim N(0, 1)$, $\sigma^{2*}_{1k} \sim N(0, 1)$, and for $t > 1$ the time series is modeled as
 \begin{align}
-  \bw^*_{tk} | \bw^*_{t-1, k} &\sim N_2\left[\phi_w \bw^*_{t-1, k}, (1 - \phi_w^2) \right] \\
-  z^*_{tk} | z^*_{t-1, k} &\sim N \left[\phi_z z^*_{t-1, k}, (1 - \phi_z^2)\right] \\
-  \sigma^{2*}_{tk} | \sigma^{2*}_{t-1, k} &\sim N \left[\phi_\sigma \sigma^{2*}_{t-1, k}, (1 - \phi_\sigma^2) \right]
+  \bw^*_{tk} | \bw^*_{t-1, k} &\sim N_2\left[\phi_w \bw^*_{t-1, k}, (1 - \phi_w^2) \right] \label{steq:ts-wstar}\\
+  z^*_{tk} | z^*_{t-1, k} &\sim N \left[\phi_z z^*_{t-1, k}, (1 - \phi_z^2)\right] \label{steq:ts-zstar} \\
+  \sigma^{2*}_{tk} | \sigma^{2*}_{t-1, k} &\sim N \left[\phi_\sigma \sigma^{2*}_{t-1, k}, (1 - \phi_\sigma^2) \right] \label{steq:ts-sigmastar}
 \end{align}
@@
 These are stationary time series models with marginal distributions \hbox{$\bw^*_{k} \sim N_2(0, 1)$}, \hbox{$z^*_{k} \sim N(0, 1)$}, and \hbox{$\sigma^{2*}_{k} \sim N(0, 1)$}.
-After transformation back to the original space, $\bw_{tk} \sim \text{Unif}(\calD)$, $z_{tk} \sim HN(0, 1)$. and $\sigma^2_{tk} \sim \text{IG}(a, b)$.
-For each day, the model is identical to the spatial-only model in (\ref{eq:partition}) by construction.
+After transformation back to the original space, $\bw_{tk} \sim \text{Unif}(\calD)$, $z_{tk} \sim HN(0, 1)$, and $\sigma^2_{tk} \sim \text{IG}(a, b)$.
+We then create the partition for day $t$ using $\bw_{t1}, \ldots, \bw_{tk}$.
+For each day, the model is identical to the spatial-only model in \eref{steq:partition} by construction.
 
-\section{Hierarchical model}\label{s:hier}
+\section{Hierarchical model}\label{sts:hier}
 Conditioned on $z_{tk}(\bs)$, $\sigma^2_{tk}(\bs)$, and $P_{tk}$, the marginal distributions of $Y(\bs)$ are Gaussian and the joint distribution of $Y(\bs)$ is multivariate Gaussian.
@@
 We model this with a Bayesian hierarchical model as follows.
-Let $\bw_{t1}, \ldots, \bw_{tK}$ be a set of daily spatial knots in a spatial domain of interest, $\calD$, and $P_{tk}$ as defined in (\ref{eq:subregions}).
-In practice, we fix $K$ at different levels, and assess its impact on prediction as described in \ref{s:modelselect}.
+Let $\bw_{t1}, \ldots, \bw_{tK}$ be a set of daily spatial knots in a spatial domain of interest, $\calD$, and $P_{tk}$ as defined in \eref{steq:subregions}.
+One approach would be to allow $K$ to be unknown and follow a Poisson process prior, but this would lead to onerous computing.
+Therefore, we elect to treat $K$ as a tuning parameter for the MCMC by fixing it at different levels and assessing its impact on prediction as described in \sref{sts:modelselect}.
 Then
 \begin{align}
-   Y_t(\bs) \mid z_{t}(\bs), \sigma_t^2(\bs), P_{tk}, \Theta &= \bX_t(\bs)^\prime \beta + \lambda \sigma_t(\bs) |z_t(\bs)| + \sigma_t(\bs) v_t(\bs) \label{eq:hier}\\
+   Y_t(\bs) \mid z_{t}(\bs), \sigma_t^2(\bs), P_{tk}, \Theta &= \bX_t(\bs)^\top \beta + \lambda \sigma_t(\bs) |z_t(\bs)| + \sigma_t(\bs) v_t(\bs) \label{steq:hier}\\
    z_t(\bs) &= z_{tk} \text{ if } \bs \in P_{tk}\nonumber\\
    \sigma^2_{t}(\bs) &= \sigma^2_{tk} \text{ if } \bs \in P_{tk}\nonumber\\
-   \lambda & \sim N(0, 20)\nonumber\\
+   \lambda & \sim N(0, \sigma^2_\lambda)\nonumber\\
    % \lambda_1 &= \left\{ \begin{array}{ll}
@@
 \end{align}
-where $\Theta = \{\varphi, \nu, \gamma\, \lambda, \beta\}$, and $\Sigma$ is a \Matern covariance matrix as described in Section \ref{s:skewt}.
+where $\Theta = \{\varphi, \nu, \gamma\, \lambda, \beta\}$, and $\Sigma$ is a \Matern covariance matrix as described in \sref{sts:skewt}.
 
-\subsection{Computation}\label{s:comp}
+\subsection{Computation}\label{sts:comp}
 We use MCMC methods to explore the posterior.
@@
 After conditioning on $\lambda$, $z_t(\bs)$ and non-censored observations, $Y_t(\bs)$ has truncated normal full conditionals.
-So we sample $Y_t(\bs) \sim N_{(-\infty, T)}(\bX_t^\prime(\bs) \beta + \lambda | z_t(\bs)|, \bSigma)$.
+So we sample $Y_t(\bs) \sim N_{(-\infty, T)}(\bX_t^\top(\bs) \beta + \lambda | z_t(\bs)|, \bSigma)$.
 
@@
 
-\section{Simulation study}\label{s:simstudy}
+\section{Simulation study}\label{sts:simstudy}
 In this section, we present the results from a simulation study to investigate how the number of partitions and the level of thresholding impact the accuracy of predictions made by the model and to compare with Gaussian and max-stable methods.
 
-\subsection{Design}\label{s:simdesign}
-For all simulation designs, we generated data from model (\ref{eq:partition}) in Section \ref{s:part} using $n_s=144$ sites and $n_t=50$ independent days.
+\subsection{Design}\label{sts:simdesign}
+For all simulation designs, we generated data from model \eref{steq:partition} in \sref{sts:part} using $n_s=144$ sites and $n_t=50$ independent days.
 The sites were generated Uniform$([0, 10] \times [0, 10])$.
-We generated data from 4 different simulation designs:
+We generated data from 5 different simulation designs:
 \begin{enumerate}[leftmargin=3em]
@@
 For design 1, we set $\lambda = 0$.
-For designs 2 and 3, $\lambda = 3$ was used as to simulate moderate skewness, and the $z_t$ were generated as described in (\ref{eq:sitezsig}).
-In designs 1 -- 3, the mean $\bX^\prime \bbeta = 10$ was assumed to be constant across space.
+For designs 2 and 3, $\lambda = 3$ was used as to simulate moderate skewness, and the $z_t$ were generated as described in Section \sref{sts:part}.
+In designs 1 -- 3, the mean $\bX^\top \bbeta = 10$ was assumed to be constant across space.
 In the fourth design, we generated from a spatial max-stable distribution \citep{Reich2012}.
@@
 where $q(0.80)$ is the 80th sample quantile of the data.
-The design matrix $\bX$ includes an intercept with a first-order spatial trend with priors of $\beta_\text{int}$, $\beta_\text{lat}$, $\beta_\text{long},  \iid \text{N}(0, 10)$.
+All methods are fit using a fully-Bayesian approach that simultaneously estimates marginal and spatial dependence parameters.
+The design matrix $\bX$ includes an intercept with a first-order spatial trend with priors of $\beta_\text{int}$, $\beta_\text{lat}$, $\beta_\text{long},  \iid \text{N}(0, 10)$ although only the intercept is used in the data generation.
 The spatial covariance parameters have priors $\log(\nu) \sim \text{N}(-1.2, 1)$, $\gamma \sim \text{Unif}(0, 1)$, $\varphi \sim \text{Unif}(15)$.
-The skewness parameter has prior $\lambda_2 \sim \text{IG}(0.1, 0.1)$.
-The residual variance terms have priors $\sigma^2_t(\bs) \sim \text{IG}(a, b)$, where $a$ has a Gamma$(0.1, 0.1)$ prior and $b$ has a discrete uniform prior on a mesh from $0.1$ to $10$ with spacing of $0.1$.
+The skewness parameter has prior $\lambda \sim N(0, 20)$.
+The residual variance terms have priors $\sigma^2_t(\bs) \sim \text{IG}(a, b)$, where $a$ has a discrete uniform prior on a mesh from $0.2$ to $20$ with spacing of $0.1$ and $b$ has a Gamma$(0.1, 0.1)$ prior.
 The knots have priors $\bw \sim \text{Unif}(\calD)$.
@@
 Each chain of the MCMC ran for 20,000 iterations with a burn-in period of 10,000 iterations.
-Parameters appear to converge properly; however, in the models with multiple partitions (i.e.
-models 4 and 5) it is hard to assess the convergence of $\bw$, $z(\bs)$, and $\sigma^2(\bs)$ because of partition label switching throughout the MCMC.
+For methods that correspond directly to a simulation design (e.g. design 2 and method 2, design 3 and method 4), $\lambda$ and $a$ appear to converge to their true values.
+It should be noted that in the models with multiple partitions (i.e.
+models 4 and 5) it is hard to assess the convergence of $\bw$, $z(\bs)$, and $\sigma^2(\bs)$ because of partition label switching throughout the MCMC; however, we are not interested in these parameters but rather spatial predictions and tail probabilities which converge well.
 
-\subsection{Cross validation}\label{s:modelselect}
+\subsection{Cross validation}\label{sts:modelselect}
 Models were compared using cross validation, with 100 sites used as training sites to fit the models, and 44 sites withheld for testing the predictions.
@@
 
-\subsection{Results}\label{s:simresults}
+\subsection{Results}\label{sts:simresults}
 We compared the Brier scores for exceeding 4 different thresholds for each dataset.
 The thresholds used for the Brier scores are extreme quantiles from the simulated data for $q(0.90)$, $q(0.95)$, $q(0.98)$, and $q(0.99)$.
-Figure \ref{fig:simbrierscores} gives the Brier score relative to the Brier score for the Gaussian method calculated as
+\fref{stfig:simbrierscores} gives the Brier score relative to the Brier score for the Gaussian method calculated as
 \begin{align}
@@
 For Friedman tests that came back with a significant p-value, we conducted a Wilcoxon-Nemenyi-McDonald-Thompson \citep{Hollander2014} test to see which of the methods had different results.
-The full results for the Wilcoxon-Nemenyi-McDonald-Thompson tests are given in \ref{a:pdiffs}.
+The full results for the Wilcoxon-Nemenyi-McDonald-Thompson tests are given in \aref{sta:pdiffs}.
 
@@
 One exception to this is the case when the generative process is max-stable.
-In this case, the max-stable method outperforms our method; however, for predictions at high quantile levels, the differences between the max-stable method and our method decrease.
+In this case, the max-stable method outperforms our method for very extreme quantiles; however, for predictions at high quantile levels, the differences between the max-stable method and our method decrease.
 The non-thresholded methods tend to outperform the thresholded methods, but this is not surprising given that in most cases, the data are generated directly from the model used in the method.
@@
   \caption{Brier scores relative to the Gaussian method for simulation study results. A ratio lower than 1 indicates that the method outperforms the Gaussian method.}
-  \label{fig:simbrierscores}
+  \label{stfig:simbrierscores}
 \end{figure}
 
-\section{Data analysis}\label{s:analysis}
-We consider daily observations of maximum 8-hour ozone measurements for the 31 days of July 2005 at 1,089 Air Quality System (AQS) monitoring sites in the United States as the response (see Figure \ref{fig:ozone-10jul}).
+\section{Data analysis}\label{sts:analysis}
+We consider daily observations of maximum 8-hour ozone measurements for the 31 days of July 2005 at 1,089 Air Quality System (AQS) monitoring sites in the United States as the response (see \fref{stfig:ozone-10jul}).
 For each site, we also have covariate information containing the estimated ozone from the Community Multi-scale Air Quality (CMAQ) modeling system.
-Initially, we fit a linear regression with $\bX_t(\bs) = [1, \text{CMAQ}_t(\bs)]^\prime$. % assuming a mean function of
+Initially, we fit a linear regression with $\bX_t(\bs) = [1, \text{CMAQ}_t(\bs)]^\top$. % assuming a mean function of
 % \begin{align}
-%   \text{E}[Y_i(\bs)] = \beta_0 + \beta_1 \cdot \text{CMAQ}_t(\bs). \label{eq:datamean}
+%   \text{E}[Y_i(\bs)] = \beta_0 + \beta_1 \cdot \text{CMAQ}_t(\bs). \label{steq:datamean}
 % \end{align}
-Figure \ref{fig:ozone-qq} shows a Q-Q plot of the residuals compared to a skew-$t$ distribution with 10 d.f. and $\alpha = 1$.
+\fref{stfig:ozone-qq} shows a Q-Q plot of the residuals compared to a skew-$t$ distribution with 10 d.f. and $\alpha = 1$.
 \begin{figure}
@@
   \caption{Q-Q plot of the residuals for a skew-$t$ distribution with 10 d.f. and $\alpha = 1$ (right)}
-  \label{fig:ozone-qq}
+  \label{stfig:ozone-qq}
 \end{figure}
@@
 Despite the fact that there is only one month of data, we can get some sense of extremal dependence between sites by looking at joint occurrences of high sample quantiles.
-For example, Figure \ref{fig:bivariateozone} suggests there is more agreement between sites that are close to one another than sites that are far from one another.
+For example, \fref{stfig:bivariateozone} suggests there is more agreement between sites that are close to one another than sites that are far from one another.
 \begin{figure}
@@
   \caption{Daily quantiles for two monitoring locations near Columbus, OH (left) and daily quantiles for a monitoring location in Los Angeles, CA and Columbus, OH (right)}
-  \label{fig:bivariateozone}
+  \label{stfig:bivariateozone}
 \end{figure}
@@
 % Finally, $\widehat{\chi}(t)$ is averaged over all sites at each of the three threshold quantiles.
-% The $\widehat{\chi}_c(h)$ and $\widehat{\chi}_c(t)$ plots in Figure \ref{fig:chi-st} show the estimated spatial and temporal dependence of the residuals for the ozone data at three quantile levels $q(0.90), q(0.95)$, and $q(0.99)$.
+% The $\widehat{\chi}_c(h)$ and $\widehat{\chi}_c(t)$ plots in Figure \ref{ststfig:chi-st} show the estimated spatial and temporal dependence of the residuals for the ozone data at three quantile levels $q(0.90), q(0.95)$, and $q(0.99)$.
 
@@
 %   \caption{$\widehat{\chi}_c(h)$ plot for the residuals (left). $\widehat{\chi}_c(t)$ plot for the residuals (right).}
-%   \label{fig:chi-st}
+%   \label{stfig:chi-st}
 % \end{figure}
@@
 Finally, as a comparison to max-stable methods, we fit the model using the hierarchical max-stable model of \citet{Reich2012} with the data thresholded at $T = 75$.
-All methods assume the mean function given in (\ref{eq:datamean}).
+All methods assume $\bX_t(\bs) = [1, \text{CMAQ}_t(\bs)]^\top$.
 To ensure that the max-stable method runs in a reasonable amount of time, we use a stratified sub-sample of 800 sites.
-We conduct two-fold cross validation using 400 training sites and 400 validation sites as described in Section \ref{s:modelselect}
+We conduct two-fold cross validation using 400 training sites and 400 validation sites as described in \sref{sts:modelselect}
 
 Each chain of the MCMC ran for 30,000 iterations with a burn-in period of 25,000 iterations.
+We use the same priors for the spatial covariance parameters, skewness parameter, and knots as in the simulation study.
+The prior for the residual variance terms is $\sigma^2_t(\bs) \sim text{IG}(a, b)$ where $a$ is the same as the simulation study, but $b$ has a Gamma$(1, 1)$ prior.
 Parameters appear to converge properly; however, as before, for models with multiple partitions it is hard to assess the convergence of $\bw$, $z(\bs)$, and $\sigma^2(\bs)$ because of partition label switching throughout the MCMC.
@@
 At a particular threshold or quantile level, the model that fits the best is the one with the lowest score.
-We then compute the relative (to Gaussian) Brier scores (see Section \ref{s:simresults}) to compare each model.
+We then compute the relative (to Gaussian) Brier scores (see \sref{sts:simresults}) to compare each model.
 
-\subsection{Results}\label{s:results}
+\subsection{Results}\label{sts:results}
 The results suggest that the skew-$t$, thresholded, partitioned, and time series models all give an improvement in predictions over the Gaussian model, whereas the max-stable method results in relative Brier scores between 1.07 and 1.15 indicating poorer performance than the Gaussian model.
-The plots in Figure \ref{fig:bs-ozone} show the relative Brier scores for time-series and non-time-series models, using $K = $ 1, 7, and 15 knots at thresholds $T = $ 0, 50, and 75 ppb.
+The plots in \fref{stfig:bs-ozone} show the relative Brier scores for time-series and non-time-series models, using $K = $ 1, 7, and 15 knots at thresholds $T = $ 0, 50, and 75 ppb.
 Most of the models perform similarly across all the Brier scores; however, for single-partition models without thresholding, performance tends to diminish in the extreme quantiles.
@@
 In particular, for all extreme quantiles, selecting a moderate number of knots (e.g. $K = 5, \ldots, 10$) tends to give the best results.
-Table \ref{tbl:ozoneresults} shows the best two models for selected extreme quantiles.
+\tref{sttbl:ozoneresults} shows the best two models for selected extreme quantiles.
 
-We illustrate the predictive capability of our model in Figure \ref{fig:ozoneq99} by plotting the 99th quantile for South Carolina and Georgia, a subset of the spatial domain, in order to study local features.
+We illustrate the predictive capability of our model in \fref{stfig:ozoneq99} by plotting the 99th quantile for South Carolina and Georgia, a subset of the spatial domain, in order to study local features.
 The four methods used are
@@
   \caption{Relative Brier scores for time-series models (left) and non-time-series models (right). Relative brier score for the max-stable model is between 1.07 and 1.15}
-  \label{fig:bs-ozone}
+  \label{stfig:bs-ozone}
 \end{figure}
@@
   \caption{Top two performing models for ozone analysis at extreme quantiles with Relative Brier score}
-  \label{tbl:ozoneresults}
+  \label{sttbl:ozoneresults}
   \centering
@@
   \caption{Panels (a) -- (d) give the posterior predictive $\widehat{q}(0.99)$ for the month of July under four different models, panel (e) gives the difference between $\widehat{q}(0.99)$ in panels (d) and (a), panel (f) gives the difference between $\widehat{q}(0.99)$ in panels (d) and (b).}
-  \label{fig:ozoneq99}
+  \label{stfig:ozoneq99}
 \end{figure}
 
-\section{Discussion}\label{s:con}
+\section{Discussion}\label{sts:con}
 In this paper we propose a new threshold exceedance approach for spatiotemporal modeling based on the skew-$t$ process.
@@
 
-\label{lastpage}
+\label{stlastpage}
 

### 2016-06-06 LaTeX/supplemental_rev.tex
manuscript changes for revisions.

--- 2016-05-16 LaTeX/supplemental.tex
@@
 \input{mycommands.sty}
+\renewcommand{\fref}[1]{Web Figure~\ref{#1}}
 
@@
 % \appendix
-\renewcommand{\thesection}{Web Appendix \Alph{section}}
+\renewcommand{\thesection}{Web Appendix~\Alph{section}}
 % \renewcommand{\tablename}{Web Table}
@@
 
-\section{MCMC details} \label{a:mcmc}
-The MCMC sampling for the model \ref{s:hier} is done using {\tt R} (http://www.r-project.org).
-Whenever possible, we select conjugate priors (see Appendix \ref{a:posterior}); however, for some of the parameters, no conjugate prior distributions exist.
+\section{MCMC details} \label{sta:mcmc}
+The MCMC sampling for the model in \sref{sts:hier} is done using {\tt R} (http://www.r-project.org).
+Whenever possible, we select conjugate priors (see \aref{sta:posterior}); however, for some of the parameters, no conjugate prior distributions exist.
 For these parameters, we use a random walk Metropolis-Hastings update step.
@@
 For each day, we update the spatial knot locations, $\bw_1, \ldots, \bw_K$, using a Metropolis-Hastings block update.
-Because the spatial domain is bounded, we generate candidate knots using the transformed knots $\bw^*_1, \ldots, \bw^*_K$ (see section \ref{s:temporal}) and a random walk bivariate Gaussian candidate distribution
+Because the spatial domain is bounded, we generate candidate knots using the transformed knots $\bw^*_1, \ldots, \bw^*_K$ (see \sref{sts:temporal}) and a random walk bivariate Gaussian candidate distribution
 \begin{align*}
@@
 \end{align*}
-where $l$ is the likelihood given in (\ref{eq:hier}), and $p(\cdot)$ is the prior either taken from the time series given in (\ref{s:temporal}) or assumed to be uniform over $\calD$.
+where $l$ is the likelihood given in \eref{steq:hier}, and $p(\cdot)$ is the prior either taken from the time series given in \eref{sts:temporal} or assumed to be uniform over $\calD$.
 The candidate knots are accepted with probability $\min\{R, 1\}$.
@@
 \subsection*{Spatial random effects}
-If there is no temporal dependence amongst the observations, we use a Gibbs update for $z_{tk}$, and the posterior distribution is given in \ref{a:posterior}.
+If there is no temporal dependence amongst the observations, we use a Gibbs update for $z_{tk}$, and the posterior distribution is given in \aref{sta:posterior}.
 If there is temporal dependence amongst the observations, then we update $z_{tk}$ using a Metropolis-Hastings update.
-Because this model uses $|z_{tk}|$, we generate candidate random effects using the $z^*_{tk}$ (see Section \ref{s:temporal}) and a random walk Gaussian candidate distribution
+Because this model uses $|z_{tk}|$, we generate candidate random effects using the $z^*_{tk}$ (see \sref{sts:temporal}) and a random walk Gaussian candidate distribution
 \begin{align*}
@@
 \end{align*}
-where $p[\cdot]$ is the prior taken from the time series given in Section \ref{s:temporal}.
+where $p[\cdot]$ is the prior taken from the time series given in \sref{sts:temporal}.
 The candidate is accepted with probability $\min\{R, 1\}$.
@@
 When there is more than one site in a partition, then we update $\sigma^2_{tk}$ using a Metropolis-Hastings update.
-First, we generate a candidate for $\sigma^2_{tk}$ using an IG$(a^*/s, b^*/s)$ candidate distribution in an independence Metropolis-Hastings update where $a^* = (n_{tk} + 1) / 2 + a$, $b^* = [Y_{tk}^T \Sigma^{-1}_{tk} Y_{tk} + z_{tk}^2] / 2 + b$, $n_{tk}$ is the number of sites in partition $k$ on day $t$, and $Y_{tk}$ and $\Sigma^{-1}_{tk}$ are the observations and precision matrix for partition $k$ on day $t$.
+First, we generate a candidate for $\sigma^2_{tk}$ using an IG$(a^*/s, b^*/s)$ candidate distribution in an independence Metropolis-Hastings update where $a^* = (n_{tk} + 1) / 2 + a$, $b^* = [Y_{tk}^\prime \Sigma^{-1}_{tk} Y_{tk} + z_{tk}^2] / 2 + b$, $n_{tk}$ is the number of sites in partition $k$ on day $t$, and $Y_{tk}$ and $\Sigma^{-1}_{tk}$ are the observations and precision matrix for partition $k$ on day $t$.
 The acceptance ratio is
@@
 \end{align*}
-where $p[\cdot]$ is the prior either taken from the time series given in Section \ref{s:temporal} or assumed to be IG$(a, b)$, and $c[\cdot]$ is the candidate distribution.
+where $p[\cdot]$ is the prior either taken from the time series given in \sref{sts:temporal} or assumed to be IG$(a, b)$, and $c[\cdot]$ is the candidate distribution.
 The candidate is accepted with probability $\min\{R, 1\}$.
@@
 
-\section{Posterior distributions} \label{a:posterior}
+\section{Posterior distributions} \label{sta:posterior}
 
-\subsection*{Conditional posterior of $z_{tk} \mid \ldots $}\label{s:mvcondu}
+\subsection*{Conditional posterior of $z_{tk} \mid \ldots $}\label{sts:mvcondu}
 If knots are independent over days, then the conditional posterior distribution of $|z_{tk}|$ is conjugate.
@@
             R_2
-        \end{array} \right)^T
+        \end{array} \right)^\prime
         \left( \begin{array}{cc}
@@
 \begin{align*}
-    \mu_l &= \lambda ( R_1^T \Omega_{11} + R_2^T \Omega_{21} )\bOne\\
-    \Lambda_l &= \lambda^2 \bOne^T \Omega_{11} \bOne + \frac{ 1 }{ \sigma^2_l }.
+    \mu_l &= \lambda ( R_1^\prime \Omega_{11} + R_2^\prime \Omega_{21} )\bOne\\
+    \Lambda_l &= \lambda^2 \bOne^\prime \Omega_{11} \bOne + \frac{ 1 }{ \sigma^2_l }.
 \end{align*}
@@
 
-\subsection*{Conditional posterior of $\beta \mid \ldots$}\label{s:betapost}
+\subsection*{Conditional posterior of $\beta \mid \ldots$}\label{sts:betapost}
 Let $\beta \sim \mbox{N}_{p}(0, \Lambda_0)$ where $\Lambda_0$ is a precision matrix. Then
 \begin{align*}
-    \pi(\beta \mid \ldots) & \propto \exp \left\{ - \frac{ 1 }{ 2 } \beta^T \Lambda_0 \beta - \frac{ 1 }{ 2 } \sum_{t = 1 }^T [\bY_t - X_t \beta - \lambda |z_t|]^T \Omega [\bY_t - X_t \beta - \lambda |z_t|] \right\}\\
-     & \propto \exp \left\{ -\frac{ 1 }{ 2 } \left[ \beta^T \Lambda_\beta \beta  - 2 \sum_{ t = 1 }^T [ \beta^T X_t^T \Omega (\bY_t - \lambda |z_t| )] \right] \right\}\\
+    \pi(\beta \mid \ldots) & \propto \exp \left\{ - \frac{ 1 }{ 2 } \beta^\prime \Lambda_0 \beta - \frac{ 1 }{ 2 } \sum_{t = 1}^{n_t} [\bY_t - X_t \beta - \lambda |z_t|]^\prime \Omega [\bY_t - X_t \beta - \lambda |z_t|] \right\}\\
+     & \propto \exp \left\{ -\frac{ 1 }{ 2 } \left[ \beta^\prime \Lambda_\beta \beta  - 2 \sum_{t = 1}^{n_t} [ \beta^\prime X_t^\prime \Omega (\bY_t - \lambda |z_t| )] \right] \right\}\\
      & \propto \mbox{N} ( \Lambda_\beta^{-1} \mu_\beta , \Lambda_\beta^{-1})
@@
 \begin{align*}
-    \mu_\beta &= \sum_{t=1}^{T} \left[ X_t^T \Omega (\bY_t - \lambda |z_t|) \right]\\
-    \Lambda_\beta &= \Lambda_0 + \sum_{ t = 1 }^{ T} X_t^T \Omega X_t.
+    \mu_\beta &= \sum_{t = 1}^{n_t} \left[ X_t^\prime \Omega (\bY_t - \lambda |z_t|) \right]\\
+    \Lambda_\beta &= \Lambda_0 + \sum_{t = 1}^{n_t} X_t^\prime \Omega X_t.
 \end{align*}
 
-\subsection*{Conditional posterior of $\sigma^2 \mid \ldots$}\label{s:sigpost}
+\subsection*{Conditional posterior of $\sigma^2 \mid \ldots$}\label{sts:sigpost}
 In the case where $L = 1$ and temporal dependence is negligible, then $\sigma^2$ has a conjugate posterior distribution.
@@
 \begin{align*}
-    \pi(\sigma^2 \mid \ldots) & \propto (\sigma^2)^{ -\alpha_0 - 1 / 2 - n / 2 - 1} \exp \left\{ -\frac{\beta_0}{\sigma^2} - \frac{ |z|^2 }{2 \sigma^2} - \frac{ (\bY - \bmu)^T \Sigma^{-1} (\bY - \bmu) }{2 \sigma^2} \right\} \\
-    & \propto (\sigma^2)^{ -\alpha_0 - 1 / 2 - n / 2 - 1} \exp \left\{ - \frac{ 1 }{ \sigma^2 } \left[\beta_0 + \frac{ |z|^2 }{ 2 } + \frac{ 1 }{ 2 } (\bY - \bmu)^T \Sigma^{-1} (\bY - \mu) \right] \right\} \\
+    \pi(\sigma^2 \mid \ldots) & \propto (\sigma^2)^{ -\alpha_0 - 1 / 2 - n / 2 - 1} \exp \left\{ -\frac{\beta_0}{\sigma^2} - \frac{ |z|^2 }{2 \sigma^2} - \frac{ (\bY - \bmu)^\prime \Sigma^{-1} (\bY - \bmu) }{2 \sigma^2} \right\} \\
+    & \propto (\sigma^2)^{ -\alpha_0 - 1 / 2 - n / 2 - 1} \exp \left\{ - \frac{ 1 }{ \sigma^2 } \left[\beta_0 + \frac{ |z|^2 }{ 2 } + \frac{ 1 }{ 2 } (\bY - \bmu)^\prime \Sigma^{-1} (\bY - \mu) \right] \right\} \\
     & \propto \mbox{IG} (\alpha^*, \beta^*)
@@
     \alpha^* &= \alpha_0 + \frac{1}{2} + \frac{n}{2} \\
-    \beta^* &= \beta_0 + \frac{ |z|^2 }{ 2 } + \frac{ 1 }{ 2 }(\bY - \bmu)^T \Sigma^{-1} (\bY - \bmu).
+    \beta^* &= \beta_0 + \frac{ |z|^2 }{ 2 } + \frac{ 1 }{ 2 }(\bY - \bmu)^\prime \Sigma^{-1} (\bY - \bmu).
 \end{align*}
@@
 
-\subsection*{Conditional posterior of $\lambda \mid \ldots$}\label{s:lambdapost}
+\subsection*{Conditional posterior of $\lambda \mid \ldots$}\label{sts:lambdapost}
+\hl{TODO: Still needs to be updated for single $\lambda$}\\
 For convergence purposes we model $\lambda = \lambda_1 \lambda_2$ where
@@
 \begin{align*}
-  \pi(\lambda_2 \mid \ldots) &\propto {\lambda_2^2}^{(-\alpha_\lambda - 1)} \exp\left\{ - \frac{ \beta_\lambda }{ \lambda_2^2} \right\} \prod_{t = 1}^{T} \prod_{k = 1}^K \frac{1} {\lambda_2} \exp\left\{ - \frac{ z^2_{tk} }{ 2 \lambda_2^2  \sigma_{tk})^2} \right\} \\
-   & \propto {\lambda_2^2}^{(-\alpha_\lambda - kt - 1)} \exp \left\{ -\frac{1}{\lambda^2_2} \left[ \beta_\lambda + \frac{ z^2 }{2 \sigma^2_{tk}} \right] \right\}
+  \pi(\lambda_2 \mid \ldots) &\propto {\lambda_2^2}^{(-\alpha_\lambda - 1)} \exp\left\{ - \frac{ \beta_\lambda }{ \lambda_2^2} \right\} \prod_{t = 1}^{n_t} \prod_{k = 1}^K \frac{1} {\lambda_2} \exp\left\{ - \frac{ z^2_{tk} }{ 2 \lambda_2^2  \sigma_{tk})^2} \right\} \\
+   & \propto {\lambda_2^2}^{(-\alpha_\lambda - K*n_t - 1)} \exp \left\{ -\frac{1}{\lambda^2_2} \left[ \beta_\lambda + \frac{ z^2 }{2 \sigma^2_{tk}} \right] \right\}
 \end{align*}
-Then $\lambda_2 \mid \ldots \sim IG \left(\alpha_\lambda + kt, \beta_\lambda + \frac{ z^2 }{2 \sigma^2_{tk}} \right)$
+Then $\lambda_2 \mid \ldots \sim IG \left(\alpha_\lambda + K * n_t, \beta_\lambda + \frac{ z^2 }{2 \sigma^2_{tk}} \right)$
 
-\section{Proof that $\lim_{h \rightarrow \infty} \pi(h) = 0$} \label{a:proofsamepartition}
+\section{Proof that $\lim_{h \rightarrow \infty} \pi(h) = 0$} \label{sta:proofsamepartition}
 Let $c$ be the midpoint of $\bs_1$ and $\bs_2$.
@@
 \begin{align*}
-  \mu(A) = \lambda_{PP} |A| = \lambda_{PP} \pi \left(\frac{h}{2}\right)^2 = \lambda_{PPA}^* h^2.
+  \mu_{PP}(A) = \lambda_{PP} |A| = \lambda_{PP} \pi \left(\frac{h}{2}\right)^2 = \lambda_{PPA}^* h^2.
 \end{align*}
-Consider a partition of $A$ into four regions, $B_1$, $B_2$, $R_1$, $R_2$ as seen in Web Figure \ref{fig:hpp}.
+Consider a partition of $A$ into four regions, $B_1$, $B_2$, $R_1$, $R_2$ as seen in \fref{stfig:hpp}.
 \begin{figure}
@@
   \caption{Illustration of the partition of $A$.}
-  \label{fig:hpp}
+  \label{stfig:hpp}
 \end{figure}
-Let $N_j$ be the number of knots in $B_j, j = 1, 2$.
+Let $N_i$ be the number of knots in $B_i$ and $L_i = l$ if $\bs_i \in P_l$ for $i = 1, 2$.
 Then
 \begin{align}
-  P(\bs_1 \in P_i, \bs_2 \in P_{j \neq i}) \ge P(N_1 > 0, N_2 > 0)
+  P(L_1 \neq L_2) \ge P(N_1 > 0, N_2 > 0)
 \end{align}
@@
 \begin{align}
-  \mu(B_1) &= \lambda_{PP} |B_1| = \lambda_{PP} \frac{h^2}{4} \left(\frac{2 \pi}{3} - \frac{\sqrt{3}}{2} \right) \nonumber \\
+  \mu_{PP}(B_1) &= \lambda_{PP} |B_1| = \lambda_{PP} \frac{h^2}{4} \left(\frac{2 \pi}{3} - \frac{\sqrt{3}}{2} \right) \nonumber \\
        &= \lambda^*_{PPB1} h^2.
@@
 \begin{align}
-  P(\bs_1 \in P_i, \bs_2 \in P_{j \neq i}) >= P(N_1 > 0)^2 = [1 - P(N_1 = 0)]^2 = [1 - \exp\left(-\lambda^*_{PPB1} h^2\right)]^2
+  P(L_1 \neq L_2) >= P(N_1 > 0)^2 = [1 - P(N_1 = 0)]^2 = [1 - \exp\left(-\lambda^*_{PPB1} h^2\right)]^2
 \end{align}
@@
 
-\section{Skew-$t$ distribution} \label{a:skewt}
+\section{Skew-$t$ distribution} \label{sta:skewt}
 \subsection*{Univariate skew-$t$ distribution}
@@
 \begin{align}
-  f_z(\bz) &= 2 t_d(\bz; \bar{\bOmega}, \eta) T \left[ \balpha^T \bz \sqrt{ \frac{\eta + d}{\nu + Q(\bz)} }; \eta + d\right] \\
+  f_z(\bz) &= 2 t_d(\bz; \bar{\bOmega}, \eta) T \left[ \balpha^\prime \bz \sqrt{ \frac{\eta + d}{\nu + Q(\bz)} }; \eta + d\right] \\
   \bz &= \bomega^{-1}(\by - \xi)
 \end{align}
-where $t_d(\bz; \bar{\bOmega}, \eta)$ is a $d$-dimensional Student's $t$-distribution with scale matrix $\bar{\bOmega}$ and degrees of freedom $\eta$, $Q(z) = \bz^T \bar{\Omega}^{-1}\bz$ and $T(\cdot; \eta)$ denotes the univariate Student's $t$ distribution function with $\eta$ degrees of freedom \citep{Azzalini2014}.
+where $t_d(\bz; \bar{\bOmega}, \eta)$ is a $d$-dimensional Student's $t$-distribution with scale matrix $\bar{\bOmega}$ and degrees of freedom $\eta$, $Q(z) = \bz^\prime \bar{\Omega}^{-1}\bz$ and $T(\cdot; \eta)$ denotes the univariate Student's $t$ distribution function with $\eta$ degrees of freedom \citep{Azzalini2014}.
 
 \subsection*{Extremal dependence}
-For a bivariate skew-$t$ random variable $\bY = [Y(\bs), Y(\bt)]^T$, the $\chi(h)$ statistic \citep{Padoan2011} is given by
-\begin{align} \label{eq:chiskew-t}
+For a bivariate skew-$t$ random variable $\bY = [Y(\bs), Y(\bt)]^\prime$, the $\chi(h)$ statistic \citep{Padoan2011} is given by
+\begin{align} \label{steq:chiskew-t}
   \chi(h) = \bar{F}_{\text{EST}}\left\{ \frac{[x_1^{1 / \eta} - \varrho(h)] \sqrt{\eta + 1} }{\sqrt{1 - \varrho(h)^2}}; 0, 1, \alpha_1, \tau_1, \eta + 1 \right\} + \bar{F}_{\text{EST}}\left\{ \frac{ [x_2^{1 / \eta} - \varrho(h)] \sqrt{\eta + 1} }{ \sqrt{1 - \varrho(h)^2} }; 0, 1, \alpha_2, \tau_2, \eta + 1 \right\},
@@
 \subsection*{Proof that $\lim_{h \rightarrow \infty} \chi(h) > 0$}
-Consider the bivariate distribution of $\bY = [Y(\bs), Y(\bt)]^T$, with $\varrho(h)$ given by (\ref{eq:matern}).
+Consider the bivariate distribution of $\bY = [Y(\bs), Y(\bt)]^\prime$, with $\varrho(h)$ given by \eref{steq:matern}.
 So, $\lim_{h \rightarrow \infty} \varrho(h) = 0$.
@@
 
-\section{Simulation study pairwise difference results} \label{a:pdiffs}
+\section{Comparisons with other parameterizations} \label{sta:otherparams}
+Various forms of multivariate skew-normal and skew-$t$ distributions have been proposed in the literature.
+In this section, we make a connection between our parameterization in \eref{steq:fullmodel} of the main text and another popular version.
+\citet{Azzalini2014} and \citet{Beranger2016} define a skew-normal process as
+\begin{align}
+  \tilde{X}(\bs) = \tilde{\lambda}|z| + (1 - \tilde{\lambda}^2)^{1 / 2} v(\bs)
+\end{align}
+where $\tilde{\lambda} \in (-1, 1)$, $z \sim N(0, 1)$, and $v(\bs)$ is a Gaussian process with mean zero, variance one, and spatial correlation function $\rho$.
+To extend this to the skew-$t$ distribution, \citet{Azzalini2003} take $\tilde{Y}(\bs) = W\tilde{X}(\bs)$ where $W^{-2} \sim $ Gamma$(a / 2, a / 2)$.
+Returning to the proposed parameterization (with $\bbeta = 0$), let $W^{-2} = \frac{2b}{a}\sigma^{-2} \sim$ Gamma$(a / 2, a / 2)$ so that \eref{steq:fullmodel} becomes
+\begin{align}
+  Y(\bs) = W \left[ \lambda \left(\frac{a}{2b}\right)^{1 / 2} |z| + \left(\frac{a}{2b}\right)^{1 / 2} v(\bs) \right].
+\end{align}
+Clearly setting $b = \frac{a}{(1 - \tilde{\lambda}^2)} > 0$, and $\lambda = \tilde{\lambda} / (1 - \tilde{\lambda}^2)^{1 / 2} \in (-\infty, \infty)$ resolves the difference in parameterizations.
+We note that our parameterization has three parameters $(a, b, \lambda)$ compared to the two parameters of the alternative parameterization $(a, \tilde{\lambda})$.
+Since we have assumed that both $v(\bs)$ and $z$ have unit scale, the additional $b$ parameter in our parameterization is required to control the precision.
+% However, if we introduce an overall scale parameter $c > 0$ into the alternative parameterization so that $\tilde{Y}(\bs) = c W \tilde{X}(\bs)$, then the two models remain equivalent by setting $a = \nu$, $b = \frac{\nu}{c^2 (1 - \tilde{\lambda}^2)}$, and $\lambda = \tilde{\lambda} / (1 - \tilde{\lambda}^2)^{1 / 2}$.
+
+\section{Simulation study pairwise difference results} \label{sta:pdiffs}
 The following tables show the methods that have significantly different Brier scores when using a Wilcoxon-Nemenyi-McDonald-Thompson test.
@@
   \caption{Setting 1 -- Gaussian marginal, $K = 1$ knot}
-  \label{tbl:gaussim}
+  \label{sttbl:gaussim}
   \begin{tabular}{|l|ccc|ccc|cccc|ccc|}
@@
   \caption{Setting 2 -- Skew-$t$ marginal, $K = 1$ knot}
-  \label{tbl:st1sim}
+  \label{sttbl:st1sim}
   \begin{tabular}{|l|ccccc|cccc|cccc|ccc|}
@@
   \caption{Setting 3 -- Skew-$t$ marginal, $K = 5$ knots}
-  \label{tbl:st5sim}
+  \label{sttbl:st5sim}
   \begin{tabular}{|l|ccc|cccc|ccc|ccc|}
@@
   \caption{Setting 4 -- Max-stable}
-  \label{tbl:mssim}
+  \label{sttbl:mssim}
   \begin{tabular}{|l|cccc|cccc|ccc|ccc|}
@@
   \caption{Setting 5 -- Transformation below $T = q(0.80)$}
-  \label{tbl:transsim}
+  \label{sttbl:transsim}
   \begin{tabular}{|l|cccc|ccc|cccc|cccc|}

### 2016-06-18 LaTeX/response.tex
manuscript and plot updates

@@
 
-\item In the simulations, you adopted a prior distribution for the $a$ and $b$ parameters in $\sigma_t^2 (\bs) \sim IG(a, b)$ as $a \sim $Gamma$(0.1, 0.1)$ and $b$ with a discrete distribution over a mesh from 0.1 to 10 with spacing $0.1$.
+\item In the simulations, you adopted a prior distribution for the $a$ and $b$ parameters in $\sigma_t^2 (\bs) \sim IG(a, b)$ as $a \sim$ Gamma$(0.1, 0.1)$ and $b$ with a discrete distribution over a mesh from 0.1 to 10 with spacing $0.1$.
 The parameter $\sigma_t^2(\bs)$ is clearly a crucial parameter as it controls the scale.
@@
 \begin{response}
-  In the original submission, the priors for $a$ and $b$ were inadvertently switched, so $a$ should have had a discrete distribution over a mesh from 0.1 to 10 with spacing $0.1$, and $b \sim $Gamma$(0.1, 0.1)$.
+  In the original submission, the priors for $a$ and $b$ were inadvertently switched, so $a$ should have had a discrete distribution over a mesh from 0.1 to 10 with spacing $0.1$, and $b \sim$ Gamma$(0.1, 0.1)$.
   With the reparameterization given in the response to Major Comment 1, this actually means that the prior for $a$ is now a discrete distribution over a mesh from 0.2 to 20 with spacing $0.2$.
-  As we now clarify in the specification of these priors, $a$ is the scale-free shape parameter, and so this prior should be uninformative in all application.
+  As we now clarify in the specification of these priors, $a$ is the scale-free shape parameter, and so this prior should be uninformative in all applications.
   In contrast, $b$ is a precision parameter, so its prior should be adapted to the scale of the response.
@@
 \begin{response}
-  \hl{TODO:}
-  Using simulated data, we have evidence to suggest that the model exhibits extremal dependence provided that the site $\bs$ is in the same partition for day $t$ and $t + 1$.
-  However, we are unable to give a formal proof that our temporal construction maintains extremal dependence.
+  Using simulated data, we have evidence to suggest that the model exhibits extremal dependence over time.
+  We are unable to give a formal proof of this, but we have included simulated $\chi$ plots for temporal dependence in \aref{sta:temporal}.
 \end{response}
@@
     We have added clarification in Section \sref{sts:temporal} to clarify that the partition is created in the $\bw$ space as opposed to the $\bw^*$ space.
-    \hl{Is this something we should be concerned about?}
+    % \hl{Is this something we should be concerned about?}
   \end{response}
@@
   \begin{response}
-    \hl{TODO:}
     Yes, we can see how this could lead to confusion.
-    In the revision, we now refer to the ``threshold'' as the value we select in the model-fitting process, i.e. $T$ in (\eref{steq:Yt}).
-    We also bolstered the description of the trade-off involved in selecting $T$ (see AE comment 4) below (\eref{steq:Yt}).
+    In the revision, we now refer to the ``threshold'' as the value we select in the model-fitting process, i.e. $T$ in \eref{steq:Yt}.
+    We also bolstered the description of the trade-off involved in selecting $T$ (see AE comment 4) below \eref{steq:Yt}.
     As mentioned in the response to the previous comment, we also want to compute probabilities of extreme events (e.g. ozone exceeding 75ppb).
@@
   \begin{response}
-    \hl{TODO:}
     Yes, as noted in the response to AE (major comment 5), we have evidence using simulated data that our temporal construction maintains extremal dependence.
-    This now appears \hl{where it will appear}.
+    This now appears in \aref{sta:temporal}.
     Thank you both for the suggestion!
@@
   \begin{response}
-    \hl{Partly done:}
     We have added clarification for these points in the description of the simulation study.
@@
   \begin{response}
-    \hl{I'm going to try to add in something about whether or not the parameters converge for skew-$t$ with 1 knot}
     As with any Bayesian mixture model, the labels of the partitions are arbitrary.
@@
     However, since the quantities of interest (spatial predictions and probability of extreme events) are defined marginally over the partitioning, label-switching does not affect convergence for these quantities.
+    We have bolstered this explanation and included coverage probabilities for $\lambda$ and the degrees of freedom for design (2) and method (2) (skew-$t$ with 1 knot) at the end of \sref{sts:simdesign}.
   \end{response}
@@
     \begin{align*}
-      P_k = \{s \in \calD : k = \argmin_i ||s−w_i||\}, \quad \text{with} \quad k= 1,2.
+      P_k = \{s \in \calD : k = \argmin_i ||s - w_i||\}, \quad \text{with} \quad k= 1,2.
     \end{align*}
@@
   \begin{response}
-    As mentioned in \sref{sts:simdesign}, some of the designs for the simulation study match the methods used to fit the datasets (e.g. design 2 and method 2, design 3 and method 4).
-    We have provided additional clarification in \sref{sts:simdesign} that for these models, we do appear to correctly estimate the skewness parameters and degrees of freedom.
+    In simulation design (2), we generate the data using a skew-$t$ distribution with a single knot, and in method (2), we fit the data using a skew-$t$ distribution with a single knot.
+    We have provided additional clarification in \sref{sts:simdesign} that in this case, we do appear to correctly estimate the skewness parameters and degrees of freedom.
     We have also added clarification that the difficulty in assessing convergence is only relevant for models with multiple partitions (i.e.
@@
   \begin{response}
-    \hl{Partially done. I got the $^\top$ for transposes, but I still need to change $T \rightarrow m$ for time points.}
     The revision uses $^\top$ for transposes and $n_t$ for the number of time points.

### 2016-06-18 LaTeX/skewt_rev.tex
manuscript and plot updates

@@
 \begin{abstract}
-To assess the compliance of air quality regulations, the Environmental Protection Agency (EPA) must know if a site exceeds a pre-specified threshold.
-In the case of ozone, the threshold for compliance is fixed at 75 parts per billion, which is high, but not extreme at all locations.
+To assess the compliance of air quality regulations, the Environmental Protection Agency (EPA) must know if a site exceeds a pre-specified level.
+In the case of ozone, the level for compliance is fixed at 75 parts per billion, which is high, but not extreme at all locations.
 We present a new method based on the spatial skew-$t$ process.
@@
 Finally, our model allows for high-dimensional Bayesian inference that is comparable in speed to traditional geostatistical methods for large datasets.
-We apply our method to an ozone analysis for July 2005, and find that our model improves over both Gaussian and max-stable methods in terms of predicting exceedances over a fixed threshold.
+We apply our method to an ozone analysis for July 2005, and find that our model improves over both Gaussian and max-stable methods in terms of predicting exceedances of a high level.
 \end{abstract}
@@
 As a result, the Environmental Protection Agency (EPA) has developed a set of standards to help reduce air pollution thereby improving air quality.
-Our study is motivated by an air pollution application where the focus is not on the average behavior, but instead the behavior over a fixed threshold determined by government regulation.
+Our study is motivated by an air pollution application where the focus is not on the average behavior, but instead the behavior over a high level determined by government regulation.
 More specifically, we consider the case of compliance for ozone.
@@
 As there are more of these values, they can unduly influence the distribution at the higher levels about which we are interested.
-Our inference method will only use data which exceed a threshold, and will impute data below the threshold, thereby tailoring the fit to the levels of interest.
+Our inference method will only use data which exceed a pre-selected threshold, and will impute data below the threshold, thereby tailoring the fit to the levels of interest.
 Second, likelihood-based spatial modeling typically assumes a Gaussian process, which is appropriate when mean behavior is of interest.
@@
 We propose a new spatiotemporal threshold exceedance model based on the skew-$t$ process \citep{Padoan2011}.
-Our model is a threshold exceedance model for the multivariate skew-$t$ distribution for a fixed threshold.
+\hl{Need to add in more references here to skew-$t$ used in a spatial setting.}
+Our model is a threshold exceedance model for the multivariate skew-$t$ distribution using a fixed threshold.
 In this setting, we describe the threshold as fixed because it is specified in advance by regulatory compliance.
 This differs from the more traditional extremes literature where a threshold is selected to be the value beyond which an extremal model is appropriate for the data.
-We use a skew-$t$ distribution because of its flexibility to model asymmetry and heavy-tailed data with the aim of predicting the probability of exceeding a high fixed threshold at an unobserved location.
-\hl{Need to add in more references here to skew-$t$ used in a spatial setting.}
+We use a skew-$t$ distribution because of its flexibility to model asymmetry and heavy-tailed data with the aim of predicting the probability of exceeding a high fixed level at an unobserved location.
 
@@
 \end{align}
-$\bX(\bs)$ is the observed covariate vector at site $\bs$, $\bbeta$ is the $p$-vector of regression parameters, $\lambda \in \calR$ is a parameter controlling skew, $z \sim N(0, 1)$, $\sigma^2 \sim \text{IG}(a / 2, b)$ is random scale parameter, IG is the distribution function of an inverse gamma random variable, $a$ is the degrees of freedom, $b$ controls the precision of the process, and $v(\bs)$ is a spatial Gaussian process with mean zero, variance one, and a positive definite correlation function $\text{Cor} \left[Y(\bs_1), Y(\bs_2)\right] = \rho(\bs_1, \bs_2)$.
+$\bX(\bs)$ is the observed covariate vector at site $\bs$, $\bbeta$ is the $p$-vector of regression parameters, $\lambda \in \calR$ is a parameter controlling skew, $z \sim N(0, 1)$, $\sigma^2 \sim \text{IG}(a / 2, b / 2)$ is random scale parameter, IG is the distribution function of an inverse gamma random variable, $a$ is the degrees of freedom, $b$ controls the precision of the process, and $v(\bs)$ is a spatial Gaussian process with mean zero, variance one, and a positive definite correlation function $\text{Cor} \left[Y(\bs_1), Y(\bs_2)\right] = \rho(\bs_1, \bs_2)$.
 Although any positive definite correlation function could be used, we choose to use the stationary isotropic \Matern correlation with
@@
 be the censored observation at site $\bs$ where $Y(\bs)$ is the uncensored observation, $\delta(\bs) = I[Y(\bs) > T]$, and $T$ is a pre-specified threshold value.
+As our goal is to predict exceedances above a high level $L$, we should select a value for $T \le L$.
+For example, in predicting ozone exceedances, we might set $T = 50$ppb in order to predict exceedances of $L = 75$ppb.
 Selecting $T$ too small may lead to bias in estimating the tail parameters; selecting $T$ to large increases variance.
@@
 In other words, $P_k$ is composed of all sites for which the closest knot is $\bw_k$.
-For all $\bs \in P_k$, with $k = 1, 2, \ldots, K$, the functions $z(\bs)$ and $\sigma(\bs)$ are equal to the constants $z_k$ and $\sigma_k$ respectively, and the $z_k$ and $\sigma^2_k$ are distributed as $z_k \iid N(0, 1)$ and $\sigma^2_k \iid$ IG$(a, b)$.
+For all $\bs \in P_k$, with $k = 1, 2, \ldots, K$, the functions $z(\bs)$ and $\sigma(\bs)$ are equal to the constants $z_k$ and $\sigma_k$ respectively, and the $z_k$ and $\sigma^2_k$ are distributed as $z_k \iid N(0, 1)$ and $\sigma^2_k \iid$ IG$(a / 2, b / 2)$.
 So, within each partition, $Y(\bs)$ follows the spatial skew-$t$ process defined in \sref{sts:spatialskew}.
@@
 \end{align}
-where $I(\cdot)$ is an indicator function, $\chi_{\text{skew-}t}(h)$ is the $\chi$ statistic for a skew-$t$ process given in \eref{steq:chiskew-t}, $\chi_{\text{Gaus}}(h)$ is the $\chi$ statistic for a Gaussian process, and $h = ||\bs_1 - \bs_2||$.
+where $I(\cdot)$ is an indicator function, $\chi_{\text{skew-}t}(h)$ is the $\chi$ statistic for a skew-$t$ process given in equation \eref{steq:chiskew-t} of \aref{sta:skewt}, $\chi_{\text{Gaus}}(h)$ is the $\chi$ statistic for a Gaussian process, and $h = ||\bs_1 - \bs_2||$.
 Therefore, sites in different subregions are asymptotically independent because $\chi_{\text{Gaus}}(h) = 0$ for all $h$.
@@
 \end{align}
-where $t \in \{1, \ldots, T\}$ denotes the day of each observation.
+where $t \in \{1, \ldots, n_t\}$ denotes the day of each observation.
 Let \hbox{$\bw_{tk} = (w_{tk1}, w_{tk2})$} be a spatial knot on day $t$, and let $\bw_{t1}, \ldots, \bw_{tK}$ be a collection of spatial knots on day $t$.
@@
 These are stationary time series models with marginal distributions \hbox{$\bw^*_{k} \sim N_2(0, 1)$}, \hbox{$z^*_{k} \sim N(0, 1)$}, and \hbox{$\sigma^{2*}_{k} \sim N(0, 1)$}.
-After transformation back to the original space, $\bw_{tk} \sim \text{Unif}(\calD)$, $z_{tk} \sim HN(0, 1)$, and $\sigma^2_{tk} \sim \text{IG}(a, b)$.
+After transformation back to the original space, $\bw_{tk} \sim \text{Unif}(\calD)$, $z_{tk} \sim HN(0, 1)$, and $\sigma^2_{tk} \sim \text{IG}(a / 2, b / 2)$.
 We then create the partition for day $t$ using $\bw_{t1}, \ldots, \bw_{tk}$.
@@
 One approach would be to allow $K$ to be unknown and follow a Poisson process prior, but this would lead to onerous computing.
-Therefore, we elect to treat $K$ as a tuning parameter for the MCMC by fixing it at different levels and assessing its impact on prediction as described in \sref{sts:modelselect}.
+Therefore, we elect to treat $K$ as a tuning parameter for the MCMC by fixing it at different values and assessing its impact on prediction as described in \sref{sts:modelselect}.
 Then
@@
 \section{Simulation study}\label{sts:simstudy}
-In this section, we present the results from a simulation study to investigate how the number of partitions and the level of thresholding impact the accuracy of predictions made by the model and to compare with Gaussian and max-stable methods.
+In this section, we present the results from a simulation study to investigate how the number of partitions and the amount of thresholding impact the accuracy of predictions made by the model and to compare with Gaussian and max-stable methods.
 
@@
   \item Skew-$t$ marginal, $K=5$ knots
-  \item Max-stable
+  \item Max-stable with GEV(1, 1, 0.2) marginals \citep{Reich2012}
   % \item Transformation below $T = q(0.80)$
@@
 The skewness parameter has prior $\lambda \sim N(0, 20)$.
-The residual variance terms have priors $\sigma^2_t(\bs) \sim \text{IG}(a, b)$, where $a$ has a discrete uniform prior on a mesh from $0.2$ to $20$ with spacing of $0.1$ and $b$ has a Gamma$(0.1, 0.1)$ prior.
+The residual variance terms have priors $\sigma^2_t(\bs) \sim \text{IG}(a / 2, b / 2)$, where $a$ has a discrete uniform prior on a mesh from $0.2$ to $20$ with spacing of $0.1$ and $b$ has a Gamma$(0.1, 0.1)$ prior.
 The knots have priors $\bw \sim \text{Unif}(\calD)$.
@@
 Each chain of the MCMC ran for 20,000 iterations with a burn-in period of 10,000 iterations.
-For methods that correspond directly to a simulation design (e.g. design 2 and method 2, design 3 and method 4), $\lambda$ and $a$ appear to converge to their true values.
+Although the goal of this simulation study is not to assess parameter estimation, for design (2) and method (2), the $\lambda$ estimate and and $a$ appear to converge to their true values (96\% coverage and 90\% coverage respectively).
 It should be noted that in the models with multiple partitions (i.e.
@@
 Models were compared using cross validation, with 100 sites used as training sites to fit the models, and 44 sites withheld for testing the predictions.
-Because one of the primary goals of this model is to predict exceedances over a fixed threshold, we use Brier scores to compare the models \citep{Gneiting2007}.
-The Brier score for predicting exceedance of a threshold $c$ is given by $[e(c) - P(c)]^2$ where $e(c) = I[y>c]$ is an indicator function indicating that a test set value, $y$, has exceeded the threshold, $c$, and $P(c)$ is the predicted probability of exceeding $c$.
+Because one of the primary goals of this model is to predict exceedances over a high level, we use Brier scores to compare the models \citep{Gneiting2007}.
+The Brier score for predicting exceedance of a level $L$ is given by $[e(L) - P(L)]^2$ where $e(L) = I[y>L]$ is an indicator function indicating that a test set value, $y$, has exceeded the level, $L$, and $P(L)$ is the predicted probability of exceeding $L$.
 We average the Brier scores over all test sites and days.
@@
 \subsection{Results}\label{sts:simresults}
-We compared the Brier scores for exceeding 4 different thresholds for each dataset.
-The thresholds used for the Brier scores are extreme quantiles from the simulated data for $q(0.90)$, $q(0.95)$, $q(0.98)$, and $q(0.99)$.
+We compared the Brier scores for exceeding 4 different high levels for each dataset.
+The levels used for the Brier scores are extreme quantiles from the simulated data for $L = q(0.90)$, $q(0.95)$, $q(0.98)$, $q(0.99)$.
 \fref{stfig:simbrierscores} gives the Brier score relative to the Brier score for the Gaussian method calculated as
@@
 \end{align}
-We analyzed the results for the simulation study using a Friedman test at $\alpha = 0.05$ to see if at least one method had a significantly different Brier score.
+We analyzed the results for the simulation study using a Friedman \citep{Hollander2014} test at $\alpha = 0.05$ to see if at least one method had a significantly different Brier score.
 For Friedman tests that came back with a significant p-value, we conducted a Wilcoxon-Nemenyi-McDonald-Thompson \citep{Hollander2014} test to see which of the methods had different results.
@@
 
+In general, we find that when the method to fit the data matches the data generation scheme, there is some improvement over other methods.
 The results show that when the data are generated from a Gaussian process, our method performs comparably to a Gaussian approach.
 In general, when the underlying process is not Gaussian, our method results in an improvement over both the max-stable and Gaussian methods.
-One exception to this is the case when the generative process is max-stable.
-In this case, the max-stable method outperforms our method for very extreme quantiles; however, for predictions at high quantile levels, the differences between the max-stable method and our method decrease.
-The non-thresholded methods tend to outperform the thresholded methods, but this is not surprising given that in most cases, the data are generated directly from the model used in the method.
+We also see that the non-thresholded methods tend to outperform the thresholded methods, but this is not surprising given that in most cases, the data are generated directly from the model used in the method.
+Finally, in the case where the data are generated using a Brown-Resnick process, we find that our method is competitive with using a max-stable model.
 % Finally, for setting 5, although the thresholded version of the single-partition model tends to perform the best across all of the extreme quantiles, the difference between the thresholded and non-thresholded methods is no longer significant in the more extreme quantiles.
@@
 We use the same priors for the spatial covariance parameters, skewness parameter, and knots as in the simulation study.
-The prior for the residual variance terms is $\sigma^2_t(\bs) \sim text{IG}(a, b)$ where $a$ is the same as the simulation study, but $b$ has a Gamma$(1, 1)$ prior.
+The prior for the residual variance terms is $\sigma^2_t(\bs) \sim \text{IG}(a / 2, b / 2)$ where $a$ is the same as the simulation study, but $b$ has a Gamma$(1, 1)$ prior.
 Parameters appear to converge properly; however, as before, for models with multiple partitions it is hard to assess the convergence of $\bw$, $z(\bs)$, and $\sigma^2(\bs)$ because of partition label switching throughout the MCMC.
 For each model, Brier scores were averaged over all sites and days to obtain a single Brier score for each dataset.
-At a particular threshold or quantile level, the model that fits the best is the one with the lowest score.
+At a particular level, the model that fits the best is the one with the lowest score.
 We then compute the relative (to Gaussian) Brier scores (see \sref{sts:simresults}) to compare each model.
@@
 \subsection{Results}\label{sts:results}
-The results suggest that the skew-$t$, thresholded, partitioned, and time series models all give an improvement in predictions over the Gaussian model, whereas the max-stable method results in relative Brier scores between 1.07 and 1.15 indicating poorer performance than the Gaussian model.
+The results suggest that the skew-$t$, thresholded, partitioned, and time series models all give an improvement in predictions over the Gaussian model, whereas the max-stable method results in relative Brier scores between 1.13 and 1.18 indicating poorer performance than the Gaussian model.
 The plots in \fref{stfig:bs-ozone} show the relative Brier scores for time-series and non-time-series models, using $K = $ 1, 7, and 15 knots at thresholds $T = $ 0, 50, and 75 ppb.
@@
   \includegraphics[width=\linewidth]{plots/bs-ozone.pdf}
-  \caption{Relative Brier scores for time-series models (left) and non-time-series models (right). Relative brier score for the max-stable model is between 1.07 and 1.15}
+  \caption{Relative Brier scores for time-series models (left) and non-time-series models (right). Relative brier score for the max-stable model is between 1.13 and 1.18}
   \label{stfig:bs-ozone}
@@
   \small
-  \caption{Top two performing models for ozone analysis at extreme quantiles with Relative Brier score}
+  \caption{Top two performing models for predicting ozone exceedance of level $L$ with Relative Brier score}
   \label{sttbl:ozoneresults}
@@
     \hline
-    $q(0.90)$  & No time series & $K=7$  & $T=0$  & BS: 0.980 &
-                 No time series & $K=9$  & $T=0$  & BS: 0.980 \\
-    $q(0.95)$  & No time series & $K=15$ & $T=50$ & BS: 0.970 &
-                 No time series & $K=9$  & $T=50$ & BS: 0.970\\
-    $q(0.98)$  & No time series & $K=5$  & $T=50$ & BS: 0.945 &
-                 No time series & $K=10$ & $T=50$ & BS: 0.946\\
-    $q(0.99)$  & Time series    & $K=10$ & $T=75$ & BS: 0.912 &
-                 Time series    & $K=6$  & $T=75$ & BS: 0.913\\
-    $q(0.995)$ & Time series    & $K=6$  & $T=75$ & BS: 0.917 &
-                 Time series    & $K=10$ & $T=75$ & BS: 0.918\\
+    $L = q(0.90)$  & No time series & $K=6$  & $T=50$  & BS: 0.992 &
+                 No time series & $K=1$  & $T=0$  & BS: 0.992 \\
+    $L = q(0.95)$  & Time series & $K=5$ & $T=50$ & BS: 0.988 &
+                 No time series & $K=6$  & $T=50$ & BS: 0.989\\
+    $L = q(0.98)$  & Time series & $K=7$  & $T=50$ & BS: 0.973 &
+                 Time series & $K=5$ & $T=50$ & BS: 0.975\\
+    $L = q(0.99)$  & No time series    & $K=8$ & $T=0$ & BS: 0.946 &
+                 Time series    & $K=9$  & $T=75$ & BS: 0.947\\
+    $L = q(0.995)$ & No time series    & $K=8$  & $T=0$ & BS: 0.951 &
+                 Time series    & $K=9$ & $T=75$ & BS: 0.956\\
     \hline

### 2016-06-18 LaTeX/supplemental_rev.tex
manuscript and plot updates

@@
 where ${\bw^*_k}^{(r - 1)}$ is the location for the transformed knot at MCMC iteration $r - 1$, $s$ is a tuning parameter, and $I_2$ is an identity matrix.
+Let $\bY_t = [Y(\bs_1), \ldots, Y(\bs_n)]$ be the vector of observed responses at each site for day $t$.
 After candidates have been generated for all $K$ knots, the acceptance ratio is
 \begin{align*}
-  R = \left\{ \frac{ l[ Y_t(\bs | \bw_1^{(c)}, \ldots, \bw_K^{(c)}, \ldots)] }{l[ Y_t(\bs | \bw_1^{(r - 1)}, \ldots, \bw_K^{(r - 1)}, \ldots)]} \right\} \times \left\{ \frac{ \prod_{k = 1}^{K}\phi(\bw_k^{(c)})}{ \prod_{k = 1}^{K}\phi(\bw_k^{(r - 1)})} \right\} \times \left\{ \frac{ \prod_{k = 1}^{K} p({\bw^*_k}^{(c)})}{ \prod_{k = 1}^{K} p({\bw^*_k}^{(r - 1)})}\right\}
+  R = \left\{ \frac{ l[ \bY_t | \bw_1^{(c)}, \ldots, \bw_K^{(c)}, \ldots] }{l[ \bY_t | \bw_1^{(r - 1)}, \ldots, \bw_K^{(r - 1)}, \ldots]} \right\} \times \left\{ \frac{ \prod_{k = 1}^{K}\phi(\bw_k^{(c)})}{ \prod_{k = 1}^{K}\phi(\bw_k^{(r - 1)})} \right\} \times \left\{ \frac{ \prod_{k = 1}^{K} p({\bw^*_k}^{(c)})}{ \prod_{k = 1}^{K} p({\bw^*_k}^{(r - 1)})}\right\}
 \end{align*}
-where $l$ is the likelihood given in \eref{steq:hier}, and $p(\cdot)$ is the prior either taken from the time series given in \eref{sts:temporal} or assumed to be uniform over $\calD$.
+where $l$ is the likelihood given in \eref{steq:hier}, and $p(\cdot)$ is the prior either taken from the time series (see \sref{sts:temporal}) or assumed to be uniform over $\calD$.
 The candidate knots are accepted with probability $\min\{R, 1\}$.
@@
 \begin{align*}
-  R = \left\{ \frac{ l[Y_t(\bs) | z_{tk}^{(c)}, \ldots] }{ l[Y_t(\bs) | z_{tk}^{(r - 1)}]} \right\} \times \left\{ \frac{ p[ z_{tk}^{(c)} ] }{ p[ z_{tk}^{(r - 1)}]}\right\}
+  R = \left\{ \frac{ l[\bY_t | z_{tk}^{(c)}, \ldots] }{ l[\bY_t | z_{tk}^{(r - 1)}]} \right\} \times \left\{ \frac{ p[ z_{tk}^{(c)} ] }{ p[ z_{tk}^{(r - 1)}]}\right\}
 \end{align*}
@@
 When there is more than one site in a partition, then we update $\sigma^2_{tk}$ using a Metropolis-Hastings update.
-First, we generate a candidate for $\sigma^2_{tk}$ using an IG$(a^*/s, b^*/s)$ candidate distribution in an independence Metropolis-Hastings update where $a^* = (n_{tk} + 1) / 2 + a$, $b^* = [Y_{tk}^\prime \Sigma^{-1}_{tk} Y_{tk} + z_{tk}^2] / 2 + b$, $n_{tk}$ is the number of sites in partition $k$ on day $t$, and $Y_{tk}$ and $\Sigma^{-1}_{tk}$ are the observations and precision matrix for partition $k$ on day $t$.
+First, we generate a candidate for $\sigma^2_{tk}$ using an IG$(a^*/s, b^*/s)$ candidate distribution in an independence Metropolis-Hastings update where $a^* = (n_{tk} + 1) / 2 + a$, $b^* = [\bY_{tk}^\top \Sigma^{-1}_{tk} \bY_{tk} + z_{tk}^2] / 2 + b$, $n_{tk}$ is the number of sites in partition $k$ on day $t$, and $\bY_{tk}$ and $\Sigma^{-1}_{tk}$ are the observations and precision matrix for partition $k$ on day $t$.
 The acceptance ratio is
@@
   R = \left\{
-    \frac{ l[Y_t(\bs) | {\sigma^2_{tk}}^{(c)}, \ldots] }
-         { l[Y_t(\bs) | {\sigma^2_{tk}}^{(r - 1)}]}
+    \frac{ l[\bY_t | {\sigma^2_{tk}}^{(c)}, \ldots] }
+         { l[\bY_t | {\sigma^2_{tk}}^{(r - 1)}]}
     \right\} \times \left\{
@@
 \begin{align*}
-	R = \left\{ \frac{ \prod_{t = 1}^{T} l[Y_t(\bs) | \rho^{(c)}, \nu^{(c)}, \gamma^{(c)}, \ldots] }{\prod_{t = 1}^{T} l[Y_t(\bs) | \rho^{(r-1)}, \nu^{(r-1)}, \gamma^{(r-1)}, \ldots] } \right\} \times \left\{ \frac{ p[\rho^{(c)}] }{ p[\rho^{(r - 1)] } } \right\} \times \left\{ \frac{ p[\nu^{(c)}] }{ p[\nu^{(r - 1)}] } \right\} \times \left\{ \frac{ p[ \gamma^{(c)} ] }{ p[\gamma^{(r - 1)} ] } \right\}.
+	R = \left\{ \frac{ \prod_{t = 1}^{n_t} l[Y_t(\bs) | \rho^{(c)}, \nu^{(c)}, \gamma^{(c)}, \ldots] }{\prod_{t = 1}^{n_t} l[Y_t(\bs) | \rho^{(r-1)}, \nu^{(r-1)}, \gamma^{(r-1)}, \ldots] } \right\} \times \left\{ \frac{ p[\rho^{(c)}] }{ p[\rho^{(r - 1)] } } \right\} \times \left\{ \frac{ p[\nu^{(c)}] }{ p[\nu^{(r - 1)}] } \right\} \times \left\{ \frac{ p[ \gamma^{(c)} ] }{ p[\gamma^{(r - 1)} ] } \right\}.
 \end{align*}
@@
 If knots are independent over days, then the conditional posterior distribution of $|z_{tk}|$ is conjugate.
-For simplicity, drop the subscript $t$, let $\tilde{z}_{tk} = |z_{tk}|$, and define
+For simplicity, drop the subscript $t$, let $\tilde{z}_{k} = |z_{k}|$, $\tilde{\bz}_{k^c}$ be the vector of $[|z(\bs_1)|, \ldots, |z(\bs_n)|]$ for $\bs \notin P_k$, $\bX = [\bX(\bs_1), \ldots, \bX(\bs_n)]^\top$, let $\bY_k$ and $\bX_k$ be the observations and covariate measurements for $\bs \in P_k$, and let $\bY_{k^c}$ and $\bX_{k^c}$ be the observations and covariate measurements for $\bs \notin P_k$ and define
 \begin{align*}
-R(\bs) = \left\{
+\bR = \left\{
     \begin{array}{ll}
-        Y(\bs) - X(\bs) \beta &s \in P_l\\[1em]
-        Y(\bs) - X(\bs) \beta - \lambda \tilde{z}(\bs) \qquad & s \notin P_l
+        \bY_k - \bX_k \bbeta & \bs \in P_k\\[1em]
+        \bY_{k^c} - \bX_{k^c} \bbeta - \lambda \tilde{\bz}_{k^c} \qquad & \bs \notin P_k
     \end{array}
@@
 \begin{align*}
-    R_1 &= \text{the vector of } R(\bs) \text{ for } s \in P_l \\
-    R_2 &= \text{the vector of } R(\bs) \text{ for } s \notin P_l \\
+    \bR_1 &= \text{the vector of } \bR \text{ for } \bs \in P_k \\
+    \bR_2 &= \text{the vector of } \bR \text{ for } \bs \notin P_k \\
     \Omega &= \Sigma^{-1}.
@@
 \begin{align*}
-    \pi(z_l | \ldots) &\propto \exp \left\{ -\frac{ 1 }{ 2 } \left[
+    \pi(z_k | \ldots) &\propto \exp \left\{ -\frac{ 1 }{ 2 } \left[
         \left( \begin{array}{c}
-            R_1 - \lambda \tilde{z}_l \bOne\\
-            R_2
-        \end{array} \right)^\prime
+            \bR_1 - \lambda \tilde{z}_k \bOne\\
+            \bR_2
+        \end{array} \right)^\top
         \left( \begin{array}{cc}
@@
         \left( \begin{array}{c}
-            R_1 - \lambda \tilde{z}_l \bOne\\
-            R_2
+            \bR_1 - \lambda \tilde{z}_k \bOne\\
+            \bR_2
         \end{array} \right)
-        +  \frac{ {\tilde{z}_l}^2 }{ \sigma_l^2 }\right]
-    \right\} I(z_l > 0) \\
-        &\propto \exp \left\{ -\frac{ 1 }{ 2 } \left[ \Lambda_l {\tilde{z}_l}^2 - 2 \mu_l \tilde{z}_l \right] \right\}
+        +  \frac{ {\tilde{z}_k}^2 }{ \sigma_k^2 }\right]
+    \right\} I(z_k > 0) \\
+        &\propto \exp \left\{ -\frac{ 1 }{ 2 } \left[ \Lambda_k {\tilde{z}_k}^2 - 2 \mu_k \tilde{z}_k \right] \right\}
 \end{align*}
@@
 \begin{align*}
-    \mu_l &= \lambda ( R_1^\prime \Omega_{11} + R_2^\prime \Omega_{21} )\bOne\\
-    \Lambda_l &= \lambda^2 \bOne^\prime \Omega_{11} \bOne + \frac{ 1 }{ \sigma^2_l }.
+    \mu_k &= \lambda ( \bR_1^\top \Omega_{11} + \bR_2^\top \Omega_{21} )\bOne\\
+    \Lambda_k &= \lambda^2 \bOne^\top \Omega_{11} \bOne + \frac{ 1 }{ \sigma^2_k }.
 \end{align*}
-Then $\tilde{Z}_l | \ldots \sim N_{(0, \infty)} (\Lambda_l^{-1} \mu_l, \Lambda_l^{-1})$
+Then $\tilde{z}_k | \ldots \sim N_{(0, \infty)} (\Lambda_k^{-1} \mu_k, \Lambda_k^{-1})$
 
-\subsection*{Conditional posterior of $\beta \mid \ldots$}\label{sts:betapost}
-Let $\beta \sim \mbox{N}_{p}(0, \Lambda_0)$ where $\Lambda_0$ is a precision matrix. Then
+\subsection*{Conditional posterior of $\bbeta, \lambda \mid \ldots$}\label{sts:betapost}
+For models that do not include a skewness parameter, we update $\bbeta$ as follows.
+Let $\bbeta \sim \mbox{N}_{p}(0, \Lambda_0)$ where $\Lambda_0$ is a precision matrix.
+Then
 \begin{align*}
-    \pi(\beta \mid \ldots) & \propto \exp \left\{ - \frac{ 1 }{ 2 } \beta^\prime \Lambda_0 \beta - \frac{ 1 }{ 2 } \sum_{t = 1}^{n_t} [\bY_t - X_t \beta - \lambda |z_t|]^\prime \Omega [\bY_t - X_t \beta - \lambda |z_t|] \right\}\\
-     & \propto \exp \left\{ -\frac{ 1 }{ 2 } \left[ \beta^\prime \Lambda_\beta \beta  - 2 \sum_{t = 1}^{n_t} [ \beta^\prime X_t^\prime \Omega (\bY_t - \lambda |z_t| )] \right] \right\}\\
+    \pi(\bbeta \mid \ldots) & \propto \exp \left\{ - \frac{ 1 }{ 2 } \bbeta^\top \Lambda_0 \bbeta - \frac{ 1 }{ 2 } \sum_{t = 1}^{n_t} [\bY_t -\bX_t \bbeta]^\top \Omega [\bY_t - \bX_t \bbeta] \right\} \\
+     & \propto \exp \left\{ -\frac{ 1 }{ 2 } \left[ \bbeta^\top \Lambda_\beta \bbeta  - 2 \sum_{t = 1}^{n_t} (\bbeta^\top \bX_t^\top \Omega \bY_t) \right] \right\} \\
      & \propto \mbox{N} ( \Lambda_\beta^{-1} \mu_\beta , \Lambda_\beta^{-1})
@@
 \begin{align*}
-    \mu_\beta &= \sum_{t = 1}^{n_t} \left[ X_t^\prime \Omega (\bY_t - \lambda |z_t|) \right]\\
-    \Lambda_\beta &= \Lambda_0 + \sum_{t = 1}^{n_t} X_t^\prime \Omega X_t.
+    \mu_\beta &= \sum_{t = 1}^{n_t} \bX_t^\top \Omega \bY_t \\
+    \Lambda_\beta &= \Lambda_0 + \sum_{t = 1}^{n_t} \bX_t^\top \Omega \bX_t.
 \end{align*}
+For models that do include a skewness parameter, a simple augmentation of the covariate matrix $\bX$ and parameter vector $\bbeta$ allows for a block update of both $\bbeta$ and $\lambda$.
+Let $\bX^*_{t} = [\bX_t, |\bz_{t}|]$ where $\bz_t = [z(\bs_1), \ldots, z(\bs_{n})]^\top$ and let $\bbeta^* = (\beta_1, \ldots, \beta_p, \lambda)^\top$.
+So to incorporate the $N(0, \sigma^2_\lambda)$ prior on $\lambda$, let $\bbeta^* \sim \mbox{N}_{p+1}(0, \Lambda^*_0)$ where
+\begin{align*}
+  \Lambda^*_0 = \left( \begin{array}{cc}
+    \Lambda_0 & 0 \\
+    0         & \sigma^{-2}_\lambda
+  \end{array}\right).
+\end{align*}
+Then the update for both $\bbeta$ and $\lambda$ is done using the conjugate prior given above with $\bX_t = \bX_t^*$ and $\bbeta = \bbeta^*$
 
@@
 In the case where $L = 1$ and temporal dependence is negligible, then $\sigma^2$ has a conjugate posterior distribution.
-Let $\sigma_t^2 \iid \mbox{IG}(\alpha_0, \beta_0)$. For simplicity, drop the subscript $t$. Then
+Let $\sigma_t^2 \iid \mbox{IG}(\alpha_0 / 2, \beta_0 / 2)$. For simplicity, drop the subscript $t$. Then
 \begin{align*}
-    \pi(\sigma^2 \mid \ldots) & \propto (\sigma^2)^{ -\alpha_0 - 1 / 2 - n / 2 - 1} \exp \left\{ -\frac{\beta_0}{\sigma^2} - \frac{ |z|^2 }{2 \sigma^2} - \frac{ (\bY - \bmu)^\prime \Sigma^{-1} (\bY - \bmu) }{2 \sigma^2} \right\} \\
-    & \propto (\sigma^2)^{ -\alpha_0 - 1 / 2 - n / 2 - 1} \exp \left\{ - \frac{ 1 }{ \sigma^2 } \left[\beta_0 + \frac{ |z|^2 }{ 2 } + \frac{ 1 }{ 2 } (\bY - \bmu)^\prime \Sigma^{-1} (\bY - \mu) \right] \right\} \\
+    \pi(\sigma^2 \mid \ldots) & \propto (\sigma^2)^{ -\alpha_0 / 2 - 1 / 2 - n / 2 - 1} \exp \left\{ -\frac{\beta_0}{2\sigma^2} - \frac{ |z|^2 }{2 \sigma^2} - \frac{ (\bY - \bmu)^\top \Sigma^{-1} (\bY - \bmu) }{2 \sigma^2} \right\} \\
+    & \propto (\sigma^2)^{ -(\alpha_0 - 1 - n) / 2 - 1} \exp \left\{ - \frac{ 1 }{ 2 \sigma^2 } \left[\beta_0 + |z|^2 + (\bY - \bmu)^\top \Sigma^{-1} (\bY - \bmu) \right] \right\} \\
     & \propto \mbox{IG} (\alpha^*, \beta^*)
@@
 \begin{align*}
-    \alpha^* &= \alpha_0 + \frac{1}{2} + \frac{n}{2} \\
-    \beta^* &= \beta_0 + \frac{ |z|^2 }{ 2 } + \frac{ 1 }{ 2 }(\bY - \bmu)^\prime \Sigma^{-1} (\bY - \bmu).
+    \alpha^* &= \frac{\alpha_0 + 1 + n}{2} \\
+    \beta^* &= \frac{1}{2} \left[ \beta_0 + |z|^2 + (\bY - \bmu)^\top \Sigma^{-1} (\bY - \bmu) \right].
 \end{align*}
-In the case that $L > 1$, a random walk Metropolis Hastings step will be used to update $\sigma^2_{lt}$.
-
-\subsection*{Conditional posterior of $\lambda \mid \ldots$}\label{sts:lambdapost}
-\hl{TODO: Still needs to be updated for single $\lambda$}\\
-For convergence purposes we model $\lambda = \lambda_1 \lambda_2$ where
-\begin{align}
-  \lambda_1 &= \left\{ \begin{array}{ll}
-      +1 \quad & \text{w.p.} 0.5\\
-      -1 \quad & \text{w.p.} 0.5
-   \end{array}\right.\\
-   \lambda^2_2 & \sim IG(\alpha_\lambda, \beta_\lambda).\\
-\end{align}
-Then
-\begin{align*}
-  \pi(\lambda_2 \mid \ldots) &\propto {\lambda_2^2}^{(-\alpha_\lambda - 1)} \exp\left\{ - \frac{ \beta_\lambda }{ \lambda_2^2} \right\} \prod_{t = 1}^{n_t} \prod_{k = 1}^K \frac{1} {\lambda_2} \exp\left\{ - \frac{ z^2_{tk} }{ 2 \lambda_2^2  \sigma_{tk})^2} \right\} \\
-   & \propto {\lambda_2^2}^{(-\alpha_\lambda - K*n_t - 1)} \exp \left\{ -\frac{1}{\lambda^2_2} \left[ \beta_\lambda + \frac{ z^2 }{2 \sigma^2_{tk}} \right] \right\}
-\end{align*}
-Then $\lambda_2 \mid \ldots \sim IG \left(\alpha_\lambda + K * n_t, \beta_\lambda + \frac{ z^2 }{2 \sigma^2_{tk}} \right)$
+In the case that $K > 1$, a random walk Metropolis Hastings step will be used to update $\sigma^2_{kt}$.
 
@@
 \begin{align}
-  f_z(\bz) &= 2 t_d(\bz; \bar{\bOmega}, \eta) T \left[ \balpha^\prime \bz \sqrt{ \frac{\eta + d}{\nu + Q(\bz)} }; \eta + d\right] \\
+  f_z(\bz) &= 2 t_d(\bz; \bar{\bOmega}, \eta) T \left[ \balpha^\top \bz \sqrt{ \frac{\eta + d}{\nu + Q(\bz)} }; \eta + d\right] \\
   \bz &= \bomega^{-1}(\by - \xi)
 \end{align}
-where $t_d(\bz; \bar{\bOmega}, \eta)$ is a $d$-dimensional Student's $t$-distribution with scale matrix $\bar{\bOmega}$ and degrees of freedom $\eta$, $Q(z) = \bz^\prime \bar{\Omega}^{-1}\bz$ and $T(\cdot; \eta)$ denotes the univariate Student's $t$ distribution function with $\eta$ degrees of freedom \citep{Azzalini2014}.
+where $t_d(\bz; \bar{\bOmega}, \eta)$ is a $d$-dimensional Student's $t$-distribution with scale matrix $\bar{\bOmega}$ and degrees of freedom $\eta$, $Q(z) = \bz^\top \bar{\Omega}^{-1}\bz$ and $T(\cdot; \eta)$ denotes the univariate Student's $t$ distribution function with $\eta$ degrees of freedom \citep{Azzalini2014}.
 
 \subsection*{Extremal dependence}
-For a bivariate skew-$t$ random variable $\bY = [Y(\bs), Y(\bt)]^\prime$, the $\chi(h)$ statistic \citep{Padoan2011} is given by
+For a bivariate skew-$t$ random variable $\bY = [Y(\bs), Y(\bt)]^\top$, the $\chi(h)$ statistic \citep{Padoan2011} is given by
 \begin{align} \label{steq:chiskew-t}
@@
 \subsection*{Proof that $\lim_{h \rightarrow \infty} \chi(h) > 0$}
-Consider the bivariate distribution of $\bY = [Y(\bs), Y(\bt)]^\prime$, with $\varrho(h)$ given by \eref{steq:matern}.
+Consider the bivariate distribution of $\bY = [Y(\bs), Y(\bt)]^\top$, with $\varrho(h)$ given by \eref{steq:matern}.
 So, $\lim_{h \rightarrow \infty} \varrho(h) = 0$.
@@
 To extend this to the skew-$t$ distribution, \citet{Azzalini2003} take $\tilde{Y}(\bs) = W\tilde{X}(\bs)$ where $W^{-2} \sim $ Gamma$(a / 2, a / 2)$.
-Returning to the proposed parameterization (with $\bbeta = 0$), let $W^{-2} = \frac{2b}{a}\sigma^{-2} \sim$ Gamma$(a / 2, a / 2)$ so that \eref{steq:fullmodel} becomes
+Returning to the proposed parameterization (with $\bbeta = 0$), let $W^{-2} = \frac{b}{a}\sigma^{-2} \sim$ Gamma$(a / 2, a / 2)$ so that \eref{steq:fullmodel} in the manuscript becomes
 \begin{align}
-  Y(\bs) = W \left[ \lambda \left(\frac{a}{2b}\right)^{1 / 2} |z| + \left(\frac{a}{2b}\right)^{1 / 2} v(\bs) \right].
+  Y(\bs) = W \left[ \lambda \left(\frac{b}{a}\right)^{1 / 2} |z| + \left(\frac{b}{a}\right)^{1 / 2} v(\bs) \right].
 \end{align}
-Clearly setting $b = \frac{a}{(1 - \tilde{\lambda}^2)} > 0$, and $\lambda = \tilde{\lambda} / (1 - \tilde{\lambda}^2)^{1 / 2} \in (-\infty, \infty)$ resolves the difference in parameterizations.
+Clearly setting $b = a (1 - \tilde{\lambda}^2) > 0$, and $\lambda = \tilde{\lambda} / (1 - \tilde{\lambda}^2)^{1 / 2} \in (-\infty, \infty)$ resolves the difference in parameterizations.
 We note that our parameterization has three parameters $(a, b, \lambda)$ compared to the two parameters of the alternative parameterization $(a, \tilde{\lambda})$.
@@
 % However, if we introduce an overall scale parameter $c > 0$ into the alternative parameterization so that $\tilde{Y}(\bs) = c W \tilde{X}(\bs)$, then the two models remain equivalent by setting $a = \nu$, $b = \frac{\nu}{c^2 (1 - \tilde{\lambda}^2)}$, and $\lambda = \tilde{\lambda} / (1 - \tilde{\lambda}^2)^{1 / 2}$.
+
+\section{Temporal dependence} \label{sta:temporal}
+It is very challenging to derive an analytical expression the temporal extremal dependence at a single site $\bs$.
+However, using simulated data, we have evidence to suggest that the model does exhibits temporal extremal dependence.
+To demonstrate that our model maintains temporal extremal dependence, we generate lag-$m$ observations for $m = 1, 3, 5, 10$ from our model setting $\phi_w = \phi_z = \phi_\sigma = \varphi$, for $\varphi = 0, 0.02, 0.04, \ldots, 1$.
+To estimate the lag-$m$ chi-statistic $\chi(m)$ we first estimate the lag-$m$ $F$-madogram $\nu_F(m)$ \citep{Cooley2006} using $\hat{\nu}_F(m) = \frac{1}{2n} \sum_{i = 1}^n | \hat{F}(y_{0}) - \hat{F}(y_{m})|$ where $\hat{F}(\cdot)$ represents an empirical CDF and $y_m$ is the lag-$m$ observation.
+The $F$-madogram is related to the $\chi$ statistic as follows
+\begin{align}
+  \chi = 2 - \frac{1 + 2 \nu_F}{1 - 2 \nu_F}.
+\end{align}
+\fref{stfig:chiphi} suggests that the extremal dependence increases as $\phi \rightarrow 1$, and that the extremal dependence decreases as $m$ increases.
+\begin{figure}
+  \centering
+  \includegraphics[width=\linewidth]{plots/chi-phi}
+  \caption{Simulated lag-$m$ $\chi$ for varying levels of $\varphi$.}
+  \label{stfig:chiphi}
+\end{figure}
 
@@
   \label{sttbl:gaussim}
-  \begin{tabular}{|l|ccc|ccc|cccc|ccc|}
-    \cline{2-14}
-    \multicolumn{1}{c}{} & \multicolumn{3}{|c}{$q(0.90)$} & \multicolumn{3}{|c}{$q(0.95)$} & \multicolumn{4}{|c}{$q(0.98)$} & \multicolumn{3}{|c|}{$q(0.99)$} \\
+  \begin{tabular}{|l|ccc|ccc|cccc|cc|}
+    \cline{2-13}
+    \multicolumn{1}{c}{} & \multicolumn{3}{|c}{$q(0.90)$} & \multicolumn{3}{|c}{$q(0.95)$} & \multicolumn{4}{|c}{$q(0.98)$} & \multicolumn{2}{|c|}{$q(0.99)$} \\
     \hline
-    Method 1 & A &   &   & A &   &   & A &   &   &   & A & B &   \\
+    Method 1 & A &   &   & A &   &   & A &   &   &   & A &   \\
     \hline
-    Method 2 & A &   &   & A &   &   & A &   &   &   & A &   &   \\
+    Method 2 & A &   &   & A &   &   & A &   &   &   & A &   \\
     \hline
-    Method 3 &   & B &   &   & B &   &   &   & C &   &   & B &   \\
+    Method 3 &   & B &   &   & B &   &   &   & C &   & A &   \\
     \hline
-    Method 4 & A &   &   & A &   &   & A & B &   &   & A & B &   \\
+    Method 4 & A &   &   & A &   &   & A & B &   &   & A &   \\
     \hline
-    Method 5 &   & B &   &   & B &   &   & B & C &   & A & B &   \\
+    Method 5 &   & B &   &   & B &   &   & B & C &   & A &   \\
     \hline
-    Method 6 &   &   & C &   &   & C &   &   &   & D &   &   & C \\
+    Method 6 &   &   & C &   &   & C &   &   &   & D &   & B \\
     \hline
@@
   \label{sttbl:st1sim}
-  \begin{tabular}{|l|ccccc|cccc|cccc|ccc|}
-    \cline{2-17}
-    \multicolumn{1}{c}{} & \multicolumn{5}{|c}{$q(0.90)$} & \multicolumn{4}{|c}{$q(0.95)$} & \multicolumn{4}{|c}{$q(0.98)$} & \multicolumn{3}{|c|}{$q(0.99)$} \\
+  \begin{tabular}{|l|cccc|cccc|cccc|ccc|}
+    \cline{2-16}
+    \multicolumn{1}{c}{} & \multicolumn{4}{|c}{$q(0.90)$} & \multicolumn{4}{|c}{$q(0.95)$} & \multicolumn{4}{|c}{$q(0.98)$} & \multicolumn{3}{|c|}{$q(0.99)$} \\
     \hline
-    Method 1 &   &   & C &   &   &   & B &   &   &   & B & C &   &   & B &   \\
+    Method 1 &   & B &   &   &   & B &   &   &   & B &   &   &   & B &   \\
     \hline
-    Method 2 & A &   &   &   &   & A &   &   &   & A &   &   &   & A &   &   \\
+    Method 2 & A &   &   &   & A &   &   &   & A &   &   &   & A &   &   \\
     \hline
-    Method 3 &   & B & C &   &   & A & B &   &   & A & B &   &   & A & B &   \\
+    Method 3 & A & B &   &   & A & B &   &   & A & B &   &   & A & B &   \\
     \hline
-    Method 4 & A & B &   &   &   &   & B &   &   &   & B &   &   & A &   &   \\
+    Method 4 & A & B &   &   & A & B &   &   & A & B &   &   & A & B &   \\
     \hline
-    Method 5 &   &   &   & D &   &   &   & C &   &   &   & C &   &   & B &   \\
+    Method 5 &   &   & C &   &   &   & C &   &   &   & C &   &   &   & C \\
     \hline
-    Method 6 &   &   &   &   & E &   &   &   & D &   &   &   & D &   &   & C \\
+    Method 6 &   &   &   & D &   &   &   & D &   &   &   & D &   &   & C \\
     \hline
@@
   \label{sttbl:st5sim}
-  \begin{tabular}{|l|ccc|cccc|ccc|ccc|}
-    \cline{2-14}
-    \multicolumn{1}{c}{} & \multicolumn{3}{|c}{$q(0.90)$} & \multicolumn{4}{|c}{$q(0.95)$} & \multicolumn{3}{|c}{$q(0.98)$} & \multicolumn{3}{|c|}{$q(0.99)$} \\
+  \begin{tabular}{|l|cccc|cccc|ccc|ccc|}
+    \cline{2-15}
+    \multicolumn{1}{c}{} & \multicolumn{4}{|c}{$q(0.90)$} & \multicolumn{4}{|c}{$q(0.95)$} & \multicolumn{3}{|c}{$q(0.98)$} & \multicolumn{3}{|c|}{$q(0.99)$} \\
     \hline
-    Method 1 &   & B &   &   &   & C &   &   & B &   &   & B &   \\
+    Method 1 &   &   & C &   &   &   & C &   &   & B &   &   & B &   \\
     \hline
-    Method 2 &   & B &   &   &   & C &   &   & B &   &   & B &   \\
+    Method 2 &   &   & C &   &   &   & C &   &   & B &   &   & B &   \\
     \hline
-    Method 3 & A &   &   &   & B &   &   &   & B &   &   & B &   \\
+    Method 3 &   & B &   &   &   & B &   &   & A &   &   & A &   &   \\
     \hline
-    Method 4 & A &   &   & A &   &   &   & A &   &   & A &   &   \\
+    Method 4 & A &   &   &   & A &   &   &   & A &   &   & A &   &   \\
     \hline
-    Method 5 & A &   &   & A &   &   &   & A &   &   & A &   &   \\
+    Method 5 & A &   &   &   & A &   &   &   & A &   &   & A &   &   \\
     \hline
-    Method 6 &   &   & C &   &   &   & D &   &   & C &   &   & C \\
+    Method 6 &   &   &   & D &   &   &   & D &   &   & C &   &   & C \\
     \hline
@@
     \hline
-    Method 2 &   & B &   &   &   & B & C &   &   & B &   &   & B & C \\
+    Method 2 &   & B &   &   &   & B &   &   &   & B &   &   & B & C \\
     \hline
@@
     \hline
-    Method 5 &   &   & C &   &   &   & C &   &   & B &   &   & B & C \\
+    Method 5 &   &   & C &   &   & B & C &   &   & B &   &   & B & C \\
     \hline
@@
   \centering
-  \caption{Setting 5 -- Transformation below $T = q(0.80)$}
+  \caption{Setting 5 -- Brown Resnick}
   \label{sttbl:transsim}
-  \begin{tabular}{|l|cccc|ccc|cccc|cccc|}
-    \cline{2-16}
-    \multicolumn{1}{c}{} & \multicolumn{4}{|c}{$q(0.90)$} & \multicolumn{3}{|c}{$q(0.95)$} & \multicolumn{4}{|c}{$q(0.98)$} & \multicolumn{4}{|c|}{$q(0.99)$} \\
+  \begin{tabular}{|l|cccc|ccc|ccc|ccc|}
+    \cline{2-14}
+    \multicolumn{1}{c}{} & \multicolumn{4}{|c}{$q(0.90)$} & \multicolumn{3}{|c}{$q(0.95)$} & \multicolumn{3}{|c}{$q(0.98)$} & \multicolumn{3}{|c|}{$q(0.99)$} \\
     \hline
-    Method 1 &   &   & C &   &   & B &   &   &   & C &   &   &   & C &   \\
+    Method 1 &   &   &   & D &   &   & C &   &   & C &   &   & C \\
     \hline
-    Method 2 &   & B &   &   &   & B &   &   & B &   &   & A & B &   &   \\
+    Method 2 &   &   &   & D &   &   & C &   &   & C &   &   & C \\
     \hline
-    Method 3 & A &   &   &   & A &   &   & A &   &   &   & A &   &   &   \\
+    Method 3 & A & B &   &   & A &   &   & A & B &   &   & B &   \\
     \hline
-    Method 4 &   & B & C &   &   & B &   &   & B &   &   &   & B & C &   \\
+    Method 4 &   &   & C &   &   & B &   &   & B &   &   & B &   \\
     \hline
-    Method 5 &   & B &   &   &   & B &   &   & B & C &   &   &   & C &   \\
+    Method 5 & A &   &   &   & A &   &   & A &   &   & A & B &   \\
     \hline
-    Method 6 &   &   &   & D &   &   & C &   &   &   & D &   &   &   & D \\
+    Method 6 &   & B & C &   & A &   &   & A &   &   & A &   &   \\
     \hline

### 2016-06-18 code/R/mcmc_cont_lambda.R

@@ -3,7 +3,7 @@
 #
 # TODO: Add in model description here
 #
-#ASSUMES x and y are in [0,1]^2
+# ASSUMES x and y are in [0,1]^2
 ################################################################################
 if (!exists("update_params_cpp_init")) {
   source('update_params_cpp.R')

### 2016-06-18 code/analysis/simstudy/results.R

@@ -208,6 +208,11 @@ comparisons <- c("gaus vs. skew t-1", "gaus vs. t-1 (T = 0.80)", "gaus vs. skew 
                  "t-1 (T = 0.80) vs. Max.stab (T = 0.80)",
                  "skew t-5 vs. t-5 (T = 0.80)", "skew t-5 vs. Max-stab (T = 0.80)",
                  "t-5 (T = 0.80) vs. Max-stab (T = 0.80)")
+comparisons <- c("1 vs. 2", "1 vs. 3", "1 vs. 4", "1 vs. 5", "1 vs. 6",
+                 "2 vs. 3", "2 vs. 4", "2 vs. 5", "2 vs. 6",
+                 "3 vs. 4", "3 vs. 5", "3 vs. 6",
+                 "4 vs. 5", "4 vs. 6",
+                 "5 vs. 6")
 
 quant.score <- savelist$quant.score
 quant.score.mean <- savelist$quant.score.mean
@@ -243,48 +248,102 @@ for (i in 1:50) {
 
 # which groups are different for different settings
 setting <- 1
+brier.score.mean[1, , setting]
+order(brier.score.mean[1, , setting])
 comparisons[which(results.wnmt[, setting, 1] >= 0.05)]  # q(0.90)
+brier.score.mean[6, , setting]
+order(brier.score.mean[6, , setting])
 comparisons[which(results.wnmt[, setting, 2] >= 0.05)]  # q(0.95)
+brier.score.mean[9, , setting]
+order(brier.score.mean[9, , setting])
 comparisons[which(results.wnmt[, setting, 3] >= 0.05)]  # q(0.98)
+brier.score.mean[10, , setting]
+order(brier.score.mean[10, , setting])
 comparisons[which(results.wnmt[, setting, 4] >= 0.05)]  # q(0.99)
+brier.score.mean[11, , setting]
 comparisons[which(results.wnmt[, setting, 5] >= 0.05)]  # q(0.995)
 
-setting <- 2
+setting <- 2  # Not using
+brier.score.mean[1, , setting]
+order(brier.score.mean[1, , setting])
 comparisons[which(results.wnmt[, setting, 1] >= 0.05)]  # q(0.90)
+brier.score.mean[6, , setting]
+order(brier.score.mean[6, , setting])
 comparisons[which(results.wnmt[, setting, 2] >= 0.05)]  # q(0.95)
[…]

### 2016-06-18 LaTeX/response.tex
manuscript updates

@@
   \begin{response}
-    \hl{TODO:}
     We agree that maps of Brier scores would be informative.
     This information could conceptually inform future monitor sites.
-    We have now added a map of Brier scores (see Web Figure), which shows \hl{XXXX}.
+    We have now added a map of Brier scores (see Web \fref{stfig:bssite}), which shows that performance is similar throughout the US with the most varied performance in California.
   \end{response}

### 2016-06-18 LaTeX/supplemental_rev.tex
manuscript updates

@@
 
+\section{Brier scores for ozone prediction} \label{sta:ozonesite}
+Because typical ozone concentration varies throughout the US, we have included Brier scores by site for two model fits (Gaussian and Skew-$t$, $K = 5$, $T = 50$) in \fref{stfig:bssite}.
+As we can see in these plots, both models seem to perform similarly across the US with the poorest performance in California.
+Other methods have similar Brier score maps to these.
+\begin{figure}
+  \centering
+  \includegraphics[width=0.8\linewidth]{plots/bs-site}
+  \caption{Map of Brier scores for Gaussian (top) vs Skew-$t$, $K = 5$, $T = 50$ (bottom).}
+  \label{stfig:bssite}
+\end{figure}
+
 \section{Simulation study pairwise difference results} \label{sta:pdiffs}

### 2016-06-19 LaTeX/response.tex
manuscript updates for revision submission

@@
     As you mention, because the transformation is non-linear and acts independently in each coordinate, the partition created using $\bw$ will be similar to, but not the same as the partition created using $\bw^*$.
-    We have added clarification in Section \sref{sts:temporal} to clarify that the partition is created in the $\bw$ space as opposed to the $\bw^*$ space.
+    We have added clarification in \sref{sts:temporal} to clarify that the partition is created in the $\bw$ space as opposed to the $\bw^*$ space.
     % \hl{Is this something we should be concerned about?}
@@
     In the extremes context, spatial prediction means using the values of monitored stations to determine whether an unmonitored site experienced an extreme event.
-    Mapping the probability of extreme events could either be the marginal probability of an event, such as ozone exceeding 75ppb, or a map of an upper quantile, such as the map of the 99th percentile in Figure \fref{stfig:ozoneq99}.
+    Mapping the probability of extreme events could either be the marginal probability of an event, such as ozone exceeding 75ppb, or a map of an upper quantile, such as the map of the 99th percentile in \fref{stfig:ozoneq99}.
   \end{response}
@@
     This now appears in \aref{sta:temporal}.
-    Thank you both for the suggestion!
+    Thank you both for the suggestion.
   \end{response}

### 2016-06-19 LaTeX/skewt_rev.tex
manuscript updates for revision submission

@@
 
-The partitioning model remove long-range dependence. Conditional on knots $\bw_1, \ldots, \bw_K$, the $\chi$ statistic for two sites $\bs_1$ and $\bs_2$ in partitions $k_1$ and $k_2$ respectively is
+The partitioning model removes long-range dependence. Conditional on knots $\bw_1, \ldots, \bw_K$, the $\chi$ statistic for two sites $\bs_1$ and $\bs_2$ in partitions $k_1$ and $k_2$ respectively is
 \begin{align}
-  \chi(h) &= I(k_1 = k_2) \chi_{\text{skew-}t}(h) + I(k_1 \neq k_2) \chi_{\text{Gaus}}(h) \nonumber \\
-         &= I(k_1 = k_2) \chi_{\text{skew-}t}(h) \label{steq:chiskewt}
+  \chi(h) &= I(k_1 = k_2) \chi_{\text{skew-}t}(h) \label{steq:chiskewt}
 \end{align}
-where $I(\cdot)$ is an indicator function, $\chi_{\text{skew-}t}(h)$ is the $\chi$ statistic for a skew-$t$ process given in equation \eref{steq:chiskew-t} of \aref{sta:skewt}, $\chi_{\text{Gaus}}(h)$ is the $\chi$ statistic for a Gaussian process, and $h = ||\bs_1 - \bs_2||$.
-Therefore, sites in different subregions are asymptotically independent because $\chi_{\text{Gaus}}(h) = 0$ for all $h$.
+where $I(\cdot)$ is an indicator function, $\chi_{\text{skew-}t}(h)$ is the $\chi$ statistic for a skew-$t$ process given in equation \eref{steq:chiskew-t} of \aref{sta:skewt}, and $h = ||\bs_1 - \bs_2||$.
 Marginally, over the knots, $\chi(h) = \pi(h) \chi_{\text{skew}-t}(h)$, where $\pi(h) = \Pr(k_1 = k_2)$ is the probability that two sites separated by distance $h$ are in the same partition.
@@
 Each chain of the MCMC ran for 20,000 iterations with a burn-in period of 10,000 iterations.
-Although the goal of this simulation study is not to assess parameter estimation, for design (2) and method (2), the $\lambda$ estimate and and $a$ appear to converge to their true values (96\% coverage and 90\% coverage respectively).
+Although the goal of this simulation study is not to assess parameter estimation, for design (2) and method (2) where data are generated and fit with a skew-$t$ distribution, the sample converges for $\lambda$ and $a$, and the empirical coverage of posterior 95\% intervals is near the nominal level (96\% coverage for $\lambda$ and 90\% for $a$).
 It should be noted that in the models with multiple partitions (i.e.

### 2016-06-19 LaTeX/supplemental_rev.tex
manuscript updates for revision submission

@@
 \end{align}
-\fref{stfig:chiphi} suggests that the extremal dependence increases as $\phi \rightarrow 1$, and that the extremal dependence decreases as $m$ increases.
+\fref{stfig:chiphi} suggests that the extremal dependence increases as $\varphi \rightarrow 1$, and that the extremal dependence decreases as $m$ increases.
 \begin{figure}

### 2016-06-28 LaTeX/skewt_rev.tex
updates to manuscript

@@
 
-\title[A Space-time Skew-$t$ Model for Threshold Exceedances]{A space-time skew-$t$ model for threshold exceedances}
+\title[A Space-time \Skewt Model for Threshold Exceedances]{A space-time \skewt model for threshold exceedances}
 \author
@@
 % \begin{center}
-% {\Large {\bf A space-time skew-$t$ model for threshold exceedances}}\\
+% {\Large {\bf A space-time \skewt model for threshold exceedances}}\\
 
@@
 In the case of ozone, the level for compliance is fixed at 75 parts per billion, which is high, but not extreme at all locations.
-We present a new method based on the spatial skew-$t$ process.
+We present a new space-time model for threshold exceedances based on the spatial \skewt process.
 Our method incorporates a random partition to permit long-distance asymptotic independence while allowing for sites that are near one another to be asymptotically dependent, and we incorporate thresholding to allow the tails of the data to speak for themselves.
@@
 \begin{keywords}
-Extreme value analysis, MCMC, Random partition, Skew-$t$, Spatio-temporal
+Extreme value analysis, Markov chain Monte Carlo, Random partition, \Skewt, Spatio-temporal modeling
 \end{keywords}
@@
 More specifically, we consider the case of compliance for ozone.
-A site is said to be in compliance if the fourth highest daily maximum 8-hour concentration averaged over 3 years does not exceed 75 parts per billion (ppb).
+A site is said to be in compliance if the fourth highest daily maximum 8-hour concentration averaged over three years does not exceed 75 parts per billion (ppb).
 \fref{stfig:ozone-10jul} shows the ozone levels from July 10, 2005, at 1089 stations across the United States.
@@
 As there are more of these values, they can unduly influence the distribution at the higher levels about which we are interested.
-Our inference method will only use data which exceed a pre-selected threshold, and will impute data below the threshold, thereby tailoring the fit to the levels of interest.
+Our inference method will only use data which exceed a pre-selected threshold and will censor data below the threshold, thereby tailoring the fit to the levels of interest.
 Second, likelihood-based spatial modeling typically assumes a Gaussian process, which is appropriate when mean behavior is of interest.
@@
 Third, we aim to capture the dependence structure when ozone is at high levels, and dependence at these levels may not be well-represented by covariances which focus again on mean behavior.
-Asymptotic dependence/independence (see \sref{sts:extdep}) are notions which describe how two random variables' probability of simultaneous exceedance of an extreme threshold behaves as one increases the threshold.
+Asymptotic dependence/independence (see \sref{sts:extdep}) are notions which describe how two random variables' probability of simultaneous exceedance of an extremely high level.
 The Gaussian distribution always exhibits asymptotic independence, except in the case of perfect dependence, thus is an inappropriate model for data which exhibits asymptotic dependence.
-To allow for more flexibility in the marginal tail and to allow for asymptotic dependence, the skew-$t$ distribution forms the basis for our model.
+To allow for more flexibility in the marginal tail and to allow for asymptotic dependence, the \skewt distribution forms the basis for our model.
 
@@
 \citet{Davison1990} considered modeling threshold exceedances of univariate time series by the generalized Pareto distribution.
-Bivariate threshold models for extreme value distributions were considered by \citet{Ledford1996} who introduced a censored approach that provides a way to deal with different types of exceedances of a bivariate threshold in terms of only one or both components.
+Threshold based inference for multivariate extreme value distributions was considered by \citet{Ledford1996} who introduced a censored approach that provides a way to deal with different types of exceedances of a threshold.
 These threshold models were extended to spatial models for extremes by \citet{Wadsworth2012} and \citet{Thibaud2013} who fit various models to spatial extremes using a censored pairwise likelihood \citep{Padoan2010} based on the approach of \citet{Ledford1996}.
 \citet{Huser2014} further extended this to space-time modeling.
-\citet{Thibaud2013a}, \citet{Engelke2014}, and \citet{Wadsworth2014}, introduced more efficient inference for threshold exceedances of extremal spatial processes with full likelihood methods.
+\citet{Wadsworth2014}, \citet{Engelke2015}, and \citet{Thibaud2013a} introduced more efficient inference for threshold exceedances of extremal spatial processes with full likelihood methods.
 The previous approaches to threshold modeling are motivated by extreme value theory and assume the threshold is high enough that extremal models are valid for the data, and for extrapolation beyond the range of observed values.
 Moreover, these approaches are computationally intensive and limited to rather small datasets.
-For example, \citet{Wadsworth2014} present a simulation study with observations at 16 sites on a regular grid, and \citet{Engelke2014} analyze a dataset with observations at 35 meteorological stations.
-Our application with ozone data does not fit into this framework because we do not focus on exceedances of a very high threshold, but on exceedances of a fixed threshold.
-Furthermore, in our application, we have observations at over 1,000 ozone monitoring locations.
+For example, \citet{Wadsworth2014} present a simulation study with observations at 16 sites on a regular grid, and \citet{Engelke2015} analyze a dataset with observations at 35 meteorological stations.
+Our application with ozone data does not fit into this framework because we do not focus on exceedances of a very high level, and we have observations at over 1,000 ozone monitoring locations.
 
-We propose a new spatiotemporal threshold exceedance model based on the skew-$t$ process \citep{Padoan2011}.
+We propose a new spatiotemporal threshold exceedance model based on the \skewt process \citep{Padoan2011}.
 \hl{Need to add in more references here to skew-$t$ used in a spatial setting.}
-Our model is a threshold exceedance model for the multivariate skew-$t$ distribution using a fixed threshold.
-In this setting, we describe the threshold as fixed because it is specified in advance by regulatory compliance.
-This differs from the more traditional extremes literature where a threshold is selected to be the value beyond which an extremal model is appropriate for the data.
-We use a skew-$t$ distribution because of its flexibility to model asymmetry and heavy-tailed data with the aim of predicting the probability of exceeding a high fixed level at an unobserved location.
-
+We use a \skewt distribution because of its flexibility to model asymmetry and heavy-tailed data with the aim of predicting the probability of exceeding a high fixed level at an unobserved location.
 Our model allows for inference and predictions using the full likelihood with computing on the order of Gaussian models.
-This allows us to use Bayesian methods to impute data below the threshold as well as to make predictions at unobserved locations.
-In a spatial setting, the multivariate skew-$t$ distribution demonstrates asymptotic dependence between observations at all sites regardless of the distance between the sites.
+This allows us to use Bayesian methods, which we use to fit the model, handle censored data below the threshold, and make predictions at unobserved locations.
+In a spatial setting, the multivariate \skewt distribution demonstrates asymptotic dependence between observations at all sites regardless of the distance between the sites.
 In order to address this concern, we introduce a random spatial partition similar to the method used by \citet{Kim2005} for non-stationary Gaussian data.
-This partitioning alleviates the asymptotic spatial dependence present in the skew-$t$ distribution for sites that are far apart.
+This partitioning alleviates the asymptotic spatial dependence present in the \skewt distribution for sites that are far apart.
 
 The paper is organized as follows.
-\sref{sts:spatialskew} is a brief review of the spatial skew-$t$ process.
-In \sref{sts:spatial}, we build upon the traditional skew-$t$ process by incorporating censoring to focus on tails, partitioning to remove long-range asymptotic dependence, and extending the model to space-time data.
+\sref{sts:spatialskew} is a brief review of the spatial \skewt process.
+In \sref{sts:spatial}, we build upon the traditional \skewt process by incorporating censoring to focus on tails, partitioning to remove long-range asymptotic dependence, and extending the model to space-time data.
 The computing is described in \sref{sts:hier}.
 In \sref{sts:simstudy}, we present a simulation study that examines the predictive capabilities of this model compared to Gaussian and max-stable methods.
-We then compare our method to Gaussian and max-stable methods with a data analysis of ozone measurements from 800 sites throughout the US in \sref{sts:analysis}.
+We then compare our method to Gaussian and max-stable methods with a data analysis of ozone measurements from 1,000 sites throughout the US in \sref{sts:analysis}.
 The final section provides brief discussion and direction for future research.
@@
 % In addition to Brown-Resnick processes, skew elliptical distributions can be used for multivariate modeling with dependent extreme values \citep{Genton2004,Zhang2010,Padoan2011}.
-% For example, the skew-normal and skew-$t$ distribution offer a flexible way to handle non-symmetric data within a framework of multivariate normal and multivariate $t$ distributions.
+% For example, the skew-normal and \skewt distribution offer a flexible way to handle non-symmetric data within a framework of multivariate normal and multivariate $t$ distributions.
 % As with multivariate Gaussian distributions, the multivariate skew-normal distribution demonstrates asymptotic independence.
-Conversely both the multivariate $t$ and skew-$t$ distributions demonstrate asymptotic dependence \citep{Padoan2011}.
-% Additionally, the limiting distribution of the maxima of skew-$t$ random vectors is the extremal skew-$t$ distribution \citep{Padoan2011} of which the extremal-$t$ \citep{Opitz2013} is a special case.
+Conversely both the multivariate $t$ and \skewt distributions demonstrate asymptotic dependence \citep{Padoan2011}.
+% Additionally, the limiting distribution of the maxima of \skewt random vectors is the extremal \skewt distribution \citep{Padoan2011} of which the extremal-$t$ \citep{Opitz2013} is a special case.
 
@@
 The skew-elliptical family of distributions provides models that are mathematically tractable while introducing a slant parameter to account for asymmetric data.
-A brief review of the additive process (\citealp{Azzalini2003}; \citealp[p. 129]{Azzalini2014}; \citealp{Beranger2016}) by which a skew-$t$ process is created is given here.
+A brief review of the additive process (\citealp{Azzalini2003}; \citealp[p. 129]{Azzalini2014}; \citealp{Beranger2016}) by which a \skewt process is created is given here.
 
-\subsection{Skew-$t$ process} \label{sts:skewt}
+\subsection{\Skewt process} \label{sts:skewt}
 Let $Y(\bs)$ be a spatial process defined for spatial location $\bs$ in a spatial domain of interest $\calD \in \calR^2$.
-The spatial skew-$t$ process can be written
+The spatial \skewt process can be written as
 \begin{align}
@@
 \end{align}
-$\bX(\bs)$ is the observed covariate vector at site $\bs$, $\bbeta$ is the $p$-vector of regression parameters, $\lambda \in \calR$ is a parameter controlling skew, $z \sim N(0, 1)$, $\sigma^2 \sim \text{IG}(a / 2, b / 2)$ is random scale parameter, IG is the distribution function of an inverse gamma random variable, $a$ is the degrees of freedom, $b$ controls the precision of the process, and $v(\bs)$ is a spatial Gaussian process with mean zero, variance one, and a positive definite correlation function $\text{Cor} \left[Y(\bs_1), Y(\bs_2)\right] = \rho(\bs_1, \bs_2)$.
+where $\bX(\bs)$ is the observed covariate vector at site $\bs$, $\bbeta$ is the $p$-vector of regression parameters, $\lambda \in \calR$ is a parameter controlling skewness, $z \sim N(0, 1)$, $\sigma^2 \sim \text{IG}(a / 2, b / 2)$ is random scale parameter, IG is the distribution function of an inverse gamma random variable, $a$ is the degrees of freedom, $b$ controls the precision of the process, and $v(\bs)$ is a spatial Gaussian process with mean zero, variance one, and a positive definite correlation function $\text{Cor} \left[Y(\bs_1), Y(\bs_2)\right] = \rho(\bs_1, \bs_2)$.
 Although any positive definite correlation function could be used, we choose to use the stationary isotropic \Matern correlation with
@@
 \end{align}
-where $\varphi$ is the spatial range, $\nu$ is the smoothness, $\gamma$ is the proportion of variance accounted for by the spatial variation, $K_\nu$ is a modified Bessel function of the second kind, and $h = || \bs_1 - \bs_2 ||$ is the Euclidean distance between sites $\bs_1$ and $\bs_2$.
+where $I(\cdot)$ is an indicator function, $\varphi$ is the spatial range, $\nu$ is the smoothness, $\gamma$ is the proportion of variance accounted for by the spatial variation, $K_\nu$ is a modified Bessel function of the second kind, and $h = || \bs_1 - \bs_2 ||$ is the Euclidean distance between sites $\bs_1$ and $\bs_2$.
 
-For a finite collection of locations $\bs_1, \ldots, \bs_n$, denote the vector of observations $\bY = [Y(\bs_1), \ldots, Y(\bs_n)]^\top$.
+For a finite collection of locations $\bs_1, \ldots, \bs_n$, denote the vector of observations $\bY = [Y(\bs_1), \ldots, Y(\bs_n)]^\top$, and the covariate matrix $\bX_{n \times p} = [\bX(\bs_1), \ldots, \bX(\bs_n)]^\top$.
 After marginalizing over both $z$ and $\sigma$, using the notation from \citet[p. 176]{Azzalini2014},
@@
 \end{align}
-that is, $\bY$ follows an $n$-dimensional skew-$t$ distribution with location $\bX \bbeta \in \calR^n$, variance matrix $\bOmega_{n \times n}$, $\balpha \in \calR^n$ slant parameters, degrees of freedom $a$, where $\bX_{n \times p} = [\bX(\bs_1)^\top, \ldots, \bX(\bs_n)^\top]$,
-$\bX(\bs_i), i = 1, \ldots, n$, are observed covariates at each site,
-$\bOmega_{n \times n} = \bomega \bar{\bOmega} \bomega$,
-$\bar{\bOmega}_{n \times n} = \frac{1}{1 + \lambda^2}(\bSigma + \lambda^2 \bOne \bOne^\top)$,
-$\bomega_{n \times n} = \text{diag}\left(\sqrt{\frac{a(1 + \lambda^2)}{ b }}, \ldots, \sqrt{\frac{a(1 + \lambda^2)}{ b }} \right)$,
-$\balpha = \lambda (1 + \lambda^2)^{1 / 2} (1 + \lambda^2 \bOne^\top \bSigma^{-1} \bOne)^{-1/2} \bSigma^{-1} \bOne$, and $\bSigma_{n \times n}$ is the positive definite correlation matrix which is obtained from $\rho(h)$.
-This process is desirable because of its flexible tail that is controlled by the skewness parameter $\lambda$ and degrees of freedom $a$.
-Furthermore, the marginal distributions at each location also follow a univariate skew-$t$ distribution \citep{Azzalini2014}.
-For comparisons with other parameterizations, please see \aref{sta:otherparams}.
+that is, $\bY$ follows an $n$-dimensional \skewt distribution with location $\bX \bbeta \in \calR^n$; covariance matrix $\bOmega_{n \times n} = \bomega \bar{\bOmega} \bomega$ where \mbox{$\bar{\bOmega}_{n \times n} = \frac{1}{1 + \lambda^2}(\bSigma + \lambda^2 \bOne \bOne^\top)$}, $\bSigma_{n \times n}$ is the positive definite correlation matrix which is obtained from $\rho(h)$, and $\bomega_{n \times n} = \text{diag}\left(\sqrt{\frac{ b }{a(1 + \lambda^2)}}, \ldots, \sqrt{\frac{ b }{a(1 + \lambda^2)}} \right)$; slant parameters $\balpha \in \calR^n = \lambda (1 + \lambda^2)^{1 / 2} (1 + \lambda^2 \bOne^\top \bSigma^{-1} \bOne)^{-1/2} \bSigma^{-1} \bOne$, and degrees of freedom $a$.
+Furthermore, the marginal distributions at each location also follow a univariate \skewt distribution \citep{Azzalini2014}.
+This process is desirable because of its flexible tail that is controlled by the skewness parameter and degrees of freedom.
+One major difference between the parameterization given in \eref{steq:fullmodel} and other parameterizations is that we characterize the skewnewss using $\lambda$ instead of the slant parameters $\balpha$.
+For a more complete comparison with other parameterizations, see \aref{sta:otherparams}.
 
@@
 \subsection{Extremal dependence}\label{sts:extdep}
-Our interest lies in spatial dependence in the tail of the skew-$t$ process.
+Our interest lies in spatial dependence in the tail of the \skewt process.
 One measure of extremal dependence is the $\chi$ statistic \citep{Coles1999}.
@@
 where $c^*$ is the upper limit of the support of $Y$.
-For the skew-$t$ distribution, $c^* = \infty$.
+For the \skewt distribution, $c^* = \infty$.
 If $\chi(h) = 0$, then observations are asymptotically independent at distance $h$.
 For Gaussian processes, $\chi(h) = 0$ regardless of the distance $h$, so they are not suitable for modeling asymptotically dependent extremes.
-Unlike the Gaussian process, the skew-$t$ process is asymptotically dependent (the explicit expression for $\chi(h)$ is given in \aref{sta:skewt}).
-However, one problem with the spatial skew-$t$ process is that $\lim_{h \rightarrow \infty} \chi(h) > 0$.
+Unlike the Gaussian process, the \skewt process is asymptotically dependent (the explicit expression for $\chi(h)$ is given in \aref{sta:skewt}).
+However, one problem with the spatial \skewt process is that $\lim_{h \rightarrow \infty} \chi(h) > 0$.
 This occurs because all observations, both near and far, share the same $z$ and $\sigma$ terms.
-Therefore, this long-range dependence feature of the skew-$t$ process is not ideal for spatial analysis of large geographic regions where we expect only local spatial dependence.
+Therefore, this long-range dependence feature of the \skewt process is not desirable for spatial analysis of large geographic regions where we expect only local spatial dependence.
 We propose a solution to this in \sref{sts:part}.
 
-\section{Spatiotemporal skew-$t$ model for threshold exceedances}\label{sts:spatial}
-In this section, we propose extensions to the skew-$t$ process to model spatial extremes over a large geographic region by introducing censoring to focus on tail behavior and a random partition to remove long-range asymptotic dependence.
+\section{Spatiotemporal \skewt model for threshold exceedances}\label{sts:spatial}
+In this section, we propose extensions to the \skewt process to model spatial extremes over a large geographic region by introducing censoring to focus on tail behavior and a random partition to remove long-range asymptotic dependence.
 For notational convenience, we introduce the model for a single replication, and then extend this model to the spatiotemporal setting in \sref{sts:temporal}.
 
-\subsection{Censoring to focus on the tails} \label{sts:censoring}
-As mentioned previously, we propose to use a censored approach because we are interested in high values and do not want the low-to-moderate values to influence the fit of the overall model.
+\subsection{Censoring to focus on the tail} \label{sts:censoring}
+As mentioned previously, we propose to use a censored approach because we do not want the low-to-moderate values to influence the fit of the overall model.
 % The censored observations below the threshold give information on the marginal probabilities to exceed the threshold and on the dependence, but their values are not used to fit the model.
@@
 \eeq
-be the censored observation at site $\bs$ where $Y(\bs)$ is the uncensored observation, $\delta(\bs) = I[Y(\bs) > T]$, and $T$ is a pre-specified threshold value.
-As our goal is to predict exceedances above a high level $L$, we should select a value for $T \le L$.
-For example, in predicting ozone exceedances, we might set $T = 50$ppb in order to predict exceedances of $L = 75$ppb.
-Selecting $T$ too small may lead to bias in estimating the tail parameters; selecting $T$ to large increases variance.
-We impute the censored values as a step in the MCMC algorithm used to fit the model described in \sref{sts:comp}, and use cross-validation to select $T$.
+be the censored process at site $\bs$ where $Y(\bs)$ is the uncensored process, $\delta(\bs) = I[Y(\bs) > T]$, and $T$ is a pre-specified threshold value.
+As our goal is to model exceedances above a high level $L$, we should select a value for $T \le L$.
+For example, in predicting ozone exceedances, we might set $T = 50$ ppb in order to predict exceedances of $L = 75$ ppb.
+Selecting $T$ too small may lead to bias in estimating the tail parameters; selecting $T$ too large increases variance.
+We impute the censored values as a step in the algorithm used to fit the model described in \sref{sts:comp}, and use cross-validation to select $T$.
 
@@
 The motivation for the partition is that for a large spatial domain, it may not be reasonable to assume sites that are far apart demonstrate asymptotic dependence.
-Modeling different levels of asymptotic dependence was discussed by \citet{Wadsworth2012} with a hybrid spatial dependence model.
+Modeling different levels of asymptotic dependence was discussed by \citet{Wadsworth2012}.
 \citet{Huser2014} also allow for varying asymptotic dependence across both space and time with a partition structure represented by random discs moving across the space for a random duration with a random velocity and random radius.
@@
 \end{align}
-To model spatial variation, consider a set of spatial knots $\bw_1, \ldots, \bw_K$ from a homogeneous Poisson process with intensity $\mu_{PP}$ over spatial domain $\calD \in \calR^2$.
+To model spatial variation, consider a set of spatial knots $\bw_1, \ldots, \bw_K$ from a homogeneous Poisson process over spatial domain $\calD \in \calR^2$.
 The knots define a random partition of $\calD$ by subregions $P_{1}, \ldots, P_{K}$ defined as
@@
 For all $\bs \in P_k$, with $k = 1, 2, \ldots, K$, the functions $z(\bs)$ and $\sigma(\bs)$ are equal to the constants $z_k$ and $\sigma_k$ respectively, and the $z_k$ and $\sigma^2_k$ are distributed as $z_k \iid N(0, 1)$ and $\sigma^2_k \iid$ IG$(a / 2, b / 2)$.
-So, within each partition, $Y(\bs)$ follows the spatial skew-$t$ process defined in \sref{sts:spatialskew}.
-Across partitions, the $Y(\bs)$ remain correlated via the correlation function for $v(\bs)$ because $v(\bs)$ spans all partitions.
-However, the bivariate distribution for sites in different partitions is neither Gaussian nor skew-$t$ and does not have asymptotic dependence.
+So, within each partition, $Y(\bs)$ follows the spatial \skewt process defined in \sref{sts:spatialskew}.
+Across partitions, the $Y(\bs)$ remain dependence via the correlation function for $v(\bs)$ because $v(\bs)$ spans all partitions.
+However, the bivariate distribution for sites in different partitions is neither Gaussian nor \skewt and does not have asymptotic dependence.
 
@@
 \end{align}
-where $I(\cdot)$ is an indicator function, $\chi_{\text{skew-}t}(h)$ is the $\chi$ statistic for a skew-$t$ process given in equation \eref{steq:chiskew-t} of \aref{sta:skewt}, and $h = ||\bs_1 - \bs_2||$.
+where $\chi_{\text{skew-}t}(h)$ is the $\chi$ statistic for a \skewt process given in equation \eref{steq:chiskew-t} of \aref{sta:skewt}, and $h = ||\bs_1 - \bs_2||$.
 Marginally, over the knots, $\chi(h) = \pi(h) \chi_{\text{skew}-t}(h)$, where $\pi(h) = \Pr(k_1 = k_2)$ is the probability that two sites separated by distance $h$ are in the same partition.
@@
 In practice we fix $K$ at a finite value and use a uniform distribution for the knots $\bw_1, \ldots, \bw_K$.
-In \fref{stfig:chi}, we give $\chi(h)$ for $K = 1, 3, 5, 10$ partitions for a skew-$t$ distribution with $\alpha = 10$, and 3 degrees of freedom.
+In \fref{stfig:chi}, we give $\chi(h)$ for $K = 1, 3, 5, 10$ partitions for a \skewt distribution with $\alpha = 10$, and 3 degrees of freedom.
 In \fref{stfig:chi} we estimate $\pi(h)$ through simulation.
@@
 In this section, we extend \eref{steq:partition} to the spatiotemporal setting.
-There are several places where temporal dependence could be incorporated in the model, including the residuals $v_t(\bs)$.
-However, we choose to allow for temporal dependence in the $\bw$, $z$, and $\sigma$ terms because these terms dictate the tail behavior which is our primary focus.
+There are several places where temporal dependence could be incorporated in the model, including the Gaussian process $v_t(\bs)$.
+However, we choose to allow for temporal dependence in the $\bw$, $z$, and $\sigma$ terms because these terms dictate the tail behavior which is our primary focus (see \aref{sta:temporal}).
 Let
@@
 We use an AR(1) time series model for $\bw_{tk}$, $z_{tk}$, and $\sigma_{tk}$.
-The time series model must be specified after a transformation to preserve the skew-$t$ process at each time point.
-For each time-varying parameter, we transform to obtain a standard normal marginal distribution, place a Gaussian prior with autocorrelation on the transformed parameter, and then transform back to the appropriate marginal distribution for the skew-$t$ process.
+The time series model must be specified after a transformation to preserve the \skewt process at each time point.
+For each time-varying parameter, we transform the parameter to obtain a standard normal marginal distribution, place a Gaussian prior with autocorrelation on the transformed parameter, and then transform back to the appropriate marginal distribution for the \skewt process.
 We first transform the spatial knots from $\calD$ to $\calR^2$ as follows.
@@
 where $|\phi_w|$, $|\phi_z|$, $|\phi_\sigma| < 1$.
-These are stationary time series models with marginal distributions \hbox{$\bw^*_{k} \sim N_2(0, 1)$}, \hbox{$z^*_{k} \sim N(0, 1)$}, and \hbox{$\sigma^{2*}_{k} \sim N(0, 1)$}.
+These are stationary time series models with marginal distributions \hbox{$\bw^*_{k} \sim N_2(\bZero, \bI_2)$}, where $\bI_2 = \text{diag}(1, 1)$, \hbox{$z^*_{k} \sim N(0, 1)$}, and \hbox{$\sigma^{2*}_{k} \sim N(0, 1)$}.
 After transformation back to the original space, $\bw_{tk} \sim \text{Unif}(\calD)$, $z_{tk} \sim HN(0, 1)$, and $\sigma^2_{tk} \sim \text{IG}(a / 2, b / 2)$.
 We then create the partition for day $t$ using $\bw_{t1}, \ldots, \bw_{tk}$.
-For each day, the model is identical to the spatial-only model in \eref{steq:partition} by construction.
+For each day, the model is identical to the spatial-only model in \eref{steq:partition} by construction.\citep{
 
@@
   % \item Symmetric-$t$ marginal, $K=5$ knots
-  \item Skew-$t$ marginal, $K=1$ knots
-  \item Skew-$t$ marginal, $K=5$ knots
+  \item \Skewt marginal, $K=1$ knots
+  \item \Skewt marginal, $K=5$ knots
   \item Max-stable with GEV(1, 1, 0.2) marginals \citep{Reich2012}
@@
   \item Gaussian marginal, $K=1$ knots
-  \item Skew-$t$ marginal, $K=1$ knots, $T=-\infty$
+  \item \Skewt marginal, $K=1$ knots, $T=-\infty$
   \item Symmetric-$t$ marginal, $K=1$ knots, $T=q(0.80)$
-  \item Skew-$t$ marginal, $K=5$ knots, $T=-\infty$
+  \item \Skewt marginal, $K=5$ knots, $T=-\infty$
   \item Symmetric-$t$ marginal, $K=5$ knots, $T=q(0.80)$
@@
 The knots have priors $\bw \sim \text{Unif}(\calD)$.
-We tried also fitting the skew-$t$ marginals for the thresholded models, but it is very challenging for the MCMC to properly identify the skewness parameter with a censored left tail.
+We tried also fitting the \skewt marginals for the thresholded models, but it is very challenging for the MCMC to properly identify the skewness parameter with a censored left tail.
 Each chain of the MCMC ran for 20,000 iterations with a burn-in period of 10,000 iterations.
-Although the goal of this simulation study is not to assess parameter estimation, for design (2) and method (2) where data are generated and fit with a skew-$t$ distribution, the sample converges for $\lambda$ and $a$, and the empirical coverage of posterior 95\% intervals is near the nominal level (96\% coverage for $\lambda$ and 90\% for $a$).
+Although the goal of this simulation study is not to assess parameter estimation, for design (2) and method (2) where data are generated and fit with a \skewt distribution, the sample converges for $\lambda$ and $a$, and the empirical coverage of posterior 95\% intervals is near the nominal level (96\% coverage for $\lambda$ and 90\% for $a$).
 It should be noted that in the models with multiple partitions (i.e.
@@
 % \end{align}
-\fref{stfig:ozone-qq} shows a Q-Q plot of the residuals compared to a skew-$t$ distribution with 10 d.f. and $\alpha = 1$.
+\fref{stfig:ozone-qq} shows a Q-Q plot of the residuals compared to a \skewt distribution with 10 d.f. and $\alpha = 1$.
 \begin{figure}
@@
   \includegraphics[width=0.5\linewidth]{plots/qq-res.pdf}
-  \caption{Q-Q plot of the residuals for a skew-$t$ distribution with 10 d.f. and $\alpha = 1$ (right)}
+  \caption{Q-Q plot of the residuals for a \skewt distribution with 10 d.f. and $\alpha = 1$ (right)}
   \label{stfig:ozone-qq}
@@
 \subsection{Model comparisons}
-We fit the model using Gaussian and skew-$t$ marginal distributions with $K=1, 5, 6, 7, 8, 9, 10, 15$ partitions.
+We fit the model using Gaussian and \skewt marginal distributions with $K=1, 5, 6, 7, 8, 9, 10, 15$ partitions.
 We choose to censor $Y(\bs)$ at $T = 0$, $T = 50$ (0.42 sample quantile), and $T = 75$ (0.92 sample quantile) ppb in order to compare results from no, moderate, and high censoring.
@@
 \subsection{Results}\label{sts:results}
-The results suggest that the skew-$t$, thresholded, partitioned, and time series models all give an improvement in predictions over the Gaussian model, whereas the max-stable method results in relative Brier scores between 1.13 and 1.18 indicating poorer performance than the Gaussian model.
+The results suggest that the \skewt, thresholded, partitioned, and time series models all give an improvement in predictions over the Gaussian model, whereas the max-stable method results in relative Brier scores between 1.13 and 1.18 indicating poorer performance than the Gaussian model.
 The plots in \fref{stfig:bs-ozone} show the relative Brier scores for time-series and non-time-series models, using $K = $ 1, 7, and 15 knots at thresholds $T = $ 0, 50, and 75 ppb.
@@
   \item Gaussian
-  \item Skew-$t$, $K =$ 1 knot, $T = $ 0, no time series
-  \item Skew-$t$, $K =$ 5 knots, $T = $ 50, no time series
+  \item \Skewt, $K =$ 1 knot, $T = $ 0, no time series
+  \item \Skewt, $K =$ 5 knots, $T = $ 50, no time series
   \item Symmetric-$t$, $K =$ 10 knots, $T = $ 75, time series.
@@
 \section{Discussion}\label{sts:con}
-In this paper we propose a new threshold exceedance approach for spatiotemporal modeling based on the skew-$t$ process.
+In this paper we propose a new threshold exceedance approach for spatiotemporal modeling based on the \skewt process.
 The proposed model gives flexible tail behavior, demonstrates asymptotic dependence for observations at sites that are near to one another, and has computation on the order of Gaussian models for large space-time datasets.

### 2016-06-28 LaTeX/supplemental_rev.tex
updates to manuscript

@@
 
-\title[Web-based Supplementary Materials for A Space-time Skew-$t$ Model for Threshold Exceedances]{Web-based Supplementary Materials for A Space-time Skew-$t$ Model for Threshold Exceedances by Morris, Reich, Thibaud, and Cooley}
+\title[Web-based Supplementary Materials for A Space-time \Skewt Model for Threshold Exceedances]{Web-based Supplementary Materials for A Space-time \Skewt Model for Threshold Exceedances by Morris, Reich, Thibaud, and Cooley}
 \author
@@
 % \begin{center}
-% {\Large {\bf Supplemental material for A space-time skew-$t$ model for threshold exceedances}}\\
+% {\Large {\bf Supplemental material for A space-time \skewt model for threshold exceedances}}\\
 
@@
 
-\section{Skew-$t$ distribution} \label{sta:skewt}
-\subsection*{Univariate skew-$t$ distribution}
-We say that $Y$ follows a univariate extended skew-$t$ distribution with location $\xi \in \calR$, scale $\omega > 0$, skew parameter $\alpha \in \calR$, and degrees of freedom $\nu$ if has distribution function
+\section{\Skewt distribution} \label{sta:skewt}
+\subsection*{Univariate \skewt distribution}
+We say that $Y$ follows a univariate extended \skewt distribution with location $\xi \in \calR$, scale $\omega > 0$, skew parameter $\alpha \in \calR$, and degrees of freedom $\nu$ if has distribution function
 \begin{align}
@@
 
-\subsection*{Multivariate skew-$t$ distribution}
-If $\bZ \sim \text{ST}_d(0, \bar{\bOmega}, \balpha, \eta)$ is a $d$-dimensional skew-$t$ distribution, and $\bY = \xi + \bomega \bZ$, where $\bomega = \text{diag}(\omega_1, \ldots, \omega_d)$, then the density of $Y$ at $y$ is
+\subsection*{Multivariate \skewt distribution}
+If $\bZ \sim \text{ST}_d(0, \bar{\bOmega}, \balpha, \eta)$ is a $d$-dimensional \skewt distribution, and $\bY = \xi + \bomega \bZ$, where $\bomega = \text{diag}(\omega_1, \ldots, \omega_d)$, then the density of $Y$ at $y$ is
 \begin{align}
@@
 \subsection*{Extremal dependence}
-For a bivariate skew-$t$ random variable $\bY = [Y(\bs), Y(\bt)]^\top$, the $\chi(h)$ statistic \citep{Padoan2011} is given by
+For a bivariate \skewt random variable $\bY = [Y(\bs), Y(\bt)]^\top$, the $\chi(h)$ statistic \citep{Padoan2011} is given by
 \begin{align} \label{steq:chiskew-t}
@@
 \end{align}
-where $\bar{F}_{\text{EST}}$ is the univariate survival extended skew-$t$ function with zero location and unit scale, \hbox{$\varrho(h) = \text{cor}[y(\bs), y(\bt)]$}, $\alpha_j = \alpha_i \sqrt{1 - \varrho^2}$, $\tau_j = \sqrt{\eta + 1}(\alpha_j + \alpha_i \varrho)$, and $x_j = F_T(\bar{\alpha}_i \sqrt{\eta + 1}; 0, 1, \eta) / F_T(\bar{\alpha}_j \sqrt{\eta + 1}; 0, 1, \eta)$ with $j = 1, 2$ and $i = 2, 1$ and where $\bar{\alpha}_j = (\alpha_j + \alpha_i \varrho) / \sqrt{ 1 + \alpha_i^2 [1 - \varrho(h)^2]}$.
+where $\bar{F}_{\text{EST}}$ is the univariate survival extended \skewt function with zero location and unit scale, \hbox{$\varrho(h) = \text{cor}[y(\bs), y(\bt)]$}, $\alpha_j = \alpha_i \sqrt{1 - \varrho^2}$, $\tau_j = \sqrt{\eta + 1}(\alpha_j + \alpha_i \varrho)$, and $x_j = F_T(\bar{\alpha}_i \sqrt{\eta + 1}; 0, 1, \eta) / F_T(\bar{\alpha}_j \sqrt{\eta + 1}; 0, 1, \eta)$ with $j = 1, 2$ and $i = 2, 1$ and where $\bar{\alpha}_j = (\alpha_j + \alpha_i \varrho) / \sqrt{ 1 + \alpha_i^2 [1 - \varrho(h)^2]}$.
 
@@
 \end{align}
-Because the extended skew-$t$ distribution is not bounded above, for all $\bar{F}_{\text{EST}}(x) = 1 - F_{\text{EST} (x)} > 0$ for all $x < \infty$.
-Therefore, for a skew-$t$ distribution, $\lim_{h \rightarrow \infty} \chi(h) > 0$.
+Because the extended \skewt distribution is not bounded above, for all $\bar{F}_{\text{EST}}(x) = 1 - F_{\text{EST} (x)} > 0$ for all $x < \infty$.
+Therefore, for a \skewt distribution, $\lim_{h \rightarrow \infty} \chi(h) > 0$.
 
 \section{Comparisons with other parameterizations} \label{sta:otherparams}
-Various forms of multivariate skew-normal and skew-$t$ distributions have been proposed in the literature.
+Various forms of multivariate skew-normal and \skewt distributions have been proposed in the literature.
 In this section, we make a connection between our parameterization in \eref{steq:fullmodel} of the main text and another popular version.
@@
 where $\tilde{\lambda} \in (-1, 1)$, $z \sim N(0, 1)$, and $v(\bs)$ is a Gaussian process with mean zero, variance one, and spatial correlation function $\rho$.
-To extend this to the skew-$t$ distribution, \citet{Azzalini2003} take $\tilde{Y}(\bs) = W\tilde{X}(\bs)$ where $W^{-2} \sim $ Gamma$(a / 2, a / 2)$.
+To extend this to the \skewt distribution, \citet{Azzalini2003} take $\tilde{Y}(\bs) = W\tilde{X}(\bs)$ where $W^{-2} \sim $ Gamma$(a / 2, a / 2)$.
 Returning to the proposed parameterization (with $\bbeta = 0$), let $W^{-2} = \frac{b}{a}\sigma^{-2} \sim$ Gamma$(a / 2, a / 2)$ so that \eref{steq:fullmodel} in the manuscript becomes
@@
 \section{Brier scores for ozone prediction} \label{sta:ozonesite}
-Because typical ozone concentration varies throughout the US, we have included Brier scores by site for two model fits (Gaussian and Skew-$t$, $K = 5$, $T = 50$) in \fref{stfig:bssite}.
+Because typical ozone concentration varies throughout the US, we have included Brier scores by site for two model fits (Gaussian and \Skewt, $K = 5$, $T = 50$) in \fref{stfig:bssite}.
 As we can see in these plots, both models seem to perform similarly across the US with the poorest performance in California.
@@
   \includegraphics[width=0.8\linewidth]{plots/bs-site}
-  \caption{Map of Brier scores for Gaussian (top) vs Skew-$t$, $K = 5$, $T = 50$ (bottom).}
+  \caption{Map of Brier scores for Gaussian (top) vs \Skewt, $K = 5$, $T = 50$ (bottom).}
   \label{stfig:bssite}
@@
   \centering
-  \caption{Setting 2 -- Skew-$t$ marginal, $K = 1$ knot}
+  \caption{Setting 2 -- \Skewt marginal, $K = 1$ knot}
   \label{sttbl:st1sim}
@@
   \centering
-  \caption{Setting 3 -- Skew-$t$ marginal, $K = 5$ knots}
+  \caption{Setting 3 -- \Skewt marginal, $K = 5$ knots}
   \label{sttbl:st5sim}

### 2016-06-29 LaTeX/response.tex
updates to manuscript for revision

@@
   \begin{response}
-    \hl{TODO:}
-    We have added several references given by Reviewer 2 to the introduction.
+    We have added several references to the introduction.
   \end{response}
@@
       However, if $\sigma_1$ and $\sigma_2$ are different and unknown, then $\bY$ is neither Gaussian nor $t$.
-      From our perspective, however, the important feature of this distribution is that it has asymptotic independence, which is sufficient to break the long-range dependence of the usual spatial-$t$ process.
+      From our perspective (as we now mention above \eref{steq:chiskewt}), however, the important feature of this distribution is that it has asymptotic independence, which is sufficient to break the long-range dependence of the usual spatial-$t$ process.
     \end{response}
@@
   \begin{response}
-    We now say throughout the proof in \aref{sta:proofsamepartition} that the point process is a homogeneous Poisson process with intensity $\mu(\cdot)$.
+    We now say throughout the proof in \aref{sta:proofsamepartition} that the point process is a homogeneous Poisson process with intensity $\lambda_{PP}$.
   \end{response}

### 2016-06-29 LaTeX/skewt_rev.tex
updates to manuscript for revision

@@
 We propose a new spatiotemporal threshold exceedance model based on the \skewt process \citep{Padoan2011}.
-\hl{Need to add in more references here to skew-$t$ used in a spatial setting.}
 We use a \skewt distribution because of its flexibility to model asymmetry and heavy-tailed data with the aim of predicting the probability of exceeding a high fixed level at an unobserved location.
@@
 This allows us to use Bayesian methods, which we use to fit the model, handle censored data below the threshold, and make predictions at unobserved locations.
+The multivariate skew normal distribution was introduced by \citet{Azzalini1996}, and this was extended to the multivariate \skewt by \citet{Branco2001}.
+These skew-elliptical distributions have been used in the spatial setting \citep{Genton2004,Kim2004}.
+\citet{Zhang2010} propose the skew-Gaussian process as a class of stationary processes that have skewed marginal distributions.
+\citet{Padoan2011} examined the usage of skew-Gaussian and \skewt distributions for multivariate extremes.
 In a spatial setting, the multivariate \skewt distribution demonstrates asymptotic dependence between observations at all sites regardless of the distance between the sites.
@@
 This process is desirable because of its flexible tail that is controlled by the skewness parameter and degrees of freedom.
-One major difference between the parameterization given in \eref{steq:fullmodel} and other parameterizations is that we characterize the skewnewss using $\lambda$ instead of the slant parameters $\balpha$.
-For a more complete comparison with other parameterizations, see \aref{sta:otherparams}.
+% One major difference between the parameterization given in \eref{steq:fullmodel} and other parameterizations is that we characterize the skewnewss using $\lambda$ instead of the slant parameters $\balpha$.
+For a comparison with other parameterizations, see \aref{sta:otherparams}.
 
@@
 where $\chi_{\text{skew-}t}(h)$ is the $\chi$ statistic for a \skewt process given in equation \eref{steq:chiskew-t} of \aref{sta:skewt}, and $h = ||\bs_1 - \bs_2||$.
-Marginally, over the knots, $\chi(h) = \pi(h) \chi_{\text{skew}-t}(h)$, where $\pi(h) = \Pr(k_1 = k_2)$ is the probability that two sites separated by distance $h$ are in the same partition.
+Marginally, over the knots, $\chi(h) = \pi(h) \chi_{\text{skew}-t}(h)$, where \mbox{$\pi(h) = \Pr(k_1 = k_2)$} is the probability that two sites separated by distance $h$ are in the same partition.
 In \aref{sta:proofsamepartition}, we show that $\lim_{h \rightarrow \infty} \pi(h) = 0$, assuming that the knots follow a homogeneous Poisson process.
@@
 \subsection{Computation}\label{sts:comp}
-We use MCMC methods to explore the posterior.
+We use MCMC methods to explore the posterior distribution of the model parameters.
 At each MCMC iteration, we first impute values below the threshold conditional on observations above the threshold.
@@
 
-Then, we update model parameters, $\Theta$, using a Metropolis-Hastings algorithm with Gibbs sampling when needed.
-The final step of the computation is to use Bayesian Kriging to generate a predictive distribution for $Y_t(\bs^*)$ at prediction location $\bs^*$.
+We update model parameters $\Theta$ using Gibbs sampling with Metropolis-Hastings steps when needed.
+In our case, we also wish to be able to make predictions at sites where we do not have data.
+We can easily implement Bayesian Kriging as a part of the algorithm to generate a predictive distribution for $Y_t(\bs^*)$ at prediction location $\bs^*$.
 This step is similar to the imputation for censored observations except that the full conditionals are no longer truncated at $T$.
@@
 For all simulation designs, we generated data from model \eref{steq:partition} in \sref{sts:part} using $n_s=144$ sites and $n_t=50$ independent days.
-The sites were generated Uniform$([0, 10] \times [0, 10])$.
-We generated data from 5 different simulation designs:
+The sites were generated uniformly on the square $[0, 10] \times [0, 10]$.
+We generated data from five different simulation designs:
 \begin{enumerate}[leftmargin=3em]
-  \item Gaussian marginal, $K=1$ knot
+  \item Gaussian, $K=1$ knot
   % \item Symmetric-$t$ marginal, $K=1$ knot
   % \item Symmetric-$t$ marginal, $K=5$ knots
-  \item \Skewt marginal, $K=1$ knots
-  \item \Skewt marginal, $K=5$ knots
-  \item Max-stable with GEV(1, 1, 0.2) marginals \citep{Reich2012}
+  \item \Skewt, $K=1$ knots
+  \item \Skewt, $K=5$ knots
+  \item Multivariate asymmetric logistic distribution \citep{Stephenson2003} %with GEV(1, 1, 0.2) marginals with dependence parameteter $\gamma_{\text{AL}} = 0.5$
   % \item Transformation below $T = q(0.80)$
-  \item A Brown-Resnick process with range 1 and smoothness 0.5
+  \item Brown-Resnick max-stable process \citep{Kabluchko2009} %with unit Fréchet marginals and stable semivariogram with range 1 and smoothness 0.5
 \end{enumerate}
@@
 In designs 1 -- 3, the mean $\bX^\top \bbeta = 10$ was assumed to be constant across space.
-In the fourth design, we generated from a spatial max-stable distribution \citep{Reich2012}.
+In the fourth design, we generated from a multivariate asymmetric logistic distribution \citep{Stephenson2003}.
 In this design, data have marginal distributions that follow a generalized extreme value distribution with location parameter $1$, scale parameter $1$, and shape parameter $0.2$.
 In this model, a random effect was used to induce spatial dependence using 144 spatial knots on a regular lattice in the square $[1, 9] \times [1, 9]$.
-For this setting, $\gamma_{\text{RS}} \in (0, 1)$, is similar to $\gamma$ in that it controls the relative strength of the nugget effect.
-We set $\gamma_{\text{RS}} = 0.5$, which represents moderate spatial dependence.
+For this setting, we set the dependence parameter $\gamma_{\text{AL}} = 0.5$ which gives moderate spatial dependence.
+For the asymmetry parameters, we use standardized Gaussian as in \citet{Reich2012} with bandwidth 1.
+The dependence parameter $\gamma_{\text{AL}} \in (0, 1)$, is similar to $\gamma$ in that it controls the relative strength of the nugget effect.
+For the final design, we generate data from a Brown-Resnick max-stable process \citep{Kabluchko2009} using \texttt{rmaxtab} in the \texttt{SpatialExtremes} package of \texttt{R} \citep{Ribatet2015}.
+For this design we give choose unit \Frechet margins, and use a range of 1 with smoothness 0.5.
 % In the final design, we generate $\tilde{y}$ using the setting from design 2, and then transform the data
@@
 The design matrix $\bX$ includes an intercept with a first-order spatial trend with priors of $\beta_\text{int}$, $\beta_\text{lat}$, $\beta_\text{long},  \iid \text{N}(0, 10)$ although only the intercept is used in the data generation.
-The spatial covariance parameters have priors $\log(\nu) \sim \text{N}(-1.2, 1)$, $\gamma \sim \text{Unif}(0, 1)$, $\varphi \sim \text{Unif}(15)$.
+The spatial covariance parameters have priors $\log(\nu) \sim \text{N}(-1.2, 1)$, $\gamma \sim \text{Unif}(0, 1)$, $\varphi \sim \text{Unif}(0, 15)$.
 The skewness parameter has prior $\lambda \sim N(0, 20)$.
@@
 Each chain of the MCMC ran for 20,000 iterations with a burn-in period of 10,000 iterations.
-Although the goal of this simulation study is not to assess parameter estimation, for design (2) and method (2) where data are generated and fit with a \skewt distribution, the sample converges for $\lambda$ and $a$, and the empirical coverage of posterior 95\% intervals is near the nominal level (96\% coverage for $\lambda$ and 90\% for $a$).
+Although the goal of this simulation study is not to assess parameter estimation, for design (2) and method (2) where data are generated and fit with a \skewt distribution, the samples converges for $\lambda$ and $a$, and the empirical coverage of posterior 95\% intervals is near the nominal level (96\% coverage for $\lambda$ and 90\% for $a$).
 It should be noted that in the models with multiple partitions (i.e.
 models 4 and 5) it is hard to assess the convergence of $\bw$, $z(\bs)$, and $\sigma^2(\bs)$ because of partition label switching throughout the MCMC; however, we are not interested in these parameters but rather spatial predictions and tail probabilities which converge well.
+Finally, we do not fit a Brown-Resnick model because we cannot use a Bayesian approach with so many sites.
 
@@
 \subsection{Results}\label{sts:simresults}
-We compared the Brier scores for exceeding 4 different high levels for each dataset.
+We compared the Brier scores for exceeding four different high levels for each dataset.
 The levels used for the Brier scores are extreme quantiles from the simulated data for $L = q(0.90)$, $q(0.95)$, $q(0.98)$, $q(0.99)$.
@@
 In general, when the underlying process is not Gaussian, our method results in an improvement over both the max-stable and Gaussian methods.
-We also see that the non-thresholded methods tend to outperform the thresholded methods, but this is not surprising given that in most cases, the data are generated directly from the model used in the method.
-Finally, in the case where the data are generated using a Brown-Resnick process, we find that our method is competitive with using a max-stable model.
+We also see that the non-thresholded methods tend to outperform the thresholded methods, but this is not surprising given that in most cases, the data are generated directly from the model.
+Finally, in the case where the data are generated from a Brown-Resnick process, we find that our method is competitive with using a max-stable model.
 % Finally, for setting 5, although the thresholded version of the single-partition model tends to perform the best across all of the extreme quantiles, the difference between the thresholded and non-thresholded methods is no longer significant in the more extreme quantiles.
-In summary, our method provides more flexibility for data that demonstrate some level of asymmetry or heavy tails, while still performing comparably to Gaussian methods when the data are symmetric and have light tails.
+In summary, our method provides great flexibility for data that demonstrate some level of asymmetry and heavy tails, while still performing comparably to Gaussian methods when the data are symmetric and have light tails.
 
@@
 % \end{align}
-\fref{stfig:ozone-qq} shows a Q-Q plot of the residuals compared to a \skewt distribution with 10 d.f. and $\alpha = 1$.
+\fref{stfig:ozone-qq} shows a Q-Q plot of the residuals compared to a \skewt distribution with $a = 10$ and $\lambda = 1$.
 \begin{figure}
   \centering
-  \includegraphics[width=0.5\linewidth]{plots/qq-res.pdf}
-  \caption{Q-Q plot of the residuals for a \skewt distribution with 10 d.f. and $\alpha = 1$ (right)}
+  \includegraphics[width=\linewidth]{plots/qq-res.pdf}
+  \caption{Gaussian Q-Q plot (left) and \skewt distribution with $a = 10$ and $\lambda = 1$ Q-Q plot (right) for of the residuals.}
   \label{stfig:ozone-qq}
@@
 \end{figure}
-Another aspect that distinguishes our approach from more traditional extremes analyses, is how the threshold is selected.
-In our example, a threshold of 75 ppb which corresponds to $q(92)$ for all observations, but marginally it represents anywhere from $q(0.06)$ to $q(1)$.
+% Another aspect that distinguishes our approach from more traditional extremes analyses, is how the threshold is selected.
+% In our example, a threshold of 75 ppb which corresponds to $q(92)$ for all observations, but marginally it represents anywhere from $q(0.06)$ to $q(1)$.
 
@@
 The most noticeable differences between the reference methods and the comparison methods is that the comparison methods tend to give higher estimates of the 99th quantile along the I-85 corridor between Charlotte and Atlanta.
+Among these methods, the fourth method demonstrates the best performance.
+For a map of Brier scores for the 99th quantile between Gaussian and the fourth method, see \aref{sta:ozonesite}.
 
@@
 \section*{Supplementary Materials}
-Web Appendices A--E are available with this paper at the Biometrics website on Wiley Online Library. \vspace*{-8pt}
+Web Appendices A--E are available with this paper at the Biometrics website on Wiley Online Library.
+\texttt{R} code implementing the spatial \skewt model for threshold exceedances is available with this paper at the Biometrics website on Wiley Online Library.\vspace*{-8pt}
 

### 2016-06-29 LaTeX/supplemental_rev.tex
updates to manuscript for revision

@@
 Define $A$ as the circle centered at $c$ with radius $h / 2$ where $h = ||\bs_1 - \bs_2||$ is the distance between sites $\bs_1$ and $\bs_2$.
-Consider a homogeneous spatial Poisson process over $A$ with intensity given by
+Consider a homogeneous spatial Poisson process over $A$ with intensity $\lambda_{PP}$, so that
 \begin{align*}
@@
 since knots in both $B_1$ and $B_2$ is sufficient, but not necessary, to ensure that $\bs_1$ and $\bs_2$ are in different partition sets.
-By definition of a Poission process, $N_1$ and $N_2$ are independent and thus $P(N_1 > 0, N_2 > 0) = P(N_1 > 0)^2$, and the intensity measure over $B_1$ is given by
+By definition of a Poission process, $N_1$ and $N_2$ are independent and thus $P(N_1 > 0, N_2 > 0) = P(N_1 > 0)^2$, and
 \begin{align}
@@
 To extend this to the \skewt distribution, \citet{Azzalini2003} take $\tilde{Y}(\bs) = W\tilde{X}(\bs)$ where $W^{-2} \sim $ Gamma$(a / 2, a / 2)$.
-Returning to the proposed parameterization (with $\bbeta = 0$), let $W^{-2} = \frac{b}{a}\sigma^{-2} \sim$ Gamma$(a / 2, a / 2)$ so that \eref{steq:fullmodel} in the manuscript becomes
+Returning to the proposed parameterization (with $\bbeta = 0$), let $W^{-2} = \frac{b}{a}\sigma^{-2}$ so that \eref{steq:fullmodel} in the manuscript becomes
 \begin{align}
@@
 \end{align}
-Clearly setting $b = a (1 - \tilde{\lambda}^2) > 0$, and $\lambda = \tilde{\lambda} / (1 - \tilde{\lambda}^2)^{1 / 2} \in (-\infty, \infty)$ resolves the difference in parameterizations.
+Clearly setting $b / a = (1 - \tilde{\lambda}^2) > 0$, and $\lambda = \tilde{\lambda} / (1 - \tilde{\lambda}^2)^{1 / 2} \in (-\infty, \infty)$ resolves the difference in parameterizations.
 We note that our parameterization has three parameters $(a, b, \lambda)$ compared to the two parameters of the alternative parameterization $(a, \tilde{\lambda})$.
@@
 \section{Brier scores for ozone prediction} \label{sta:ozonesite}
-Because typical ozone concentration varies throughout the US, we have included Brier scores by site for two model fits (Gaussian and \Skewt, $K = 5$, $T = 50$) in \fref{stfig:bssite}.
+Because typical ozone concentration varies throughout the US, we have included Brier scores for exceedance of the 99th quantile by site for two model fits (Gaussian and Symmetric-$t$, $K = $ 10 knots, $T = $ 75, time series) in \fref{stfig:bssite}.
 As we can see in these plots, both models seem to perform similarly across the US with the poorest performance in California.
@@
   \includegraphics[width=0.8\linewidth]{plots/bs-site}
-  \caption{Map of Brier scores for Gaussian (top) vs \Skewt, $K = 5$, $T = 50$ (bottom).}
+  \caption{Map of Brier scores for Gaussian (top) vs Symmetric-$t$, $K = $ 10 knots, $T = $ 75, time series (bottom).}
   \label{stfig:bssite}
@@
   \centering
-  \caption{Setting 4 -- Max-stable}
+  \caption{Setting 4 -- Max-stable, Asymmetric logistic}
   \label{sttbl:mssim}
@@
   \centering
-  \caption{Setting 5 -- Brown Resnick}
+  \caption{Setting 5 -- Max-stable, Brown-Resnick}
   \label{sttbl:transsim}

### 2016-06-29 code/analysis/ozone/US-all/us-all-setup.R

@@ -323,17 +323,34 @@ omega  <- exp(fit[2])  # logscale
 lambda <- fit[3]  # slant
 nu     <- fit[4]  # degrees of freedom
 
-xplot <- qst(theory.qq, xi = xi, omega = omega, alpha = lambda, nu = nu)
+these <- c(1:1000, sample(1001:22984, size = 2000), 22985:length(xplot))
+par(mfrow = c(1, 2))
+xplot <- qnorm(theory.qq)
+xplot <- (xplot - mean(xplot))
+plot(xplot[these], res.qq[these],
+     xlab = "Theoretical Quantile", ylab = "Observed Quantile",
+     cex = 1.5, cex.lab = 1.5, cex.axis = 1.5, cex.main = 2)
+abline(0, 1)
+
+xplot <- qt(theory.qq, 10)
+xplot <- (xplot - mean(xplot))
+plot(xplot[these], res.qq[these],
+     xlab = "Theoretical Quantile", ylab = "Observed Quantile",
+     cex = 1.5, cex.lab = 1.5, cex.axis = 1.5, cex.main = 2)
+abline(0, 1)
+
+# xplot <- qst(theory.qq, xi = xi, omega = omega, alpha = lambda, nu = nu)
 xplot <- qst(theory.qq, alpha = 1, nu = 10)
 # xplot <- qst(theory.qq, alpha = 0.304, nu = 10)
-these <- c(1:1000, sample(1001:22984, size = 2000), 22985:length(xplot))
+
 xplot <- (xplot - mean(xplot))
-plot(xplot[these], res.qq[these], 
-     xlab="Theoretical Quantile", ylab="Observed Quantile", 
+plot(xplot[these], res.qq[these],
+     xlab="Theoretical Quantile", ylab="Observed Quantile",
      #main="Q-Q plot: Skew-t with 10 d.f. and alpha = 1"
      cex=1.5, cex.lab=1.5, cex.axis=1.5, cex.main=2)
 abline(0, 1)
-dev.print(width = 6, height = 6, device = pdf, file = "plots/qq-res.pdf")
+
+dev.print(width = 12, height = 6, device = pdf, file = "plots/qq-res.pdf")
 
 # ML - Skew-t fit and diagnostics
 library(fields)

### 2016-06-30 LaTeX/response.tex
edits to prepare for submitting the revision

@@
   With the reparameterization given in the response to Major Comment 1, this actually means that the prior for $a$ is now a discrete distribution over a mesh from 0.2 to 20 with spacing $0.2$.
-  As we now clarify in the specification of these priors, $a$ is the scale-free shape parameter, and so this prior should be uninformative in all applications.
+  As we now clarify in the specification of these priors both below equation \eref{steq:fullmodel} of the main text as well as in \sref{sts:simdesign} when describing the models we fit, $a$ is the degrees of freedom, and so this prior should be uninformative in all applications.
   In contrast, $b$ is a precision parameter, so its prior should be adapted to the scale of the response.
@@
 
-  \item Page 8, line -1: a partition in the $\bw^∗$ space induces a partition in the $\bw^∗$.
-The property (7) is used in the $\bw^∗$ space.
+  \item Page 8, line -1: a partition in the $\bw^*$ space induces a partition in the $\bw^*$.
+The property (7) is used in the $\bw^*$ space.
 Is this property still respected in the $\bw$ space since the transformation is non-linear and acts independently in each coordinate?
@@
   \begin{response}
-    When we create the partition, it is created in the space of the original data (i.e. using $\bw$).
-    As you mention, because the transformation is non-linear and acts independently in each coordinate, the partition created using $\bw$ will be similar to, but not the same as the partition created using $\bw^*$.
-    We have added clarification in \sref{sts:temporal} to clarify that the partition is created in the $\bw$ space as opposed to the $\bw^*$ space.
+    As we now clarify in the final sentence of \sref{sts:temporal}, by construction the entire process including the prior on the partition is the same as the space-only model of \sref{sts:part}
     % \hl{Is this something we should be concerned about?}
@@
     Yes, we can see how this could lead to confusion.
-    In the revision, we now refer to the ``threshold'' as the value we select in the model-fitting process, i.e. $T$ in \eref{steq:Yt}.
-    We also bolstered the description of the trade-off involved in selecting $T$ (see AE comment 4) below \eref{steq:Yt}.
+    In the revision, we now refer to the ``threshold'' as the value we select in the model-fitting process, i.e. $T$ in \sref{sts:censoring}.
+    We also bolstered the description of the trade-off involved in selecting $T$ (see AE comment 4) in this section.
     As mentioned in the response to the previous comment, we also want to compute probabilities of extreme events (e.g. ozone exceeding 75ppb).

### 2016-06-30 LaTeX/skewt_rev.tex
edits to prepare for submitting the revision

@@
 In the case of ozone, the level for compliance is fixed at 75 parts per billion, which is high, but not extreme at all locations.
-We present a new space-time model for threshold exceedances based on the spatial \skewt process.
+We present a new space-time model for threshold exceedances based on the \skewt process.
 Our method incorporates a random partition to permit long-distance asymptotic independence while allowing for sites that are near one another to be asymptotically dependent, and we incorporate thresholding to allow the tails of the data to speak for themselves.
 We also introduce a transformed AR(1) time-series to allow for temporal dependence.
-Finally, our model allows for high-dimensional Bayesian inference that is comparable in speed to traditional geostatistical methods for large datasets.
+Finally, our model allows for high-dimensional Bayesian inference that is comparable in computation time to traditional geostatistical methods for large datasets.
 We apply our method to an ozone analysis for July 2005, and find that our model improves over both Gaussian and max-stable methods in terms of predicting exceedances of a high level.
@@
 As a result, the Environmental Protection Agency (EPA) has developed a set of standards to help reduce air pollution thereby improving air quality.
-Our study is motivated by an air pollution application where the focus is not on the average behavior, but instead the behavior over a high level determined by government regulation.
+Our study is motivated by an air pollution application where the focus is not on the average behavior, but on the behavior over a high level determined by government regulation.
 More specifically, we consider the case of compliance for ozone.
@@
 Threshold based inference for multivariate extreme value distributions was considered by \citet{Ledford1996} who introduced a censored approach that provides a way to deal with different types of exceedances of a threshold.
-These threshold models were extended to spatial models for extremes by \citet{Wadsworth2012} and \citet{Thibaud2013} who fit various models to spatial extremes using a censored pairwise likelihood \citep{Padoan2010} based on the approach of \citet{Ledford1996}.
+These models were extended to spatial models for threshold exceedances by \citet{Wadsworth2012} and \citet{Thibaud2013} who fit various models to spatial extremes using a censored pairwise likelihood \citep{Padoan2010} based on the approach of \citet{Ledford1996}.
 \citet{Huser2014} further extended this to space-time modeling.
 \citet{Wadsworth2014}, \citet{Engelke2015}, and \citet{Thibaud2013a} introduced more efficient inference for threshold exceedances of extremal spatial processes with full likelihood methods.
-The previous approaches to threshold modeling are motivated by extreme value theory and assume the threshold is high enough that extremal models are valid for the data, and for extrapolation beyond the range of observed values.
+The previous approaches to threshold modeling are motivated by extreme value theory and assume the threshold is high enough that extremal models are valid for the data and for extrapolation beyond the range of observed values.
 Moreover, these approaches are computationally intensive and limited to rather small datasets.
-For example, \citet{Wadsworth2014} present a simulation study with observations at 16 sites on a regular grid, and \citet{Engelke2015} analyze a dataset with observations at 35 meteorological stations.
-Our application with ozone data does not fit into this framework because we do not focus on exceedances of a very high level, and we have observations at over 1,000 ozone monitoring locations.
+% For example, \citet{Wadsworth2014} present a simulation study with observations at 16 sites on a regular grid, and \citet{Engelke2015} analyze a dataset with observations at 35 meteorological stations.
+Our application with ozone data does not fit into this framework because we do not focus on exceedances of a very high level, and we have observations at 1,089 ozone monitoring locations.
 
 We propose a new spatiotemporal threshold exceedance model based on the \skewt process \citep{Padoan2011}.
-We use a \skewt distribution because of its flexibility to model asymmetry and heavy-tailed data with the aim of predicting the probability of exceeding a high fixed level at an unobserved location.
+We use a \skewt distribution because of its flexibility to model asymmetry and heavy-tailed data with the aim of modeling exceedances of a high fixed level at an unobserved location.
 Our model allows for inference and predictions using the full likelihood with computing on the order of Gaussian models.
@@
 In order to address this concern, we introduce a random spatial partition similar to the method used by \citet{Kim2005} for non-stationary Gaussian data.
-This partitioning alleviates the asymptotic spatial dependence present in the \skewt distribution for sites that are far apart.
+% This partitioning alleviates the asymptotic spatial dependence present in the \skewt distribution for sites that are far apart.
 
@@
 \sref{sts:spatialskew} is a brief review of the spatial \skewt process.
-In \sref{sts:spatial}, we build upon the traditional \skewt process by incorporating censoring to focus on tails, partitioning to remove long-range asymptotic dependence, and extending the model to space-time data.
+In \sref{sts:spatial}, we build upon the traditional \skewt process by incorporating censoring, partitioning, and extending the model to space-time data.
 The computing is described in \sref{sts:hier}.
 In \sref{sts:simstudy}, we present a simulation study that examines the predictive capabilities of this model compared to Gaussian and max-stable methods.
-We then compare our method to Gaussian and max-stable methods with a data analysis of ozone measurements from 1,000 sites throughout the US in \sref{sts:analysis}.
-The final section provides brief discussion and direction for future research.
+We compare our method to Gaussian and max-stable methods with a data analysis of ozone measurements throughout the US in \sref{sts:analysis}.
+% The final section provides brief discussion and direction for future research.
 
@@
 \end{align}
-where $\bX(\bs)$ is the observed covariate vector at site $\bs$, $\bbeta$ is the $p$-vector of regression parameters, $\lambda \in \calR$ is a parameter controlling skewness, $z \sim N(0, 1)$, $\sigma^2 \sim \text{IG}(a / 2, b / 2)$ is random scale parameter, IG is the distribution function of an inverse gamma random variable, $a$ is the degrees of freedom, $b$ controls the precision of the process, and $v(\bs)$ is a spatial Gaussian process with mean zero, variance one, and a positive definite correlation function $\text{Cor} \left[Y(\bs_1), Y(\bs_2)\right] = \rho(\bs_1, \bs_2)$.
+where $\bX(\bs)$ is the observed covariate vector at site $\bs$, $\bbeta$ is the $p$-vector of regression parameters, $\lambda \in \calR$ is a parameter controlling skewness, $z \sim N(0, 1)$, $\sigma^2 \sim \text{IG}(a / 2, b / 2)$ is random scale parameter, IG is the distribution function of an inverse gamma random variable, $a$ is the degrees of freedom, $b$ controls the precision of the process, and $v(\bs)$ is a standard Gaussian process with positive definite correlation function $\text{Cor} \left[Y(\bs_1), Y(\bs_2)\right] = \rho(\bs_1, \bs_2)$.
 Although any positive definite correlation function could be used, we choose to use the stationary isotropic \Matern correlation with
@@
 \end{align}
-where $I(\cdot)$ is an indicator function, $\varphi$ is the spatial range, $\nu$ is the smoothness, $\gamma$ is the proportion of variance accounted for by the spatial variation, $K_\nu$ is a modified Bessel function of the second kind, and $h = || \bs_1 - \bs_2 ||$ is the Euclidean distance between sites $\bs_1$ and $\bs_2$.
+where $I(\cdot)$ is an indicator function, $\varphi > 0$ is the spatial range, $\nu > 0$ is the smoothness, $\gamma \in [0, 1]$ is the proportion of variance accounted for by the spatial variation, $K_\nu$ is a modified Bessel function of the second kind, and $h = || \bs_1 - \bs_2 ||$ is the Euclidean distance between sites $\bs_1$ and $\bs_2$.
 
@@
 \end{align}
-that is, $\bY$ follows an $n$-dimensional \skewt distribution with location $\bX \bbeta \in \calR^n$; covariance matrix $\bOmega_{n \times n} = \bomega \bar{\bOmega} \bomega$ where \mbox{$\bar{\bOmega}_{n \times n} = \frac{1}{1 + \lambda^2}(\bSigma + \lambda^2 \bOne \bOne^\top)$}, $\bSigma_{n \times n}$ is the positive definite correlation matrix which is obtained from $\rho(h)$, and $\bomega_{n \times n} = \text{diag}\left(\sqrt{\frac{ b }{a(1 + \lambda^2)}}, \ldots, \sqrt{\frac{ b }{a(1 + \lambda^2)}} \right)$; slant parameters $\balpha \in \calR^n = \lambda (1 + \lambda^2)^{1 / 2} (1 + \lambda^2 \bOne^\top \bSigma^{-1} \bOne)^{-1/2} \bSigma^{-1} \bOne$, and degrees of freedom $a$.
+that is, $\bY$ follows an $n$-dimensional \skewt distribution with location $\bX \bbeta \in \calR^n$; covariance matrix $\bOmega_{n \times n} = \bomega \left[ \frac{1}{1 + \lambda^2}(\bSigma + \lambda^2 \bOne \bOne^\top) \right] \bomega$, $\bSigma_{n \times n}$ is the positive definite correlation matrix which is obtained from $\rho(h)$, and $\bomega_{n \times n} = \text{diag}\left(\sqrt{\frac{ b }{a(1 + \lambda^2)}}, \ldots, \sqrt{\frac{ b }{a(1 + \lambda^2)}} \right)$; slant parameters $\balpha \in \calR^n = \lambda (1 + \lambda^2)^{1 / 2} (1 + \lambda^2 \bOne^\top \bSigma^{-1} \bOne)^{-1/2} \bSigma^{-1} \bOne$, and degrees of freedom $a$.
 Furthermore, the marginal distributions at each location also follow a univariate \skewt distribution \citep{Azzalini2014}.
-This process is desirable because of its flexible tail that is controlled by the skewness parameter and degrees of freedom.
+This process is desirable because it is heavy tailed with tail index $a$, and the shape of the distribution is controlled by the skewness parameter.
 % One major difference between the parameterization given in \eref{steq:fullmodel} and other parameterizations is that we characterize the skewnewss using $\lambda$ instead of the slant parameters $\balpha$.
@@
 One measure of extremal dependence is the $\chi$ statistic \citep{Coles1999}.
-For a stationary and isotropic spatial process, the $\chi$ statistic for locations $\bs_1$ and $\bs_2$ separated by distance $h = || \bs_1 - \bs_2 ||$ with identical marginal distributions is
+For a stationary and isotropic spatial process, the $\chi$ statistic for two locations separated by distance $h$ is
 \begin{align} \label{steq:chih}
-  \chi(h) = \lim_{c \rightarrow c^*} \Pr[Y(\bs_1) > c | Y(\bs_2) > c]
+  \chi(h) = \lim_{c \rightarrow c^*} \Pr[Y(\bs + h) > c | Y(\bs) > c]
 \end{align}
-where $c^*$ is the upper limit of the support of $Y$.
-For the \skewt distribution, $c^* = \infty$.
+where $c^*$ is the upper limit of the support of $Y$; for the \skewt distribution $c^* = \infty$.
 If $\chi(h) = 0$, then observations are asymptotically independent at distance $h$.
@@
 \subsection{Censoring to focus on the tail} \label{sts:censoring}
-As mentioned previously, we propose to use a censored approach because we do not want the low-to-moderate values to influence the fit of the overall model.
-% The censored observations below the threshold give information on the marginal probabilities to exceed the threshold and on the dependence, but their values are not used to fit the model.
-Let
-\beq \label{steq:Yt}
-  \widetilde{Y}(\bs) = \left\{ \begin{array}{ll}
-      Y(\bs) \quad & \delta(\bs) = 1 \\
-      T & \delta(\bs) = 0
-  \end{array} \right.
-\eeq
-be the censored process at site $\bs$ where $Y(\bs)$ is the uncensored process, $\delta(\bs) = I[Y(\bs) > T]$, and $T$ is a pre-specified threshold value.
+We do not want the low-to-moderate values to influence the fit of the model.
+We propose the use of a censored approach to fit threshold exceedances only.
+More specifically, we assume our \skewt model $Y(\bs)$ is valid at each location $\bs$ above a threshold $T$, and censor the values below $T$ for which we don't assume the model to be valid.
+We define our partially censored observations as $\widetilde{\bY} = [\widetilde{Y}(\bs_1), \ldots, \widetilde{Y}(\bs_n)]^\top$ where $\widetilde{Y}(\bs) = \max\{Y(\bs), T\}$, and fit the \skewt process to these $\widetilde{\bY}$.
+In our Bayesian framework, inference can be easily performed by imputing censored observations below $T$ (see \sref{sts:comp}).
+% We propose to use a censored approach because we do not want the low-to-moderate values to influence the fit of the model.
+% % The censored observations below the threshold give information on the marginal probabilities to exceed the threshold and on the dependence, but their values are not used to fit the model.
+% % \beq \label{steq:Yt}
+% %   \widetilde{Y}(\bs) = \left\{ \begin{array}{ll}
+% %       Y(\bs) \quad & \delta(\bs) = 1 \\
+% %       T & \delta(\bs) = 0
+% %   \end{array} \right.
+% % \eeq
+% We will assume our data were generated from the censored process $\widetilde{Y}(\bs) = \max\{Y(\bs), T\}$ where $T$ is a pre-specified threshold value.
+% Then from the censored process, the vector of observations becomes $\widetilde{\bY} = [\widetilde{Y}(\bs_1), \ldots, \widetilde{Y}(\bs_n)]^\top$.
+
 As our goal is to model exceedances above a high level $L$, we should select a value for $T \le L$.
@@
 \subsection{Partitioning to remove long-range asymptotic dependence}\label{sts:part}
-The motivation for the partition is that for a large spatial domain, it may not be reasonable to assume sites that are far apart demonstrate asymptotic dependence.
-Modeling different levels of asymptotic dependence was discussed by \citet{Wadsworth2012}.
-\citet{Huser2014} also allow for varying asymptotic dependence across both space and time with a partition structure represented by random discs moving across the space for a random duration with a random velocity and random radius.
-We handle the problem of long-range asymptotic dependence with a random partition.
+For a large spatial domain, it may not be reasonable to assume sites that are far apart demonstrate asymptotic dependence.%, so we handle the problem of long-range asymptotic dependence with a random partition.
+% Modeling different levels of asymptotic dependence was discussed by \citet{Wadsworth2012}.
+% \citet{Huser2014} also allow for varying asymptotic dependence across both space and time with a partition structure represented by random discs moving across the space for a random duration with a random velocity and random radius.
 As discussed in \sref{sts:spatialskew}, the source of long-range dependence is the shared $z$ and $\sigma$.
-Therefore, to alleviate this dependence, we allow $z$ and $\sigma$ to vary by site.
+Therefore, to alleviate this dependence, we allow $z$ and $\sigma$ to vary by site using a partitioning approach.
 The model becomes
@@
 So, within each partition, $Y(\bs)$ follows the spatial \skewt process defined in \sref{sts:spatialskew}.
-Across partitions, the $Y(\bs)$ remain dependence via the correlation function for $v(\bs)$ because $v(\bs)$ spans all partitions.
+Across partitions, the $Y(\bs)$ remain dependent via the correlation function for $v(\bs)$ because $v(\bs)$ spans all partitions.
 However, the bivariate distribution for sites in different partitions is neither Gaussian nor \skewt and does not have asymptotic dependence.
 
-The partitioning model removes long-range dependence. Conditional on knots $\bw_1, \ldots, \bw_K$, the $\chi$ statistic for two sites $\bs_1$ and $\bs_2$ in partitions $k_1$ and $k_2$ respectively is
+The partitioning model removes long-range dependence.
+Conditional on knots $\bw_1, \ldots, \bw_K$, the $\chi$ statistic for two sites $\bs_1$ and $\bs_2$ in partitions $k_1$ and $k_2$ respectively is
 \begin{align}
@@
 Marginally, over the knots, $\chi(h) = \pi(h) \chi_{\text{skew}-t}(h)$, where \mbox{$\pi(h) = \Pr(k_1 = k_2)$} is the probability that two sites separated by distance $h$ are in the same partition.
-In \aref{sta:proofsamepartition}, we show that $\lim_{h \rightarrow \infty} \pi(h) = 0$, assuming that the knots follow a homogeneous Poisson process.
-This implies that $\lim_{h \rightarrow \infty} \chi(h) = 0$.
+In \aref{sta:proofsamepartition}, we show that assuming the knots follow a homogeneous Poisson process, $\lim_{h \rightarrow \infty} \pi(h) = 0$, and thus long-range dependence is removed.
+% In \aref{sta:proofsamepartition}, we show that $\lim_{h \rightarrow \infty} \pi(h) = 0$, assuming that the knots follow a homogeneous Poisson process.
+% This implies that $\lim_{h \rightarrow \infty} \chi(h) = 0$.
 In practice we fix $K$ at a finite value and use a uniform distribution for the knots $\bw_1, \ldots, \bw_K$.
-In \fref{stfig:chi}, we give $\chi(h)$ for $K = 1, 3, 5, 10$ partitions for a \skewt distribution with $\alpha = 10$, and 3 degrees of freedom.
-In \fref{stfig:chi} we estimate $\pi(h)$ through simulation.
+In \fref{stfig:chi}, we estimate $\chi(h)$ for $K = 1, 3, 5, 10$ partitions for a \skewt distribution with $\alpha = 10$, and 3 degrees of freedom.
+% In \fref{stfig:chi} we estimate $\pi(h)$ through simulation.
 % To estimate $\pi(h)$, we generate 500 sites uniformly over the unit-square.
@@
 % This plot demonstrates how partitioning helps to reduce extremal dependence as $h$ increases.
-
 \begin{figure}
@@
 There are several places where temporal dependence could be incorporated in the model, including the Gaussian process $v_t(\bs)$.
-However, we choose to allow for temporal dependence in the $\bw$, $z$, and $\sigma$ terms because these terms dictate the tail behavior which is our primary focus (see \aref{sta:temporal}).
+However, we choose to allow for temporal dependence in the $\bw$, $z$, and $\sigma$ terms because these terms dictate the tail behavior of the process which is our primary focus (see \aref{sta:temporal} for a discussion of the induced temporal asymptotic dependence).
 Let
@@
 where $t \in \{1, \ldots, n_t\}$ denotes the day of each observation.
-Let \hbox{$\bw_{tk} = (w_{tk1}, w_{tk2})$} be a spatial knot on day $t$, and let $\bw_{t1}, \ldots, \bw_{tK}$ be a collection of spatial knots on day $t$.
+Let \hbox{$\bw_{tk} = (w_{tk1}, w_{tk2})$} be a spatial knot on day $t$, and let $\bw_{t1}, \ldots, \bw_{tK}$ be the collection of spatial knots on day $t$.
 As in \sref{sts:part}, these knots define a daily partition $P_{t1}, \ldots, P_{tK}$, and for $\bs \in P_{tk}$,
@@
 where IG is defined as before.
-The AR(1) process for each tail parameter is $\bw^*_{1k} \sim N_w(0, 1)$, $z^*_{1k} \sim N(0, 1)$, $\sigma^{2*}_{1k} \sim N(0, 1)$, and for $t > 1$ the time series is modeled as
+The AR(1) process for each tail parameter is $\bw^*_{1k} \sim N_2(\bZero, \bI_2)$ where $\bI_2 = \text{diag}(1, 1)$, $z^*_{1k} \sim N(0, 1)$, $\sigma^{2*}_{1k} \sim N(0, 1)$, and for $t > 1$ the time series is modeled as
 \begin{align}
-  \bw^*_{tk} | \bw^*_{t-1, k} &\sim N_2\left[\phi_w \bw^*_{t-1, k}, (1 - \phi_w^2) \right] \label{steq:ts-wstar}\\
+  \bw^*_{tk} | \bw^*_{t-1, k} &\sim N_2\left[\phi_w \bw^*_{t-1, k}, (1 - \phi_w^2) \bI_2 \right] \label{steq:ts-wstar}\\
   z^*_{tk} | z^*_{t-1, k} &\sim N \left[\phi_z z^*_{t-1, k}, (1 - \phi_z^2)\right] \label{steq:ts-zstar} \\
@@
 where $|\phi_w|$, $|\phi_z|$, $|\phi_\sigma| < 1$.
-These are stationary time series models with marginal distributions \hbox{$\bw^*_{k} \sim N_2(\bZero, \bI_2)$}, where $\bI_2 = \text{diag}(1, 1)$, \hbox{$z^*_{k} \sim N(0, 1)$}, and \hbox{$\sigma^{2*}_{k} \sim N(0, 1)$}.
-After transformation back to the original space, $\bw_{tk} \sim \text{Unif}(\calD)$, $z_{tk} \sim HN(0, 1)$, and $\sigma^2_{tk} \sim \text{IG}(a / 2, b / 2)$.
-We then create the partition for day $t$ using $\bw_{t1}, \ldots, \bw_{tk}$.
-For each day, the model is identical to the spatial-only model in \eref{steq:partition} by construction.\citep{
+These are stationary time series models with marginal distributions \hbox{$\bw^*_{k} \sim N_2(\bZero, \bI_2)$} \hbox{$z^*_{k} \sim N(0, 1)$}, and \hbox{$\sigma^{2*}_{k} \sim N(0, 1)$}.
+After transformation back to the original space, $\bw_{tk} \sim \text{Unif}(\calD)$, $z_{tk} \sim $ HN$(0, 1)$, and $\sigma^2_{tk} \sim \text{IG}(a / 2, b / 2)$.
+We then create the partition for day $t$ using $\bw_{t1}, \ldots, \bw_{tK}$.
+For each day, the model is identical to the spatial-only model in \eref{steq:partition} by construction.
 
 \section{Hierarchical model}\label{sts:hier}
-Conditioned on $z_{tk}(\bs)$, $\sigma^2_{tk}(\bs)$, and $P_{tk}$, the marginal distributions of $Y(\bs)$ are Gaussian and the joint distribution of $Y(\bs)$ is multivariate Gaussian.
-However, we do not fix the partitions, they are treated as unknown and updated in the MCMC algorithm.
-We model this with a Bayesian hierarchical model as follows.
-Let $\bw_{t1}, \ldots, \bw_{tK}$ be a set of daily spatial knots in a spatial domain of interest, $\calD$, and $P_{tk}$ as defined in \eref{steq:subregions}.
+We define a Bayesian hierarchical model based on the \skewt process and use a Markov chain Monte Carlo (MCMC) algorithm to fit the data.
+We model our data as partially censored observations (see \sref{sts:censoring}) from the \skewt model defined in \sref{sts:spatial}.
+In the first step of the MCMC algorithm, we impute censored values of $\widetilde{Y}$ such that the estimation of model parameters can be based on the completed $\bY$.
+% We model our data as observations from the censored \skewt model defined in \sref{sts:censoring}.
+Conditioned on $z_{tk}(\bs)$, $\sigma^2_{tk}(\bs)$, and $P_{tk}$, joint distribution of $\bY$ is multivariate Gaussian.
+We do not fix the partitions; instead, the locations of the $K$ knots are treated as unknown and random.
+% The first step in the algorithm is to update the knot locations, $z_{tk}(\bs)$, $\sigma^2_{tk}(\bs)$.
+% The knot locations are treated as unknown and random in the MCMC algorithm.
+% For level
+%  $\bw_{t1}, \ldots, \bw_{tK}$ be a set of daily spatial knots in a spatial domain of interest, $\calD$, and $P_{tk}$ as defined in \eref{steq:subregions}.
 One approach would be to allow $K$ to be unknown and follow a Poisson process prior, but this would lead to onerous computing.
 Therefore, we elect to treat $K$ as a tuning parameter for the MCMC by fixing it at different values and assessing its impact on prediction as described in \sref{sts:modelselect}.
-Then
+Then the hierarchical model is given as
 \begin{align}
@@
    \sigma^2_{t}(\bs) &= \sigma^2_{tk} \text{ if } \bs \in P_{tk}\nonumber\\
-   \lambda & \sim N(0, \sigma^2_\lambda)\nonumber\\
+   \lambda & \sim N(0, \sigma^2_\lambda)\nonumber \\
    % \lambda_1 &= \left\{ \begin{array}{ll}
@@
    % \lambda^2_2 & \sim IG(a, b)\nonumber\\
-   v_t(\bs) \mid \Theta &\sim \Matern(0, \Sigma)\nonumber\\
-   z^*_{tk} \mid z^*_{t-1, k} &\sim N(\phi_z z^*_{t-1, k}, (1 - \phi_z^2))\nonumber\\
-   \sigma^{2*}_{tk} \mid \sigma^{2*}_{t-1, k} &\sim N(\phi_\sigma \sigma^{2*}_{t-1, k}, (1 - \phi_\sigma^2))\nonumber\\
-   \bw^*_{tk} \mid \bw^*_{t-1, k} &\sim N_2(\phi_w \bw^*_{t-1, k}, (1 - \phi_w^2)) \nonumber
+   v_t(\bs) \mid \Theta &\sim \Matern(\bZero, \bSigma)\nonumber
 \end{align}
-where $\Theta = \{\varphi, \nu, \gamma\, \lambda, \beta\}$, and $\Sigma$ is a \Matern covariance matrix as described in \sref{sts:skewt}.
+% \begin{align*}
+%    z^*_{tk} \mid z^*_{t-1, k} &\sim N(\phi_z z^*_{t-1, k}, (1 - \phi_z^2))\nonumber\\
+%    \sigma^{2*}_{tk} \mid \sigma^{2*}_{t-1, k} &\sim N(\phi_\sigma \sigma^{2*}_{t-1, k}, (1 - \phi_\sigma^2))\nonumber\\
+%    \bw^*_{tk} \mid \bw^*_{t-1, k} &\sim N_2(\phi_w \bw^*_{t-1, k}, \bI_2 (1 - \phi_w^2) ) \nonumber
+% \end{align*}
+where $\Theta = \{\varphi, \nu, \gamma\, \lambda, \bbeta\}$, $\bSigma$ is a \Matern covariance matrix as described in \sref{sts:skewt}, and the priors on $\bw^*_{tk}$, $z^*_{tk}$, and $\sigma^{2*}_{tk}$ are given in \eref{steq:ts-wstar} -- \eref{steq:ts-sigmastar}.
 
 \subsection{Computation}\label{sts:comp}
-We use MCMC methods to explore the posterior distribution of the model parameters.
+We use MCMC methods to estimate the posterior distribution of the model parameters.
 At each MCMC iteration, we first impute values below the threshold conditional on observations above the threshold.
-This is feasible for large datasets with our model because for a single day, conditional on the model parameters, we only need to draw from a truncated multivariate normal distribution.
-We can use Gibbs sampling to update $Y_t(\bs)$ for censored observations that are below the threshold $T$.
-After conditioning on $\lambda$, $z_t(\bs)$ and non-censored observations, $Y_t(\bs)$ has truncated normal full conditionals.
-So we sample $Y_t(\bs) \sim N_{(-\infty, T)}(\bX_t^\top(\bs) \beta + \lambda | z_t(\bs)|, \bSigma)$.
+% This is feasible for large datasets with our model because for a single day, conditional on the model parameters, we only need to draw from a truncated multivariate normal distribution.
+% We can use Gibbs sampling to update $Y_t(\bs)$ for censored observations that are below the threshold $T$.
+After conditioning on $\lambda$, $z_t(\bs)$ and non-censored observations, $Y_t(\bs)$ has truncated normal full conditionals $Y_t(\bs) \sim N_{(-\infty, T)}(\bX_t^\top(\bs) \beta + \lambda | z_t(\bs)|, \bSigma)$.
 
@@
   \item \Skewt, $K=5$ knots
-  \item Multivariate asymmetric logistic distribution \citep{Stephenson2003} %with GEV(1, 1, 0.2) marginals with dependence parameteter $\gamma_{\text{AL}} = 0.5$
+  \item \citet{Reich2012} max-stable process %with GEV(1, 1, 0.2) marginals with dependence parameteter $\gamma_{\text{AL}} = 0.5$
   % \item Transformation below $T = q(0.80)$
@@
 \end{enumerate}
-In the first three designs, the $v_t(\bs)$ terms were generated using a \Matern covariance with smoothness parameter $\nu = 0.5$ and spatial range $\varphi = 1$.
-For the covariance matrices in designs 1 -- 3, the proportion of the variance accounted for by the spatial variation was $\gamma = 0.9$ while the proportion of the variance accounted for by the nugget effect was $0.1$.
+In the first three designs, the realizations from $v_t(\bs)$ were generated using a \Matern covariance with smoothness parameter $\nu = 0.5$, spatial range $\varphi = 1$ and $\gamma = 0.9$.
 In the first design, $\sigma^2 = 2$ was used for all days which results in a Gaussian distribution.
@@
 For design 1, we set $\lambda = 0$.
-For designs 2 and 3, $\lambda = 3$ was used as to simulate moderate skewness, and the $z_t$ were generated as described in Section \sref{sts:part}.
+For designs 2 and 3, $\lambda = 3$ was used as to simulate moderate skewness, and the $z_t$ were generated as described in \sref{sts:part}.
 In designs 1 -- 3, the mean $\bX^\top \bbeta = 10$ was assumed to be constant across space.
-In the fourth design, we generated from a multivariate asymmetric logistic distribution \citep{Stephenson2003}.
-In this design, data have marginal distributions that follow a generalized extreme value distribution with location parameter $1$, scale parameter $1$, and shape parameter $0.2$.
-In this model, a random effect was used to induce spatial dependence using 144 spatial knots on a regular lattice in the square $[1, 9] \times [1, 9]$.
-For this setting, we set the dependence parameter $\gamma_{\text{AL}} = 0.5$ which gives moderate spatial dependence.
-For the asymmetry parameters, we use standardized Gaussian as in \citet{Reich2012} with bandwidth 1.
-The dependence parameter $\gamma_{\text{AL}} \in (0, 1)$, is similar to $\gamma$ in that it controls the relative strength of the nugget effect.
-For the final design, we generate data from a Brown-Resnick max-stable process \citep{Kabluchko2009} using \texttt{rmaxtab} in the \texttt{SpatialExtremes} package of \texttt{R} \citep{Ribatet2015}.
-For this design we give choose unit \Frechet margins, and use a range of 1 with smoothness 0.5.
+In the fourth design, we generated from the max-stable model of \citet{Reich2012}.
+The marginal distributions follow a generalized extreme value distribution with location parameter 1, scale parameter 1, and shape parameter 0.2.
+Spatial dependence in the form an asymmetric logistic dependence function is induced by random effects for kernel basis functions associated with 144 spatial knots defined on a square grid on $[1,9] \times [1,9]$.
+We set the dependence parameter ($\alpha$ in Reich and Shaby, 2012) to 0.5 which represents moderate spatial dependence.
+For the final design, we generated data from a Brown-Resnick max-stable process  using \texttt{rmaxtab} in the \texttt{SpatialExtremes} package of \texttt{R} \citep{Ribatet2015}.
+For this design we fixed unit \Frechet margins, and we used a range of 1 and smoothness 0.5.
 % In the final design, we generate $\tilde{y}$ using the setting from design 2, and then transform the data
@@
 
-$M = 50$ data sets are generated for each design.
+$M = 50$ data sets were generated for each design.
 For each data set we fit the data using six models
@@
   \item Symmetric-$t$ marginal, $K=5$ knots, $T=q(0.80)$
-  \item A max-stable model based on \citet{Reich2012} thresholded at $T = q(0.80)$
+  \item \citet{Reich2012} max-table model thresholded at $T = q(0.80)$
 \end{enumerate}
 where $q(0.80)$ is the 80th sample quantile of the data.
-All methods are fit using a fully-Bayesian approach that simultaneously estimates marginal and spatial dependence parameters.
+All methods were fit using a fully-Bayesian approach that simultaneously estimates marginal and spatial dependence parameters.
 The design matrix $\bX$ includes an intercept with a first-order spatial trend with priors of $\beta_\text{int}$, $\beta_\text{lat}$, $\beta_\text{long},  \iid \text{N}(0, 10)$ although only the intercept is used in the data generation.
@@
 The residual variance terms have priors $\sigma^2_t(\bs) \sim \text{IG}(a / 2, b / 2)$, where $a$ has a discrete uniform prior on a mesh from $0.2$ to $20$ with spacing of $0.1$ and $b$ has a Gamma$(0.1, 0.1)$ prior.
+As described in \sref{sts:skewt}, these priors are meant to be fairly uninformative to allow the data to dictate both the degrees of freedom $a$ and precision $b$ of the process.
 The knots have priors $\bw \sim \text{Unif}(\calD)$.
@@
 models 4 and 5) it is hard to assess the convergence of $\bw$, $z(\bs)$, and $\sigma^2(\bs)$ because of partition label switching throughout the MCMC; however, we are not interested in these parameters but rather spatial predictions and tail probabilities which converge well.
-Finally, we do not fit a Brown-Resnick model because we cannot use a Bayesian approach with so many sites.
+Finally, we did not fit a Brown-Resnick model because we cannot use a Bayesian approach with so many sites.
 
@@
 % \end{align}
-\fref{stfig:ozone-qq} shows a Q-Q plot of the residuals compared to a \skewt distribution with $a = 10$ and $\lambda = 1$.
+\fref{stfig:ozone-qq} shows a Q-Q plot of the residuals compared to a \skewt distribution with $a = 10$ and $\lambda = 1$, suggesting the data are heavy tailed.
 \begin{figure}
@@
   \includegraphics[width=\linewidth]{plots/qq-res.pdf}
-  \caption{Gaussian Q-Q plot (left) and \skewt distribution with $a = 10$ and $\lambda = 1$ Q-Q plot (right) for of the residuals.}
+  \caption{Gaussian Q-Q plot (left) and \skewt with $a = 10$ and $\lambda = 1$ Q-Q plot (right) of the residuals.}
   \label{stfig:ozone-qq}
@@
 We fit the model using Gaussian and \skewt marginal distributions with $K=1, 5, 6, 7, 8, 9, 10, 15$ partitions.
-We choose to censor $Y(\bs)$ at $T = 0$, $T = 50$ (0.42 sample quantile), and $T = 75$ (0.92 sample quantile) ppb in order to compare results from no, moderate, and high censoring.
+We censored $Y(\bs)$ at $T = 0$, $T = 50$ (0.42 sample quantile), and $T = 75$ (0.92 sample quantile) ppb in order to compare results from no, moderate, and high censoring.
 The upper threshold of 75 ppb was used because the current air quality standard is based on exceedance of 75 ppb.
-As with the simulation study, for models with a threshold of $T = 75$, we use a symmetric-$t$ marginal distribution.
-We also compare models with no time series to models that include the time series.
+As with the simulation study, for models with a threshold of $T = 75$, we used a symmetric-$t$ marginal distribution.
+We also compared models with no time series to models that included the time series.
 Finally, as a comparison to max-stable methods, we fit the model using the hierarchical max-stable model of \citet{Reich2012} with the data thresholded at $T = 75$.
-All methods assume $\bX_t(\bs) = [1, \text{CMAQ}_t(\bs)]^\top$.
-To ensure that the max-stable method runs in a reasonable amount of time, we use a stratified sub-sample of 800 sites.
-We conduct two-fold cross validation using 400 training sites and 400 validation sites as described in \sref{sts:modelselect}
+All methods assumed $\bX_t(\bs) = [1, \text{CMAQ}_t(\bs)]^\top$.
+To ensure that the max-stable method ran in a reasonable amount of time, we used a stratified sub-sample of 800 sites.
+We conducted two-fold cross validation using 400 training sites and 400 validation sites as described in \sref{sts:modelselect}
 
 Each chain of the MCMC ran for 30,000 iterations with a burn-in period of 25,000 iterations.
-We use the same priors for the spatial covariance parameters, skewness parameter, and knots as in the simulation study.
-The prior for the residual variance terms is $\sigma^2_t(\bs) \sim \text{IG}(a / 2, b / 2)$ where $a$ is the same as the simulation study, but $b$ has a Gamma$(1, 1)$ prior.
-Parameters appear to converge properly; however, as before, for models with multiple partitions it is hard to assess the convergence of $\bw$, $z(\bs)$, and $\sigma^2(\bs)$ because of partition label switching throughout the MCMC.
-For each model, Brier scores were averaged over all sites and days to obtain a single Brier score for each dataset.
-At a particular level, the model that fits the best is the one with the lowest score.
-We then compute the relative (to Gaussian) Brier scores (see \sref{sts:simresults}) to compare each model.
+We used the same priors for the spatial covariance parameters, skewness parameter, and knots as in the simulation study.
+The prior for the residual variance terms was $\sigma^2_t(\bs) \sim \text{IG}(a / 2, b / 2)$ where $a$ was the same as the simulation study, but $b$ had a Gamma$(1, 1)$ prior.
+Parameters appeared to converge properly; however, as before, for models with multiple partitions it was hard to assess the convergence of $\bw$, $z(\bs)$, and $\sigma^2(\bs)$ because of partition label switching throughout the MCMC.
+For each model, we averaged Brier scores over all sites and days to obtain a single Brier score for each dataset.
+At a particular level, the model that fit the best was the one with the lowest score.
+We then computed the relative (to Gaussian) Brier scores (see \sref{sts:simresults}) to compare each model.
 

### 2016-06-30 code/biometrics/auxfunctions.R

@@ -1,0 +1,520 @@
+################################################################################
+# Common data transformations
+################################################################################
+transform <- list(
+  logit = function(x, lower = 0, upper = 1) {
+    x <- (x - lower) / (upper - lower)
+    return(log(x / (1 - x)))
+  },
+  inv.logit = function(x, lower = 0, upper = 1) {
+    p <- exp(x) / (1 + exp(x))
+    p <- p * (upper - lower) + lower
+    return(p)
+  },
+  probit = function(x, lower = 0, upper = 1) {
+    x <- (x - lower) / (upper - lower)
+    return(qnorm(x))
+  },
+  inv.probit = function(x, lower = 0, upper = 1) {
+    p <- pnorm(x)
+    p <- p * (upper - lower) + lower
+    return(p)
+  },
+  log = function(x) log(x),
+  exp = function(x) exp(x),
+  copula = function(dens) {
+    this.dens <- paste("p", dens, sep = "")
+    function(x, ...) qnorm(do.call(this.dens, args=list(x, ...)))
+  },
+  inv.copula = function(dens) {
+    this.dens <- paste("q", dens, sep = "")
+    function(x, ...) do.call(this.dens, args=list(pnorm(x), ...))
+  }
+)
+
+################################################################################
+# Useful densities
+################################################################################
+
+#################################################################
+# Half normal: Z* = |Z| is HN(0, sig) when Z ~ N(0, sig)
+# Note: the location parameter must be 0
+#################################################################
+dhn <- function(x, sig = 1, log = FALSE) {
+  if (sum(x < 0) > 0) {
+    stop("x must be non-negative")
+  }
+  density <- dnorm(x, 0, sig) / 2
+
+  if (!log) {
+    return(density)
+  } else {
+    return(log(density))
+  }
+}
+
+phn <- function(x, sig = 1, lower.tail = TRUE) {
+  if (sum(x < 0) > 0) {
+    stop("x must be non-negative")
+  }
+  p.val <- 2 * pnorm(x, 0, sig) - 1
[…]

### 2016-06-30 code/biometrics/update_params_cpp.R

@@ -1,0 +1,170 @@
+library(inline)
+
+update_params_cpp_init <- function(x) { NULL }
+
+#### update for z when continuous prior on lambda
+
+code <- '
+arma::mat tau_g_c = Rcpp::as<arma::mat>(taug);
+arma::mat tau_c = Rcpp::as<arma::mat>(tau);
+arma::mat y_c = Rcpp::as<arma::mat>(y);
+arma::mat xbeta_c = Rcpp::as<arma::mat>(x_beta);
+arma::mat mu_c = Rcpp::as<arma::mat>(mu);
+arma::mat g_c = Rcpp::as<arma::mat>(g);
+arma::mat prec_c = Rcpp::as<arma::mat>(prec);
+arma::mat prec_11; arma::mat prec_21;
+double lambda_c = Rcpp::as<double>(lambda);
+arma::mat zg_c = Rcpp::as<arma::mat>(zg);
+int nknots = tau_c.n_rows; int nt = tau_c.n_cols; int ns = tau_g_c.n_rows;
+int nparts;
+arma::mat z(nknots, nt);
+arma::uvec these; arma::uvec nthese; arma::vec tau_g_t (ns);
+arma::vec r_1; arma::vec r_2;
+arma::vec y_t(ns); arma::vec xbeta_t(ns); arma::vec mu_t(ns);
+arma::mat mmm(nknots, nt); arma::mat sd(nknots, nt);
+double mmmkt; double vvvkt; double sdkt; double zkt;
+
+for (int t = 0; t < nt; t++) {
+// extract the current values for day t
+tau_g_t = sqrt(tau_g_c.col(t));
+y_t = y_c.col(t); xbeta_t = xbeta_c.col(t); mu_t = mu_c.col(t);
+
+for (int k = 0; k < nknots; k++) {
+these  = find(g_c.col(t) == (k+1));
+nthese = find(g_c.col(t) != (k+1));  // nthese is "not these"
+nparts = these.n_elem;
+
+// recompute the mean for the day to account for updates to other knots
+mu_t = xbeta_t + lambda_c * zg_c.col(t);
+
+r_1 = (y_t(these) - xbeta_t(these));
+r_2 = (y_t(nthese) - mu_t(nthese)) % tau_g_t(nthese) / sqrt(tau_c(k, t));
+
+prec_11 = prec_c(these, these); prec_21 = prec_c(nthese, these);
+
+mmmkt = lambda_c * sum(r_1.t() * prec_11 + r_2.t() * prec_21);
[…]

### 2016-08-06 LaTeX/jsm2016/HighDimEx.tex
updates with plots for JSM

\documentclass{beamer}
% \usepackage{beamerthemesplit}
\usetheme{Madrid}
% \usepackage{pstricks}
\usepackage{graphicx}
\usepackage{mdwlist}
\usepackage{lineno, hyperref}
\usepackage{stackrel}
\usepackage{changepage}

\usepackage{amssymb,latexsym,amsmath,amsthm,bbm}
\usepackage{hyperref}
\usepackage{tikz}
\usepackage[english]{babel}
\usepackage[latin1]{inputenc}
\usepackage[authoryear]{natbib}
\usepackage{multirow}
\usepackage{verbatim}
\usepackage{alltt}
\usepackage{mycommands}

% \usepackage{cmbright}
\renewcommand*\familydefault{\sfdefault}
\usepackage[T1]{fontenc}

\definecolor{wp-red}{RGB}{204,0,0}
\definecolor{wp-gray}{RGB}{51,51,51}
\definecolor{reynolds-red}{RGB}{153,0,0}
\definecolor{pyroman-flame}{RGB}{209,81,34}
\definecolor{hunt-yellow}{RGB}{253,215,38}
\definecolor{genomic-green}{RGB}{125,140,31}
\definecolor{innovation-blue}{RGB}{66,126,147}
\definecolor{bio-indigo}{RGB}{65,86,161}

\setbeamercolor{structure}{fg=wp-red}
\setbeamercolor{title}{bg=white, fg=wp-red}  % changes color on title page
\setbeamerfont{title}{series=\bfseries, size=\huge}
\setbeamerfont{author}{series=\bfseries, size=\large}
\setbeamerfont{institute}{series=\mdseries, size=\large}

\setbeamercolor{frametitle}{bg=wp-red, fg=white}  % changes color at top of frame
\setbeamerfont{frametitle}{series=\bfseries}
\setbeamercolor{title in head/foot}{fg=white, bg=wp-red}  % changes color for title in footer
\setbeamerfont{title in head/foot}{series=\bfseries}
\setbeamercolor{author in head/foot}{fg=white,bg=wp-gray}  % changes color for author in footer
\setbeamerfont{author in head/foot}{series=\bfseries}

\newcommand{\btheta}{ \mbox{\boldmath $\theta$}}
\newcommand{\blambda}{ \mbox{\boldmath $\lambda$}}
\newcommand{\bmu}{ \mbox{\boldmath $\mu$}}
\newcommand{\bsigma}{ \mbox{\boldmath $\sigma$}}
\newcommand{\balpha}{ \mbox{\boldmath $\alpha$}}
\newcommand{\bbeta}{ \mbox{\boldmath $\beta$}}
\newcommand{\bdelta}{ \mbox{\boldmath $\delta$}}
\newcommand{\bgamma}{ \mbox{\boldmath $\gamma$}}
\newcommand{\brho}{ \mbox{\boldmath $\rho$}}
\newcommand{\bnu}{ \mbox{\boldmath $\nu$}}
\newcommand{\bpsi}{ \mbox{\boldmath $\psi$}}
\newcommand{\btau}{ \mbox{\boldmath $\tau$}}
\newcommand{\bepsilon}{ \mbox{\boldmath $\epsilon$}}
\newcommand{\beps}{ \mbox{\boldmath $\varepsilon$}}
\newcommand{\bX}{ \mbox{\boldmath X}}
\newcommand{\bx}{ {\bf x} }
\newcommand{\by}{ \mbox{\boldmath y}}
\newcommand{\bY}{ \mbox{\boldmath Y}}
\newcommand{\bz}{ \mbox{\boldmath z}}
\newcommand{\bZ}{ \mbox{\boldmath Z}}
\newcommand{\bW}{ \mbox{\boldmath W}}
\newcommand{\bb}{ \mbox{\boldmath b}}
\newcommand{\bg}{ \mbox{\boldmath g}}
\newcommand{\bL}{ \mbox{\boldmath L}}
\renewcommand{\bs}{ \mbox{\boldmath s}}
\newcommand{\bt}{ \mbox{\boldmath t}}
\newcommand{\bu}{ \mbox{\boldmath u}}
\newcommand{\bv}{ \mbox{\boldmath v}}
\newcommand{\ba}{ \mbox{\boldmath a}}
\newcommand{\bfe}{ \mbox{\boldmath e}}
\newcommand{\bE}{ \mbox{\boldmath E}}
\newcommand{\bV}{ \mbox{\boldmath V}}
\newcommand{\bM}{ \mbox{\boldmath M}}
\newcommand{\bA}{ \mbox{\boldmath A}}
\newcommand{\skewt}{skew-$t$ }
\newcommand{\Skewt}{Skew-$t$ }

\newcommand{\STGP}{ \cal{STGP}}
\newcommand{\GP}{ \cal{GP}}

% \newcommand{\argmin}{{\mathop{\rm arg\, min}}}
% \newcommand{\argmax}{{\mathop{\rm arg\, max}}}
% \newcommand{\iid}{\stackrel{iid}{\sim}}
\newcommand{\bit}{\begin{itemize}}
  \newcommand{\eit}{\end{itemize}}

% \newcommand*\oldmacro{}
% \let\oldmacro\insertshorttitle
% \renewcommand*\insertshorttitle{
%   \oldmacro\hfill
%   \insertframenumber\,/\,\inserttotalframenumber}

\title[\Skewt model for threshold exceedances]{A Space-time \skewt model for threshold exceedances}
\author[S. Morris]{Samuel A. Morris}
\institute[]{North Carolina State University}
\date[]{August 3, 2016}

% \subject{Talks}

%%% NCSU BLOCK %%
% \pgfdeclareimage[height=1.6cm]{ncsu-block}{ncsu-block}
% \logo{\pgfuseimage{ncsu-block}}

%%% NCSU LOGO %%
% \pgfdeclareimage[height=1.5cm]{75year}{75year}
% \logo{\vspace{195px}\pgfuseimage{75year}}

% \AtBeginSection[]
% {
%   \begin{frame}<beamer>
%   \frametitle{Outline}
%   \tableofcontents[currentsection,currentsubsection]
%   \end{frame}
% }

% \AtBeginSubsection[]
% {
%   \begin{frame}<beamer>
%   \frametitle{Outline}
%   \tableofcontents[currentsection,currentsubsection]
%   \end{frame}
% }

% If you wish to uncover everything in a step-wise fashion, uncomment
% the following command:

%\beamerdefaultoverlayspecification{<+->}
% \newcommand{\Href}[2]{\href{#1}{\textcolor{blue}{#2}}}

\begin{document}

\begin{frame}
  \titlepage

\begin{center}
   Joint work with Brian Reich (NCSU), Emeric Thibaud (CSU), and Dan Cooley (CSU)
\end{center}
\end{frame}

% \begin{frame}\frametitle{I reached my 5-year return level!}
%   \begin{center}
%     \includegraphics[height=0.7\textheight]{Pic/tunnel_mountain}
%   \end{center}
%  Thanks to the organizers. The meeting has be great.
% \end{frame}

% \begin{frame}\frametitle{Sam Morris}
%   \begin{center}
%     \includegraphics[height=0.7\textheight]{Pic/Sam}
%   \end{center}
% \end{frame}

\begin{frame}\frametitle{Spatial extremes}
 \bit\setlength\itemsep{\fill}
  \item Extreme Values Analysis (EVA) can benefit greatly from spatial methods
  \item Spatial methods can map risk and borrow strength over space to estimate rare-event probabilities
  \item Accounting for spatial dependence is necessary for valid inference
  \item Methods and software in this area are developing rapidly to meet a growing demand
 \eit
\end{frame}

\begin{frame}\frametitle{Max-stable processes}
  \bit\setlength\itemsep{\fill}
  \item Let $Y_1(\bs)$, ..., $Y_m(\bs)$ be iid spatial processes
  \item The pointwise maximum process is $${\tilde Y}(\bs) =\bigvee_{l=1}^m Y_l(\bs)$$
  \item If there exist constants $a_m$ and $b_m$ so that $$Z(\bs)= a_m + b_m {\tilde Y}(\bs)$$
  converges to a valid process as $m\rightarrow\infty$, then $Z$ is max-stable
  \eit
\end{frame}

\begin{frame}\frametitle{Max-stable processes}
  \bit\setlength\itemsep{\fill}
  \item The marginal distribution of $Z$ at each $\bs$ follows generalized extreme value (GEV) distribution
  \item Distribution function is $F(z) = \exp\{-t(z)\}$ where
  \begin{align*}
      t(z) = \left\{ \begin{array}{ll}
        \left[1 + \xi \displaystyle \left(\frac{z - \mu}{\sigma}\right)\right]^{-1 / \xi}, \quad &\xi \neq 0 \\ [1em]
        \exp\left\{- \displaystyle \frac{z - \mu}{\sigma}\right\}, &\xi = 0.
    \end{array}\right.
  \end{align*}
  % \item Three parameters:
  % \begin{itemize}
    \item Location: $\mu \in \mathcal{R}$
    \item Scale: $\sigma > 0$
    \item Shape: $\xi \in \mathcal{R}$
  % \end{itemize}
  \eit
\end{frame}

\begin{frame}\frametitle{Challenges in EVA}
  \bit\setlength\itemsep{\fill}
  \item Covariance focuses on deviations around the mean and not the extremes
  \item Want dependence measure to capture likelihood of seeing values that are jointly extreme
  \item Two common measures of dependence: \vspace{0.5em}
  \begin{itemize}
    \item Extremal coefficient $\vartheta \in (1, 2)$: $$\mbox{Prob}[Z(\bs_1)<c, Z(\bs_2)<c] = \mbox{Prob}[Z(\bs_1)<c]^{\vartheta(\bs_1,\bs_2)}$$
    \item $\chi \in (0, 1)$: $$\chi(\bs_1,\bs_2) = \lim_{c\rightarrow \infty}\mbox{Prob}[Z(\bs_1)>c|Z(\bs_2)>c]$$
  \end{itemize}
  \eit
\end{frame}

\begin{frame}\frametitle{Current approaches and limitations}
  \bit\setlength\itemsep{\fill}
   \item Theory suggests that a max-stable process is a good option for spatial extremes
   \item The max-stable process gives a complicated likelihood function with no closed-form except in trivial cases
   \item Current Bayesian approaches can handle only a moderate number of spatial locations
   \item This is limiting because most modern applications have hundreds or thousands of stations
   \item Because of these challenges advanced methods for e.g., multivariate or nonstationary data are limited
\eit
\end{frame}

\begin{frame}\frametitle{Application -- Air pollution (ozone)}
 \begin{center}
  \includegraphics[width=0.8\linewidth, trim={0, 3em, 0, 0}]{plots/ozone-10jul-us}

  Ozone measurements on July 10, 2005, at 1,089 stations across the US
 \end{center}
\end{frame}

% \begin{frame}\frametitle{Low-rank positive-stable representation}
%   \bit\setlength\itemsep{\fill}
%      \item $Z_t$ is max-stable marginally over the random effects $A_{tl}$
%      \item The joint is GEV - asymmetric Laplace
%      \item Dependence is measured by the extremal coefficient $\vartheta$, defined via $$\mbox{Prob}[Z_t(\bs_1)<c, Z_t(\bs_2)<c] = \mbox{Prob}[Z_t(\bs_1)<c]^{\vartheta(\bs_1,\bs_2)}$$
%       \item For the low-rank PS model $$\vartheta(\bs_1,\bs_2) = \sum_{l=1}^L\left[B_l(\bs_1)^{1/\alpha}+B_l(\bs_2)^{1/\alpha}\right]^{\alpha}\in[1,2] $$
%   \eit
% \end{frame}

% \begin{frame}\frametitle{Estimating the EBFs, $B_l(\bs)$}
%   \begin{enumerate}\setlength\itemsep{\fill}
%     \item Use a rank transformation to standardize data for each $\bs$
%     \item Estimate the extremal dependence between each pair of sites (using $\chi$ or madogram), ${\hat \vartheta}(\bs_i,\bs_j)$
%     \item Spatially (4D) smooth the sample dependence measures
%     \item Constrained least squares (next slide) to minimize the distance between sample (${\hat \vartheta}$) and model ($\vartheta$ as a function of the $B$) spatial dependence
%     \item Order the terms by $v_l = \sum_{\bs}B_l(\bs)$
%   \end{enumerate}
% \end{frame}

% \begin{frame}\frametitle{Estimating the EBFs, $B_l(\bs)$}
%   \begin{itemize}\setlength\itemsep{\fill}
%     \item The objective function to estimate the $B_l$ is
%     $$ \sum_{i<j}\left[\hat{\vartheta}(\bs_i,\bs_j) - \vartheta(\bs_i,\bs_j)\right]^2$$
%     where $\vartheta(\bs_i,\bs_j)$ is a function of $B_l$
%     \item The EBFs must satisfy $B_l(\bs)>0$ and $\sum_{l}B_l(\bs)=1$ for all $\bs$
%     \item The solution is approximated by cycling through the sites and solving a series of constrained optimization problems
%   \end{itemize}
% \end{frame}

% \begin{frame}\frametitle{Contrasts with PCA}
%   \bit\setlength\itemsep{\fill}
%   \item Basis functions are not orthonormal
%   \item Loadings are positive stable, not Gaussian
%   \item Loadings $A_{lt}$ may not be independent
%   \item Computing $A$ and $B$ is not as simple as a few matrix operations
%   \eit
% \end{frame}

% \begin{frame}\frametitle{Analogy with PCA}
%   \bit\setlength\itemsep{\fill}
%   \item Reduces the dimension from $n$ to $L$
%   \item Maps of $B_l(\bs)$ tell us about the most important spatial patterns
%   \item Captures a non-stationary spatial dependence structure
%   \item The $v_l$ tell us how many important features are present
%   \item Loadings $A_{lt}$ can be estimated and fed into future analyses
%   \eit
% \end{frame}

% \begin{frame}\frametitle{Bayesian implementation}
%   \bit\setlength\itemsep{\fill}
%   \item Given the basis function $B_l(\bs)$ we can proceed with\\ MCMC to estimate the remaining parameters
%   \item GEV location:  $\mu_t(\bs) = \beta_{\mu,int}(\bs) + \beta_{\mu,time}(\bs)t$
%   \item GEV scale:  $\log[\sigma_t(\bs)] = \beta_{\sigma,int}(\bs) + \beta_{\sigma,time}(\bs)t$
%   \item GEV shape: $\xi$ for all $\bs$ and $t$
%   \item The $\beta$ have Gaussian process priors
%   \item We use cross-validation (quantile and Bries scores) to select $L$
%   \item Alternative: select $L$ so that $\sum_{l=1}^Lv_l = 0.8$
%   \eit
% \end{frame}

% \begin{frame}\frametitle{Application 1 - forest fires in GA}
%   \bit\setlength\itemsep{\fill}
%   \item The data are the number of acres burned by forest fires each year (1965-2014) in each county of Georgia
%   \item We censor the data at the local 95$^{th}$ percentile, $T(\bs)$
%   \item The censored data are modeled as GEV with spatially-varying location and scale
%   \item The objectives are to map fire risk and determine if it is changing with time
%   \eit
% \end{frame}

% \begin{frame}\frametitle{GA Fires -- time series for each county }
%  \begin{center}
%   \includegraphics[width=1.0\textwidth]{Pic/fire-spag-rand-25}
%  \end{center}
% \end{frame}

% \begin{frame}\frametitle{GA Fires -- picking the threshold}
%  \begin{center}
%   \includegraphics[width=1\textwidth]{Pic/fire-mrl-plots}
%  \end{center}
% \end{frame}

% \begin{frame}\frametitle{GA Fires -- 95th percentile by county, $T(\bs)$}
% \vspace{-40pt}
% \begin{center}
%   \includegraphics[height=1.2\textheight]{Pic/fire-spatial-q95}
%  \end{center}
% \end{frame}

% \begin{frame}\frametitle{Model comparisons}
% \begin{center}
%   {\Large \begin{tabular}{l|cc}
%       $L$ & Brier Score & Quantile score\\ \hline
%       5   & 5.64 & 135.7 \\
%       10  & 5.33 & 127.3 \\
%       15  & 5.00 & 128.3 \\
%       20  & 4.93 & 122.4 \\
%       {\bf 25}  & {\bf 4.78} & {\bf 116.9} \\
%       40  & 4.72 & 115.7
%   \end{tabular}}
%   \end{center}
% \end{frame}

% \begin{frame}\frametitle{GA Fires -- EBF weights, $v_l$}
%   \begin{center}
%     \includegraphics[height=0.9\textheight]{Pic/firev}
%   \end{center}
% \end{frame}

% \begin{frame}\frametitle{GA Fires -- EBF's $B_l(\bs)$}
% \vspace{22pt}
% \begin{center}
%   \includegraphics[width=0.99\textwidth]{Pic/ga-ebf-panel}
% \end{center}
% \end{frame}

% \begin{frame}\frametitle{GA Fires -- Ecoregions}
%   \begin{center}
%     \includegraphics[height=0.9\textheight]{Pic/ga_eco_regions}
%   \end{center}
% \end{frame}

% \begin{frame}\frametitle{Time trend ($\beta_{\mu,time}$, $\beta_{\sigma,time}$) -- posterior mean}
%  \begin{center}
%   \includegraphics[width=1.1\textwidth]{Pic/fire_post_mean}
%  \end{center}
% \end{frame}

% \begin{frame}\frametitle{Time trend ($\beta_{\mu,time}$, $\beta_{\sigma,time}$) -- posterior SD}
%   \begin{center}
%     \includegraphics[width=1.1\textwidth]{Pic/fire_post_sd}
%   \end{center}
% \end{frame}

% \begin{frame}\frametitle{Time trend ($\beta_{\mu,time}$, $\beta_{\sigma,time}$) -- prob $>$ 0}
%   \begin{center}
%     \includegraphics[width=1.1\textwidth]{Pic/fire_post_pos}
%   \end{center}
% \end{frame}

% \begin{frame}\frametitle{Application 2 -- NARCCAP climate model output}
%   \bit\setlength\itemsep{\fill}
%   \item Data consist of annual maximum precipitation at \\697 grid cells in the Eastern US
%   \item The model is run separately for 1969-2000 and 2039-2077
%   \item The objective is the compare the extremes in the two climate periods
%   \item We fit the same model as for the fire data except without censoring
%   \item We fit the model separately for the two periods
%   \eit
% \end{frame}

% \begin{frame}\frametitle{Climate model output for 1969}
%  \begin{center}
%   \includegraphics[height=0.99\textheight]{Pic/1969}
%  \end{center}
% \end{frame}

% \begin{frame}\frametitle{Precip -- EBF weights, $v_l$}
%   \begin{center}
%     \includegraphics[height=0.9\textheight]{Pic/precipv}
%   \end{center}
% \end{frame}

% \begin{frame}\frametitle{Precip -- EBFs $B_l(\bs)$}
% \vspace{22pt}
%  \begin{center}
%   \includegraphics[width=0.99\textwidth]{Pic/precip-ebf-panel}
%  \end{center}
% \end{frame}

\begin{frame}\frametitle{Back to a Gaussian process model}
  \bit\setlength\itemsep{\fill}
  \item The max-stable process is an elegant approach, but does that mean it's the right model?
  \item In reality, it is only an approximation
  \item There are less complicated approximations
  \item For example, we could model daily data as a Gaussian process (GP)
  \item If the goal is spatial interpolation, perhaps this is competitive
\eit
\end{frame}

\begin{frame}\frametitle{GP - asymptotic independence}
  \bit\setlength\itemsep{\fill}
  \item A GP leads to simple interpretation and computing, but asymptotic independence
  \item If $Y(\bs_1)$ and $Y(\bs_2)$ are bivariate normal then $\chi(\bs_1,\bs_2)=0$, i.e., asymptotic independence
  \item This suggests Kriging will not capture extremes
  \item But so much is known for the Gaussian case: nonstationarity, multivariate, numerical approximations, \ldots
  \item Rather than toss it out, can we patch it up?
  \eit
\end{frame}

\begin{frame}\frametitle{Spatial \skewt process}
  A spatial \skewt process (Azzalini and Capitanio, 2014)  resembles a GP but exhibits asymptotic dependence
 \begin{eqnarray}
 Y_t(\bs) &=& \bX(\bs)^\top\bbeta +\lambda\sigma_t|z_t| + \sigma_tv_t(\bs)\nonumber\\
 z_t &\sim& \mbox{Normal}(0,1)\nonumber\\
\sigma_t^2&\sim& \mbox{InvGamma}(a/2,b/2)\nonumber\\
v_t&\sim& \mbox{Spatial GP}\nonumber
 \end{eqnarray}
  \bit\setlength\itemsep{\fill}
  \item Location: $\bX(\bs)^\top\bbeta$
  \item Scale: $b>0$
  \item Skewness: $\lambda\in{\cal R}$
  \item Degrees of freedom: $a>0$
  \eit
\end{frame}

\begin{frame}\frametitle{Good properties}
  \bit\setlength\itemsep{\fill}
  \item Flexible $t$ marginal distribution with four parameters including the degrees of freedom which allows for heavy tails ($a=1$ gives a Cauchy)
  \item Computation on the order of a GP; the only extra steps are $z_t$ and $\sigma_t$ which have conjugate full conditionals
  \item Asymptotic dependence: $\chi(\bs_1,\bs_2)>0$ for all $\bs_1$ and $\bs_2$
  \eit
\end{frame}

\begin{frame}\frametitle{Bad properties and remedies}
  \bit\setlength\itemsep{1em}
  \item Modeling all the data (bulk and extreme) can lead to poor tail probability estimates if the model is misspecified
  \item Long-range dependence: $\chi(\bs_1,\bs_2)>0$ for all $\bs_1$ and $\bs_2$ even if $\bs_1$ and $\bs_2$ are far apart
  \item This occurs because all sites share $z_t$ and $\sigma_t$
  \item Remedies:
  \bit
    \item We use a censored likelihood to focus on the tails
    \item We propose a local \skewt process
  \eit
  \eit
\end{frame}

\begin{frame}\frametitle{Censored likelihood}
  \bit\setlength\itemsep{\fill}
  \item Censored likelihood: We censor the data
  $${\tilde Y}_t(\bs) =\left\{\begin{array}{ll}
  T & \text{for } Y_t(\bs)\le T\\
  Y_t(\bs) & \text{for } Y_t(\bs)>T
  \end{array}\right.$$
  \item Censoring is handled using standard Bayesian imputation methods
  \item The threshold $T$ is chosen by cross-validation
  \item If $T$ is moderately extreme in the distribution (e.g. $q(0.75)$), set $\lambda = 0$
  \eit
\end{frame}

\begin{frame}\frametitle{Local \skewt process}
  \bit\setlength\itemsep{\fill}
  \item Let the knots $\bv_{t1},...,\bv_{tK}$ follow a homogeneous Poisson process over the domain of interest (in practice we fix $K$)
  \item Associated with each is
  \begin{itemize}
    \item $z_{tk}\sim\mbox{Normal}(0,1)$
    \item $\sigma_{tk}^2\sim\mbox{InvGamma}(a/2,b/2)$
  \end{itemize}
  \item The knots partition the domain if we assign location $\bs$ to subregion $k=\argmin_{l}||\bs-\bv_{tl}||$
  \item If $\bs$ is in subregion $k$ then $$Y_t(\bs) = \bX(\bs)^T\bbeta +\lambda\sigma_{tk}|z_{tk}| + \sigma_{tk}v_t(\bs)\nonumber$$
  \item The marginal distribution remains a $t$, but partitioning breaks long-range spatial dependence
  \eit
\end{frame}

\begin{frame}\frametitle{$\chi$-statistic by $h=||\bs_1-\bs_2||$}
 \begin{center}
  \includegraphics[height=0.8\textheight]{plots/chi-h}
 \end{center}
\end{frame}

\begin{frame}\frametitle{Temporal dependence}
  \bit\setlength\itemsep{1em}
    \item It may not be reasonable to assume that observations are temporally independent (e.g. flooding, high temperatures)
    \item Temporal dependence is handled through the $z_{tk}$, $\sigma_{tk}$ and $\bv_{tl}$
    \item Method:
    \bit\setlength\itemsep{0.5em}
      \item Use a copula to transform parameters to \emph{nice} space (i.e. $\mathcal{R}$)
      \item AR(1) structure imposed on parameters in transformed space
      \item Transform back to original parameter space to preserve \skewt
    \eit
  \eit
\end{frame}

\begin{frame}\frametitle{Results of a simulation study}
  In terms of Brier scores for spatial prediction:\vspace{1em}
  \bit\setlength\itemsep{1em}
  \item Data generated as a GP:
  \bit
    \item \skewt is close to GP
    \item max-stable is 15\% -- 30\% worse than GP
  \eit
  \item Data generated as a \skewt with multiple partitions:
  \bit
    \item \skewt is 15\% better than GP
    \item max-stable is 30\% worse than GP
  \eit
  \item Data generated as asymmetric logistic (max-stable):
  \bit
    \item \skewt is close to GP
    \item max-stable performs 10\% better than GP
  \eit
  \item Data generated as Brown-Resnick (max-stable):
  \bit
    \item \skewt performs 40\% -- 60\% better than GP
    \item max-stable performs 40\% -- 60\% better than GP
  \eit
  \eit
\end{frame}

\begin{frame}\frametitle{Application to ozone}
  \bit\setlength\itemsep{\fill}
  \item The USEPA has an extensive network of ozone monitors throughout the US
  \item We will analyze ozone for 31 days in July, 2005 at $n=1,089$ stations
  \item Currently the EPA regulates the annual 99$^{th}$ percentile
  \item Our objective is to map the probability of an extreme ozone event
  \eit
\end{frame}

\begin{frame}\frametitle{Ozone on July 10}
  \begin{center}
    \includegraphics[width=1.2\textheight]{plots/ozone-10jul-us}
  \end{center}
\end{frame}

\begin{frame}\frametitle{Q-Q plots}
  \begin{center}
    \includegraphics[width=0.9\textheight]{plots/qq-res}

    Gaussian Q-Q plot (left) and skew-$t$ with $a = 10$ and $\lambda = 1$ Q-Q plot (right)
  \end{center}
\end{frame}

\begin{frame}\frametitle{Cross-validation}
  \bit\setlength\itemsep{\fill}
  \item We split the sites into training and testing
  \item We found that $K=15$ knots and censoring at $T$ equal to the median with no time series gave the best results
  \item Results were not sensitive to these tuning parameters
  \item This model was 5\% more accurate (Brier score) than GP
  \item The max-stable model fit was 15\% less accurate than GP
  \eit
\end{frame}

\begin{frame}\frametitle{Fitted 99$^{th}$ percentile - Gaussian}
  \begin{center}
    \includegraphics[width=0.45\linewidth]{plots/q99gaus}
    \includegraphics[width=0.45\linewidth]{plots/q99skewtTS}

    Gaussian (left) Symmetric-$t$, 10 knots, T = 75, Time series (right)
  \end{center}
\end{frame}

% \begin{frame}\frametitle{Fitted 99$^{th}$ percentile - Thresholded $t$}
%   \begin{center}

%   \end{center}
% \end{frame}

\begin{frame}\frametitle{Difference (Thresholded $t$ - Gaussian)}
  \begin{center}
    \includegraphics[width=0.55\linewidth]{plots/q99diffgausskewt}

    Difference between Symmetric-$t$, 10 knots, T = 75 and Gaussian
  \end{center}
\end{frame}

\begin{frame}\frametitle{Summary}
  \bit\setlength\itemsep{\fill}
  \item Our proposed method can handle large datasets
  \item The proposed model gives a balance between theoretical properties and computational feasibility
  % \item This should at least be used as a benchmark for more sophisticated approaches
  \item Work supported by NSF, NIH, DOI, and EPA
  \item Thanks!
  \eit
\end{frame}

\end{document}

### 2016-12-09 LaTeX/response_2.tex
edits for final submission

\documentclass[11pt]{article}
\usepackage[letterpaper, margin=1in]{geometry}
% \documentclass[useAMS,usenatbib,referee]{biomweb}
\usepackage{amssymb, amsthm, amsmath}
\usepackage{amssymb, amsmath}
\usepackage{bm}
\usepackage{graphicx}
\usepackage[authoryear]{natbib}
\usepackage{bm}
\usepackage{verbatim}
\usepackage{lineno}
\usepackage{soul}
\usepackage{color}
% \usepackage{enumitem}
\usepackage{enumerate}
\usepackage{setspace}
\usepackage{appendix}
% \usepackage{caption}
% \captionsetup[table]{name=Web Table}

% cross-referencing appendices
\usepackage{xr}
\externaldocument{skewt_rev}
\externaldocument{supplemental_rev}

% \usepackage[left=1in,top=1in,right=1in]{geometry}
% \pdfpageheight 11in
% \pdfpagewidth 8.5in
% \linespread{2.0}
\input{mycommands.sty}
\begin{document}

\begin{center}
Point-by-Point response to Report on revised manuscript\\
Biometrics MS \#150926M\\
\today
\end{center}

Again, we wish to thank the reviewers and editors for their helpful comments and feedback regarding our manuscript. 
To address the concerns, we edited the manuscript to address the concerns given in the revised report.

\section*{Response to the comments of the Associate Editor}
\subsection*{Remaining comments}
\begin{enumerate}[1.]
	\item p.2 1.1-2: ``Asymptotic \ldots level'' this sentence doesn't make sense

	\begin{response}
	We have changed this sentence to read ``Asymptotic dependence/independence are notions which describe the probability that two random variables simultaneously exceed an extremely high level.''
	\end{response}

	\item p.6 Section 3.2 1.2: space missing after .

	\begin{response}
	We have added a space after the `.'
	\end{response}

	\item p.7 Figure 2: is this really an \emph{estimate} of $\chi(h)$; are these not theoretical values?

	\begin{response}
	We have changed the wording on p.7 (2 lines above the suggested location of Figure 2) text describing Figure 2 from ``we estimate $\chi(h)$ \ldots'' to ``we show $\chi(h)$ \ldots''.
	\end{response}

	\item The phrasing ``we use a transformation of a Gaussian random variable on $z_t$\ldots'' sounds quite strange. Why not just say ``we transform $z_t$ to the Gaussian scale using the probability integral transform''? And similarly for $\sigma^2_t$.

	\begin{response}
	We have adopted the wording suggested by the reviewer.
	\end{response}

	\item p.7 Section 4.1: What method do you use to simulate from the truncated multivariate Gaussian - rejection? It would be useful to give an idea of how difficult or otherwise this step is.

	\begin{response}
	We have added clarification that we use the conditional normal distribution to sample from the univariate truncated normal distribution.
	\end{response}

	\item p.11 Section 5.1 / Fig. 3: the title of the 4th panel in Figure 3 is ``Asymmetric logistic'' whereas (4) in 5.1 is ``Reich and Shaby'' process

	\begin{response}
	We have changed the title of the plot to ``Reich and Shaby'' to be consistent with the text.
	\end{response}

	\item p.11 1-2. \texttt{rmaxstab}

	\begin{response}
	This has been corrected.
	\end{response}

	\item p.12 item (5) typo ``max-table''

	\begin{response}
	This has been corrected.
	\end{response}

	\item p.15 Section 6.1: when giving the burn in period of 25,000, is this taken out of the 30,000 iterations, or in addition to them?

	\begin{response}
	We have adjusted the wording (as well as the corresponding part of the simulation study in 5.1) here to reflect that the burn-in period is taken out of the number of iterations reported.
	\end{response}

	\item Web Appendix C, p.6 final line: typo "Poission"; p.7 eq.(3) typo $>=$

	\begin{response}
	These have been corrected.		
	\end{response}

\end{enumerate}
\bibliographystyle{biom}
\bibliography{library}

\end{document}

### 2016-12-09 LaTeX/skewt_rev_2.tex
edits for final submission

--- 2016-06-30 LaTeX/skewt_rev.tex
@@
 \usepackage{xr}
-\externaldocument{supplemental_rev}
+\externaldocument{supplemental_rev_2}
 
@@
 Third, we aim to capture the dependence structure when ozone is at high levels, and dependence at these levels may not be well-represented by covariances which focus again on mean behavior.
-Asymptotic dependence/independence (see \sref{sts:extdep}) are notions which describe how two random variables' probability of simultaneous exceedance of an extremely high level.
+Asymptotic dependence/independence (see \sref{sts:extdep}) are notions which describe the probability that two random variables simultaneously exceed of an extremely high level.
 The Gaussian distribution always exhibits asymptotic independence, except in the case of perfect dependence, thus is an inappropriate model for data which exhibits asymptotic dependence.
@@
 \subsection{Partitioning to remove long-range asymptotic dependence}\label{sts:part}
-For a large spatial domain, it may not be reasonable to assume sites that are far apart demonstrate asymptotic dependence.%, so we handle the problem of long-range asymptotic dependence with a random partition.
+For a large spatial domain, it may not be reasonable to assume sites that are far apart demonstrate asymptotic dependence.
+%, so we handle the problem of long-range asymptotic dependence with a random partition.
 % Modeling different levels of asymptotic dependence was discussed by \citet{Wadsworth2012}.
@@
 In practice we fix $K$ at a finite value and use a uniform distribution for the knots $\bw_1, \ldots, \bw_K$.
-In \fref{stfig:chi}, we estimate $\chi(h)$ for $K = 1, 3, 5, 10$ partitions for a \skewt distribution with $\alpha = 10$, and 3 degrees of freedom.
+In \fref{stfig:chi}, we show $\chi(h)$ for $K = 1, 3, 5, 10$ partitions for a \skewt distribution with $\alpha = 10$, and 3 degrees of freedom.
 % In \fref{stfig:chi} we estimate $\pi(h)$ through simulation.
@@
 Then the transformed knots $\bw^*_{tk} \in \calR^2$.
-We use a transformation of a Gaussian random variable on $z_{t}(\bs)$ to ensure that the marginal distributions of $z_t(\bs)$ are half-normal.
+We transform $z_{t}(\bs)$ to the Gaussian scale using the probability integral transform to ensure that the marginal distributions of $z_t(\bs)$ are half-normal.
 Let
@@
 where HN is the distribution function of a half-normal random variable.
-We also use a transformation of a Gaussian random variable on $\sigma^2_t(\bs)$ to ensure that the marginal distributions of $\sigma^2_t(\bs)$ are inverse gamma.
+We also transform $\sigma^2_t(\bs)$ to the Gaussian scale using the probability integral transform to ensure that the marginal distributions of $\sigma^2_t(\bs)$ are inverse gamma.
 Let
@@
 After conditioning on $\lambda$, $z_t(\bs)$ and non-censored observations, $Y_t(\bs)$ has truncated normal full conditionals $Y_t(\bs) \sim N_{(-\infty, T)}(\bX_t^\top(\bs) \beta + \lambda | z_t(\bs)|, \bSigma)$.
+To impute the censored observations, at each MCMC iteration, for censored site $i$ and day $t$, we use the conditional distribution of $\widetilde{Y}_t(\bs_i)$ given the current imputed values for all censored sites and true values for all non-censored sites $\widetilde{Y}^{(c)}_t(\bs_{-i})$.
+Conditional on $\widetilde{Y}^{(c)}_t(\bs_{-i})$, the imputation of $\widetilde{Y}_t(\bs_i)$ is done using a univariate truncated normal likelihood with mean and standard deviation from the conditional multivariate normal distribution of $\widetilde{Y}_t(\bs)$.
 
@@
 We set the dependence parameter ($\alpha$ in Reich and Shaby, 2012) to 0.5 which represents moderate spatial dependence.
-For the final design, we generated data from a Brown-Resnick max-stable process  using \texttt{rmaxtab} in the \texttt{SpatialExtremes} package of \texttt{R} \citep{Ribatet2015}.
+For the final design, we generated data from a Brown-Resnick max-stable process  using \texttt{rmaxstab} in the \texttt{SpatialExtremes} package of \texttt{R} \citep{Ribatet2015}.
 For this design we fixed unit \Frechet margins, and we used a range of 1 and smoothness 0.5.
@@
   \item Symmetric-$t$ marginal, $K=5$ knots, $T=q(0.80)$
-  \item \citet{Reich2012} max-table model thresholded at $T = q(0.80)$
+  \item \citet{Reich2012} max-stable model thresholded at $T = q(0.80)$
 \end{enumerate}
@@
 We tried also fitting the \skewt marginals for the thresholded models, but it is very challenging for the MCMC to properly identify the skewness parameter with a censored left tail.
-Each chain of the MCMC ran for 20,000 iterations with a burn-in period of 10,000 iterations.
+Each chain of the MCMC ran for 20,000 iterations which includes a burn-in period of 10,000 iterations.
 Although the goal of this simulation study is not to assess parameter estimation, for design (2) and method (2) where data are generated and fit with a \skewt distribution, the samples converges for $\lambda$ and $a$, and the empirical coverage of posterior 95\% intervals is near the nominal level (96\% coverage for $\lambda$ and 90\% for $a$).
@@
 Despite the fact that there is only one month of data, we can get some sense of extremal dependence between sites by looking at joint occurrences of high sample quantiles.
-For example, \fref{stfig:bivariateozone} suggests there is more agreement between sites that are close to one another than sites that are far from one another.
-\begin{figure}
-  \centering
-  \includegraphics[width=\linewidth]{plots/daily-quantiles-ozone.pdf}
-  \caption{Daily quantiles for two monitoring locations near Columbus, OH (left) and daily quantiles for a monitoring location in Los Angeles, CA and Columbus, OH (right)}
-  \label{stfig:bivariateozone}
-\end{figure}
+For example, in \aref{sta:ozonespatdep}, we present a plot that suggests there is more agreement between sites that are close to one another than sites that are far from one another.
+
 % Another aspect that distinguishes our approach from more traditional extremes analyses, is how the threshold is selected.
@@
 
-Each chain of the MCMC ran for 30,000 iterations with a burn-in period of 25,000 iterations.
+Each chain of the MCMC ran for 30,000 iterations which includes a burn-in period of 25,000 iterations.
 We used the same priors for the spatial covariance parameters, skewness parameter, and knots as in the simulation study.
@@
 In particular, for all extreme quantiles, selecting a moderate number of knots (e.g. $K = 5, \ldots, 10$) tends to give the best results.
-\tref{sttbl:ozoneresults} shows the best two models for selected extreme quantiles.
+See \aref{sta:ozonemodelperf} for the performance of the best two models for selected extreme quantiles.
 
@@
 \end{figure}
-\begin{table}
-  \small
-  \caption{Top two performing models for predicting ozone exceedance of level $L$ with Relative Brier score}
-  \label{sttbl:ozoneresults}
-  \centering
-  \begin{tabular}{|l|l l l c|l l l c|}
-    \cline{2-9}
-    \multicolumn{1}{c|}{}  & \multicolumn{4}{c|}{1st} & \multicolumn{4}{c|}{2nd} \\
-    \hline
-    $L = q(0.90)$  & No time series & $K=6$  & $T=50$  & BS: 0.992 &
-                 No time series & $K=1$  & $T=0$  & BS: 0.992 \\
-    $L = q(0.95)$  & Time series & $K=5$ & $T=50$ & BS: 0.988 &
-                 No time series & $K=6$  & $T=50$ & BS: 0.989\\
-    $L = q(0.98)$  & Time series & $K=7$  & $T=50$ & BS: 0.973 &
-                 Time series & $K=5$ & $T=50$ & BS: 0.975\\
-    $L = q(0.99)$  & No time series    & $K=8$ & $T=0$ & BS: 0.946 &
-                 Time series    & $K=9$  & $T=75$ & BS: 0.947\\
-    $L = q(0.995)$ & No time series    & $K=8$  & $T=0$ & BS: 0.951 &
-                 Time series    & $K=9$ & $T=75$ & BS: 0.956\\
-    \hline
-  \end{tabular}
-\end{table}
 
@@
 
+\section{Supplementary Materials}
+Web Appendices, Tables, and Figures referenced in Sections 2.1, 2.2, 3.2, 3.3, 5.3, 6, and 6.2 are available with this paper at the Biometrics website on Wiley Online Library.
+\texttt{R} code implementing the spatial \skewt model for threshold exceedances is available with this paper at the Biometrics website on Wiley Online Library.\vspace*{-8pt}
+
 \section*{Acknowledgments}
@@
 The authors also received high-performance computing support from Yellowstone (ark:/85065/d7wd3xhc) provided by NCAR's Computational and Information Systems Laboratory, sponsored by the National Science Foundation.
-
-\section*{Supplementary Materials}
-Web Appendices A--E are available with this paper at the Biometrics website on Wiley Online Library.
-\texttt{R} code implementing the spatial \skewt model for threshold exceedances is available with this paper at the Biometrics website on Wiley Online Library.\vspace*{-8pt}
 

### 2016-12-09 LaTeX/supplemental_rev_2.tex
edits for final submission

--- 2016-06-29 LaTeX/supplemental_rev.tex
@@
 \usepackage{xr}
-\externaldocument{skewt_rev}
+\externaldocument{skewt_rev_2}
 
@@
 since knots in both $B_1$ and $B_2$ is sufficient, but not necessary, to ensure that $\bs_1$ and $\bs_2$ are in different partition sets.
-By definition of a Poission process, $N_1$ and $N_2$ are independent and thus $P(N_1 > 0, N_2 > 0) = P(N_1 > 0)^2$, and
+By definition of a Poisson process, $N_1$ and $N_2$ are independent and thus $P(N_1 > 0, N_2 > 0) = P(N_1 > 0)^2$, and
 \begin{align}
@@
 \begin{align}
-  P(L_1 \neq L_2) >= P(N_1 > 0)^2 = [1 - P(N_1 = 0)]^2 = [1 - \exp\left(-\lambda^*_{PPB1} h^2\right)]^2
+  P(L_1 \neq L_2) \ge P(N_1 > 0)^2 = [1 - P(N_1 = 0)]^2 = [1 - \exp\left(-\lambda^*_{PPB1} h^2\right)]^2
 \end{align}
@@
 \end{figure}
+
+% \section{Generating from the max-stable model of \citet{Reich2012}} \label{sta:reichshaby}
+% The max-stable model of \citet{Reich2012} uses the \citet{Smith1990} max-stable process with a nugget term that follows a GEV(1, $\alpha$, $\alpha$).
+% The purpose of the nuggest is to address criticism of the Smith max-stable process as unrealistically smooth.
+% To generate from the max-stable model of \citet{Reich2012}, we use Algorithm 1.2 given of \citet{Stephenson2003}.
+% In step 1, we generate from the symmetric logistic distribution using the \texttt{rmevd} function in the \texttt{evd} package in \texttt{R} with dependence parameter 0.5 \citep{Stephenson2002}.
+% The dependence parameter, is similar to $\gamma$ in that it controls the relative strength of the nugget effect, so a setting of $0.5$ should give moderate spatial dependence.
+% For the asymmetry parameters, we introduce 144 spatial knots on a regular lattice in the square $[1, 9] \times [1, 9]$.
+% One contraint for this model is that the asymmetry parameters must sum to 1, so we use standardized Gaussian kernels with bandwidth 1 as in \citet[equations (2.5) and (2.6)]{Reich2012}.
+
+\section{Spatial dependence of ozone} \label{sta:ozonespatdep}
+In this appendix, we include a plot suggesting spatial dependence in ozone measurements.
+\begin{figure}
+	\centering
+	\includegraphics[width=\linewidth]{plots/daily-quantiles-ozone.pdf}
+	\caption{Daily quantiles for two monitoring locations near Columbus, OH (left) and daily quantiles for a monitoring location in Los Angeles, CA and Columbus, OH (right)}
+	\label{stfig:bivariateozone}
+\end{figure}
+
+\section{Best performing models} \label{sta:ozonemodelperf}
+This appendix includes a table with the top two performing models for predicting ozone exceedance of level $L$.
+\begin{table}
+	\small
+	\caption{Top two performing models for predicting ozone exceedance of level $L$ with Relative Brier score}
+	\label{sttbl:ozoneresults}
+	\centering
+	\begin{tabular}{|l|l l l c|l l l c|}
+		\cline{2-9}
+		\multicolumn{1}{c|}{}  & \multicolumn{4}{c|}{1st} & \multicolumn{4}{c|}{2nd} \\
+		\hline
+		$L = q(0.90)$  & No time series & $K=6$  & $T=50$  & BS: 0.992 &
+		No time series & $K=1$  & $T=0$  & BS: 0.992 \\
+		$L = q(0.95)$  & Time series & $K=5$ & $T=50$ & BS: 0.988 &
+		No time series & $K=6$  & $T=50$ & BS: 0.989\\
+		$L = q(0.98)$  & Time series & $K=7$  & $T=50$ & BS: 0.973 &
+		Time series & $K=5$ & $T=50$ & BS: 0.975\\
+		$L = q(0.99)$  & No time series    & $K=8$ & $T=0$ & BS: 0.946 &
+		Time series    & $K=9$  & $T=75$ & BS: 0.947\\
+		$L = q(0.995)$ & No time series    & $K=8$  & $T=0$ & BS: 0.951 &
+		Time series    & $K=9$ & $T=75$ & BS: 0.956\\
+		\hline
+	\end{tabular}
+\end{table}
 
